### 2021-01-25 cluster_paper/authors.tex
stage cluster_paper and planet_paper

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0001-8732-6166]{J. D. Hartman}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0001-7204-6727]{G. \'A. Bakos}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

### 2021-01-25 cluster_paper/ms.tex
stage cluster_paper and planet_paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\cn}{NGC\,2516} % cluster name

\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13

\newcommand{\kms}{\,km\,s$^{-1}$}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{Core \& Halo of NGC\,2516}

\begin{document}

\defcitealias{bouma_wasp4b_2019}{B19}

\title{
  Cluster Difference Imaging Photometric Survey. III.
  The Coeval Halo and Core of NGC\,2516
}

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

\begin{abstract}
  foo,bar.
\end{abstract}

\keywords{
	stellar ages (1581),
	young star clusters (1833),
  open clusters (1160),
  stellar rotation (1629)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

In the traditional picture of star formation molecular clouds go
gravitationally unstable, and collapse into little knots.  The knots
produce many stars close to each other.  The resulting ``open star
clusters'' are, as with any first-order guess, spherical.

How does the cluster evolve?  The answer depends
on the total stellar mass.
% What exactly does this mean?
The smallest 90\% of star clusters disperse during the ``embedded
phase'' (CITE).  This is driven mostly by {\it process A} (CITE).  The
more massive clusters make it to maybe 100 Myr. Maybe a bit longer.
Their evaporation is thought to be driven by collisions with molecular
clouds (CITE, Spitzer 1958), and the galactic tide (CITE), and {\it
process A}, and {\it process B}.  Of course, the expected evaporation
time should depend on factors including the mass of the cluster
itself, how many high-mass (O and B) stars form, since their winds
[and maybe supernovae] clear out most of the cloud, and the initial
density of the gas cloud to begin with.

Assuming that the cluster achieves virial equilibrium, stellar fly-bys
then conspire to segregate the stellar mass distribution within the
cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
assuming equipartition of the specific ``thermal energy'' per star,
then kT = 0.5$mv^2$ at lower masses requires higher velocities).

Identification of the stars that disperse into the galactic field is
an important task for understanding the conditions under which stars
and star clusters form, and for understanding how they subsequently
evolve.  For instance, how does the process of radial migration across
the galactic disk affect cluster dispersal?  Did the Sun form in an
open cluster? If so, how massive was its cluster, and is there any
hope at identifying the stars that formed near the Sun?
Qualitatively, the dispersal of open clusters also provides perhaps
the best ``test case'' for the concept of ``chemical tagging'', also
referred to as ``galactic archaeology'' (CITE).

Outside of the issue of star formation, identifying the remnant halos
of open clusters is important for a separate project: that of
discovering young transiting planets.  Young transiting planets are
hard to find because young stars are rare (CITE), and often reside in
crowded regions of the sky (CITE).  If the halos of nearby star
clusters can be reliably identified, this could expand the census of
nearby sub-Gyr stars by a factor of 2 or even 3 (CITE Meingast).  A
fortuitous benefit of searching for planets in cluster environments is
that issues with stellar crowding are also alleviated.

Recent clustering studies using Gaia have begun to report the
identification of structures that could correspond to low-density
halos of stars that may have evaporated from open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  However, different clustering
methods on the Gaia data tend to give different results (CITE: Hunt \&
Reffert 2020).  Employing say Gaussian Mixture Modelling, or any
analogous method that requires ``clusters'' to be ellipses in
spatial/velocity phase-space unsurpsingly yields open clusters that
are roughly elliptical (CITE: Cantat-Gaudin+18). Unsupervised
clustering methods such as HDBScan (e.g., Kounkel+19) have been found
to yield additional structures, particularly in lower density regions
such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19, Newton+21),
but also more generally around many open clusters (Kounkel+19).
Unsupervised approaches that incorporate physical constraints (e.g.,
imposing a maximum velocity dispersion on putative members) yield
similar results, potentially with higher purity (Meingast+21).

We've recently been making TESS light curves of age-dated stars across
the sky, as part of a Cluster Difference Imaging Photometry Survey
(CDIPS, CITE Bouma+19).  Our analysis of Cycle 1 (Sectors 1-13)
yielded light curves of \numsouthernuniqlcs candidate cluster members
in the Southern Ecliptic hemisphere, available on MAST
[\href{https://archive.stsci.edu/hlsp/cdips}{1},\href{https://ui.adsabs.harvard.edu/abs/2019ApJS..245...13B/abstract}{2}].
% TODO: make this number actually quantitative...
Based on rotation periods, $\approx$25\% appear to be bonafide cluster
members.

As part of a broader project of identifying a large and clean sample
of young stars for a transit search, we take the opportunity of this
paper to answer a rather modest question: in just a single rich
southern open cluster, is the cluster halo coeval with the core?  The
cluster we chose for this analysis was NGC\,2516, since it was young
($\sim$110\,Myr) and close ($d=400$\,pc) enough to facilitate rotation
measurements using TESS, and some degree of spectroscopic analysis.
We want to know: is the halo real? To what extent can we use Gaia
alone to reliably identify age-dated needles in the haystack of boring
field stars?  And more generally, what are the implications for the
evolution of open clusters if they do have halos?

Section~\ref{sec:gaia} presents the astrometric and photometric data
from Gaia, and clarifies our usage of the terms ``core'' and ``halo''.
Section~\ref{sec:agedate} age-dates the halo of NGC\,2516, using TESS
gyrochronology (Section~\ref{sec:rotation}), and lithium depletion
(Section~\ref{sec:lithium}).  In Section~\ref{sec:discussion} we
discuss the implications of this analysis for NGC\,2516 specifically
and stellar spin-down and open cluster evolution generally.
Section~\ref{sec:conclusion} presents our conclusions.

\section{A 250\,pc Halo around a Core?}
\label{sec:gaia}
%
% What is the core? What is the halo?
%

\subsection{Gaia Astrometry}
\label{subsec:astrometry}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Reported components of NGC\,2516 in position and
    velocity space.}
    The ``core'', identified by \citet{cantatgaudin_gaia_2018} using
    Gaia DR2, is visually coincident
    with where you would think the cluster is if you looked at it through a pair
    of binoculars.
    The ``halo'' was identified by \citet{kounkel_untangling_2019}
    using a less restrictive membership assignment algorithm (discussed
    in the appendices).
    The ``field'' is a set of randomly drawn and non-overlapping stars
    within a
    $(\alpha, \delta, \pi)$ cone centered on the cluster.
		\label{fig:gaia6d}
	}
\end{figure*}

Figure~\ref{fig:gaia6d} shows the problem we would like to address.
In this figure, the ``core'' comprises \cn members that
\citet{cantatgaudin_gaia_2018} reported to have ``membership
probability'' exceeding 10\%, based on the Gaia DR2 astrometric data.
The exact meaning of this probability, and details concerning their
clustering algorithm, are discussed in Appendix~\ref{app:clustering}.
The ``halo'' comprises \cn members that
\citet{kounkel_untangling_2019} reported as members, also based on the
Gaia DR2 data.  That study did not report continuous membership
probabilities, instead opting for the binary ``member'' or ``not''.
These two different clustering methods yielded wildly different
results.  What is the true structure of \cn?
Are the core and halo truly coeval?

\section{Age-Dating the Halo of NGC\,2516}
\label{sec:agedate}

\subsection{HR Diagram from Gaia}
\label{subsec:hr}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f2a.pdf}
		}

		\subfloat{
			\includegraphics[width=0.8\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516, using Gaia EDR3 photometry.}
    {\it Top:} The core (black) shows a clean sequence consistent with
    stars with a fixed age and metallicity, and varying mass.  The
    halo (blue) is similar, but somewhat noisier.  The faintest M
    dwarfs in the core and halo are brighter than in the field star
    comparison sample (gray), consistent with these stars having not
    yet reached the ZAMS.
    {\it Bottom:} Reported members of the halo, as a function of
    galactic latitude. Can the additional scatter in the halo be
    understood through differential reddening?
    {\bf Maybe}.
    \label{fig:hr}
  }
\end{figure*}

The first check on whether this membership assignment is plausible was
already performed by \citet{kounkel_untangling_2019} and more recently
by \citet{meingast_2021}.  That check is to see whether the HR
diagrams of these cluster components that were selected based on
positions and velocities photometrically support the claim that they
are coeval.

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.
The halo members are roughly consistent with this, but they do show
greater scatter.  One possible explanation for this scatter is that
the halo is more contaminated by field stars.
Another, explored in the other panel of the figure, is differential
reddening.
The halo is reported to span 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun
by up to 200\,pc (Figure~\ref{fig:gaia6d}).

%TODO 
Comparing these HR diagrams to the PARSEC isochrone models, we find
that X, Y, and Z.
In particular, ``the faintest M dwarfs in the core and halo are
brighter than in the field star comparison sample, consistent with
these stars having not yet reached the ZAMS.''
This is consistent with the 
main-sequence turn-off being at $Bp-Rp\approx0.05$, which implies an
age of XXX.
The resulting photometric age we adopt for the core is XXX.
For the halo, the claimed age from photometry is YYY.
Applying the same procedure to the field star comparison sample,
we get an age of ZZZ.

\subsection{Rotation from TESS}
\label{subsec:tess}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f3a.pdf}
		}

		\subfloat{
			\includegraphics[width=0.8\textwidth]{f3b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn in the space of rotation
    period and Gaia color.}
    The {\it top} plot shows periods against a linear scale, while the
    {\it bottom}  shows them against a logarithmic scale.
		\label{fig:rot}
	}
\end{figure*}

%TODO: what are these numbers?
Duly motivated, we collected light curves from TESS photometry
for YYYY of the ZZZZ cluster members.
%TODO write the words to describe the procedure, the detrending, the
%selection function you used for these janky periods, etc.

\subsection{Lithium from Gaia-ESO and GALAH}
\label{subsec:spectra}

\section{Discussion}
\label{sec:discussion}

  \subsection{How did it form?}
  \subsection{Mass differences between center and outer reaches?}
  \subsection{Fast rotators: are we going faster than other clusters?}

\section{Conclusion}
\label{sec:conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\acknowledgements
\raggedbottom

The authors thank X and Y for fruitful discussions.
%
L.G.B. and J.H. acknowledge support by the TESS GI Program, program
NUMBER, through NASA grant NUMBER.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
2020A-0146; 2020B-NUMBER PI: L{.}~Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
ACKNOWLEDGE PFS / CAMPANAS.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
Latham, Sara Seager, Josh Winn, Jon Jenkins) and the many TESS team
members for their efforts to make the mission a continued success.
%

%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

% %
% This research made use of the NASA Exoplanet Archive, which is
% operated by the California Institute of Technology, under contract
% with the National Aeronautics and Space Administration under the
% Exoplanet Exploration Program.
% %

% Resources supporting this work were provided by the NASA High-End
% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
% Division at Ames Research Center for the production of the SPOC data
% products.
%

% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
% Chilean Ministry of Economy. A.J.\ acknowledges additional support
% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
% from FONDECYT Post-doctoral Fellowship Project 3180246.
% %
% C.T.\ and C.B\ acknowledge support from Australian Research Council
% grants LE150100087, LE160100014, LE180100165, DP170103491 and
% DP190103688.
% %
% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
% Astronomy \& Astrophysics, funded through an endowment established by
% the Dunlap family and the University of Toronto.
% %
% D.D.\ acknowledges support through the TESS Guest Investigator Program
% Grant 80NSSC19K1727.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey,
    SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	%{\bf VLT (number), UVES and GIRAFFE} (CITE: Pasquini et al 2002)
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

% \input{TOI837_phot_table.tex}
% \input{TOI837_rv_table.tex}
% \input{ic2602_ages.tex}
% \input{starparams.tex}
% \input{compparams.tex}
% \input{model_posterior_table.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\appendix
\section{Clustering method details}
\label{app:clustering}

%
% TODO: the following two paragraphs, when quoted, are ENTIRELY FROM
% the relevant papers. they should probably be nixed in favor of
% concision.
%
\citet{cantatgaudin_gaia_2018} applied a procedure that ultimately
yielded what we call ``the core''.  Their procedure was to first query
a Gaia DR2 cone around the previously reported RA and dec of the
cluster, and within $\pm 0.5$\,mas of its previously reported
distance. 
% TODO: which was it?
The outer radius of their cone was either r2 from MWSC (CITE
Kharchenko 2013), or twice the ``cluster radius'' listed by Dias.  No
proper motion cut was applied.  They then applied an unsupervised
classification scheme called UPMASK to $G<18$\,mag stars within this
cone (CITE).  The steps of UPMASK are first to perform a k-means
clustering in the ``astrometric space'' ($\mu_{\alpha'}, \mu_\delta,
\pi$).  Then, a ``veto'' step is applied to assess whether the groups
of stars output from the k-means clustering are or are not more
concentrated than a random distribution.  This is implemented by
``comparing the total branch length of the minimum spanning tree
connecting the stars with the expected branch length for a random
uniform distribution covering the investigated field of view''.  ``To
turn this yes/no flag to a membership probability,
\citet{cantatgaudin_gaia_2018} then redraw new values of
($\mu_{\alpha'}, \mu_\delta, \pi$) for each source based on the listed
value, uncertainty, and covariance.  After a certain number of
redrawings, the final probability is the frequency with which a given
star passes the veto".  In the case of \cn, this yielded a reported
``\texttt{r50}'' within which half of the cluster members were found
to be within 0.496$^\circ$.

\citet{kounkel_untangling_2019} applied a different unsupervised
clustering method to the 5-dimensional Gaia DR2 data (omitting radial
velocities, due to their sparsity).  Their selection function (see
their Section~2) yielded $\approx 2\times 10^7$ stars, mostly within
$\approx 1$\,kpc and typically with $G<18$\,mag.  Their clustering
algorithm was the ``hierarchical density-based spatial clustering of
applications with noise'' (HDBSCAN, CITE McInnes17).  The classical
DBSCAN algorithm ``identifies clusters as overdensities in a
multi-dimensional space in which the number of sources exceeds the
required minimum number of points within a neighborhood of a
particular linking length $\epsilon$.  HDBSCAN does not depend on
$\epsilon$; instead it condenses the minimum spanning tree by pruning
off the nodes that do not meet the minimum number of sources in a
cluster and reanalyzing the nodes that do. Depending on the chosen
algorithm, it would then either find the most persistent structure
(through the excess of mass method), or return clusters as the leaves
of the tree (which results in somewhat more homogeneous clusters). In
both cases it is more effective at finding structures of varying
densities in a given data set than DBSCAN.''
`` The two main parameters that control HDBSCAN are the number of
sources in a cluster and the number of samples. The former is the
parameter that rejects groupings that are too small; the latter sets
the threshold of how conservative the algorithm is in its
considerations of the background noise (even if the resulting noisy
groupings do meet the minimum cluster size).  By default, the sample
size is set to the same value as the cluster size, but it is possible
to adjust them separately.''

\listofchanges

%\allauthors
\end{document}

### 2021-01-25 planet_paper/authors.tex
stage cluster_paper and planet_paper

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

%
% key contributions
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0001-8732-6166]{J. D. Hartman}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

% PENDING
\author[0000-0001-7409-5688]{G. Stefansson} % halpha, bisectors, spec analysis
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%PENDING
\author[0000-0002-4891-3517]{G. Zhou} % metallicity analysis++
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

%PENDING
\author[0000-0002-9158-7315]{R. Brahm} %rafael.brahm@uai.cl coordination, WINE board
\affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

%PENDING
\author[0000-0002-5674-2404]{P. Evans}
\affiliation{El Sauce Observatory, Coquimbo Province, Chile}

%PENDING
\author[0000-0001-6588-9574]{K. A. Collins} % karenacollins@outlook.com
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

%PENDING
\author[0000-0001-8045-1765]{E. Flowers}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%PENDING
\author[0000-0002-8964-8377]{S. N. Quinn} % squinn@cfa.harvard.edu, CHIRON vetting spectrum
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

% %PENDING
% \author{J. de Leon}
% \affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ky, Tokyo 113-0033, Japan}
% 
% %PENDING
% \author[0000-0002-4881-3620]{J. Livingston}
% \affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ky, Tokyo 113-0033, Japan}
%
% %PENDING
% \author[0000-0003-2989-7774]{C. Bergmann}
% \affiliation{Exoplanetary Science at UNSW, School of Physics, UNSW Sydney, NSW 2052, Australia}
% \affiliation{Deutsches Zentrum f\"ur Luft- und Raumfahrt, M\"unchener Str. 20, 82234 Wessling, Germany}

%PENDING
\author[0000-0002-3481-9052]{K. G. Stassun}
\affiliation{Vanderbilt University, Department of Physics \& Astronomy, 6301 Stevenson Center Lane, Nashville, TN 37235, USA}
\affiliation{Fisk University, Department of Physics, 1000 17th Avenue N., Nashville, TN 37208, USA}
%

%
% Princeton team
%
%PENDING
\author[0000-0002-0628-0088]{W. Bhatti}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0001-7204-6727]{G. \'A. Bakos}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%
% SG1 contributors
%

% % begin (ASTEP team)
% %PENDING
% \author{L. Abe} % Lyu
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %%PENDING->OMITTED
% %\author{A. Agabi} % Abdelkrim 
% %\affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author[0000-0001-7866-8738]{N. Crouzet} % Nicolas
% \affiliation{European Space Agency, European Space Research and Technology Centre (ESA/ESTEC), Keplerlaan 1, 2201 AZ Noordwijk, The Netherlands}
% 
% %PENDING
% \author[0000-0002-3937-630X]{G. Dransfield} % Georgina
% \affiliation{School of Physics \& Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
% 
% \author[0000-0002-7188-8428]{T. Guillot} % Tristan 
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author{W. Marie-Sainte} % Wenceslas
% \affiliation{Institut Paul \'{E}mile Victor, Concordia Station, Antarctica}
% 
% %PENDING
% \author{D. M\'ekarnia} % Djamel
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author[0000-0002-5510-8751]{A. H.M.J. Triaud} % Amaury
% \affiliation{School of Physics \& Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
% % end (ASTEP team)

%
% SG2/SG3 contributors
%
\author{J.~Teske} % PFS
\affiliation{Earth and Planets Laboratory, Carnegie Institution for Science, 5241 Broad Branch Road, NW, Washington, DC 20015, USA}

\author{S.~X.~Wang}% PFS
\affiliation{Department of Astronomy, Tsinghua University, Beijing 100084, People's Republic of China}

\author{R. P. Butler}% PFS
\affiliation{Earth and Planets Laboratory, Carnegie Institution for Science, 5241 Broad Branch Road, NW, Washington, DC 20015, USA}

\author{J. D. Crane}% PFS
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\author{S. A. Shectman}% PFS
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

% %PENDING
% \author[0000-0002-7595-0970]{C.~G.~Tinney}
% \affiliation{Exoplanetary Science at UNSW, School of Physics, UNSW Sydney, NSW 2052, Australia}

% %
% % WINE MPIA team
% %
% 
% %PENDING
% \author{T. Henning} %Thomas henning@mpia.de FEROS, WINE board
% \affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }
% 
% %PENDING
% \author[0000-0001-9513-1449]{N. Espinoza} %N\'estor nespinoza@stsci.edu WINE board
% \affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}
% 
% %PENDING
% \author[0000-0002-5389-3944]{A. Jord\'an} %andres.jordan@gmail.com WINE board
% \affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
% \affiliation{Millennium Institute for Astrophysics, Chile}
% 
% %PENDING
% \author{M. Barbieri} %Maruo mauro.barbieri@uda.cl FEROS observations
% \affiliation{INCT, Universidad de Atacama, calle Copayapu 485, Copiap\'o, Atacama, Chile}
% 
% %PENDING
% \author{S. Nandakumar} %sangeetha.nandakumar@postgrados.uda.cl  FEROS observations
% \affiliation{INCT, Universidad de Atacama, calle Copayapu 485, Copiap\'o, Atacama, Chile}
% 
% %PENDING
% \author{T. Trifonov} %Trifon trifonov@mpia.de FEROS observations
% \affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }
% 
% %PENDING
% \author[0000-0002-1896-2377]{J.~I.~Vines} %Jose jose.vines.l@gmail.com FEROS observations
% \affiliation{Departamento de Astronom\'ia, Universidad de Chile, Camino El Observatorio 1515, Las Condes, Santiago, Chile}
% 
% %PENDING
% \author{M. Vuckovic} %Maja maja.vuckovic@uv.cl FEROS observations
% \affiliation{Instituto de F\'isica y Astronom\'ia, Universidad de Vapara\'iso, Casilla 5030, Valpara\'iso, Chile}

%
% SG4 contributors
%
%PENDING
\author[0000-0002-0619-7639]{C.~Ziegler} % carl.ziegler@dunlap.utoronto.ca
\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, Ontario M5S 3H4, Canada}

%%OMITTED
%\author{C.~Brice\~{n}o} % cbriceno@ctio.noao.edu
%\affiliation{Cerro Tololo Inter-American Observatory, Casilla 603, La Serena, Chile}

%PENDING
\author{N.~Law} % nlaw@unc.edu
\affiliation{Department of Physics and Astronomy, The University of North Carolina at Chapel Hill, Chapel Hill, NC 27599-3255, USA}

%PENDING
\author[0000-0003-3654-1602]{A.~W.~Mann} % awmann@unc.edu
\affiliation{Department of Physics and Astronomy, The University of North Carolina at Chapel Hill, Chapel Hill, NC 27599-3255, USA}

%PENDING
\author[0000-0002-2532-2853]{S. B. Howell} % non-detection, but write & inquire
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}

%PENDING
\author[0000-0001-7233-7508]{R. A. Matson}
\affiliation{U.S. Naval Observatory, Washington, DC 20392, USA}
%

% 
%-------------------------------------
% TESS Mission Architects:
% These authors should be listed in this order
% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
%-------------------------------------
%
%PENDING
\author{G. R. Ricker} % grr@space.mit.edu
\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu
\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
%
%PENDING
\author{S. Seager} % seager@mit.edu
\affiliation{Department of Earth, Atmospheric, and Planetary Sciences, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}

%
%-------------------------------------
% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
%These 9 authors should be listed in alphabetical order
%-------------------------------------

% %PENDING
% \author[0000-0002-7754-9486]{C.~J.~Burke} % cjburke@mit.edu
% \affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %PENDING
% \author[0000-0003-2313-467X]{D.~Dragomir}
% \affiliation{Department of Physics and Astronomy, University of New Mexico, Albuquerque, NM, USA}
% 	
% %%PENDING->OMITTED
% %\author[0000-0003-0918-7484]{C.~X.~Huang}
% %\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %%PENDING->OMITTED
% %\author{R.~C.~Kidwell, Jr.} % rkidwell@stsci.edu
% %\affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore MD 21218 }
% 
% %PENDING
% \author[0000-0001-8172-0453]{A.~M.~Levine} % aml@space.mit.edu
% \affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %PENDING
% \author{E.~V.~Quintana} % elisa.quintana@nasa.gov
% \affiliation{NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
% 
% %PENDING
% \author[0000-0001-8812-0565]{J.~E.~Rodriguez}
% \affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
% 
% %PENDING
% \author[0000-0002-6148-7903]{J. C. Smith} % jeffrey.c.smith-1@nasa.gov
% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
% \affiliation{SETI Institute, Mountain View, CA 94043, USA}
% 
% %PENDING
% \author[0000-0002-5402-9613]{B. Wohler} % bill.wohler@nasa.gov
% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
% \affiliation{SETI Institute, Mountain View, CA 94043, USA}

### 2021-01-25 planet_paper/ms.tex
stage cluster_paper and planet_paper

--- 2021-01-25 cluster_paper/ms.tex
@@
 
-\newcommand{\cn}{NGC\,2516} % cluster name
-
-\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+\newcommand{\tn}{TOI~1937} % target star name
+\newcommand{\pn}{TOI~1937b} % planet name
+\newcommand{\cn}{NGC~2516} % cluster name
 
 \newcommand{\kms}{\,km\,s$^{-1}$}
+
+\newcommand{\stscilink}{\textsc{\url{archive.stsci.edu/hlsp/cdips}}}
+\newcommand{\datasetlink}{\textsc{\dataset[doi.org/10.17909/t9-ayd0-k727]{https://doi.org/10.17909/t9-ayd0-k727}}}
 
@@
 \accepted{---}
-\submitjournal{AAS journals.}
-\shorttitle{Core \& Halo of NGC\,2516}
+\submitjournal{TBD.}%AAS journals.}
+\shorttitle{TDB.}
 
@@
   Cluster Difference Imaging Photometric Survey. III.
-  The Coeval Halo and Core of NGC\,2516
+  Subtitle.
 }
@@
 \keywords{
-	stellar ages (1581),
-	young star clusters (1833),
-  open clusters (1160),
-  stellar rotation (1629)
+	Exoplanets (498),
+  Transits (1711),
+	Exoplanet evolution (491),
+	Stellar ages (1581),
+	Young star clusters (1833)
 }
@@
 
-In the traditional picture of star formation molecular clouds go
-gravitationally unstable, and collapse into little knots.  The knots
-produce many stars close to each other.  The resulting ``open star
-clusters'' are, as with any first-order guess, spherical.
+Lorem ipsum.
 
-How does the cluster evolve?  The answer depends
-on the total stellar mass.
-% What exactly does this mean?
-The smallest 90\% of star clusters disperse during the ``embedded
-phase'' (CITE).  This is driven mostly by {\it process A} (CITE).  The
-more massive clusters make it to maybe 100 Myr. Maybe a bit longer.
-Their evaporation is thought to be driven by collisions with molecular
-clouds (CITE, Spitzer 1958), and the galactic tide (CITE), and {\it
-process A}, and {\it process B}.  Of course, the expected evaporation
-time should depend on factors including the mass of the cluster
-itself, how many high-mass (O and B) stars form, since their winds
-[and maybe supernovae] clear out most of the cloud, and the initial
-density of the gas cloud to begin with.
+Section~\ref{sec:observations} describes the identification of the
+candidate, and the follow-up observations that led to its confirmation.
 
-Assuming that the cluster achieves virial equilibrium, stellar fly-bys
-then conspire to segregate the stellar mass distribution within the
-cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
-assuming equipartition of the specific ``thermal energy'' per star,
-then kT = 0.5$mv^2$ at lower masses requires higher velocities).
+The planet can only be understood with respect to its (putative)
+host cluster, so in turn we analyze the
+available six-dimensional positions and kinematics (Section~\ref{sec:gaia6d}),
+the rotation periods of stars in \cn\ 
+(Section~\ref{sec:rotation}), and the available lithium measurements
+(Section~\ref{sec:lithium}).
+We synthesize this data in
+Section~\ref{sec:system} to present our best interpretation of the
+system itself, in turn presenting our interpretation of the
+cluster
+(Section~\ref{subsec:cluster}), the star (Section~\ref{subsec:star})
+and the planet (Section~\ref{subsec:planet}).  We conclude in
+Section~\ref{sec:discussion} by discussing which questions we have
+been able to answer, and which remain open.
 
-Identification of the stars that disperse into the galactic field is
-an important task for understanding the conditions under which stars
-and star clusters form, and for understanding how they subsequently
-evolve.  For instance, how does the process of radial migration across
-the galactic disk affect cluster dispersal?  Did the Sun form in an
-open cluster? If so, how massive was its cluster, and is there any
-hope at identifying the stars that formed near the Sun?
-Qualitatively, the dispersal of open clusters also provides perhaps
-the best ``test case'' for the concept of ``chemical tagging'', also
-referred to as ``galactic archaeology'' (CITE).
+\section{Identification and Follow-up Observations}
+\label{sec:observations}
 
-Outside of the issue of star formation, identifying the remnant halos
-of open clusters is important for a separate project: that of
-discovering young transiting planets.  Young transiting planets are
-hard to find because young stars are rare (CITE), and often reside in
-crowded regions of the sky (CITE).  If the halos of nearby star
-clusters can be reliably identified, this could expand the census of
-nearby sub-Gyr stars by a factor of 2 or even 3 (CITE Meingast).  A
-fortuitous benefit of searching for planets in cluster environments is
-that issues with stellar crowding are also alleviated.
+  \subsection{TESS Photometry}
+  \label{subsec:tess}
+  TESS: S7 + S9
 
-Recent clustering studies using Gaia have begun to report the
-identification of structures that could correspond to low-density
-halos of stars that may have evaporated from open clusters (e.g.,
-Kounkel+19, Kounkel+20, Meingast+21).  However, different clustering
-methods on the Gaia data tend to give different results (CITE: Hunt \&
-Reffert 2020).  Employing say Gaussian Mixture Modelling, or any
-analogous method that requires ``clusters'' to be ellipses in
-spatial/velocity phase-space unsurpsingly yields open clusters that
-are roughly elliptical (CITE: Cantat-Gaudin+18). Unsupervised
-clustering methods such as HDBScan (e.g., Kounkel+19) have been found
-to yield additional structures, particularly in lower density regions
-such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19, Newton+21),
-but also more generally around many open clusters (Kounkel+19).
-Unsupervised approaches that incorporate physical constraints (e.g.,
-imposing a maximum velocity dispersion on putative members) yield
-similar results, potentially with higher purity (Meingast+21).
+  \subsection{Gaia Astrometry and Imaging}
+  \label{subsec:gaia}
+  Gaia:
+    Within 20 arcsec, there are three sources.
 
-We've recently been making TESS light curves of age-dated stars across
-the sky, as part of a Cluster Difference Imaging Photometry Survey
-(CDIPS, CITE Bouma+19).  Our analysis of Cycle 1 (Sectors 1-13)
-yielded light curves of \numsouthernuniqlcs candidate cluster members
-in the Southern Ecliptic hemisphere, available on MAST
-[\href{https://archive.stsci.edu/hlsp/cdips}{1},\href{https://ui.adsabs.harvard.edu/abs/2019ApJS..245...13B/abstract}{2}].
-% TODO: make this number actually quantitative...
-Based on rotation periods, $\approx$25\% appear to be bonafide cluster
-members.
+  % 5489726768531119616 (target, G=13.02, plx=2.38 +/- 0.017 mas, so Bp-Rp=1.00).
+  %   Note RV = 27.77km/s, +/- 9.99 km/s. Quotes E(Bp-Rp) = 0.1925
 
-As part of a broader project of identifying a large and clean sample
-of young stars for a transit search, we take the opportunity of this
-paper to answer a rather modest question: in just a single rich
-southern open cluster, is the cluster halo coeval with the core?  The
-cluster we chose for this analysis was NGC\,2516, since it was young
-($\sim$110\,Myr) and close ($d=400$\,pc) enough to facilitate rotation
-measurements using TESS, and some degree of spectroscopic analysis.
-We want to know: is the halo real? To what extent can we use Gaia
-alone to reliably identify age-dated needles in the haystack of boring
-field stars?  And more generally, what are the implications for the
-evolution of open clusters if they do have halos?
+  % 5489726768531118848 (G=17.59, Bp=17.7, Rp=16.2, Bp-Rp=1.502,...
+  %   but has plx=2.32 +/- 0.118 mas, and VERY SIMILAR proper motions. So, it's a
+  %   binary. Presumably this is the companion that Ziegler found.)
 
-Section~\ref{sec:gaia} presents the astrometric and photometric data
-from Gaia, and clarifies our usage of the terms ``core'' and ``halo''.
-Section~\ref{sec:agedate} age-dates the halo of NGC\,2516, using TESS
-gyrochronology (Section~\ref{sec:rotation}), and lithium depletion
-(Section~\ref{sec:lithium}).  In Section~\ref{sec:discussion} we
-discuss the implications of this analysis for NGC\,2516 specifically
-and stellar spin-down and open cluster evolution generally.
-Section~\ref{sec:conclusion} presents our conclusions.
+  % There's also 5489726768531122560, a G=19.5 companion at further sep.
 
-\section{A 250\,pc Halo around a Core?}
-\label{sec:gaia}
-%
-% What is the core? What is the halo?
-%
+  % ...
 
-\subsection{Gaia Astrometry}
-\label{subsec:astrometry}
+  % Within 30 arcsec, there are... 8 sources. The others don't have parallaxes or
+  % proper motions of significant interest.
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f1.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Reported components of NGC\,2516 in position and
-    velocity space.}
-    The ``core'', identified by \citet{cantatgaudin_gaia_2018} using
-    Gaia DR2, is visually coincident
-    with where you would think the cluster is if you looked at it through a pair
-    of binoculars.
-    The ``halo'' was identified by \citet{kounkel_untangling_2019}
-    using a less restrictive membership assignment algorithm (discussed
-    in the appendices).
-    The ``field'' is a set of randomly drawn and non-overlapping stars
-    within a
-    $(\alpha, \delta, \pi)$ cone centered on the cluster.
-		\label{fig:gaia6d}
-	}
-\end{figure*}
+  \subsection{High-Resolution Imaging}
+  \label{subsec:speckle}
+  Howell et al took some, for me, and had a non-detection.
 
-Figure~\ref{fig:gaia6d} shows the problem we would like to address.
-In this figure, the ``core'' comprises \cn members that
-\citet{cantatgaudin_gaia_2018} reported to have ``membership
-probability'' exceeding 10\%, based on the Gaia DR2 astrometric data.
-The exact meaning of this probability, and details concerning their
-clustering algorithm, are discussed in Appendix~\ref{app:clustering}.
-The ``halo'' comprises \cn members that
-\citet{kounkel_untangling_2019} reported as members, also based on the
-Gaia DR2 data.  That study did not report continuous membership
-probabilities, instead opting for the binary ``member'' or ``not''.
-These two different clustering methods yielded wildly different
-results.  What is the true structure of \cn?
-Are the core and halo truly coeval?
+  Also, Ziegler et al took some.
+    They get the 2.1 arcsec separation, faint companion.
 
-\section{Age-Dating the Halo of NGC\,2516}
-\label{sec:agedate}
+  \subsection{Ground-based Time-Series Photometric Follow-up}
+  \label{subsec:groundphot}
+  LCOGT
+    5 light curves. 3 i band, 1 g band, 1 z band
+  El Sauce
+    1 r band.
 
-\subsection{HR Diagram from Gaia}
+  \subsection{Spectroscopic Follow-up}
+  \label{subsec:spectra}
+
+  \subsubsection{SMARTS 1.5$\,$m / CHIRON}
+  \label{subsec:chiron}
+  CHIRON x1 recon 2020-02-04, another circa 2021.
+
+  \subsubsection{PFS}
+  PFS template + RVs
+
+  %\input{TOI837_rv_table.tex}
+
+\section{Kinematics}
+
+  \subsection{NGC 2516 broadly}
+  \subsection{TOI 1937 specifically}
+
+\section{Rotation}
+
+  \subsection{NGC 2516 broadly}
+  \subsection{TOI 1937 specifically}
+
+\section{Lithium}
+
+  \subsection{NGC 2516 broadly}
+  \subsection{TOI 1937 specifically}
+
+\section{System Modeling}
+\label{sec:system}
+
+\subsection{The Cluster}
+\label{subsec:cluster}
+
+\subsubsection{Physical Characteristics}
+\label{subsec:clusterchar}
+
+\subsubsection{HR Diagram}
 \label{subsec:hr}
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f2a.pdf}
-		}
+\subsection{The Star}
+\label{subsec:star}
 
-		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f2b.pdf}
-		}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf HR diagrams of NGC\,2516, using Gaia EDR3 photometry.}
-    {\it Top:} The core (black) shows a clean sequence consistent with
-    stars with a fixed age and metallicity, and varying mass.  The
-    halo (blue) is similar, but somewhat noisier.  The faintest M
-    dwarfs in the core and halo are brighter than in the field star
-    comparison sample (gray), consistent with these stars having not
-    yet reached the ZAMS.
-    {\it Bottom:} Reported members of the halo, as a function of
-    galactic latitude. Can the additional scatter in the halo be
-    understood through differential reddening?
-    {\bf Maybe}.
-    \label{fig:hr}
-  }
-\end{figure*}
+\subsubsection{Membership of \tn\ in \cn}
+\label{subsec:member}
 
-The first check on whether this membership assignment is plausible was
-already performed by \citet{kounkel_untangling_2019} and more recently
-by \citet{meingast_2021}.  That check is to see whether the HR
-diagrams of these cluster components that were selected based on
-positions and velocities photometrically support the claim that they
-are coeval.
+\subsubsection{Rotation}
 
-Figure~\ref{fig:hr} presents similar results to what these
-investigators have already found.  The core members of the cluster
-show a clean sequence consistent with stars with a fixed age and
-metallicity, and varying mass.
-The halo members are roughly consistent with this, but they do show
-greater scatter.  One possible explanation for this scatter is that
-the halo is more contaminated by field stars.
-Another, explored in the other panel of the figure, is differential
-reddening.
-The halo is reported to span 20$^\circ$ on-sky, and varies in position from
-about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
-galactic plane also being further from the Sun
-by up to 200\,pc (Figure~\ref{fig:gaia6d}).
+\subsubsection{Lithium}
 
-%TODO 
-Comparing these HR diagrams to the PARSEC isochrone models, we find
-that X, Y, and Z.
-In particular, ``the faintest M dwarfs in the core and halo are
-brighter than in the field star comparison sample, consistent with
-these stars having not yet reached the ZAMS.''
-This is consistent with the 
-main-sequence turn-off being at $Bp-Rp\approx0.05$, which implies an
-age of XXX.
-The resulting photometric age we adopt for the core is XXX.
-For the halo, the claimed age from photometry is YYY.
-Applying the same procedure to the field star comparison sample,
-we get an age of ZZZ.
+\subsubsection{Stellar Parameters}
+\label{subsec:starparams}
 
-\subsection{Rotation from TESS}
-\label{subsec:tess}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f3a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f3b.pdf}
-		}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf The core and halo of \cn in the space of rotation
-    period and Gaia color.}
-    The {\it top} plot shows periods against a linear scale, while the
-    {\it bottom}  shows them against a logarithmic scale.
-		\label{fig:rot}
-	}
-\end{figure*}
-
-%TODO: what are these numbers?
-Duly motivated, we collected light curves from TESS photometry
-for YYYY of the ZZZZ cluster members.
-%TODO write the words to describe the procedure, the detrending, the
-%selection function you used for these janky periods, etc.
-
-\subsection{Lithium from Gaia-ESO and GALAH}
-\label{subsec:spectra}
+\subsection{The Planet}
+\label{subsec:planet}
 
@@
 
-  \subsection{How did it form?}
-  \subsection{Mass differences between center and outer reaches?}
-  \subsection{Fast rotators: are we going faster than other clusters?}
-
-\section{Conclusion}
-\label{sec:conclusion}
+The future:
+- Detailed multi-abundance modelling.
+  i.e,. Get spectra of other cluster stars using PFS, and compare the metallicities.
+- 2$\farc$5 separate, multiband imager. (With LCOGT, 2m in Australia,
+  z-band)->Do PSF fitting.
+- K-band image. MagAO? Multicolor imager.
+- Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.
 
@@
 %
-We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
-Latham, Sara Seager, Josh Winn, Jon Jenkins) and the many TESS team
-members for their efforts to make the mission a continued success.
-%
 
+% The ASTEP project acknowledges support from the French and Italian
+% Polar Agencies, IPEV and PNRA, and from Universit\'e C\^ote d'Azur
+% under Idex UCAJEDI (ANR-15-IDEX-01). We thank the dedicated staff at
+% Concordia for their continuous presence and support throughout the
+% Austral winter.
+% %
+% This research received funding from the European Research Council
+% (ERC) under the European Union's Horizon 2020 research and innovation
+% programme (grant n$^\circ$ 803193/BEBOP), and from the
+% Science and Technology Facilities Council (STFC; grant n$^\circ$
+% ST/S00193X/1).
 %
@@
 
+This research was based in part on observations obtained at the
+Southern Astrophysical Research (SOAR) telescope, which is a joint
+project of the Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e
+Inova\c{c}\~{o}es (MCTI/LNA) do Brasil, the US National Science
+Foundation's NOIRLab, the University of North Carolina at Chapel Hill
+(UNC), and Michigan State University (MSU).
+
+This research made use of the Exoplanet Follow-up Observation
+Program website, which is operated by the California Institute of
+Technology, under contract with the National Aeronautics and Space
+Administration under the Exoplanet Exploration Program.
 % %
@@
 
-% Resources supporting this work were provided by the NASA High-End
-% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
-% Division at Ames Research Center for the production of the SPOC data
-% products.
+This research made use of the SVO Filter Profile Service
+(\url{http://svo2.cab.inta-csic.es/theory/fps/}) supported from the Spanish
+MINECO through grant AYA2017-84089.
+
+Resources supporting this work were provided by the NASA High-End
+Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
+Division at Ames Research Center for the production of the SPOC data
+products.
 %
@@
 \software{
-  %\texttt{arviz} \citep{arviz_2019},
+  \texttt{arviz} \citep{arviz_2019},
   \texttt{astrobase} \citep{bhatti_astrobase_2018},
   %\texttt{astroplan} \citep{astroplan2018},
-	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
+	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
   \texttt{astropy} \citep{astropy_2018},
@@
   %\texttt{BATMAN} \citep{kreidberg_batman_2015},
-  %\texttt{ceres} \citep{brahm_2017_ceres},
+  \texttt{ceres} \citep{brahm_2017_ceres},
   \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
@@
   %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
-  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
-  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
-  % 	exoplanet:theano},
+  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
+  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
+  	exoplanet:theano},
 	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
@@
   \texttt{pandas} \citep{mckinney-proc-scipy-2010},
-  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
-  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
-  %\texttt{radvel} \citep{fulton_radvel_2018},
+  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
+  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
+  \texttt{radvel} \citep{fulton_radvel_2018},
   %\texttt{scikit-learn} \citep{scikit-learn},
   \texttt{scipy} \citep{jones_scipy_2001},
-  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
+  \texttt{tesscut} \citep{brasseur_astrocut_2019},
 	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
@@
 	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
-  %PFS ({\bf CITE}),
+  PFS ({\bf CITE}),
   %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
-	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
+	AAT~(Veloce; \citealt{gilbert_veloce_2018}).
  	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
@@
 % 	Danish 1.54m Telescope,
-%	  El Sauce:0.356$\,$m,
+	  El Sauce:0.356$\,$m,
 % 	Elizabeth 1.0m at SAAO,
@@
 % \input{ic2602_ages.tex}
-% \input{starparams.tex}
-% \input{compparams.tex}
+\input{starparams.tex}
+\input{compparams.tex}
 % \input{model_posterior_table.tex}
@@
 
-\appendix
-\section{Clustering method details}
-\label{app:clustering}
-
-%
-% TODO: the following two paragraphs, when quoted, are ENTIRELY FROM
-% the relevant papers. they should probably be nixed in favor of
-% concision.
-%
-\citet{cantatgaudin_gaia_2018} applied a procedure that ultimately
-yielded what we call ``the core''.  Their procedure was to first query
-a Gaia DR2 cone around the previously reported RA and dec of the
-cluster, and within $\pm 0.5$\,mas of its previously reported
-distance. 
-% TODO: which was it?
-The outer radius of their cone was either r2 from MWSC (CITE
-Kharchenko 2013), or twice the ``cluster radius'' listed by Dias.  No
-proper motion cut was applied.  They then applied an unsupervised
-classification scheme called UPMASK to $G<18$\,mag stars within this
-cone (CITE).  The steps of UPMASK are first to perform a k-means
-clustering in the ``astrometric space'' ($\mu_{\alpha'}, \mu_\delta,
-\pi$).  Then, a ``veto'' step is applied to assess whether the groups
-of stars output from the k-means clustering are or are not more
-concentrated than a random distribution.  This is implemented by
-``comparing the total branch length of the minimum spanning tree
-connecting the stars with the expected branch length for a random
-uniform distribution covering the investigated field of view''.  ``To
-turn this yes/no flag to a membership probability,
-\citet{cantatgaudin_gaia_2018} then redraw new values of
-($\mu_{\alpha'}, \mu_\delta, \pi$) for each source based on the listed
-value, uncertainty, and covariance.  After a certain number of
-redrawings, the final probability is the frequency with which a given
-star passes the veto".  In the case of \cn, this yielded a reported
-``\texttt{r50}'' within which half of the cluster members were found
-to be within 0.496$^\circ$.
-
-\citet{kounkel_untangling_2019} applied a different unsupervised
-clustering method to the 5-dimensional Gaia DR2 data (omitting radial
-velocities, due to their sparsity).  Their selection function (see
-their Section~2) yielded $\approx 2\times 10^7$ stars, mostly within
-$\approx 1$\,kpc and typically with $G<18$\,mag.  Their clustering
-algorithm was the ``hierarchical density-based spatial clustering of
-applications with noise'' (HDBSCAN, CITE McInnes17).  The classical
-DBSCAN algorithm ``identifies clusters as overdensities in a
-multi-dimensional space in which the number of sources exceeds the
-required minimum number of points within a neighborhood of a
-particular linking length $\epsilon$.  HDBSCAN does not depend on
-$\epsilon$; instead it condenses the minimum spanning tree by pruning
-off the nodes that do not meet the minimum number of sources in a
-cluster and reanalyzing the nodes that do. Depending on the chosen
-algorithm, it would then either find the most persistent structure
-(through the excess of mass method), or return clusters as the leaves
-of the tree (which results in somewhat more homogeneous clusters). In
-both cases it is more effective at finding structures of varying
-densities in a given data set than DBSCAN.''
-`` The two main parameters that control HDBSCAN are the number of
-sources in a cluster and the number of samples. The former is the
-parameter that rejects groupings that are too small; the latter sets
-the threshold of how conservative the algorithm is in its
-considerations of the background noise (even if the resulting noisy
-groupings do meet the minimum cluster size).  By default, the sample
-size is set to the same value as the cluster size, but it is possible
-to adjust them separately.''
-
 \listofchanges

### 2021-01-25 cluster_paper/collect.sh

@@ -1,0 +1,8 @@
+
+cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
+
+cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f2a.pdf
+cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2b.pdf
+
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear.pdf f3a.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf f3b.pdf

### 2021-01-27 cluster_paper/ms.tex
add words

@@
 
+%TODO:
+%
+% n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
+% n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports
+
 \newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
@@
 \accepted{---}
-\submitjournal{AAS journals.}
+\submitjournal{ApJL.}
 \shorttitle{Core \& Halo of NGC\,2516}
@@
 
+% Context, aims, methods, results, conclusions.
 \begin{abstract}
-  foo,bar.
+  Recent analyses of the Gaia data have reported the existence of
+  diffuse stellar populations (``halos'') surrounding nearby open clusters.  The stars in these
+  halos could have escaped from the cores of the cluster, or they
+  could be remnants of a larger and lower-density star formation
+  complex that has since dispersed.  They could also be field
+  stars---false positives given by overly aggressive clustering algorithms.
+  In this study, we focus on a single reported 250\,pc-long halo
+  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal radius of
+  this cluster is $\sim$XX\,pc.  Combining photometry from Gaia,
+  rotation periods from TESS, and lithium measurements from
+  the Gaia-ESO and GALAH surveys, we find that: i) the halo of
+  NGC\,2516 is real and coeval with the core; ii) at fixed stellar
+  mass and age, rapidly rotating stars have greater lithium abundances
+  than slowly rotating stars; and iii) given Gaia-selected halo
+  members, the contamination rate of field stars in the halo (XX\%) is
+  larger than that in the core (YY\%).  This work expands the set of
+  confirmed NGC\,2516 members by a factor of $\approx$2, and
+  quantifies the degree to which Gaia-based analyses can currently
+  identify dispersed stellar populations against a background of field
+  star contaminants.  Implications for spectroscopic survey targeting, open cluster dispersal, and planet searches around young
+  stars are discussed.
 \end{abstract}
@@
 gravitationally unstable, and collapse into little knots.  The knots
-produce many stars close to each other.  The resulting ``open star
+produce groups of stars close to each other.  The resulting ``open star
 clusters'' are, as with any first-order guess, spherical.
@@
 phase'' (CITE).  This is driven mostly by {\it process A} (CITE).  The
-more massive clusters make it to maybe 100 Myr. Maybe a bit longer.
+more massive clusters make it maybe 100 Myr. Maybe a bit longer.
 Their evaporation is thought to be driven by collisions with molecular
@@
 time should depend on factors including the mass of the cluster
-itself, how many high-mass (O and B) stars form, since their winds
+itself, and how many high-mass (O and B) stars form, since their winds
 [and maybe supernovae] clear out most of the cloud, and the initial
@@
 yielded light curves of \numsouthernuniqlcs candidate cluster members
-in the Southern Ecliptic hemisphere, available on MAST
-[\href{https://archive.stsci.edu/hlsp/cdips}{1},\href{https://ui.adsabs.harvard.edu/abs/2019ApJS..245...13B/abstract}{2}].
+in the Southern Ecliptic hemisphere, available on MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
 % TODO: make this number actually quantitative...
@@
 As part of a broader project of identifying a large and clean sample
-of young stars for a transit search, we take the opportunity of this
-paper to answer a rather modest question: in just a single rich
+of young stars for a transit search, we focus in his
+paper on a rather modest question: in just a single rich
 southern open cluster, is the cluster halo coeval with the core?  The
 cluster we chose for this analysis was NGC\,2516, since it was young
-($\sim$110\,Myr) and close ($d=400$\,pc) enough to facilitate rotation
+($\sim$112\,Myr) and close ($d=400$\,pc) enough to facilitate rotation
 measurements using TESS, and some degree of spectroscopic analysis.
@@
 from Gaia, and clarifies our usage of the terms ``core'' and ``halo''.
-Section~\ref{sec:agedate} age-dates the halo of NGC\,2516, using TESS
-gyrochronology (Section~\ref{sec:rotation}), and lithium depletion
-(Section~\ref{sec:lithium}).  In Section~\ref{sec:discussion} we
+Section~\ref{sec:agedate} age-dates the halo of NGC\,2516,
+based on Gaia photometry (Section~\ref{subsec:hr}),
+TESS gyrochronology (Section~\ref{subsec:tess}), and lithium depletion
+(Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
 discuss the implications of this analysis for NGC\,2516 specifically
@@
 		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f2a.pdf}
+			\includegraphics[width=0.75\textwidth]{f2a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f2b.pdf}
+			\includegraphics[width=0.75\textwidth]{f2b.pdf}
 		}
@@
 age of XXX.
-The resulting photometric age we adopt for the core is XXX.
+The resulting photometric age we calculate for the core is XXX.
 For the halo, the claimed age from photometry is YYY.
@@
 		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f3a.pdf}
+			\includegraphics[width=0.75\textwidth]{f3a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f3b.pdf}
+			\includegraphics[width=0.75\textwidth]{f3b.pdf}
 		}
@@
 
+First, we collected the XXXX Gaia DR2 source\_ids corresponding to the
+core and halo members.  For each source, we first retrieved all
+available CDIPS light curves, on a per-sector basis.  Then, we
+detrended the systematics in each sector individually, and stitched
+together the resulting light curves before searching for the
+periodicity.  Some details regarding our detrending approach are
+discussed in Appendix~\ref{app:detrending}.  After applying a
+detrending step aimed at removing systematic trends, we proceeding
+with a few small cleaning steps aimed at improving the purity of the
+rotation period measurements: we masked 0.7 days at the beginning and
+end of each spacecraft orbit, and ran a sliding standard-deviation
+rejection window over the light curve, which removed any outlying
+points within $\pm3\times$MAD of the median in each window.
+
+We then measured the rotation period of the resulting light curve.  We
+used the the aperture radius that, based on theoretical expectations,
+was expected to give the optimal balance between light from the target
+and background-light (CITE Sullivan15).  This typically resulted in an
+aperture radius of either 1 or 1.5 pixels.  To measure the periods, we
+used the periodogram implementations in \texttt{astrobase}, in
+particular the Stellingwerf PDM periodogram (CITE), along with the
+more traditional Lomb-Scargle (CITE).  We recorded the top five
+periodogram peaks from each method, and their corresponding powers.
+Finally, as a check on crowding, we also recorded the number of stars
+within the aperture of equal brightness, and of brightness with 1.25
+and 2.5 TESS magnitudes of the target star.
+
+Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
+on this plot are light curves for which the peak Lomb Scargle
+periodogram period was below 15 days, had power exceeding 0.08, and
+had at most one equal-brightness companion within the aperture.  These
+selection criteria are entirely heuristic.  
+But splitting the sample into the ``core'' and the ``halo'', they
+suggest that the halo is real.
+
+The resulting gyrochronology age we find for the core is XXX.
+For the halo, the claimed age from gyrochronology is YYY.
+Applying the same procedure to the field star comparison sample,
+we get an age of ZZZ.
+
+\subsubsection{Kinematics $\otimes$ Rotation}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
+    velocity space, cross-matched against the rotators.}
+		\label{fig:gaia6d_x_rotn}
+	}
+\end{figure*}
+
+How far away from the core, in position and velocity space, does the
+halo actually extend?  We can explore this by crossmatching the
+rotator sample against the Gaia data.  Figure~\ref{fig:gaia6d_x_rotn}
+shows the result.
+
 \subsection{Lithium from Gaia-ESO and GALAH}
-\label{subsec:spectra}
+\label{subsec:lithium}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.75\textwidth]{f5a.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.75\textwidth]{f5b.pdf}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Lithium equivalent widths in NGC\,2516,}
+  plotted against extinction-corrected color ({\it top}), and the
+  combination of rotation period and extinction-corrected color ({\it
+  bottom}).
+  \label{fig:lithium}
+  }
+\end{figure*}
+
+Figure~\ref{fig:lithium} shows the results.  At fixed stellar mass
+(and age), it seems that the rapid rotators tend to show elevated
+lithium equivalent widths.  Similar trends have been previously noted
+in the Pleiades by \citet{soderblom_evolution_1993} and
+\citet{bouvier_pleiades_lirot_2018}, in the Psc-Eri stream by
+\citet{arancibia_2020}, and in M\,35 by \citet{jeffries_m35_li_2020}.
+
+% TODO: this paragraph more or less parrots the processes described by
+% Bouvier+20. are there others? is there a better way to describe
+% these?
+The trend contradicts what might be the naive expectation that rapidly
+rotating stars should show more vigorous convection, leading to
+earlier destruction of photospheric Li.  Possible physical
+explanations include processes both internal and external to the star
+(CITE see the recent review by Bouvier+20).  Internal processes would
+be tied to the interplay between surface rotation, differential
+rotation in the interior, and the convective mixing efficiency (e.g.,
+CITE Siess + Livio 1997, Baraffe+2017).  Internal processes could also
+include the possibility that stronger magnetic fields in the stellar
+interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
+Somers + Pinsonneault 2014).  An external process that could also be
+important is the effect of star-disk magnetic braking during the PMS
+phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
+the star's outer convective zone being ``locked'' for longer while the
+radiative core continues contracting.  The resulting differential
+rotation and rotational mixing could drive the lithium depletion
+(CITE: Bouvier 08, Eggenberger+12).
 
@@
 
-  \subsection{How did it form?}
+  \subsection{What is the contamination fraction in the halo? Does it change vs. location?}
+  \subsection{Are the ``very slow rotators'' bonafide members?}
+  Probably not. They are not isotropically distributed around the
+  cluster.
+  \subsection{How did the halo form?}
   \subsection{Mass differences between center and outer reaches?}
@@
 
+\section{Detrending details}
+\label{app:detrending}
+
+In ``detrending'' for our general variability search, our goal was to
+preserve astrophysical variability, while removing systematic
+variability.  One particular concern for the TESS light curves is
+systematic variability at the timescale of the 14-day satellite orbit,
+mostly induced by scattered light from the Earth and Moon.
+
+We therefore turned to the principal components (i.e., the
+eigenvectors) calculated following the procedure described by
+\citet{bouma_cdipsI_2019}. In brief, these vectors are computed using
+a set of ``trend stars'' selected from across each CCD according to
+ad-hoc heuristics that (hopefully) lead them to be dominated by {\it
+systematic} variability (Sec~3.7.2).
+
+The principal component vectors, also referred to as the eigenvectors,
+are rank-ordered by the degree of variance that they predict in the
+training set (of ``trend stars'').
+
+We then posit that any given target star's light curve is described as
+a linear combination of the eigenvectors.  Optionally, we also
+considered the inclusion of additional systematic vectors that could
+affect the light curve, such as the CCD temperature, the flux level
+measured in the background annulus, and the centroid positions of the
+stars on the CCDs.  These can be treated as additional ``features'' in
+the linear model.
+
+There are many methods for determining the coefficients of the linear
+model, after the full set of eigenvectors (plus optionally
+``sytematic'' vectors) has been asssembled.  We explored two: ordinary
+least squares, and ridge regression. Ridge regression is the same as
+ordinary least squares, except it includes an L2 norm with a regularization
+coefficient. The regularization coefficient that best applied for any
+given target light curve was solved for using a cross-validation grid
+search, using \texttt{sklearn.linear\_modelRidgeCV} (CITE).
+
+Each target light curve was mean-subtracted and normalized by its
+standard deviation, as were the eigenvectors. The linear problem was
+then solved, and the light curve was reconstructed by re-adding the
+original mean, and re-multiplying by the standard deviation to ensure
+that the variance of the light curve did not change.
+
+We found, somewhat to our surprise, that the choice of using ordinary
+least squares versus ridge regression did not seem to significantly
+affect the resulting light curves. In other words, the inclusion (or
+lack thereof) of a regularization did not strongly alter the
+best-fitting coefficients.
+
+A few other choices seemed to be more important:
+
+\begin{itemize}
+
+  \item {\it To smooth, or to not smooth the eigenvectors}.
+  Ideally, the eigenvectors should be smooth (and not contain
+    residuals from {\it e.g.}, eclipsing binaries that snuck their way
+    into the template set). However this does not always apply.
+    Sometimes, the eigenvectors are also noisy, which leads them to
+    induce extra variability into the PCA ``detrended'' light curves.
+    To address this problem, we opted to smooth the eigenvectors using
+    a windowed filter (with a "biweight" weight scheme, implemented in
+    \texttt{wotan} by \citet{hippke_wotan_2019}; window length 1 day, cval 6).
+  One issue with this is that systematic sharp features (captured e.g., in
+  "spike vectors") no longer are captured, so they end up in the "PCA
+  detrended" light curves. They can be filtered out relatively easily though
+  (using rolling outlier rejection), and we prefer this approach to having
+  systematic features {\it injected by} the PCA detrending.
+
+  \item {\it How many eigenvectors to use}.
+    A larger number always leads to greater whitening.  In
+    \citet{bouma_cdipsI_2019}, we performed a Factor Analysis
+    cross-validation to determine the number of eigenvectors to use.
+    The typical number adopted based on this analysis was 10--15.
+    Even though this approach was chosen to prevent over-fitting, in
+    our experience, for stellar rotation it still often lead to
+    overfitting, especially for rotation signals with periodicities of
+    $\gtrsim 3$ days.  (Shorter signals typically are not distorted,
+    since the eigenvectors generally do not contain the high-frequency
+    content that leads to the distortions).  For this analysis, we
+    therefore impose the maximum number of eigenvectors to be 5.
+
+  \item {\it Which supplementary systematics vectors to use}.
+    We considered using the \texttt{BGV}, \texttt{CCDTEMP},
+    \texttt{XIC}, \texttt{YIC}, and \texttt{BGV} vectors, packaged
+    with the CDIPS light curves. we found that the background value
+    measured in an annulus centered on the aperture, \texttt{BGV},
+    tended to produce the best independent information from the PCA
+    eigenvectors, and so we adopted it as our only ``supplementary''
+    trend vector.  We opted to not smooth it (in hopes that it would
+    provide direct complement to the smoothed PCA vectors; 1 sharp
+    vector containing literally the background information, plus 5
+    smooth vectors).
+\end{itemize}
+
+For every ``target star'', we then decorrelated the raw
+(image-subtracted and background-subtracted) light curve
+using a linear model with ordinary least squares.
+
+\section{Rotation $\otimes$ RUWE}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Rotation versus color, colored by RUWE.}
+		Looks like on the slow sequence, there's more yellow at the bottom?
+		\label{fig:rotn_X_RUWE}
+	}
+\end{figure*}
+
 \listofchanges

### 2021-01-27 cluster_paper/collect.sh

@@ -1,4 +1,3 @@
-
 cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
 
 cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f2a.pdf
@@ -6,3 +5,17 @@ cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2b.pdf
 
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear.pdf f3a.pdf
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf f3b.pdf
+
+cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
+
+cp ../results/lithium/randich_lithium_vs_BpmRp_xmatch_goldrot_corehalosplit.pdf f5a.pdf
+cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_randich18_lithium_viridis.pdf f5b.pdf
+
+# supplementary
+cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis.pdf f6.pdf
+
+# cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
+# cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf
+# 
+# cp ../results/glue_ngc2516_manual_rotator_viz/slow_rotators_are_lower_declination_and_closer_to_sun.png temp0.png
+# cp ../results/glue_ngc2516_manual_rotator_viz/fast_rotators_are_spread_throughout_core_and_halo.png temp1.png

### 2021-01-27 cluster_paper/ms.tex
more words

@@
 \defcitealias{bouma_wasp4b_2019}{B19}
+\defcitealias{cantatgaudin_gaia_2018}{CG18}
+\defcitealias{kounkel_untangling_2019}{KC19}
 
@@
 	stellar ages (1581),
-	young star clusters (1833),
+  stellar associations (1582),
   open clusters (1160),
+  %stellar dynamics (1596),
   stellar rotation (1629)
@@
 
-How does the cluster evolve?  The answer depends
-on the total stellar mass.
-% What exactly does this mean?
-The smallest 90\% of star clusters disperse during the ``embedded
-phase'' (CITE).  This is driven mostly by {\it process A} (CITE).  The
-more massive clusters make it maybe 100 Myr. Maybe a bit longer.
-Their evaporation is thought to be driven by collisions with molecular
-clouds (CITE, Spitzer 1958), and the galactic tide (CITE), and {\it
-process A}, and {\it process B}.  Of course, the expected evaporation
-time should depend on factors including the mass of the cluster
-itself, and how many high-mass (O and B) stars form, since their winds
-[and maybe supernovae] clear out most of the cloud, and the initial
-density of the gas cloud to begin with.
+How does the cluster evolve?  The answer depends on the total stellar
+mass.  The smallest 90\% of star clusters disperse during the ``embedded
+phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
+by {\it process A} (CITE).  The more massive clusters make it maybe 100
+Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
+collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
+Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
+process B}.  Of course, the expected evaporation time should depend on
+factors including the mass of the cluster itself, and how many high-mass
+(O and B) stars form, since their winds [and maybe supernovae] clear out
+most of the cloud, and the initial density of the gas cloud to begin
+with.
 
@@
 cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
-assuming equipartition of the specific ``thermal energy'' per star,
-then kT = 0.5$mv^2$ at lower masses requires higher velocities).
+assuming equipartition of the specific ``thermal energy'' per star, then
+kT = 0.5$mv^2$ at lower masses requires higher velocities).
 
-Identification of the stars that disperse into the galactic field is
-an important task for understanding the conditions under which stars
-and star clusters form, and for understanding how they subsequently
-evolve.  For instance, how does the process of radial migration across
-the galactic disk affect cluster dispersal?  Did the Sun form in an
-open cluster? If so, how massive was its cluster, and is there any
-hope at identifying the stars that formed near the Sun?
-Qualitatively, the dispersal of open clusters also provides perhaps
-the best ``test case'' for the concept of ``chemical tagging'', also
-referred to as ``galactic archaeology'' (CITE).
+Identification of the stars that disperse into the galactic field is an
+important task for understanding the conditions under which stars and
+star clusters form, and for understanding how they subsequently evolve.
+For instance, how does the process of radial migration across the
+galactic disk affect cluster dispersal?  Did the Sun form in an open
+cluster? If so, how massive was its cluster, and is there any hope at
+identifying the stars that formed near the Sun?  Qualitatively, the
+dispersal of open clusters also provides perhaps the best test case for
+the concept of ``chemical tagging'', aka.\ ``galactic archaeology''
+(CITE).  If you can't chemically tag stars that are kinematically
+associated, why would you be able to do so once they've dispersed
+throughout the galactic disk?
 
-Outside of the issue of star formation, identifying the remnant halos
-of open clusters is important for a separate project: that of
-discovering young transiting planets.  Young transiting planets are
-hard to find because young stars are rare (CITE), and often reside in
-crowded regions of the sky (CITE).  If the halos of nearby star
-clusters can be reliably identified, this could expand the census of
-nearby sub-Gyr stars by a factor of 2 or even 3 (CITE Meingast).  A
-fortuitous benefit of searching for planets in cluster environments is
-that issues with stellar crowding are also alleviated.
+Outside of the issue of star formation, identifying the remnant halos of
+open clusters is important for a separate project: that of discovering
+young transiting planets.  Young transiting planets are hard to find
+because young stars are rare (CITE), and often reside in crowded regions
+of the sky (CITE).  If the halos of nearby star clusters can be reliably
+identified, this could expand the census of nearby sub-Gyr stars by a
+factor of 2 or 3 (CITE).  A fortuitous benefit of searching for planets
+in cluster environments is that issues with stellar crowding are also
+alleviated.
 
 Recent clustering studies using Gaia have begun to report the
-identification of structures that could correspond to low-density
-halos of stars that may have evaporated from open clusters (e.g.,
-Kounkel+19, Kounkel+20, Meingast+21).  However, different clustering
-methods on the Gaia data tend to give different results (CITE: Hunt \&
-Reffert 2020).  Employing say Gaussian Mixture Modelling, or any
-analogous method that requires ``clusters'' to be ellipses in
-spatial/velocity phase-space unsurpsingly yields open clusters that
-are roughly elliptical (CITE: Cantat-Gaudin+18). Unsupervised
-clustering methods such as HDBScan (e.g., Kounkel+19) have been found
-to yield additional structures, particularly in lower density regions
-such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19, Newton+21),
-but also more generally around many open clusters (Kounkel+19).
-Unsupervised approaches that incorporate physical constraints (e.g.,
-imposing a maximum velocity dispersion on putative members) yield
-similar results, potentially with higher purity (Meingast+21).
+identification of structures that could correspond to low-density halos
+of stars that may have evaporated from open clusters (e.g., Kounkel+19,
+Kounkel+20, Meingast+21).  However, different clustering methods on the
+Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
+Employing say Gaussian Mixture Modelling, or any analogous method that
+requires ``clusters'' to be ellipses in phase-space (position or
+velocity) unsurpsingly yields open clusters that are roughly elliptical.
+Some unsupervised clustering methods lead to comparable ``roughly
+elliptical'' results \citep[][hereafter
+\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
+unsupervised methods such as HDBScan (e.g., \citep[][hereafter
+\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
+been found to yield additional structures, particularly in lower density
+regions such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19,
+Newton+21), but also more generally around many open clusters
+\citep{kounkel_untangling_2019}.  Unsupervised approaches that
+incorporate physical constraints like imposing a maximum velocity
+dispersion on putative members yield similar results, potentially with
+higher purity (Meingast+21).
 
+% TODO: make the bonafide number below quantitative
 We've recently been making TESS light curves of age-dated stars across
 the sky, as part of a Cluster Difference Imaging Photometry Survey
-(CDIPS, CITE Bouma+19).  Our analysis of Cycle 1 (Sectors 1-13)
-yielded light curves of \numsouthernuniqlcs candidate cluster members
-in the Southern Ecliptic hemisphere, available on MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
-% TODO: make this number actually quantitative...
-Based on rotation periods, $\approx$25\% appear to be bonafide cluster
-members.
+(CDIPS, CITE Bouma+19).  Our analysis of Cycle 1 (Sectors 1-13) yielded
+light curves of \numsouthernuniqlcs candidate cluster members in the
+Southern Ecliptic hemisphere, available on
+MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.  Based on
+rotation periods, $\approx$25\% appear to be bonafide cluster members.
 
-As part of a broader project of identifying a large and clean sample
-of young stars for a transit search, we focus in his
-paper on a rather modest question: in just a single rich
-southern open cluster, is the cluster halo coeval with the core?  The
-cluster we chose for this analysis was NGC\,2516, since it was young
-($\sim$112\,Myr) and close ($d=400$\,pc) enough to facilitate rotation
-measurements using TESS, and some degree of spectroscopic analysis.
-We want to know: is the halo real? To what extent can we use Gaia
-alone to reliably identify age-dated needles in the haystack of boring
-field stars?  And more generally, what are the implications for the
-evolution of open clusters if they do have halos?
+As part of a broader project of identifying a large and clean sample of
+young stars for a transit search, we focus in his paper on a rather
+modest question: in just a single rich southern open cluster, is the
+cluster halo coeval with the core?  The cluster we chose for this
+analysis was NGC\,2516, since it was young ($\sim$120\,Myr) and close
+($d\approx400$\,pc) enough to facilitate rotation measurements using
+TESS, and some degree of spectroscopic analysis.  We want to know: is
+the halo real? To what extent can we use Gaia alone to reliably identify
+age-dated needles in the haystack of boring field stars?  And more
+generally, what are the implications for the evolution of open clusters
+if they do have halos?
 
-Section~\ref{sec:gaia} presents the astrometric and photometric data
-from Gaia, and clarifies our usage of the terms ``core'' and ``halo''.
-Section~\ref{sec:agedate} age-dates the halo of NGC\,2516,
-based on Gaia photometry (Section~\ref{subsec:hr}),
-TESS gyrochronology (Section~\ref{subsec:tess}), and lithium depletion
+Section~\ref{sec:gaia} presents the astrometric data from Gaia, and
+clarifies our usage of the terms ``core'' and ``halo''.
+Section~\ref{sec:agedate} age-dates the halo of NGC\,2516, based on Gaia
+photometry (Section~\ref{subsec:hr}), TESS gyrochronology
+(Section~\ref{subsec:tess}), and lithium depletion
 (Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
-discuss the implications of this analysis for NGC\,2516 specifically
-and stellar spin-down and open cluster evolution generally.
-Section~\ref{sec:conclusion} presents our conclusions.
+discuss the implications of this analysis for NGC\,2516 specifically and
+stellar spin-down and open cluster evolution generally.
+Section~\ref{sec:conclusion} gives our conclusions.
 
@@
 
-Figure~\ref{fig:gaia6d} shows the problem we would like to address.
-In this figure, the ``core'' comprises \cn members that
-\citet{cantatgaudin_gaia_2018} reported to have ``membership
-probability'' exceeding 10\%, based on the Gaia DR2 astrometric data.
-The exact meaning of this probability, and details concerning their
-clustering algorithm, are discussed in Appendix~\ref{app:clustering}.
-The ``halo'' comprises \cn members that
-\citet{kounkel_untangling_2019} reported as members, also based on the
-Gaia DR2 data.  That study did not report continuous membership
-probabilities, instead opting for the binary ``member'' or ``not''.
-These two different clustering methods yielded wildly different
-results.  What is the true structure of \cn?
-Are the core and halo truly coeval?
+We chose the stars to focus on based on previously reported
+memberships of NGC\,2516, available in the literature.  While a number
+of pre-Gaia catalogs were available, the purity and accuracy of the
+Gaia-derived results are expected to be the current state of the art.
+We therefore adopted what we viewed as the most interesting clustering
+samples to compare: those of \citet{cantatgaudin_gaia_2018}
+(\citetalias{cantatgaudin_gaia_2018}) and
+\citet{kounkel_untangling_2019}
+(\citetalias{kounkel_untangling_2019})\footnote{A third sample was
+also recently reported by \citet{meingast_2021}. Although the results
+of this study were made publicly available too late to be included in
+our analysis,  1577 of the 1860 \citet{meingast_2021} sources (85\%)
+were included in the \citet{kounkel_untangling_2019} sample.  A
+complete set of \cn members could very well include some of the newly
+reported candidate members from \citet{meingast_2021}.}. While we
+could have performed our own clustering analysis based on the Gaia
+data, such an effort would be replicating the work of these
+investigators. And why would our clustering methods be better than
+theirs? We opt instead to use their studies as starting points.
+
+% Theia 613
+\citetalias{cantatgaudin_gaia_2018} reported 1106 candidate members of
+\cn, brighter than $G=18$\,mag.  \citet{kounkel_untangling_2019}
+reported 3003 candidate members, and included stars up to
+$\approx$1\,mag fainter.  Each study used unsupervised clustering
+based on the second Gaia Data Release, detailed in
+Appendix~\ref{app:clustering}.
+
+%TODO calculate these physical sizes carefully
+Figure~\ref{fig:gaia6d} shows the cluster members reported by each
+study.  The \citetalias{cantatgaudin_gaia_2018} members are all within
+a few degrees of the cluster center, while the
+\citetalias{kounkel_untangling_2019} members span tens of degrees.  In
+physical units, this corresponds to a size difference of XX\,pc versus
+YY\,pc.  This difference is somewhat tautological, because
+\citetalias{cantatgaudin_gaia_2018} did not extend their search for
+members out to tens of degrees from the cluster center, since their
+clustering algorithm relied on the contrast between cluster and field
+stars in the astrometric space of ($\mu_{\alpha'}, \mu_\delta, \pi$).
+Nonetheless, the membership catalog of \citetalias{cantatgaudin_gaia_2018}
+echoes that of many previous investigators (CITE, CITE, CITE), and is
+consistent with the general {\it visual} impression that one has when
+looking at \cn visually: it seems to be at most {\bf X$^\circ$}, and
+not a big stream.
+This discrepancy begs the question.  What is the true structure of
+\cn?  Are the core and halo truly coeval?
+
+To answer these questions, we adopt the
+\citetalias{cantatgaudin_gaia_2018} stars as candidate ``core''
+members and the \citetalias{kounkel_untangling_2019} stars as
+candidate ``halo'' members.  If a star is in both lists, it is a
+``core'' member.  1091 of the 1106 core members are in both lists.
+This leaves 1912 candidate members in the halo.
+Concatenating the two lists yields 3018 candidate members.
+
+In this work, we also define a set of nearby field stars in the
+``neighborhood'' of \cn.  Based on the observed distribution of halo
+members, we draw these stars randomly from the following ``cube'' in
+right ascension, declination, and parallax:
+\begin{align}
+  $\alpha$\,[{\rm deg}] &\in [108, 132], \\
+  $\delta$\,[{\rm deg}] &\in [-76, -45], \\
+  $\pi$\,[{\rm mas}] &\in [1.5, 4.0].
+\end{align}
+We imposed a magnitude limit of $G=19$\,mag, and ran the queries
+using the \texttt{astroquery.gaia} module (CITE).
+% TODO: why?
+We allowed the number of stars in the comparison sample to exceed that
+in the cluster sample by a factor of a few, to ensure broad sampling
+of stellar masses and evolutionary states.
+We also required the comparison sample to not overlap with the cluster
+sample, which led to the omission of 1.1\% of the stars drawn over the
+volume noted above.
 
@@
 The first check on whether this membership assignment is plausible was
-already performed by \citet{kounkel_untangling_2019} and more recently
-by \citet{meingast_2021}.  That check is to see whether the HR
-diagrams of these cluster components that were selected based on
-positions and velocities photometrically support the claim that they
-are coeval.
+already performed by \citetalias{cantatgaudin_gaia_2018},
+\citetalias{kounkel_untangling_2019}, and more recently by
+\citet{meingast_2021}.  That check is to see whether the HR diagrams
+of the cluster members support the claim that they are coeval.
 
@@
 %
-\citet{cantatgaudin_gaia_2018} applied a procedure that ultimately
+\citetalias{cantatgaudin_gaia_2018} applied a procedure that ultimately
 yielded what we call ``the core''.  Their procedure was to first query
@@
 to be within 0.496$^\circ$.
+When we selected candidate \cn members from the results of
+\citetalias{cantatgaudin_gaia_2018}, we opted to include all candidate members
+with reported membership probability exceeding 10\%.
+While this seems {\it a priori} low, our results ({\bf SECTION XXX})
+show that this ``membership probabilty'' severely underestimates the
+purity of the \citetalias{cantatgaudin_gaia_2018} sample for \cn. 
+%TODO: quantify
+Their false positive rate across the sample is more like 1-5\%.
 
-\citet{kounkel_untangling_2019} applied a different unsupervised
+\citetalias{kounkel_untangling_2019}
+applied a different unsupervised
 clustering method to the 5-dimensional Gaia DR2 data (omitting radial
@@
 to adjust them separately.''
+Regarding membership proabilities,
+\citetalias{kounkel_untangling_2019} did not report continuous membership
+probabilities, instead opting for the binary ``member'' or ``not''.
 

### 2021-01-27 drivers/plot_full_kinematics.py

@@ -6,7 +6,7 @@ PLOTDIR = os.path.join(RESULTSDIR, 'cluster_membership')
 
 for show1937 in [0,1]:
     for galacticframe in [0,1]:
-        ep.plot_full_kinematics(PLOTDIR, basedata='bright', show1937=show1937, galacticframe=galacticframe) # standard
         ep.plot_full_kinematics(PLOTDIR, basedata='fullfaint', show1937=show1937, galacticframe=galacticframe) # deeper
+        ep.plot_full_kinematics(PLOTDIR, basedata='bright', show1937=show1937, galacticframe=galacticframe) # standard
         ep.plot_full_kinematics(PLOTDIR, basedata='fullfaint_edr3', show1937=show1937, galacticframe=galacticframe) # deeper + EDR3
-    # ep.plot_full_kinematics(PLOTDIR, basedata='extinctioncorrected', show1937=show1937) # NOTE not implemented
+        # ep.plot_full_kinematics(PLOTDIR, basedata='extinctioncorrected', show1937=show1937) # NOTE not implemented

### 2021-01-29 cluster_paper/ms.tex
paper updates

@@
 
+\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
+
 \newcommand{\kms}{\,km\,s$^{-1}$}
@@
   Cluster Difference Imaging Photometric Survey. III.
-  The Coeval Halo and Core of NGC\,2516
+  %The Coeval Core \& Halo of NGC\,2516
+  %The Halo of NGC\,2516 Extends To At Least 250\,pc
+  An Open Cluster With a 250\,pc Halo
+  %NGC\,2516  250\,pc
 }
@@
   halos could have escaped from the cores of the cluster, or they
-  could be remnants of a larger and lower-density star formation
+  could be remnants of a star formation
   complex that has since dispersed.  They could also be field
-  stars---false positives given by overly aggressive clustering algorithms.
+  stars---false positives originating from overly aggressive clustering algorithms.
   In this study, we focus on a single reported 250\,pc-long halo
@@
   rotation periods from TESS, and lithium measurements from
-  the Gaia-ESO and GALAH surveys, we find that: i) the halo of
-  NGC\,2516 is real and coeval with the core; ii) at fixed stellar
-  mass and age, rapidly rotating stars have greater lithium abundances
-  than slowly rotating stars; and iii) given Gaia-selected halo
+  the Gaia-ESO and GALAH surveys, we find that the halo of
+  NGC\,2516 is real and coeval with the core. At fixed stellar
+  mass, the rotation rate and
+  lithium abundance are strongly anti-correlated. And finally, given Gaia-selected halo
   members, the contamination rate of field stars in the halo (XX\%) is
@@
   confirmed NGC\,2516 members by a factor of $\approx$2, and
-  quantifies the degree to which Gaia-based analyses can currently
+  quantifies the degree to which Gaia-based analyses can 
   identify dispersed stellar populations against a background of field
-  star contaminants.  Implications for spectroscopic survey targeting, open cluster dispersal, and planet searches around young
-  stars are discussed.
+  star contaminants.  We highlight the implications for spectroscopic
+  survey targeting, open cluster dispersal, and planet searches around
+  young stars.
 \end{abstract}
@@
 requires ``clusters'' to be ellipses in phase-space (position or
-velocity) unsurpsingly yields open clusters that are roughly elliptical.
+velocity) unsurprisingly yields open clusters that are roughly elliptical.
 Some unsupervised clustering methods lead to comparable ``roughly
@@
 regions such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19,
-Newton+21), but also more generally around many open clusters
-\citep{kounkel_untangling_2019}.  Unsupervised approaches that
-incorporate physical constraints like imposing a maximum velocity
-dispersion on putative members yield similar results, potentially with
-higher purity (Meingast+21).
+Newton+21, but reword bc discovery didn't use HDBScan), but also more
+generally around many open clusters \citep{kounkel_untangling_2019}.
+Unsupervised approaches that incorporate physical constraints like
+imposing a maximum velocity dispersion on putative members yield similar
+results, potentially with higher purity (Meingast+21).
 
@@
 As part of a broader project of identifying a large and clean sample of
-young stars for a transit search, we focus in his paper on a rather
+young stars for a transit search, we focus in this paper on a 
 modest question: in just a single rich southern open cluster, is the
 cluster halo coeval with the core?  The cluster we chose for this
-analysis was NGC\,2516, since it was young ($\sim$120\,Myr) and close
+analysis was NGC\,2516, since it is young ($\sim$120\,Myr) and close
 ($d\approx400$\,pc) enough to facilitate rotation measurements using
@@
 were included in the \citet{kounkel_untangling_2019} sample.  A
-complete set of \cn members could very well include some of the newly
+complete set of \cn\ members could very well include some of the newly
 reported candidate members from \citet{meingast_2021}.}. While we
@@
 consistent with the general {\it visual} impression that one has when
-looking at \cn visually: it seems to be at most {\bf X$^\circ$}, and
+looking at \cn\ visually: it seems to be at most {\bf X$^\circ$}, and
 not a big stream.
@@
 This leaves 1912 candidate members in the halo.
-Concatenating the two lists yields 3018 candidate members.
+Concatenating the two lists yields \nkinematic candidate members.
 
@@
 \begin{align}
-  $\alpha$\,[{\rm deg}] &\in [108, 132], \\
-  $\delta$\,[{\rm deg}] &\in [-76, -45], \\
-  $\pi$\,[{\rm mas}] &\in [1.5, 4.0].
+  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
+  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
+  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
 \end{align}
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf The core and halo of \cn in the space of rotation
+	\caption{ {\bf The core and halo of \cn\ in the space of rotation
     period and Gaia color.}
@@
 Finally, as a check on crowding, we also recorded the number of stars
-within the aperture of equal brightness, and of brightness with 1.25
-and 2.5 TESS magnitudes of the target star.
+within the aperture of equal brightness to the target stars, and of
+brightness with 1.25 and 2.5 TESS magnitudes of the target star.
 
 Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
-on this plot are light curves for which the peak Lomb Scargle
-periodogram period was below 15 days, had power exceeding 0.08, and
-had at most one equal-brightness companion within the aperture.  These
+on this plot are the \nautorot light curves for which the peak Lomb Scargle
+periodogram period was below 15 days, the LS power exceeded 0.08, and
+at most one equal-brightness companion was within the aperture.  These
 selection criteria are entirely heuristic.  
@@
 halo actually extend?  We can explore this by crossmatching the
-rotator sample against the Gaia data.  Figure~\ref{fig:gaia6d_x_rotn}
-shows the result.
+rotator sample against \nkinematic Gaia DR2 kinematic candidate members.
+
+Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
+in interpreting this plot is that we show only stars with
+$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
+detrended light curves.  In other words---the base ``core'' and
+``halo'' samples are the stars for which we could have plausibly
+detected a rotation period, if the star were really a cluster member.
+Faint stars with $Rp>16$ were beyond our light curve selection
+limit, and crowded stars ({\it e.g.,} near saturated stars
+in the cluster center) do not appear.
+We selected the color limit above based on Figure~\ref{fig:rot},
+since the density of period detections seems to decrease once 
+$(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types later
+than $\approx$K4V.
+
+Qualitatively, the fraction of kinematic members with TESS-detection rotation 
+periods with increasing distance from the halo core---both in
+position and in velocity space.
+To quantify this, we computed the distances.
 
@@
 
+The third and final approach we took for age-dating the stellar
+population was using the lithium deplition technique (CITE, CITE,
+CITE).
+
+Reviewing the literature, two spectroscopic datasets for \cn\ seemed
+particularly important: Gaia-ESO 
+\citep{randich_gaiaeso_2018} and GALAH DR3 \citep{buder_galah_2020}.
+The target selection and results from each survey were as follows.
+
+Gaia-ESO selected {\it cluster candidates} to be observed with the
+GIRAFFE and UVES spectrographs based on previously reported literature
+members and publicly available optical and NIR photometry.  This
+target selection approach included many stars in the spatial proximity
+of previously known members.  Since the possible existence of the \cn
+halo was not known at the time of targetting, very few ``halo'' stars
+are in their ultimate sample.  For \cn, \citet{randich_gaiaeso_2018}
+ultimately reported stellar parameters (including lithium equivalent
+widths and metallicities) for 796 candidate members.  Cross-matching
+against our kinematic list of \nkinematic candidate members by position and
+imposing a 0$\farcs$5 maximum separation limit yields 492 kinematic
+members for which a Gaia-ESO spectrum was acquired and stellar
+parameters were reported.  15 of these (all with $(Bp-Rp)_0 > 2.0$)
+are spurious matches based on the Gaia color and GES effective
+temperature, and we remove them.  This yields 477 stars, of which 436
+are in the core, and 41 are in the halo.  This ratio is primarily
+caused by the Gaia-ESO targetting selection function.
+
+% GALAH DR3 selected XXX/YYY.  GALAH-DR3 = GALAH-main + GALAH-faint +
+% K2-HERMES + TESS-HERMES + HERMES other.  Data taken both with the 2dF
+% fiber positioner, and the HERMES spectrograph at the 3.9 AAT
+% (Lewis+02, Sheinis+15).  GALAH-main is 12<V<14, $\delta<+10^\circ$,
+% $|b|>10^\circ$, in regions of sky with at least 400 targets in $\pi$
+% square degrees (the 2dF FoV).  Fields with more than 400 targets were
+% targetted multiple times, with different fiber arrangements.
+% TESS-HERMES: 10.0 < V < 13.1. TESS Southern CVZ, within $12^\circ$ of
+% the southern ecliptic pole.  54354 in the ``HERMES other'' program are
+% from targeted observations of stars in open clusters, the GALAH Pilot
+% Survey (Martell et al. 2017),
+
+The GALAH DR3 target selection process is discussed in detail by
+\citet{buder_galah_2020}.  The most relevant aspects for our projects
+are that the ``main'' survey targeted $12.0<V<14.0$ stars at
+$\delta<+10^\circ$, provided they were at least ten degrees from the
+galactic plane.  The survey is spatially inhomogeneous\footnote{See
+the footprint at
+\url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
+targetting for stars in the TESS southern continuous viewing zone, and
+for known open cluster members was also performed.  We identified the
+candidate \cn\ members
+for which spectra had been obtained by searching
+the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after excluding
+stars with the stellar parameter bit flags 1, 2, 3.  This excludes
+spectra with unreliable broadening, low S/N, and unreliable wavelength
+solutions (see Table~6 of \citealt{buder_galah_2020}).  Since our main
+focus is measuring equivalent widths for the Li\,6708\ \AA line, these
+were the most relevant bit flags for our purpose.  Of our \nkinematic
+candidate \cn\ members, 106 had spectra in GALAH DR3.  51 were in the
+``core''; 55 were in the ``halo''.
+
+%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
+% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
+We downloaded the GALAH DR3 spectra for all 106 entries by doing PROCESS X.
+We then measured the lithium equivalent widths using PROCESS Y.
+We then concatenated the GALAH DR3 and Gaia-ESO results together.
+
+% TODO: fix the fact that the figure currently only shows Randich+18
 Figure~\ref{fig:lithium} shows the results.  At fixed stellar mass
-(and age), it seems that the rapid rotators tend to show elevated
-lithium equivalent widths.  Similar trends have been previously noted
-in the Pleiades by \citet{soderblom_evolution_1993} and
-\citet{bouvier_pleiades_lirot_2018}, in the Psc-Eri stream by
-\citet{arancibia_2020}, and in M\,35 by \citet{jeffries_m35_li_2020}.
+(and age), the rapid rotators tend to show elevated lithium equivalent
+widths. This effect is mostly apparent in the K dwarfs.  Similar
+trends in the Pleiades were noted nearly three decades ago by
+\citet{soderblom_evolution_1993}.  More recent studies have found
+comparable results in the Pleiades, the Psc-Eri stream, and M\,35
+\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
+The fact that we see the same trend across the core and halo of \cn
+{\it a)} supports the conclusion that the halo is coeval with the
+core, and {\it b)} suggests that the lithium-rotation correlation is
+not explained by environmental effects such as the density of the cluster
+during its embedded phase, but is instead tied to physical processes
+acting on all low-mass pre-main sequence stars.
 
@@
 % these?
-The trend contradicts what might be the naive expectation that rapidly
-rotating stars should show more vigorous convection, leading to
-earlier destruction of photospheric Li.  Possible physical
-explanations include processes both internal and external to the star
-(CITE see the recent review by Bouvier+20).  Internal processes would
-be tied to the interplay between surface rotation, differential
-rotation in the interior, and the convective mixing efficiency (e.g.,
-CITE Siess + Livio 1997, Baraffe+2017).  Internal processes could also
-include the possibility that stronger magnetic fields in the stellar
+Both internal and external processes have been suggested to explain
+the lithium-rotation trend (CITE see the recent review by Bouvier+20).
+One explanation based in the stellar interior could be that the
+convective mixing efficiency is anticorrelated with the surface
+rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
+possibility could be that stronger magnetic fields in the star's
 interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
-Somers + Pinsonneault 2014).  An external process that could also be
-important is the effect of star-disk magnetic braking during the PMS
+Somers + Pinsonneault 2014).  An external process that might also be
+important is the effect of star-disk magnetic locking during the PMS
 phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
-the star's outer convective zone being ``locked'' for longer while the
-radiative core continues contracting.  The resulting differential
-rotation and rotational mixing could drive the lithium depletion
-(CITE: Bouvier 08, Eggenberger+12).
+the star's outer convective zone being locked for longer while the
+radiative core contracts.  The resulting differential rotation and
+rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
+Eggenberger+12).
+We have no preference between these possibilities---we simply note
+that the \cn\ sample could be another helpful data point in
+distinguishing them.
 
@@
 
+\subsection{How did the halo form?}
+``The stars in these halos could have escaped from the cores of the
+cluster, or they could be remnants of a larger and lower-density star
+formation complex that has since dispersed.''
+Another possibility is {\it triggered star formation}.
+
+\subsection{Are the ``very slow rotators'' bonafide members?}
+Unsure. They are not isotropically distributed around the
+cluster... so either triggered star formation (CITE, CITE), or they're
+actually field stars.
+
 \subsection{What is the contamination fraction in the halo? Does it change vs. location?}
-  \subsection{Are the ``very slow rotators'' bonafide members?}
-  Probably not. They are not isotropically distributed around the
-  cluster.
-  \subsection{How did the halo form?}
+
 \subsection{Mass differences between center and outer reaches?}
-  \subsection{Fast rotators: are we going faster than other clusters?}
+
+\subsection{Fast and blue rotators: are we going faster than other clusters?}
 
@@
 \label{sec:conclusion}
+
+We analyzed X, Y, Z.
+Our main results are as follows.
+\begin{itemize}
+	\item 
+\end{itemize}
 
@@
 %
-\citetalias{cantatgaudin_gaia_2018} applied a procedure that ultimately
+\citetalias{cantatgaudin_gaia_2018} applied a procedure that 
 yielded what we call ``the core''.  Their procedure was to first query
@@
 The outer radius of their cone was either r2 from MWSC (CITE
-Kharchenko 2013), or twice the ``cluster radius'' listed by Dias.  No
+Kharchenko 2013), or twice the ``cluster radius'' listed by DAML/Dias (CITE).  No
 proper motion cut was applied.  They then applied an unsupervised
@@
 to be within 0.496$^\circ$.
-When we selected candidate \cn members from the results of
+When we selected candidate \cn\ members from the results of
 \citetalias{cantatgaudin_gaia_2018}, we opted to include all candidate members
@@
 $\approx 1$\,kpc and typically with $G<18$\,mag.  Their clustering
-algorithm was the ``hierarchical density-based spatial clustering of
+algorithm, which was run on this entire  stellar sample,
+ was the ``hierarchical density-based spatial clustering of
 applications with noise'' (HDBSCAN, CITE McInnes17).  The classical
@@
 
-There are many methods for determining the coefficients of the linear
-model, after the full set of eigenvectors (plus optionally
-``sytematic'' vectors) has been asssembled.  We explored two: ordinary
+To determine the coefficients of the linear
+model after the full set of eigenvectors (plus optionally
+``sytematic'' vectors) had been asssembled,  we explored two possible
+methods: ordinary
 least squares, and ridge regression. Ridge regression is the same as
@@
 standard deviation, as were the eigenvectors. The linear problem was
-then solved, and the light curve was reconstructed by re-adding the
+then solved numerically, and the light curve was reconstructed by re-adding the
 original mean, and re-multiplying by the standard deviation to ensure
@@
 
-We found, somewhat to our surprise, that the choice of using ordinary
+We found that the choice of using ordinary
 least squares versus ridge regression did not seem to significantly
 affect the resulting light curves. In other words, the inclusion (or
-lack thereof) of a regularization did not strongly alter the
+lack thereof) of a regularization term did not strongly alter the
 best-fitting coefficients.
+In the spirit of ``KISS'', we opted to use ordinary least squares.
 
@@
   \item {\it To smooth, or to not smooth the eigenvectors}.
-  Ideally, the eigenvectors should be smooth (and not contain
+  Ideally, the eigenvectors should be smooth in time. They should not contain
     residuals from {\it e.g.}, eclipsing binaries that snuck their way
-    into the template set). However this does not always apply.
-    Sometimes, the eigenvectors are also noisy, which leads them to
+    into the template set, and they should also not be intrinsically noisier
+    than the target star. If either of these is the case (and we found that
+    it sometimes was), it can
     induce extra variability into the PCA ``detrended'' light curves.
     To address this problem, we opted to smooth the eigenvectors using
-    a windowed filter (with a "biweight" weight scheme, implemented in
+    a time-windowed filter (with a "biweight" weight scheme, implemented in
     \texttt{wotan} by \citet{hippke_wotan_2019}; window length 1 day, cval 6).
@@
   "spike vectors") no longer are captured, so they end up in the "PCA
-  detrended" light curves. They can be filtered out relatively easily though
-  (using rolling outlier rejection), and we prefer this approach to having
-  systematic features {\it injected by} the PCA detrending.
+  detrended" light curves. They can then be filtered out
+  ({\it e.g.}, using rolling outlier rejection), and we prefer this approach to having
+  systematic features being {\it injected by} the PCA detrending.
 
@@
     The typical number adopted based on this analysis was 10--15.
-    Even though this approach was chosen to prevent over-fitting, in
+    While this approach should in theory prevent over-fitting, in
     our experience, for stellar rotation it still often lead to
-    overfitting, especially for rotation signals with periodicities of
+    distorts the signals, especially for rotation signals with small amplitudes
+    and periodicities of
     $\gtrsim 3$ days.  (Shorter signals typically are not distorted,
-    since the eigenvectors generally do not contain the high-frequency
-    content that leads to the distortions).  For this analysis, we
+    since the eigenvectors do not contain the high-frequency
+    content that leads to the distortions).  For the present analysis, we
     therefore impose the maximum number of eigenvectors to be 5.
@@
     \texttt{XIC}, \texttt{YIC}, and \texttt{BGV} vectors, packaged
-    with the CDIPS light curves. we found that the background value
+    with the CDIPS light curves. We found that the background value
     measured in an annulus centered on the aperture, \texttt{BGV},
@@
 
+Figure~\ref{fig:rotn_X_RUWE}.
+
+\section{Kinematics $\otimes$ Rotation in EDR3}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f7.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
+    velocity space, cross-matched against the rotators. Analog of
+    Figure~\ref{fig:gaia6d_x_rotn}, but showing EDR3 kinematics.}
+		\label{fig:gaia6d_x_rotn_EDR3}
+	}
+\end{figure*}
+
+How did we do?
+Figure~\ref{fig:gaia6d_x_rotn_EDR3} lets us check with a more precise
+astrometric solution.
+
 \listofchanges

### 2021-01-29 cluster_paper/collect.sh

@@ -6,13 +6,16 @@ cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2b.pdf
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear.pdf f3a.pdf
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf f3b.pdf
 
-cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
+#cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
+cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_galactic.pdf f4.pdf
 
 cp ../results/lithium/randich_lithium_vs_BpmRp_xmatch_goldrot_corehalosplit.pdf f5a.pdf
 cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_randich18_lithium_viridis.pdf f5b.pdf
 
 # supplementary
 cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis.pdf f6.pdf
+
+cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_edr3_galactic.pdf f7.pdf
 
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf

### 2021-01-30 cluster_paper/ms.tex
update paper

@@
   complex that has since dispersed.  They could also be field
-  stars---false positives originating from overly aggressive clustering algorithms.
+  stars---the false positives of overly aggressive clustering algorithms.
   In this study, we focus on a single reported 250\,pc-long halo
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.75\textwidth]{f8a.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.75\textwidth]{f8b.pdf}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
+    velocity space, cross-matched against the rotators, but now in
+    physical units.}
+    Top: relevant projections. Bottom: binned versus 3d position and
+    2d velocity difference.
+		\label{fig:physical_x_rotn}
+	}
+\end{figure*}
+
 How far away from the core, in position and velocity space, does the
@@
 ``halo'' samples are the stars for which we could have plausibly
-detected a rotation period, if the star were really a cluster member.
+detected a rotation period.
 Faint stars with $Rp>16$ were beyond our light curve selection
@@
 
-Qualitatively, the fraction of kinematic members with TESS-detection rotation 
-periods with increasing distance from the halo core---both in
+Qualitatively, the fraction of kinematic members with TESS-detected rotation 
+periods seems like it may increase with increasing distance from the halo core---both in
 position and in velocity space.
-To quantify this, we computed the distances.
+To quantify this, we computed the relelvant distances in physical coordinates.
+This entailed first transforming from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) which we did 
+assuming the AstroPy v4.0 coordinate standard (CITE).
+We then converted the on-sky proper motions to physical units by dividing by the distance of each star.
+We then computed the SEPARATIONS of all of these positions and
+2d-velocities from the ``median core member'', which was selected by
+taking \citetalias{cantatgaudin_gaia_2018} members with membership
+probability exceeding 70\%.
+Again, we only considered stars for which we could have plausibly
+detected a rotation period --- the same color and observability limits
+discussed above were also applied.
+
+The results are shown in Figure~\ref{fig:physical_x_rotn}.
+The halo extends to about 250\,pc in physical space, and out to at
+least 10 of km/s in tangential velocity space.
+\citet{meingast_2021} by comparison required a physically motivated cut
+in tangential velocity space of 1.5 km per sec.
+This probably improves purity, but clearly doesn't get everthing for
+completeness.
+
+% [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0   0]
+% [495 116  65  46  28  20  22  12  11   8   5   1   0   1   0   2]
+% [384 109  45  27  15   4   0   0   0   0   0   0   0   0   0   0   0   0
+%    0   0]
+% [533 153  67  40  23  10   3   1   2   0   0   0   0   0   0   0   0   0
+%    0   0]
 

### 2021-01-30 cluster_paper/collect.sh

@@ -9,6 +9,9 @@ cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf
 #cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
 cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_galactic.pdf f4.pdf
 
+cp ../results/physical_X_rotation/physical_X_rotation_fullfaint.pdf f8a.pdf
+cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f8b.pdf
+
 cp ../results/lithium/randich_lithium_vs_BpmRp_xmatch_goldrot_corehalosplit.pdf f5a.pdf
 cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_randich18_lithium_viridis.pdf f5b.pdf
 

### 2021-02-02 cluster_paper/ms.tex
paper updates

@@
   Recent analyses of the Gaia data have reported the existence of
-  diffuse stellar populations (``halos'') surrounding nearby open clusters.  The stars in these
-  halos could have escaped from the cores of the cluster, or they
-  could be remnants of a star formation
-  complex that has since dispersed.  They could also be field
-  stars---the false positives of overly aggressive clustering algorithms.
-  In this study, we focus on a single reported 250\,pc-long halo
-  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal radius of
-  this cluster is $\sim$XX\,pc.  Combining photometry from Gaia,
-  rotation periods from TESS, and lithium measurements from
-  the Gaia-ESO and GALAH surveys, we find that the halo of
-  NGC\,2516 is real and coeval with the core. At fixed stellar
-  mass, the rotation rate and
-  lithium abundance are strongly anti-correlated. And finally, given Gaia-selected halo
-  members, the contamination rate of field stars in the halo (XX\%) is
-  larger than that in the core (YY\%).  This work expands the set of
-  confirmed NGC\,2516 members by a factor of $\approx$2, and
-  quantifies the degree to which Gaia-based analyses can 
-  identify dispersed stellar populations against a background of field
-  star contaminants.  We highlight the implications for spectroscopic
-  survey targeting, open cluster dispersal, and planet searches around
-  young stars.
+  diffuse stellar populations surrounding nearby open clusters.  The
+  stars in these ``halos'' could have escaped from the cores of the
+  cluster, or they could be remnants of a star formation complex that
+  has since dispersed.  They could also be field stars---false
+  positives from overly aggressive clustering algorithms.  In this
+  study, we focus on a 250\,pc halo reported
+  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
+  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
+  from Gaia, rotation periods from TESS, and lithium measurements from
+  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
+  coeval with the core.  Two in three kinematically-selected halo
+  members 25-250\,pc from the cluster center show rotation periods
+  consistent with a gyrochronological age of 120\,Myr.  Stars in the
+  inner 25\,pc show higher period detection rates than the
+  halo (73\% vs.\ 66\%), but both vastly exceed our field star
+  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
+  are also anti-correlated at fixed stellar mass, as is the case with
+  comparable clusters.  This work expands the set of confirmed
+  NGC\,2516 members by a factor of $\approx$2, and quantifies the
+  degree to which Gaia-based analyses can identify dispersed stellar
+  populations against a background of field star contaminants.  We
+  highlight the implications for spectroscopic survey targeting, open
+  cluster dispersal, and planet searches around young stars.
 \end{abstract}
+% Rotators < 25pc : 0.72929 +/- 0.00013
+% Rotators > 25pc : 0.66172 +/- 0.00023
 
@@
 
+{\bf Emphasize that figure 1 does not clearly tell us whether or not the halo is really distinct from the field, by eye.}
+
 \begin{figure*}[t]
@@
     2d velocity difference.
+    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
+    velocity.
 		\label{fig:physical_x_rotn}
@@
 halo actually extend?  We can explore this by crossmatching the
-rotator sample against \nkinematic Gaia DR2 kinematic candidate members.
+rotator sample against \nkinematic Gaia DR2 kinematic candidate
+members.
 
@@
 $0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
-detrended light curves.  In other words---the base ``core'' and
-``halo'' samples are the stars for which we could have plausibly
-detected a rotation period.
-Faint stars with $Rp>16$ were beyond our light curve selection
-limit, and crowded stars ({\it e.g.,} near saturated stars
-in the cluster center) do not appear.
-We selected the color limit above based on Figure~\ref{fig:rot},
-since the density of period detections seems to decrease once 
-$(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types later
-than $\approx$K4V.
+detrended light curves.  In other words---the base sample is the stars
+for which we could have plausibly detected a rotation period.  Faint
+stars with $Rp>16$ were beyond our light curve selection limit, and
+crowded stars ({\it e.g.,} near saturated stars in the cluster center)
+do not appear.  We selected the color limit above based on
+Figure~\ref{fig:rot}, since the density of period detections seems to
+decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types
+later than $\approx$K4V.
+Our expectation is that completion effects become important for
+fainter stars.
+At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
+for which our reductions of the TESS images yield a precision of
+$\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
+{\bf 
+The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
+REBULL 2020, FIGURE 10).
+}
+
+%G=18 is like Abs G 9.8, Bp-Rp not corrected of 2.7. 
+% (Bp-Rp )_0 of 1.25 is (Bp-Rp) of 1.38... or Abs G of ~= 6.8.
+% or G-Rp ~= 0.7
+% So, G~=15... meaning Rp~=14.3.
+
+% IN POSITION
+% Bin width: 25
+% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
+% 0   0   0 0   0]
+% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
+% 0   2   0   0 0   0]
+% Rotators < 25pc : 0.72929 +/- 0.00013
+% Rotators > 25pc : 0.66172 +/- 0.00023
+% ... where numerator and denom are 223/337
 

### 2021-02-03 cluster_paper/ms.tex
update paper

@@
 
+\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
 \newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
-\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
+
+\newcommand{\ncalibration}{9{,}619{,}784} % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
+\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
+\newcommand{\nnbhdcaliblc}{10{,}314\ } % make_compstar_NGC_2516_sourcelist.py
+\newcommand{\nnbhdcalibstar}{2{,}212\ } % make_compstar_NGC_2516_sourcelist.py
 
@@
   Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
-  coeval with the core.  Two in three kinematically-selected halo
+  coeval with the core.  Two in three kinematically selected halo
   members 25-250\,pc from the cluster center show rotation periods
@@
 We allowed the number of stars in the comparison sample to exceed that
-in the cluster sample by a factor of a few, to ensure broad sampling
+in the cluster sample by a factor of $\approx$5, to ensure broad sampling
 of stellar masses and evolutionary states.
@@
 		\leavevmode
-		\subfloat{
 			\includegraphics[width=0.75\textwidth]{f2a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f2b.pdf}
-		}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf HR diagrams of NGC\,2516, using Gaia EDR3 photometry.}
-    {\it Top:} The core (black) shows a clean sequence consistent with
+  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
+    The core (black) shows a clean sequence consistent with
     stars with a fixed age and metallicity, and varying mass.  The
-    halo (blue) is similar, but somewhat noisier.  The faintest M
+    halo (blue) is similar, though somewhat noisier.  The faintest M
     dwarfs in the core and halo are brighter than in the field star
@@
     yet reached the ZAMS.
-    {\it Bottom:} Reported members of the halo, as a function of
-    galactic latitude. Can the additional scatter in the halo be
-    understood through differential reddening?
-    {\bf Maybe}.
     \label{fig:hr}
@@
 show a clean sequence consistent with stars with a fixed age and
-metallicity, and varying mass.
-The halo members are roughly consistent with this, but they do show
-greater scatter.  One possible explanation for this scatter is that
-the halo is more contaminated by field stars.
-Another, explored in the other panel of the figure, is differential
-reddening.
-The halo is reported to span 20$^\circ$ on-sky, and varies in position from
-about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
-galactic plane also being further from the Sun
-by up to 200\,pc (Figure~\ref{fig:gaia6d}).
+metallicity, and varying mass.  The halo members show a roughly
+consistent main sequence, but with somewhat greater scatter.  One
+possible explanation for this scatter is that the halo is more
+contaminated by field stars.  Another possible explanation is
+differential reddening.  The halo is reported to span 20$^\circ$
+on-sky, and varies in position from about $b=-12^\circ$ to
+$b=-20^\circ$, with the stars closest to the galactic plane also being
+further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
+Empirical evidence for differential reddening is discussed further in
+Appendix~\ref{app:diffred}. A priori, both field star contamination
+and differential reddening could well play a role.
 
@@
 we get an age of ZZZ.
+
+\paragraph{Comparison sample}
+To provide a basis for comparison, we also opted to search the
+``calibration'' light curves ($Rp<13$) that were created as a part of
+the CDIPS project. 
+Over the southern sky (Sectors 1-13 of TESS), this corresponded to a sample of \ncalibration stars.
+Cross-matching these against \nnbhd randomly drawn stars in the
+neighborhood of \cn yielded \nnbhdcalibstar unique stars, with
+\nnbhdcaliblc sector-specific TESS light curves.
 
@@
 
+\section{Differential Reddening}
+\label{app:diffred}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+			\includegraphics[width=0.75\textwidth]{f2b.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
+    {\it Bottom:} Reported members of the halo, as a function of
+    galactic latitude. Can the additional scatter in the halo be
+    understood through differential reddening?
+    {\bf Maybe}.
+    \label{fig:diffred}
+  }
+\end{figure*}
+
+Figure~\ref{fig:diffred}.
+
 \section{Rotation $\otimes$ RUWE}
+\label{app:ruwe}
 

### 2021-02-05 cluster_paper/ms.tex
update paper

@@
 
-\newcommand{\ncalibration}{9{,}619{,}784} % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
+\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
 \newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
@@
   populations against a background of field star contaminants.  We
-  highlight the implications for spectroscopic survey targeting, open
+  highlight implications for spectroscopic survey targeting, open
   cluster dispersal, and planet searches around young stars.

### 2021-02-05 cluster_paper/ms.tex
update paper

@@
 
-\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
+\newcommand{\nautorot}{1{,}144\ } % % depends on which cuts are adopted. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+
 \newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
@@
   %The Coeval Core \& Halo of NGC\,2516
-  %The Halo of NGC\,2516 Extends To At Least 250\,pc
-  An Open Cluster With a 250\,pc Halo
-  %NGC\,2516  250\,pc
+  %The Halo of NGC\,2516 Extends To At Least 300\,pc
+  An Open Cluster With a 300\,pc Halo
+  %NGC\,2516  300\,pc
 }
@@
   positives from overly aggressive clustering algorithms.  In this
-  study, we focus on a 250\,pc halo reported
+  study, we focus on a 300\,pc halo reported
   around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
@@
 
-\section{A 250\,pc Halo around a Core?}
+\section{A 300\,pc Halo around a Core?}
 \label{sec:gaia}
@@
 periodogram period was below 15 days, the LS power exceeded 0.08, and
-at most one equal-brightness companion was within the aperture.  These
+for which no equal-brightness companions were within the aperture.  These
 selection criteria are entirely heuristic.  
@@
 The results are shown in Figure~\ref{fig:physical_x_rotn}.
-The halo extends to about 250\,pc in physical space, and out to at
-least 10 of km/s in tangential velocity space.
-\citet{meingast_2021} by comparison required a physically motivated cut
-in tangential velocity space of 1.5 km per sec.
-This probably improves purity, but clearly doesn't get everthing for
-completeness.
+The halo extends to separations of about 200\,pc in physical space
+from the cluster core.
+This corresponds to a length of $\approx$300-350\,pc, depending on
+which members of the halo are chosen as the ``tips'' on either
+end.
+
+In tangential velocity space, the fraction of stars with rotation
+period detections remains high out to roughly 10\,km/s.
+\citet{meingast_2021} by comparison required a physically motivated
+cut in tangential velocity space of 1.5 km per sec.  This probably
+improves purity, but clearly doesn't get everthing for completeness.
 
@@
   bottom}).
+  {\bf TODO: Do this against the autorot sample; Add the GALAH EWs}
   \label{fig:lithium}

### 2021-02-09 cluster_paper/ms.tex
update paper

@@
 
-\newcommand{\nautorot}{1{,}144\ } % % depends on which cuts are adopted. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\nautorot}{987\ } % % For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
 
@@
   around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
-  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
+  tidal radius of the cluster is $\sim$10\,pc.  Using photometry
   from Gaia, rotation periods from TESS, and lithium measurements from
@@
 identified, this could expand the census of nearby sub-Gyr stars by a
-factor of 2 or 3 (CITE).  A fortuitous benefit of searching for planets
+factor of 2 or 3 (CITE).  
+One path to identify these young stars is to combine photometric
+selection criteria with spectroscopic observations (e.g., CITE Zerjal
+2019, 2020).
+A fortuitous benefit of searching for planets
 in cluster environments is that issues with stellar crowding are also
@@
 Recent clustering studies using Gaia have begun to report the
-identification of structures that could correspond to low-density halos
-of stars that may have evaporated from open clusters (e.g., Kounkel+19,
-Kounkel+20, Meingast+21).  However, different clustering methods on the
+identification of structures that could correspond to low-density
+halos of stars that may have evaporated from open clusters (e.g.,
+Kounkel+19, Kounkel+20, Meingast+21).  There is evidence that in some
+cases, these halos correspond to tidal tails, e.g., in the Hyades
+(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
+group (CITE Gagne+20).  However, different clustering methods on the
 Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
@@
 requires ``clusters'' to be ellipses in phase-space (position or
-velocity) unsurprisingly yields open clusters that are roughly elliptical.
-Some unsupervised clustering methods lead to comparable ``roughly
-elliptical'' results \citep[][hereafter
+velocity) unsurprisingly yields open clusters that are roughly
+elliptical.  Some unsupervised clustering methods lead to comparable
+``roughly elliptical'' results \citep[][hereafter
 \citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
@@
 \citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
-been found to yield additional structures, particularly in lower density
-regions such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19,
-Newton+21, but reword bc discovery didn't use HDBScan), but also more
-generally around many open clusters \citep{kounkel_untangling_2019}.
-Unsupervised approaches that incorporate physical constraints like
-imposing a maximum velocity dispersion on putative members yield similar
-results, potentially with higher purity (Meingast+21).
+been found to yield additional structures, particularly in lower
+density regions such as the Psc-Eri stream (CITE: Meingast+18,
+Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
+also more generally around many open clusters
+\citep{kounkel_untangling_2019}.  Unsupervised approaches that
+incorporate physical constraints like imposing a maximum velocity
+dispersion on putative members yield similar results, potentially with
+higher purity (Meingast+21).
 
@@
 Appendix~\ref{app:clustering}.
+%FIXME 
+\citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
+XX DEGREES} centered on the cluster's position.
+Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
+Piskunov+2008, which is 8$^\circ$ at the cluster distance of
+$\approx$400\,pc, perhaps they should have considered stars further
+out.
 
@@
 Another possibility is {\it triggered star formation}.
+
+One relevant connection may be to the ``Mamajek-2'' stellar group
+(see Jilinski+2009).
 

### 2021-02-09 planet_paper/ms.tex
minor
paper updates
update paper
planet paper: minor notes

@@
 \title{
-  Cluster Difference Imaging Photometric Survey. III.
-  Subtitle.
+  Cluster Difference Imaging Photometric Survey. IV.
+  TOI 1937Ab. Young Hot Jupiter, or Just Tidally Spun Up?
 }
@@
   \subsection{TOI 1937 specifically}
+  If you back-integrate 20 Myr, it gets closer. But then, it gets
+  further.
 
@@
 
+\pararaph{Metallicity}
+(Quoting Jilinski+09)
+``After an initial period where several authors (see Terndrup et al.
+2002) found for NGC 2516 a metallicity of a few tenths dex below the
+solar, more recent analysis (Irwin et al. 2007; Jeffries et al.  2001;
+Sciortino et al. 2001; Sung et al. 2002) places NGC 2516 with a
+metallicity close to solar. ''
+
 \subsubsection{Physical Characteristics}
 \label{subsec:clusterchar}
+
+\paragraph{Mass} 1.8Mjup
+
+\paragraph{Obliquity} Less than 20 deg or so. Consistent with
+expectations given its short orbital period (e.g., Anderson et al
+2021).
 
@@
 
-The future:
+\subsection{The present}
+
+The minimum orbital period beyond which a planet is tidally sheared
+apart depends on the concentration of mass within the planet
+(Rappaport+2013).
+For a VERY CENTRALLY CONCENTRATED planet with \tn's
+mean density of 0.XX g/cm3, this period is XX.X hours.
+For the limit of a uniform density distribution, the period would be
+XX.X hours.
+
+is
+
+XX.X hr (e.g., Rappaport+ 2013).
+Given ,
+
+The Roche limit
+
+Roche limit for a close-orbiting planet can
+be expressed as a minimum orbital period depending
+on the density distribution of the planet (Rappaport
+et al. 2013). For WASP-12b, with a mean density of
+0.46 g cm􀀀3, the Roche-limiting orbital period is 14.2 hr
+assuming the mass of the planet to be concentrated near
+the center, and 18.6 hr in the opposite limit of a spher-
+ical and incompressible planet. The
+
+\subsection{The future}:
 - Detailed multi-abundance modelling.

### 2021-02-09 cluster_paper/ms.tex
minor paper update

@@
 The GALAH DR3 target selection process is discussed in detail by
-\citet{buder_galah_2020}.  The most relevant aspects for our projects
+\citet{buder_galah_2020}.  The most relevant aspects for our project
 are that the ``main'' survey targeted $12.0<V<14.0$ stars at
-$\delta<+10^\circ$, provided they were at least ten degrees from the
+$\delta<+10^\circ$, provided the stars were at least ten degrees from the
 galactic plane.  The survey is spatially inhomogeneous\footnote{See
@@
 \url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
-targetting for stars in the TESS southern continuous viewing zone, and
+targeting for stars in the TESS southern continuous viewing zone, and
 for known open cluster members was also performed.  We identified the
@@
 ``core''; 55 were in the ``halo''.
+77 had ``finite'' lithium, i.e., no lithium flag set.
 
@@
 % https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
-We downloaded the GALAH DR3 spectra for all 106 entries by doing PROCESS X.
+We downloaded\footnote{Via \url{datacentral.org.au/services/download},
+using the \texttt{sobject_id} identifiers.}
+the GALAH DR3 spectra for all 106 entries.
 We then measured the lithium equivalent widths using PROCESS Y.

### 2021-02-10 cluster_paper/ms.tex
update paper

@@
 %\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
-\newcommand{\nautorot}{987\ } % % For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
 
@@
 \newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
-\newcommand{\nnbhdcaliblc}{10{,}314\ } % make_compstar_NGC_2516_sourcelist.py
-\newcommand{\nnbhdcalibstar}{2{,}212\ } % make_compstar_NGC_2516_sourcelist.py
+\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
+\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py
+
+\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
+\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
+\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
+\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
+\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
+\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py
 
@@
   diffuse stellar populations surrounding nearby open clusters.  The
-  stars in these ``halos'' could have escaped from the cores of the
+  stars in these halos could have escaped from the cores of the
   cluster, or they could be remnants of a star formation complex that
@@
   positives from overly aggressive clustering algorithms.  In this
-  study, we focus on a 300\,pc halo reported
-  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
-  tidal radius of the cluster is $\sim$10\,pc.  Using photometry
-  from Gaia, rotation periods from TESS, and lithium measurements from
-  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
-  coeval with the core.  Two in three kinematically selected halo
-  members 25-250\,pc from the cluster center show rotation periods
-  consistent with a gyrochronological age of 120\,Myr.  Stars in the
-  inner 25\,pc show higher period detection rates than the
-  halo (73\% vs.\ 66\%), but both vastly exceed our field star
-  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
-  are also anti-correlated at fixed stellar mass, as is the case with
-  comparable clusters.  This work expands the set of confirmed
-  NGC\,2516 members by a factor of $\approx$2, and quantifies the
-  degree to which Gaia-based analyses can identify dispersed stellar
-  populations against a background of field star contaminants.  We
-  highlight implications for spectroscopic survey targeting, open
-  cluster dispersal, and planet searches around young stars.
+  study, we focus on a halo with a length of 300\,pc reported around
+  the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal
+  radius of the cluster is $\sim$10\,pc.  Using photometry from Gaia,
+  rotation periods from TESS, and lithium measurements from Gaia-ESO
+  and GALAH, we find that the halo of NGC\,2516 is real and coeval
+  with the core.  Two in three kinematically selected halo members
+  25-250\,pc from the cluster center show rotation periods consistent
+  with a gyrochronological age of 120\,Myr.  A comparison sample of
+  field stars shows no such trend.  The rotation period and lithium
+  abundance for the cluster stars are also anti-correlated at fixed
+  stellar mass, as is the case with comparable stellar populations.
+  This work expands the set of confirmed NGC\,2516 members by a factor
+  of $\approx$2, and quantifies the degree to which Gaia-based
+  analyses can identify dispersed stellar populations against a
+  background of field star contaminants.  We highlight implications
+  for spectroscopic survey targeting, open cluster dispersal, and
+  planet searches around young stars.
 \end{abstract}
-% Rotators < 25pc : 0.72929 +/- 0.00013
-% Rotators > 25pc : 0.66172 +/- 0.00023
 
@@
 
+\subsubsection{Cluster Star Sample}
+
 \begin{figure*}[t]
@@
 Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
-on this plot are the \nautorot light curves for which the peak Lomb Scargle
+on this plot are the \nautorotnumerator light curves for which the peak Lomb Scargle
 periodogram period was below 15 days, the LS power exceeded 0.08, and
-for which no equal-brightness companions were within the aperture.  These
-selection criteria are entirely heuristic.  
-But splitting the sample into the ``core'' and the ``halo'', they
+for which no equal-brightness or greater companions were within the aperture. 
+We also required that at most one companion with brightness exceeding 
+one tenth of the target star could reside in the aperture.
+These
+selection criteria are entirely heuristic, and yielded
+\nautorotdenominator stars for which rotation periods might have been
+detected.  
+By exploring the different possible selection criteria, we found that
+the ones we chose provided a
+reasonable balance between excluding stars on the slow sequence, {\it
+vs.} reporting spurious rotation periods in the ``void'' below the
+slow sequence, where the dominant signal was in fact from a
+neighboring star.
+Splitting the resulting sample into the ``core'' and the ``halo'', they
 suggest that the halo is real.
@@
 
-\paragraph{Comparison sample}
+\subsubsection{Comparison Star Sample}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+    \includegraphics[width=0.49\textwidth]{f9.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Field stars show a different rotation period
+  distribution than the kinematically selected members.}
+		\label{fig:compstar}
+	}
+\end{figure}
+
 To provide a basis for comparison, we also opted to search the
 ``calibration'' light curves ($Rp<13$) that were created as a part of
-the CDIPS project. 
-Over the southern sky (Sectors 1-13 of TESS), this corresponded to a sample of \ncalibration stars.
-Cross-matching these against \nnbhd randomly drawn stars in the
-neighborhood of \cn yielded \nnbhdcalibstar unique stars, with
-\nnbhdcaliblc sector-specific TESS light curves.
+the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
+corresponded to a sample of \ncalibration stars.  Cross-matching these
+against the \nnbhd randomly drawn stars in the neighborhood of \cn yielded
+\nnbhdcalibstar unique stars, with a cumulative total of \nnbhdcaliblc TESS
+sectors observed.
+The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
+corresponds to an extinction-corrected color cutoff of $(Bp-Rp)_0
+\approx 0.80$, or spectral types of $\approx$G1V.
+This reaches sufficiently far down the main sequence to enable a
+comparison to the cluster star sample.
+
+We performed the same light curve stitching and period-search
+procedure on the field comparison stars.
+Imposing the same requirements for crowding resulted in
+\ncompstardenominator stars for which rotation periods could have been
+detected.
+Imposing the same Lomb Scargle power cutoff, and period upper limit,
+yielded \ncompstarnumerator period detections (\ncompfrac).
+Within the same brightness cutoff, 
+\nautovscompstarnumerator of \nautovscompstardenominator cluster stars
+yielded period detections (\nautofrac).
+Though the detection fractions are frankly not very different
+(likely because of the brightness cutoff), the period {\it
+vs} color distributions are quite different
+(Figure~\ref{fig:compstar}).
 
@@
 We downloaded\footnote{Via \url{datacentral.org.au/services/download},
-using the \texttt{sobject_id} identifiers.}
+using the \texttt{sobject\_id} identifiers.}
 the GALAH DR3 spectra for all 106 entries.

### 2021-02-10 cluster_paper/collect.sh

@@ -20,6 +20,9 @@ cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis_linear.pdf f6.pdf
 
 cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_edr3_galactic.pdf f7.pdf
 
+cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear.pdf f9.pdf
+
+
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf
 # 

### 2021-02-11 cluster_paper/ms.tex
update paper

@@
 % n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
+% ... which came from ~/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split/OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv
 % n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports
@@
 \newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\ncdips}{2,238\ } % e.g., from OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv. checked with plot_target_vs_actual_list_sanity_checks.py
 
@@
 
-%TODO: what are these numbers?
-Duly motivated, we collected light curves from TESS photometry
-for YYYY of the ZZZZ cluster members.
-%TODO write the words to describe the procedure, the detrending, the
-%selection function you used for these janky periods, etc.
-
-First, we collected the XXXX Gaia DR2 source\_ids corresponding to the
-core and halo members.  For each source, we first retrieved all
-available CDIPS light curves, on a per-sector basis.  Then, we
-detrended the systematics in each sector individually, and stitched
-together the resulting light curves before searching for the
+First, we collected the \nkinematic Gaia DR2 source\_ids corresponding to the
+kinematically-identified core and halo members.  For each source, we first retrieved all
+available CDIPS light curves, on a per-sector basis. 
+This yielded
+\ncdips light curves, all brighter than $Rp=16$\,mag.
+The $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
+roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
+the distance of NGC\,2516.
+
+We cleaned and stitched the light curves as follows.
+First, we detrended the systematics in each sector individually, and
+stitched together the resulting light curves before searching for the
 periodicity.  Some details regarding our detrending approach are

### 2021-02-11 drivers/measure_galahdr3_Li_EWs.py

@@ -26,7 +26,7 @@ df = pd.read_csv(inpath)
 #
 # get the available spectra
 #
-workingdir = os.path.join(DATADIR, 'galah_dr3_spectra', 'EXAMPLE', 'galah',
+workingdir = os.path.join(DATADIR, 'galah_dr3_spectra', 'galah',
                           'dr3', 'spectra', 'hermes')
 
 red_specpaths = glob(os.path.join(workingdir, '*3.fits'))
@@ -60,7 +60,12 @@ for _, r in mdf.iterrows():
     assert len(specpath) == 1
     specpath = specpath[0]
 
-
     outpath = os.path.join(OUTDIR,
                            f'gaiadr2_{dr2_source_id}_GALAHDR3_Li6708_EW.png')
-    get_Li_6708_EW(specpath, outpath=outpath)
+
+    LiI_a = 6707.76
+    LiI_b = 6707.91
+    Li_avg = (LiI_a + LiI_b)/2
+
+    xshift = 6708.30 - Li_avg
+    get_Li_6708_EW(specpath, xshift=xshift, outpath=outpath)

### 2021-02-11 cluster_paper/ms.tex
lithium: minor plot updates

@@
   field stars shows no such trend.  The rotation period and lithium
-  abundance for the cluster stars are also anti-correlated at fixed
+  abundances of the cluster stars are also anti-correlated at fixed
   stellar mass, as is the case with comparable stellar populations.

### 2021-02-11 cluster_paper/collect.sh

@@ -12,8 +12,8 @@ cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_ga
 cp ../results/physical_X_rotation/physical_X_rotation_fullfaint.pdf f8a.pdf
 cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f8b.pdf
 
-cp ../results/lithium/randich_lithium_vs_BpmRp_xmatch_goldrot_corehalosplit.pdf f5a.pdf
-cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_randich18_lithium_viridis.pdf f5b.pdf
+cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit.pdf f5a.pdf
+cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_lithium_viridis_gaiaeso_galahdr3.pdf f5b.pdf
 
 # supplementary
 cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis_linear.pdf f6.pdf

### 2021-02-11 drivers/plot_lithium.py

@@ -8,7 +8,8 @@ if not os.path.exists(PLOTDIR):
 
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=0)
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=0, galahdr3=1)
-ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1)
+ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=0)
+ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1)
 
 ep.plot_galah_dr3_lithium_abundance(PLOTDIR)
 ep.plot_galah_dr3_lithium_abundance(PLOTDIR, corehalosplit=1)

### 2021-02-11 earhart/plotting.py

@@ -61,6 +61,8 @@ from earhart.physicalpositions import (
 )
 from earhart.lithium import _get_lithium_EW_df
 
+from earhart.priors import TEFF, P_ROT, AVG_EBpmRp
+
 def plot_TIC268_nbhd_small(outdir=RESULTSDIR):
 
     basedata = 'bright'
@@ -837,8 +839,6 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
 
 def plot_rotation(outdir, BpmRp=0, include_ngc2516=0, ngc_core_halo=0):
 
-    from earhart.priors import TEFF, P_ROT, AVG_EBpmRp
-
     set_style()
 
     rotdir = os.path.join(DATADIR, 'rotation')
@@ -964,8 +964,6 @@ def plot_rotation(outdir, BpmRp=0, include_ngc2516=0, ngc_core_halo=0):
 
 
 def plot_skypositions_x_rotn(outdir):
-
-    from earhart.priors import AVG_EBpmRp
 
     runid = "NGC_2516"
     rot_df, lc_df = get_autorotation_dataframe(runid='NGC_2516', returnbase=True)
@@ -1268,8 +1266,6 @@ def plot_galah_dr3_lithium_abundance(outdir, corehalosplit=0):
     # make tha plot 
     ##########
 
-    from earhart.priors import TEFF, P_ROT, AVG_EBpmRp
-
     set_style()
 
     plt.close('all')
@@ -1383,7 +1379,6 @@ def plot_randich_lithium(outdir, vs_rotators=1, corehalosplit=0):
         'epoch_photometry_url', 'subcluster']
     """
 
-    from earhart.priors import AVG_EBpmRp
     assert abs(AVG_EBpmRp - 0.1343) < 1e-4 # used by KC19
 
     set_style()
@@ -1486,7 +1481,7 @@ def plot_randich_lithium(outdir, vs_rotators=1, corehalosplit=0):
         savefig(f, outpath)
 
 
-def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0):
+def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0, corehalosplit=0):
     """
     Plot Li EW vs color for Randich+18's Gaia ESO lithium stars, crossmatched
     against the "fullfaint kinematic" sample.
[…]

### 2021-02-12 cluster_paper/ms.tex
update paper

@@
   Recent analyses of the Gaia data have reported the existence of
-  diffuse stellar populations surrounding nearby open clusters.  The
-  stars in these halos could have escaped from the cores of the
+  diffuse stellar populations surrounding nearby open clusters. 
+  The stars in these halos could have escaped from the cores of the
   cluster, or they could be remnants of a star formation complex that
-  has since dispersed.  They could also be field stars---false
-  positives from overly aggressive clustering algorithms.  In this
-  study, we focus on a halo with a length of 300\,pc reported around
-  the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal
-  radius of the cluster is $\sim$10\,pc.  Using photometry from Gaia,
-  rotation periods from TESS, and lithium measurements from Gaia-ESO
-  and GALAH, we find that the halo of NGC\,2516 is real and coeval
-  with the core.  Two in three kinematically selected halo members
-  25-250\,pc from the cluster center show rotation periods consistent
-  with a gyrochronological age of 120\,Myr.  A comparison sample of
-  field stars shows no such trend.  The rotation period and lithium
-  abundances of the cluster stars are also anti-correlated at fixed
-  stellar mass, as is the case with comparable stellar populations.
+  has since dispersed.
+  They could also be field stars---the false positives of overly
+  aggressive clustering algorithms.
+  In this study, we focus on a halo with a length of 300\,pc reported
+  around the $\approx$120\,Myr open cluster NGC\,2516.
+  The classical tidal radius of the cluster is $\sim$10\,pc.
+  Combining photometry from Gaia, rotation periods from TESS, and
+  lithium measurements from Gaia-ESO and GALAH, we find that the halo
+  of NGC\,2516 is real and coeval with the core.
+  Two in three kinematically selected halo members 25-250\,pc from the
+  cluster center show rotation periods consistent with a
+  gyrochronological age of 120\,Myr.
+  A comparison sample of field stars shows no such trend.
+  The rotation period and lithium abundances of the cluster stars are
+  also anti-correlated at fixed stellar mass, as is the case with
+  comparable stellar populations.
   This work expands the set of confirmed NGC\,2516 members by a factor
@@
   analyses can identify dispersed stellar populations against a
-  background of field star contaminants.  We highlight implications
-  for spectroscopic survey targeting, open cluster dispersal, and
-  planet searches around young stars.
+  background of field star contaminants.
+  We highlight the implications for spectroscopic survey targeting,
+  open cluster dispersal, and planet searches around young stars.
 \end{abstract}
@@
 
-In the traditional picture of star formation molecular clouds go
-gravitationally unstable, and collapse into little knots.  The knots
-produce groups of stars close to each other.  The resulting ``open star
-clusters'' are, as with any first-order guess, spherical.
-
-How does the cluster evolve?  The answer depends on the total stellar
-mass.  The smallest 90\% of star clusters disperse during the ``embedded
-phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
-by {\it process A} (CITE).  The more massive clusters make it maybe 100
-Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
-collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
-Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
-process B}.  Of course, the expected evaporation time should depend on
-factors including the mass of the cluster itself, and how many high-mass
-(O and B) stars form, since their winds [and maybe supernovae] clear out
-most of the cloud, and the initial density of the gas cloud to begin
-with.
-
-Assuming that the cluster achieves virial equilibrium, stellar fly-bys
-then conspire to segregate the stellar mass distribution within the
-cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
-assuming equipartition of the specific ``thermal energy'' per star, then
-kT = 0.5$mv^2$ at lower masses requires higher velocities).
+Star clusters form in overdensities of hierarchically structured
+molecular clouds (CITE).  Over the first $\approx$10\,Myr,
+``feedback'' effects including protostellar outflows, photoionization,
+radiation pressure, and supernova shocks conspire to disperse the gas
+out of the cloud (CITE Krumholz+19).  Since only a small fraction of
+the cloud mass is converted into stars ($\approx 1\%$; see Krumholz+19
+Figure 10), this gas dispersal drives the majority ($\sim$90\%) of the
+stars in the cluster to become gravitationally unbound (e.g., Lada \&
+Lada 03).
+
+From 10 to 1000\,Myr, the portion of the cluster that survives gas
+dispersal suffers an onslaught that almost always succeeds in
+destroying the cluster.  Internal to the cluster, mass loss from
+supernovae and AGB winds, and close stellar encounters
+(``relaxation'') are important over 10s and 100s of megayears,
+respectively.  External to the cluster, collisions with GMCs (Spitzer
+1958), and perturbations from the galactic tide in both the radial
+(CITE) and vertical (CITE) dimensions further drive stars toward the
+``high velocity tails'' of the stellar energy distribution.
+
+% In the traditional picture of star formation molecular clouds go
+% gravitationally unstable, and collapse into little knots.  The knots
+% produce groups of stars close to each other.  The resulting ``open star
+% clusters'' are, as with any first-order guess, spherical.
+
+% How does the cluster evolve?  The answer depends on the total stellar
+% mass.  The smallest 90\% of star clusters disperse during the ``embedded
+% phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
+% by {\it process A} (CITE).  The more massive clusters make it maybe 100
+% Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
+% collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
+% Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
+% process B}.  Of course, the expected evaporation time should depend on
+% factors including the mass of the cluster itself, and how many high-mass
+% (O and B) stars form, since their winds [and maybe supernovae] clear out
+% most of the cloud, and the initial density of the gas cloud to begin
+% with.
+% 
+% Assuming that the cluster achieves virial equilibrium, stellar fly-bys
+% then conspire to segregate the stellar mass distribution within the
+% cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
+% assuming equipartition of the specific ``thermal energy'' per star, then
+% kT = 0.5$mv^2$ at lower masses requires higher velocities).
 
 Identification of the stars that disperse into the galactic field is an
-important task for understanding the conditions under which stars and
+important aspect of understanding the conditions under which stars and
 star clusters form, and for understanding how they subsequently evolve.
@@
 
-Outside of the issue of star formation, identifying the remnant halos of
-open clusters is important for a separate project: that of discovering
-young transiting planets.  Young transiting planets are hard to find
-because young stars are rare (CITE), and often reside in crowded regions
-of the sky (CITE).  If the halos of nearby star clusters can be reliably
-identified, this could expand the census of nearby sub-Gyr stars by a
-factor of 2 or 3 (CITE).  
-One path to identify these young stars is to combine photometric
-selection criteria with spectroscopic observations (e.g., CITE Zerjal
-2019, 2020).
-A fortuitous benefit of searching for planets
-in cluster environments is that issues with stellar crowding are also
-alleviated.
-
-Recent clustering studies using Gaia have begun to report the
-identification of structures that could correspond to low-density
-halos of stars that may have evaporated from open clusters (e.g.,
-Kounkel+19, Kounkel+20, Meingast+21).  There is evidence that in some
+Outside of the issue of star formation, identifying the remnant
+halos\footnote{ The low-density outer regions of open clusters have
+been referred to in general terms as ``halos'' and ``coronae''; we
+adopt the former (CITE, CITE). } of open clusters is important for a
+separate project: that of discovering young transiting planets.  Young
+transiting planets are hard to find because young stars are rare
+(e.g., CITE Piskunov 2018), and often reside in crowded regions of the
+sky (e.g., CITE ANY OPEN CLUSTER CATALOG).  If the dispersed halos of
+nearby star clusters could be reliably identified in a complete
+manner, this could expand the census of nearby star clusters by up to
+a factor of $10$, based on the expected fraction of stars thought to
+be lost during gas dispersal.  Path to identifying these young stars
+include photometric criteria (CITE Zari+18), spectroscopic criteria
+(CITE Berger+18), and combinations thereof (e.g., CITE Zerjal 2019,
+2020, Zhou's work).  A benefit of searching for planets in
+dispersed stellar environments is that issues with stellar crowding
+are also alleviated.
+
+The identification of low-density (GIVE A NUMBER DENSITY) stellar
+associations, already possible before Gaia (e.g.,
+Zuckerman+ CITE), has reached a breakthrough pace with the Gaia data.
+Some low-density populations (``moving groups'') have no discernable
+cores (e.g., USco CITE, Psc Eri CITE, mu Tau CITE).
+However other
+low-density
+halos seem to be centered on previously reported open clusters (e.g.,
+Kounkel+19, Kounkel+20, Meingast+21).  In some
 cases, these halos correspond to tidal tails, e.g., in the Hyades
 (CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
-group (CITE Gagne+20).  However, different clustering methods on the
+group (CITE Gagne+20). 
+However, different clustering methods on the
 Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).

### 2021-02-16 cluster_paper/ms.tex
update paper

@@
   The stars in these halos could have escaped from the cores of the
-  cluster, or they could be remnants of a star formation complex that
-  has since dispersed.
+  cluster, or they could be dispersed remnants of the initial star formation complex.
   They could also be field stars---the false positives of overly
@@
   around the $\approx$120\,Myr open cluster NGC\,2516.
+  %TODO: says who?
   The classical tidal radius of the cluster is $\sim$10\,pc.
@@
 
-Star clusters form in overdensities of hierarchically structured
-molecular clouds (CITE).  Over the first $\approx$10\,Myr,
-``feedback'' effects including protostellar outflows, photoionization,
-radiation pressure, and supernova shocks conspire to disperse the gas
-out of the cloud (CITE Krumholz+19).  Since only a small fraction of
-the cloud mass is converted into stars ($\approx 1\%$; see Krumholz+19
-Figure 10), this gas dispersal drives the majority ($\sim$90\%) of the
-stars in the cluster to become gravitationally unbound (e.g., Lada \&
-Lada 03).
+Star clusters form within hierarchically structured molecular clouds
+(CITE).  Over the first $\approx$10\,Myr, ``feedback'' effects
+including protostellar outflows, photoionization, radiation pressure,
+and supernova shocks disperse the gas out of the cloud (CITE
+Krumholz+19).  Since only a small fraction of the cloud mass is
+converted into stars ($\approx 1\%$; see Krumholz+19 Figure 10), the
+gas dispersal drives the majority ($\approx$90\%) of stars in the
+cluster to become gravitationally unbound (e.g., Lada \& Lada 03).
 
 From 10 to 1000\,Myr, the portion of the cluster that survives gas
-dispersal suffers an onslaught that almost always succeeds in
-destroying the cluster.  Internal to the cluster, mass loss from
-supernovae and AGB winds, and close stellar encounters
-(``relaxation'') are important over 10s and 100s of megayears,
-respectively.  External to the cluster, collisions with GMCs (Spitzer
-1958), and perturbations from the galactic tide in both the radial
-(CITE) and vertical (CITE) dimensions further drive stars toward the
-``high velocity tails'' of the stellar energy distribution.
+dispersal suffers an onslaught that typically unbinds the cluster.
+Internal to the cluster, mass loss from supernovae and AGB winds, and
+close stellar encounters (``relaxation'') are important over 10s and
+100s of megayears, respectively (CITE, CITE).  External to the
+cluster, collisions with GMCs (Spitzer 1958), and perturbations from
+the galactic tide in both the radial (CITE) and vertical (CITE)
+dimensions further drive stars toward the ``high velocity tails'' of
+the stellar energy distribution.
 
@@
 
-Identification of the stars that disperse into the galactic field is an
-important aspect of understanding the conditions under which stars and
-star clusters form, and for understanding how they subsequently evolve.
-For instance, how does the process of radial migration across the
-galactic disk affect cluster dispersal?  Did the Sun form in an open
-cluster? If so, how massive was its cluster, and is there any hope at
-identifying the stars that formed near the Sun?  Qualitatively, the
-dispersal of open clusters also provides perhaps the best test case for
-the concept of ``chemical tagging'', aka.\ ``galactic archaeology''
-(CITE).  If you can't chemically tag stars that are kinematically
-associated, why would you be able to do so once they've dispersed
-throughout the galactic disk?
-
-Outside of the issue of star formation, identifying the remnant
+Identifying the stars that have dispersed from their natal clusters
+into the galactic field is of interest for a few reasons.  One 
+reason is understanding the predominant conditions under which stars and star
+clusters form (e.g., Wright Mamajek 2018 ScoCen), as well as
+the processes that drive cluster dissolution (including gas dispersal,
+internal and external heating mechanisms, and radial migration across
+the galactic disk).
+Another is to understand the Sun's birth conditions---
+did it form in an open
+cluster? If so, how massive was cluster, and is there any hope at
+identifying the stars that formed near the Sun (e.g., Adams 2010)?
+Finally, the dispersal of open clusters also provides perhaps
+the best test case for the concept of ``chemical tagging'', aka.\
+``galactic archaeology'' (CITE).  If you can't chemically tag stars
+that are kinematically associated, why would you be able to do so once
+they've dispersed throughout the galactic disk?
+
+Identifying the remnant
 halos\footnote{ The low-density outer regions of open clusters have
 been referred to in general terms as ``halos'' and ``coronae''; we
-adopt the former (CITE, CITE). } of open clusters is important for a
+adopt the former (CITE, CITE). } of open clusters is also important for a
 separate project: that of discovering young transiting planets.  Young
 transiting planets are hard to find because young stars are rare
-(e.g., CITE Piskunov 2018), and often reside in crowded regions of the
-sky (e.g., CITE ANY OPEN CLUSTER CATALOG).  If the dispersed halos of
+(e.g., CITE Piskunov 2018), and reside in crowded regions of the
+sky (e.g., Kharchenko 2013).  If the dispersed halos of
 nearby star clusters could be reliably identified in a complete
-manner, this could expand the census of nearby star clusters by up to
+manner, this could expand the census of nearby young stars by up to
 a factor of $10$, based on the expected fraction of stars thought to
-be lost during gas dispersal.  Path to identifying these young stars
-include photometric criteria (CITE Zari+18), spectroscopic criteria
-(CITE Berger+18), and combinations thereof (e.g., CITE Zerjal 2019,
-2020, Zhou's work).  A benefit of searching for planets in
-dispersed stellar environments is that issues with stellar crowding
-are also alleviated.
+be lost during gas dispersal.
 
 The identification of low-density (GIVE A NUMBER DENSITY) stellar
-associations, already possible before Gaia (e.g.,
-Zuckerman+ CITE), has reached a breakthrough pace with the Gaia data.
-Some low-density populations (``moving groups'') have no discernable
-cores (e.g., USco CITE, Psc Eri CITE, mu Tau CITE).
+associations, though possible before Gaia (e.g.,
+Zuckerman+ CITE, Bergond+01), has reached a breakthrough pace in the present era.
+Some low-density moving groups have no discernable
+cores (e.g., Psc Eri CITE, mu Tau CITE), or are hierarchically
+sub-structured (e.g., USco CITE, Taurus[?]).
 However other
 low-density
-halos seem to be centered on previously reported open clusters (e.g.,
+halos are associated with previously known open clusters (e.g.,
 Kounkel+19, Kounkel+20, Meingast+21).  In some
 cases, these halos correspond to tidal tails, e.g., in the Hyades
-(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
+(CITE Meingast+Alves 20, Roser+20), and in the Ursa Major moving
 group (CITE Gagne+20). 
-However, different clustering methods on the
-Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
-Employing say Gaussian Mixture Modelling, or any analogous method that
-requires ``clusters'' to be ellipses in phase-space (position or
-velocity) unsurprisingly yields open clusters that are roughly
-elliptical.  Some unsupervised clustering methods lead to comparable
-``roughly elliptical'' results \citep[][hereafter
+However, applying different clustering methods on the
+Gaia data yields different results (CITE: Hunt \& Reffert 2020).
+Using say Gaussian Mixture Modelling, or any analogous method that
+requires ``clusters'' to be ellipses in phase-space unsurprisingly
+yields open clusters that are roughly elliptical.  Some unsupervised
+clustering methods lead to similar results \citep[][hereafter
 \citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
-unsupervised methods such as HDBScan (e.g., \citep[][hereafter
-\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
-been found to yield additional structures, particularly in lower
-density regions such as the Psc-Eri stream (CITE: Meingast+18,
-Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
-also more generally around many open clusters
-\citep{kounkel_untangling_2019}.  Unsupervised approaches that
-incorporate physical constraints like imposing a maximum velocity
-dispersion on putative members yield similar results, potentially with
-higher purity (Meingast+21).
+unsupervised methods
+have been found to yield dispersed and asymmetric structures
+with number densities down to 100 times less than the field, spread
+over many hundreds of parsecs
+(e.g., \citep[][hereafter
+\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}; Meingast+21).
 
@@
 modest question: in just a single rich southern open cluster, is the
-cluster halo coeval with the core?  The cluster we chose for this
-analysis was NGC\,2516, since it is young ($\sim$120\,Myr) and close
+cluster halo coeval with the core?  We chose for this
+NGC\,2516 for this analysis since it is young ($\sim$120\,Myr; B3V at
+the turnoff) and close
 ($d\approx400$\,pc) enough to facilitate rotation measurements using
-TESS, and some degree of spectroscopic analysis.  We want to know: is
+TESS, and some degree of spectroscopic analysis.  
+Its halo, spanning $\approx$300\,pc, is also quite large.
+
+We want to know: is
 the halo real? To what extent can we use Gaia alone to reliably identify
@@
 % these?
-Both internal and external processes have been suggested to explain
+Processes both internal and external to the star have been suggested to explain
 the lithium-rotation trend (CITE see the recent review by Bouvier+20).
@@
 \subsection{How did the halo form?}
-``The stars in these halos could have escaped from the cores of the
-cluster, or they could be remnants of a larger and lower-density star
-formation complex that has since dispersed.''
+Another way to frame this question: should we call these stars a tidal
+tail?
+% TODO: read Gagne+20 on this...
+The classical explanation of such a structure (e.g., Krumholz+19) is
+that stars escape out of L1 and L2, and then form leading and lagging
+arms due to differential rotation in the galaxy.
+The leading and lagging arms of NGC\,2516's halo qualitatively
+match this picture, in that they are oriented roughly toward and away
+the direction of galactic rotation (CITE, Meingast21, link to their
+website).
+
+A second explanation though invokes
+the idea that the cluster formed in a larger and lower-density star
+formation complex, and the stars we see in the halo did not in fact
+form in say the same ``clump'' as those in the cluster core.
+
+% ``The stars in these halos could have escaped from the cores of the
+% cluster, or they could be remnants of a larger and lower-density star
+% formation complex that has since dispersed.''
 Another possibility is {\it triggered star formation}.
@@
 \subsection{Fast and blue rotators: are we going faster than other clusters?}
+
+\subsection{Ways of doing this at different ages}
+Some final discussion is warranted on the applicability of our
+approach more generally.  This work starts with a kinematically
+identified population, and then uses photometry (Gaia HR diagrams;
+TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
+confirm youth in the stars.  Other approaches are also possible.
+
+One can start with photometric criteria (e.g., Gaia HR diagram of
+everything within some distance; Zari+18), or with spectroscopic
+criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
+Zerjal 2019, 2020, Zhou's work).  
+It might even be possible to start using chemical abundance signatures
+(CITE, GALAHDR3?).
+
+These paths might in certain cases succeed in identifying more
+complete samples of dispersed members of young populations, since they
+do not require kinematic proximity.
 

### 2021-02-16 planet_paper/ms.tex
update paper

@@
   Cluster Difference Imaging Photometric Survey. IV.
-  TOI 1937Ab. Young Hot Jupiter, or Just Tidally Spun Up?
+  TOI 1937Ab. Youngest Hot Jupiter, or Just Tidally Spun Up?
 }
@@
 \begin{abstract}
-  foo,bar.
+  We report the discovery and confirmation of a hot Jupiter,
+  TOI\,1937Ab, and present the evidence for and against its youth
+  ($\approx$120\,Myr).
+  We found the planet in images taken by the NASA TESS mission
+  using the CDIPS pipeline.
+  We measured its mass (1.X Mjup) and orbital obliquity
+  (0$\pm$XX$^\circ$) using PFS at Magellan.
+  Gaia kinematics suggest that the star is in the halo of NGC\,2516.
+  The possible youth is corroborated by the {\bf G2V} star's 6.X day
+  rotation period, the planet's 22.X hour orbital period, and the
+  star's metallicity ([Fe/H] X.X$\pm$Y.Y dex) being consistent with
+  that of the cluster (X.X$\pm$Z.Z dex).
+  However, the star's spectrum does not show lithium, which argues
+  against it being truly young.
+  We outline tests that could support or refute the youth of the
+  system, and outline the reasons why few, if any, sub-100 Myr hot
+  Jupiters have been securely detected.
 \end{abstract}
@@
 - K-band image. MagAO? Multicolor imager.
+  E.g., Keck-NIRC2 (if it were in the North).
+  ???What is the prediction for the G-K color???
 - Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.

### 2021-02-17 cluster_paper/ms.tex
update figures

@@
 
-{\bf Emphasize that figure 1 does not clearly tell us whether or not the halo is really distinct from the field, by eye.}
-
 \begin{figure*}[t]
@@
   \caption{ {\bf Reported components of NGC\,2516 in position and
-    velocity space.}
-    The ``core'', identified by \citet{cantatgaudin_gaia_2018} using
-    Gaia DR2, is visually coincident
-    with where you would think the cluster is if you looked at it through a pair
-    of binoculars.
-    The ``halo'' was identified by \citet{kounkel_untangling_2019}
-    using a less restrictive membership assignment algorithm (discussed
-    in the appendices).
-    The ``field'' is a set of randomly drawn and non-overlapping stars
-    within a
-    $(\alpha, \delta, \pi)$ cone centered on the cluster.
+  velocity space.} The core, identified by
+  \citet{cantatgaudin_gaia_2018} using Gaia DR2, is coincident with
+  the visual overdensity of stars canonically accepted as ``the
+  cluster''.  The halo was identified by
+  \citet{kounkel_untangling_2019} using a less restrictive membership
+  assignment algorithm (see Appendix~\ref{app:clustering}).  The field
+  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
+  centered on the cluster.  It is difficult to visually distinguish
+  the field and the halo samples due to the low volume density of the
+  halo stars.
   \label{fig:gaia6d}
@@
 		\leavevmode
-			\includegraphics[width=0.75\textwidth]{f2a.pdf}
+			\includegraphics[width=0.85\textwidth]{f2a.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
-    The core (black) shows a clean sequence consistent with
-    stars with a fixed age and metallicity, and varying mass.  The
-    halo (blue) is similar, though somewhat noisier.  The faintest M
-    dwarfs in the core and halo are brighter than in the field star
-    comparison sample (gray), consistent with these stars having not
-    yet reached the ZAMS.
+  \caption{ {\bf HR diagram of NGC\,2516.}
+    No reddening correction has been applied; the photometric data are
+    from Gaia EDR3.  The core (black) shows a main sequence and
+    turnoff consistent with stars of a fixed age and metallicity.  The
+    halo (blue) is similar, though the binary sequence is less
+    well-defined.  The faintest M dwarfs in the core and halo are
+    brighter than in the field star comparison sample (gray),
+    consistent with these stars having not yet reached the ZAMS.
     \label{fig:hr}
@@
 		\leavevmode
-		\subfloat{
     \includegraphics[width=0.75\textwidth]{f3a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f3b.pdf}
-		}
+		% \subfloat{
+		% 	\includegraphics[width=0.75\textwidth]{f3a.pdf}
+		% }
+
+		% \subfloat{
+		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
+		% }
 	\end{center}
@@
 	\caption{ {\bf The core and halo of \cn\ in the space of rotation
-    period and Gaia color.}
-    The {\it top} plot shows periods against a linear scale, while the
-    {\it bottom}  shows them against a logarithmic scale.
+    period and dereddened Gaia color.}
+    % The {\it top} plot shows periods against a linear scale, while the
+    % {\it bottom}  shows them against a logarithmic scale.
+    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
+    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
+    are shown for reference.
+    Stars in the core and the halo of NGC\,2516 are
+    gyrochronologically coeval with the Pleiades.
 		\label{fig:rot}
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Field stars show a different rotation period
-  distribution than the kinematically selected members.}
+	\caption{ {\bf Rotation in NGC\,2516 compared to the field.}
+  Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
+  considered. (Fainter stars in the field sample were not studied as part
+  of the broader CDIPS project).
+  The field stars show a different rotation period distribution than
+  the kinematically selected NGC\,2516 members.
   \label{fig:compstar}
@@
     velocity space, cross-matched against the rotators.}
+    The plotted stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the
+    crowding restrictions described in Section~\ref{subsec:tess}: they
+    should have been sufficiently bright and non-crowded to enable
+    rotation period detection.
 		\label{fig:gaia6d_x_rotn}
@@
 	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
-    velocity space, cross-matched against the rotators, but now in
-    physical units.}
-    Top: relevant projections. Bottom: binned versus 3d position and
+    velocity space, cross-matched against the rotators, in physical
+    units.}
+    {\it Top:} projections in position and 2d-tangential velocity
+    space. {\it Bottom:} radially binned versus 3d position and
     2d velocity difference.

### 2021-02-17 drivers/plot_full_kinematics.py

@@ -6,7 +6,7 @@ PLOTDIR = os.path.join(RESULTSDIR, 'cluster_membership')
 
 for show1937 in [0,1]:
     for galacticframe in [0,1]:
-        ep.plot_full_kinematics(PLOTDIR, basedata='fullfaint', show1937=show1937, galacticframe=galacticframe) # deeper
         ep.plot_full_kinematics(PLOTDIR, basedata='bright', show1937=show1937, galacticframe=galacticframe) # standard
+        ep.plot_full_kinematics(PLOTDIR, basedata='fullfaint', show1937=show1937, galacticframe=galacticframe) # deeper
         ep.plot_full_kinematics(PLOTDIR, basedata='fullfaint_edr3', show1937=show1937, galacticframe=galacticframe) # deeper + EDR3
         # ep.plot_full_kinematics(PLOTDIR, basedata='extinctioncorrected', show1937=show1937) # NOTE not implemented

### 2021-02-17 cluster_paper/ms.tex
update paper

@@
   diffuse stellar populations surrounding nearby open clusters. 
-  The stars in these halos could have escaped from the cores of the
+  The stars in these diffuse halos could have escaped from the cores of the
   cluster, or they could be dispersed remnants of the initial star formation complex.
@@
   %TODO: says who?
-  The classical tidal radius of the cluster is $\sim$10\,pc.
+  The classical tidal radius of the cluster is $\approx$10\,pc.
   Combining photometry from Gaia, rotation periods from TESS, and
@@
   The rotation period and lithium abundances of the cluster stars are
-  also anti-correlated at fixed stellar mass, as is the case with
-  comparable stellar populations.
+  also anti-correlated at fixed stellar mass, as has been observed in
+  similar stellar populations.
   This work expands the set of confirmed NGC\,2516 members by a factor
@@
 From 10 to 1000\,Myr, the portion of the cluster that survives gas
-dispersal suffers an onslaught that typically unbinds the cluster.
+dispersal suffers an onslaught that typically unbinds the remaining stars.
 Internal to the cluster, mass loss from supernovae and AGB winds, and
@@
 into the galactic field is of interest for a few reasons.  One 
-reason is understanding the predominant conditions under which stars and star
-clusters form (e.g., Wright Mamajek 2018 ScoCen), as well as
-the processes that drive cluster dissolution (including gas dispersal,
-internal and external heating mechanisms, and radial migration across
-the galactic disk).
+is to understand the predominant conditions under which stars and star
+clusters form (e.g., Wright Mamajek 2018 ScoCen), in addition to
+the processes that drive cluster dissolution.
 Another is to understand the Sun's birth conditions---
 did it form in an open
-cluster? If so, how massive was cluster, and is there any hope at
+cluster, and if so, what type? Is there any hope at
 identifying the stars that formed near the Sun (e.g., Adams 2010)?
-Finally, the dispersal of open clusters also provides perhaps
-the best test case for the concept of ``chemical tagging'', aka.\
-``galactic archaeology'' (CITE).  If you can't chemically tag stars
-that are kinematically associated, why would you be able to do so once
-they've dispersed throughout the galactic disk?
+Finally, the dispersal of open clusters also provides 
+a strong test case for ``chemical tagging'', aka.\
+``galactic archaeology'' for stars in the thin and thick disk of
+the Galaxy (CITE).  If one cannot chemically tag stars
+that are known to be kinematically associated, why would 
+it be possible to do so once
+they are radially mixed throughout the galactic disk?
 
@@
 Some low-density moving groups have no discernable
-cores (e.g., Psc Eri CITE, mu Tau CITE), or are hierarchically
+cores (e.g., Psc Eri CITE, mu Tau CITE); others are hierarchically
 sub-structured (e.g., USco CITE, Taurus[?]).
-However other
+The focus of this work is
 low-density
-halos are associated with previously known open clusters (e.g.,
+halos that are associated with previously known open clusters (e.g.,
 Kounkel+19, Kounkel+20, Meingast+21).  In some
@@
 unsupervised methods
-have been found to yield dispersed and asymmetric structures
+have yielded dispersed and asymmetric structures
 with number densities down to 100 times less than the field, spread
@@
 modest question: in just a single rich southern open cluster, is the
-cluster halo coeval with the core?  We chose for this
+cluster halo coeval with the core?  We chose
 NGC\,2516 for this analysis since it is young ($\sim$120\,Myr; B3V at
@@
 Its halo, spanning $\approx$300\,pc, is also quite large.
-
 We want to know: is

### 2021-02-19 cluster_paper/ms.tex
update paper

@@
 By exploring the different possible selection criteria, we found that
-the ones we chose provided a
+the ones we adopted provided a
 reasonable balance between excluding stars on the slow sequence, {\it
@@
 
+To put the clusters on the same color scale, we needed to adopt
+reddening corrections.
+We assumed {X} for the Pleiades, and {Y} for Praesepe.
+For NGC\,2516, we queried the 2MASS DUST service to find the
+extinction parameters for the positions of each NGC\,2516 member
+(CITE: astrobase.services.dust; 2MASS DUST).  The mean and standard
+deviation of the E(B-V) values for the Schlegel+98 and SF11 maps were
+consistent, so we considered those of Schlegel+1998:
+$E(B-V)=0.206\pm0.039$.  Bonifacio+2000 noted though that Schlegel+98
+overestimate the reddening values when the color excess exceeds about
+0.10 mag. We applied their correction (Eq 1 from Bonifacio+2000),
+which led to an adopted $E(B-V)=0.102\pm0.019$.  Finally, to convert
+to Gaia colors we adopted the calibration of Stassun et al 2019, and
+assumed $E(Bp-Rp)=1.31 E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded
+$E(Bp-Rp)=0.134\pm0.025$, which is used in the plots.
+
 The resulting gyrochronology age we find for the core is XXX.
 For the halo, the claimed age from gyrochronology is YYY.
-Applying the same procedure to the field star comparison sample,
-we get an age of ZZZ.
-
-\subsubsection{Comparison Star Sample}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-    \includegraphics[width=0.49\textwidth]{f9.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Rotation in NGC\,2516 compared to the field.}
-  Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
-  considered. (Fainter stars in the field sample were not studied as part
-  of the broader CDIPS project).
-  The field stars show a different rotation period distribution than
-  the kinematically selected NGC\,2516 members.
-  \label{fig:compstar}
-	}
-\end{figure}
-
-To provide a basis for comparison, we also opted to search the
-``calibration'' light curves ($Rp<13$) that were created as a part of
-the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
-corresponded to a sample of \ncalibration stars.  Cross-matching these
-against the \nnbhd randomly drawn stars in the neighborhood of \cn yielded
-\nnbhdcalibstar unique stars, with a cumulative total of \nnbhdcaliblc TESS
-sectors observed.
-The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
-corresponds to an extinction-corrected color cutoff of $(Bp-Rp)_0
-\approx 0.80$, or spectral types of $\approx$G1V.
-This reaches sufficiently far down the main sequence to enable a
-comparison to the cluster star sample.
-
-We performed the same light curve stitching and period-search
-procedure on the field comparison stars.
-Imposing the same requirements for crowding resulted in
-\ncompstardenominator stars for which rotation periods could have been
-detected.
-Imposing the same Lomb Scargle power cutoff, and period upper limit,
-yielded \ncompstarnumerator period detections (\ncompfrac).
-Within the same brightness cutoff, 
-\nautovscompstarnumerator of \nautovscompstardenominator cluster stars
-yielded period detections (\nautofrac).
-Though the detection fractions are frankly not very different
-(likely because of the brightness cutoff), the period {\it
-vs} color distributions are quite different
-(Figure~\ref{fig:compstar}).
+Applying the same procedure to the field star comparison sample (see
+Appendix~\ref{app:compstar}),
+we get an age of ZZZ+/-AAA.
 
@@
 
+\section{Comparison Star Rotation Periods}
+\label{app:compstar}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+    \includegraphics[width=0.49\textwidth]{f9.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Rotation in NGC\,2516 compared to the field.}
+  Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
+  considered. (Fainter stars in the field sample were not studied as part
+  of the broader CDIPS project).
+  The field stars show a different rotation period distribution than
+  the kinematically selected NGC\,2516 members.
+  \label{fig:compstar}
+	}
+\end{figure}
+
+To provide a basis for comparison, we also opted to search the
+``calibration'' light curves ($Rp<13$) that were created as a part of
+the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
+corresponded to a sample of \ncalibration stars.  Cross-matching these
+against the \nnbhd randomly drawn stars in the neighborhood of \cn yielded
+\nnbhdcalibstar unique stars, with a cumulative total of \nnbhdcaliblc TESS
+sectors observed.
+The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
+corresponds to an extinction-corrected color cutoff of $(Bp-Rp)_0
+\approx 0.80$, or spectral types of $\approx$G1V.
+This reaches sufficiently far down the main sequence to enable a
+comparison to the cluster star sample.
+
+We performed the same light curve stitching and period-search
+procedure on the field comparison stars.
+Imposing the same requirements for crowding resulted in
+\ncompstardenominator stars for which rotation periods could have been
+detected.
+Imposing the same Lomb Scargle power cutoff, and period upper limit,
+yielded \ncompstarnumerator period detections (\ncompfrac).
+Within the same brightness cutoff, 
+\nautovscompstarnumerator of \nautovscompstardenominator cluster stars
+yielded period detections (\nautofrac).
+Though the detection fractions are frankly not very different
+(likely because of the brightness cutoff), the period {\it
+vs} color distributions are quite different
+(Figure~\ref{fig:compstar}).
+
 \section{Differential Reddening}

### 2021-02-22 cluster_paper/ms.tex
update paper
the introduction is no longer garbage

@@
 \defcitealias{kounkel_untangling_2019}{KC19}
+\defcitealias{meingast_2021}{M21}
 
@@
   diffuse stellar populations surrounding nearby open clusters. 
-  The stars in these diffuse halos could have escaped from the cores of the
-  cluster, or they could be dispersed remnants of the initial star formation complex.
+  The stars in these diffuse halos could be the tidal escapees of the
+  cluster cores, or they could be remnants of an initially
+  dispersed star formation complex.
   They could also be field stars---the false positives of overly
@@
 
-Star clusters form within hierarchically structured molecular clouds
-(CITE).  Over the first $\approx$10\,Myr, ``feedback'' effects
+Star clusters form in hierarchically structured molecular clouds
+\citep{shu_star_1987}.  Over the first 10\,Myr, ``feedback'' effects
 including protostellar outflows, photoionization, radiation pressure,
-and supernova shocks disperse the gas out of the cloud (CITE
-Krumholz+19).  Since only a small fraction of the cloud mass is
-converted into stars ($\approx 1\%$; see Krumholz+19 Figure 10), the
-gas dispersal drives the majority ($\approx$90\%) of stars in the
-cluster to become gravitationally unbound (e.g., Lada \& Lada 03).
-
-From 10 to 1000\,Myr, the portion of the cluster that survives gas
-dispersal suffers an onslaught that typically unbinds the remaining stars.
-Internal to the cluster, mass loss from supernovae and AGB winds, and
-close stellar encounters (``relaxation'') are important over 10s and
-100s of megayears, respectively (CITE, CITE).  External to the
-cluster, collisions with GMCs (Spitzer 1958), and perturbations from
-the galactic tide in both the radial (CITE) and vertical (CITE)
-dimensions further drive stars toward the ``high velocity tails'' of
-the stellar energy distribution.
-
-% In the traditional picture of star formation molecular clouds go
-% gravitationally unstable, and collapse into little knots.  The knots
-% produce groups of stars close to each other.  The resulting ``open star
-% clusters'' are, as with any first-order guess, spherical.
-
-% How does the cluster evolve?  The answer depends on the total stellar
-% mass.  The smallest 90\% of star clusters disperse during the ``embedded
-% phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
-% by {\it process A} (CITE).  The more massive clusters make it maybe 100
-% Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
-% collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
-% Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
-% process B}.  Of course, the expected evaporation time should depend on
-% factors including the mass of the cluster itself, and how many high-mass
-% (O and B) stars form, since their winds [and maybe supernovae] clear out
-% most of the cloud, and the initial density of the gas cloud to begin
-% with.
-% 
-% Assuming that the cluster achieves virial equilibrium, stellar fly-bys
-% then conspire to segregate the stellar mass distribution within the
-% cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
-% assuming equipartition of the specific ``thermal energy'' per star, then
-% kT = 0.5$mv^2$ at lower masses requires higher velocities).
-
-Identifying the stars that have dispersed from their natal clusters
-into the galactic field is of interest for a few reasons.  One 
-is to understand the predominant conditions under which stars and star
-clusters form (e.g., Wright Mamajek 2018 ScoCen), in addition to
-the processes that drive cluster dissolution.
-Another is to understand the Sun's birth conditions---
-did it form in an open
-cluster, and if so, what type? Is there any hope at
-identifying the stars that formed near the Sun (e.g., Adams 2010)?
-Finally, the dispersal of open clusters also provides 
-a strong test case for ``chemical tagging'', aka.\
-``galactic archaeology'' for stars in the thin and thick disk of
-the Galaxy (CITE).  If one cannot chemically tag stars
-that are known to be kinematically associated, why would 
-it be possible to do so once
-they are radially mixed throughout the galactic disk?
-
-Identifying the remnant
-halos\footnote{ The low-density outer regions of open clusters have
-been referred to in general terms as ``halos'' and ``coronae''; we
-adopt the former (CITE, CITE). } of open clusters is also important for a
-separate project: that of discovering young transiting planets.  Young
-transiting planets are hard to find because young stars are rare
-(e.g., CITE Piskunov 2018), and reside in crowded regions of the
-sky (e.g., Kharchenko 2013).  If the dispersed halos of
-nearby star clusters could be reliably identified in a complete
-manner, this could expand the census of nearby young stars by up to
-a factor of $10$, based on the expected fraction of stars thought to
-be lost during gas dispersal.
-
-The identification of low-density (GIVE A NUMBER DENSITY) stellar
-associations, though possible before Gaia (e.g.,
-Zuckerman+ CITE, Bergond+01), has reached a breakthrough pace in the present era.
-Some low-density moving groups have no discernable
-cores (e.g., Psc Eri CITE, mu Tau CITE); others are hierarchically
-sub-structured (e.g., USco CITE, Taurus[?]).
-The focus of this work is
-low-density
-halos that are associated with previously known open clusters (e.g.,
-Kounkel+19, Kounkel+20, Meingast+21).  In some
-cases, these halos correspond to tidal tails, e.g., in the Hyades
-(CITE Meingast+Alves 20, Roser+20), and in the Ursa Major moving
-group (CITE Gagne+20). 
-However, applying different clustering methods on the
-Gaia data yields different results (CITE: Hunt \& Reffert 2020).
-Using say Gaussian Mixture Modelling, or any analogous method that
-requires ``clusters'' to be ellipses in phase-space unsurprisingly
-yields open clusters that are roughly elliptical.  Some unsupervised
-clustering methods lead to similar results \citep[][hereafter
-\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
-unsupervised methods
-have yielded dispersed and asymmetric structures
-with number densities down to 100 times less than the field, spread
-over many hundreds of parsecs
-(e.g., \citep[][hereafter
-\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}; Meingast+21).
-
-% TODO: make the bonafide number below quantitative
-We've recently been making TESS light curves of age-dated stars across
-the sky, as part of a Cluster Difference Imaging Photometry Survey
-(CDIPS, CITE Bouma+19).  Our analysis of Cycle 1 (Sectors 1-13) yielded
-light curves of \numsouthernuniqlcs candidate cluster members in the
-Southern Ecliptic hemisphere, available on
-MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.  Based on
-rotation periods, $\approx$25\% appear to be bonafide cluster members.
-
-As part of a broader project of identifying a large and clean sample of
-young stars for a transit search, we focus in this paper on a 
-modest question: in just a single rich southern open cluster, is the
-cluster halo coeval with the core?  We chose
-NGC\,2516 for this analysis since it is young ($\sim$120\,Myr; B3V at
-the turnoff) and close
-($d\approx400$\,pc) enough to facilitate rotation measurements using
-TESS, and some degree of spectroscopic analysis.  
-Its halo, spanning $\approx$300\,pc, is also quite large.
-We want to know: is
-the halo real? To what extent can we use Gaia alone to reliably identify
-age-dated needles in the haystack of boring field stars?  And more
-generally, what are the implications for the evolution of open clusters
-if they do have halos?
-
-Section~\ref{sec:gaia} presents the astrometric data from Gaia, and
-clarifies our usage of the terms ``core'' and ``halo''.
-Section~\ref{sec:agedate} age-dates the halo of NGC\,2516, based on Gaia
-photometry (Section~\ref{subsec:hr}), TESS gyrochronology
+and supernova shocks disperse the gas out of the cloud
+\citep{krumholz_star_2019}.  Since only a small fraction of the cloud
+mass is converted into stars ($\sim 1\%$), gas dispersal enables the
+majority ($\sim$90\%) of stars in the cluster to escape the cluster's
+gravitational well \citep{lada_embedded_2003}.
+
+From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
+suffer an onslaught that almost always leads to dissolution.
+Within the cluster supernovae, AGB winds, and close stellar encounters
+lead to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
+the cluster, collisions with giant molecular clouds
+\citep{spitzer_disruption_1958}, and perturbations from the galactic
+tide in both the radial and vertical dimensions further promote
+stellar escape \citep[{\it
+e.g.},][]{fukushige_timescale_2000,bergond_gravitational_2001}.
+
+Finding the stars that have dispersed from their natal clusters
+into the galactic field is of interest for a few reasons.  One is to
+understand the predominant conditions under which stars and star
+clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another is
+to understand the Sun's birth environment
+\citep{adams_birth_2010}.  Did the Sun form in an open cluster,
+and if so, what type? Is there any hope at identifying the stars that
+formed near the Sun?  Open clusters and their dispersed
+remnants also provide a test case for ``chemical tagging'' \citep[{\it
+e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
+The converse holds too: one could use chemical tagging to test the
+viability of kinematic clustering methods.
+
+A separate project that benefits from the new discoveries of dispersed
+stellar populations is that of finding young transiting exoplanets.
+Young transiting planets are hard to find because young stars are
+rare, and mostly reside in crowded galactic plane \citep[see {\it e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
+dispersed halos of nearby star clusters could be reliably identified,
+this could expand the census of nearby young stars by up to a factor
+of $10$, based on the expected fraction of stars thought to be lost
+during gas dispersal.
+
+The detection of dispersed stellar associations, although
+possible before Gaia, has now reached a breakthrough pace \citep[{\it
+e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
+The resuls can be summarized as follows. On one end are coeval populations with no
+discernable cores ({\it e.g.}, Psc Eri and $\mu$ Tau,
+\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
+the other are hierarchically structured populations with many over and
+under-densities  \citep[{\it e.g.} the Sco-Cen and Vela OB2
+associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here,
+we focus on the middle: low-density halos associated with a single
+overdensity, typically an open cluster that was known before Gaia
+\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
+In some cases, these halos could correspond to tidal tails, as have been
+reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
+the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
+Berenices \citep{tang_comaber_2019}.
+
+One point of difficulty however is that
+different clustering methods yield differing levels of both
+sensitivity and precision \citep{hunt_clustering_2020}.  Using say a
+Gaussian Mixture Model tautologically yields open clusters that
+are elliptical \citep[{\it e.g.},][]{wallace_m4_2018}.  Some
+unsupervised methods yield dispersed and asymmetric structures with
+number densities down to 100 times less than the field around the same
+regions (see {\it e.g.} \citealt{kounkel_untangling_2019}, hereafter
+\citetalias{kounkel_untangling_2019}], and
+\citealt{meingast_2021}, hereafter \citetalias{meingast_2021}).
+Discrepancies between different clustering methods leads to
+the question of how contaminated the resulting samples are,
+and in some cases whether the reported structures are truly
+coeval stellar populations.
+
+We have recently been making TESS light curves of stars reported
+to be members of such populations
+(CDIPS, \citealt{bouma_cdipsI_2019}).  Our analysis of TESS Cycle 1
+(Sectors 1-13) yielded light curves of \numsouthernuniqlcs candidate
+cluster members in the Southern Ecliptic hemisphere, available on
+MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
+
+%; B3V at the turnoff
+As part of this broader project, we focus here on a modest question:
+in just a single rich southern open cluster, is the cluster halo truly
+coeval with the core?  We chose NGC\,2516 for this analysis since it
+is young ($\approx$120\,Myr) and sufficiently nearby
+($d\approx400$\,pc) to facilitate photometric rotation measurements
+using TESS, and some degree of spectroscopic analysis.  Its halo is
+decidedly non-spherical: \citetalias{kounkel_untangling_2019} reported
+its dimensions as $\approx20\times10\times350$\,pc.  We want
+to know: is the halo real? To what extent can we use Gaia alone to
+reliably identify age-dated needles in the haystack of field
+stars?  What are the implications, observationally and theoretically,
+if we can identify the dispersed halos of open clusters?
+
+A brief
+note on terminology.  Low-density stellar associations connected to a
+dense population (the ``core'') have been described as ``halos'',
+``strings'', ``coronae'', ``snakes'', and ``tidal tails'' \citep[{\it
+	e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
+The latter term implies a particular model for the
+formation of the association.  We opt to use ``halo'' because it is
+concise and model agnostic.  It suffers however from the suggestion of
+spherical symmetry.  The halo of NGC\,2516 is anything but.
+
+Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
+data that led to our interest in NGC\,2516.
+Section~\ref{sec:agedate} age-dates the cluster's halo and core using
+Gaia photometry (Section~\ref{subsec:hr}), TESS gyrochronology
 (Section~\ref{subsec:tess}), and lithium depletion
 (Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
-discuss the implications of this analysis for NGC\,2516 specifically and
-stellar spin-down and open cluster evolution generally.
+discuss the implications of our analysis for NGC\,2516 specifically
+and stellar spin-down and open cluster evolution generally.
 Section~\ref{sec:conclusion} gives our conclusions.
@@
 target selection approach included many stars in the spatial proximity
-of previously known members.  Since the possible existence of the \cn
+of previously known members.  Since the possible existence of the \cn\ 
 halo was not known at the time of targetting, very few ``halo'' stars
@@
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
-The fact that we see the same trend across the core and halo of \cn
+The fact that we see the same trend across the core and halo of \cn\ 
 {\it a)} supports the conclusion that the halo is coeval with the

### 2021-02-22 planet_paper/ms.tex
update figures
update paper
the introduction is no longer garbage

@@
 
+\paragraph{Isn't the rotation period a little slow?}
+Yes. But hot Jupiters affect the star's rotation, and it's hard to be
+sure how much. Consider TOI 1431b, an Am-stype with vsini=8km/s
+hosting a retrograde hot Jupiter. 
+Generally speaking, retrograde systems will spin down the star (e.g.,
+Anderson+2021), even if the planet is quickly ``realigned''.
+The latter realignment would be needed in our case to understand TOI
+1937Ab, since the observed orbit is prograde.
+
 \subsection{The present}
@@
   z-band)->Do PSF fitting.
+
 - K-band image. MagAO? Multicolor imager.
@@
   ???What is the prediction for the G-K color???
+  Doesn't seem all that promising. G-mag would predict something like
+  M2.5V or M3.5V (0.40Msun or 0.30Msun from mamajek).
+  At 100 Myr, a 0.20Msun star is 18Gcm vs 16Gcm (12\% inflated;
+  Burrows+01). This is not a whole lot. And it is worse for more massive
+  stars.
 - Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.
+  Going whole hog, maybe SPHERE could do it.
 

### 2021-02-24 planet_paper/ms.tex
planet paper: sketch some intro

@@
 
-Lorem ipsum.
+Processes that have been suggested to produce hot Jupiters occur on
+timescales ranging from megayears to gigayears (CITE, CITE, CITE).
+One way to distinguish between these processes is therefore to find
+young ($\lesssim$100\,Myr) hot Jupiters.  However the search for
+sub-100\,Myr hot Jupiters has yielded remarkably few, if any, secure
+detections.
+
+Transits have provided a pile of tantalizing candidates.  HIP\,67522b,
+with an age, orbital period, and size of XX\,Myr, X.X\,days, and
+TOOSMALL\,$R_\oplus$ respectively, comes close
+\citep{rizzuto_tess_2020}.  Its
+size however is smaller than expected for a Jovian-mass object given
+its age, by a factor of $\approx1.5\times$ (CITE Burrows, Fortney,
+Thorngren models, CITE Owen 2020 entropy).  A mass measurement is
+therefore needed to resolve its status as either a hot Jupiter, or an
+inflated Neptunian world.  No other planet comes as close to fitting
+the bill.  V1298 Tau b, part of a unique system with at least four
+transiting planets, has the correct size and age. However its orbital
+period (XX\,days) is too long to be a hot Jupiter (CITE David 2019,
+2020).  The dips in PTFO 8-8695, long-interpreted as a candidate hot
+Jupiter, are chromatic and show a changing orbital phase.  This is
+inconsistent with a planetary interpretation (van Eyken 2012, Yu+15,
+Onitsuka+17, Tanimoto+20, Bouma+20).
+
+Reports of very young hot Jupiters have come from the radial velocity
+(RV) technique, but these reports are difficult to verify.  TW Hya
+(age=X\,Myr), for instance, showed a radial velocity semi-amplitude of
+$\approx$200\,m\,s$^{-1}$ with a period of 3.6\,days, as measured in
+optical spectra \citep{setiawan_young_2008} (CITE age reference).
+Even though \citet{setiawan_young_2008} observed no significant
+correlation between activity indicators and the RV signal, subsequent
+infrared velocities measured by \citet{huelamo_tw_2008} showed small
+variations $\lesssim 35\,$m\,s$^{-1}$.  \citet{huelamo_tw_2008}
+ultimately concluded that activity-induced variations were the source
+of the optical signal.  More recently, very young hot Jupiters have
+been reported around CI Tau, V830 Tau, and TAP 26
+\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,biddle_k2_2018,flagg_co_2019}.
+The planetary nature of at least two of these signals (CI Tau and V830
+Tau) has been debated \citep{donati_magnetic_2020,damasso_gaps_2020}.
+More generally, the Bayesian model comparison problem of ``stellar
+activity alone'' versus ``stellar activity plus planet'' has been
+shown to require significant amounts of both data and statistical
+machinery \citep[{\it
+e.g.},][]{barragan_radial_2019,klein_simulated_2020} To date, none of
+the reported young hot Jupiter detections have performed this style of
+analysis.  Such an analysis, or else the acquisition of multi-color
+radial velocities, seems like a requirement given the challenges of
+detecting small signals in the presence of significant stellar
+variability.
+
+So where does that leave the search for young hot Jupiters?
+A quick query to the NASA Exoplanet Archive yields the current state
+of the search.
+(FIGURE: Rp vs Age, and Mp\,sini vs Age).
 

### 2021-02-24 cluster_paper/ms.tex
cluster paper: manuscript

@@
 From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
-suffer an onslaught that almost always leads to dissolution.
-Within the cluster supernovae, AGB winds, and close stellar encounters
-lead to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
+suffer an onslaught that almost always leads to dissolution.  Within
+the cluster supernovae, AGB winds, and close stellar encounters lead
+to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
 the cluster, collisions with giant molecular clouds
@@
 
-Finding the stars that have dispersed from their natal clusters
-into the galactic field is of interest for a few reasons.  One is to
+Finding the stars that have dispersed from their natal clusters into
+the galactic field is of interest for a few reasons.  One is to
 understand the predominant conditions under which stars and star
-clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another is
-to understand the Sun's birth environment
-\citep{adams_birth_2010}.  Did the Sun form in an open cluster,
-and if so, what type? Is there any hope at identifying the stars that
-formed near the Sun?  Open clusters and their dispersed
-remnants also provide a test case for ``chemical tagging'' \citep[{\it
+clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
+is to understand the Sun's birth environment \citep{adams_birth_2010}.
+Did the Sun form in an open cluster, and if so, what type? Is there
+any hope at identifying the stars that formed near the Sun?  Open
+clusters and their dispersed remnants also provide a test case for
+``chemical tagging'' \citep[{\it
 e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
@@
 Young transiting planets are hard to find because young stars are
-rare, and mostly reside in crowded galactic plane \citep[see {\it e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
+rare, and mostly reside in crowded galactic plane \citep[see {\it
+e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
 dispersed halos of nearby star clusters could be reliably identified,
@@
 
-The detection of dispersed stellar associations, although
-possible before Gaia, has now reached a breakthrough pace \citep[{\it
+The detection of dispersed stellar associations, although possible
+before Gaia, has now reached a breakthrough pace \citep[{\it
 e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
-The resuls can be summarized as follows. On one end are coeval populations with no
-discernable cores ({\it e.g.}, Psc Eri and $\mu$ Tau,
+The resuls can be summarized as follows. On one end are coeval
+populations with no discernable cores ({\it e.g.}, Psc Eri and $\mu$
+Tau,
 \citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
@@
 under-densities  \citep[{\it e.g.} the Sco-Cen and Vela OB2
-associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here,
-we focus on the middle: low-density halos associated with a single
+associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
+focus on the middle: low-density halos associated with a single
 overdensity, typically an open cluster that was known before Gaia
 \citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
-In some cases, these halos could correspond to tidal tails, as have been
-reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
-the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
-Berenices \citep{tang_comaber_2019}.
-
-One point of difficulty however is that
-different clustering methods yield differing levels of both
-sensitivity and precision \citep{hunt_clustering_2020}.  Using say a
-Gaussian Mixture Model tautologically yields open clusters that
-are elliptical \citep[{\it e.g.},][]{wallace_m4_2018}.  Some
-unsupervised methods yield dispersed and asymmetric structures with
-number densities down to 100 times less than the field around the same
-regions (see {\it e.g.} \citealt{kounkel_untangling_2019}, hereafter
-\citetalias{kounkel_untangling_2019}], and
-\citealt{meingast_2021}, hereafter \citetalias{meingast_2021}).
-Discrepancies between different clustering methods leads to
-the question of how contaminated the resulting samples are,
-and in some cases whether the reported structures are truly
-coeval stellar populations.
-
-We have recently been making TESS light curves of stars reported
-to be members of such populations
-(CDIPS, \citealt{bouma_cdipsI_2019}).  Our analysis of TESS Cycle 1
-(Sectors 1-13) yielded light curves of \numsouthernuniqlcs candidate
-cluster members in the Southern Ecliptic hemisphere, available on
+In some cases, these halos could correspond to tidal tails, as have
+been reported in the Hyades
+\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
+group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
+\citep{tang_comaber_2019}.
+
+One point of difficulty however is that different clustering methods
+yield differing levels of both sensitivity and precision
+\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
+tautologically yields open clusters that are elliptical \citep[{\it
+e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
+and asymmetric structures with number densities down to 100 times less
+than the field around the same regions (see {\it e.g.}
+\citealt{kounkel_untangling_2019}, hereafter
+\citetalias{kounkel_untangling_2019}], and \citealt{meingast_2021},
+hereafter \citetalias{meingast_2021}).  Discrepancies between
+different clustering methods leads to the question of how contaminated
+the resulting samples are, and in some cases whether the reported
+structures are truly coeval stellar populations.
+
+We have recently been making TESS light curves of stars reported to be
+members of such populations (CDIPS, \citealt{bouma_cdipsI_2019}).  Our
+analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
+\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
+hemisphere, available on
 MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
@@
 decidedly non-spherical: \citetalias{kounkel_untangling_2019} reported
-its dimensions as $\approx20\times10\times350$\,pc.  We want
-to know: is the halo real? To what extent can we use Gaia alone to
-reliably identify age-dated needles in the haystack of field
-stars?  What are the implications, observationally and theoretically,
-if we can identify the dispersed halos of open clusters?
-
-A brief
-note on terminology.  Low-density stellar associations connected to a
-dense population (the ``core'') have been described as ``halos'',
-``strings'', ``coronae'', ``snakes'', and ``tidal tails'' \citep[{\it
+its dimensions as $\approx20\times10\times350$\,pc.  We want to know:
+is the halo real? To what extent can we use Gaia alone to reliably
+identify age-dated needles in the haystack of field stars?  What are
+the implications, observationally and theoretically, if we can
+identify the dispersed halos of open clusters?
+
+A brief note on terminology.  Low-density stellar associations
+connected to a dense population (the ``core'') have been described as
+being in ``halos'', ``strings'', ``coronae'', ``snakes'', ``the
+outskirts'', and ``tidal tails'' \citep[{\it
 e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
-The latter term implies a particular model for the
-formation of the association.  We opt to use ``halo'' because it is
-concise and model agnostic.  It suffers however from the suggestion of
-spherical symmetry.  The halo of NGC\,2516 is anything but.
+The latter term implies a particular model for the formation of the
+association.  We opt to use ``halo'' because it is concise and model
+agnostic.  It suffers however from the suggestion of spherical
+symmetry.  The halo of NGC\,2516 is anything but.
 
 Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
-data that led to our interest in NGC\,2516.
-Section~\ref{sec:agedate} age-dates the cluster's halo and core using
-Gaia photometry (Section~\ref{subsec:hr}), TESS gyrochronology
+data that led to our interest in NGC\,2516.  Section~\ref{sec:agedate}
+age-dates the cluster's halo and core using Gaia photometry
+(Section~\ref{subsec:hr}), TESS gyrochronology
 (Section~\ref{subsec:tess}), and lithium depletion
@@
 
-We chose the stars to focus on based on previously reported
-memberships of NGC\,2516, available in the literature.  While a number
-of pre-Gaia catalogs were available, the purity and accuracy of the
-Gaia-derived results are expected to be the current state of the art.
+We selected candidate \cn\ members based on those available in the
+literature.  While a number of pre-Gaia catalogs were available, the
+purity and accuracy of the Gaia-derived results are expected to be the
+current state of the art.
 We therefore adopted what we viewed as the most interesting clustering
@@
 complete set of \cn\ members could very well include some of the newly
-reported candidate members from \citet{meingast_2021}.}. While we
-could have performed our own clustering analysis based on the Gaia
-data, such an effort would be replicating the work of these
+reported candidate members from \citet{meingast_2021}.}.  A useful
+visualization of these samples is available online\footnote{
+  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
+  made by \citet{meingast_2021}.  }.
+While we could have performed our own clustering analysis based on the
+Gaia data, such an effort would be replicating the work of these
 investigators. And why would our clustering methods be better than

### 2021-02-26 cluster_paper/ms.tex
update paper

@@
   The stars in these diffuse halos could be the tidal escapees of the
-  cluster cores, or they could be remnants of an initially
-  dispersed star formation complex.
+  cluster cores, or they could be remnants of an initially dispersed
+  star formation complex.
   They could also be field stars---the false positives of overly
@@
 
-%; B3V at the turnoff
 As part of this broader project, we focus here on a modest question:
@@
 \label{sec:gaia}
-%
-% What is the core? What is the halo?
-%
 
@@
 literature.  While a number of pre-Gaia catalogs were available, the
-purity and accuracy of the Gaia-derived results are expected to be the
-current state of the art.
+purity and accuracy of the Gaia-derived results are the current state
+of the art.
 We therefore adopted what we viewed as the most interesting clustering
 samples to compare: those of \citet{cantatgaudin_gaia_2018}
-(\citetalias{cantatgaudin_gaia_2018}) and
+(\citetalias{cantatgaudin_gaia_2018}),
 \citet{kounkel_untangling_2019}
-(\citetalias{kounkel_untangling_2019})\footnote{A third sample was
-also recently reported by \citet{meingast_2021}. Although the results
-of this study were made publicly available too late to be included in
-our analysis,  1577 of the 1860 \citet{meingast_2021} sources (85\%)
-were included in the \citet{kounkel_untangling_2019} sample.  A
-complete set of \cn\ members could very well include some of the newly
-reported candidate members from \citet{meingast_2021}.}.  A useful
-visualization of these samples is available online\footnote{
+(\citetalias{kounkel_untangling_2019}),
+and \citet{meingast_2021} (\citetalias{meingast_2021}).
+A useful visualization of these samples is available online\footnote{
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
@@
 While we could have performed our own clustering analysis based on the
-Gaia data, such an effort would be replicating the work of these
-investigators. And why would our clustering methods be better than
-theirs? We opt instead to use their studies as starting points.
+Gaia data, such an effort would be only replicating the work of these
+investigators.  We opt to instead use their studies as starting
+points.
+
+%FIXME FIXME FIXME FROM HERE
 
@@
 reported 3003 candidate members, and included stars up to
-$\approx$1\,mag fainter.  Each study used unsupervised clustering
-based on the second Gaia Data Release, detailed in
+$\approx$1\,mag fainter.  
+Finally, 1577 of the 1860 \citet{meingast_2021} sources (85\%)
+were included in the \citet{kounkel_untangling_2019} sample.  
+Each study applied unsupervised clustering techniques
+to the second Gaia Data Release, as detailed in
 Appendix~\ref{app:clustering}.
+
 %FIXME 
@@
   \caption{ {\bf HR diagram of NGC\,2516.}
-    No reddening correction has been applied; the photometric data are
-    from Gaia EDR3.  The core (black) shows a main sequence and
+    The photometric data are from Gaia EDR3; no reddening correction
+    has been applied.  The core (black) shows a main-sequence and
     turnoff consistent with stars of a fixed age and metallicity.  The
     halo (blue) is similar, though the binary sequence is less
-    well-defined.  The faintest M dwarfs in the core and halo are
-    brighter than in the field star comparison sample (gray),
-    consistent with these stars having not yet reached the ZAMS.
+    defined.  At fixed color, the faintest M dwarfs in the core and
+    halo are brighter than in the field star comparison sample (gray),
+    consistent with these stars having not yet reached the
+    main-sequence.
     \label{fig:hr}
@@
 
-The first check on whether this membership assignment is plausible was
-already performed by \citetalias{cantatgaudin_gaia_2018},
+The first check on the plausibility of the candidate cluster members
+being coeval was already performed by
+\citetalias{cantatgaudin_gaia_2018},
 \citetalias{kounkel_untangling_2019}, and more recently by
@@
 metallicity, and varying mass.  The halo members show a roughly
-consistent main sequence, but with somewhat greater scatter.  One
+consistent main-sequence, but with somewhat greater scatter.  One
 possible explanation for this scatter is that the halo is more
@@
 during its embedded phase, but is instead tied to physical processes
-acting on all low-mass pre-main sequence stars.
+acting on all low-mass pre-main-sequence stars.
 
@@
 NUMBER, through NASA grant NUMBER.
+L.G.B. was also supported by a XXXX Fellowship from Princeton
+University.
 %
@@
 %
-% TODO: the following two paragraphs, when quoted, are ENTIRELY FROM
+% TODO: the following two paragraphs, WHEN QUOTED, are ENTIRELY FROM
 % the relevant papers. they should probably be nixed in favor of
@@
 
+\citetalias{meingast_2021} did the same as
+\citetalias{kounkel_untangling_2019}, but required A) the tangential
+velocity dispersions to be smaller, and B) deconvolved the distances.
+{\bf FIXME FIXME FIXME}
+{\bf FIXME FIXME FIXME}
+{\bf FIXME FIXME FIXME}
+
 \section{Detrending details}
@@
 \approx 0.80$, or spectral types of $\approx$G1V.
-This reaches sufficiently far down the main sequence to enable a
+This reaches sufficiently far down the main-sequence to enable a
 comparison to the cluster star sample.

### 2021-02-26 cluster_paper/ms.tex
HR diagrams: making them awesome

@@
 	\vspace{-0.7cm}
-  \caption{ {\bf Reported components of NGC\,2516 in position and
-  velocity space.} The core, identified by
-  \citet{cantatgaudin_gaia_2018} using Gaia DR2, is coincident with
+  \caption{ {\bf NGC\,2516 has multiple reported components in position and
+  velocity space.} The core (black) was identified by
+  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
   the visual overdensity of stars canonically accepted as ``the
   cluster''.  The halo was identified by
-  \citet{kounkel_untangling_2019} using a less restrictive membership
-  assignment algorithm (see Appendix~\ref{app:clustering}).  The field
+  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
+  assignment algorithms (see Appendix~\ref{app:clustering}).  The field
   sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
-  centered on the cluster.  It is difficult to visually distinguish
-  the field and the halo samples due to the low volume density of the
-  halo stars.
+  centered on the cluster.  The low volume density of the
+  halo stars makes it difficult to visually distinguish
+  the field and the halo samples.
   \label{fig:gaia6d}
@@
 		\leavevmode
-			\includegraphics[width=0.85\textwidth]{f2a.pdf}
+			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f2a.pdf}
+			\includegraphics[width=0.49\textwidth]{f2b.pdf}
+		}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf HR diagram of NGC\,2516.}
-    The photometric data are from Gaia EDR3; no reddening correction
+  \caption{ {\bf HR diagrams of NGC\,2516.}
+    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
     has been applied.  The core (black) shows a main-sequence and
@@
     main-sequence.
+    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
+    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$ with a
+    reddening correction applied.
+    The main-sequence turn-off location and evolved stars suggests a
+    cluster age slightly older than the Pleiades.  The models diverge
+    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
+    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
     \label{fig:hr}
@@
 \citetalias{cantatgaudin_gaia_2018},
-\citetalias{kounkel_untangling_2019}, and more recently by
-\citet{meingast_2021}.  That check is to see whether the HR diagrams
-of the cluster members support the claim that they are coeval.
+\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.
+That check is to see whether the HR diagrams of the cluster members
+support the claim that they are coeval.
 
@@
 metallicity, and varying mass.  The halo members show a roughly
-consistent main-sequence, but with somewhat greater scatter.  One
-possible explanation for this scatter is that the halo is more
-contaminated by field stars.  Another possible explanation is
-differential reddening.  The halo is reported to span 20$^\circ$
-on-sky, and varies in position from about $b=-12^\circ$ to
+consistent main-sequence, perhaps with greater scatter.  One
+possible explanation for the additional scatter could be that the halo
+is more contaminated by field stars.  
+For instance, $\approx5$ evolved halo stars are visible in the
+field-star portion of the red giant branch (RGB).
+Given the expected age of the system, these stars must be field
+interlopers.
+
+A separate possible explanation for scatter in the halo's HR diagram
+could be differential reddening.  The halo is reported to span
+20$^\circ$ on-sky, and varies in position from about $b=-12^\circ$ to
 $b=-20^\circ$, with the stars closest to the galactic plane also being
 further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
-Empirical evidence for differential reddening is discussed further in
-Appendix~\ref{app:diffred}. A priori, both field star contamination
-and differential reddening could well play a role.
-
-%TODO 
-Comparing these HR diagrams to the PARSEC isochrone models, we find
-that X, Y, and Z.
-In particular, ``the faintest M dwarfs in the core and halo are
-brighter than in the field star comparison sample, consistent with
-these stars having not yet reached the ZAMS.''
-This is consistent with the 
-main-sequence turn-off being at $Bp-Rp\approx0.05$, which implies an
-age of XXX.
-The resulting photometric age we calculate for the core is XXX.
-For the halo, the claimed age from photometry is YYY.
-Applying the same procedure to the field star comparison sample,
-we get an age of ZZZ.
+Some empirical evidence for differential reddening is discussed
+further in Appendix~\ref{app:diffred}. Based on the available
+evidence, we expect that both field star contamination and
+differential reddening could play a role.
+A third possibility that we have not explored is whether the binary
+fraction could differ in spatially different regions of the cluster.
+
+In the right panel of Figure~\ref{fig:hr}, we compare the observed
+photometry with PARSEC isochrones
+\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
+We used the web
+interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
+\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
+at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and then applied the
+reddening correction described in Section~\ref{subsec:tess}, in the
+sense of adding the $E(Bp-Rp)$ correction to the synthetic Gaia EDR3
+photometry.  The data and models generally agree for masses above
+approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
+Bp-Rp$\approx2.2$, the agreement is worse.  The MIST isochrones show a
+comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
+both starspots and incomplete molecular line lists have been proposed
+\citep[{\it
+e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
+For purposes of age-dating the cluster, we prefer to simply focus on
+the main-sequence turn-off (MSTO), since it is where the models have maximal
+predictive power.
+
+Our photometric age-dating procedure is as follows.  We selected all
+stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
+excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
+outliers from the rest of the cluster.  Using an isochrone grid
+spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
+spacing, we then numerically evaluated the $\chi^2$ fit for each
+possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
+percentiles are used as summary statstics.
+
+% NOTE: probably asymmetric.
+The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
+For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
+yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
+the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
 
@@
 		\leavevmode
-			\includegraphics[width=0.75\textwidth]{f2b.pdf}
+			\includegraphics[width=0.75\textwidth]{f2c.pdf}
 	\end{center}

### 2021-02-26 cluster_paper/collect.sh

@@ -1,7 +1,8 @@
 cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
 
 cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f2a.pdf
-cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2b.pdf
+cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f2b.pdf
+cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2c.pdf
 
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear.pdf f3a.pdf
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf f3b.pdf

### 2021-02-26 drivers/plot_hr.py

@@ -3,35 +3,36 @@ import earhart.plotting as ep
 from earhart.paths import RESULTSDIR
 
 basedatas = ['fullfaint_edr3'] #, 'bright', 'fullfaint', 'denis_fullfaint_edr3']
-isochrone = 'parsec' # could be None, mist, parsec.
+isochrones = ['parsec', None, 'mist'] # could be None, mist, parsec.
 
 PLOTDIR = os.path.join(RESULTSDIR, 'hr')
 
 for basedata in basedatas:
+    for isochrone in isochrones:
 
-    if 'denis' not in basedata:
+        if 'denis' not in basedata:
 
-        for c in ['phot_bp_mean_mag', 'phot_g_mean_mag']:
+            for c in ['phot_bp_mean_mag', 'phot_g_mean_mag']:
+
+                # base HR diagrams
+                for show1937 in [0,1]:
+                    for highlight_companion in [0,1]:
+                        if highlight_companion and not show1937:
+                            continue
+                        ep.plot_hr(PLOTDIR, isochrone=isochrone, color0=c,
+                                   highlight_companion=highlight_companion,
+                                   basedata=basedata, show1937=show1937)
+
+                if isochrone is None:
+                    # check for differential extinction
+                    ep.plot_hr(PLOTDIR, color0=c,
+                               highlight_companion=0, basedata=basedata, colorhalobyglat=1)
+
+
+        else:
 
             # base HR diagrams
-            for show1937 in [0,1]:
-                for highlight_companion in [0,1]:
-                    if highlight_companion and not show1937:
-                        continue
-                    ep.plot_hr(PLOTDIR, isochrone=isochrone, color0=c,
-                               highlight_companion=highlight_companion,
-                               basedata=basedata, show1937=show1937)
-
-            if isochrone is None:
-                # check for differential extinction
-                ep.plot_hr(PLOTDIR, color0=c,
-                           highlight_companion=0, basedata=basedata, colorhalobyglat=1)
-
-
-    else:
-
-        # base HR diagrams
-        for highlight_companion in [1,0]:
-            ep.plot_hr(PLOTDIR, color0=None,
-                       highlight_companion=highlight_companion, basedata=basedata)
+            for highlight_companion in [1,0]:
+                ep.plot_hr(PLOTDIR, color0=None,
+                           highlight_companion=highlight_companion, basedata=basedata)
 

### 2021-02-26 earhart/plotting.py

@@ -529,7 +529,7 @@ def plot_edr3_blending_vs_apparentmag(outdir, basedata='fullfaint', num=None):
 
 def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
             basedata='fullfaint', highlight_companion=0, colorhalobyglat=0,
-            show1937=1):
+            show1937=1, rasterized=False):
     """
     basedata (str): any of ['bright', 'extinctioncorrected', 'fullfaint',
     'fullfaint_edr3'], where each defines a different set of neighborhood /
@@ -579,7 +579,7 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
 
     plt.close('all')
 
-    f, ax = plt.subplots(figsize=(4,3))
+    f, ax = plt.subplots(figsize=(1.5*2,1.5*3))
 
     get_yval = (
         lambda _df: np.array(
@@ -610,13 +610,18 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
 
 
     if not colorhalobyglat:
+        if isochrone is None:
+            l0,l1 = 'Field', 'Halo'
+        else:
+            l0,l1 = None, None
+        # mixed rasterizing along layers b/c we keep the loading times nice
         ax.scatter(
             get_xval(nbhd_df), get_yval(nbhd_df), c='gray', alpha=0.5, zorder=2,
-            s=5, rasterized=True, linewidths=0, label='Field', marker='.'
+            s=6, rasterized=False, linewidths=0, label=l0, marker='.'
         )
         ax.scatter(
             get_xval(kc19_df), get_yval(kc19_df), c='lightskyblue', alpha=1,
-            zorder=3, s=5, rasterized=True, linewidths=0.15, label='Halo',
+            zorder=3, s=6, rasterized=rasterized, linewidths=0.03, label=l1,
             marker='.', edgecolors='k'
         )
 
@@ -629,7 +634,7 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
 
         cax = ax.scatter(
             get_xval(kc19_df), get_yval(kc19_df), c=nparr(kc19_df[glatkey]),
-            alpha=1, zorder=3, s=2, rasterized=True, linewidths=0.1,
+            alpha=1, zorder=3, s=2, rasterized=rasterized, linewidths=0.1,
[…]

### 2021-03-01 cluster_paper/ms.tex
update reddening

@@
   In this study, we focus on a halo with a length of 300\,pc reported
-  around the $\approx$120\,Myr open cluster NGC\,2516.
+  around the $\approx$175\,Myr open cluster NGC\,2516.
   %TODO: says who?
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf NGC\,2516 has multiple reported components in position and
+  \caption{ {\bf The dense and diffuse components of NGC\,2516 across position and
   velocity space.} The core (black) was identified by
@@
   the visual overdensity of stars canonically accepted as ``the
-  cluster''.  The halo was identified by
+  cluster''.  The halo (blue) was identified by
   \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
-  assignment algorithms (see Appendix~\ref{app:clustering}).  The field
+  assignment algorithms described in Appendix~\ref{app:clustering}.  The field
   sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
@@
     {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
-    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$ with a
-    reddening correction applied.
+    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
+    The arrow represents the average reddening correction applied
+    to the models.
     The main-sequence turn-off location and evolved stars suggests a
@@
 \texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
-at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and then applied the
-reddening correction described in Section~\ref{subsec:tess}, in the
-sense of adding the $E(Bp-Rp)$ correction to the synthetic Gaia EDR3
-photometry.  The data and models generally agree for masses above
+at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied the
+reddening correction described in Section~\ref{subsec:tess}. This
+entailed by including using our corrected value of
+$E(B-V)=0.102\pm0.019$, assuming $R_V=3.1$, and applying the
+\citet{odonnell_1994} SED-dependent reddening law.
+The data and models generally agree for masses above
 approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
-Bp-Rp$\approx2.2$, the agreement is worse.  The MIST isochrones show a
+Bp-Rp$\approx2.2$, the data and models diverge.  The MIST isochrones show a
 comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
@@
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
+We adopt the 
+PARSEC isochrones since they diverge at slightly lower mass than the
+MIST models, due to the temperature-opacity calibration
+discussed by \citet{chen_improving_2014}.
 For purposes of age-dating the cluster, we prefer to simply focus on
-the main-sequence turn-off (MSTO), since it is where the models have maximal
-predictive power.
-
-Our photometric age-dating procedure is as follows.  We selected all
-stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
-excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
-outliers from the rest of the cluster.  Using an isochrone grid
-spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
-spacing, we then numerically evaluated the $\chi^2$ fit for each
-possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
-percentiles are used as summary statstics.
-
-% NOTE: probably asymmetric.
-The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
-For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
-yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
-the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
+the main-sequence turn-off (MSTO), since it is where the models have
+maximal predictive power.
+
+\citet{cummings_2018} presented techniques for mitigating some of the
+complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
+rotation, high binarity fractions, and the presence of blue
+stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
+cluster memberships, they found MSTO ages for NGC\,2516 ranging from
+165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
+PARSEC, MIST, or SYCLIST).  The statistical uncertainties ($\approx
+25$\,Myr) were comparable to the systematic uncertainties.
+
+Our primary goal is to determine whether the ages of the core and halo
+are consistent.  The left panel of Figure~\ref{fig:hr} shows
+empirically that they seem to be: stars past the turnoff in both the
+halo and core are on a consistent locus.  The right panel of
+Figure~\ref{fig:hr} suggests the precision with which the claim can be
+made.  The MSTO stars are consistent with the 178 and 316 Myr models,
+but are bluer than the 100 Myr model.  The RGB stars are consistent
+only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
+the assumption that the width of the MSTO can be attributed to binary
+stars, we are most interested in the blue edge. This seems most
+compatible with the 178 Myr model.
+
+% Our photometric age-dating procedure is as follows.  We selected all
+% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
+% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
+% outliers from the rest of the cluster.  Using an isochrone grid
+% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
+% spacing, we then numerically evaluated the $\chi^2$ fit for each
+% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
+% percentiles are used as summary statstics.
+% 
+% % NOTE: probably asymmetric.
+% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
+% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
+% yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
+% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
 
@@
     are shown for reference.
-    Stars in the core and the halo of NGC\,2516 are
-    gyrochronologically coeval with the Pleiades.
+    Stars in the core and the halo of NGC\,2516 appear
+    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
+    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
+    in which some NGC\,2516 members appear elevated with respect to
+    the Pleiades, suggesting a slightly older age.
 		\label{fig:rot}

### 2021-03-01 planet_paper/ms.tex
update reddening

@@
 
-Processes that have been suggested to produce hot Jupiters occur on
-timescales ranging from megayears to gigayears (CITE, CITE, CITE).
-One way to distinguish between these processes is therefore to find
-young ($\lesssim$100\,Myr) hot Jupiters.  However the search for
-sub-100\,Myr hot Jupiters has yielded remarkably few, if any, secure
-detections.
+Depending on the process that produces them, hot Jupiters arrive on
+their tiny orbits on timescales of anywhere between megayears and
+gigayears (CITE, CITE, CITE).
+One way to distinguish between these processes is to find
+young ($\lesssim$100\,Myr) hot Jupiters.
+However the search for such hot Jupiters has yielded remarkably few,
+if any, secure detections.
+What is the youngest hot Jupiter known?
 
@@
 the bill.  V1298 Tau b, part of a unique system with at least four
-transiting planets, has the correct size and age. However its orbital
+transiting planets, is appropriately large and young. However its orbital
 period (XX\,days) is too long to be a hot Jupiter (CITE David 2019,
@@
 $\approx$200\,m\,s$^{-1}$ with a period of 3.6\,days, as measured in
-optical spectra \citep{setiawan_young_2008} (CITE age reference).
+optical spectra \citep{setiawan_young_2008} (CITEP ALSO age reference).
 Even though \citet{setiawan_young_2008} observed no significant
@@
 
+%TODO: make figure
 So where does that leave the search for young hot Jupiters?
 A quick query to the NASA Exoplanet Archive yields the current state
-of the search.
+of the search, showcased in Figure~\ref{fig:rp_vs_age}.
 (FIGURE: Rp vs Age, and Mp\,sini vs Age).
+The youngest transiting hot Jupiter ($R_p>R_{\rm Jup}$, $P<10\,{\rm
+days}$) appears to be Qatar-4b. 
+Discovered by (CITE), the age was reported to be very low because of
+gyrochrones.
+Not great (CITE CITE).
 

### 2021-03-01 earhart/plotting.py

@@ -567,11 +567,15 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
 
         elif isochrone == 'parsec':
 			# v1
-            #isopath = os.path.join(DATADIR, 'isochrones',
-            #                       'output624293709713.dat')
+            # isopath = os.path.join(DATADIR, 'isochrones',
+            #                        'output624293709713.dat')
             # v2
             isopath = os.path.join(DATADIR, 'isochrones',
                                    'output911305443923.dat')
+            nored_iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
+            # v3
+            isopath = os.path.join(DATADIR, 'isochrones',
+                                   'output4767372113.dat')
             iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
 
 
@@ -619,11 +623,26 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
             get_xval(nbhd_df), get_yval(nbhd_df), c='gray', alpha=0.5, zorder=2,
             s=6, rasterized=False, linewidths=0, label=l0, marker='.'
         )
+        _s = 6
+
+
+        # wonky way to get output lines...
         ax.scatter(
             get_xval(kc19_df), get_yval(kc19_df), c='lightskyblue', alpha=1,
-            zorder=3, s=6, rasterized=rasterized, linewidths=0.03, label=l1,
+            zorder=4, s=_s, rasterized=rasterized, linewidths=0, label=None,
             marker='.', edgecolors='k'
         )
+        ax.scatter(
+            get_xval(kc19_df), get_yval(kc19_df), c='k', alpha=1,
+            zorder=3, s=_s+1, rasterized=rasterized, linewidths=0, label=None,
+            marker='.', edgecolors='k'
+        )
+        ax.scatter(
+            -99, -99, c='lightskyblue', alpha=1,
+            zorder=4, s=_s, rasterized=rasterized, linewidths=0.2, label=l1,
+            marker='.', edgecolors='k'
+        )
+
 
     else:
         glatkey = 'b'
@@ -646,7 +665,7 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
         _l = 'Core' if isochrone is None else None
         ax.scatter(
             get_xval(cg18_df), get_yval(cg18_df), c='k', alpha=0.9,
-            zorder=4, s=6, rasterized=rasterized, linewidths=0, label=_l, marker='.'
[…]

### 2021-03-02 cluster_paper/ms.tex
update text

@@
   \caption{ {\bf HR diagrams of NGC\,2516.}
-    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
+    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
     has been applied.  The core (black) shows a main-sequence and
@@
     halo are brighter than in the field star comparison sample (gray),
-    consistent with these stars having not yet reached the
+    consistent with these stars not having yet reached the
     main-sequence.
-    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
+    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
     at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
@@
     The main-sequence turn-off location and evolved stars suggests a
-    cluster age slightly older than the Pleiades.  The models diverge
+    cluster age slightly older than the Pleiades (dark blue line).  The models diverge
     from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
@@
 metallicity, and varying mass.  The halo members show a roughly
-consistent main-sequence, perhaps with greater scatter.  One
-possible explanation for the additional scatter could be that the halo
-is more contaminated by field stars.  
-For instance, $\approx5$ evolved halo stars are visible in the
-field-star portion of the red giant branch (RGB).
-Given the expected age of the system, these stars must be field
-interlopers.
+consistent main-sequence, with somewhat greater scatter.  One
+possible explanation for the additional scatter could be that the
+halo is more contaminated by field stars.  For instance, $\approx5$
+evolved halo stars are visible in the field-star portion of the red
+giant branch (RGB).  Given the age of the system, these stars must be
+field interlopers.
 
 A separate possible explanation for scatter in the halo's HR diagram
-could be differential reddening.  The halo is reported to span
-20$^\circ$ on-sky, and varies in position from about $b=-12^\circ$ to
-$b=-20^\circ$, with the stars closest to the galactic plane also being
-further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
-Some empirical evidence for differential reddening is discussed
-further in Appendix~\ref{app:diffred}. Based on the available
-evidence, we expect that both field star contamination and
-differential reddening could play a role.
-A third possibility that we have not explored is whether the binary
-fraction could differ in spatially different regions of the cluster.
+could be differential reddening across different sightlines.  The halo
+is reported to span 20$^\circ$ on-sky, and varies in position from
+about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
+galactic plane also being further from the Sun by up to 200\,pc
+(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
+reddening is discussed further in Appendix~\ref{app:diffred}. Based on
+the available evidence, we expect that both field star contamination
+and differential reddening could play a role in the larger scatter of
+the halo relative to the core.  A third possibility that we have not
+explored is whether the binary fraction could also differ between the
+different regions of the cluster.
 
@@
 interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
-\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
-at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied the
-reddening correction described in Section~\ref{subsec:tess}. This
-entailed by including using our corrected value of
-$E(B-V)=0.102\pm0.019$, assuming $R_V=3.1$, and applying the
-\citet{odonnell_1994} SED-dependent reddening law.
-The data and models generally agree for masses above
-approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
-Bp-Rp$\approx2.2$, the data and models diverge.  The MIST isochrones show a
-comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
-both starspots and incomplete molecular line lists have been proposed
-\citep[{\it
+\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
+corrections as in \citet{chen_2019_YBC}.} to interpolate these
+isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
+the reddening correction described in Section~\ref{subsec:tess}.  We
+assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applyied
+the \citet{odonnell_1994} SED-dependent extinction law star-by-star
+through the PARSEC interface.  For the metallicity, we considered a
+range of super and sub-solar metallicities to fit as much of the locus
+from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
+super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
+predictions too blue along the main sequence by $\approx$0.1\,mag.
+Our adopted metallicity agrees with that found by
+\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
+abundance recently determined by the Gaia-ESO team
+\citep{baratella_gaiaeso_2020}, which represented an up-revision from
+an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
+\citet{bailey_rv_2018} report a slightly sub-solar metallicity for
+the cluster, we prefer super-solar based on the goodness of fit from
+the photometry.
+
+Overall, the data and models agree for masses above
+approximately 0.45\,$M_\odot$.  However below this mass, the data and
+models diverge at colors redder than Bp-Rp$\approx2.2$.  The MIST
+isochrones showed a comparable disagreement \citep{choi_mesa_2016}.
+Explanations invoking both starspots and incomplete molecular line
+lists have been proposed \citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
-We adopt the 
-PARSEC isochrones since they diverge at slightly lower mass than the
-MIST models, due to the temperature-opacity calibration
-discussed by \citet{chen_improving_2014}.
-For purposes of age-dating the cluster, we prefer to simply focus on
-the main-sequence turn-off (MSTO), since it is where the models have
-maximal predictive power.
+We adopt the PARSEC isochrones since they diverge at slightly lower
+mass than the MIST models, due to the temperature-opacity calibration
+implemented by \citet{chen_improving_2014}.  For purposes of
+age-dating the cluster, we prefer to simply focus on the main-sequence
+turn-off (MSTO), since it is where the models have maximal predictive
+power.
 
@@
 are consistent.  The left panel of Figure~\ref{fig:hr} shows
-empirically that they seem to be: stars past the turnoff in both the
-halo and core are on a consistent locus.  The right panel of
-Figure~\ref{fig:hr} suggests the precision with which the claim can be
+empirically that they are: stars past the turnoff in both the halo and
+core are on a consistent locus.  The right panel of
+Figure~\ref{fig:hr} demonstrates the precision with which the claim can be
 made.  The MSTO stars are consistent with the 178 and 316 Myr models,
@@
 the assumption that the width of the MSTO can be attributed to binary
-stars, we are most interested in the blue edge. This seems most
-compatible with the 178 Myr model.
-
+stars, we are most interested in the blue edge. This appears most
+compatible with the 178 Myr model.  These results are consistent with
+the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
+and we refer to their work for a more precise model-dependent
+comparison.
+
+% FIXME: do you want to do the procedure below? Seems like overkill
+% given the Cummings study.
 % Our photometric age-dating procedure is as follows.  We selected all

### 2021-03-02 earhart/plotting.py

@@ -566,17 +566,36 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
             assert len(isocmd.isocmds) > 1
 
         elif isochrone == 'parsec':
+            #
+            # see /doc/20210226_isochrones_theory.txt
+            #
 			# v1
             # isopath = os.path.join(DATADIR, 'isochrones',
             #                        'output624293709713.dat')
-            # v2
+
+            # # v2
+            # isopath = os.path.join(DATADIR, 'isochrones',
+            #                        'output911305443923.dat')
+            # nored_iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
+            # # v3
+            # isopath = os.path.join(DATADIR, 'isochrones',
+            #                        'output4767372113.dat')
+            # iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
+
+            # v4
             isopath = os.path.join(DATADIR, 'isochrones',
-                                   'output911305443923.dat')
+                                   'output360587063784.dat')
             nored_iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
-            # v3
+
+            # v5
             isopath = os.path.join(DATADIR, 'isochrones',
-                                   'output4767372113.dat')
+                                   'output813364732851.dat')
             iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
+
+            # # v6
+            # isopath = os.path.join(DATADIR, 'isochrones',
+            #                        'output659369601708.dat')
+            # iso_df = pd.read_csv(isopath, delim_whitespace=True, comment='#')
 
 
     ##########
@@ -744,9 +763,9 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
                     (iso_df.Mass < 7)
                 )
 
-                #corr = 7.85
+                corr = 7.80
                 #corr = 7.65
-                corr = 7.60
+                #corr = 7.60
                 _yval = (
                     iso_df[sel]['Gmag'] +
                     5*np.log10(np.nanmedian(cg18_df['parallax']/1e3)) + 5
@@ -766,6 +785,11 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
                     iso_df[sel][sel2][_c0]-iso_df[sel][sel2]['G_RPmag']
                 )
 
+                ax.plot(
[…]

### 2021-03-02 cluster_paper/ms.tex
update manuscript

@@
 
-%TODO:
-%
+\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+
+\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
+\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes
+
 % n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
@@
 % n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports
-
-\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-
-%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
-
-\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
 \newcommand{\ncdips}{2,238\ } % e.g., from OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv. checked with plot_target_vs_actual_list_sanity_checks.py
@@
 \newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
-\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
 \newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
@@
 
+%FIXME these change depending on your cutoff. ensure the final numbers agree!
 \newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
@@
   around the $\approx$175\,Myr open cluster NGC\,2516.
-  %TODO: says who?
+  %TODO: Says who? Says Piskunov+08
   The classical tidal radius of the cluster is $\approx$10\,pc.
@@
 Star clusters form in hierarchically structured molecular clouds
-\citep{shu_star_1987}.  Over the first 10\,Myr, ``feedback'' effects
+\citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
 including protostellar outflows, photoionization, radiation pressure,
@@
 
+% NOTE: words in this paragraph need work
 Finding the stars that have dispersed from their natal clusters into
@@
 e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
-The converse holds too: one could use chemical tagging to test the
+The converse holds too: one can use chemical tagging to test the
 viability of kinematic clustering methods.
@@
 e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
-The resuls can be summarized as follows. On one end are coeval
-populations with no discernable cores ({\it e.g.}, Psc Eri and $\mu$
-Tau,
+The populations found by these studies can be summarized as follows.
+On one end are coeval groups with no discernable cores ({\it e.g.},
+Psc Eri and $\mu$ Tau,
 \citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
-the other are hierarchically structured populations with many over and
-under-densities  \citep[{\it e.g.} the Sco-Cen and Vela OB2
-associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
-focus on the middle: low-density halos associated with a single
+the other are hierarchically structured associations with many over
+and under-densities  \citep[{\it e.g.} Sco-Cen and Vela
+OB2][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
+an intermediate regime: low-density halos associated with a single
 overdensity, typically an open cluster that was known before Gaia
@@
 than the field around the same regions (see {\it e.g.}
-\citealt{kounkel_untangling_2019}, hereafter
-\citetalias{kounkel_untangling_2019}], and \citealt{meingast_2021},
-hereafter \citetalias{meingast_2021}).  Discrepancies between
-different clustering methods leads to the question of how contaminated
-the resulting samples are, and in some cases whether the reported
-structures are truly coeval stellar populations.
+\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
+Difference between different clustering methods raises the issue of
+how contaminated the resulting samples are, and in some cases whether
+the reported structures are truly coeval stellar populations at all.
 
 We have recently been making TESS light curves of stars reported to be
-members of such populations (CDIPS, \citealt{bouma_cdipsI_2019}).  Our
-analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
+members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
+Our analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
 \numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
-hemisphere, available on
+hemisphere, which are available on
 MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
@@
 coeval with the core?  We chose NGC\,2516 for this analysis since it
-is young ($\approx$120\,Myr) and sufficiently nearby
-($d\approx400$\,pc) to facilitate photometric rotation measurements
-using TESS, and some degree of spectroscopic analysis.  Its halo is
-decidedly non-spherical: \citetalias{kounkel_untangling_2019} reported
-its dimensions as $\approx20\times10\times350$\,pc.  We want to know:
-is the halo real? To what extent can we use Gaia alone to reliably
-identify age-dated needles in the haystack of field stars?  What are
-the implications, observationally and theoretically, if we can
-identify the dispersed halos of open clusters?
-
+is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
+to facilitate photometric rotation measurements using TESS, and some
+degree of spectroscopic analysis.  Its halo is decidedly
+non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
+as $\approx20\times10\times350$\,pc.  We want to know: is the halo
+real? To what extent can we use Gaia alone to reliably identify
+age-dated needles in the haystack of field stars?  What are the
+implications, observationally and theoretically, if we can identify
+the dispersed halos of open clusters?
+
+%TODO: double check this owrding.
 A brief note on terminology.  Low-density stellar associations
 connected to a dense population (the ``core'') have been described as
-being in ``halos'', ``strings'', ``coronae'', ``snakes'', ``the
-outskirts'', and ``tidal tails'' \citep[{\it
+being in ``halos'', ``strings'', ``coronae'', ``snakes'',
+``outskirts'', and ``tidal tails'' \citep[{\it
 e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
@@
 agnostic.  It suffers however from the suggestion of spherical
-symmetry.  The halo of NGC\,2516 is anything but.
+symmetry.  The halo of NGC\,2516 might be better-described as ``tidal
+tails'', but the primary goal of this study however is
+not to verify this claim.
 
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf The dense and diffuse components of NGC\,2516 across position and
-  velocity space.} The core (black) was identified by
-  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
-  the visual overdensity of stars canonically accepted as ``the
-  cluster''.  The halo (blue) was identified by
-  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
-  assignment algorithms described in Appendix~\ref{app:clustering}.  The field
-  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
-  centered on the cluster.  The low volume density of the
-  halo stars makes it difficult to visually distinguish
-  the field and the halo samples.
+  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
+  position and velocity space.} The core (black) was analyzed by
+  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
+  with the visual overdensity of stars canonically accepted as ``the
+  cluster''.  The halo (blue) is a concatenation of the studies by
+  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
+  used less restrictive membership assignment algorithms described in
+  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
+  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
+  low volume density of the halo stars makes it difficult to visually
+  distinguish the field and the halo samples.
   \label{fig:gaia6d}
@@
 purity and accuracy of the Gaia-derived results are the current state
-of the art.
-We therefore adopted what we viewed as the most interesting clustering
-samples to compare: those of \citet{cantatgaudin_gaia_2018}
-(\citetalias{cantatgaudin_gaia_2018}),
-\citet{kounkel_untangling_2019}
-(\citetalias{kounkel_untangling_2019}),
-and \citet{meingast_2021} (\citetalias{meingast_2021}).
-A useful visualization of these samples is available online\footnote{
+of the art.  We therefore adopted what we viewed as the most
+interesting clustering samples to compare: those of
+\citet{cantatgaudin_gaia_2018}, \citet{kounkel_untangling_2019} and
+\citet{meingast_2021}.  We refer to these studies in the following as
+\citetalias{cantatgaudin_gaia_2018},
+\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
+respectively.  A useful visualization of the different samples is
+available online\footnote{
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}.  }.
-While we could have performed our own clustering analysis based on the
-Gaia data, such an effort would be only replicating the work of these
-investigators.  We opt to instead use their studies as starting
-points.
-
-%FIXME FIXME FIXME FROM HERE
-
-% Theia 613
-\citetalias{cantatgaudin_gaia_2018} reported 1106 candidate members of
-\cn, brighter than $G=18$\,mag.  \citet{kounkel_untangling_2019}
-reported 3003 candidate members, and included stars up to
-$\approx$1\,mag fainter.  
-Finally, 1577 of the 1860 \citet{meingast_2021} sources (85\%)
-were included in the \citet{kounkel_untangling_2019} sample.  
-Each study applied unsupervised clustering techniques
-to the second Gaia Data Release, as detailed in
-Appendix~\ref{app:clustering}.
-
-%FIXME 
-\citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
-XX DEGREES} centered on the cluster's position.
-Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
-Piskunov+2008, which is 8$^\circ$ at the cluster distance of
-$\approx$400\,pc, perhaps they should have considered stars further
-out.
-
-%TODO calculate these physical sizes carefully
+  made by \citet{meingast_2021}.}.  While we considering performing
+our own clustering analysis using the Gaia data, such an effort would
+in effect only replicate the work of these investigators.  We opt to
+instead use their studies as starting points.
+
+The Gaia clustering studies each used slightly different selection
+functions, with different results.
+\citetalias{cantatgaudin_gaia_2018} considered stars brighter than
+$G=18$\,mag within a few degrees of the cluster center, and reported
+1106 candidate members of \cn.  \citetalias{kounkel_untangling_2019}
+and \citetalias{meingast_2021} considered stars up to $\approx$1\,mag
+fainter, and reported 3003 and 1860 members respectively.  The
+unsupervised clustering techniques that each of these studies applied
+to the second Gaia Data Release are discussed
+Appendix~\ref{app:clustering}, as is the overlap between their
+resulting membership samples.
+
+% %FIXME 
+% \citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
+% XX DEGREES} centered on the cluster's position.
+% Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
+% Piskunov+2008, which is 8$^\circ$ at the cluster distance of
+% $\approx$400\,pc, perhaps they should have considered stars further
+% out.
+
 Figure~\ref{fig:gaia6d} shows the cluster members reported by each
@@
 a few degrees of the cluster center, while the
-\citetalias{kounkel_untangling_2019} members span tens of degrees.  In
-physical units, this corresponds to a size difference of XX\,pc versus
-YY\,pc.  This difference is somewhat tautological, because
-\citetalias{cantatgaudin_gaia_2018} did not extend their search for
-members out to tens of degrees from the cluster center, since their
-clustering algorithm relied on the contrast between cluster and field
-stars in the astrometric space of ($\mu_{\alpha'}, \mu_\delta, \pi$).
-Nonetheless, the membership catalog of \citetalias{cantatgaudin_gaia_2018}
-echoes that of many previous investigators (CITE, CITE, CITE), and is
-consistent with the general {\it visual} impression that one has when
-looking at \cn\ visually: it seems to be at most {\bf X$^\circ$}, and
-not a big stream.
-This discrepancy begs the question.  What is the true structure of
-\cn?  Are the core and halo truly coeval?
-
-To answer these questions, we adopt the
-\citetalias{cantatgaudin_gaia_2018} stars as candidate ``core''
-members and the \citetalias{kounkel_untangling_2019} stars as
-candidate ``halo'' members.  If a star is in both lists, it is a
-``core'' member.  1091 of the 1106 core members are in both lists.
-This leaves 1912 candidate members in the halo.
-Concatenating the two lists yields \nkinematic candidate members.
-
-In this work, we also define a set of nearby field stars in the
-``neighborhood'' of \cn.  Based on the observed distribution of halo
-members, we draw these stars randomly from the following ``cube'' in
-right ascension, declination, and parallax:
+\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
+members span tens of degrees.  In physical units, the physical size of
+the entire structure is $\approx20\times10\times350$\,pc.
+As a side note, stars in the upper right quadrant of $\mu_{\alpha'}$
+{\it vs.} $\mu_\delta$ were reported almost entirely by
+\citet{meingast_2021}.
+
+In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
+describe the \citetalias{cantatgaudin_gaia_2018} members as the
+``core'' of the cluster, and any non-overlapping
+\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
+members as the ``halo''.  These definitions imply that there are
+\ncore core members, and \nhalo halo members.  This distinction is
+perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
+not extend their search for members out to tens of degrees from the
+cluster center.  Nonetheless, the membership catalog of
+\citetalias{cantatgaudin_gaia_2018} echoes that of many previous
+investigators (CITE, CITE, CITE), and is consistent with the general
+{\it visual} impression that one has when looking at \cn\ in the night
+sky: it appears to span $\approx$2$^\circ$, and not $\approx20^\circ$.
+
+The discrepancy between our canonical understanding of open clusters
+and the possibility of a diffuse population existing around \cn\
+prompted our initial interest in this analysis.  What is the true
+structure of \cn?  Are the core and halo truly coeval?
+
+Outside of the core and halo, we also define a set of nearby field
+stars in the ``neighborhood'' of \cn.  Based on the observed
+distribution of halo members, we drew these stars randomly from the
+following ``cube'' in right ascension, declination, and parallax:
 \begin{align}
@@
 \end{align}
-We imposed a magnitude limit of $G=19$\,mag, and ran the queries
-using the \texttt{astroquery.gaia} module (CITE).
-% TODO: why?
-We allowed the number of stars in the comparison sample to exceed that
-in the cluster sample by a factor of $\approx$5, to ensure broad sampling
-of stellar masses and evolutionary states.
-We also required the comparison sample to not overlap with the cluster
-sample, which led to the omission of 1.1\% of the stars drawn over the
-volume noted above.
+We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
+the \texttt{astroquery.gaia} module (CITE).  We allowed the number of
+stars in the comparison sample to exceed that in the cluster sample by
+a factor of $\approx$5, to ensure broad sampling of stellar masses and
+evolutionary states.  We also required the comparison sample to not
+overlap with the cluster sample, which led to the omission of 1.1\% of
+the stars drawn from the volume noted above.
 
@@
 The first check on the plausibility of the candidate cluster members
-being coeval was already performed by
-\citetalias{cantatgaudin_gaia_2018},
+being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.  Comparable analyses
+have already been performed by \citetalias{cantatgaudin_gaia_2018},
 \citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.
-That check is to see whether the HR diagrams of the cluster members
-support the claim that they are coeval.
-
-Figure~\ref{fig:hr} presents similar results to what these
-investigators have already found.  The core members of the cluster
-show a clean sequence consistent with stars with a fixed age and
-metallicity, and varying mass.  The halo members show a roughly
-consistent main-sequence, with somewhat greater scatter.  One
-possible explanation for the additional scatter could be that the
-halo is more contaminated by field stars.  For instance, $\approx5$
-evolved halo stars are visible in the field-star portion of the red
-giant branch (RGB).  Given the age of the system, these stars must be
-field interlopers.
+
+Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
+$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
+members of the cluster show a clean sequence consistent with stars
+with a fixed age and metallicity, and varying mass.  The halo members
+mostly show a consistent main-sequence that has somewhat greater
+scatter.  One possible explanation for the additional scatter could be
+that the halo is more contaminated by field stars.  For instance,
+$\approx5$ evolved halo stars are visible in the field-star portion of
+the red giant branch (RGB).  Given the age of the system, these stars
+must be field interlopers.
 
 A separate possible explanation for scatter in the halo's HR diagram
-could be differential reddening across different sightlines.  The halo
-is reported to span 20$^\circ$ on-sky, and varies in position from
+could be differential reddening across different sightlines.  The
+reported halo spans 20$^\circ$ on-sky, and varies in position from
 about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
@@
 (Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
-reddening is discussed further in Appendix~\ref{app:diffred}. Based on
-the available evidence, we expect that both field star contamination
-and differential reddening could play a role in the larger scatter of
-the halo relative to the core.  A third possibility that we have not
+reddening is discussed in Appendix~\ref{app:diffred}. Based on the
+available evidence, we expect that both field star contamination and
+differential reddening could play a role in the larger scatter of the
+halo relative to the core.  A third possibility that we have not
 explored is whether the binary fraction could also differ between the
@@
 In the right panel of Figure~\ref{fig:hr}, we compare the observed
-photometry with PARSEC isochrones
+Gaia EDR3 photometry with PARSEC isochrones
 \citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
@@
 the reddening correction described in Section~\ref{subsec:tess}.  We
-assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applyied
+assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
 the \citet{odonnell_1994} SED-dependent extinction law star-by-star
@@
 an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
-\citet{bailey_rv_2018} report a slightly sub-solar metallicity for
-the cluster, we prefer super-solar based on the goodness of fit from
-the photometry.
-
-Overall, the data and models agree for masses above
-approximately 0.45\,$M_\odot$.  However below this mass, the data and
-models diverge at colors redder than Bp-Rp$\approx2.2$.  The MIST
-isochrones showed a comparable disagreement \citep{choi_mesa_2016}.
-Explanations invoking both starspots and incomplete molecular line
-lists have been proposed \citep[{\it
+\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
+cluster, we prefer super-solar based on the goodness of fit from the
+photometry.
+
+Overall, the data and models agree for masses above approximately
+0.45\,$M_\odot$.  However below this mass, the data and models diverge
+at colors redder than Bp-Rp$\approx2.2$.  The MIST isochrones showed a
+comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
+both starspots and incomplete molecular line lists have been proposed
+\citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
@@
 Our primary goal is to determine whether the ages of the core and halo
-are consistent.  The left panel of Figure~\ref{fig:hr} shows
-empirically that they are: stars past the turnoff in both the halo and
-core are on a consistent locus.  The right panel of
-Figure~\ref{fig:hr} demonstrates the precision with which the claim can be
-made.  The MSTO stars are consistent with the 178 and 316 Myr models,
-but are bluer than the 100 Myr model.  The RGB stars are consistent
-only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
-the assumption that the width of the MSTO can be attributed to binary
-stars, we are most interested in the blue edge. This appears most
-compatible with the 178 Myr model.  These results are consistent with
-the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
-and we refer to their work for a more precise model-dependent
-comparison.
+are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
+they are: stars past the turnoff in both the halo and core are on a
+consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
+the precision with which the claim can be made.  The MSTO stars are
+consistent with the 178 and 316 Myr models, but are bluer than the 100
+Myr model.  The RGB stars are consistent only with the 178 Myr
+($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
+the width of the MSTO can be attributed to binary stars, we are most
+interested in the blue edge. This appears most compatible with the 178
+Myr model.  These results are consistent with the MSTO age range of
+165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
+work for a more precise model-dependent comparison.
 
@@
 % given the Cummings study.
+
 % Our photometric age-dating procedure is as follows.  We selected all
@@
 % For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
-% yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
+% yields $XXX\pm YYY$\,Myr.  Finally, applying the same procedure to
 % the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
+% [nb. the procedure should be simple and general enough that you can
+% apply it to the field sample, and it'll just give a broad
+% distribution...]
 
@@
 
+\citet{meingast_2021} reported a 
+Finally, 1577 of the 1860 \citet{meingast_2021} sources (85\%)
+were included in the \citet{kounkel_untangling_2019} sample.
+
+{\bf TODO: ADD A VENN DIAGRAM}
+
 \section{Detrending details}

### 2021-03-03 cluster_paper/collect.sh
add venn diagram

@@ -23,6 +23,7 @@ cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_ed
 
 cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear.pdf f9.pdf
 
+cp ../results/venn/venn.pdf f10.pdf
 
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf

### 2021-03-03 drivers/plot_venn.py

@@ -1,0 +1,13 @@
+"""
+Plot venn diagram of overlapping cluster memberships
+"""
+
+import os
+import earhart.plotting as ep
+from earhart.paths import RESULTSDIR
+
+PLOTDIR = os.path.join(RESULTSDIR, 'venn')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+ep.plot_venn(PLOTDIR)

### 2021-03-03 earhart/helpers.py

@@ -406,7 +406,11 @@ def _get_fullfaint_dataframes():
     print(f'Got {len(full_df)} unique sources in the cluster.')
 
     full_df['in_CG18'] = full_df.source_id.isin(cg18_df.source_id)
+    kc19_df = pd.read_csv(kc19path)
+    kc19_df = kc19_df[kc19_df.group_id == 613]
     full_df['in_KC19'] = full_df.source_id.isin(kc19_df.source_id)
+    m21_df = Table(fits.open(m21path)[1].data).to_pandas()
+    m21_df = m21_df.rename(mapper={'GaiaDR2': 'source_id'}, axis=1)
     full_df['in_M21'] = full_df.source_id.isin(m21_df.source_id)
 
     return nbhd_df, core_df, halo_df, full_df, target_df

### 2021-03-03 earhart/plotting.py

@@ -24,6 +24,7 @@ Contents:
         plot_edr3_blending_vs_apparentmag
         plot_bisector_span_vs_RV
         plot_backintegration_ngc2516
+        plot_venn
 """
 import os, corner, pickle
 from glob import glob
@@ -2647,3 +2648,31 @@ def plot_physical_X_rotation(outdir, basedata='bright', show1937=0,
         outpath = os.path.join(outdir, f'rot_df_physical_X_rotation{s}.csv')
         rot_df.to_csv(outpath, index=False)
         print(f'Made {outpath}')
+
+
+def plot_venn(outdir):
+
+    basedata = 'fullfaint'
+    nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
+
+    cg18_set = set(full_df[full_df.in_CG18].source_id)
+    kc19_set = set(full_df[full_df.in_KC19].source_id)
+    m21_set = set(full_df[full_df.in_M21].source_id)
+
+    from matplotlib_venn import venn3, venn3_circles
+
+    fig, ax = plt.subplots(figsize=(4,4))
+
+    v = venn3(
+        [cg18_set, kc19_set, m21_set],
+        set_labels=('CG18', 'KC19', 'M21'),
+        normalize_to=1e1
+    )
+
+    v.get_patch_by_id('100').set_alpha(0.2)
+    v.get_patch_by_id('010').set_alpha(0.2)
+    v.get_patch_by_id('001').set_alpha(0.2)
+
+    outpath = os.path.join(outdir, f'venn.png')
+    savefig(fig, outpath)
+

### 2021-03-04 cluster_paper/ms.tex
update manuscript
add venn diagram
update ms

@@
 
-% Context, aims, methods, results, conclusions.
+% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
+% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
+% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
+% 7.70pc.
 \begin{abstract}
@@
   around the $\approx$175\,Myr open cluster NGC\,2516.
-  %TODO: Says who? Says Piskunov+08
-  The classical tidal radius of the cluster is $\approx$10\,pc.
+  The classical tidal radius of the cluster is $\lesssim$10\,pc.
   Combining photometry from Gaia, rotation periods from TESS, and
@@
 
-Star clusters form in hierarchically structured molecular clouds
+Star clusters form from the densest filaments of 
+hierarchically structured molecular clouds
 \citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
@@
 and supernova shocks disperse the gas out of the cloud
-\citep{krumholz_star_2019}.  Since only a small fraction of the cloud
-mass is converted into stars ($\sim 1\%$), gas dispersal enables the
+\citep{krumholz_star_2019}.  Since only $\sim 1\%$ of the cloud
+mass is converted into stars, gas dispersal enables the
 majority ($\sim$90\%) of stars in the cluster to escape the cluster's
@@
 From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
-suffer an onslaught that almost always leads to dissolution.  Within
+suffer an onslaught that almost always leads to dissolution.  Inside
 the cluster supernovae, AGB winds, and close stellar encounters lead
@@
 
-% NOTE: words in this paragraph need work
 Finding the stars that have dispersed from their natal clusters into
-the galactic field is of interest for a few reasons.  One is to
-understand the predominant conditions under which stars and star
-clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
+the galactic field is a pressing topic for a few reasons.  One is to
+understand the spatial extent and duration of star formation events
+\citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
 is to understand the Sun's birth environment \citep{adams_birth_2010}.
@@
 ``chemical tagging'' \citep[{\it
-e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
+e.g.},][]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
 The converse holds too: one can use chemical tagging to test the
@@
 the other are hierarchically structured associations with many over
-and under-densities  \citep[{\it e.g.} Sco-Cen and Vela
-OB2][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
+and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
+associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
 an intermediate regime: low-density halos associated with a single
@@
 \citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
-Difference between different clustering methods raises the issue of
+Differences between different clustering methods raise the issue of
 how contaminated the resulting samples are, and in some cases whether
@@
 members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
-Our analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
+Our analysis of TESS Sectors 1-13 yielded light curves of
 \numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
@@
 As part of this broader project, we focus here on a modest question:
-in just a single rich southern open cluster, is the cluster halo truly
+in just a single southern open cluster, is the cluster halo truly
 coeval with the core?  We chose NGC\,2516 for this analysis since it
@@
 The latter term implies a particular model for the formation of the
-association.  We opt to use ``halo'' because it is concise and model
+association.  We opt for ``halo'' because it is concise and model
 agnostic.  It suffers however from the suggestion of spherical
-symmetry.  The halo of NGC\,2516 might be better-described as ``tidal
-tails'', but the primary goal of this study however is
-not to verify this claim.
+symmetry.  The halo of NGC\,2516 might be best-described as ``tails'',
+but the primary goal of this study is
+not to verify whether these tails are tidal.
 
@@
 
-\subsection{Gaia Astrometry}
-\label{subsec:astrometry}
+%\subsection{Gaia Astrometry}
+%\label{subsec:astrometry}
 
@@
 
-We selected candidate \cn\ members based on those available in the
-literature.  While a number of pre-Gaia catalogs were available, the
+We selected candidate \cn\ members based on those reported in the
+literature.  While a number of pre-Gaia 
+studies have been performed, the
 purity and accuracy of the Gaia-derived results are the current state
-of the art.  We therefore adopted what we viewed as the most
+of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore adopted what we viewed as the most
 interesting clustering samples to compare: those of
@@
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}.}.  While we considering performing
+  made by \citet{meingast_2021}.}.  While we considered performing
 our own clustering analysis using the Gaia data, such an effort would
@@
 
-The Gaia clustering studies each used slightly different selection
-functions, with different results.
-\citetalias{cantatgaudin_gaia_2018} considered stars brighter than
-$G=18$\,mag within a few degrees of the cluster center, and reported
-1106 candidate members of \cn.  \citetalias{kounkel_untangling_2019}
-and \citetalias{meingast_2021} considered stars up to $\approx$1\,mag
-fainter, and reported 3003 and 1860 members respectively.  The
-unsupervised clustering techniques that each of these studies applied
-to the second Gaia Data Release are discussed
-Appendix~\ref{app:clustering}, as is the overlap between their
-resulting membership samples.
-
-% %FIXME 
-% \citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
-% XX DEGREES} centered on the cluster's position.
-% Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
-% Piskunov+2008, which is 8$^\circ$ at the cluster distance of
-% $\approx$400\,pc, perhaps they should have considered stars further
-% out.
+The Gaia clustering studies each used different selection functions,
+which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
+considered stars brighter than $G=18$\,mag within a few degrees of the
+cluster center, and reported 1106 candidate members of \cn.
+\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
+considered stars up to $\approx$1\,mag fainter, and reported 3003 and
+1860 members respectively.  The unsupervised clustering techniques
+that each of these studies applied to the second Gaia data release are
+discussed Appendix~\ref{app:clustering}, as is the overlap between
+their resulting membership samples.
 
 Figure~\ref{fig:gaia6d} shows the cluster members reported by each
-study.  The \citetalias{cantatgaudin_gaia_2018} members are all within
-a few degrees of the cluster center, while the
+study in the space of positions, proper motions, and radial velocity
+when available.  The \citetalias{cantatgaudin_gaia_2018} members are
+all within a few degrees of the cluster center, while the
 \citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
-members span tens of degrees.  In physical units, the physical size of
-the entire structure is $\approx20\times10\times350$\,pc.
-As a side note, stars in the upper right quadrant of $\mu_{\alpha'}$
-{\it vs.} $\mu_\delta$ were reported almost entirely by
+members (the ``halo'') span tens of degrees. 
+Inverting the parallax shows that members have been reported from
+line-of-sight distances ranging from $300$ to $600\,$pc.
+Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
+$\mu_\delta$ projection were reported almost entirely by
 \citet{meingast_2021}.
@@
 \citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
-members as the ``halo''.  These definitions imply that there are
-\ncore core members, and \nhalo halo members.  This distinction is
+members as the ``halo''.  This defintion implies that there are
+\ncore\ core members, and \nhalo\ halo members.  The distinction is
 perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
 not extend their search for members out to tens of degrees from the
-cluster center.  Nonetheless, the membership catalog of
-\citetalias{cantatgaudin_gaia_2018} echoes that of many previous
-investigators (CITE, CITE, CITE), and is consistent with the general
-{\it visual} impression that one has when looking at \cn\ in the night
-sky: it appears to span $\approx$2$^\circ$, and not $\approx20^\circ$.
+cluster center.  Nonetheless, the \citetalias{cantatgaudin_gaia_2018}
+membership catalog is consistent with that of many previous
+investigators \citep[{\it
+e.g.},][]{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, and is
+consistent with the general {\it visual} impression that one has when
+looking at \cn\ in the night sky: it appears to span
+$\approx$2$^\circ$, not $\approx20^\circ$.
 
@@
 We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
-the \texttt{astroquery.gaia} module (CITE).  We allowed the number of
-stars in the comparison sample to exceed that in the cluster sample by
-a factor of $\approx$5, to ensure broad sampling of stellar masses and
-evolutionary states.  We also required the comparison sample to not
-overlap with the cluster sample, which led to the omission of 1.1\% of
-the stars drawn from the volume noted above.
+the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
+allowed the number of stars in the comparison sample to exceed that in
+the cluster sample by a factor of $\approx$5, to ensure broad sampling
+of stellar masses and evolutionary states.  We also required the
+comparison sample to not overlap with the cluster sample, which led to
+the omission of 1.1\% of the stars drawn from the volume noted above.
 
@@
     to the models.
-    The main-sequence turn-off location and evolved stars suggests a
-    cluster age slightly older than the Pleiades (dark blue line).  The models diverge
+     The models diverge
     from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
-    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
+    $Bp-Rp\approx2.2$).
     \label{fig:hr}
@@
 about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
-galactic plane also being further from the Sun by up to 200\,pc
+galactic plane also being further from the Sun by up to 300\,pc
 (Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
@@
 work for a more precise model-dependent comparison.
+This implies an absolute age slightly older than the Pleiades
+\citep[{\it cf.}][]{mermilliod_comparative_1981}.
 
@@
     period and dereddened Gaia color.}
-    % The {\it top} plot shows periods against a linear scale, while the
-    % {\it bottom}  shows them against a logarithmic scale.
     The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
@@
 
-First, we collected the \nkinematic Gaia DR2 source\_ids corresponding to the
-kinematically-identified core and halo members.  For each source, we first retrieved all
-available CDIPS light curves, on a per-sector basis. 
-This yielded
-\ncdips light curves, all brighter than $Rp=16$\,mag.
-The $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
-roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
-the distance of NGC\,2516.
+%FIXME: except actually, will need to redo with the new source list
+%too .......
+First, we collected the \nkinematic Gaia DR2 source\_ids corresponding
+to the kinematically-identified core and halo members.  For each
+source, we first retrieved all available CDIPS light curves, on a
+per-sector basis.  This yielded \ncdips light curves, all brighter
+than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
+processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
+type of $\approx$M2V, at the distance of NGC\,2516.
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f8a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f8b.pdf}
-		}
+			\includegraphics[width=0.49\textwidth]{f8a.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
-    velocity space, cross-matched against the rotators, in physical
-    units.}
-    {\it Top:} projections in position and 2d-tangential velocity
-    space. {\it Bottom:} radially binned versus 3d position and
-    2d velocity difference.
-    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
-    velocity.
+  \caption{ {\bf Kinematic members of NGC\,2516, cross-matched against
+  rotational members.}
+    The projections are in galactic position coordinates and 2d-tangential velocity
+    space.
     \label{fig:physical_x_rotn}
 	}
-\end{figure*}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{center}
+			\includegraphics[width=0.49\textwidth]{f8b.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
+  versus 3d separation from the core and 2d velocity difference.} Bins have width
+  25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
+  {\bf Uncertainties in the fraction are taken as the quadrature sum of
+  the rel unc of the numerator and denominator (FIXME)}.
+  Bonafide members appear to exist hundreds of parsecs from the core, and
+  up to $\approx$10\,km\,s$^{-1}$ in tangential velocity separation.
+  \label{fig:hist_physical_x_rotn}
+	}
+\end{figure}
+
+%\begin{figure*}[t]
+%	\begin{center}
+%		\leavevmode
+%		\subfloat{
+%			\includegraphics[width=0.75\textwidth]{f8a.pdf}
+%		}
+%	
+%		\subfloat{
+%			\includegraphics[width=0.75\textwidth]{f8b.pdf}
+%		}
+%	\end{center}
+%	\vspace{-0.7cm}
+%	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
+%    velocity space, cross-matched against the rotators, in physical
+%    units.}
+%    {\it Top:} projections in position and 2d-tangential velocity
+%    space. {\it Bottom:} radially binned versus 3d position and
+%    2d velocity difference.
+%    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
+%    velocity.
+%		\label{fig:physical_x_rotn}
+%	}
+%\end{figure*}
 
@@
 \appendix
-\section{Clustering method details}
+\section{Clustering methods and outcomes}
 \label{app:clustering}
+
+\begin{figure*}[h]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.4\textwidth]{f10.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Different clustering techniques yield different
+  outcomes.} \nkinematic unique candidate cluster members found
+  using three different techniques are considered in this study.
+  \citetalias{cantatgaudin_gaia_2018}: \citet{cantatgaudin_gaia_2018},
+  \citetalias{kounkel_untangling_2019}:
+  \citet{kounkel_untangling_2019}, \citetalias{meingast_2021}:
+  \citet{meingast_2021}.
+  \label{fig:venn}
+	}
+\end{figure*}
+
+Figure~\ref{fig:venn} is a Venn diagram showing the overlap between
+the three membership catalogs concatenated in this study.  99\% of the
+\citetalias{cantatgaudin_gaia_2018} sample overlaps with
+\citetalias{kounkel_untangling_2019}, \citetalias{meingast_2021}, or
+both.  By comparison, 36\% of the \citetalias{kounkel_untangling_2019}
+sample, and 15\% of the \citetalias{meingast_2021} sample are
+non-overlapping.
+The data---Gaia DR2---used by all the studies was the same.
+What are the differences in methodology that produce these different
+outcomes?
 
@@
 
-{\bf TODO: ADD A VENN DIAGRAM}
-
 \section{Detrending details}

### 2021-03-04 cluster_paper/collect.sh

@@ -1,4 +1,5 @@
-cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
+#cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
+cp ../results/cluster_membership/full_kinematics_fullfaint_galactic.pdf f1.pdf
 
 cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f2a.pdf
 cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f2b.pdf

### 2021-03-05 cluster_paper/ms.tex
update ms

@@
   In this study, we focus on a halo with a length of 300\,pc reported
-  around the $\approx$175\,Myr open cluster NGC\,2516.
+  around the $\approx$150\,Myr open cluster NGC\,2516.
   The classical tidal radius of the cluster is $\lesssim$10\,pc.
@@
 from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
-super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
-predictions too blue along the main sequence by $\approx$0.1\,mag.
-Our adopted metallicity agrees with that found by
+super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.  Sub-solar
+metallicities led to model predictions too blue along the main
+sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees with
+the spectroscopic metallicity from
 \citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
@@
 0.45\,$M_\odot$.  However below this mass, the data and models diverge
-at colors redder than Bp-Rp$\approx2.2$.  The MIST isochrones showed a
-comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
-both starspots and incomplete molecular line lists have been proposed
-\citep[{\it
+at colors redder than Bp-Rp$\approx2.2$, in the sense that the model
+isochrones have lower luminosities and bluer colors than the
+observations.  The MIST isochrones showed a comparable disagreement
+\citep{choi_mesa_2016}.  Explanations invoking both starspots and
+incomplete molecular line lists have been proposed \citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
@@
 implemented by \citet{chen_improving_2014}.  For purposes of
-age-dating the cluster, we prefer to simply focus on the main-sequence
-turn-off (MSTO), since it is where the models have maximal predictive
+age-dating the cluster, we prefer to focus on the main-sequence
+turn-off (MSTO), since this is where the models have maximal predictive
 power.
@@
 
+We began our rotation analysis by considering all \nkinematic\
+candidate cluster members.  For each source, we first retrieved all
+available CDIPS light curves, on a per-sector basis.
+
 %FIXME: except actually, will need to redo with the new source list
 %too .......
-First, we collected the \nkinematic Gaia DR2 source\_ids corresponding
-to the kinematically-identified core and halo members.  For each
-source, we first retrieved all available CDIPS light curves, on a
-per-sector basis.  This yielded \ncdips light curves, all brighter
+
+This yielded \ncdips light curves, all brighter
 than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
@@
 reddening corrections.
+% FIXME: where did these reference tables come from???
 We assumed {X} for the Pleiades, and {Y} for Praesepe.
-For NGC\,2516, we queried the 2MASS DUST service to find the
-extinction parameters for the positions of each NGC\,2516 member
-(CITE: astrobase.services.dust; 2MASS DUST).  The mean and standard
-deviation of the E(B-V) values for the Schlegel+98 and SF11 maps were
-consistent, so we considered those of Schlegel+1998:
-$E(B-V)=0.206\pm0.039$.  Bonifacio+2000 noted though that Schlegel+98
-overestimate the reddening values when the color excess exceeds about
-0.10 mag. We applied their correction (Eq 1 from Bonifacio+2000),
-which led to an adopted $E(B-V)=0.102\pm0.019$.  Finally, to convert
-to Gaia colors we adopted the calibration of Stassun et al 2019, and
-assumed $E(Bp-Rp)=1.31 E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded
-$E(Bp-Rp)=0.134\pm0.025$, which is used in the plots.
+For NGC\,2516, we used the \texttt{astrobase.services.dust} module to
+query the 2MASS DUST
+service\footnote{\url{http://irsa.ipac.caltech.edu/applications/DUST}}
+and retrieve the extinction parameters for the positions of each
+NGC\,2516 member \citep{bhatti_astrobase_2018}.  The mean and standard
+deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
+and \citet{schlafly_measuring_2011} maps were consistent, so we
+adopted the average from \citet{schlegel_maps_1998}:
+$E(B-V)=0.206\pm0.039$.  \citet{bonifacio_search_2000}
+\citet{schlegel_maps_1998} maps overestimate the reddening values when
+the color excess exceeds about 0.10 mag. Applying the correction
+proposed by \citet{bonifacio_search_2000} leads to our adopted
+$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we adopted
+the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
+E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
+which is used in the plots.
 
@@
 
+\subsection{What is the contamination fraction in the halo? Does it change vs. location?}
+
 \subsection{Are the ``very slow rotators'' bonafide members?}
@@
 
-\subsection{What is the contamination fraction in the halo? Does it change vs. location?}
-
 \subsection{Mass differences between center and outer reaches?}
@@
 \subsection{Fast and blue rotators: are we going faster than other clusters?}
+
+\subsection{Absolute age of NGC 2516}
+Cummings & Kalirai (2018) favor 150 Myr based on MSTO.
+Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
+"""
+Similar to Cummings et al. (2016), Cummings & Kalirai (2018),
+and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
+&
+McCullough 2020)
+"""
 

### 2021-03-05 cluster_paper/ms.tex
verified Gaia target list vs CDIPS results

@@
 
-% n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
-% ... which came from ~/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split/OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv
-% n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports
-\newcommand{\ncdips}{2,238\ } % e.g., from OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv. checked with plot_target_vs_actual_list_sanity_checks.py
+% from drivers.cluster_rotation.get_auto_rotation_periods.py
+% checked with plot_target_vs_actual_list_sanity_checks.py
+\newcommand{\ncdips}{2{,}205\ }
+\newcommand{\nkinematicrpltsixteen}{2{,}270\ }
+\newcommand{\nkinematicCGandKC}{2{,}236\ }
 
@@
 We began our rotation analysis by considering all \nkinematic\
-candidate cluster members.  For each source, we first retrieved all
-available CDIPS light curves, on a per-sector basis.
-
-%FIXME: except actually, will need to redo with the new source list
-%too .......
-
-This yielded \ncdips light curves, all brighter
-than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
+candidate members of \cn.  For each source, we first retrieved all
+available CDIPS light curves, on a per-sector basis.  This yielded
+\ncdips stars with at least one sector from a CDIPS light curve, all
+brighter than $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the
+stars from the source list have $Rp<16$.  The difference is comprised
+of 34 stars unique to \citet{meingast_2021} which were not
+available at the time of the CDIPS reductions and stars falling on
+the TESS chip gaps.  The $Rp=16$\,mag cutoff imposed during the CDIPS
 processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
@@
 
-We cleaned and stitched the light curves as follows.
-First, we detrended the systematics in each sector individually, and
+We removed systematic trends shared between
+all light curves on each CCD in each sector individually, and
 stitched together the resulting light curves before searching for the
-periodicity.  Some details regarding our detrending approach are
-discussed in Appendix~\ref{app:detrending}.  After applying a
+periodicity.  Details regarding our detrending approach are
+discussed in Appendix~\ref{app:detrending}.  After applying the
+initial
 detrending step aimed at removing systematic trends, we proceeding
@@
 We then measured the rotation period of the resulting light curve.  We
-used the the aperture radius that, based on theoretical expectations,
+used the CDIPS aperture radius that, based on theoretical expectations,
 was expected to give the optimal balance between light from the target
-and background-light (CITE Sullivan15).  This typically resulted in an
+and background-light \citep{Sullivan_2015}.  This typically resulted in an
 aperture radius of either 1 or 1.5 pixels.  To measure the periods, we
@@
 \subsection{Absolute age of NGC 2516}
-Cummings & Kalirai (2018) favor 150 Myr based on MSTO.
+Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
 Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
 """
-Similar to Cummings et al. (2016), Cummings & Kalirai (2018),
+Similar to Cummings et al. (2016), Cummings \& Kalirai (2018),
 and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
-&
+\&
 McCullough 2020)

### 2021-03-05 drivers/cluster_rotation/get_auto_rotation_periods.py

@@ -51,10 +51,11 @@ def get_auto_rotation_periods(
 
     # retrieve the LS periods. only top period; since we're not bothering with
     # the "second period" classification option.
-    periods, lspvals, nequal, nclose, nfaint = [], [], [], [], []
+    periods, lspvals, nequal, nclose, nfaint, nsectors = [], [], [], [], [], []
     if get_spdm:
         spdmperiods, spdmvals = [], []
 
+    ix = 0
     for source_id, logpath in zip(source_ids, logfiles):
 
         s = load_status(logpath)
@@ -70,6 +71,8 @@ def get_auto_rotation_periods(
                 spdmperiods.append(float(d['spdm']['bestperiod']))
                 spdmvals.append(float(d['spdm']['bestlspval']))
 
+        n_sectors = int(s['lc_info']['n_sectors'])
+        nsectors.append(n_sectors)
         try:
             periods.append(float(s['report_info']['ls_period']))
             lspvals.append(float(s['report_info']['bestlspval']))
@@ -83,9 +86,9 @@ def get_auto_rotation_periods(
             nclose.append(np.nan)
             nfaint.append(np.nan)
 
-
     period_df = pd.DataFrame({
         'source_id': source_ids,
+        'n_cdips_sector': nsectors,
         'period': periods,
         'lspval': lspvals,
         'nequal': nequal,
@@ -96,14 +99,20 @@ def get_auto_rotation_periods(
         period_df['spdmperiod'] = spdmperiods
         period_df['spdmval'] = spdmvals
 
+    print(f'Got {len(period_df[period_df.n_cdips_sector > 0])} sources with at least 1 cdips sector')
     print(f'Got {len(period_df[~pd.isnull(period_df.period)])} periods')
 
     # get the runid's source list
     if 'compstar' not in runid:
-        sourcelistpath = os.path.join(
-            '/Users/luke/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split',
-            f'OC_MG_FINAL_v0.4_publishable_CUT_{runid}.csv'
-        )
+        if runid == 'NGC_2516':
+            sourcelistpath = (
+                '/Users/luke/Dropbox/proj/cdips/data/cluster_data/NGC_2516_full_fullfaint_20210305.csv'
+            )
+        else:
+            sourcelistpath = os.path.join(
[…]

### 2021-03-05 drivers/plot_target_vs_actual_list_sanity_checks.py

@@ -1,16 +1,19 @@
 """
-The 3018 candidate cluster members are split into the "core" and
-"halo" samples, below. Source ID's are from Gaia DR2.
+The 3298 candidate cluster members are split into the "core" and "halo" samples
+(CG18, KC19+M21 respectively). Source ID's are from Gaia DR2.
 
-For rotation periods, 2238 of the stars had light curves produced.
+How many had CDIPS light curves produced?
 
-Where are the other 780?
+Which ones did not, and why?
 """
 
 import os
 import numpy as np, matplotlib.pyplot as plt, pandas as pd
+from earhart.helpers import get_gaia_basedata
 from earhart.paths import RESULTSDIR, DATADIR
 from aesthetic.plot import savefig, format_ax, set_style
+
+set_style()
 
 PLOTDIR = os.path.join(
     RESULTSDIR, 'target_vs_actual_list_sanity_checks'
@@ -18,21 +21,16 @@ PLOTDIR = os.path.join(
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-halo_df = pd.read_csv(
-    os.path.join(RESULTSDIR, 'tables', 'NGC_2516_halo_fullfaint.csv')
-)
-core_df = pd.read_csv(
-    os.path.join(RESULTSDIR, 'tables', 'NGC_2516_core_fullfaint.csv')
-)
-ngc_df = pd.concat((core_df, halo_df))
-print(f'NGC candidate member total: {len(ngc_df)}')
 
+runid = "NGC_2516"
+basedata = 'fullfaint'
+nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
+
+# made by drivers.cluster_rotation.get_auto_rotation_periods
 rot_df = pd.read_csv(
     os.path.join(DATADIR, 'rotation', 'NGC_2516_rotation_periods.csv')
 )
-lccut_df = pd.read_csv(
-    '/Users/luke/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split/OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv'
-)
+lccut_df = rot_df[rot_df.n_cdips_sector>0]
[…]

### 2021-03-05 drivers/write_NGC2516_sourcelist_table.py

@@ -14,16 +14,13 @@ neighborhood members too.
 import os, corner, pickle
 import numpy as np, matplotlib.pyplot as plt, pandas as pd
 
-from earhart.helpers import (
-    get_gaia_basedata
-)
+from earhart.helpers import get_gaia_basedata
 from earhart.paths import RESULTSDIR
 
 OUTDIR = os.path.join(RESULTSDIR, 'tables')
 
 runid = "NGC_2516"
 basedata = 'fullfaint'
-
 nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
 
 dfdict = {

### 2021-03-06 cluster_paper/ms.tex
update ms

@@
 
-\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-
+% gaia target sample numbers
 \newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
@@
 
-% from drivers.cluster_rotation.get_auto_rotation_periods.py
-% checked with plot_target_vs_actual_list_sanity_checks.py
+% from drivers.cluster_rotation.get_auto_rotation_periods.py;  checked with plot_target_vs_actual_list_sanity_checks.py
 \newcommand{\ncdips}{2{,}205\ }
@@
 
+% cdips results numbers
+\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+
 \newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
@@
 
-%FIXME these change depending on your cutoff. ensure the final numbers agree!
 \newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
 \newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\nautorotnumeratormatching}{507\ } % plot_autorotation.py, LSP=SPDM period
 \newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
@@
 corrections as in \citet{chen_2019_YBC}.} to interpolate these
-isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
-the reddening correction described in Section~\ref{subsec:tess}.  We
-assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
-the \citet{odonnell_1994} SED-dependent extinction law star-by-star
-through the PARSEC interface.  For the metallicity, we considered a
-range of super and sub-solar metallicities to fit as much of the locus
-from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
-super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.  Sub-solar
-metallicities led to model predictions too blue along the main
-sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees with
-the spectroscopic metallicity from
+isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
+
+To determine the reddening correction across the NGC\,2516 cluster, we
+queried the 2MASS
+DUST
+service\footnote{
+	\url{http://irsa.ipac.caltech.edu/applications/DUST};
+	query performed using the \texttt{astrobase.services.dust} module
+	\citep{bhatti_astrobase_2018}.
+}
+and retrieved the extinction parameters at the positions of each
+NGC\,2516 member.  The mean and standard
+deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
+and \citet{schlafly_measuring_2011} maps were consistent, so we
+adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
+S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
+the \citet{schlegel_maps_1998} maps overestimate the reddening values
+when the color excess exceeds about 0.10\,mag. Applying the correction
+proposed by \citet{bonifacio_search_2000} leads to our adopted
+$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
+the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
+E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
+which is used in the plots.  To redden the isochrones, we assumed
+$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
+extinction law star-by-star through the PARSEC interface.
+
+For the metallicity, we considered a range of super and sub-solar
+metallicities to fit as much of the locus from $0.5<Bp-Rp<1.5$ as
+possible, and settled on the slightly super-solar $[M/H]=0.06$
+\citep{cummings_2011_li_iron}.  Sub-solar metallicities led to model
+predictions too blue along the main sequence by $\approx$0.1\,mag.
+Our adopted metallicity agrees with the spectroscopic metallicity from
 \citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
@@
 		\leavevmode
-    \includegraphics[width=0.75\textwidth]{f3a.pdf}
-		% \subfloat{
-		% 	\includegraphics[width=0.75\textwidth]{f3a.pdf}
-		% }
-
-		% \subfloat{
-		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
-		% }
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f3a.pdf}
+			\includegraphics[width=0.49\textwidth]{f3b.pdf}
+		}
+	   %
+		%\subfloat{
+		%	\includegraphics[width=0.6\textwidth]{f3b.pdf}
+		%}
 	\end{center}
@@
     period and dereddened Gaia color.}
-    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
-    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
-    are shown for reference.
+    Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
+    undergo successive degrees of automated cleaning, described in the
+    text.
+    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a} and
+    Praesepe \citep[650\,Myr][]{douglas_poking_2017} are shown for
+    reference.
     Stars in the core and the halo of NGC\,2516 appear
@@
 \ncdips stars with at least one sector from a CDIPS light curve, all
-brighter than $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the
-stars from the source list have $Rp<16$.  The difference is comprised
-of 34 stars unique to \citet{meingast_2021} which were not
-available at the time of the CDIPS reductions and stars falling on
-the TESS chip gaps.  The $Rp=16$\,mag cutoff imposed during the CDIPS
-processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
-type of $\approx$M2V, at the distance of NGC\,2516.
-
-We removed systematic trends shared between
-all light curves on each CCD in each sector individually, and
-stitched together the resulting light curves before searching for the
-periodicity.  Details regarding our detrending approach are
-discussed in Appendix~\ref{app:detrending}.  After applying the
-initial
-detrending step aimed at removing systematic trends, we proceeding
-with a few small cleaning steps aimed at improving the purity of the
-rotation period measurements: we masked 0.7 days at the beginning and
-end of each spacecraft orbit, and ran a sliding standard-deviation
-rejection window over the light curve, which removed any outlying
-points within $\pm3\times$MAD of the median in each window.
-
-We then measured the rotation period of the resulting light curve.  We
-used the CDIPS aperture radius that, based on theoretical expectations,
-was expected to give the optimal balance between light from the target
-and background-light \citep{Sullivan_2015}.  This typically resulted in an
-aperture radius of either 1 or 1.5 pixels.  To measure the periods, we
-used the periodogram implementations in \texttt{astrobase}, in
-particular the Stellingwerf PDM periodogram (CITE), along with the
-more traditional Lomb-Scargle (CITE).  We recorded the top five
-periodogram peaks from each method, and their corresponding powers.
-Finally, as a check on crowding, we also recorded the number of stars
-within the aperture of equal brightness to the target stars, and of
-brightness with 1.25 and 2.5 TESS magnitudes of the target star.
-
-Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
-on this plot are the \nautorotnumerator light curves for which the peak Lomb Scargle
-periodogram period was below 15 days, the LS power exceeded 0.08, and
-for which no equal-brightness or greater companions were within the aperture. 
-We also required that at most one companion with brightness exceeding 
-one tenth of the target star could reside in the aperture.
-These
-selection criteria are entirely heuristic, and yielded
-\nautorotdenominator stars for which rotation periods might have been
-detected.  
-By exploring the different possible selection criteria, we found that
-the ones we adopted provided a
-reasonable balance between excluding stars on the slow sequence, {\it
-vs.} reporting spurious rotation periods in the ``void'' below the
-slow sequence, where the dominant signal was in fact from a
-neighboring star.
-Splitting the resulting sample into the ``core'' and the ``halo'', they
-suggest that the halo is real.
-
+brighter than $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from
+the source list have $Rp<16$.  The difference is comprised of 34 stars
+unique to \citet{meingast_2021} which were not available at the time
+of the CDIPS reductions and stars falling on the TESS chip gaps.  The
+$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
+roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
+the distance of NGC\,2516.
+
+We removed systematic trends shared between all light curves on each
+CCD in each sector individually, and stitched together the resulting
+light curves before searching for the periodicity.  Details regarding
+our detrending approach are discussed in
+Appendix~\ref{app:detrending}.  After applying the initial detrending
+step aimed at removing systematic trends, we proceeding with a few
+small cleaning steps aimed at improving the purity of the rotation
+period measurements: we masked 0.7 days at the beginning and end of
+each spacecraft orbit, and ran a sliding standard-deviation rejection
+window over the light curve, which removed any outlying points within
+$\pm3\times$MAD of the median in each window.
+
+We then measured the rotation period from the resulting light curve.
+We used the CDIPS aperture radius that, based on theoretical
+expectations, was expected to give the optimal balance between light
+from the target and background-light \citep{Sullivan_2015}.  This
+typically resulted in an aperture radius of either 1 or 1.5 pixels.
+To measure the periods, we used the periodogram implementations in
+\texttt{astrobase}, in particular the Stellingwerf phase-dispersion
+minimization periodogram, along with the more traditional Lomb-Scargle
+\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
+We recorded the top five periodogram peaks from each method, and their
+corresponding powers.  Finally, as a check on crowding, we also
+recorded the number of stars within the aperture of equal brightness
+to the target stars, and of brightness with 1.25 and 2.5 TESS
+magnitudes of the target star.
+
+Figure~\ref{fig:rot} shows the resulting rotation periods, following
+three different sets of cleaning criteria.  The first and second sets
+of criteria were entirely automated, and yielded subsets we denote
+$\mathcal{A}$ and $\mathcal{B}$.  For both subsets, we considered
+light curves for which the peak Lomb-Scargle periodogram period was
+below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
+for which no equal-brightness or greater companions were within the
+aperture.  We also required that the target star be the brightest star
+in the aperture, and that at most one companion with brightness
+exceeding one tenth of the target star's brightness be present in the
+aperture according to the Gaia DR2 source catalog.  For subset
+$\mathcal{A}$, this yielded \nautorotnumerator light curves, out of
+\nautorotdenominator stars for which rotation periods might plausibly
+have been detected.  For subset $\mathcal{B}$, we additionally
+required that the Lomb-Scargle and Stellingwerf phase-dispersion
+methods yielded periods identical to within 10\%.  This yielded
+\nautorotnumeratormatching light curves.
+Subset $\mathcal{A}$ is more complete than 
+subset $\mathcal{B}$, at the expense of greater contamination.
+
+{\bf TODO CORRECTLY}
 To put the clusters on the same color scale, we needed to adopt
 reddening corrections.
+For \cn, we used the same average reddening as described in
+Section~\ref{subsec:hr}.
 % FIXME: where did these reference tables come from???
-We assumed {X} for the Pleiades, and {Y} for Praesepe.
-For NGC\,2516, we used the \texttt{astrobase.services.dust} module to
-query the 2MASS DUST
-service\footnote{\url{http://irsa.ipac.caltech.edu/applications/DUST}}
-and retrieve the extinction parameters for the positions of each
-NGC\,2516 member \citep{bhatti_astrobase_2018}.  The mean and standard
-deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
-and \citet{schlafly_measuring_2011} maps were consistent, so we
-adopted the average from \citet{schlegel_maps_1998}:
-$E(B-V)=0.206\pm0.039$.  \citet{bonifacio_search_2000}
-\citet{schlegel_maps_1998} maps overestimate the reddening values when
-the color excess exceeds about 0.10 mag. Applying the correction
-proposed by \citet{bonifacio_search_2000} leads to our adopted
-$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we adopted
-the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
-E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
-which is used in the plots.
-
+For the Pleiades, and for Praesepe, we used {\bf XXX, YYY}.
+
+{\bf TODO CORRECTLY}
 The resulting gyrochronology age we find for the core is XXX.

### 2021-03-06 cluster_paper/collect.sh

@@ -6,7 +6,7 @@ cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f2b.pdf
 cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2c.pdf
 
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear.pdf f3a.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_log.pdf f3b.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear_periodogram_match.pdf f3b.pdf
 
 #cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
 cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_galactic.pdf f4.pdf

### 2021-03-06 earhart/plotting.py

@@ -1091,7 +1091,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     # make plot
     plt.close('all')
 
-    f, ax = plt.subplots(figsize=(4,3))
+    f, ax = plt.subplots(figsize=(4,5))
 
     classes = ['pleiades', 'praesepe', f'{runid}']
     colors = ['gray', 'gray', 'k']

### 2021-03-08 cluster_paper/ms.tex
rotation references: use Curtis+20 table 5

@@
     from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
-    $Bp-Rp\approx2.2$).
+    $Bp-Rp\approx$2.2).
     \label{fig:hr}
@@
 0.45\,$M_\odot$.  However below this mass, the data and models diverge
-at colors redder than Bp-Rp$\approx2.2$, in the sense that the model
+at colors redder than $Bp-Rp\approx2.2$, in the sense that the model
 isochrones have lower luminosities and bluer colors than the
@@
 This implies an absolute age slightly older than the Pleiades
-\citep[{\it cf.}][]{mermilliod_comparative_1981}.
+\citep[cf.][]{mermilliod_comparative_1981}.
 
@@
 \subsubsection{Cluster Star Sample}
+\label{subsubsec:cluster}
 
@@
     Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
-    undergo successive degrees of automated cleaning, described in the
-    text.
-    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a} and
-    Praesepe \citep[650\,Myr][]{douglas_poking_2017} are shown for
+    undergo successive degrees of automated cleaning, described in Section~\ref{subsubsec:cluster}.
+    The Pleiades \citep[120\,Myr;][]{rebull_rotation_2016a} and
+    Praesepe \citep[650\,Myr;][]{douglas_poking_2017} are shown for
     reference.
     Stars in the core and the halo of NGC\,2516 appear
-    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
-    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
+    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
+    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$,
     in which some NGC\,2516 members appear elevated with respect to
@@
 for which no equal-brightness or greater companions were within the
-aperture.  We also required that the target star be the brightest star
-in the aperture, and that at most one companion with brightness
-exceeding one tenth of the target star's brightness be present in the
-aperture according to the Gaia DR2 source catalog.  For subset
-$\mathcal{A}$, this yielded \nautorotnumerator light curves, out of
-\nautorotdenominator stars for which rotation periods might plausibly
-have been detected.  For subset $\mathcal{B}$, we additionally
-required that the Lomb-Scargle and Stellingwerf phase-dispersion
-methods yielded periods identical to within 10\%.  This yielded
-\nautorotnumeratormatching light curves.
-Subset $\mathcal{A}$ is more complete than 
-subset $\mathcal{B}$, at the expense of greater contamination.
-
-{\bf TODO CORRECTLY}
-To put the clusters on the same color scale, we needed to adopt
-reddening corrections.
-For \cn, we used the same average reddening as described in
-Section~\ref{subsec:hr}.
-% FIXME: where did these reference tables come from???
-For the Pleiades, and for Praesepe, we used {\bf XXX, YYY}.
-
-{\bf TODO CORRECTLY}
+aperture.  Beyond requiring that the target star be the brightest star
+in the aperture, we also required that at most one companion with
+brightness exceeding one tenth of the target star's brightness be
+present in the aperture according to the Gaia DR2 source catalog.  For
+set $\mathcal{A}$, this yielded \nautorotnumerator light curves,
+out of \nautorotdenominator stars for which rotation periods might
+plausibly have been detected.  For set $\mathcal{B}$, we
+then additionally required that the Lomb-Scargle and Stellingwerf
+phase-dispersion methods yielded identical periods to within 10\%.
+This yielded \nautorotnumeratormatching light curves.
+
+The comparison to the Pleiades and Praesepe is made using the rotation
+periods and dereddened colors from Table~5 of \citet{curtis_rup147_2020}, which
+were based data from \citet{rebull_rotation_2016a} and
+\citet{douglas_k2_2019} respectively.
+
+A few conclusions can be drawn by comparing
+the \cn\ rotation periods with those of
+the Pleiades and Praesepe.
+The first is that in both stages of cleaning,
+the core and halo of \cn\ both show
+gyrochronological ``slow sequences''.
+These sequences appear to be
+coeval with each other, and at least to first order
+appear to overlap with the Pleiades sequence.
+
+Second, subset
+$\mathcal{A}$ is more complete than subset $\mathcal{B}$.
+This completeness comes at the
+expense of greater contamination.
+The slow sequence
+has a factor of $\sim$twice as many stars in subset
+$\mathcal{A}$ than $\mathcal{B}$.
+However stars above the slow sequence, which could
+well be field contaminants, for the most part disappear in
+subset $\mathcal{B}$.
+The ``void'' underneath the slow sequence
+is also more pronounced in the cleaner
+subset,
+though it seems to not be as empty as in
+the Pleiades.
+
+Finally, comparing the slow sequences of the Pleaides
+and \cn, they appear to overlap from roughly
+$0.5<(Bp-Rp)_0<1.2$.
+For stars that are redder,
+the dispersion in rotation periods increases.
+From
+$1.2<(Bp-Rp)_0<1.7$, subset
+$\mathcal{B}$ of \cn\ has stars that overlap with 
+the Pleiades.
+However, it also has about ten stars at longer rotation periods---extending
+up to $\approx$11 days, rather than the
+$\approx$8.5day limit for the Pleiades.
+This is consistent with \cn\ being slightly
+older than the Pleiades.
+
+{\bf We explore this quantitatively by fitting
+interpolated models}[?].
 The resulting gyrochronology age we find for the core is XXX.

### 2021-03-08 earhart/plotting.py

@@ -1109,7 +1109,16 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     ):
 
         if f'{runid}' not in _cls:
-            df = pd.read_csv(os.path.join(rotdir, f'curtis19_{_cls}.csv'))
+            t = Table.read(
+                os.path.join(rotdir, 'Curtis_2020_apjabbf58t5_mrt.txt'),
+                format='cds'
+            )
+            if _cls == 'pleiades':
+                df = t[t['Cluster'] == 'Pleiades'].to_pandas()
+            elif _cls == 'praesepe':
+                df = t[t['Cluster'] == 'Praesepe'].to_pandas()
+            else:
+                raise NotImplementedError
 
         else:
             df = get_autorotation_dataframe(
@@ -1121,18 +1130,21 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
             print(42*'-')
 
         if f'{runid}' not in _cls:
-            xval = get_interp_BpmRp_from_Teff(df['teff'])
-            df['BpmRp_interp'] = xval
-            df.to_csv(
-                os.path.join(rotdir, f'curtis19_{_cls}_BpmRpinterp.csv'),
-                index=False
-            )
+            xval = df['(BP-RP)0']
+
+            # NOTE: deprecated; based on the webplotdigitzing approach
+            # xval = get_interp_BpmRp_from_Teff(df['teff'])
+            # df['BpmRp_interp'] = xval
+            # df.to_csv(
+            #     os.path.join(rotdir, f'curtis19_{_cls}_BpmRpinterp.csv'),
+            #     index=False
+            # )
         else:
             xval = (
                 df['phot_bp_mean_mag'] - df['phot_rp_mean_mag'] - E_BpmRp
             )
 
-        ykey = 'prot' if f'{runid}' not in _cls else 'period'
+        ykey = 'Prot' if f'{runid}' not in _cls else 'period'
 
         if core_halo and f'{runid}' in _cls:
             sel = (df.subcluster == 'core')

### 2021-03-09 cluster_paper/ms.tex
improved. manuscript.
* sanity checking my coordinate systems
* adding sick vectors
* add v_tangential check. yeah, it does matter.

@@
   %The Halo of NGC\,2516 Extends To At Least 300\,pc
-  An Open Cluster With a 300\,pc Halo
+  %An Open Cluster With a 300\,pc Halo
+  An Open Cluster Spread Across 500\,pc
   %NGC\,2516  300\,pc
@@
   diffuse stellar populations surrounding nearby open clusters. 
-  The stars in these diffuse halos could be the tidal escapees of the
+  The stars in these diffuse halos could be tidal escapees from the
   cluster cores, or they could be remnants of an initially dispersed
@@
   They could also be field stars---the false positives of overly
-  aggressive clustering algorithms.
-  In this study, we focus on a halo with a length of 300\,pc reported
-  around the $\approx$150\,Myr open cluster NGC\,2516.
+  sensitive clustering algorithms.
+  In this study, we focus on a halo with a tip-to-tip length of
+  500\,pc reported around the $\approx$150\,Myr open cluster
+  NGC\,2516.
   The classical tidal radius of the cluster is $\lesssim$10\,pc.
@@
   cluster center show rotation periods consistent with a
-  gyrochronological age of 120\,Myr.
+  gyrochronological age of 150\,Myr.
   A comparison sample of field stars shows no such trend.
@@
   similar stellar populations.
-  This work expands the set of confirmed NGC\,2516 members by a factor
-  of $\approx$2, and quantifies the degree to which Gaia-based
+  This work quantifies the degree to which Gaia-based
   analyses can identify dispersed stellar populations against a
-  background of field star contaminants.
+  background of field star contaminants, and
+  expands the set of confirmed NGC\,2516 members by a factor
+  of $\approx$2.
   We highlight the implications for spectroscopic survey targeting,
@@
   open clusters (1160),
-  %stellar dynamics (1596),
+  stellar dynamics (1596),
   stellar rotation (1629)
@@
 
-Star clusters form from the densest filaments of 
-hierarchically structured molecular clouds
-\citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
-including protostellar outflows, photoionization, radiation pressure,
-and supernova shocks disperse the gas out of the cloud
-\citep{krumholz_star_2019}.  Since only $\sim 1\%$ of the cloud
-mass is converted into stars, gas dispersal enables the
-majority ($\sim$90\%) of stars in the cluster to escape the cluster's
-gravitational well \citep{lada_embedded_2003}.
+Star clusters form when dense filaments in hierarchically structured
+molecular clouds collide and collapse \citep{shu_star_1987}.  Over the
+first 10\,Myr, feedback effects including protostellar outflows,
+photoionization, radiation pressure, and supernova shocks disperse the
+gas out of the cloud \citep{krumholz_star_2019}.  Since only $\sim
+1\%$ of the cloud mass is converted into stars, gas dispersal enables
+the majority ($\sim$90\%) of stars in the cluster to escape the
+cluster's gravitational well \citep{lada_embedded_2003}.
 
@@
 agnostic.  It suffers however from the suggestion of spherical
-symmetry.  The halo of NGC\,2516 might be best-described as ``tails'',
-but the primary goal of this study is
-not to verify whether these tails are tidal.
+symmetry.  The halo of NGC\,2516 might be better described as ``tails'',
+but the primary goal of this study is not to verify whether these
+tails are tidal.
 
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f11.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf NGC\,2516 in the galactic context.} 
+  Points denote halo members with $\pi/\sigma_\pi>20$.
+  In our coordinate system, the Sun is at
+  $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
+  The gray line traces the past and future cluster orbit.  In
+  different projections, the line spans different time intervals: the
+  longest time window is visible in the $(Z,X)$ projection.
+  The large blue arrows denote the median cluster velocity after
+  subtracting the local standard of rest: 
+  $\{v_{\rm X}, v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.
+  Their sizes are internally proportionate.
+  By comparison, $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
+  \label{fig:XYZ}
+	}
+\end{figure*}
+
 We selected candidate \cn\ members based on those reported in the
-literature.  While a number of pre-Gaia 
-studies have been performed, the
-purity and accuracy of the Gaia-derived results are the current state
-of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore adopted what we viewed as the most
-interesting clustering samples to compare: those of
-\citet{cantatgaudin_gaia_2018}, \citet{kounkel_untangling_2019} and
-\citet{meingast_2021}.  We refer to these studies in the following as
+literature.  While a number of pre-Gaia studies have been performed,
+the purity and accuracy of the Gaia-derived results are the current
+state of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.
+We therefore adopted what we viewed as the most interesting clustering
+samples to compare: those of \citet{cantatgaudin_gaia_2018},
+\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
+to these studies in the following as
 \citetalias{cantatgaudin_gaia_2018},
@@
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}.}.  While we considered performing
-our own clustering analysis using the Gaia data, such an effort would
-in effect only replicate the work of these investigators.  We opt to
+  made by \citet{meingast_2021}.}.  While we considered performing our
+own clustering analysis using the Gaia data, such an effort would in
+effect only replicate the work of these investigators.  We opt to
 instead use their studies as starting points.
@@
 $\approx$2$^\circ$, not $\approx20^\circ$.
-
-The discrepancy between our canonical understanding of open clusters
-and the possibility of a diffuse population existing around \cn\
-prompted our initial interest in this analysis.  What is the true
-structure of \cn?  Are the core and halo truly coeval?
 
@@
 the omission of 1.1\% of the stars drawn from the volume noted above.
+
+The discrepancy between our canonical understanding of open clusters
+and the possibility of a diffuse population existing around \cn\
+prompted our initial interest in this analysis.  What is the true
+structure of \cn?  Are the core and halo truly coeval?
+
+% FIXME FIXME FIXME
+{\bf
+To at least visualize the thing better...
+}
+Figure~\ref{fig:XYZ}.
 
@@
 
+{\bf Curtis TODO: set $\mathcal{C}$ was created through a manual
+curation process.}
+
 The comparison to the Pleiades and Praesepe is made using the rotation
@@
 
-A few conclusions can be drawn by comparing
-the \cn\ rotation periods with those of
-the Pleiades and Praesepe.
-The first is that in both stages of cleaning,
-the core and halo of \cn\ both show
-gyrochronological ``slow sequences''.
-These sequences appear to be
-coeval with each other, and at least to first order
-appear to overlap with the Pleiades sequence.
-
-Second, subset
-$\mathcal{A}$ is more complete than subset $\mathcal{B}$.
-This completeness comes at the
-expense of greater contamination.
-The slow sequence
-has a factor of $\sim$twice as many stars in subset
-$\mathcal{A}$ than $\mathcal{B}$.
-However stars above the slow sequence, which could
-well be field contaminants, for the most part disappear in
-subset $\mathcal{B}$.
-The ``void'' underneath the slow sequence
-is also more pronounced in the cleaner
-subset,
-though it seems to not be as empty as in
-the Pleiades.
-
-Finally, comparing the slow sequences of the Pleaides
-and \cn, they appear to overlap from roughly
-$0.5<(Bp-Rp)_0<1.2$.
-For stars that are redder,
-the dispersion in rotation periods increases.
-From
-$1.2<(Bp-Rp)_0<1.7$, subset
-$\mathcal{B}$ of \cn\ has stars that overlap with 
-the Pleiades.
-However, it also has about ten stars at longer rotation periods---extending
-up to $\approx$11 days, rather than the
-$\approx$8.5day limit for the Pleiades.
-This is consistent with \cn\ being slightly
-older than the Pleiades.
+A few conclusions can be drawn by comparing the \cn\ rotation periods
+with those of the Pleiades and Praesepe.  The first is that in both
+stages of cleaning, the core and halo of \cn\ both show
+gyrochronological ``slow sequences''.  These sequences appear to be
+coeval with each other, and at least to first order appear to overlap
+with the Pleiades sequence.
+
+Second, subset $\mathcal{A}$ is more complete than subset
+$\mathcal{B}$.  This completeness comes at the expense of greater
+contamination.  The slow sequence has a factor of $\sim$twice as many
+stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
+the slow sequence, which could well be field contaminants, for the
+most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
+the slow sequence is also more pronounced in the cleaner subset,
+though it seems to not be as empty as in the Pleiades.
+
+Finally, comparing the slow sequences of the Pleaides and \cn, they
+appear to overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  For stars that
+are redder, the dispersion in rotation periods increases.  From
+$1.2<(Bp-Rp)_0<1.7$, subset $\mathcal{B}$ of \cn\ has stars that
+overlap with the Pleiades.  However, it also has about ten stars at
+longer rotation periods---extending up to $\approx$11 days, rather
+than the $\approx$8.5day limit for the Pleiades.  This is consistent
+with \cn\ being slightly older than the Pleiades.
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f4.pdf}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f12a.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f12b.pdf}
+		}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
-    velocity space, cross-matched against the rotators.}
-    The plotted stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the
-    crowding restrictions described in Section~\ref{subsec:tess}: they
-    should have been sufficiently bright and non-crowded to enable
-    rotation period detection.
-		\label{fig:gaia6d_x_rotn}
-	}
-\end{figure*}
+  \caption{ {\bf Projection effects for co-moving populations
+  		across the sky.}
+  	The map is colored by the velocity difference
+  	a star co-moving with \cn\ would exhibit across
+  	the equatorial sphere.
+  	Actual positions of candidate \cn\ members are shown with points; the star denotes the cluster center.
+	\label{fig:vproj}
+	}
+\end{figure}
 
@@
   \caption{ {\bf Kinematic members of NGC\,2516, cross-matched against
-  rotational members.}
-    The projections are in galactic position coordinates and 2d-tangential velocity
-    space.
+  rotational members.} The projections are in galactic position
+  coordinates and 2d-tangential velocity space.
   \label{fig:physical_x_rotn}
@@
   \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
-  versus 3d separation from the core and 2d velocity difference.} Bins have width
-  25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
-  {\bf Uncertainties in the fraction are taken as the quadrature sum of
-  the rel unc of the numerator and denominator (FIXME)}.
-  Bonafide members appear to exist hundreds of parsecs from the core, and
-  up to $\approx$10\,km\,s$^{-1}$ in tangential velocity separation.
+  versus 3d separation from the core and 2d velocity difference.} Bins
+  have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
+  {\bf Uncertainties in the fraction are taken as the quadrature sum
+  of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
+  members appear to exist hundreds of parsecs from the core, and up to
+  $\approx$10\kms in tangential velocity separation.
   \label{fig:hist_physical_x_rotn}
@@
 
-%\begin{figure*}[t]
-%	\begin{center}
-%		\leavevmode
-%		\subfloat{
-%			\includegraphics[width=0.75\textwidth]{f8a.pdf}
-%		}
-%	
-%		\subfloat{
-%			\includegraphics[width=0.75\textwidth]{f8b.pdf}
-%		}
-%	\end{center}
-%	\vspace{-0.7cm}
-%	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
-%    velocity space, cross-matched against the rotators, in physical
-%    units.}
-%    {\it Top:} projections in position and 2d-tangential velocity
-%    space. {\it Bottom:} radially binned versus 3d position and
-%    2d velocity difference.
-%    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
-%    velocity.
-%		\label{fig:physical_x_rotn}
-%	}
-%\end{figure*}
-
 How far away from the core, in position and velocity space, does the
-halo actually extend?  We can explore this by crossmatching the
-rotator sample against \nkinematic Gaia DR2 kinematic candidate
-members.
-
-Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
-in interpreting this plot is that we show only stars with
-$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
-detrended light curves.  In other words---the base sample is the stars
-for which we could have plausibly detected a rotation period.  Faint
-stars with $Rp>16$ were beyond our light curve selection limit, and
-crowded stars ({\it e.g.,} near saturated stars in the cluster center)
-do not appear.  We selected the color limit above based on
-Figure~\ref{fig:rot}, since the density of period detections seems to
-decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types
-later than $\approx$K4V.
+halo extend?  We can explore this by crossmatching the stars with
+detected rotation periods against our original target list of
+\nkinematic Gaia DR2 kinematic candidate members.  For a fair
+comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
+which our TESS pipeline successfully made detrended light curves.  In
+other words---the base sample is the stars for which we could have
+plausibly detected a rotation period.  Faint stars with $Rp>16$ were
+beyond our light curve selection limit, and crowded stars ({\it e.g.,}
+near saturated stars in the cluster center) are similarly omitted.  We
+selected the color limit above based on Figure~\ref{fig:rot}, since
+the density of period detections seems to decrease once $(Bp-Rp)_0
+\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
 Our expectation is that completion effects become important for
 fainter stars.
-At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
-for which our reductions of the TESS images yield a precision of
-$\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
-{\bf 
-The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
-REBULL 2020, FIGURE 10).
-}
-
+%
+% At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
+% for which our reductions of the TESS images yield a precision of
+% $\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
+% {\bf 
+% The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
+% REBULL 2020, FIGURE 10).
+% }
+%
 %G=18 is like Abs G 9.8, Bp-Rp not corrected of 2.7. 
@@
 % So, G~=15... meaning Rp~=14.3. 
-
+%
 % IN POSITION
@@
 % Rotators > 25pc : 0.66172 +/- 0.00023
-% ... where numerator and denom are 223/337
-
-Qualitatively, the fraction of kinematic members with TESS-detected rotation 
-periods seems like it may increase with increasing distance from the halo core---both in
-position and in velocity space.
-To quantify this, we computed the relelvant distances in physical coordinates.
-This entailed first transforming from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) which we did 
-assuming the AstroPy v4.0 coordinate standard (CITE).
-We then converted the on-sky proper motions to physical units by dividing by the distance of each star.
-We then computed the SEPARATIONS of all of these positions and
-2d-velocities from the ``median core member'', which was selected by
-taking \citetalias{cantatgaudin_gaia_2018} members with membership
-probability exceeding 70\%.
-Again, we only considered stars for which we could have plausibly
-detected a rotation period --- the same color and observability limits
-discussed above were also applied.
-
-The results are shown in Figure~\ref{fig:physical_x_rotn}.
-The halo extends to separations of about 200\,pc in physical space
-from the cluster core.
-This corresponds to a length of $\approx$300-350\,pc, depending on
-which members of the halo are chosen as the ``tips'' on either
-end.
+%
+
+The results are shown using physical coordinates in
+Figure~\ref{fig:physical_x_rotn}.  Appendix~\ref{app:gaia6d_x_rotn}
+includes a supplemental visualization in the non-transformed space of
+Gaia measurements.  To compute the physical positions, we transformed
+from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) assuming the
+\texttt{astropy v4.0} coordinate standard \citep{astropy_2018}.  We
+then converted the on-sky proper motions to physical units by dividing
+by the distance of each star, taken as the inverse parallax.  The
+inverse parallax approximation does not introduce significant biases
+for this sample: the mean $\pi/\sigma_\pi$ and its standard deviation
+are 122 and 40 respectively; the minimum is 19.5.
+Qualitatively, the fraction of kinematic members with TESS-detected
+rotation periods seems like it may decrease with increasing distance
+from the halo core---both in position and in velocity space.
+
+To quantify this, we computed the separation of the positions and 2D
+velocities from the median core member, selected by taking the median
+across \citetalias{cantatgaudin_gaia_2018} members with membership
+probabilities exceeding 70\%.  The result is shown in
+Figure~\ref{fig:hist_physical_x_rotn}.  The halo extends to
+separations of about 200\,pc in physical space from the cluster core.
+This corresponds to total a length of $\approx$300-350\,pc, depending on
+which members of the halo are chosen as the ``tips'' on either end.
+Broadly speaking, this agrees with the results of
+\citetalias{kounkel_untangling_2019}.
 
 In tangential velocity space, the fraction of stars with rotation
-period detections remains high out to roughly 10\,km/s.
+period detections remains high out to roughly 10\kms.
 \citet{meingast_2021} by comparison required a physically motivated
-cut in tangential velocity space of 1.5 km per sec.  This probably
-improves purity, but clearly doesn't get everthing for completeness.
+cut in tangential velocity space of 1.5\kms.
+
+Of course, an important aspect of the tangential velocity cut needs to
+be understood in light of 
+PROJECTION EFFECTS.
+
+... So uhhhhh yeah I need to compute that map.
+
+While this 
+improves purity, it is somewhat lacking in completeness...
 
@@
 
+\section{Kinematics $\otimes$ Rotation}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Gaia-based components of NGC\,2516 in position and
+  velocity space, cross-matched against the rotators.} The plotted
+  stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
+  restrictions described in Section~\ref{subsec:tess}: they should
+  have been sufficiently bright and non-crowded to enable rotation
+  period detection.  We caution that the appearance of fewer
+  non-rotators being present toward the core is due the layering of
+  the plot: quantitatively, stars toward the cluster center do not all
+  show rotation periods in our analysis (see
+  Figure~\ref{fig:hist_physical_x_rotn}).
+  \label{fig:gaia6d_x_rotn}
+	}
+\end{figure*}
+
+Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
+data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
+in this diagram correspond to ``Set $\mathcal{A}$'' as described in
+Section~\ref{subsec:tess}.  Many stars in the outer region of the
+cluster show rotation periods consistent with a young age.  One issue
+with the visualization as displayed however is that it hides the
+amount of non-rotators toward the cluster center: we did not detect
+rotation periods for roughly one in four stars at the cluster
+center (see Figures~\ref{fig:physical_x_rotn}
+and~\ref{fig:hist_physical_x_rotn}).
+
 \section{Rotation $\otimes$ RUWE}

### 2021-03-09 cluster_paper/collect.sh

@@ -17,6 +17,10 @@ cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f8
 cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit.pdf f5a.pdf
 cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_lithium_viridis_gaiaeso_galahdr3.pdf f5b.pdf
 
+cp ../results/vtangential_projection/XYZ_with_orbit.pdf f11.pdf
+cp ../results/vtangential_projection/skymap_tangential_velocity_pmdec.pdf f12a.pdf
+cp ../results/vtangential_projection/skymap_tangential_velocity_pmra_cosdec.pdf f12b.pdf
+
 # supplementary
 cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis_linear.pdf f6.pdf
 

### 2021-03-09 earhart/helpers.py

@@ -10,6 +10,7 @@ get_gaia_basedata
     _get_fullfaint_edr3_dataframes
     _get_denis_fullfaint_edr3_dataframes
     _get_extinction_dataframes
+    _get_median_ngc2516_core_params
 get_denis_xmatch
 """
 import os, collections, pickle
@@ -886,3 +887,28 @@ def get_autorotation_dataframe(runid='NGC_2516', verbose=1, returnbase=0,
     else:
         return df[sel], df[ref_sel]
 
+
+def _get_median_ngc2516_core_params(core_df, basedata, CUTOFF_PROB=0.7):
+    """
+    To get median parameters of the NGC 2516 cluster, select the high
+    probability CG18 members, and take a median.
+    """
+
+    core_path = os.path.join(DATADIR, 'gaia', 'CantatGaudin2018_vizier_only_NGC2516.fits')
+    t_df = Table(fits.open(core_path)[1].data).to_pandas()
+    if basedata == 'fullfaint':
+        assert len(t_df) == len(core_df)
+
+    sel = (t_df.PMemb > CUTOFF_PROB)
+    s_t_df = t_df[sel]
+    assert np.all(s_t_df.Source.isin(core_df.source_id))
+    sel = core_df.source_id.isin(s_t_df.Source)
+    s_core_df = core_df[sel]
+    rvkey = (
+        'radial_velocity' if 'edr3' not in basedata else 'dr2_radial_velocity'
+    )
+    getcols = ['ra', 'dec', 'parallax', 'pmra', 'pmdec', rvkey]
+    med_df = pd.DataFrame(s_core_df[getcols].median()).T
+    std_df = pd.DataFrame(s_core_df[getcols].std()).T
+
+    return med_df, std_df

### 2021-03-09 earhart/physicalpositions.py

@@ -1,5 +1,10 @@
 """
 Tools for working in XYZ, UVW, and Δμ_δ, Δμ_α' [km/s] coordinates.
+
+Contents:
+    append_physicalpositions: (α,δ,π)->(X,Y,Z).
+    given_gaia_df_get_icrs_arr: DataFrame -> SkyCoord conversion
+    calc_dist: between two 3d cartesian points.
 """
 
 import numpy as np
@@ -22,6 +27,11 @@ def append_physicalpositions(df, median_df):
     the XYZ coordinates, the "delta_pmra_km_s" and "delta_pmdec_km_s"
     coordinates, the "delta_r_pc" (3d separation) and "delta_mu_km_s" (velocity
     separation) coordinates, and append them to the dataframe.
+
+    Returns:
+        DataFrame with ['x_pc', 'y_pc', 'z_pc', 'delta_r_pc',
+        'delta_mu_km_s', 'delta_pmra_km_s' ,'delta_pmdec_km_s']
+        appended.
     """
 
     # get XYZ in galactocentric
@@ -49,10 +59,16 @@ def append_physicalpositions(df, median_df):
         df[new_key] = new_val
 
         if c in ['pmra', 'pmdec']:
+            #
+            # naive approach: just take the different to the median
+            # pmra/pmdec of the cluster. (ignoring the spatial
+            # projection effect).
+            #
             # recall:
             # θ_arcsec = r_AU / d_pc
             # mu_arcsec/yr = v_AU/yr / d_pc
             # v_AU/yr = mu_arcsec/yr * d_pc
+            #
             d_pc = (nparr(1/(df.parallax*1e-3))*u.pc).value
             c_AU_per_yr = ((( nparr(df[c]) - nparr(median_df[c]) )*1e-3) * d_pc)*(u.AU/u.yr)
             c_km_per_sec = c_AU_per_yr.to(u.km/u.second)
@@ -86,13 +102,18 @@ def given_gaia_df_get_icrs_arr(df, zero_rv=0):
         )
 
     else:
+        rvkey = 'dr2_radial_velocity'
+        if rvkey in df:
+            pass
+        else:
+            rvkey = 'radial_velocity'
         return coord.SkyCoord(
             ra=nparr(df.ra)*u.deg,
             dec=nparr(df.dec)*u.deg,
             distance=nparr(1/(df.parallax*1e-3))*u.pc,
[…]

### 2021-03-09 earhart/plotting.py

@@ -25,6 +25,7 @@ Contents:
         plot_bisector_span_vs_RV
         plot_backintegration_ngc2516
         plot_venn
+        plot_vtangential_projection
 """
 import os, corner, pickle
 from glob import glob
@@ -33,6 +34,7 @@ import numpy as np, matplotlib.pyplot as plt, pandas as pd, pymc3 as pm
 from numpy import array as nparr
 
 import matplotlib as mpl
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 from astropy import units as u, constants as const
 from astropy.coordinates import SkyCoord
@@ -55,7 +57,8 @@ from cdips.utils.mamajek import get_interp_BpmRp_from_Teff
 
 from earhart.paths import DATADIR, RESULTSDIR
 from earhart.helpers import (
-    get_gaia_basedata, get_autorotation_dataframe
+    get_gaia_basedata, get_autorotation_dataframe,
+    _get_median_ngc2516_core_params
 )
 from earhart.physicalpositions import (
     given_gaia_df_get_icrs_arr, calc_dist
@@ -2259,11 +2262,11 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
     params = [xkey, ykey, 'parallax', 'pmra', 'pmdec', rvkey]
     # whether to limit axis by 5/95th percetile
     qlimd = {
-        xkey: 0, ykey: 0, 'parallax': 0, 'pmra': 1, 'pmdec': 1, rvkey: 1
+        xkey: 0, ykey: 0, 'parallax': 0, 'pmra': 0, 'pmdec': 0, rvkey: 0
     }
     # whether to limit axis by 99th percentile
     nnlimd = {
-        xkey: 1, ykey: 1, 'parallax': 1, 'pmra': 0, 'pmdec': 0, rvkey: 0
+        xkey:0, ykey:0, 'parallax':0, 'pmra':0, 'pmdec':0, rvkey:0
     }
     ldict = {
         xkey: xl, ykey: yl,
@@ -2298,11 +2301,11 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
             yv = params[i+1]
             print(i,j,xv,yv)
 
-            axs[i,j].scatter(
-                nbhd_df[sel_color(nbhd_df)][xv], nbhd_df[sel_color(nbhd_df)][yv],
-                c='gray', alpha=0.9, zorder=2, s=5, rasterized=True,
[…]

### 2021-03-09 cluster_paper/ms.tex
propagate better vtang subtraction

@@
   %An Open Cluster With a 300\,pc Halo
-  An Open Cluster Spread Across 500\,pc
+  An Open Cluster Spread Across 400\,pc
   %NGC\,2516  300\,pc
@@
   In this study, we focus on a halo with a tip-to-tip length of
-  500\,pc reported around the $\approx$150\,Myr open cluster
+  $\approx$400\,pc reported around the $\approx$150\,Myr open cluster
   NGC\,2516.
@@
 
-\section{A 300\,pc Halo around a Core?}
+\section{A Dispersed Halo and a Core?}
 \label{sec:gaia}
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf NGC\,2516 in the galactic context.} 
-  Points denote halo members with $\pi/\sigma_\pi>20$.
-  In our coordinate system, the Sun is at
-  $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
-  The gray line traces the past and future cluster orbit.  In
-  different projections, the line spans different time intervals: the
-  longest time window is visible in the $(Z,X)$ projection.
-  The large blue arrows denote the median cluster velocity after
-  subtracting the local standard of rest: 
-  $\{v_{\rm X}, v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.
-  Their sizes are internally proportionate.
-  By comparison, $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
+  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
+  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
+  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
+  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
+  cluster orbit.  In different projections, the line spans different
+  time intervals: the longest time window is visible in the $(Z,X)$
+  projection.  The large blue arrows denote the median cluster
+  velocity after subtracting the local standard of rest: $\{v_{\rm X},
+  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.  The sizes of
+  the blue arrows are internally proportionate.  By comparison,
+  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
   \label{fig:XYZ}
@@
 Figure~\ref{fig:gaia6d} shows the cluster members reported by each
-study in the space of positions, proper motions, and radial velocity
-when available.  The \citetalias{cantatgaudin_gaia_2018} members are
-all within a few degrees of the cluster center, while the
+study in the space of observed positions, proper motions, and radial
+velocity when available.  The \citetalias{cantatgaudin_gaia_2018}
+members are all within a few degrees of the cluster center, while the
 \citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
-members (the ``halo'') span tens of degrees. 
-Inverting the parallax shows that members have been reported from
-line-of-sight distances ranging from $300$ to $600\,$pc.
-Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
-$\mu_\delta$ projection were reported almost entirely by
-\citet{meingast_2021}.
+members (the ``halo'') span tens of degrees.  Inverting the parallax
+shows that members have been reported from line-of-sight distances
+ranging from $300$ to $600\,$pc.  Stars in the upper right quadrant of
+the $\mu_{\alpha'}$ {\it vs.} $\mu_\delta$ projection were reported
+almost entirely by \citet{meingast_2021}.
 
@@
 distribution of halo members, we drew these stars randomly from the
-following ``cube'' in right ascension, declination, and parallax:
+following intervals of right ascension, declination, and parallax:
 \begin{align}
@@
 and the possibility of a diffuse population existing around \cn\
-prompted our initial interest in this analysis.  What is the true
-structure of \cn?  Are the core and halo truly coeval?
-
-% FIXME FIXME FIXME
-{\bf
-To at least visualize the thing better...
-}
-Figure~\ref{fig:XYZ}.
+prompted our interest in understanding the reality and origin of the
+diffuse population.  Figure~\ref{fig:XYZ} helps in visualizing the
+proposed structure and kinematics of the halo.  We computed the
+physical positions by transforming from $(\alpha, \delta, \pi)$ to
+galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
+standard \citep{astropy_2018}.  The direction of galactic rotation is
+$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
+median parameters for core members for which
+\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
+exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
+the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.
+The elongation of the cluster in both the $(X,Y)$ and $(Z,Y)$ planes
+is correlated with the direction of the LSR-corrected median cluster
+velocity.  A similar correlation was noted in Coma Ber by
+\citet{tang_comaber_2019}.  Possible explanations are discussed in
+Section~\ref{subsec:origin}.
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
-		\leavevmode
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f12a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f12b.pdf}
-		}
+			\includegraphics[width=0.5\textwidth]{f8a.pdf}
+			\includegraphics[width=0.5\textwidth]{f8b.pdf}
+		}
+	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf Projection effects for co-moving populations
-  		across the sky.}
-  	The map is colored by the velocity difference
-  	a star co-moving with \cn\ would exhibit across
-  	the equatorial sphere.
-  	Actual positions of candidate \cn\ members are shown with points; the star denotes the cluster center.
-	\label{fig:vproj}
-	}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-			\includegraphics[width=0.49\textwidth]{f8a.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Kinematic members of NGC\,2516, cross-matched against
-  rotational members.} The projections are in galactic position
-  coordinates and 2d-tangential velocity space.
+  \caption{ 
+  {\bf Members of NGC\,2516 exist hundreds of parsecs from the core,
+  and up to $\approx$5\kms\ from the core in tangential velocity.} Stars
+  that meet both kinematic and rotational membership criteria are shown in
+  blue.  Stars for which rotation periods should have been detectable,
+  but were not detected, are shown in orange.  Projections are shown in galactic coordinates and
+  (projection-corrected) 2D-tangential velocity space.  The histogram
+  on the right bins versus 3D separation from the core and
+  2D velocity difference. Bins have width 25\,pc
+  in position, and 1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties
+  in the fraction are taken as the quadrature sum of the rel unc of
+  the numerator and denominator (FIXME)}
   \label{fig:physical_x_rotn}
 	}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-			\includegraphics[width=0.49\textwidth]{f8b.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
-  versus 3d separation from the core and 2d velocity difference.} Bins
-  have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
-  {\bf Uncertainties in the fraction are taken as the quadrature sum
-  of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
-  members appear to exist hundreds of parsecs from the core, and up to
-  $\approx$10\kms in tangential velocity separation.
-  \label{fig:hist_physical_x_rotn}
-	}
-\end{figure}
+\end{figure*}
+
+% \begin{figure}[t]
+% 	\begin{center}
+% 			\includegraphics[width=0.49\textwidth]{f8b.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+%   \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
+%   versus 3d separation from the core and projection-corrected 2d velocity difference.} Bins
+%   have width 25\,pc in position, and 1.0\,km\,s$^{-1}$ in velocity.
+%   {\bf Uncertainties in the fraction are taken as the quadrature sum
+%   of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
+%   members appear to exist hundreds of parsecs from the core, and up to
+%   $\approx$5\kms in tangential velocity separation.
+%   \label{fig:hist_physical_x_rotn}
+% 	}
+% \end{figure}
 
@@
 fainter stars.
-%
-% At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
-% for which our reductions of the TESS images yield a precision of
-% $\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
-% {\bf 
-% The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
-% REBULL 2020, FIGURE 10).
-% }
-%
-%G=18 is like Abs G 9.8, Bp-Rp not corrected of 2.7. 
-% (Bp-Rp )_0 of 1.25 is (Bp-Rp) of 1.38... or Abs G of ~= 6.8.
-% or G-Rp ~= 0.7
-% So, G~=15... meaning Rp~=14.3. 
-%
+
 % IN POSITION
@@
 Figure~\ref{fig:physical_x_rotn}.  Appendix~\ref{app:gaia6d_x_rotn}
-includes a supplemental visualization in the non-transformed space of
-Gaia measurements.  To compute the physical positions, we transformed
-from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) assuming the
-\texttt{astropy v4.0} coordinate standard \citep{astropy_2018}.  We
-then converted the on-sky proper motions to physical units by dividing
-by the distance of each star, taken as the inverse parallax.  The
-inverse parallax approximation does not introduce significant biases
-for this sample: the mean $\pi/\sigma_\pi$ and its standard deviation
-are 122 and 40 respectively; the minimum is 19.5.
-Qualitatively, the fraction of kinematic members with TESS-detected
-rotation periods seems like it may decrease with increasing distance
-from the halo core---both in position and in velocity space.
-
-To quantify this, we computed the separation of the positions and 2D
-velocities from the median core member, selected by taking the median
-across \citetalias{cantatgaudin_gaia_2018} members with membership
-probabilities exceeding 70\%.  The result is shown in
-Figure~\ref{fig:hist_physical_x_rotn}.  The halo extends to
-separations of about 200\,pc in physical space from the cluster core.
-This corresponds to total a length of $\approx$300-350\,pc, depending on
-which members of the halo are chosen as the ``tips'' on either end.
-Broadly speaking, this agrees with the results of
+includes a supplemental visualization in the observed space of Gaia
+measurements.  The observed proper motions however suffer from
+projection effects, discussed in Appendix~\ref{app:vproj}.  These
+effects are important in the equatorial coordinate system due to the
+proximity of some \cn\ members to the southern ecliptic pole.  For a
+star at position $(\alpha, \delta)$, we correct for the projection
+effect by comparing the observed proper motion $(\mu_{\alpha'},
+\mu_\delta)$ to what the proper motion at the star's position would
+have been if the star were comoving with the core of \cn.  This yields
+a quantity we denote $\Delta \mu^{*}$, following
+\citetalias{meingast_2021}.  We then convert these proper motion
+differences to physical units by dividing by the distance of each
+star, taken as the inverse parallax.  The inverse parallax
+approximation does not introduce significant biases for this sample,
+since the mean $\pi/\sigma_\pi$ and its standard deviation are 122 and
+40 respectively; the minimum is 19.5.  An inspection of the results
+suggests that the fraction of kinematic members with TESS-detected
+rotation periods may decrease with increasing distance from the halo
+core---both in position and in velocity space.
+
+To quantify this, we compute the separation of the positions and 2D
+velocities from the median \citetalias{cantatgaudin_gaia_2018} core
+member.  The result is shown on the right side of
+Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
+about 200\,pc in physical space from the cluster core.  This
+corresponds to total a length of $\approx$400\,pc, depending on which
+members of the halo are chosen as the ``tips'' on either end.  This
+agrees broadly speaking with the structure of the halo reported by
 \citetalias{kounkel_untangling_2019}.
 
-In tangential velocity space, the fraction of stars with rotation
-period detections remains high out to roughly 10\kms.
-\citet{meingast_2021} by comparison required a physically motivated
-cut in tangential velocity space of 1.5\kms.
-
-Of course, an important aspect of the tangential velocity cut needs to
-be understood in light of 
-PROJECTION EFFECTS.
-
-... So uhhhhh yeah I need to compute that map.
-
-While this 
-improves purity, it is somewhat lacking in completeness...
-
-% [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0   0]
-% [495 116  65  46  28  20  22  12  11   8   5   1   0   1   0   2]
-% [384 109  45  27  15   4   0   0   0   0   0   0   0   0   0   0   0   0
-%    0   0]
-% [533 153  67  40  23  10   3   1   2   0   0   0   0   0   0   0   0   0
-%    0   0]
+In projection-corrected tangential velocity space, the fraction of
+stars with rotation period detections remains high out to roughly
+5\kms.  \citet{meingast_2021} by comparison required a physically
+motivated cut in tangential velocity space of 1.5\kms.  
+Our results show that at the expense of higher field star
+contamination rates, bonafide members can be identified even out to
+higher velocity separations.
+{\bf todo: anything else to mention? e.g., the actual numbers?}
 
@@
 \subsection{How did the halo form?}
+\label{subsec:origin}
+
+Differential rotation in the galaxy is clearly important. Meingast+21
+established this; in ten different cluster halos, nine showed leading
+and trailing arms oriented in line with the differential rotation of
+the Galaxy.
+(See e.g., their Figure 13).
+
 Another way to frame this question: should we call these stars a tidal
@@
   % 	exoplanet:theano},
+	\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
 	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
@@
 
+\section{Projection effects for 2D velocities}
+\label{app:vproj}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f12a.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f12b.pdf}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Projection effects for co-moving populations across
+  the sky.} The map is colored by the velocity difference a star
+  co-moving with \cn\ would exhibit across the equatorial sphere.
+  Actual positions of candidate \cn\ members are shown with points;
+  the star denotes the cluster center.
+	\label{fig:vproj}
+	}
+\end{figure}
+
+Figure~\ref{fig:vproj} shows why the correct approach for comparing
+any star's proper motions against the cluster's mean depends on the
+position of the star being considered.
+
 \section{Differential Reddening}

### 2021-03-09 earhart/physicalpositions.py

@@ -39,6 +39,11 @@ def append_physicalpositions(df, median_df):
     c_median = get_galcen(given_gaia_df_get_icrs_arr(median_df, zero_rv=1))
     c_df = get_galcen(given_gaia_df_get_icrs_arr(df, zero_rv=1))
 
+    # ditto, but in icrs. used in the comoving projection trick.
+    # as in plotting.plot_vtangential_projection
+    icrs_c_median = given_gaia_df_get_icrs_arr(median_df, zero_rv=0)
+    icrs_c_df = given_gaia_df_get_icrs_arr(df, zero_allvelocities=1)
+
     # get 3D separation from cluster median
     sep_3d_position_pc = (c_median.separation_3d(c_df)).to(u.pc).value
 
@@ -77,19 +82,50 @@ def append_physicalpositions(df, median_df):
             new_val = c_km_per_sec
             df[new_key] = new_val
 
+            #
+            # sophisticated thing: actually correct for the projection
+            # effect.
+            #
+            vel_to_add = np.ones(len(df))*icrs_c_median.velocity
+            _ = icrs_c_df.data.to_cartesian().with_differentials(vel_to_add)
+            c_full = icrs_c_df.realize_frame(_)
+
+            if c == 'pmra':
+                c_AU_per_yr = ((( nparr(df[c]) - nparr(c_full.pm_ra_cosdec) )*1e-3) * d_pc)*(u.AU/u.yr)
+            elif c == 'pmdec':
+                c_AU_per_yr = ((( nparr(df[c]) - nparr(c_full.pm_dec) )*1e-3) * d_pc)*(u.AU/u.yr)
+
+            c_km_per_sec = c_AU_per_yr.to(u.km/u.second)
+            new_key = 'delta_'+c+'_prime_km_s'
+            new_val = c_km_per_sec
+            df[new_key] = new_val
+
     df['delta_mu_km_s'] = np.sqrt(
         df['delta_pmra_km_s']**2 + df['delta_pmdec_km_s']**2
+    )
+    df['delta_mu_prime_km_s'] = np.sqrt(
+        df['delta_pmra_prime_km_s']**2 + df['delta_pmdec_prime_km_s']**2
     )
 
     return df
 
 
-def given_gaia_df_get_icrs_arr(df, zero_rv=0):
+def given_gaia_df_get_icrs_arr(df, zero_rv=0, zero_allvelocities=0):
     """
[…]

### 2021-03-09 earhart/plotting.py

@@ -2451,29 +2451,29 @@ def plot_physical_X_rotation(outdir, basedata=None, show1937=0,
 
     # make it!
     plt.close('all')
-    f, axs = plt.subplots(figsize=(4,3), nrows=2, ncols=2)
+    f, axs = plt.subplots(figsize=(4,4), nrows=2, ncols=2)
     axs = axs.flatten()
 
     xytuples = [
-        ('delta_x_pc', 'delta_y_pc'),
-        ('delta_x_pc', 'delta_z_pc'),
-        ('delta_y_pc', 'delta_z_pc'),
-        ('delta_pmra_km_s', 'delta_pmdec_km_s')
+        ('x_pc', 'y_pc'),
+        ('x_pc', 'z_pc'),
+        ('y_pc', 'z_pc'),
+        ('delta_pmra_prime_km_s', 'delta_pmdec_prime_km_s')
     ]
     ldict = {
-        'delta_x_pc':'$\Delta$X [pc]',
-        'delta_y_pc':'$\Delta$Y [pc]',
-        'delta_z_pc':'$\Delta$Z [pc]',
-        'delta_pmra_km_s': r"$\Delta \mu_{{\alpha'}}$ [km$\,$s$^{-1}$]",
-        'delta_pmdec_km_s': r"$\Delta \mu_{\delta}$ [km$\,$s$^{-1}$]"
+        'x_pc':'X [pc]',
+        'y_pc':'Y [pc]',
+        'z_pc':'Z [pc]',
+        'delta_pmra_prime_km_s': r"$\Delta \mu_{{\alpha'}}^{*}$ [km$\,$s$^{-1}$]",
+        'delta_pmdec_prime_km_s': r"$\Delta \mu_{\delta}^{*}$ [km$\,$s$^{-1}$]"
     }
 
     # limit axis by 99th percentile or iqr
     qlimd = {
-        'delta_x_pc': 0, 'delta_y_pc': 0, 'delta_z_pc': 0, 'delta_pmra_km_s': 0, 'delta_pmdec_km_s': 0
+        'x_pc': 0, 'y_pc': 0, 'z_pc': 0, 'delta_pmra_km_s': 0, 'delta_pmdec_km_s': 0
     }
     nnlimd = {
-        'delta_x_pc': 1, 'delta_y_pc': 1, 'delta_z_pc': 1, 'delta_pmra_km_s': 1, 'delta_pmdec_km_s': 1
[…]

### 2021-03-10 cluster_paper/ms.tex
update ms

@@
   {\bf Members of NGC\,2516 exist hundreds of parsecs from the core,
-  and up to $\approx$5\kms\ from the core in tangential velocity.} Stars
+  and up to $\approx$5\kms\ from the core in tangential velocity.}
+  Halo members
   that meet both kinematic and rotational membership criteria are shown in
-  blue.  Stars for which rotation periods should have been detectable,
+  blue.  Halo members for which rotation periods should have been detectable,
   but were not detected, are shown in orange.  Projections are shown in galactic coordinates and

### 2021-03-10 cluster_paper/ms.tex
lithium: fixed the interpolation bug

@@
 
-The third and final approach we took for age-dating the stellar
-population was using the lithium deplition technique (CITE, CITE,
-CITE).
-
-Reviewing the literature, two spectroscopic datasets for \cn\ seemed
-particularly important: Gaia-ESO 
-\citep{randich_gaiaeso_2018} and GALAH DR3 \citep{buder_galah_2020}.
-The target selection and results from each survey were as follows.
+The third and final approach we took for confirming the youth of the
+halo population of \cn\ was an analysis of the Li\,\textsc{I}
+6708\,\AA doublet.  The method is reviewed by
+\citet{soderblom_ages_2010}, and its connection to stellar rotation
+has recently been surveyed by \citet{bouvier_lithium-rotation_2020}.
+For \cn, two spectroscopic datasets  seemed particularly important:
+Gaia-ESO \citep{randich_gaiaeso_2018} and GALAH DR3
+\citep{buder_galah_2020}.  The target selection and results from each
+survey were as follows.
 
@@
 members and publicly available optical and NIR photometry.  This
-target selection approach included many stars in the spatial proximity
-of previously known members.  Since the possible existence of the \cn\ 
-halo was not known at the time of targetting, very few ``halo'' stars
-are in their ultimate sample.  For \cn, \citet{randich_gaiaeso_2018}
-ultimately reported stellar parameters (including lithium equivalent
-widths and metallicities) for 796 candidate members.  Cross-matching
-against our kinematic list of \nkinematic candidate members by position and
-imposing a 0$\farcs$5 maximum separation limit yields 492 kinematic
-members for which a Gaia-ESO spectrum was acquired and stellar
-parameters were reported.  15 of these (all with $(Bp-Rp)_0 > 2.0$)
-are spurious matches based on the Gaia color and GES effective
-temperature, and we remove them.  This yields 477 stars, of which 436
-are in the core, and 41 are in the halo.  This ratio is primarily
-caused by the Gaia-ESO targetting selection function.
+target selection approach included some stars in the spatial proximity
+of previously known members.  Since the existence of the \cn\ halo was
+not known at the time of targetting, very few ``halo'' stars are in
+their sample.  For \cn, \citet{randich_gaiaeso_2018} ultimately
+reported stellar parameters (including lithium equivalent widths and
+metallicities) for 796 stars that they considered possible \cn\
+members.  Cross-matching against our kinematic list of \nkinematic
+candidate members by position and imposing a 0$\farcs$5 maximum
+separation limit yields 492 kinematic members for which a Gaia-ESO
+spectrum was acquired and stellar parameters were reported.  15 of
+these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on the
+Gaia color and GES effective temperature, and we remove them.  This
+yields 477 stars, of which 436 are in the core, and 41 are in the
+halo.  The lop-sided ratio is primarily caused by the Gaia-ESO
+selection function.
 
@@
 are that the ``main'' survey targeted $12.0<V<14.0$ stars at
-$\delta<+10^\circ$, provided the stars were at least ten degrees from the
-galactic plane.  The survey is spatially inhomogeneous\footnote{See
-the footprint at
+$\delta<+10^\circ$, provided the stars were at least ten degrees from
+the galactic plane.  The survey is spatially
+inhomogeneous\footnote{See the footprint at
 \url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
@@
 for known open cluster members was also performed.  We identified the
-candidate \cn\ members
-for which spectra had been obtained by searching
-the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after excluding
-stars with the stellar parameter bit flags 1, 2, 3.  This excludes
-spectra with unreliable broadening, low S/N, and unreliable wavelength
-solutions (see Table~6 of \citealt{buder_galah_2020}).  Since our main
-focus is measuring equivalent widths for the Li\,6708\ \AA line, these
-were the most relevant bit flags for our purpose.  Of our \nkinematic
-candidate \cn\ members, 106 had spectra in GALAH DR3.  51 were in the
-``core''; 55 were in the ``halo''.
-77 had ``finite'' lithium, i.e., no lithium flag set.
-
+candidate \cn\ members for which spectra had been obtained by
+searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
+excluding stars with the stellar parameter bit flags 1, 2, 3.  This
+excludes spectra with unreliable broadening, low S/N, and unreliable
+wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
+Since our main focus is measuring equivalent widths for the
+Li\,6708\,\AA doublet, these were the most relevant bit flags.  Of our
+\nkinematic candidate \cn\ members, 107 had spectra in GALAH DR3.  51
+were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
+lithium, i.e., no lithium flag set.
+
+%FIXME FIXME FIXME FIXME
+%FIXME FIXME FIXME FIXME
+
+%
 %/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
+%
 % https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
 We downloaded\footnote{Via \url{datacentral.org.au/services/download},
-using the \texttt{sobject\_id} identifiers.}
-the GALAH DR3 spectra for all 106 entries.
+using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
+all 107 entries.
 We then measured the lithium equivalent widths using PROCESS Y.

### 2021-03-10 drivers/measure_galah_dr3_Li_EWs.py

@@ -75,6 +75,7 @@ for _, r in mdf.iterrows():
                        writecsvresults=True)
 
     else:
+        print(f'found {csv_outpath}')
         pass
 
 #
@@ -90,32 +91,36 @@ for csvpath in result_csvs:
 
 li_df = pd.concat(dfs)
 li_df['source_id'] = sourceids
-FUDGE = 65
-li_df['Fitted_Li_EW_mA_plus_fudge'] = li_df['Fitted_Li_EW_mA'] + FUDGE
 
 galah_li_path = os.path.join(
-    DATADIR, 'lithium', 'galahdr3_fullfaintkinematic_xmatch_20210211.csv'
+    DATADIR, 'lithium', 'galahdr3_fullfaintkinematic_xmatch_20210310.csv'
 )
 li_df.to_csv(galah_li_path, index=False)
 print(f'Saved {galah_li_path}')
 
 datapath = os.path.join(DATADIR, 'lithium',
-                        'randich_fullfaintkinematic_xmatch_20210128.csv')
+                        'randich_fullfaintkinematic_xmatch_20210310.csv')
 if not os.path.exists(datapath):
     from earhart.lithium import _make_Randich18_xmatch
     _make_Randich18_xmatch(datapath, vs_rotators=0)
 rdf = pd.read_csv(datapath)
 
 mdf = li_df.merge(rdf, how='inner', on='source_id')
+mdf = mdf.drop_duplicates(subset='source_id', keep='first')
+mdf = mdf[(~pd.isnull(mdf.EWLi))]
+print(f'Got {len(mdf)} calibration stars')
 
 plt.close('all')
-f,ax = plt.subplots(figsize=(4,3))
+f,ax = plt.subplots(figsize=(4,4))
 ax.scatter(mdf.EWLi, mdf.Fitted_Li_EW_mA)
-ax.scatter(mdf.EWLi, mdf.Fitted_Li_EW_mA+FUDGE)
 ax.plot(np.arange(-50,250,1), np.arange(-50,250,1), 'k--')
 ax.set_xlabel('Randich+18 EW Li from Gaia-ESO [mA]')
 ax.set_ylabel('My EW Li from GALAHDR3 [mA]')
 outpath = '../results/lithium/validate_my_GALAH_EWs_vs_Randich18.png'
 f.savefig(outpath, bbox_inches='tight', dpi=400)
+print(f'Made {outpath}')
 plt.close('all')
 
+print(42*'-')
[…]

### 2021-03-10 drivers/plot_lithium.py

@@ -6,13 +6,13 @@ PLOTDIR = os.path.join(RESULTSDIR, 'lithium')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+ep.plot_galah_dr3_lithium_abundance(PLOTDIR)
+ep.plot_galah_dr3_lithium_abundance(PLOTDIR, corehalosplit=1)
+
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=0)
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=0, galahdr3=1)
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=0)
 ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1)
-
-ep.plot_galah_dr3_lithium_abundance(PLOTDIR)
-ep.plot_galah_dr3_lithium_abundance(PLOTDIR, corehalosplit=1)
 
 ep.plot_randich_lithium(PLOTDIR)
 ep.plot_randich_lithium(PLOTDIR, corehalosplit=1)

### 2021-03-10 earhart/lithium.py

@@ -6,6 +6,7 @@ Contents:
 """
 import os
 import numpy as np, pandas as pd
+from numpy import array as nparr
 from astropy.io import fits
 from astropy.table import Table
 from astropy import units as u, constants as const
@@ -16,7 +17,7 @@ from earhart.paths import DATADIR, RESULTSDIR
 def _get_lithium_EW_df(gaiaeso, galahdr3):
 
     datapath = os.path.join(DATADIR, 'lithium',
-                            'randich_fullfaintkinematic_xmatch_20210128.csv')
+                            'randich_fullfaintkinematic_xmatch_20210310.csv')
     if not os.path.exists(datapath):
         from earhart.lithium import _make_Randich18_xmatch
         _make_Randich18_xmatch(datapath, vs_rotators=0)
@@ -147,6 +148,7 @@ def _make_Randich18_xmatch(datapath, vs_rotators=1, RADIUS=0.5):
     rdf = get_Randich18_NGC2516()
 
     if vs_rotators:
+        raise DeprecationWarning
         rotdir = os.path.join(DATADIR, 'rotation')
         rot_df = pd.read_csv(
             os.path.join(rotdir, 'ngc2516_rotation_periods.csv')
@@ -156,11 +158,9 @@ def _make_Randich18_xmatch(datapath, vs_rotators=1, RADIUS=0.5):
     else:
 
         from earhart.helpers import _get_fullfaint_dataframes
-        nbhd_df, cg18_df, kc19_df, target_df = _get_fullfaint_dataframes()
-        cg18_df['subcluster'] = 'core'
-        kc19_df['subcluster'] = 'halo'
-        comp_df = pd.concat((cg18_df, kc19_df))
-        print('Comparing vs the "fullfaint" kinematic NGC2516 rotators sample (core + halo)...')
+        nbhd_df, core_df, halo_df, full_df, target_df = _get_fullfaint_dataframes()
+        comp_df = full_df
+        print(f'Comparing vs the {len(comp_df)} "fullfaint" kinematic NGC2516 rotators sample (core + halo)...')
 
     c_comp = SkyCoord(ra=nparr(comp_df.ra)*u.deg, dec=nparr(comp_df.dec)*u.deg)
     c_r18 = SkyCoord(ra=nparr(rdf._RA)*u.deg, dec=nparr(rdf._DE)*u.deg)
[…]

### 2021-03-10 earhart/plotting.py

@@ -1299,11 +1299,9 @@ def plot_galah_dr3_lithium_abundance(outdir, corehalosplit=0):
     g_df = pd.DataFrame(g_dict)
 
     basedata = 'fullfaint'
-    nbhd_df, cg18_df, kc19_df, trgt_df = get_gaia_basedata(basedata)
-    cg18_df['subcluster'] = 'core'
-    kc19_df['subcluster'] = 'halo'
-    comp_df = pd.concat((cg18_df, kc19_df))
-    print('Comparing vs the "fullfaint" kinematic NGC2516 rotators sample (core + halo)...')
+    nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
+    comp_df = full_df
+    print(f'Comparing vs the {len(comp_df)} "fullfaint" kinematic NGC2516 rotators sample (core + halo)...')
     assert type(comp_df.source_id.iloc[0]) == np.int64
 
     mdf = comp_df.merge(g_df, on='source_id', how='inner')
@@ -1633,7 +1631,7 @@ def plot_rotation_X_lithium(outdir, cmapname, gaiaeso=0, galahdr3=0):
     Plot Prot vs (Bp-Rp)_0, color points by Li EW.
     """
 
-    assert abs(AVG_EBpmRp - 0.1343) < 1e-4 # used by KC19
+    assert abs(AVG_EBpmRp - 0.1343) < 1e-4
 
     set_style()
 

### 2021-03-10 cluster_paper/ms.tex
update ms

@@
 
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		%\subfloat{
+			\includegraphics[width=0.49\textwidth]{f5a.pdf}
+		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
+		%}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
+  Equivalent width of the 6708\,\AA\ doublet is shown
+  versus extinction-corrected color for all kinematic \cn\ members
+  with Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.
+  The GALAH spectra comprise slightly over half of the halo stars,
+  due to the non-targeted selection function of that survey.
+  Points with ${\rm EW}\approx0\,$m\AA\ are non-detections.
+  \label{fig:lithiumcorehalo}
+  }
+\end{figure}
 \begin{figure*}[t]
@@
 		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f5a.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.75\textwidth]{f5b.pdf}
-		}
+		%\subfloat{
+		\includegraphics[width=1\textwidth]{f5b.pdf}
+		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
+		%}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf Lithium equivalent widths in NGC\,2516,}
-  plotted against extinction-corrected color ({\it top}), and the
-  combination of rotation period and extinction-corrected color ({\it
-  bottom}).
-  {\bf TODO: Do this against the autorot sample; Add the GALAH EWs}
+	\caption{ {\bf Lithium and rotation in NGC\,2516.}
+		Points show all kinematic \cn\ members with TESS rotation
+		periods in the Gaia-ESO or GALAH surveys, and are colored by
+		equivalent width.
+	   Non-detections appear with ${\rm EW}\approx0\,$m\AA.
+	The K-dwarfs shows a similar lithium-rotation correlation
+		to what has been previously observed,
+		and is discussed in Section~\ref{subsec:lithium}.
 		\label{fig:lithium}
@@
 halo population of \cn\ was an analysis of the Li\,\textsc{I}
-6708\,\AA doublet.  The method is reviewed by
+6708\,\AA\ doublet.  The method has been reviewed by
 \citet{soderblom_ages_2010}, and its connection to stellar rotation
@@
 of previously known members.  Since the existence of the \cn\ halo was
-not known at the time of targetting, very few ``halo'' stars are in
-their sample.  For \cn, \citet{randich_gaiaeso_2018} ultimately
+not known at the time of target selection, very few ``halo'' stars are in
+the sample.  For \cn, \citet{randich_gaiaeso_2018} ultimately
 reported stellar parameters (including lithium equivalent widths and
@@
 selection function.
+%FIXME VERIFY THAT THESE NUMBERS DON'T INCLUDE THEIR NANS...
 
@@
 \section{Kinematics $\otimes$ Rotation}
+\label{app:gaia6d_x_rotn}
 

### 2021-03-10 earhart/lithium.py

@@ -14,7 +14,7 @@ from astropy.coordinates import SkyCoord
 
 from earhart.paths import DATADIR, RESULTSDIR
 
-def _get_lithium_EW_df(gaiaeso, galahdr3):
+def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99):
 
     datapath = os.path.join(DATADIR, 'lithium',
                             'randich_fullfaintkinematic_xmatch_20210310.csv')
@@ -36,12 +36,18 @@ def _get_lithium_EW_df(gaiaeso, galahdr3):
 
     if gaiaeso and galahdr3:
         df = pd.concat((s_gaiaeso_df, s_galah_df))
+        df = df[df.Li_EW_mA > EW_CUTOFF_mA]
+        print(f'Gaia-ESO + GALAH: got {len(df[~pd.isnull(df.Li_EW_mA)])} finite Li EWs > {EW_CUTOFF_mA} mA')
     if gaiaeso and not galahdr3:
         df = s_gaiaeso_df
+        df = df[df.Li_EW_mA > EW_CUTOFF_mA]
+        print(f'Gaia-ESO: got {len(df[~pd.isnull(df.Li_EW_mA)])} finite Li EWs > {EW_CUTOFF_mA} mA')
     if not gaiaeso and galahdr3:
         df = s_galah_df
+        df = df[df.Li_EW_mA > EW_CUTOFF_mA]
+        print(f'GALAH: got {len(df[~pd.isnull(df.Li_EW_mA)])} finite Li EWs > {EW_CUTOFF_mA} mA')
 
-    return df
+    return df[df.Li_EW_mA > EW_CUTOFF_mA]
 
 def get_GalahDR3_li_EWs(verbose=1):
     """

### 2021-03-10 earhart/plotting.py

@@ -1647,7 +1647,7 @@ def plot_rotation_X_lithium(outdir, cmapname, gaiaeso=0, galahdr3=0):
     # check crossmatch quality
     #
     plt.close('all')
-    f, ax = plt.subplots(figsize=(4.5,3))
+    f, ax = plt.subplots(figsize=(6,3))
 
     # color scheme
     if cmapname == 'nipy_spectral':
@@ -1661,7 +1661,7 @@ def plot_rotation_X_lithium(outdir, cmapname, gaiaeso=0, galahdr3=0):
     cax = ax.scatter(
         mdf[sel]['phot_bp_mean_mag'] - mdf[sel]['phot_rp_mean_mag'] - AVG_EBpmRp,
         mdf[sel]['period'],
-        c=nparr(mdf[sel]['Li_EW_mA']), alpha=1, zorder=2, s=10, edgecolors='k',
+        c=nparr(mdf[sel]['Li_EW_mA']), alpha=1, zorder=2, s=15, edgecolors='k',
         marker='o', cmap=cmap, norm=norm, linewidths=0.3
     )
 
@@ -1672,7 +1672,8 @@ def plot_rotation_X_lithium(outdir, cmapname, gaiaeso=0, galahdr3=0):
 
     ax.set_ylabel('Rotation Period [days]')
     ax.set_xlabel('(Bp-Rp)$_0$ [mag]')
-    ax.set_xlim((0.5, 1.5))
+    # ax.set_xlim((0.5, 1.5))
+    # ax.set_ylim((0.5, 1.5))
 
     format_ax(ax)
     outstr = '_' + cmapname

### 2021-03-10 cluster_paper/ms.tex
update ms

@@
 For \cn, two spectroscopic datasets seemed particularly important:
-Gaia-ESO \citep{randich_gaiaeso_2018} and GALAH DR3
-\citep{buder_galah_2020}.  The target selection and results from each
-survey were as follows.
-
-Gaia-ESO selected {\it cluster candidates} to be observed with the
-GIRAFFE and UVES spectrographs based on previously reported literature
-members and publicly available optical and NIR photometry.  This
-target selection approach included some stars in the spatial proximity
-of previously known members.  Since the existence of the \cn\ halo was
-not known at the time of target selection, very few ``halo'' stars are in
-the sample.  For \cn, \citet{randich_gaiaeso_2018} ultimately
-reported stellar parameters (including lithium equivalent widths and
-metallicities) for 796 stars that they considered possible \cn\
-members.  Cross-matching against our kinematic list of \nkinematic
-candidate members by position and imposing a 0$\farcs$5 maximum
-separation limit yields 492 kinematic members for which a Gaia-ESO
-spectrum was acquired and stellar parameters were reported.  15 of
-these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on the
-Gaia color and GES effective temperature, and we remove them.  This
-yields 477 stars, of which 436 are in the core, and 41 are in the
-halo.  The lop-sided ratio is primarily caused by the Gaia-ESO
-selection function.
-%FIXME VERIFY THAT THESE NUMBERS DON'T INCLUDE THEIR NANS...
+Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
+\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
+GALAH DR3 were the most relevant, respectively \citep[{\it
+e.g.},][]{randich_gaiaeso_2018,buder_galah_2020}.  The target
+selection and results from each survey were as follows.
+
+Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
+and UVES spectrographs based on previously reported literature members
+and publicly available photometry.  Since the existence of the \cn\
+halo was not known at the time of target selection, very few ``halo''
+stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
+ultimately reported stellar parameters (including lithium equivalent
+widths and metallicities) for 796 stars that they considered possible
+\cn\ members.  Cross-matching against our kinematic list of
+\nkinematic candidate members by position and imposing a 0$\farcs$5
+maximum separation limit yields 492 kinematic members for which a
+Gaia-ESO spectrum was acquired and stellar parameters were reported.
+15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
+the Gaia color and GES effective temperature, and we remove them.
+This yields 477 stars, of which 436 are in the core, and 41 are in the
+halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
+We verified by directly querying the ESO archive that
+\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
+spectra as part of their \cn\ sample, and so we opted to use their
+results rather than repeating their analysis.
 
@@
 
-The GALAH DR3 target selection process is discussed in detail by
-\citet{buder_galah_2020}.  The most relevant aspects for our project
-are that the ``main'' survey targeted $12.0<V<14.0$ stars at
+The GALAH DR3 target selection is discussed in depth by
+\citet{buder_galah_2020}.  The relevant aspects for our analysis
+are that the ``main'' survey targeted $12<V<14$ stars at
 $\delta<+10^\circ$, provided the stars were at least ten degrees from
@@
 
-%FIXME FIXME FIXME FIXME
-%FIXME FIXME FIXME FIXME
-
 %
@@
 using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
-all 107 entries.
+all 107 entries.  We then measured the lithium equivalent widths using
+the AstroPy-affiliated \texttt{specutils} package
+\citep{specutils_v1pt1}.
+
+%FIXME TODO: put it into words
+% circa ~line 1224 from ~/Dropbox/proj/cdips_followup/cdips_followup/spectools.py
+Our procedure was as follow...
+
 We then measured the lithium equivalent widths using PROCESS Y.

### 2021-03-12 cluster_paper/ms.tex
update ms

@@
   %An Open Cluster With a 300\,pc Halo
-  An Open Cluster Spread Across 400\,pc
+  An Open Cluster Spread Across 500\,pc
   %NGC\,2516  300\,pc
@@
   In this study, we focus on a halo with a tip-to-tip length of
-  $\approx$400\,pc reported around the $\approx$150\,Myr open cluster
+  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
   NGC\,2516.
@@
   of NGC\,2516 is real and coeval with the core.
-  Two in three kinematically selected halo members 25-250\,pc from the
+  Two in three kinematically selected halo members out to 250\,pc from the
   cluster center show rotation periods consistent with a
@@
 Young transiting planets are hard to find because young stars are
-rare, and mostly reside in crowded galactic plane \citep[see {\it
+rare, and mostly reside in the crowded galactic plane \citep[see {\it
 e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
@@
 and asymmetric structures with number densities down to 100 times less
-than the field around the same regions (see {\it e.g.}
+than the field around the same regions (see {\it e.g.},
 \citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
@@
 agnostic.  It suffers however from the suggestion of spherical
-symmetry.  The halo of NGC\,2516 might be better described as ``tails'',
+symmetry.  The halo of NGC\,2516 might be best described as ``tails'',
 but the primary goal of this study is not to verify whether these
@@
 We selected candidate \cn\ members based on those reported in the
-literature.  While a number of pre-Gaia studies have been performed,
+literature.  While some pre-Gaia analyses were available,
 the purity and accuracy of the Gaia-derived results are the current
@@
 considered stars brighter than $G=18$\,mag within a few degrees of the
-cluster center, and reported 1106 candidate members of \cn.
+cluster center, and reported 1106 candidate \cn\ members.
 \citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
@@
 about 200\,pc in physical space from the cluster core.  This
-corresponds to total a length of $\approx$400\,pc, depending on which
+corresponds to total a length of $\approx$500\,pc, depending on which
 members of the halo are chosen as the ``tips'' on either end.  This

### 2021-03-12 cluster_paper/ms.tex
update ms

@@
 \citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
-spectra as part of their \cn\ sample, and so we opted to use their
-results rather than repeating their analysis.
-
-% GALAH DR3 selected XXX/YYY.  GALAH-DR3 = GALAH-main + GALAH-faint +
-% K2-HERMES + TESS-HERMES + HERMES other.  Data taken both with the 2dF
-% fiber positioner, and the HERMES spectrograph at the 3.9 AAT
-% (Lewis+02, Sheinis+15).  GALAH-main is 12<V<14, $\delta<+10^\circ$,
-% $|b|>10^\circ$, in regions of sky with at least 400 targets in $\pi$
-% square degrees (the 2dF FoV).  Fields with more than 400 targets were
-% targetted multiple times, with different fiber arrangements.
-% TESS-HERMES: 10.0 < V < 13.1. TESS Southern CVZ, within $12^\circ$ of
-% the southern ecliptic pole.  54354 in the ``HERMES other'' program are
-% from targeted observations of stars in open clusters, the GALAH Pilot
-% Survey (Martell et al. 2017),
+Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
+their results rather than repeating their analysis.
 
@@
 all 107 entries.  We then measured the lithium equivalent widths using
-the AstroPy-affiliated \texttt{specutils} package
-\citep{specutils_v1pt1}.
-
-%FIXME TODO: put it into words
-% circa ~line 1224 from ~/Dropbox/proj/cdips_followup/cdips_followup/spectools.py
-Our procedure was as follow...
-
-We then measured the lithium equivalent widths using PROCESS Y.
-We then concatenated the GALAH DR3 and Gaia-ESO results together.
-
-% TODO: fix the fact that the figure currently only shows Randich+18
-Figure~\ref{fig:lithium} shows the results.  At fixed stellar mass
+the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
+by first focusing in on a 15\AA\ window centered on the 6707.835\AA
+lithium doublet.  We then continuum normalized the spectra using a
+third-degree Chebyshev series, while excluding any regions that showed
+absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
+6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
+to the continuum normalized spectra, considering only a $\pm 1\,$\AA
+window centered on the Li doublet.  The equivalent widths were then
+evaluated by numerically integrating the fitted model over the same
+window.  Our approach therefore includes the Fe\,6707.44 blend in the
+reported Li equivalent widths---which leads to systematic
+overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
+e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
+uncertainties on the EWs, we repeated the procedure twenty times,
+but added noise to the spectra, drawn from a normal distribution with
+a scale set by the standard deviation of the continuum.
+The reported uncertainties are then drawn from the $84^{\rm th}$ and
+$16^{\rm th}$ percentiles of the resulting EW distribution.
+We verified the overall scale of our results by comparing the 21
+stars that overlapped between GALAH and Gaia-ESO.
+17 of the 21 measurements agreed within 1-$\sigma$.
+The remaining four were stars with $T_{\rm eff}>6200\,{\rm K}$, for
+which \citet{randich_gaiaeso_2018} reported Li EW detections
+ranging from $50$ to $75\,$m\AA, and for which we found 
+either null results or only slightly significant
+($\approx$30m\AA) detections.
+
+% The outliers:
+%    Fitted_Li_EW_mA     EWLi            source_id  Teff
+%    24           -1.490   76.100  5290730488348194304  6786
+%    34           19.485   71.300  5290935341107132800  6702
+%    15           20.251   50.867  5290722310730727040  6197
+%    25           29.378   52.500  5290724681552584192  6511
+
+Figure~\ref{fig:lithium} shows the results.  
+At fixed stellar mass
 (and age), the rapid rotators tend to show elevated lithium equivalent
@@
 The authors thank X and Y for fruitful discussions.
+
+B.~Tofflemire, % lithium suggestoins
+
 %

### 2021-03-13 cluster_paper/ms.tex
update ms
with lithium

@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
@@
 		%\subfloat{
-			\includegraphics[width=0.49\textwidth]{f5a.pdf}
+			\includegraphics[width=1\textwidth]{f5a.pdf}
 		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
@@
   due to the non-targeted selection function of that survey.
+  The gray points are Keck-HIRES measurements of stars with detected
+  transiting planets from the Kepler mission
+  \citep{berger_identifying_2018}.
   Points with ${\rm EW}\approx0\,$m\AA\ are non-detections.
@@
   }
-\end{figure}
+\end{figure*}
 \begin{figure*}[t]
@@
 		and is discussed in Section~\ref{subsec:lithium}.
-		\label{fig:lithium}
+		\label{fig:lithiumrot}
 	}
@@
 
-Figure~\ref{fig:lithium} shows the results.  
+Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
+Gaia-ESO and GALAH results, with the points colored by whether the
+stars are in the core or the halo of the cluster.  The GALAH spectra
+span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
+cutoff of the survey.  The overall depletion of stars redder than
+$(Bp-Rp)_0\gtrsim1.4$, {\it i.e.}, later than K4.5V (0.71\,$M_\odot$),
+is visible relative to the earlier K dwarfs.  For comparison, we have
+also plotted the Li EWs measured by \citet{berger_identifying_2018}
+for planet-hosting stars in the Kepler field (we interpolated from
+their effective temperatures using the \citet{pecaut_mamajek_2013}
+table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
+version 2019.3.22.}.
+The field star distribution peaks at late F-type stars.
+The majority of kinematically selected stars in the core and the halo
+show lithium equivalent widths substantially in excess of these field
+stars.
+
+We can go one step further, and match our sample of kinematically
+selected stars with spectra against the TESS rotators.
+The result is shown in Figure~\ref{fig:lithiumrot}
 At fixed stellar mass
@@
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
-The fact that we see the same trend across the core and halo of \cn\ 
-{\it a)} supports the conclusion that the halo is coeval with the
-core, and {\it b)} suggests that the lithium-rotation correlation is
-not explained by environmental effects such as the density of the cluster
-during its embedded phase, but is instead tied to physical processes
-acting on all low-mass pre-main-sequence stars.
+
+% The fact that we see the same trend across the core and halo of \cn\ 
+% {\it a)} supports the interpretation that the halo is coeval with the
+% core, and {\it b)} suggests that the lithium-rotation correlation is
+% not explained by environmental effects such as the density of the cluster
+% during its embedded phase, but is instead tied to physical processes
+% acting on all low-mass pre-main-sequence stars.
 
@@
 % these?
+
+An important point regarding the lithium-rotation trend is that it has
+been shown to carry over beyond just equivalent widths, and to the
+actual abundance of lithium in the photosphere, A(Li).  A priori, one
+might worry that the trend is tautological: perhaps some feature the
+stellar equatorial velocity, or the microturbulent velocity at the
+stellar surface, are larger for faster rotators and this could affect
+the formation conditions of the lithium doublet.
+%FIXME improve these words
+Soderblom+1993 calculated the relevant LTE curves of growth for the
+lithium line, and argued that this could not be the case.
+Since then, LindAsplund+2009 have calculated the relevant non-LTE
+correction, and has shown that it affects the abundances by
+$<0.5\,{\rm dex}$.
+The order of magnitude of the rotation-abundance trend is 2-3\,dex.
+
+%     * OK, then to convert to an ABUNDANCE, follow Bouvier+2018. In other words,
+%       just interpolate against the LTE curve of growth tables from
+%       Soderblom+1993.
+% 
+%     * Nb. the Non-LTE correction has been shown to be <0.5 dex (but
+%       >0.1 dex, for sun-like stars with Li abundances above 2 dex,
+%       >which is the regime we're interested in for Pleaides and NGC
+%       >2516).
+%       CITE: Lind, Asplund, + 2009
+%       https://ui.adsabs.harvard.edu/abs/2009A%26A...503..541L/abstract
+
 Processes both internal and external to the star have been suggested to explain
@@
 \label{subsec:origin}
+% Should we be calling the halo a tail?
 
@@
 
-\subsection{Mass differences between center and outer reaches?}
+% \subsection{Mass differences between center and outer reaches?}
 
@@
 Some final discussion is warranted on the applicability of our
-approach more generally.  This work starts with a kinematically
-identified population, and then uses photometry (Gaia HR diagrams;
+approach more generally.  Given a kinematically
+identified population, this study uses photometry (Gaia HR diagrams;
 TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
@@
 
-One can start with photometric criteria (e.g., Gaia HR diagram of
-everything within some distance; Zari+18), or with spectroscopic
+One can use photometric criteria (e.g., Gaia HR diagram of
+everything within some distance; Zari+18), or spectroscopic
 criteria (CITE Berger+18), or with combinations thereof (e.g., CITE

### 2021-03-13 cluster_paper/collect.sh

@@ -14,7 +14,7 @@ cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_ga
 cp ../results/physical_X_rotation/physical_X_rotation_fullfaint.pdf f8a.pdf
 cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f8b.pdf
 
-cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit.pdf f5a.pdf
+cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit_showkepfield.pdf f5a.pdf
 cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_lithium_viridis_gaiaeso_galahdr3.pdf f5b.pdf
 
 cp ../results/vtangential_projection/XYZ_with_orbit.pdf f11.pdf

### 2021-03-13 drivers/plot_lithium.py

@@ -6,10 +6,21 @@ PLOTDIR = os.path.join(RESULTSDIR, 'lithium')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=0)
-ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=0, galahdr3=1)
-ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=0)
-ep.plot_lithium_EW_vs_color(PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1)
+ep.plot_lithium_EW_vs_color(
+    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1
+)
+ep.plot_lithium_EW_vs_color(
+    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1, showkepfield=1
+)
+ep.plot_lithium_EW_vs_color(
+    PLOTDIR, gaiaeso=1, galahdr3=0
+)
+ep.plot_lithium_EW_vs_color(
+    PLOTDIR, gaiaeso=0, galahdr3=1
+)
+ep.plot_lithium_EW_vs_color(
+    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=0
+)
 
 ep.plot_galah_dr3_lithium_abundance(PLOTDIR)
 ep.plot_galah_dr3_lithium_abundance(PLOTDIR, corehalosplit=1)

### 2021-03-13 earhart/lithium.py

@@ -1,5 +1,6 @@
 """
 Contents:
+    _get_lithium_EW_df
     get_Randich18_NGC2516
     get_GalahDR3_lithium
     get_GalahDR3_li_EWs
@@ -48,6 +49,7 @@ def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99):
         print(f'GALAH: got {len(df[~pd.isnull(df.Li_EW_mA)])} finite Li EWs > {EW_CUTOFF_mA} mA')
 
     return df[df.Li_EW_mA > EW_CUTOFF_mA]
+
 
 def get_GalahDR3_li_EWs(verbose=1):
     """

### 2021-03-13 earhart/plotting.py

@@ -1545,15 +1545,12 @@ def plot_randich_lithium(outdir, vs_rotators=1, corehalosplit=0):
         savefig(f, outpath)
 
 
-def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0, corehalosplit=0):
+def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
+                             corehalosplit=0, showkepfield=0):
     """
-    Plot Li EW vs color for Randich+18's Gaia ESO lithium stars, crossmatched
-    against the "fullfaint kinematic" sample.
-
-    Relevant R+18 columns:
-        'CName', 'Cluster', 'Inst', 'Teff', 'e_Teff', 'logg', 'e_logg',
-        '__Fe_H_', 'e__Fe_H_', 'l_EWLi', 'EWLi', 'e_EWLi', 'f_EWLi'
-
+    Plot Li EW vs color for a) Randich+18's Gaia ESO lithium stars, b)
+    the GALAH DR3 EWs, or c) both, after crossmatching against the
+    "fullfaint kinematic" sample.
     """
 
     assert abs(AVG_EBpmRp - 0.1343) < 1e-4 # used by KC19
@@ -1572,7 +1569,18 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0, corehalosplit=0):
     # check crossmatch quality
     #
     plt.close('all')
-    f, ax = plt.subplots(figsize=(4,3))
+    f, ax = plt.subplots(figsize=(6,3))
+
+    if showkepfield:
+        from timmy.lithium import get_Berger18_lithium
+        bdf = get_Berger18_lithium()
+        bdf['BpmRp0'] = get_interp_BpmRp_from_Teff(bdf['Teff'])
+
+        ax.scatter(
+            bdf['BpmRp0'], bdf['EW_Li_'], c='gray', alpha=1,
+            zorder=-1, s=5, edgecolors='gray', marker='.',
+            linewidths=0, label='BHB18'
+        )
 
     if not corehalosplit:
         ax.scatter(
@@ -1618,6 +1626,8 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0, corehalosplit=0):
         outstr += '_galahdr3'
     if corehalosplit:
         outstr += '_corehalosplit'
+    if showkepfield:
[…]

### 2021-03-16 cluster_paper/ms.tex
update ms

@@
   %An Open Cluster With a 300\,pc Halo
-  An Open Cluster Spread Across 500\,pc
+  An Open Cluster Spanning 500\,Parsecs
   %NGC\,2516  300\,pc
@@
   of NGC\,2516 is real and coeval with the core.
-  Two in three kinematically selected halo members out to 250\,pc from the
-  cluster center show rotation periods consistent with a
+  Two thirds of kinematically selected halo members out to 250\,pc from the
+  cluster center have rotation periods consistent with a
   gyrochronological age of 150\,Myr.
   A comparison sample of field stars shows no such trend.
-  The rotation period and lithium abundances of the cluster stars are
-  also anti-correlated at fixed stellar mass, as has been observed in
-  similar stellar populations.
-  This work quantifies the degree to which Gaia-based
-  analyses can identify dispersed stellar populations against a
-  background of field star contaminants, and
-  expands the set of confirmed NGC\,2516 members by a factor
-  of $\approx$2.
+  In addition, the lithium abundances of stars in the halo are
+  elevated relative to the field, and are correlated with the stellar
+  rotation rate in a manner consistent with previous observations of the
+  Pleiades and Psc Eri.
+  This work expands the set of confirmed NGC 2516 members by a factor
+  of two, and is consistent with a paradigm in which the halos of many
+  open clusters are more populous than their cores.
   We highlight the implications for spectroscopic survey targeting,
@@
 
-We analyzed X, Y, Z.
-Our main results are as follows.
+The combination of Gaia, TESS, and the GALAH and Gaia-ESO
+spectroscopic surveys has clarified a few 
+things about \cn.
 \begin{itemize}
-	\item 
+	\item {\it Overdensities observed by Gaia in five-dimensional (3-D position + 2-D velocity)
+		phase space
+		revealed a halo of stars spanning 500 parsecs in the plane of the galaxy.}
+	The earlier reference to the full structure that we can find is by CITE KOUNKEL.
+	The halo is more precisely described as a leading and trailing tail
+	(Figure~\ref{fig:XYZ}), with the leading end on the interior of the
+	cluster's
+	galactic orbit.
+	%
+	\item {\it Isochronal, rotational, and lithium age-dating show that the halo is real and coeval with the core of \cn.}
+    The Gaia EDR3 photometry and astrometry show a main-sequence turnoff consistent with
+    an isochronal age of 150\,Myr for both the core and halo.
+    The faintest M dwarfs in the core and halo are also elevated relative to a field star comparison sample, but using
+    this to isochronally measure the age is challenging due to uncertainties in M dwarf atmospheric opacities.
+    Gyrochronally, FIXME FIXME FIXME
+    %
+    \item {\it The identified halo extends to $\approx$250\,pc in separate from the cluster center, and out to $\approx$5\kms\ in velocity separation.}
+    Figure~\ref{fig:physical_x_rotn} shows the overlap between kinematic
+    and rotational cluster members. 
+    %
+    \item {\it The field star contamination rate increases at larger physical and 
+    velocity separations from the cluster core.}
+	Figure~\ref{fig:physical_x_rotn}
+
+    The stars further from the c
+    As a reminder, a star with a 1\kms\ velocity difference moves 1\,pc per Myr.
+
+	More specifically, ``two thirds of kinematically selected halo members out to 250 pc from the cluster center have rotation periods consistent with a gyrochronological age of 150 Myr''.
+
+	\item {\it Differential rotation in the Milky Way likely plays an important role in shaping the halos of nearby open clusters.}
+	The Milky Way's differential rotation rate has a comparable
+	order of magnitude ($\approx$-0.2\,deg\,Myr$^{-1}$\,kpc$^{-1}$) to that measured across the halo \citep{meingast_2021}.
+	It therefore likely plays an important role in understanding the halo's formation.
+
 \end{itemize}

### 2021-03-16 cluster_paper/ms.tex
update ms

@@
 before Gaia, has now reached a breakthrough pace \citep[{\it
-e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
+e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
 The populations found by these studies can be summarized as follows.
@@
 formation complex, and the stars we see in the halo did not in fact
-form in say the same ``clump'' as those in the cluster core.
+form the same ``clump'' as those in the cluster core.
+
+Consider for instance the Orion\,A cloud.  \citet{grosschedl_3d_2018} showed
+using YSOs that formed in the past few megayears as tracers that Orion\,A is
+best described with two components:
+a dense ``head'', and a lower density 75\,pc long ``tail''.
+While this filament currently totals 90\,pc in length, its future evolutionary state could very well resemble something like \cn.
+A direct numerical exploration would be necessary to be sure {\bf FIXME}.
+
+CITE ALSO: ZUCKER 2017 BONES CATALOG.
+
+CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
+THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
+LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
+
+Similarly Zari et al 2019 studied the entire Orion complex, and found that
+different regions of the complex (one toward the Belt, one toward 25\,Ori)
+began forming stars roughly simultaneously. No evidence was found of it
+propagating from these initial sites.
+
+ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
+
+cometary-like cloud,
+oriented toward the Galactic plane, with two distinct components:
+a denser and enhanced star-forming (bent) Head, and a
+lower density and star-formation quieter 75 pc long Tail.
 

### 2021-03-18 cluster_paper/ms.tex
made li plot better

@@
 		equivalent width.
-	   Non-detections appear with ${\rm EW}\approx0\,$m\AA.
-	The K-dwarfs shows a similar lithium-rotation correlation
-		to what has been previously observed,
-		and is discussed in Section~\ref{subsec:lithium}.
+    Points with reported equivalent widths below 20\,m\AA\ (including
+    non-detections) are shown with gray crosses.
+    The lithium-rotation correlation for the K-dwarfs
+    has been observed in other young clusters, and is discussed in
+    Section~\ref{subsec:lithium}.
 		\label{fig:lithiumrot}

### 2021-03-18 drivers/plot_lithium.py

@@ -7,10 +7,10 @@ if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
 ep.plot_lithium_EW_vs_color(
-    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1
+    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1, showkepfield=1
 )
 ep.plot_lithium_EW_vs_color(
-    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1, showkepfield=1
+    PLOTDIR, gaiaeso=1, galahdr3=1, corehalosplit=1
 )
 ep.plot_lithium_EW_vs_color(
     PLOTDIR, gaiaeso=1, galahdr3=0

### 2021-03-18 earhart/plotting.py

@@ -1235,7 +1235,7 @@ def plot_compstar_rotation(outdir, E_BpmRp=0.1343, yscale=None):
             df = pd.read_csv(os.path.join(rotdir, f'curtis19_{_cls}.csv'))
 
         else:
-            df = get_autorotation_dataframe(_cls)
+            df = get_autorotation_dataframe(_cls, cleaning='defaultcleaning')
             df = df[df.phot_rp_mean_mag < 13]
             print(42*'-')
             print(f'Applying E(Bp-Rp) = {E_BpmRp:.4f}')
@@ -1555,6 +1555,9 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
 
     set_style()
 
+    #
+    # get data
+    #
     df = _get_lithium_EW_df(gaiaeso, galahdr3)
 
     basedata = 'fullfaint'
@@ -1563,17 +1566,18 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
 
     df = df.merge(mdf, how='left', on='source_id')
 
-    #
-    # check crossmatch quality
-    #
-    plt.close('all')
-    f, ax = plt.subplots(figsize=(6,3))
-
     if showkepfield:
         from timmy.lithium import get_Berger18_lithium
         bdf = get_Berger18_lithium()
         bdf['BpmRp0'] = get_interp_BpmRp_from_Teff(bdf['Teff'])
 
+    #
+    # make plot
+    #
+    plt.close('all')
+    f, ax = plt.subplots(figsize=(6,3))
+
+    if showkepfield:
         ax.scatter(
             bdf['BpmRp0'], bdf['EW_Li_'], c='gray', alpha=1,
             zorder=-1, s=5, edgecolors='gray', marker='.',
@@ -1588,20 +1592,19 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
             linewidths=0.3
         )
     else:
-        sel = (df.subcluster == 'core')
-        ax.scatter(
-            df[sel]['phot_bp_mean_mag'] - df[sel]['phot_rp_mean_mag'] - AVG_EBpmRp,
-            df[sel]['Li_EW_mA'],
-            c='k', alpha=1, zorder=2, s=8, edgecolors='k', marker='o',
-            linewidths=0.3, label='Core'
-        )
-        sel = (df.subcluster == 'halo')
[…]

### 2021-03-25 cluster_paper/ms.tex
update ms

@@
 \newcommand{\kms}{\,km\,s$^{-1}$}
+\newcommand{\ms}{\,m\,s$^{-1}$}
 
@@
 \accepted{---}
-\submitjournal{ApJL.}
+\submitjournal{AAS journals.}
 \shorttitle{Core \& Halo of NGC\,2516}
@@
   elevated relative to the field, and are correlated with the stellar
-  rotation rate in a manner consistent with previous observations of the
-  Pleiades and Psc Eri.
+  rotation rate in a manner consistent with previous observations.
   This work expands the set of confirmed NGC 2516 members by a factor
@@
   velocity after subtracting the local standard of rest: $\{v_{\rm X},
-  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.  The sizes of
-  the blue arrows are internally proportionate.  By comparison,
-  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
+  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By comparison,
+  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the inertial
+  frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
+  direction.
+  The sizes of the blue arrows are internally proportionate. 
   \label{fig:XYZ}
@@
 present in the aperture according to the Gaia DR2 source catalog.  For
-set $\mathcal{A}$, this yielded \nautorotnumerator light curves,
-out of \nautorotdenominator stars for which rotation periods might
-plausibly have been detected.  For set $\mathcal{B}$, we
-then additionally required that the Lomb-Scargle and Stellingwerf
+set $\mathcal{A}$, this yielded \nautorotnumerator light curves, out
+of \nautorotdenominator stars for which rotation periods might
+plausibly have been detected.  For set $\mathcal{B}$, we then
+additionally required that the Lomb-Scargle and Stellingwerf
 phase-dispersion methods yielded identical periods to within 10\%.
@@
 The comparison to the Pleiades and Praesepe is made using the rotation
-periods and dereddened colors from Table~5 of \citet{curtis_rup147_2020}, which
-were based data from \citet{rebull_rotation_2016a} and
-\citet{douglas_k2_2019} respectively.
+periods and dereddened colors from Table~5 of
+\citet{curtis_rup147_2020}, which were based data from
+\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
+respectively.
 
@@
 
-{\bf We explore this quantitatively by fitting
-interpolated models}[?].
-The resulting gyrochronology age we find for the core is XXX.
-For the halo, the claimed age from gyrochronology is YYY.
+{\bf We explore this quantitatively by fitting interpolated
+models}[?].  The resulting gyrochronology age we find for the core is
+XXX.  For the halo, the claimed age from gyrochronology is YYY.
 Applying the same procedure to the field star comparison sample (see
-Appendix~\ref{app:compstar}),
-we get an age of ZZZ+/-AAA.
+Appendix~\ref{app:compstar}), we get an age of ZZZ+/-AAA.
 
@@
   \caption{ 
-  {\bf Members of NGC\,2516 exist hundreds of parsecs from the core,
+  {\bf Members of \cn\ exist hundreds of parsecs from the core,
   and up to $\approx$5\kms\ from the core in tangential velocity.}
-  Halo members
-  that meet both kinematic and rotational membership criteria are shown in
-  blue.  Halo members for which rotation periods should have been detectable,
-  but were not detected, are shown in orange.  Projections are shown in galactic coordinates and
-  (projection-corrected) 2D-tangential velocity space.  The histogram
-  on the right bins versus 3D separation from the core and
-  2D velocity difference. Bins have width 25\,pc
-  in position, and 1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties
-  in the fraction are taken as the quadrature sum of the rel unc of
-  the numerator and denominator (FIXME)}
+  Halo members that meet both kinematic and rotational membership
+  criteria (``Subset~$\mathcal{A}$'') are shown in blue.  Halo members for which rotation periods
+  should have been detectable, but were not detected, are shown in
+  orange.  
+  The plotted stars are restricted to $0.5<(Bp-Rp)_0<1.2$.
+  The {\it left panels} show galactic coordinates and
+  2D-tangential velocity space, after
+  correcting for projection effects
+  (Appendix~\ref{app:vproj}).
+  The histogram on the right bins versus 3D separation from the core
+  and 2D velocity difference.  Bins have width 25\,pc in position, and
+  1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties in the fraction
+  are taken as the quadrature sum of the rel unc of the numerator and
+  denominator (FIXME)}
   \label{fig:physical_x_rotn}
@@
 
+% some wide separation pairs and dist_pc:
+% 5263077805266480000  5506439929389838592  421.083675
+% 5506623616550497792  5273740456715689088  427.053947
+% 5506439929389838592  5213536216500903680  427.557268
+% 5273740456715689088  5488984598182200704  428.952934
+% 5506439929389838592  5273740456715689088  438.906581
+
 To quantify this, we compute the separation of the positions and 2D
@@
 Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
-about 200\,pc in physical space from the cluster core.  This
-corresponds to total a length of $\approx$500\,pc, depending on which
-members of the halo are chosen as the ``tips'' on either end.  This
-agrees broadly speaking with the structure of the halo reported by
-\citetalias{kounkel_untangling_2019}.
+about 200\,pc in physical space from the cluster core.  The most
+widely separated F2V-K2V stars with rotation periods (blue points in
+Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
+The total length of the structure is therefore 400--500\,pc, depending
+on which members of the halo are chosen as the ``tips'' on either end.
+Broadly speaking, this agrees with the structure of the halo reported
+by \citetalias{kounkel_untangling_2019}.
 
@@
 5\kms.  \citet{meingast_2021} by comparison required a physically
-motivated cut in tangential velocity space of 1.5\kms.  
-Our results show that at the expense of higher field star
-contamination rates, bonafide members can be identified even out to
-higher velocity separations.
-{\bf todo: anything else to mention? e.g., the actual numbers?}
+motivated cut in tangential velocity space of 1.5\kms.  Our results
+show that at the expense of higher field star contamination rates,
+bonafide members can be identified even out to higher velocity
+separations.  {\bf todo: anything else to mention? e.g., the actual
+numbers?}
 
@@
   \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
-  Equivalent width of the 6708\,\AA\ doublet is shown
-  versus extinction-corrected color for all kinematic \cn\ members
-  with Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.
-  The GALAH spectra comprise slightly over half of the halo stars,
-  due to the non-targeted selection function of that survey.
-  The gray points are Keck-HIRES measurements of stars with detected
-  transiting planets from the Kepler mission
-  \citep{berger_identifying_2018}.
-  Points with ${\rm EW}\approx0\,$m\AA\ are non-detections.
+  Equivalent width of the 6708\,\AA\ doublet is shown versus
+  extinction-corrected color for all kinematic \cn\ members with
+  Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
+  spectra comprise slightly over half of the halo stars, due to the
+  non-targeted selection function of that survey.  The gray points are
+  Keck-HIRES measurements of stars with detected transiting planets
+  from the Kepler mission \citep{berger_identifying_2018}.  Points
+  with ${\rm EW}\approx0\,$m\AA\ are non-detections.
   \label{fig:lithiumcorehalo}
@@
 	\caption{ {\bf Lithium and rotation in NGC\,2516.}
-		Points show all kinematic \cn\ members with TESS rotation
-		periods in the Gaia-ESO or GALAH surveys, and are colored by
-		equivalent width.
-    Points with reported equivalent widths below 20\,m\AA\ (including
-    non-detections) are shown with gray crosses.
-    The lithium-rotation correlation for the K-dwarfs
-    has been observed in other young clusters, and is discussed in
+    Points show all kinematic \cn\ members with TESS rotation periods
+    in the Gaia-ESO or GALAH surveys, and are colored by equivalent
+    width.  Points with reported equivalent widths below 20\,m\AA\
+    (including non-detections) are shown with gray crosses.  The
+    lithium-rotation correlation for the K-dwarfs has been observed in
+    other young clusters, and is discussed in
     Section~\ref{subsec:lithium}.
@@
 The GALAH DR3 target selection is discussed in depth by
-\citet{buder_galah_2020}.  The relevant aspects for our analysis
-are that the ``main'' survey targeted $12<V<14$ stars at
+\citet{buder_galah_2020}.  The relevant aspects for our analysis are
+that the ``main'' survey targeted $12<V<14$ stars at
 $\delta<+10^\circ$, provided the stars were at least ten degrees from
@@
 e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
-uncertainties on the EWs, we repeated the procedure twenty times,
-but added noise to the spectra, drawn from a normal distribution with
-a scale set by the standard deviation of the continuum.
-The reported uncertainties are then drawn from the $84^{\rm th}$ and
-$16^{\rm th}$ percentiles of the resulting EW distribution.
-We verified the overall scale of our results by comparing the 21
-stars that overlapped between GALAH and Gaia-ESO.
-17 of the 21 measurements agreed within 1-$\sigma$.
-The remaining four were stars with $T_{\rm eff}>6200\,{\rm K}$, for
-which \citet{randich_gaiaeso_2018} reported Li EW detections
-ranging from $50$ to $75\,$m\AA, and for which we found 
-either null results or only slightly significant
+uncertainties on the EWs, we repeated the procedure twenty times, but
+added noise to the spectra, drawn from a normal distribution with a
+scale set by the standard deviation of the continuum.  The reported
+uncertainties are then drawn from the $84^{\rm th}$ and $16^{\rm th}$
+percentiles of the resulting EW distribution.  We verified the overall
+scale of our results by comparing the 21 stars that overlapped between
+GALAH and Gaia-ESO.  17 of the 21 measurements agreed within
+1-$\sigma$.  The remaining four were stars with $T_{\rm
+eff}>6200\,{\rm K}$, for which \citet{randich_gaiaeso_2018} reported
+Li EW detections ranging from $50$ to $75\,$m\AA, and for which we
+found either null results or only slightly significant
 ($\approx$30m\AA) detections.
@@
 span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
-cutoff of the survey.  The overall depletion of stars redder than
-$(Bp-Rp)_0\gtrsim1.4$, {\it i.e.}, later than K4.5V (0.71\,$M_\odot$),
-is visible relative to the earlier K dwarfs.  For comparison, we have
-also plotted the Li EWs measured by \citet{berger_identifying_2018}
-for planet-hosting stars in the Kepler field (we interpolated from
-their effective temperatures using the \citet{pecaut_mamajek_2013}
+cutoff of the survey.  The overall increase of Li EW from
+$0<(Bp-Rp)_0<1.0$ is thought to be driven by the temperatures in the
+stellar chromospheres: hotter stars fully ionize Li out of its ground
+state \citep[{\it e.g.},Figure~4 of][]{soderblom_evolution_1993}.  The
+depletion of stars redder than $(Bp-Rp)_0\gtrsim1.4$, {\it i.e.},
+later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
+K dwarfs.  For comparison, we have also plotted the Li EWs measured by
+\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
+field (we interpolated from their effective temperatures using the
+\citet{pecaut_mamajek_2013}
 table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version 2019.3.22.}.
-The field star distribution peaks at late F-type stars.
-The majority of kinematically selected stars in the core and the halo
-show lithium equivalent widths substantially in excess of these field
-stars.
+version 2019.3.22.}.  The field star distribution peaks at late F-type
+stars.  The majority of kinematically selected stars in the core and
+the halo show lithium equivalent widths substantially in excess of
+these field stars.
 
 We can go one step further, and match our sample of kinematically
-selected stars with spectra against the TESS rotators.
-The result is shown in Figure~\ref{fig:lithiumrot}
-At fixed stellar mass
-(and age), the rapid rotators tend to show elevated lithium equivalent
-widths. This effect is mostly apparent in the K dwarfs.  Similar
-trends in the Pleiades were noted nearly three decades ago by
-\citet{soderblom_evolution_1993}.  More recent studies have found
-comparable results in the Pleiades, the Psc-Eri stream, and M\,35
+selected stars with spectra against the TESS rotators.  The result is
+shown in Figure~\ref{fig:lithiumrot}. At fixed stellar mass, the rapid
+rotators have lithium equivalent widths an order of magnitude larger
+than the slow rotators.  This effect is mostly apparent in the K
+dwarfs.  Similar trends were noted in the Pleiades over three decades
+ago \citep{butler_pleiades_1987}, and are most likely caused by
+intrinsic differences in lithium abundance
+\citep{soderblom_evolution_1993}.  Alternative explanations, such as
+differences in line-formation conditions ({\it e.g.,} chromospheric
+temperatures, microturbulent velocities, or the presence of starspots)
+are incompatible with the available data.  The lithium-rotation
+correlation is discussed further below
+(Section~\ref{discussion:lithium}).
+
+\section{Discussion}
+\label{sec:discussion}
+
+%\subsection{Contamination fraction across the halo}
+%\subsection{How did the halo form?}
+%\subsection{Rotation in NGC\,2516}
+%\subsection{The lithium-rotation correlation}
+%\subsection{Age of NGC\,2516}
+
+\subsection{Contamination fraction across the halo}
+
+\subsection{How did the halo form?}
+\label{subsec:origin}
+% Should we be calling the halo a tail?
+
+Differential rotation in the galaxy is clearly important. Meingast+21
+established this; in ten different cluster halos, nine showed leading
+and trailing arms oriented in line with the differential rotation of
+the Galaxy.
+(See e.g., their Figure 13).
+
+Another way to frame this question: should we call these stars a tidal
+tail?
+% TODO: read Gagne+20 on this...
+The classical explanation of such a structure (e.g., Krumholz+19) is
+that stars escape out of L1 and L2, and then form leading and lagging
+arms due to differential rotation in the galaxy.
+The leading and lagging arms of NGC\,2516's halo qualitatively
+match this picture, in that they are oriented roughly toward and away
+the direction of galactic rotation (CITE, Meingast21, link to their
+website).
+
+A second explanation though invokes
+the idea that the cluster formed in a larger and lower-density star
+formation complex, and the stars we see in the halo did not in fact
+form the same ``clump'' as those in the cluster core.
+
+Consider for instance the Orion\,A cloud.  \citet{grosschedl_3d_2018} showed
+using YSOs that formed in the past few megayears as tracers that Orion\,A is
+best described with two components:
+a dense ``head'', and a lower density 75\,pc long ``tail''.
+While this filament currently totals 90\,pc in length, its future evolutionary state could very well resemble something like \cn.
+A direct numerical exploration would be necessary to be sure {\bf FIXME}.
+
+CITE ALSO: ZUCKER 2017 BONES CATALOG.
+
+CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
+THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
+LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
+
+Similarly Zari et al 2019 studied the entire Orion complex, and found that
+different regions of the complex (one toward the Belt, one toward 25\,Ori)
+began forming stars roughly simultaneously. No evidence was found of it
+propagating from these initial sites.
+
+ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
+
+cometary-like cloud,
+oriented toward the Galactic plane, with two distinct components:
+a denser and enhanced star-forming (bent) Head, and a
+lower density and star-formation quieter 75 pc long Tail.
+
+% ``The stars in these halos could have escaped from the cores of the
+% cluster, or they could be remnants of a larger and lower-density star
+% formation complex that has since dispersed.''
+Another possibility is {\it triggered star formation}.
+
+One relevant connection may be to the ``Mamajek-2'' stellar group
+(see Jilinski+2009).
+
+\subsection{Rotation in NGC\,2516}
+
+Gotta cite Fritzewski+2020.  (Maybe even do the Blanco1, Pleiades,
+NGC2516, PscEri concatenation?)
+
+  \paragraph{Slow sequence and fast sequence}
+  %Why is there a slow and fast sequence?
+  Slow: Skumanich spindown, Weber and Davis magnetic braking.
+  This doesn't explain everything.
+  Core envelope decoupling handwaves away the fast rotators? (Irwin,
+  Aigrain, Monitor).
+  Another fast rotator explanation: Naoz's merging of giant planets
+  (could explain some fraction of the fast rotators -- not all of them).
+
+  \paragraph{Stalling}
+  Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
+  spindown is not monotonic in time.
+  We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
+  much change.
+
+  \paragraph{Are the ``very slow rotators'' bonafide members?}
+  Unsure. They are not isotropically distributed around the
+  cluster... so either triggered star formation (CITE, CITE), or they're
+  actually field stars.
+
+\subsection{The lithium-rotation correlation}
+\label{discussion:lithium}
+
+An important point regarding the lithium-rotation trend is that it has
+been shown to carry over beyond just equivalent widths, and to the
+actual abundance of lithium in the photosphere, A(Li).
+\citet{soderblom_evolution_1993} calculated the relevant LTE curves of
+growth for the Li\,6708\,\AA\ doublet, which can in principle be used to
+convert our measured equivalent widths to abundances.  Non-LTE
+corrections affect the abundances by $<0.5\,{\rm dex}$ and do not
+siginificantly affect the global relation between Li abundance and
+stellar temperature or rotation
+\citep{carlsson_1994,lind_departures_2009}.
+
+More recent studies of lithium and rotation since 
+\citet{soderblom_evolution_1993} have expanded these results in the
+Pleiades, the Psc-Eri stream, and M\,35, among other clusters
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
+In \cn\ specifically, \citet{jeffries_rotation_1998} observed a
+similar trend that that seen in Figure~\ref{fig:lithiumrot} when
+analyzing 24 stars in the core of \cn.  Our concatenation of the
+\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
+spectra represent a significant expanasion in volume and color range
+since these earlier studies.
 
@@
 
-% TODO: this paragraph more or less parrots the processes described by
-% Bouvier+20. are there others? is there a better way to describe
-% these?
-
-An important point regarding the lithium-rotation trend is that it has
-been shown to carry over beyond just equivalent widths, and to the
-actual abundance of lithium in the photosphere, A(Li).  A priori, one
-might worry that the trend is tautological: perhaps some feature the
-stellar equatorial velocity, or the microturbulent velocity at the
-stellar surface, are larger for faster rotators and this could affect
-the formation conditions of the lithium doublet.
-%FIXME improve these words
-Soderblom+1993 calculated the relevant LTE curves of growth for the
-lithium line, and argued that this could not be the case.
-Since then, LindAsplund+2009 have calculated the relevant non-LTE
-correction, and has shown that it affects the abundances by
-$<0.5\,{\rm dex}$.
-The order of magnitude of the rotation-abundance trend is 2-3\,dex.
-
-%     * OK, then to convert to an ABUNDANCE, follow Bouvier+2018. In other words,
-%       just interpolate against the LTE curve of growth tables from
-%       Soderblom+1993.
-% 
-%     * Nb. the Non-LTE correction has been shown to be <0.5 dex (but
-%       >0.1 dex, for sun-like stars with Li abundances above 2 dex,
-%       >which is the regime we're interested in for Pleaides and NGC
-%       >2516).
-%       CITE: Lind, Asplund, + 2009
-%       https://ui.adsabs.harvard.edu/abs/2009A%26A...503..541L/abstract
-
+What causes the correlation between lithium abundance and rotation at
+fixed stellar mass?
+{\bf FIXME: improve this paragraph}
+%FIXME IMPROVE THIS PARAGRAPH.
 Processes both internal and external to the star have been suggested to explain
@@
 
-\section{Discussion}
-\label{sec:discussion}
-
-\subsection{How did the halo form?}
-\label{subsec:origin}
-% Should we be calling the halo a tail?
-
-Differential rotation in the galaxy is clearly important. Meingast+21
-established this; in ten different cluster halos, nine showed leading
-and trailing arms oriented in line with the differential rotation of
-the Galaxy.
-(See e.g., their Figure 13).
-
-Another way to frame this question: should we call these stars a tidal
-tail?
-% TODO: read Gagne+20 on this...
-The classical explanation of such a structure (e.g., Krumholz+19) is
-that stars escape out of L1 and L2, and then form leading and lagging
-arms due to differential rotation in the galaxy.
-The leading and lagging arms of NGC\,2516's halo qualitatively
-match this picture, in that they are oriented roughly toward and away
-the direction of galactic rotation (CITE, Meingast21, link to their
-website).
-
-A second explanation though invokes
-the idea that the cluster formed in a larger and lower-density star
-formation complex, and the stars we see in the halo did not in fact
-form the same ``clump'' as those in the cluster core.
-
-Consider for instance the Orion\,A cloud.  \citet{grosschedl_3d_2018} showed
-using YSOs that formed in the past few megayears as tracers that Orion\,A is
-best described with two components:
-a dense ``head'', and a lower density 75\,pc long ``tail''.
-While this filament currently totals 90\,pc in length, its future evolutionary state could very well resemble something like \cn.
-A direct numerical exploration would be necessary to be sure {\bf FIXME}.
-
-CITE ALSO: ZUCKER 2017 BONES CATALOG.
-
-CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
-THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
-LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
-
-Similarly Zari et al 2019 studied the entire Orion complex, and found that
-different regions of the complex (one toward the Belt, one toward 25\,Ori)
-began forming stars roughly simultaneously. No evidence was found of it
-propagating from these initial sites.
-
-ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
-
-cometary-like cloud,
-oriented toward the Galactic plane, with two distinct components:
-a denser and enhanced star-forming (bent) Head, and a
-lower density and star-formation quieter 75 pc long Tail.
-
-% ``The stars in these halos could have escaped from the cores of the
-% cluster, or they could be remnants of a larger and lower-density star
-% formation complex that has since dispersed.''
-Another possibility is {\it triggered star formation}.
-
-One relevant connection may be to the ``Mamajek-2'' stellar group
-(see Jilinski+2009).
-
-\subsection{What is the contamination fraction in the halo? Does it change vs. location?}
-
-\subsection{Are the ``very slow rotators'' bonafide members?}
-Unsure. They are not isotropically distributed around the
-cluster... so either triggered star formation (CITE, CITE), or they're
-actually field stars.
-
-% \subsection{Mass differences between center and outer reaches?}
-
-\subsection{Fast and blue rotators: are we going faster than other clusters?}
-
-\subsection{Absolute age of NGC 2516}
+\subsection{Age of NGC\,2516}
 Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
@@
 
-\subsection{Ways of doing this at different ages}
-Some final discussion is warranted on the applicability of our
-approach more generally.  Given a kinematically
-identified population, this study uses photometry (Gaia HR diagrams;
-TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
-confirm youth in the stars.  Other approaches are also possible.
-
-One can use photometric criteria (e.g., Gaia HR diagram of
-everything within some distance; Zari+18), or spectroscopic
-criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
-Zerjal 2019, 2020, Zhou's work).  
-It might even be possible to start using chemical abundance signatures
-(CITE, GALAHDR3?).
-
-These paths might in certain cases succeed in identifying more
-complete samples of dispersed members of young populations, since they
-do not require kinematic proximity.
+% \subsection{Ways of doing this at different ages}
+% Some final discussion is warranted on the applicability of our
+% approach more generally.  Given a kinematically
+% identified population, this study uses photometry (Gaia HR diagrams;
+% TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
+% confirm youth in the stars.  Other approaches are also possible.
+% 
+% One can use photometric criteria (e.g., Gaia HR diagram of
+% everything within some distance; Zari+18), or spectroscopic
+% criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
+% Zerjal 2019, 2020, Zhou's work).  
+% It might even be possible to start using chemical abundance signatures
+% (CITE, GALAHDR3?).
+% 
+% These paths might in certain cases succeed in identifying more
+% complete samples of dispersed members of young populations, since they
+% do not require kinematic proximity.
 
@@
 
-The combination of Gaia, TESS, and the GALAH and Gaia-ESO
-spectroscopic surveys has clarified a few 
-things about \cn.
+The combination of astrometry from Gaia, photometry from TESS, and
+spectroscopy from GALAH and Gaia-ESO has clarified a few things about
+the structure of \cn.
 \begin{itemize}
-	\item {\it Overdensities observed by Gaia in five-dimensional (3-D position + 2-D velocity)
-		phase space
-		revealed a halo of stars spanning 500 parsecs in the plane of the galaxy.}
-	The earlier reference to the full structure that we can find is by CITE KOUNKEL.
-	The halo is more precisely described as a leading and trailing tail
-	(Figure~\ref{fig:XYZ}), with the leading end on the interior of the
-	cluster's
-	galactic orbit.
-	%
-	\item {\it Isochronal, rotational, and lithium age-dating show that the halo is real and coeval with the core of \cn.}
-    The Gaia EDR3 photometry and astrometry show a main-sequence turnoff consistent with
-    an isochronal age of 150\,Myr for both the core and halo.
-    The faintest M dwarfs in the core and halo are also elevated relative to a field star comparison sample, but using
-    this to isochronally measure the age is challenging due to uncertainties in M dwarf atmospheric opacities.
-    Gyrochronally, FIXME FIXME FIXME
-    %
-    \item {\it The identified halo extends to $\approx$250\,pc in separate from the cluster center, and out to $\approx$5\kms\ in velocity separation.}
-    Figure~\ref{fig:physical_x_rotn} shows the overlap between kinematic
-    and rotational cluster members. 
-    %
-    \item {\it The field star contamination rate increases at larger physical and 
-    velocity separations from the cluster core.}
-	Figure~\ref{fig:physical_x_rotn}
-
-    The stars further from the c
-    As a reminder, a star with a 1\kms\ velocity difference moves 1\,pc per Myr.
-
-	More specifically, ``two thirds of kinematically selected halo members out to 250 pc from the cluster center have rotation periods consistent with a gyrochronological age of 150 Myr''.
-
-	\item {\it Differential rotation in the Milky Way likely plays an important role in shaping the halos of nearby open clusters.}
-	The Milky Way's differential rotation rate has a comparable
-	order of magnitude ($\approx$-0.2\,deg\,Myr$^{-1}$\,kpc$^{-1}$) to that measured across the halo \citep{meingast_2021}.
-	It therefore likely plays an important role in understanding the halo's formation.
-
+  %
+  \item {\it Over-densities observed in the Gaia 3-D positions and 2-D
+    velocities revealed a halo of stars spanning $\approx$500 parsecs
+    in the plane of the galaxy.} The earliest reference to this halo
+    that we can find is by \citet{kounkel_untangling_2019}.  The halo
+    is more precisely described as a leading and trailing tail
+    (Figure~\ref{fig:XYZ}), with the leading edge angled toward the
+    galactic center, relative to the cluster's orbit in the Galaxy.
+    This is consistent with the direction and amplitude of the Milky
+    Way's differential rotation
+    ($\approx$-0.2\,deg\,Myr$^{-1}$\,kpc$^{-1}$), and similar
+    halos/tails have been observed around $\approx$10 other nearby
+    open clusters \citep{meingast_2021}.
+	%
+  \item {\it Isochronal, rotational, and lithium age-dating show that
+    the halo of \cn\ is coeval with its core.} In other words, all
+    the data are consistent with the halo being real.  The Gaia EDR3
+    photometry and astrometry show a main-sequence turnoff consistent
+    with an isochronal age of 150\,Myr for both the core and halo
+    (Figure~\ref{fig:hr}).  The faintest M dwarfs in the core and halo
+    are also brighter than comparison field stars of the same color.
+    Gyrochronally, stars in \cn\ spanning $0.5<(Bp-Rp)_0<1.2$
+    (spectral types F2V-K3V) overlap with the slow sequence of the
+    Pleaides (Figure~\ref{fig:rot}).  At slightly redder
+    colors, from $1.2<(Bp-Rp)_0<1.7$ (spectral types K3V-K6V,
+    $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
+    dispersion in rotation period.  The upper envelope of the rotation
+    period distribution at these redder colors extends to
+    longer rotation periods than in the Pleiades (11 {\it
+    vs.}\ 8 day rotation periods).
+    This also supports our {\bf age determination of $XXX$\,Myr for
+    the cluster}.
+  %
+  \item {\it Bonafide members exist out to $\approx$250\,pc in
+    separation from the cluster center, and out to $\approx$5\kms\ in
+    velocity separation.} Figure~\ref{fig:physical_x_rotn} shows the
+    overlap between kinematic and rotational cluster members used to
+    make this assessment.
+  %
+  \item {\it The field star contamination rate increases at larger
+    physical and velocity separations from the cluster core.} The
+    right panel of Figure~\ref{fig:physical_x_rotn} quantifies the
+    statement: $\approx$75\% of the ``inner-most'' members show
+    rotation periods consistent with the age of the cluster.  This
+    fraction decreases to $\approx$60\% at physical separations of
+    $\approx$200\,pc.  Given the number density ratios of cluster to
+    field stars at these wide separations ($\approx 10^{-3}$,
+    \citealt{meingast_2021}), the implied contamination fractions are
+    not as bad as one might fear.
 \end{itemize}
+
+Broadly speaking, the existence of the halo is perhaps not very
+surprising.  A star with a 1\kms\ velocity difference from the cluster
+average will move away from the center at a rate of 1\,pc\,Myr$^{-1}$.
+For a $\sim$100\,Myr old cluster, a halo with characteristic size
+$\sim$100\,pc is therefore expected, given that the typical velocity
+dispersions of open clusters are of the order of kilometers per
+second.
+
+On the other hand, our ability to reliably identify members of these
+halos is rather surprising.  These stars have
+remained hidden in the background of the Galaxy throughout centuries
+of modern astronomy.  The consequences that these halo stars will have
+in expanding our understanding of planetary and stellar origins are
+only just beginning to be understood.
 

### 2021-03-25 earhart/plotting.py

@@ -2431,7 +2431,9 @@ def plot_physical_X_rotation(outdir, basedata=None, show1937=0,
     """
 
     nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
-    rot_df, lc_df = get_autorotation_dataframe(runid='NGC_2516', returnbase=True)
+    rot_df, lc_df = get_autorotation_dataframe(
+        runid='NGC_2516', returnbase=True, cleaning='defaultcleaning'
+    )
     med_df, _ = _get_median_ngc2516_core_params(core_df, basedata)
 
     from earhart.physicalpositions import append_physicalpositions

### 2021-03-29 cluster_paper/ms.tex
update ms

@@
   Halo members that meet both kinematic and rotational membership
-  criteria (``Subset~$\mathcal{A}$'') are shown in blue.  Halo members for which rotation periods
+  criteria (Subset~$\mathcal{A}$) are shown in blue.  Halo members for which rotation periods
   should have been detectable, but were not detected, are shown in
@@
   This doesn't explain everything.
+  Originally described as ``interface / convective'' dynamo for each.
+  Now disfavored (but why?).
   Core envelope decoupling handwaves away the fast rotators? (Irwin,
@@
   (could explain some fraction of the fast rotators -- not all of them).
+  Tim Brown, ``metastable dynamo''?
+
+  Perhaps most compelling: some kind of binary related explanation.
+  Gillen+20 in Blanco-1 found most of them to be photometric binaries.
+  Simonian+19 in the Kepler field found most rapid rotators to have
+  photometric excesses.
 

### 2021-03-29 cluster_paper/ms.tex
update ms with phot/astrom binarity

@@
 		}
-	   %
-		%\subfloat{
-		%	\includegraphics[width=0.6\textwidth]{f3b.pdf}
-		%}
+
+        \vspace{-0.1cm}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f3c.pdf}
+			\includegraphics[width=0.49\textwidth]{f3d.pdf}
+		}
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf The core and halo of \cn\ in the space of rotation
-    period and dereddened Gaia color.}
-    Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
-    undergo successive degrees of automated cleaning, described in Section~\ref{subsubsec:cluster}.
-    The Pleiades \citep[120\,Myr;][]{rebull_rotation_2016a} and
-    Praesepe \citep[650\,Myr;][]{douglas_poking_2017} are shown for
-    reference.
+  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
+    {\it Top row}: \cn\ split into its core and halo.  Subset
+    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right columns}) undergo
+    successive degrees of automated cleaning, described in
+    Section~\ref{subsubsec:cluster}.  The Pleiades
+    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
+    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
     Stars in the core and the halo of NGC\,2516 appear
@@
     There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$,
-    in which some NGC\,2516 members appear elevated with respect to
-    the Pleiades, suggesting a slightly older age.
+    in which some NGC\,2516 members appear slightly elevated with respect to
+    the Pleiades, suggesting an older age.
+    {\it Bottom row}: \cn\ colored by binarity indicators.
+    Sources with $\mathrm{RUWE}>1.2$ appear in red (astrometric
+    binaries).
+    Overplotted in orange are sources at least 0.3\,mag above an
+    empirically fit cluster isochrone (photometric binaries).
 		\label{fig:rot}
@@
 
-Figure~\ref{fig:rot} shows the resulting rotation periods, following
-three different sets of cleaning criteria.  The first and second sets
-of criteria were entirely automated, and yielded subsets we denote
-$\mathcal{A}$ and $\mathcal{B}$.  For both subsets, we considered
-light curves for which the peak Lomb-Scargle periodogram period was
-below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
-for which no equal-brightness or greater companions were within the
-aperture.  Beyond requiring that the target star be the brightest star
-in the aperture, we also required that at most one companion with
-brightness exceeding one tenth of the target star's brightness be
-present in the aperture according to the Gaia DR2 source catalog.  For
-set $\mathcal{A}$, this yielded \nautorotnumerator light curves, out
-of \nautorotdenominator stars for which rotation periods might
-plausibly have been detected.  For set $\mathcal{B}$, we then
+The top row of Figure~\ref{fig:rot} shows the resulting rotation
+periods, following three different sets of cleaning criteria.  The
+first and second sets of criteria were entirely automated, and yielded
+subsets we denote $\mathcal{A}$ and $\mathcal{B}$.  For both subsets,
+we considered light curves for which the peak Lomb-Scargle periodogram
+period was below 15 days, the normalized Lomb-Scargle power exceeded
+0.08, and for which no equal-brightness or greater companions were
+within the aperture.  Beyond requiring that the target star be the
+brightest star in the aperture, we also required that at most one
+companion with brightness exceeding one tenth of the target star's
+brightness be present in the aperture according to the Gaia DR2 source
+catalog.  For set $\mathcal{A}$, this yielded \nautorotnumerator light
+curves, out of \nautorotdenominator stars for which rotation periods
+might plausibly have been detected.  For set $\mathcal{B}$, we 
 additionally required that the Lomb-Scargle and Stellingwerf
@@
 
-{\bf Curtis TODO: set $\mathcal{C}$ was created through a manual
-curation process.}
-
 The comparison to the Pleiades and Praesepe is made using the rotation
 periods and dereddened colors from Table~5 of
-\citet{curtis_rup147_2020}, which were based data from
+\citet{curtis_rup147_2020}, which used data from
 \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
 respectively.
-
 A few conclusions can be drawn by comparing the \cn\ rotation periods
@@
 with \cn\ being slightly older than the Pleiades.
-
 {\bf We explore this quantitatively by fitting interpolated
@@
 Appendix~\ref{app:compstar}), we get an age of ZZZ+/-AAA.
+
+The bottom row of Figure~\ref{fig:rot} shows the same data, with the
+points colored by indicators of binarity.
+To assess astrometric binarity, we used the renormalized unit weight
+error (RUWE) from Gaia EDR3.
+For plotting purposes, we labeled anything with RUWE exceeding 1.2 as
+an astrometric binary (11\% of the overall cluster
+sample).
+A cautionary point is that poor astrometric fits can also be caused by
+photometric variability.
+To assess photometric binarity, we fitted a spline to
+Figure~\ref{fig:hr} by hand, and labelled any points brighter than
+0.3\,mag above the cluster sequence as photometric binaries.
+This included 20\% of the overall cluster sample.
+Many of the resulting candidate binaries, though not all of them,
+appear underneath the ``slow sequence''.
+We speculate on possible explanations in Section~\ref{disc:rotn}.
 
@@
 \subsection{Rotation in NGC\,2516}
+\label{disc:rotn}
 

### 2021-03-29 cluster_paper/collect.sh

@@ -7,6 +7,9 @@ cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2c.pdf
 
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear_defaultcleaning.pdf f3a.pdf
 cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear_periodogram_match.pdf f3b.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_linear_defaultcleaning.pdf f3c.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_linear_periodogram_match.pdf f3d.pdf
+
 
 #cp ../results/skypositions/skypositions_x_rotn.pdf f4.pdf
 cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_galactic.pdf f4.pdf

### 2021-03-29 earhart/plotting.py

@@ -379,7 +379,7 @@ def plot_ruwe_vs_apparentmag(outdir, basedata='fullfaint', smallylim=False):
 
     if not basedata == 'fullfaint_edr3':
         raise NotImplementedError('only EDR3 has ruwe built in')
-    nbhd_df, cg18_df, kc19_df, trgt_df = get_gaia_basedata(basedata)
+    nbhd_df, cg18_df, kc19_df, full_df, trgt_df = get_gaia_basedata(basedata)
 
     plt.close('all')
 

### 2021-04-01 cluster_paper/ms.tex
update ms

@@
   +20.8\}\,{\rm pc}$.  The gray line traces the past and future
-  cluster orbit.  In different projections, the line spans different
+  cluster orbit.
+  In different projections, the gray line spans different
   time intervals: the longest time window is visible in the $(Z,X)$
@@
   \caption{ {\bf HR diagrams of NGC\,2516.}
-    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
-    has been applied.  The core (black) shows a main-sequence and
-    turnoff consistent with stars of a fixed age and metallicity.  The
-    halo (blue) is similar, though the binary sequence is less
-    defined.  At fixed color, the faintest M dwarfs in the core and
-    halo are brighter than in the field star comparison sample (gray),
-    consistent with these stars not having yet reached the
-    main-sequence.
-    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
-    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
-    The arrow represents the average reddening correction applied
-    to the models.
-     The models diverge
-    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
-    $Bp-Rp\approx$2.2).
+    {\it Left}: Photometric data from Gaia EDR3; no reddening
+    correction has been applied.  The core (black) shows a
+    main-sequence and turnoff consistent with stars of a fixed age and
+    metallicity.  The halo (blue) is similar, though the binary
+    sequence is less defined.  The faintest M dwarfs in the core and
+    halo are brighter than field stars (gray) with the same color,
+    consistent with the cluster M dwarfs not having yet reached the
+    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
+    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
+    8.5\}$.  The arrow represents the average reddening correction
+    applied to the models.  The opacity of the M-dwarf model
+    atmospheres is less well-calibrated for $M_\star \lesssim
+    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
+    are shown with dashed lines.
     \label{fig:hr}
@@
 The first check on the plausibility of the candidate cluster members
-being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.  Comparable analyses
-have already been performed by \citetalias{cantatgaudin_gaia_2018},
+being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.
+Comparable analyses have already been performed by
+\citetalias{cantatgaudin_gaia_2018},
 \citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.
@@
 with a fixed age and metallicity, and varying mass.  The halo members
-mostly show a consistent main-sequence that has somewhat greater
-scatter.  One possible explanation for the additional scatter could be
-that the halo is more contaminated by field stars.  For instance,
-$\approx5$ evolved halo stars are visible in the field-star portion of
-the red giant branch (RGB).  Given the age of the system, these stars
-must be field interlopers.
-
-A separate possible explanation for scatter in the halo's HR diagram
+show a consistent sequence, though with greater scatter.  One possible
+explanation for the additional scatter could be that the halo is more
+contaminated by field stars.  For instance, $\approx5$ halo
+stars are present in portions of the red giant branch
+(RGB) that imply stellar ages inconsistent with that of the cluster.
+These stars must be field interlopers.
+
+A second possible explanation for scatter in the halo's HR diagram
 could be differential reddening across different sightlines.  The
@@
 Overall, the data and models agree for masses above approximately
-0.45\,$M_\odot$.  However below this mass, the data and models diverge
-at colors redder than $Bp-Rp\approx2.2$, in the sense that the model
+0.45\,$M_\odot$.  Below this mass, the data and models diverge at
+colors redder than $Bp-Rp\approx2.2$, in the sense that the model
 isochrones have lower luminosities and bluer colors than the
-observations.  The MIST isochrones showed a comparable disagreement
-\citep{choi_mesa_2016}.  Explanations invoking both starspots and
-incomplete molecular line lists have been proposed \citep[{\it
+observations.\footnote{The 100 Myr isochrone does overlap well with the
+lower main sequence, but it fails to fit the upper main sequence and
+red giants.}  The MIST isochrones showed a comparable disagreement
+\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
+between the models and observations have include starspots and
+incomplete molecular line lists\citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
-We adopt the PARSEC isochrones since they diverge at slightly lower
-mass than the MIST models, due to the temperature-opacity calibration
-implemented by \citet{chen_improving_2014}.  For purposes of
-age-dating the cluster, we prefer to focus on the main-sequence
-turn-off (MSTO), since this is where the models have maximal predictive
-power.
+Ultimately, we adopted the PARSEC isochrones because
+\citet{chen_improving_2014} implemented an empirical
+temperature-opacity calibration, which leads the PARSEC isochrones to
+diverge at slightly lower mass than the MIST models.  Regardless, for
+purposes of age-dating the cluster we focus on the main-sequence
+turn-off (MSTO), since this is where the models have maximal
+predictive power.
 
@@
 165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
-PARSEC, MIST, or SYCLIST).  The statistical uncertainties ($\approx
+PARSEC, MIST, or SYCLIST).  Their statistical uncertainties ($\approx
 25$\,Myr) were comparable to the systematic uncertainties.
@@
 Our primary goal is to determine whether the ages of the core and halo
-are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
-they are: stars past the turnoff in both the halo and core are on a
-consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
-the precision with which the claim can be made.  The MSTO stars are
-consistent with the 178 and 316 Myr models, but are bluer than the 100
-Myr model.  The RGB stars are consistent only with the 178 Myr
-($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
-the width of the MSTO can be attributed to binary stars, we are most
-interested in the blue edge. This appears most compatible with the 178
-Myr model.  These results are consistent with the MSTO age range of
-165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
-work for a more precise model-dependent comparison.
-This implies an absolute age slightly older than the Pleiades
-\citep[cf.][]{mermilliod_comparative_1981}.
-
-% FIXME: do you want to do the procedure below? Seems like overkill
-% given the Cummings study.
-
-% Our photometric age-dating procedure is as follows.  We selected all
-% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
-% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
-% outliers from the rest of the cluster.  Using an isochrone grid
-% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
-% spacing, we then numerically evaluated the $\chi^2$ fit for each
-% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
-% percentiles are used as summary statstics.
-% 
-% % NOTE: probably asymmetric.
-% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
-% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
-% yields $XXX\pm YYY$\,Myr.  Finally, applying the same procedure to
-% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
-% [nb. the procedure should be simple and general enough that you can
-% apply it to the field sample, and it'll just give a broad
-% distribution...]
+are consistent.  The left panel of Figure~\ref{fig:hr} shows that
+isochronally, they are: stars past the turnoff in both the halo and
+core are on a consistent locus.  The right panel of
+Figure~\ref{fig:hr} demonstrates the precision with which the claim
+can be made.  The MSTO stars are consistent with the 178 and 316 Myr
+models, but are bluer than the 100 Myr model.  The RGB stars are
+consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
+model.  Based on the assumption that the width of the MSTO can be
+attributed to binary stars, we are most interested in the blue edge
+(blue stragglers are, however, a concern).  The blue edge appears most
+compatible with the 178 Myr model.  These results are consistent with
+the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
+and we refer to their work for a more precise model-dependent
+comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
+adopt an PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
+our quoted uncertainty is a quadrature sum of their quoted statistical
+and systematic components.  The relative age of ``slightly older than
+the Pleiades'' is likely the more robust statement, consistent with
+reports since \citet{mermilliod_comparative_1981}.
 
@@
 
-        \vspace{-0.1cm}
+        \vspace{-0.5cm}
 		\subfloat{
@@
     {\it Top row}: \cn\ split into its core and halo.  Subset
-    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right columns}) undergo
-    successive degrees of automated cleaning, described in
-    Section~\ref{subsubsec:cluster}.  The Pleiades
+    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
+    columns}) undergo successive degrees of automated cleaning,
+    described in Section~\ref{subsubsec:cluster}.  The Pleiades
     \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
@@
     gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
-    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$,
-    in which some NGC\,2516 members appear slightly elevated with respect to
-    the Pleiades, suggesting an older age.
-    {\it Bottom row}: \cn\ colored by binarity indicators.
-    Sources with $\mathrm{RUWE}>1.2$ appear in red (astrometric
-    binaries).
+    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
+    which some NGC\,2516 members appear slightly elevated with respect
+    to the Pleiades, suggesting an older age.
+    {\it Bottom row}: \cn\ colored by binarity indicators.  Sources
+    with $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\%
+    of stars).
     Overplotted in orange are sources at least 0.3\,mag above an
-    empirically fit cluster isochrone (photometric binaries).
+    empirically fit cluster isochrone (photometric binaries; 20\% of stars).
 		\label{fig:rot}
@@
 This yielded \nautorotnumeratormatching light curves. 
+While subset $\mathcal{A}$ is more complete than subset $\mathcal{B}$,
+this completeness comes at the expense of greater contamination.
+% For instance, stars above the slow sequence (possible field
+% contaminants) for the most part disappear in subset $\mathcal{B}$
 
@@
 with those of the Pleiades and Praesepe.  The first is that in both
-stages of cleaning, the core and halo of \cn\ both show
+stages of cleaning, the core and halo of \cn\ each show
 gyrochronological ``slow sequences''.  These sequences appear to be
-coeval with each other, and at least to first order appear to overlap
+coeval with each other, and at first order appear to overlap
 with the Pleiades sequence.
 
-Second, subset $\mathcal{A}$ is more complete than subset
-$\mathcal{B}$.  This completeness comes at the expense of greater
-contamination.  The slow sequence has a factor of $\sim$twice as many
-stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
-the slow sequence, which could well be field contaminants, for the
-most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
-the slow sequence is also more pronounced in the cleaner subset,
-though it seems to not be as empty as in the Pleiades.
-
-Finally, comparing the slow sequences of the Pleaides and \cn, they
-appear to overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  For stars that
-are redder, the dispersion in rotation periods increases.  From
-$1.2<(Bp-Rp)_0<1.7$, subset $\mathcal{B}$ of \cn\ has stars that
-overlap with the Pleiades.  However, it also has about ten stars at
-longer rotation periods---extending up to $\approx$11 days, rather
-than the $\approx$8.5day limit for the Pleiades.  This is consistent
-with \cn\ being slightly older than the Pleiades.
-{\bf We explore this quantitatively by fitting interpolated
-models}[?].  The resulting gyrochronology age we find for the core is
-XXX.  For the halo, the claimed age from gyrochronology is YYY.
-Applying the same procedure to the field star comparison sample (see
-Appendix~\ref{app:compstar}), we get an age of ZZZ+/-AAA.
+% Second, subset $\mathcal{A}$ is more complete than subset
+% $\mathcal{B}$.  This completeness comes at the expense of greater
+% contamination.  The slow sequence has a factor of $\sim$twice as many
+% stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
+% the slow sequence, which could be field contaminants, for the
+% most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
+% the slow sequence is also more pronounced in the cleaner subset,
+% though it seems to not be as empty as in the Pleiades.
+
+%FIXME this belongs in the DISCUSSION!!
+%FIXME this belongs in the DISCUSSION!!
+%FIXME this belongs in the DISCUSSION!!
+%FIXME this belongs in the DISCUSSION!!
+Comparing the slow sequence of the Pleiades and \cn\ 
+more closely, the top row of Figure~\ref{fig:rot} shows that the two
+sequences
+overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
+($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
+The maximum rotation periods seen in \cn, in both subsets
+$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
+than the $\approx$8.5day limit seen in the Pleiades.  This is
+consistent with \cn\ being slightly older than the Pleiades.  Fitting
+a model to substantiate this statement \citep[{\it e.g.}, any
+of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020} is
+an exercise in tautology, since the aforementioned models are empirically fit using
+the Pleiades and Praesepe.  In other words, at $(Bp-Rp)_0=1.35$, we
+see rotation periods in \cn\ ranging from 7 to 10 days.  The 7 day
+periods overlap with the Pleiades; the 10 day periods do not.  Fitting
+formulae from the previously cited studies give ages for
+a star with $(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}$ of 204 Myr,
+316 Myr, and 107 Myr respectively.  Therefore, only the
+\citet{spada_competing_2020} model is well-calibrated at these ages
+and colors.  For an 8 day rotation period of a star with the same
+color, this model quotes a 193 Myr age; at 9 days, 388 Myr.
+Given the model uncertainty, we prefer the relative statements that
+\cn\ appears slightly gyrochronologically older than the Pleaides, and
+much younger than Praesepe.
 
 The bottom row of Figure~\ref{fig:rot} shows the same data, with the
-points colored by indicators of binarity.
-To assess astrometric binarity, we used the renormalized unit weight
-error (RUWE) from Gaia EDR3.
-For plotting purposes, we labeled anything with RUWE exceeding 1.2 as
-an astrometric binary (11\% of the overall cluster
-sample).
-A cautionary point is that poor astrometric fits can also be caused by
-photometric variability.
-To assess photometric binarity, we fitted a spline to
-Figure~\ref{fig:hr} by hand, and labelled any points brighter than
-0.3\,mag above the cluster sequence as photometric binaries.
-This included 20\% of the overall cluster sample.
-Many of the resulting candidate binaries, though not all of them,
-appear underneath the ``slow sequence''.
-We speculate on possible explanations in Section~\ref{disc:rotn}.
+points colored by indicators of binarity.  To assess astrometric
+binarity, we used the renormalized unit weight error (RUWE) from Gaia
+EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
+1.2 as an astrometric binary (11\% of the overall cluster sample).  A
+cautionary point is that poor astrometric fits can also be caused by
+photometric variability.  To assess photometric binarity, we fitted a
+spline to Figure~\ref{fig:hr} by hand, and labelled any points
+brighter than 0.3\,mag above the cluster sequence as photometric
+binaries.  This included 20\% of the overall cluster sample.  Many of
+the resulting candidate binaries, though not all of them, appear
+underneath the ``slow sequence''.  We speculate on possible
+explanations in Section~\ref{disc:rotn}.
 

### 2021-04-01 cluster_paper/ms.tex
update ms

@@
 % gaia target sample numbers
-\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
-\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
+\newcommand{\nkinematic}{3{,}298} % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nnbhd}{13{,}843} % make_compstar_NGC_2516_sourcelist.py
 \newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
@@
 % from drivers.cluster_rotation.get_auto_rotation_periods.py;  checked with plot_target_vs_actual_list_sanity_checks.py
-\newcommand{\ncdips}{2{,}205\ }
-\newcommand{\nkinematicrpltsixteen}{2{,}270\ }
-\newcommand{\nkinematicCGandKC}{2{,}236\ }
+\newcommand{\ncdips}{2{,}205}
+\newcommand{\nkinematicrpltsixteen}{2{,}270}
+\newcommand{\nkinematicCGandKC}{2{,}236}
 
 % cdips results numbers
-\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
-
-\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
-\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
-\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py
-
-\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
-\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
-\newcommand{\nautorotnumeratormatching}{507\ } % plot_autorotation.py, LSP=SPDM period
-\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
-\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
+\newcommand{\numcdipslcs}{671{,}901} % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+\newcommand{\numsouthernuniqlcs}{483{,}407} % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
+
+\newcommand{\ncalibration}{9{,}619{,}784} % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
+\newcommand{\nnbhdcaliblc}{9{,}790} % get_auto_rotation_periods.py
+\newcommand{\nnbhdcalibstar}{1{,}987} % get_auto_rotation_periods.py
+
+\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
+\newcommand{\nautorotnumeratormatching}{507} % plot_autorotation.py, LSP=SPDM period
+\newcommand{\ncompstardenominator}{820} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
+\newcommand{\ncompstarnumerator}{365} % plot_autorotation.py
 \newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
-\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
-\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
+\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
+\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
 \newcommand{\nautofrac}{53.1\%} % plot_autorotation.py
@@
 Our analysis of TESS Sectors 1-13 yielded light curves of
-\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
+\numsouthernuniqlcs\ candidate cluster members in the Southern Ecliptic
 hemisphere, which are available on
@@
 We selected candidate \cn\ members based on those reported in the
-literature.  While some pre-Gaia analyses were available,
-the purity and accuracy of the Gaia-derived results are the current
-state of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.
-We therefore adopted what we viewed as the most interesting clustering
-samples to compare: those of \citet{cantatgaudin_gaia_2018},
+literature.  While some pre-Gaia analyses were available, the purity
+and accuracy of the Gaia-derived results are the current state of the
+art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore
+adopted what we viewed as the most interesting clustering samples to
+compare: those of \citet{cantatgaudin_gaia_2018},
 \citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
@@
 		\leavevmode
-			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
 		\subfloat{
@@
   \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
-    {\it Top row}: \cn\ split into its core and halo.  Subset
+    {\it Top row}: \cn\ split into its core and halo.  Sets
     $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
@@
     to the Pleiades, suggesting an older age.  
-    {\it Bottom row}: \cn\ colored by binarity indicators.  Sources
-    with $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\%
-    of stars).
-    Overplotted in orange are sources at least 0.3\,mag above an
-    empirically fit cluster isochrone (photometric binaries; 20\% of stars).
+    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
+    are fainter than the magnitude limit of our survey.
+    {\it Bottom row}: \cn\
+    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
+    appear in red (astrometric binaries; 11\% of stars).  Overplotted
+    in orange are sources at least 0.3\,mag above an empirically fit
+    cluster isochrone (photometric binaries; 20\% of stars).
 		\label{fig:rot}
@@
 We began our rotation analysis by considering all \nkinematic\
-candidate members of \cn.  For each source, we first retrieved all
-available CDIPS light curves, on a per-sector basis.  This yielded
-\ncdips stars with at least one sector from a CDIPS light curve, all
-brighter than $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from
-the source list have $Rp<16$.  The difference is comprised of 34 stars
-unique to \citet{meingast_2021} which were not available at the time
-of the CDIPS reductions and stars falling on the TESS chip gaps.  The
+candidate members of \cn.  For each source, we retrieved all available
+CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
+stars with at least one sector from a CDIPS light curve.  Each of
+these stars is brighter than the CDIPS magnitude limit of
+$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
+source list had $Rp<16$.  The difference is due to 34 stars unique to
+\citet{meingast_2021} which were not available at the time of the
+CDIPS reductions, and a few stars falling on the TESS chip gaps.  The
 $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
 roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
-the distance of NGC\,2516.
+the distance of \cn.
 
@@
 CCD in each sector individually, and stitched together the resulting
-light curves before searching for the periodicity.  Details regarding
-our detrending approach are discussed in
-Appendix~\ref{app:detrending}.  After applying the initial detrending
+light curves before searching for rotation-induced periodicity.
+Details regarding our detrending approach are presented in
+Appendix~\ref{app:detrending}.  After applying an initial detrending
 step aimed at removing systematic trends, we proceeding with a few
-small cleaning steps aimed at improving the purity of the rotation
-period measurements: we masked 0.7 days at the beginning and end of
-each spacecraft orbit, and ran a sliding standard-deviation rejection
+small cleaning steps to improve the purity of the rotation period
+measurements: we masked 0.7 days at the beginning and end of each
+spacecraft orbit, and ran a sliding standard-deviation rejection
 window over the light curve, which removed any outlying points within
@@
 
-We then measured the rotation period from the resulting light curve.
+We then measured the rotation period from the resulting light curve
+using the periodogram implementations in \texttt{astrobase}, in
+particular the Stellingwerf phase-dispersion minimization periodogram,
+along with the more traditional Lomb-Scargle
+\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
 We used the CDIPS aperture radius that, based on theoretical
@@
 typically resulted in an aperture radius of either 1 or 1.5 pixels.
-To measure the periods, we used the periodogram implementations in
-\texttt{astrobase}, in particular the Stellingwerf phase-dispersion
-minimization periodogram, along with the more traditional Lomb-Scargle
-\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
 We recorded the top five periodogram peaks from each method, and their
-corresponding powers.  Finally, as a check on crowding, we also
-recorded the number of stars within the aperture of equal brightness
-to the target stars, and of brightness with 1.25 and 2.5 TESS
-magnitudes of the target star.
+corresponding powers.  Finally, as a check on crowding, we recorded
+the number of stars within the aperture with equal or greater
+brightness than the target star, and with brightness within 1.25 and
+2.5 TESS magnitudes of the target star.
+
+To clean the resulting rotation period measurements, we devised a set
+of automated cleaning criteria, which yielded sets we denote
+$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered
+light curves for which the peak Lomb-Scargle periodogram period was
+below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
+for which no equal-brightness or greater companions were within the
+aperture.  Beyond requiring that the target star be the brightest star
+in the aperture, we also required that at most one companion with
+brightness exceeding one tenth of the target star's brightness be
+present in the aperture according to the Gaia DR2 source catalog.  For
+set $\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out
+of \nautorotdenominator\ stars for which rotation periods might
+plausibly have been detected.  For set $\mathcal{B}$, we additionally
+required that the Lomb-Scargle and Stellingwerf phase-dispersion
+methods yielded identical periods to within 10\%.  This yielded
+\nautorotnumeratormatching\ light curves.
 
 The top row of Figure~\ref{fig:rot} shows the resulting rotation
-periods, following three different sets of cleaning criteria.  The
-first and second sets of criteria were entirely automated, and yielded
-subsets we denote $\mathcal{A}$ and $\mathcal{B}$.  For both subsets,
-we considered light curves for which the peak Lomb-Scargle periodogram
-period was below 15 days, the normalized Lomb-Scargle power exceeded
-0.08, and for which no equal-brightness or greater companions were
-within the aperture.  Beyond requiring that the target star be the
-brightest star in the aperture, we also required that at most one
-companion with brightness exceeding one tenth of the target star's
-brightness be present in the aperture according to the Gaia DR2 source
-catalog.  For set $\mathcal{A}$, this yielded \nautorotnumerator light
-curves, out of \nautorotdenominator stars for which rotation periods
-might plausibly have been detected.  For set $\mathcal{B}$, we 
-additionally required that the Lomb-Scargle and Stellingwerf
-phase-dispersion methods yielded identical periods to within 10\%.
-This yielded \nautorotnumeratormatching light curves. 
-While subset $\mathcal{A}$ is more complete than subset $\mathcal{B}$,
-this completeness comes at the expense of greater contamination.
-% For instance, stars above the slow sequence (possible field
-% contaminants) for the most part disappear in subset $\mathcal{B}$
-
-The comparison to the Pleiades and Praesepe is made using the rotation
-periods and dereddened colors from Table~5 of
-\citet{curtis_rup147_2020}, which used data from
-\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
-respectively.
-A few conclusions can be drawn by comparing the \cn\ rotation periods
-with those of the Pleiades and Praesepe.  The first is that in both
-stages of cleaning, the core and halo of \cn\ each show
-gyrochronological ``slow sequences''.  These sequences appear to be
-coeval with each other, and at first order appear to overlap
-with the Pleiades sequence.
-
-% Second, subset $\mathcal{A}$ is more complete than subset
-% $\mathcal{B}$.  This completeness comes at the expense of greater
-% contamination.  The slow sequence has a factor of $\sim$twice as many
-% stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
-% the slow sequence, which could be field contaminants, for the
-% most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
-% the slow sequence is also more pronounced in the cleaner subset,
-% though it seems to not be as empty as in the Pleiades.
-
-%FIXME this belongs in the DISCUSSION!!
-%FIXME this belongs in the DISCUSSION!!
-%FIXME this belongs in the DISCUSSION!!
-%FIXME this belongs in the DISCUSSION!!
-Comparing the slow sequence of the Pleiades and \cn\ 
-more closely, the top row of Figure~\ref{fig:rot} shows that the two
-sequences
-overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
-($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
-The maximum rotation periods seen in \cn, in both subsets
-$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
-than the $\approx$8.5day limit seen in the Pleiades.  This is
-consistent with \cn\ being slightly older than the Pleiades.  Fitting
-a model to substantiate this statement \citep[{\it e.g.}, any
-of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020} is
-an exercise in tautology, since the aforementioned models are empirically fit using
-the Pleiades and Praesepe.  In other words, at $(Bp-Rp)_0=1.35$, we
-see rotation periods in \cn\ ranging from 7 to 10 days.  The 7 day
-periods overlap with the Pleiades; the 10 day periods do not.  Fitting
-formulae from the previously cited studies give ages for
-a star with $(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}$ of 204 Myr,
-316 Myr, and 107 Myr respectively.  Therefore, only the
-\citet{spada_competing_2020} model is well-calibrated at these ages
-and colors.  For an 8 day rotation period of a star with the same
-color, this model quotes a 193 Myr age; at 9 days, 388 Myr.
-Given the model uncertainty, we prefer the relative statements that
-\cn\ appears slightly gyrochronologically older than the Pleaides, and
-much younger than Praesepe.
+periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
+the right.  While set $\mathcal{A}$ is more complete than set
+$\mathcal{B}$, this completeness comes at the expense of greater
+contamination.  For instance, stars above the slow sequence (possible
+field contaminants) for the most part disappear in set $\mathcal{B}$.
+The ``void'' or ``gap'' underneath the slow sequence is also more
+pronounced in set $\mathcal{B}$.  To facilitate a comparison against
+the Pleiades and Praesepe, we used the rotation periods and dereddened
+colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
+from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
+respectively.  The first order conclusion is that the Pleiades and
+\cn\ appear gyrochronologically coeval for colors spanning
+$0.5<(Bp-Rp)_0<1.2$.  The implications of this comparison for the age
+of \cn\ are assessed in greater detail in Section~\ref{disc:absage}.
 
 The bottom row of Figure~\ref{fig:rot} shows the same data, with the
-points colored by indicators of binarity.  To assess astrometric
+points colored with indicators of binarity.  To assess astrometric
 binarity, we used the renormalized unit weight error (RUWE) from Gaia
 EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
-1.2 as an astrometric binary (11\% of the overall cluster sample).  A
-cautionary point is that poor astrometric fits can also be caused by
-photometric variability.  To assess photometric binarity, we fitted a
-spline to Figure~\ref{fig:hr} by hand, and labelled any points
-brighter than 0.3\,mag above the cluster sequence as photometric
-binaries.  This included 20\% of the overall cluster sample.  Many of
-the resulting candidate binaries, though not all of them, appear
-underneath the ``slow sequence''.  We speculate on possible
-explanations in Section~\ref{disc:rotn}.
+1.2 as an astrometric binary (11\% of the overall cluster sample; see
+{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
+photometric binarity, we fitted a spline to Figure~\ref{fig:hr} by
+hand, and labelled any points brighter than 0.3\,mag above the cluster
+sequence as photometric binaries.  This included 20\% of the overall
+cluster sample.  Many of the resulting candidate binaries, though not
+all of them, appear underneath the ``slow sequence''.  We speculate on
+possible explanations in Section~\ref{disc:rotn}.
 
@@
 		}
-	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
 	\end{center}
@@
 detected rotation periods against our original target list of
-\nkinematic Gaia DR2 kinematic candidate members.  For a fair
+\nkinematic\ Gaia DR2 kinematic candidate members.  For a fair
 comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
@@
 \cn\ members.  Cross-matching against our kinematic list of
-\nkinematic candidate members by position and imposing a 0$\farcs$5
+\nkinematic\ candidate members by position and imposing a 0$\farcs$5
 maximum separation limit yields 492 kinematic members for which a
@@
 Li\,6708\,\AA doublet, these were the most relevant bit flags.  Of our
-\nkinematic candidate \cn\ members, 107 had spectra in GALAH DR3.  51
+\nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.  51
 were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
@@
 
-\subsection{Rotation in NGC\,2516}
+\subsection{Rotation at 100-200\,Myr}
 \label{disc:rotn}
 
-Gotta cite Fritzewski+2020.  (Maybe even do the Blanco1, Pleiades,
-NGC2516, PscEri concatenation?)
-
-  \paragraph{Slow sequence and fast sequence}
-  %Why is there a slow and fast sequence?
-  Slow: Skumanich spindown, Weber and Davis magnetic braking.
-  This doesn't explain everything.
-  Originally described as ``interface / convective'' dynamo for each.
-  Now disfavored (but why?).
-  Core envelope decoupling handwaves away the fast rotators? (Irwin,
-  Aigrain, Monitor).
-  Another fast rotator explanation: Naoz's merging of giant planets
-  (could explain some fraction of the fast rotators -- not all of them).
-  Tim Brown, ``metastable dynamo''?
-
-  Perhaps most compelling: some kind of binary related explanation.
-  Gillen+20 in Blanco-1 found most of them to be photometric binaries.
-  Simonian+19 in the Kepler field found most rapid rotators to have
-  photometric excesses.
+From their rotation period analysis of \cn,
+\citet{fritzewski_rotation_2020} have recently solidified our
+understanding that at ages between 100 and 200\,Myr, young open
+clusters ({Blanco\,1, the Pleiades, Psc\,Eri, M\,35, and M\,50) all
+seem to show similar rotation period distributions.  The FGK stars
+show a ``slow sequence'', a population of ``rapid rotators'', and
+in some cases stars in the ``void'' between the two
+populations ({\it e.g.,}, Psc\,Eri and \cn\ show such populations).
+The late M dwarfs with detected rotation periods tend to rotate very
+rapidly ($<3$\,days) --- these are unfortunately beyond the magnitude
+limit of the present analysis.  In this section, we briefly discuss
+these features, and focus on the main difference between \cn\
+and the Pleiades: the longer maximum rotation periods seen in the
+K-dwarfs.
+
+\subsubsection{The slow and fast sequences}
+The existence of a slow sequence is the main premise of gyrochronology
+\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
+spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
+\citet{skumanich_time_1972} has not worked in detail, the fact that it
+is simple and can be understood to first order through magnetic
+breaking \citep{weber_angular_1967} has led to more systematic efforts
+at age-dating stars through their rotation periods.
+
+The other features of the rotation-color diagram are in some sense
+affronts to the magnetic breaking model, and the general question of
+what produces the rapid rotators has not yet been definitively
+answered.  Models that introduces variable core-envelope decoupling
+timescales can reproduce each type of population
+\citep{Irwin_NGC2516_2007}, but the question then shifts to why these
+decoupling timescales should be differ for stars that are otherwise
+identical.  Another possible explanation invokes a spontaneous and
+random change of magnetic field topology that cause the dynamo to
+couple more strongly with the stellar wind
+\citep{brown_metastable_2014}.  Other processes could be external the
+star entirely: \citet{qureshi_signature_2018} for instance have
+reported that mergers between giant planets and the star could explain
+a non-neglible fraction of the rapid rotators.
+
+A separate hypothesis that we can test from the existing data is the
+idea that 
+the rapid rotators are somehow linked to binarity.
+Analyses by {\it e.g.}, \citet{gillen_ngts_2020} and
+\citet{simonian_rapid_2019} have reported that
+rapid rotators in young clusters and in the field are correlated
+with binarity.
+Related effects could include tidal synchronization for the shortest
+period binaries,
+or some form of disk truncation that when combined with magnetic locking
+(CITE)
+could produce the rotation period anomaly.
+The lower panels of Figure~\ref{fig:rot} show our best attempt to
+identify unresolved binaries in the \cn\ rotation sample.
+Binaries resolved by Gaia were already excluded in our initial
+definitions of the samples.
+WHAT IS THE FRACTION OF FAST ROTATORS THAT ARE BINARY?
+WHAT CONCLUSION CAN BE MADE?
+%FIXME CALCULATE!!!!
+%FIXME CALCULATE!!!!
+%FIXME CALCULATE!!!!
+%FIXME CALCULATE!!!!
+
+\subsubsection{The slowest K dwarfs of NGC\,2516}
+%FIXME need to rewrite this paragraph? (Or maybe not, because it's
+%already in the absolute age section...
+A few conclusions can be drawn by comparing the \cn\ rotation periods
+with those of the Pleiades and Praesepe.  The first is that in both
+stages of cleaning, the core and halo of \cn\ each show
+gyrochronological ``slow sequences''.  These sequences appear to be
+coeval with each other, and at first order appear to overlap
+with the Pleiades sequence.
 
@@
 
-  \paragraph{Are the ``very slow rotators'' bonafide members?}
-  Unsure. They are not isotropically distributed around the
-  cluster... so either triggered star formation (CITE, CITE), or they're
-  actually field stars.
+%   \paragraph{Are the ``very slow rotators'' bonafide members?}
+%   Unsure. They are not isotropically distributed around the
+%   cluster... so either triggered star formation (CITE, CITE), or they're
+%   actually field stars.
 
@@
 \subsection{Age of NGC\,2516}
+\label{disc:absage}
 Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
@@
 
-% \subsection{Ways of doing this at different ages}
-% Some final discussion is warranted on the applicability of our
-% approach more generally.  Given a kinematically
-% identified population, this study uses photometry (Gaia HR diagrams;
-% TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
-% confirm youth in the stars.  Other approaches are also possible.
-% 
-% One can use photometric criteria (e.g., Gaia HR diagram of
-% everything within some distance; Zari+18), or spectroscopic
-% criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
-% Zerjal 2019, 2020, Zhou's work).  
-% It might even be possible to start using chemical abundance signatures
-% (CITE, GALAHDR3?).
-% 
-% These paths might in certain cases succeed in identifying more
-% complete samples of dispersed members of young populations, since they
-% do not require kinematic proximity.
+\paragraph{Gyrochronological Age}
+Comparing the slow sequence of the Pleiades and \cn\ more closely, the
+top row of Figure~\ref{fig:rot} shows that the two sequences overlap
+from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
+($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
+The maximum rotation periods seen in \cn across both sets
+$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
+than the $\approx$8.5day upper limit seen in the Pleiades.  This is
+consistent with \cn\ being slightly older than the Pleiades.
+
+Fitting a model to substantiate the claim that \cn\ is
+gyrochronologically older than the Pleiades \citep[{\it e.g.}, any
+of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
+is, unfortunately, an exercise in tautology.  The aforementioned
+models are empirically calibrated using the Pleiades and Praesepe.  In
+other words, at $(Bp-Rp)_0=1.35$, we observe rotation periods in \cn\
+ranging from 7 to 10 days.  The 7 day rotation periods are consistent
+with those observed in the Pleiades; the 10 day rotation periods are
+not.  Fitting formulae from the previously cited studies give ages for
+a star with $\{(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}\}$ of 204
+Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
+\citet{spada_competing_2020} model has an absolute scale that is
+well-calibrated at these ages and colors.  For an 8 day rotation
+period of a star with the same color, this model quotes a 193 Myr age;
+at 9 days, 388 Myr.  Given this degree of model uncertainty, we prefer
+the relative statement that \cn\ appears slightly gyrochronologically
+older than the Pleaides, and much younger than Praesepe.  Given a
+Pleiades age of $130\pm15$\,Myr (an average of MSTO and
+rotation-corrected lithium-depletion results;
+\citealt{cummings_2018}), a gyrochronology-derived age estimate for \cn\
+between 150 and 200\,Myr would therefore appear reasonable.  Much
+older, and the issue of why the F and G dwarfs do not spin down
+becomes pressing.
 
@@
   \caption{ {\bf Different clustering techniques yield different
-  outcomes.} \nkinematic unique candidate cluster members found
+  outcomes.} \nkinematic\ unique candidate cluster members found
   using three different techniques are considered in this study.
@@
 the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
-corresponded to a sample of \ncalibration stars.  Cross-matching these
-against the \nnbhd randomly drawn stars in the neighborhood of \cn yielded
-\nnbhdcalibstar unique stars, with a cumulative total of \nnbhdcaliblc TESS
+corresponded to a sample of \ncalibration\ stars.  Cross-matching these
+against the \nnbhd\ randomly drawn stars in the neighborhood of \cn yielded
+\nnbhdcalibstar\ unique stars, with a cumulative total of
+\nnbhdcaliblc\ TESS
 sectors observed.
@@
 Imposing the same requirements for crowding resulted in
-\ncompstardenominator stars for which rotation periods could have been
+\ncompstardenominator\ stars for which rotation periods could have been
 detected.
 Imposing the same Lomb Scargle power cutoff, and period upper limit,
-yielded \ncompstarnumerator period detections (\ncompfrac).
+yielded \ncompstarnumerator\ period detections (\ncompfrac).
 Within the same brightness cutoff, 
-\nautovscompstarnumerator of \nautovscompstardenominator cluster stars
+\nautovscompstarnumerator\ of \nautovscompstardenominator\ cluster stars
 yielded period detections (\nautofrac).

### 2021-04-01 earhart/plotting.py

@@ -1116,7 +1116,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     lws = [0., 0.3, 0.2]
     mews= [0.5, 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
-    ss = [3, 4.5, 9]
+    ss = [5, 6, 10]
     labels = ['Pleaides', 'Praesepe', f'{runid}']
 
     # plot vals
@@ -1233,7 +1233,8 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     ax.set_ylabel('Rotation Period [days]', fontsize='large')
 
     ax.set_xlabel('(Bp-Rp)$_0$ [mag]', fontsize='large')
-    ax.set_xlim((0.25, 2.0))
+    ax.set_xlim((0.2, 2.4))
+    #ax.set_xlim((0.25, 2.0))
 
     if yscale == 'linear':
         ax.set_ylim((0,15))

### 2021-04-05 cluster_paper/ms.tex
update ms

@@
     \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
-    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
+    \citep[650\,Myr;][]{douglas_poking_2017} are shown.
     Stars in the core and the halo of NGC\,2516 appear
@@
 cluster sample.  Many of the resulting candidate binaries, though not
-all of them, appear underneath the ``slow sequence''.  We speculate on
-possible explanations in Section~\ref{disc:rotn}.
+all of them, appear underneath the slow rotation sequence.  In set
+$\mathcal{B}$, over the color range where the slow and fast sequence
+are present ($0.5<(Bp-Rp)_0<1.3$), there are 281 stars, of which 103
+(37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
+and ``slow'' sequences by eye,  the fraction of stars showing signs of
+binarity in the slow and fast rotation subsamples are 30\% (63/210)
+and 56\% (40/71), respectively.  We speculate on possible explanations
+in Section~\ref{disc:rotn}.
 
@@
 spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
-\citet{skumanich_time_1972} has not worked in detail, the fact that it
-is simple and can be understood to first order through magnetic
-breaking \citep{weber_angular_1967} has led to more systematic efforts
-at age-dating stars through their rotation periods.
+\citet{skumanich_time_1972} has not worked in detail, the 
+general idea is correct, and the first-order effect can be understood through magnetic
+breaking \citep{weber_angular_1967}.
+%More systematic efforts at age-dating stars through their rotation periods.
 
@@
 what produces the rapid rotators has not yet been definitively
-answered.  Models that introduces variable core-envelope decoupling
+answered.  Models that introduce variable core-envelope decoupling
 timescales can reproduce each type of population
@@
 identical.  Another possible explanation invokes a spontaneous and
-random change of magnetic field topology that cause the dynamo to
+random change of magnetic field topology that causes the dynamo to
 couple more strongly with the stellar wind
@@
 
-A separate hypothesis that we can test from the existing data is the
-idea that 
-the rapid rotators are somehow linked to binarity.
-Analyses by {\it e.g.}, \citet{gillen_ngts_2020} and
-\citet{simonian_rapid_2019} have reported that
-rapid rotators in young clusters and in the field are correlated
-with binarity.
-Related effects could include tidal synchronization for the shortest
-period binaries,
-or some form of disk truncation that when combined with magnetic locking
-(CITE)
-could produce the rotation period anomaly.
-The lower panels of Figure~\ref{fig:rot} show our best attempt to
-identify unresolved binaries in the \cn\ rotation sample.
+A separate hypothesis that we can test using the existing data is the
+idea that rapid rotation can be explained through binarity.  Analyses
+by {\it e.g.}, \citet{gillen_ngts_2020} and
+\citet{simonian_rapid_2019} have reported that rapid
+rotation in young clusters and in the field is correlated with
+binarity.  Related effects could include tidal synchronization for the
+shortest period binaries, or alternatively disk truncation that when
+combined with magnetic locking {\bf (CITE)} could produce the rotation
+period bimodality.
+
+The lower panels of Figure~\ref{fig:rot} show our attempt at
+identifying unresolved binaries in the \cn\ rotation sample.
 Binaries resolved by Gaia were already excluded in our initial
 definitions of the samples.
-WHAT IS THE FRACTION OF FAST ROTATORS THAT ARE BINARY?
-WHAT CONCLUSION CAN BE MADE?
-%FIXME CALCULATE!!!!
-%FIXME CALCULATE!!!!
-%FIXME CALCULATE!!!!
-%FIXME CALCULATE!!!!
-
-\subsubsection{The slowest K dwarfs of NGC\,2516}
-%FIXME need to rewrite this paragraph? (Or maybe not, because it's
-%already in the absolute age section...
-A few conclusions can be drawn by comparing the \cn\ rotation periods
-with those of the Pleiades and Praesepe.  The first is that in both
-stages of cleaning, the core and halo of \cn\ each show
-gyrochronological ``slow sequences''.  These sequences appear to be
-coeval with each other, and at first order appear to overlap
-with the Pleiades sequence.
-
-  \paragraph{Stalling}
-  Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
-  spindown is not monotonic in time.
-  We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
-  much change.
+In \cn\, we see that
+the fraction of stars showing signs of binarity in the slow and fast
+rotation subsamples are 30\% (63/210) and 56\% (40/71); the fast
+rotators show a preference for binarity.
+This correlation is in line with the earlier findings noted above,
+though we emphasize many rapid rotators in \cn\ (34\%) have no 
+evidence for being binary.
+
+Comparing the explanations of disk-locking against tidal
+synchronization, the population statistics do seem to strain the
+possibility of tidal synchronization.  A quarter of the \cn\ members
+are rapidly rotating.  In comparison, in the field half of Sun-like
+stars are binaries, and $\approx$9\% have periods below 100\,days
+\citep{raghavan_survey_2010}.  If we assume  that all
+binaries with sub-100\,day periods become tidally synchronized, we
+can only explain a rapid rotator occurrence rate of
+$\approx$5\%.
+
+A separate effect is that on the slow rotation sequence, 
+Figure~\ref{fig:rot}} shows that the
+binary stars appear to be either preferentially redder, or to have
+faster rotation periods than the single stars.  One likely
+explanation for this could be that the unresolved binaries have a
+component contributing additional red light to the system, skewing
+the color measurement of the primary.  Whether any physical
+effects could be at play remains a question for future
+work.
+
+% \subsubsection{The slowest K dwarfs of NGC\,2516}
+% %FIXME need to rewrite this paragraph? (Or maybe not, because it's
+% %already in the absolute age section...
+% A few conclusions can be drawn by comparing the \cn\ rotation periods
+% with those of the Pleiades and Praesepe.  The first is that in both
+% stages of cleaning, the core and halo of \cn\ each show
+% gyrochronological ``slow sequences''.  These sequences appear to be
+% coeval with each other, and at first order appear to overlap
+% with the Pleiades sequence.
+% 
+%   \paragraph{Stalling}
+%   Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
+%   spindown is not monotonic in time.
+%   We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
+%   much change.
 

### 2021-04-06 cluster_paper/ms.tex
update ms

@@
   Recent analyses of the Gaia data have reported the existence of
-  diffuse stellar populations surrounding nearby open clusters. 
-  The stars in these diffuse halos could be tidal escapees from the
-  cluster cores, or they could be remnants of an initially dispersed
-  star formation complex.
-  They could also be field stars---the false positives of overly
-  sensitive clustering algorithms.
-  In this study, we focus on a halo with a tip-to-tip length of
-  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
-  NGC\,2516.
-  The classical tidal radius of the cluster is $\lesssim$10\,pc.
-  Combining photometry from Gaia, rotation periods from TESS, and
-  lithium measurements from Gaia-ESO and GALAH, we find that the halo
-  of NGC\,2516 is real and coeval with the core.
-  Two thirds of kinematically selected halo members out to 250\,pc from the
-  cluster center have rotation periods consistent with a
-  gyrochronological age of 150\,Myr.
-  A comparison sample of field stars shows no such trend.
-  In addition, the lithium abundances of stars in the halo are
+  diffuse stellar populations surrounding nearby open clusters.  It is
+  important to verify that these ``halos'', ``tails'', and ``strings''
+  are of similar ages and compositions as stars in the denser part of
+  the cluster.  We present an analysis of NGC\,2516
+  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
+  and an apparent halo of stars spanning 500\,pc.  Combining
+  photometry from Gaia, rotation periods from TESS, and lithium
+  measurements from Gaia-ESO and GALAH, we find that the halo of
+  NGC\,2516 is real and coeval with the core.  Two thirds of
+  kinematically selected halo members out to 250\,pc from the cluster
+  center have rotation periods consistent with a gyrochronological age
+  of 150\,Myr.  A comparison sample of field stars shows no such
+  trend.  In addition, the lithium abundances of stars in the halo are
   elevated relative to the field, and are correlated with the stellar
-  rotation rate in a manner consistent with previous observations.
-  This work expands the set of confirmed NGC 2516 members by a factor
-  of two, and is consistent with a paradigm in which the halos of many
-  open clusters are more populous than their cores.
-  We highlight the implications for spectroscopic survey targeting,
-  open cluster dispersal, and planet searches around young stars.
+  rotation rate in a manner consistent with previous observations of
+  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
+  paradigm in which the halos of many open clusters are more populous
+  than their cores.  We highlight the implications for spectroscopic
+  survey targeting, open cluster dispersal, and planet searches around
+  young stars.
 \end{abstract}
@@
 understand the spatial extent and duration of star formation events
-\citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
-is to understand the Sun's birth environment \citep{adams_birth_2010}.
-Did the Sun form in an open cluster, and if so, what type? Is there
-any hope at identifying the stars that formed near the Sun?  Open
-clusters and their dispersed remnants also provide a test case for
-``chemical tagging'' \citep[{\it
+\citep[{\it e.g.},][]{wright_kinematics_2018}.  Another is to
+understand the Sun's birth environment \citep{adams_birth_2010}.  Did
+the Sun form in an open cluster, and if so, what type? Is there any
+hope at identifying the stars that formed near the Sun?  Open clusters
+and their dispersed remnants also provide a test case for ``chemical
+tagging'' \citep[{\it
 e.g.},][]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
@@
 One point of difficulty however is that different clustering methods
-yield differing levels of both sensitivity and precision
+yield differing levels of sensitivity and precision
 \citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
@@
 is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
-to facilitate photometric rotation measurements using TESS, and some
-degree of spectroscopic analysis.  Its halo is decidedly
-non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
-as $\approx20\times10\times350$\,pc.  We want to know: is the halo
-real? To what extent can we use Gaia alone to reliably identify
-age-dated needles in the haystack of field stars?  What are the
-implications, observationally and theoretically, if we can identify
-the dispersed halos of open clusters?
-
-%TODO: double check this owrding.
+to facilitate photometric rotation measurements using TESS.  Its halo
+was recently reported by \citet{kounkel_untangling_2019} to span
+$\approx20\times10\times350$\,pc.  We want to know: is this halo real?
+To what extent can we use Gaia alone to reliably identify age-dated
+needles in the haystack of field stars?  What are the implications,
+observationally and theoretically, if we can identify the dispersed
+halos of open clusters?
+
 A brief note on terminology.  Low-density stellar associations
-connected to a dense population (the ``core'') have been described as
+connected to a dense population (a ``core'') have been described as
 being in ``halos'', ``strings'', ``coronae'', ``snakes'',
@@
 The latter term implies a particular model for the formation of the
-association.  We opt for ``halo'' because it is concise and model
-agnostic.  It suffers however from the suggestion of spherical
-symmetry.  The halo of NGC\,2516 might be best described as ``tails'',
-but the primary goal of this study is not to verify whether these
-tails are tidal.
+dispersed group.  ``Halo'' is concise and model agnostic, but suffers
+from the suggestion of spherical symmetry.  The halo of NGC\,2516 is
+most accurately described as consisting of a leading and trailing
+tail. However, the goal of this study is to clarify whether the stars
+in these tails are the same age as stars in the core, and not to make
+a ruling on the origin of the overall structure.  For this reason, we
+adopt the model-agnostic ``halo''.
 
 Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
-data that led to our interest in NGC\,2516.  Section~\ref{sec:agedate}
-age-dates the cluster's halo and core using Gaia photometry
+data that led to our interest in \cn.  Section~\ref{sec:agedate}
+measures the age of the cluster's halo and core using Gaia photometry
 (Section~\ref{subsec:hr}), TESS gyrochronology
@@
 \label{sec:gaia}
-
-%\subsection{Gaia Astrometry}
-%\label{subsec:astrometry}
 
@@
   +20.8\}\,{\rm pc}$.  The gray line traces the past and future
-  cluster orbit.
-  In different projections, the gray line spans different
-  time intervals: the longest time window is visible in the $(Z,X)$
-  projection.  The large blue arrows denote the median cluster
+  cluster orbit.  In different projections, the gray line spans
+  different time intervals: the longest time window is visible in the
+  $(Z,X)$ projection.  The large blue arrows denote the median cluster
   velocity after subtracting the local standard of rest: $\{v_{\rm X},
-  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By comparison,
-  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the inertial
-  frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
-  direction.
-  The sizes of the blue arrows are internally proportionate. 
+  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By
+  comparison, $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the
+  inertial frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
+  direction.  The sizes of the blue arrows are internally
+  proportionate. 
   \label{fig:XYZ}
@@
 We selected candidate \cn\ members based on those reported in the
-literature.  While some pre-Gaia analyses were available, the purity
-and accuracy of the Gaia-derived results are the current state of the
-art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore
-adopted what we viewed as the most interesting clustering samples to
-compare: those of \citet{cantatgaudin_gaia_2018},
+literature.  While some pre-Gaia analyses were available
+\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
+accuracy of the Gaia-derived results are the current state of the art.
+We therefore adopted what we viewed as the most important clustering
+samples to compare: those of \citet{cantatgaudin_gaia_2018},
 \citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
@@
 \citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
-respectively.  A useful visualization of the different samples is
+respectively.  A useful visualization of these samples is
 available online\footnote{
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}.}.  While we considered performing our
-own clustering analysis using the Gaia data, such an effort would in
-effect only replicate the work of these investigators.  We opt to
-instead use their studies as starting points.
+  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}.
+While we considered performing our own clustering analysis on the
+Gaia data, such an effort would in effect only replicate the work of
+these investigators.  We opt to instead use their studies as a starting
+point.
 
@@
 considered stars brighter than $G=18$\,mag within a few degrees of the
-cluster center, and reported 1106 candidate \cn\ members.
+center of \cn, and reported 1106 candidate cluster members.
 \citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
@@
 that each of these studies applied to the second Gaia data release are
-discussed Appendix~\ref{app:clustering}, as is the overlap between
-their resulting membership samples.
+discussed in Appendix~\ref{app:clustering}, as is the overlap between
+their resulting membership catalogs.
 
@@
 study in the space of observed positions, proper motions, and radial
-velocity when available.  The \citetalias{cantatgaudin_gaia_2018}
-members are all within a few degrees of the cluster center, while the
-\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
-members (the ``halo'') span tens of degrees.  Inverting the parallax
-shows that members have been reported from line-of-sight distances
-ranging from $300$ to $600\,$pc.  Stars in the upper right quadrant of
-the $\mu_{\alpha'}$ {\it vs.} $\mu_\delta$ projection were reported
-almost entirely by \citet{meingast_2021}.
-
+velocity when available.
 In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
@@
 \ncore\ core members, and \nhalo\ halo members.  The distinction is
-perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
+tautological, in the sense that \citetalias{cantatgaudin_gaia_2018} did
 not extend their search for members out to tens of degrees from the
 cluster center.  Nonetheless, the \citetalias{cantatgaudin_gaia_2018}
-membership catalog is consistent with that of many previous
+membership catalog is consistent with that of many earlier
 investigators \citep[{\it
 e.g.},][]{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, and is
-consistent with the general {\it visual} impression that one has when
-looking at \cn\ in the night sky: it appears to span
-$\approx$2$^\circ$, not $\approx20^\circ$.
+consistent with the general visual overdensity one sees when viewing
+\cn\ in the sky: it appears to span $\approx$2$^\circ$, not
+$\approx20^\circ$.
 
@@
 of stellar masses and evolutionary states.  We also required the
-comparison sample to not overlap with the cluster sample, which led to
-the omission of 1.1\% of the stars drawn from the volume noted above.
-
-The discrepancy between our canonical understanding of open clusters
-and the possibility of a diffuse population existing around \cn\
-prompted our interest in understanding the reality and origin of the
-diffuse population.  Figure~\ref{fig:XYZ} helps in visualizing the
-proposed structure and kinematics of the halo.  We computed the
-physical positions by transforming from $(\alpha, \delta, \pi)$ to
+comparison sample to not overlap with the cluster sample.%, which led to
+%the omission of 1.1\% of the stars drawn from the volume noted above.
+
+The style of visualization given in Figure~\ref{fig:gaia6d} does not
+make it clear, in our eyes, why the clustering algorithms have decided
+to associate certain stars with the halo.  Canonically, open clusters
+are spherical, and span $\approx$10\,pc \citep[{\it
+e.g.},][]{Kharchenko_et_al_2013}.  Inverting the parallaxes in
+Figure~\ref{fig:gaia6d} shows that members have been reported from
+line-of-sight distances ranging from $\approx$$300$ to
+$\approx$$600\,$pc. 
+Is this structure real? What explains its existence?
+
+% Are these halo stars truly associated with the cluster's core?
+% What is the 
+% The opportunity of verifying and understanding the diffuse
+% population of \cn\ prompted our interest in understanding
+% the reality and origin of the halo.
+
+%Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
+%$\mu_\delta$ projection were reported almost entirely by
+%\citet{meingast_2021}.
+
+An initial step in visualizing the reported structure and kinematics
+of the candidate cluster members is to consider them in Cartesian
+coordinates (Figure~\ref{fig:XYZ}).  We computed the physical
+positions by transforming from $(\alpha, \delta, \pi)$ to
 galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
@@
 the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.
-The elongation of the cluster in both the $(X,Y)$ and $(Z,Y)$ planes
-is correlated with the direction of the LSR-corrected median cluster
-velocity.  A similar correlation was noted in Coma Ber by
-\citet{tang_comaber_2019}.  Possible explanations are discussed in
-Section~\ref{subsec:origin}.
+
+The general shape of the cluster itself seems to include a central overdensity
+and two tails (the halo).  One tail is leading the cluster's orbit,
+and is angled toward the center of the galaxy when viewed top-down.
+The other tail is trailing the cluster's orbit, and is pointed away
+from the center of the galaxy.  The elongation of the cluster in both
+the $(X,Y)$ and $(Z,Y)$ planes is correlated with the direction of the
+LSR-corrected median cluster velocity.   Possible explanations for
+this overall morphology are discussed in Section~\ref{subsec:origin}.
 
@@
 
-The first check on the plausibility of the candidate cluster members
-being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.
-Comparable analyses have already been performed by
+The first check on whether the candidate cluster members
+share the same age is to analyze the Gaia Hertzsprung-Russell (HR)
+diagrams.  Comparable analyses have previously been performed by
 \citetalias{cantatgaudin_gaia_2018},
@@
 Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
-$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
-members of the cluster show a clean sequence consistent with stars
-with a fixed age and metallicity, and varying mass.  The halo members
-show a consistent sequence, though with greater scatter.  One possible
+$G$ magnitude as a function of observed $Bp-Rp$ color.  Stars in
+the core appear consistent with
+having a fixed age and metallicity, and varying mass.  The halo stars
+show a similar sequence, though with greater scatter.  One possible
 explanation for the additional scatter could be that the halo is more
-contaminated by field stars.  For instance, $\approx5$ halo
-stars are present in portions of the red giant branch
-(RGB) that imply stellar ages inconsistent with that of the cluster.
-These stars must be field interlopers.
+contaminated by field stars.  For instance, $\approx5$ red
+giants in the halo
+must be field interlopers, because their isochronally implied
+ages would be inconsistent with that of the cluster.
 
@@
 
-To determine the reddening correction across the NGC\,2516 cluster, we
+To determine the reddening correction across \cn, we
 queried the 2MASS
@@
 
-Overall, the data and models agree for masses above approximately
-0.45\,$M_\odot$.  Below this mass, the data and models diverge at
+Overall, the data and models agree for masses above
+$\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
 colors redder than $Bp-Rp\approx2.2$, in the sense that the model
@@
 165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
-PARSEC, MIST, or SYCLIST).  Their statistical uncertainties ($\approx
-25$\,Myr) were comparable to the systematic uncertainties.
-
-Our primary goal is to determine whether the ages of the core and halo
-are consistent.  The left panel of Figure~\ref{fig:hr} shows that
+PARSEC, MIST, or SYCLIST).  %Their statistical uncertainties ($\approx
+%25$\,Myr) were comparable to the systematic uncertainties.
+
+Our goal is to determine whether the ages of the core and halo
+are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
 isochronally, they are: stars past the turnoff in both the halo and
@@
 can be made.  The MSTO stars are consistent with the 178 and 316 Myr
-models, but are bluer than the 100 Myr model.  The RGB stars are
+models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
 consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
@@
 the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
-and we refer to their work for a more precise model-dependent
+and we refer to their work for the more precise model-dependent
 comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
-adopt an PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
-our quoted uncertainty is a quadrature sum of their quoted statistical
-and systematic components.  The relative age of ``slightly older than
-the Pleiades'' is likely the more robust statement, consistent with
-reports since \citet{mermilliod_comparative_1981}.
+adopt a PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
+our quoted uncertainty is a quadrature sum of the statistical
+and systematic components from \cite{cummings_2018}. 
+%The relative age of ``slightly older than
+%the Pleiades'' is likely the more robust statement, consistent with
+%reports since \citet{mermilliod_comparative_1981}.
 
@@
     {\it Top row}: \cn\ split into its core and halo.  Sets
-    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
-    columns}) undergo successive degrees of automated cleaning,
-    described in Section~\ref{subsubsec:cluster}.  The Pleiades
+    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
+    (see Section~\ref{subsubsec:cluster}).  The Pleiades
     \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
-    \citep[650\,Myr;][]{douglas_poking_2017} are shown.
-    Stars in the core and the halo of NGC\,2516 appear
-    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
-    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
-    which some NGC\,2516 members appear slightly elevated with respect
-    to the Pleiades, suggesting an older age.  
-    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
-    are fainter than the magnitude limit of our survey.
+    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
+    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $(Bp-Rp)_0$$<$1.2.
+    From $1.2<(Bp-Rp)_0<1.7$, 
+    some NGC\,2516 members have longer rotation periods than Pleaides members.  
     {\it Bottom row}: \cn\
@@
 $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
-source list had $Rp<16$.  The difference is due to 34 stars unique to
+source list had $Rp<16$.  The difference (65 stars)
+is due to 35 stars from 
 \citet{meingast_2021} which were not available at the time of the
-CDIPS reductions, and a few stars falling on the TESS chip gaps.  The
+CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
 $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
@@
 Details regarding our detrending approach are presented in
-Appendix~\ref{app:detrending}.  After applying an initial detrending
-step aimed at removing systematic trends, we proceeding with a few
-small cleaning steps to improve the purity of the rotation period
-measurements: we masked 0.7 days at the beginning and end of each
+Appendix~\ref{app:detrending}.  After applying the initial
+ensemble decorrelation to remove systematic trends, we proceeded with a few
+cleaning steps: we masked 0.7 days at the beginning and end of each
 spacecraft orbit, and ran a sliding standard-deviation rejection
@@
 We then measured the rotation period from the resulting light curve
-using the periodogram implementations in \texttt{astrobase}, in
-particular the Stellingwerf phase-dispersion minimization periodogram,
-along with the more traditional Lomb-Scargle
+using the periodogram implementations in \texttt{astrobase}, namely
+the traditional Lomb-Scargle, and
+the Stellingwerf phase-dispersion minimization algorithm
 \citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
@@
 expectations, was expected to give the optimal balance between light
-from the target and background-light \citep{Sullivan_2015}.  This
+from the target star and background-light \citep{Sullivan_2015}.  This
 typically resulted in an aperture radius of either 1 or 1.5 pixels.
@@
 
-To clean the resulting rotation period measurements, we devised a set
+To clean the resulting rotation period measurements, we designed a set
 of automated cleaning criteria, which yielded sets we denote
@@
 $\mathcal{B}$, this completeness comes at the expense of greater
-contamination.  For instance, stars above the slow sequence (possible
-field contaminants) for the most part disappear in set $\mathcal{B}$.
+contamination.  For instance, stars above the slow sequence (perhaps
+field contaminants) mostly disappear in set $\mathcal{B}$.
 The ``void'' or ``gap'' underneath the slow sequence is also more
@@
 $0.5<(Bp-Rp)_0<1.2$.  The implications of this comparison for the age
-of \cn\ are assessed in greater detail in Section~\ref{disc:absage}.
+of \cn\ are discussed in Section~\ref{disc:absage}.
+%An important note is that stars redder than $(Bp-Rp)_0\gtrsim 2.2$
+%(later than $\approx$M2V)  are beyond the magnitude limit of our survey.
 
@@
 {\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
-photometric binarity, we fitted a spline to Figure~\ref{fig:hr} by
-hand, and labelled any points brighter than 0.3\,mag above the cluster
+photometric binarity, we fitted a spline to Figure~\ref{fig:hr} and fixed
+the nodes by hand.  We then labeled any points brighter than 0.3\,mag above the cluster
 sequence as photometric binaries.  This included 20\% of the overall
 cluster sample.  Many of the resulting candidate binaries, though not
-all of them, appear underneath the slow rotation sequence.  In set
+all of them, appear below the slow rotation sequence.  In set
 $\mathcal{B}$, over the color range where the slow and fast sequence
@@
 and ``slow'' sequences by eye,  the fraction of stars showing signs of
-binarity in the slow and fast rotation subsamples are 30\% (63/210)
+binarity in the slow and fast subsamples are 30\% (63/210)
 and 56\% (40/71), respectively.  We speculate on possible explanations
@@
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f8a.pdf}
+			\includegraphics[width=0.48\textwidth]{f8a.pdf}
 			\includegraphics[width=0.5\textwidth]{f8b.pdf}
 		}
+
+		\subfloat{
+			\includegraphics[width=0.9\textwidth]{f13.pdf}
+		}
+
 	\end{center}
@@
   \caption{ 
-  {\bf Members of \cn\ exist hundreds of parsecs from the core,
-  and up to $\approx$5\kms\ from the core in tangential velocity.}
-  Halo members that meet both kinematic and rotational membership
-  criteria (Subset~$\mathcal{A}$) are shown in blue.  Halo members for which rotation periods
-  should have been detectable, but were not detected, are shown in
-  orange.  
-  The plotted stars are restricted to $0.5<(Bp-Rp)_0<1.2$.
-  The {\it left panels} show galactic coordinates and
-  2D-tangential velocity space, after
-  correcting for projection effects
-  (Appendix~\ref{app:vproj}).
-  The histogram on the right bins versus 3D separation from the core
-  and 2D velocity difference.  Bins have width 25\,pc in position, and
-  1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties in the fraction
-  are taken as the quadrature sum of the rel unc of the numerator and
-  denominator (FIXME)}
+  {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
+  to $\approx$5\kms\ from the core in tangential velocity.}
+  {\it Top left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
+  2D-tangential velocities for halo members that meet kinematic
+  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
+  rotation periods are shown in blue;  halo members for which rotation
+  periods should have been detectable, but were not detected, are
+  shown in orange.  Stars in the core are shown as black points.
+  Computing the 2D-tangential velocity requires a correction for
+  projection effects (Appendix~\ref{app:vproj}).
+  In the {\it top right}, the histograms on bin versus 3D separation
+  from the core and 2D tangential velocity difference.  Bin widths are
+  chosen to enforce the same number of stars in the denominator (54)
+  per bin.
+  In the {\it bottom panels}, the innermost and outermost 100 stars in
+  position and velocity space are displayed in the space of rotation
+  periods and dereddened colors.
   \label{fig:physical_x_rotn}
@@
 
-% \begin{figure}[t]
-% 	\begin{center}
-% 			\includegraphics[width=0.49\textwidth]{f8b.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-%   \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
-%   versus 3d separation from the core and projection-corrected 2d velocity difference.} Bins
-%   have width 25\,pc in position, and 1.0\,km\,s$^{-1}$ in velocity.
-%   {\bf Uncertainties in the fraction are taken as the quadrature sum
-%   of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
-%   members appear to exist hundreds of parsecs from the core, and up to
-%   $\approx$5\kms in tangential velocity separation.
-%   \label{fig:hist_physical_x_rotn}
-% 	}
-% \end{figure}
-
-How far away from the core, in position and velocity space, does the
-halo extend?  We can explore this by crossmatching the stars with
-detected rotation periods against our original target list of
-\nkinematic\ Gaia DR2 kinematic candidate members.  For a fair
-comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
-which our TESS pipeline successfully made detrended light curves.  In
-other words---the base sample is the stars for which we could have
-plausibly detected a rotation period.  Faint stars with $Rp>16$ were
-beyond our light curve selection limit, and crowded stars ({\it e.g.,}
-near saturated stars in the cluster center) are similarly omitted.  We
-selected the color limit above based on Figure~\ref{fig:rot}, since
-the density of period detections seems to decrease once $(Bp-Rp)_0
-\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
-Our expectation is that completion effects become important for
-fainter stars.
+Given the rotation periods from TESS, we can explore how far away from
+the core we observe cluster members from the core in position and
+velocity space.  To do this, we cross-matched the stars with detected
+rotation periods against our original target list of \nkinematic\ Gaia
+DR2 kinematic candidate members.  For a fair comparison, we considered
+only the stars with $0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline
+successfully made detrended light curves.  In other words---the base
+sample is the stars for which we could have plausibly detected a
+rotation period.  Faint stars with $Rp>16$ were beyond our light curve
+selection limit, and stars not meeting the crowding requirements
+previously described are omitted.  We selected the color range above
+based on Figure~\ref{fig:rot}, since the density of period detections
+seems to decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for
+spectral types later than $\approx$K4V.
 
@@
 
-The results are shown using physical coordinates in
-Figure~\ref{fig:physical_x_rotn}.  Appendix~\ref{app:gaia6d_x_rotn}
-includes a supplemental visualization in the observed space of Gaia
-measurements.  The observed proper motions however suffer from
-projection effects, discussed in Appendix~\ref{app:vproj}.  These
+The results are shown in Figure~\ref{fig:physical_x_rotn}.
+Appendix~\ref{app:gaia6d_x_rotn} includes a supplemental visualization
+in the observed space of Gaia measurements, rather than Cartesian
+coordinates and tangential velocities.  The observed proper motions
+however suffer from projection effects, discussed in
+Appendix~\ref{app:vproj} {\bf TODO: merge these appendices}.  These
 effects are important in the equatorial coordinate system due to the
@@
 \citetalias{meingast_2021}.  We then convert these proper motion
-differences to physical units by dividing by the distance of each
-star, taken as the inverse parallax.  The inverse parallax
-approximation does not introduce significant biases for this sample,
-since the mean $\pi/\sigma_\pi$ and its standard deviation are 122 and
-40 respectively; the minimum is 19.5.  An inspection of the results
-suggests that the fraction of kinematic members with TESS-detected
-rotation periods may decrease with increasing distance from the halo
-core---both in position and in velocity space.
-
-% some wide separation pairs and dist_pc:
-% 5263077805266480000  5506439929389838592  421.083675
-% 5506623616550497792  5273740456715689088  427.053947
-% 5506439929389838592  5213536216500903680  427.557268
-% 5273740456715689088  5488984598182200704  428.952934
-% 5506439929389838592  5273740456715689088  438.906581
-
-To quantify this, we compute the separation of the positions and 2D
+differences to physical units by dividing by the distance to each
+star, taken as the inverse parallax.
+
+{\bf FIXME FIXME FIXME} 
+The results suggest that the
+fraction of kinematic members with TESS-detected rotation periods may
+decrease with increasing distance from the halo core---both in
+position and in velocity space.
+
+%The inverse parallax approximation does not introduce significant
+%biases for this sample, since the mean $\pi/\sigma_\pi$ and its
+%standard deviation are 122 and 40 respectively; the minimum is 19.5.
+
+We quantified this by computing the separation of the positions and 2D
 velocities from the median \citetalias{cantatgaudin_gaia_2018} core
-member.  The result is shown on the right side of
+member {\bf TODO: clarify binning choice}.  The result is shown on the right side of
 Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
@@
 
+{\bf FIXME update this text, make it less crappy... ditto the figure}
 In projection-corrected tangential velocity space, the fraction of
@@
 Since our main focus is measuring equivalent widths for the
-Li\,6708\,\AA doublet, these were the most relevant bit flags.  Of our
+Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of our
 \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.  51
@@
 stellar chromospheres: hotter stars fully ionize Li out of its ground
-state \citep[{\it e.g.},Figure~4 of][]{soderblom_evolution_1993}.  The
+state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
 depletion of stars redder than $(Bp-Rp)_0\gtrsim1.4$, {\it i.e.},
@@
 
+The main aim this study is to verify whether stars
+in the halo of \cn\ are the same age as stars in the core, and
+secondarily to quantify the contamination rate.
+Figure~\ref{fig:physical_x_rotn}
+
 \subsection{How did the halo form?}
@@
 
+The halo, comprised of a leading and trailing tail, extends 
+over at least 500\,pc.  Are these structures in fact tidal tails?  The
+canonical idea behind such a structure \citep[{\it e.g.},][]{krumholz_star_2019} is that
+stars escape out of the Lagrange points imposed by the Milky Way's
+potential, and then form leading and lagging arms due to differential
+rotation in the galaxy.  The leading and lagging arms of \cn\ 
+qualitatively match this picture, in that they are oriented
+roughly toward and away from the direction of galactic rotation.
+
+A second possible explanation invokes the idea that the cluster formed
+in a larger and more dispersed star formation complex, and that the
+stars currently in the halo need not have formed in the same ``clump''
+as those in the cluster core.  In other words, the formation
+environment might be better described a giant molecular filament,
+rather than a giant molecular cloud.  A sample of such filaments has
+recently been collected and analyzed by \citet{zucker_physical_2018}:
+given the current $\approx$20:1 aspect ratio of \cn, its structure
+would match the aspect ratios of what they term either ``elongated
+dense core complexes'', or ``bone candidates''.  A more immediate
+example could be the Orion\,A cloud.  \citet{grosschedl_3d_2018}
+showed using YSOs as tracers that Orion\,A is well-described as a
+90\,pc long filament with a dense ``head'', and a lower density 75\,pc
+long ``tail''. Whether the future evolutionary state of this filament,
+or those in the \citet{zucker_physical_2018} sample, could resemble
+something like \cn\ is a topic worthy of future numerical exploration.
+
+%The general shape of the cluster itself seems to include a central
+%overdensity and two tails (the halo).  One tail is leading the
+%cluster's orbit, and is angled toward the center of the galaxy when
+%viewed top-down.  The other tail is trailing the cluster's orbit, and
+%is pointed away from the center of the galaxy.  The elongation of the
+%cluster in both the $(X,Y)$ and $(Z,Y)$ planes is correlated with the
+%direction of the LSR-corrected median cluster velocity.   Possible
+%explanations for this overall morphology are discussed in
+%Section~\ref{subsec:origin}.
+
+TODO: quiver map velocities?
+
 Differential rotation in the galaxy is clearly important. Meingast+21
@@
 and trailing arms oriented in line with the differential rotation of
-the Galaxy.
-(See e.g., their Figure 13).
-
-Another way to frame this question: should we call these stars a tidal
-tail?
-% TODO: read Gagne+20 on this...
-The classical explanation of such a structure (e.g., Krumholz+19) is
-that stars escape out of L1 and L2, and then form leading and lagging
-arms due to differential rotation in the galaxy.
-The leading and lagging arms of NGC\,2516's halo qualitatively
-match this picture, in that they are oriented roughly toward and away
-the direction of galactic rotation (CITE, Meingast21, link to their
-website).
-
-A second explanation though invokes
-the idea that the cluster formed in a larger and lower-density star
-formation complex, and the stars we see in the halo did not in fact
-form the same ``clump'' as those in the cluster core.
-
-Consider for instance the Orion\,A cloud.  \citet{grosschedl_3d_2018} showed
-using YSOs that formed in the past few megayears as tracers that Orion\,A is
-best described with two components:
-a dense ``head'', and a lower density 75\,pc long ``tail''.
-While this filament currently totals 90\,pc in length, its future evolutionary state could very well resemble something like \cn.
-A direct numerical exploration would be necessary to be sure {\bf FIXME}.
-
-CITE ALSO: ZUCKER 2017 BONES CATALOG.
+the Galaxy.  (See e.g., their Figure 13).
+
+{\it Correlation between mean cluster velocity and elongation axes}
+Figure~\ref{fig:XYZ} shows that after subtracting the local standard
+of rest, the axes along which \cn\ is elongated are correlated with
+its average stellar velocity.
+A similar correlation was noted in Coma Ber by
+\citet{tang_comaber_2019}.
 
@@
 
-ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
-
-cometary-like cloud,
-oriented toward the Galactic plane, with two distinct components:
-a denser and enhanced star-forming (bent) Head, and a
-lower density and star-formation quieter 75 pc long Tail.
+% ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
+
+% cometary-like cloud,
+% oriented toward the Galactic plane, with two distinct components:
+% a denser and enhanced star-forming (bent) Head, and a
+% lower density and star-formation quieter 75 pc long Tail.
 
@@
 % formation complex that has since dispersed.''
-Another possibility is {\it triggered star formation}.
-
-One relevant connection may be to the ``Mamajek-2'' stellar group
-(see Jilinski+2009).
+% Another possibility is {\it triggered star formation}.
+
+% One relevant connection may be to the ``Mamajek-2'' stellar group
+% (see Jilinski+2009).
 
@@
 From their rotation period analysis of \cn,
-\citet{fritzewski_rotation_2020} have recently solidified our
-understanding that at ages between 100 and 200\,Myr, young open
-clusters ({Blanco\,1, the Pleiades, Psc\,Eri, M\,35, and M\,50) all
-seem to show similar rotation period distributions.  The FGK stars
+\citet{fritzewski_rotation_2020} have recently
+demonstrated that at ages between 100 and 200\,Myr, young open
+clusters ({\cn, Blanco\,1, the Pleiades, Psc\,Eri, M\,35, and M\,50) all
+show similar rotation period distributions.  The FGK stars
 show a ``slow sequence'', a population of ``rapid rotators'', and
 in some cases stars in the ``void'' between the two
-populations ({\it e.g.,}, Psc\,Eri and \cn\ show such populations).
+populations ({\it e.g.,}, Psc\,Eri and \cn\ show more stars in the void
+than the Pleiades).
 The late M dwarfs with detected rotation periods tend to rotate very
-rapidly ($<3$\,days) --- these are unfortunately beyond the magnitude
-limit of the present analysis.  In this section, we briefly discuss
-these features, and focus on the main difference between \cn\
-and the Pleiades: the longer maximum rotation periods seen in the
-K-dwarfs.
-
-\subsubsection{The slow and fast sequences}
+rapidly ($<3$\,days) --- these are beyond the magnitude
+limit of the present analysis.  This section briefly discusses
+the origins of these feature.
+
 The existence of a slow sequence is the main premise of gyrochronology
@@
 
-An important point regarding the lithium-rotation trend is that it has
-been shown to carry over beyond just equivalent widths, and to the
-actual abundance of lithium in the photosphere, A(Li).
+The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) 
+has been shown to carry over beyond
+just equivalent widths, and to the actual abundance of lithium in the
+photosphere, A(Li).
 \citet{soderblom_evolution_1993} calculated the relevant LTE curves of
-growth for the Li\,6708\,\AA\ doublet, which can in principle be used to
+growth for the Li\,6708\,\AA\ doublet, and used them to
 convert our measured equivalent widths to abundances.  Non-LTE
@@
 \label{disc:absage}
+
 Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
@@
 top row of Figure~\ref{fig:rot} shows that the two sequences overlap
-from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
+from $0.5<(Bp-Rp)_0<1.2$.  At redder colors
 ($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
-The maximum rotation periods seen in \cn across both sets
+The maximum rotation periods seen in \cn\ across both sets
 $\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
@@
   \item {\it Isochronal, rotational, and lithium age-dating show that
-    the halo of \cn\ is coeval with its core.} In other words, all
+    the halo of \cn\ is coeval with its core.} In short, all
     the data are consistent with the halo being real.  The Gaia EDR3
-    photometry and astrometry show a main-sequence turnoff consistent
-    with an isochronal age of 150\,Myr for both the core and halo
+    photometry and astrometry show a main-sequence turnoff that suggest
+    an isochronal age of 150\,Myr for both the core and halo
     (Figure~\ref{fig:hr}).  The faintest M dwarfs in the core and halo
@@
     (spectral types F2V-K3V) overlap with the slow sequence of the
-    Pleaides (Figure~\ref{fig:rot}).  At slightly redder
+    Pleaides (Figure~\ref{fig:rot}).  At redder
     colors, from $1.2<(Bp-Rp)_0<1.7$ (spectral types K3V-K6V,
@@
     vs.}\ 8 day rotation periods).
-    This also supports our {\bf age determination of $XXX$\,Myr for
-    the cluster}.
-  %
-  \item {\it Bonafide members exist out to $\approx$250\,pc in
-    separation from the cluster center, and out to $\approx$5\kms\ in
-    velocity separation.} Figure~\ref{fig:physical_x_rotn} shows the
+    This is consistent with \cn\ being slightly older than the Pleiades
+    ({\it i.e.}, $\approx$150\,Myr).
+    The lithium equivalent widths of the kinematic cluster members (Figure~\ref{fig:lithiumcorehalo})
+    are consistent with this assessment.
+  %
+  \item {\it Bonafide \cn\ members exist out to $\approx$250\,pc in
+    separation and out to $\approx$5\kms\ in
+    tangential velocity separation from the cluster core.} Figure~\ref{fig:physical_x_rotn} shows the
     overlap between kinematic and rotational cluster members used to
@@
     not as bad as one might fear.
+   %
+   \item {\it The rapid rotators show elevated lithium abundances, and elevated
+   binarity fractions}.
+   Each trend has been noted in comparable stellar populations ({\it e.g.,} \citealt{soderblom_evolution_1993}, \citealt{gillen_ngts_2020}).
+   The rough orders of magnitude of the effects are $\approx$1\,dex in lithium abundance 
+   for a factor of 10 in rotation period, and a factor of $\approx2\times$ enhancement
+   in binarity fraction for fast {\it vs.} slowly rotating G and early K dwarfs.
 \end{itemize}
 
-Broadly speaking, the existence of the halo is perhaps not very
+Generally speaking, the existence of the halo itself should not be not very
 surprising.  A star with a 1\kms\ velocity difference from the cluster
@@
 and~\ref{fig:hist_physical_x_rotn}).
-
-\section{Rotation $\otimes$ RUWE}
-\label{app:ruwe}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Rotation versus color, colored by RUWE.}
-		Looks like on the slow sequence, there's more yellow at the bottom?
-		\label{fig:rotn_X_RUWE}
-	}
-\end{figure*}
-
-Figure~\ref{fig:rotn_X_RUWE}.
-
-\section{Kinematics $\otimes$ Rotation in EDR3}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f7.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
-    velocity space, cross-matched against the rotators. Analog of
-    Figure~\ref{fig:gaia6d_x_rotn}, but showing EDR3 kinematics.}
-		\label{fig:gaia6d_x_rotn_EDR3}
-	}
-\end{figure*}
-
-How did we do?
-Figure~\ref{fig:gaia6d_x_rotn_EDR3} lets us check with a more precise
-astrometric solution.
+% 
+% 
+% 
+% \section{Rotation $\otimes$ RUWE}
+% \label{app:ruwe}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.95\textwidth]{f6.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{ {\bf Rotation versus color, colored by RUWE.}
+% 		Looks like on the slow sequence, there's more yellow at the bottom?
+% 		\label{fig:rotn_X_RUWE}
+% 	}
+% \end{figure*}
+% 
+% Figure~\ref{fig:rotn_X_RUWE}.
+% 
+% 
+% \section{Kinematics $\otimes$ Rotation in EDR3}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.95\textwidth]{f7.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
+%     velocity space, cross-matched against the rotators. Analog of
+%     Figure~\ref{fig:gaia6d_x_rotn}, but showing EDR3 kinematics.}
+% 		\label{fig:gaia6d_x_rotn_EDR3}
+% 	}
+% \end{figure*}
+% 
+% How did we do?
+% Figure~\ref{fig:gaia6d_x_rotn_EDR3} lets us check with a more precise
+% astrometric solution.
 

### 2021-04-06 cluster_paper/collect.sh

@@ -24,6 +24,8 @@ cp ../results/vtangential_projection/XYZ_with_orbit.pdf f11.pdf
 cp ../results/vtangential_projection/skymap_tangential_velocity_pmdec.pdf f12a.pdf
 cp ../results/vtangential_projection/skymap_tangential_velocity_pmra_cosdec.pdf f12b.pdf
 
+cp ../results/rotation_X_positions/rotation_X_positions_approach2_defaultcleaning.pdf f13.pdf
+
 # supplementary
 cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis_linear.pdf f6.pdf
 

### 2021-04-06 drivers/plot_rotation_X_positions.py

@@ -1,0 +1,22 @@
+import os
+import earhart.plotting as ep
+from earhart.paths import RESULTSDIR
+
+PLOTDIR = os.path.join(RESULTSDIR, 'rotation_X_positions')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+for approach in [2,1]:
+    for cleaning in ['defaultcleaning','periodogram_match']:
+        if approach == 1:
+            ep.plot_rotation_X_positions(PLOTDIR, cmapname='magma',
+                                         approach=approach, cleaning=cleaning)
+            ep.plot_rotation_X_positions(PLOTDIR, cmapname='cividis',
+                                         approach=approach, cleaning=cleaning)
+            ep.plot_rotation_X_positions(PLOTDIR, cmapname='viridis',
+                                         approach=approach, cleaning=cleaning)
+            ep.plot_rotation_X_positions(PLOTDIR, cmapname='plasma',
+                                         approach=approach, cleaning=cleaning)
+        else:
+            ep.plot_rotation_X_positions(PLOTDIR, cmapname=None, approach=approach,
+                                         cleaning=cleaning)

### 2021-04-06 earhart/plotting.py

@@ -14,6 +14,7 @@ Contents:
         plot_full_kinematics_X_rotation
         plot_physical_X_rotation (+ histogram_physical_X_rotation)
         plot_slowfast_photbinarity_comparison
+        plot_rotation_X_positions
     Lithium:
         plot_lithium_EW_vs_color
         plot_rotation_X_lithium
@@ -36,6 +37,10 @@ from numpy import array as nparr
 from scipy.interpolate import interp1d
 
 import matplotlib as mpl
+from matplotlib.ticker import (
+    MultipleLocator, AutoMinorLocator, LogLocator
+)
+
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 from astropy import units as u, constants as const
@@ -1114,7 +1119,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     colors = ['gray', 'gray', 'k']
     zorders = [-2, -3, -1]
     markers = ['s', 'x', 'o']
-    lws = [0., 0.3, 0.2]
+    lws = [0., 0.1, 0.1]
     mews= [0.5, 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
     ss = [5, 6, 10]
@@ -2487,6 +2492,207 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
     savefig(f, outpath)
 
 
+def plot_rotation_X_positions(outdir, basedata='fullfaint', cmapname=None,
+                              approach=None, cleaning='defaultcleaning'):
+    """
+    Prot vs (Bp-Rp)_0, colored by position in the cluster.
+    """
+
+    if approach == 1:
+        assert isinstance(cmapname, str)
+    assert isinstance(approach, int)
+
+    _, core_df, _, full_df, trgt_df = get_gaia_basedata(basedata)
+    rot_df, lc_df = get_autorotation_dataframe(
+        runid='NGC_2516', returnbase=True, cleaning=cleaning
+    )
+    med_df, _ = _get_median_ngc2516_core_params(core_df, basedata)
+
+    from earhart.physicalpositions import append_physicalpositions
+    full_df = append_physicalpositions(full_df, med_df)
[…]

### 2021-04-06 cluster_paper/ms.tex
update ms

@@
 
-\subsubsection{Cluster Star Sample}
+\subsubsection{Methodology}
 \label{subsubsec:cluster}
@@
 \nautorotnumeratormatching\ light curves.
+
+\subsubsection{Rotation-Color Diagrams}
 
@@
 
-\subsubsection{Kinematics $\otimes$ Rotation}
+\subsubsection{Comparing Kinematics and Rotation}
 
@@
   {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
-  to $\approx$5\kms\ from the core in tangential velocity.}
-  {\it Top left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
+  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
+  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
   2D-tangential velocities for halo members that meet kinematic
@@
   Computing the 2D-tangential velocity requires a correction for
-  projection effects (Appendix~\ref{app:vproj}).
-  In the {\it top right}, the histograms on bin versus 3D separation
-  from the core and 2D tangential velocity difference.  Bin widths are
-  chosen to enforce the same number of stars in the denominator (54)
-  per bin.
-  In the {\it bottom panels}, the innermost and outermost 100 stars in
+  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
+  right}, the histograms on bin versus 3D separation from the core and
+  2D tangential velocity difference.  Bin widths are chosen to enforce
+  the same number of stars in the denominator (54) per bin.  In the
+  {\it bottom panels}, the innermost and outermost 100 stars in
   position and velocity space are displayed in the space of rotation
@@
 
-Given the rotation periods from TESS, we can explore how far away from
-the core we observe cluster members from the core in position and
-velocity space.  To do this, we cross-matched the stars with detected
-rotation periods against our original target list of \nkinematic\ Gaia
-DR2 kinematic candidate members.  For a fair comparison, we considered
-only the stars with $0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline
-successfully made detrended light curves.  In other words---the base
-sample is the stars for which we could have plausibly detected a
-rotation period.  Faint stars with $Rp>16$ were beyond our light curve
-selection limit, and stars not meeting the crowding requirements
-previously described are omitted.  We selected the color range above
-based on Figure~\ref{fig:rot}, since the density of period detections
-seems to decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for
-spectral types later than $\approx$K4V.
-
-% IN POSITION
-% Bin width: 25
-% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
-% 0   0   0 0   0]
-% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
-% 0   2   0   0 0   0]
-% Rotators < 25pc : 0.72929 +/- 0.00013
-% Rotators > 25pc : 0.66172 +/- 0.00023
-%
-
-The results are shown in Figure~\ref{fig:physical_x_rotn}.
-Appendix~\ref{app:gaia6d_x_rotn} includes a supplemental visualization
-in the observed space of Gaia measurements, rather than Cartesian
-coordinates and tangential velocities.  The observed proper motions
-however suffer from projection effects, discussed in
-Appendix~\ref{app:vproj} {\bf TODO: merge these appendices}.  These
-effects are important in the equatorial coordinate system due to the
-proximity of some \cn\ members to the southern ecliptic pole.  For a
-star at position $(\alpha, \delta)$, we correct for the projection
-effect by comparing the observed proper motion $(\mu_{\alpha'},
-\mu_\delta)$ to what the proper motion at the star's position would
-have been if the star were comoving with the core of \cn.  This yields
-a quantity we denote $\Delta \mu^{*}$, following
-\citetalias{meingast_2021}.  We then convert these proper motion
-differences to physical units by dividing by the distance to each
-star, taken as the inverse parallax.
-
-{\bf FIXME FIXME FIXME} 
-The results suggest that the
-fraction of kinematic members with TESS-detected rotation periods may
-decrease with increasing distance from the halo core---both in
-position and in velocity space.
-
-%The inverse parallax approximation does not introduce significant
-%biases for this sample, since the mean $\pi/\sigma_\pi$ and its
-%standard deviation are 122 and 40 respectively; the minimum is 19.5.
-
-We quantified this by computing the separation of the positions and 2D
-velocities from the median \citetalias{cantatgaudin_gaia_2018} core
-member {\bf TODO: clarify binning choice}.  The result is shown on the right side of
-Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
-about 200\,pc in physical space from the cluster core.  The most
-widely separated F2V-K2V stars with rotation periods (blue points in
+The TESS rotation periods can provide an independent check on the
+Gaia-derived kinematic cluster memberships.  We explored this by
+cross-matching the stars with detected TESS rotation periods against
+our original target list of \nkinematic\ Gaia DR2 members.
+
+To visualize the results, we again opted for Cartesian coordinates,
+and supplemented them by calculating the tangential stellar velocities
+relative to the cluster center.  Appendix~\ref{app:vproj} discusses
+the projection effect correction required for calculating the
+tangential velocities: for a star at position $(\alpha, \delta)$, we
+compare the observed proper motion $(\mu_{\alpha'}, \mu_\delta)$ with
+what the proper motion at the star's position would have been if the
+star were comoving with the core of \cn.  This yields a quantity we
+denote $\Delta v^{*}$, per \citetalias{meingast_2021}.  We convert
+these proper motion differences to physical units by dividing by the
+distance to each star, taken as the inverse parallax.
+
+The results are shown in the top two rows of
+Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
+comparison, we only show stars with $0.5<(Bp-Rp)_0<1.2$ for which our
+TESS pipeline succeeding in making light curves.  In other words---the
+base sample is the stars for which we could have plausibly detected a
+rotation period.  This color range is preferred since the the slow
+sequence becomes less defined for spectral types later than
+$\approx$K4V.  The bins in the histograms (top-right) were chosen to
+ensure that the same number of stars were in each bin.  Rotation
+periods are detected for 66\% of stars at least 25\,pc from the
+cluster core.  The average detection rate inside of 25\,pc, 73\%, is
+slightly higher.
+
+The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
+view of how field star contamination affects the outskirts of the
+cluster.  Of the outermost 100 stars in position space with detected
+rotation periods, $\approx$8 appear as outliers above the slow
+sequence, compared to just a few for the innermost 100 stars.  For the
+outermost 100 stars in velocity space, $\approx$13 appear as outliers.
+After accounting for the stars that were expected to show rotation
+periods and did not, this suggests field contaminant rates of
+$\approx$40-50\% for the outermost cluster members in our target list.
+
+Speaking in the positive though, our rotation period measurements show
+that the halo extends to separations of $\approx$250\,pc in physical
+space from the cluster core.  The most widely separated F2V-K2V stars
+with rotation periods (blue points in
 Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
@@
 on which members of the halo are chosen as the ``tips'' on either end.
-Broadly speaking, this agrees with the structure of the halo reported
-by \citetalias{kounkel_untangling_2019}.
-
-{\bf FIXME update this text, make it less crappy... ditto the figure}
+This agrees with the overall structure of the halo reported by
+\citetalias{kounkel_untangling_2019}, with some minor caveats visible
+in the top-left panels of Figure~\ref{fig:physical_x_rotn}.
+
 In projection-corrected tangential velocity space, the fraction of
@@
 show that at the expense of higher field star contamination rates,
-bonafide members can be identified even out to higher velocity
-separations.  {\bf todo: anything else to mention? e.g., the actual
-numbers?}
+bonafide members can be identified even out at higher velocity
+separations.
 
@@
 
-\section{Kinematics $\otimes$ Rotation}
-\label{app:gaia6d_x_rotn}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Gaia-based components of NGC\,2516 in position and
-  velocity space, cross-matched against the rotators.} The plotted
-  stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
-  restrictions described in Section~\ref{subsec:tess}: they should
-  have been sufficiently bright and non-crowded to enable rotation
-  period detection.  We caution that the appearance of fewer
-  non-rotators being present toward the core is due the layering of
-  the plot: quantitatively, stars toward the cluster center do not all
-  show rotation periods in our analysis (see
-  Figure~\ref{fig:hist_physical_x_rotn}).
-  \label{fig:gaia6d_x_rotn}
-	}
-\end{figure*}
-
-Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
-data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
-in this diagram correspond to ``Set $\mathcal{A}$'' as described in
-Section~\ref{subsec:tess}.  Many stars in the outer region of the
-cluster show rotation periods consistent with a young age.  One issue
-with the visualization as displayed however is that it hides the
-amount of non-rotators toward the cluster center: we did not detect
-rotation periods for roughly one in four stars at the cluster
-center (see Figures~\ref{fig:physical_x_rotn}
-and~\ref{fig:hist_physical_x_rotn}).
+% \section{Kinematics $\otimes$ Rotation}
+% \label{app:gaia6d_x_rotn}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.95\textwidth]{f4.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+%   \caption{ {\bf Gaia-based components of NGC\,2516 in position and
+%   velocity space, cross-matched against the rotators.} The plotted
+%   stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
+%   restrictions described in Section~\ref{subsec:tess}: they should
+%   have been sufficiently bright and non-crowded to enable rotation
+%   period detection.  We caution that the appearance of fewer
+%   non-rotators being present toward the core is due the layering of
+%   the plot: quantitatively, stars toward the cluster center do not all
+%   show rotation periods in our analysis (see
+%   Figure~\ref{fig:hist_physical_x_rotn}).
+%   \label{fig:gaia6d_x_rotn}
+% 	}
+% \end{figure*}
+% 
+% Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
+% data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
+% in this diagram correspond to ``Set $\mathcal{A}$'' as described in
+% Section~\ref{subsec:tess}.  Many stars in the outer region of the
+% cluster show rotation periods consistent with a young age.  One issue
+% with the visualization as displayed however is that it hides the
+% amount of non-rotators toward the cluster center: we did not detect
+% rotation periods for roughly one in four stars at the cluster
+% center (see Figures~\ref{fig:physical_x_rotn}
+% and~\ref{fig:hist_physical_x_rotn}).
 % 

### 2021-04-06 cluster_paper/ms.tex
update ms

@@
 \newcommand{\ms}{\,m\,s$^{-1}$}
+\newcommand{\bpmrp}{(\mathrm{Bp}-\mathrm{Rp})_0}
 
@@
   with the visual overdensity of stars canonically accepted as ``the
-  cluster''.  The halo (blue) is a concatenation of the studies by
+  cluster''.  The halo (blue) is a concatenation of studies by
   \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
@@
     atmospheres is less well-calibrated for $M_\star \lesssim
-    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
+    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2); these model values
     are shown with dashed lines.
@@
 Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
-$G$ magnitude as a function of observed $Bp-Rp$ color.  Stars in
+$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$ color.  Stars in
 the core appear consistent with
@@
 NGC\,2516 member.  The mean and standard
-deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
+deviation of the $E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
 and \citet{schlafly_measuring_2011} maps were consistent, so we
-adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
+adopted the average from \citet{schlegel_maps_1998}: $E(\mathrm{B}-\mathrm{V})_{\rm
 S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
@@
 proposed by \citet{bonifacio_search_2000} leads to our adopted
-$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
-the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
-E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
+$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
+the calibration of \citet{stassun_TIC8_2019}, namely $E(\mathrm{Bp}-\mathrm{Rp})=1.31
+E(\mathrm{B}-\mathrm{V})$ and $A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded $E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$,
 which is used in the plots.  To redden the isochrones, we assumed
@@
 For the metallicity, we considered a range of super and sub-solar
-metallicities to fit as much of the locus from $0.5<Bp-Rp<1.5$ as
+metallicities to fit as much of the locus from $0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as
 possible, and settled on the slightly super-solar $[M/H]=0.06$
@@
 $\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
-colors redder than $Bp-Rp\approx2.2$, in the sense that the model
+colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in the sense that the model
 isochrones have lower luminosities and bluer colors than the
@@
     \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
-    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $(Bp-Rp)_0$$<$1.2.
-    From $1.2<(Bp-Rp)_0<1.7$, 
+    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $\bpmrp$$<$1.2.
+    From $1.2<\bpmrp<1.7$, 
     some NGC\,2516 members have longer rotation periods than Pleaides members.  
@@
 $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
-roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
+roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at
 the distance of \cn.
@@
 \cn\ appear gyrochronologically coeval for colors spanning
-$0.5<(Bp-Rp)_0<1.2$.  The implications of this comparison for the age
+$0.5<\bpmrp<1.2$.  The implications of this comparison for the age
 of \cn\ are discussed in Section~\ref{disc:absage}.
-%An important note is that stars redder than $(Bp-Rp)_0\gtrsim 2.2$
+%An important note is that stars redder than $\bpmrp\gtrsim 2.2$
 %(later than $\approx$M2V)  are beyond the magnitude limit of our survey.
@@
 {\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
-photometric binarity, we fitted a spline to Figure~\ref{fig:hr} and fixed
+photometric binarity, we fitted a spline to Figure~\ref{fig:hr}, fixing
 the nodes by hand.  We then labeled any points brighter than 0.3\,mag above the cluster
@@
 $\mathcal{B}$, over the color range where the slow and fast sequence
-are present ($0.5<(Bp-Rp)_0<1.3$), there are 281 stars, of which 103
+are present ($0.5<\bpmrp<1.3$), there are 281 stars, of which 103
 (37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
@@
   2D tangential velocity difference.  Bin widths are chosen to enforce
-  the same number of stars in the denominator (54) per bin.  In the
+  the same number of stars in the denominator ($N=54$) per bin.  In the
   {\it bottom panels}, the innermost and outermost 100 stars in
@@
 Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
-comparison, we only show stars with $0.5<(Bp-Rp)_0<1.2$ for which our
+comparison, we only show stars with $0.5<\bpmrp<1.2$ for which our
 TESS pipeline succeeding in making light curves.  In other words---the
@@
 
-Speaking in the positive though, our rotation period measurements show
+Even with contaminants, our rotation period measurements show
 that the halo extends to separations of $\approx$250\,pc in physical
@@
   Equivalent width of the 6708\,\AA\ doublet is shown versus
-  extinction-corrected color for all kinematic \cn\ members with
+  dereddened color for all kinematic \cn\ members with
   Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
@@
 Gaia-ESO spectrum was acquired and stellar parameters were reported.
-15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
+15 of these (all with $\bpmrp > 2.0$) are spurious matches based on
 the Gaia color and GES effective temperature, and we remove them.
@@
 stars are in the core or the halo of the cluster.  The GALAH spectra
-span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
+span a color range of $0<\bpmrp<1.5$, due to the $V<14$ brightness
 cutoff of the survey.  The overall increase of Li EW from
-$0<(Bp-Rp)_0<1.0$ is thought to be driven by the temperatures in the
+$0<\bpmrp<1.0$ is thought to be driven by the temperatures in the
 stellar chromospheres: hotter stars fully ionize Li out of its ground
 state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
-depletion of stars redder than $(Bp-Rp)_0\gtrsim1.4$, {\it i.e.},
+depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.},
 later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
@@
 
-\subsection{Contamination fraction across the halo}
-
-The main aim this study is to verify whether stars
-in the halo of \cn\ are the same age as stars in the core, and
-secondarily to quantify the contamination rate.
-Figure~\ref{fig:physical_x_rotn}
+% \subsection{Contamination fraction across the halo}
+% 
+% The main aim this study is to verify whether stars
+% in the halo of \cn\ are the same age as stars in the core, and
+% secondarily to quantify the contamination rate.
+% Figure~\ref{fig:physical_x_rotn}
 
@@
 
-The halo, comprised of a leading and trailing tail, extends 
-over at least 500\,pc.  Are these structures in fact tidal tails?  The
-canonical idea behind such a structure \citep[{\it e.g.},][]{krumholz_star_2019} is that
+Any theory for the structure of \cn, the Psc\,Eri stream, and
+the strings and halos found by
+\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
+needs to explain a few observations.  First is their aspect
+ratios.  The halo of \cn, for instance,
+extends over at least 500\,pc, despite being only $\approx$25\,pc
+wide.  Second is their axial tilts in the plane of the Milky Way -- why
+their leading and trailing arms tend to conform to the Galaxy's
+differential rotation ({\it e.g.}, Figure~\ref{fig:XYZ} or Figure~13
+of \citetalias{meingast_2021}).  Third is the correlation between the
+mean LSR-subtracted cluster velocity and elongation axes (blue arrows in
+Figure~\ref{fig:XYZ}).  Such a correlation has also been noted in Coma
+Ber by \citet{tang_comaber_2019}.  The other halos may show
+similar trends.
+
+One possible explanation is that these structures are tidal tails
+\citep[{\it e.g.},][]{krumholz_star_2019}. The idea would be that
 stars escape out of the Lagrange points imposed by the Milky Way's
 potential, and then form leading and lagging arms due to differential
-rotation in the galaxy.  The leading and lagging arms of \cn\ 
-qualitatively match this picture, in that they are oriented
-roughly toward and away from the direction of galactic rotation.
-
-A second possible explanation invokes the idea that the cluster formed
-in a larger and more dispersed star formation complex, and that the
-stars currently in the halo need not have formed in the same ``clump''
-as those in the cluster core.  In other words, the formation
-environment might be better described a giant molecular filament,
-rather than a giant molecular cloud.  A sample of such filaments has
-recently been collected and analyzed by \citet{zucker_physical_2018}:
-given the current $\approx$20:1 aspect ratio of \cn, its structure
-would match the aspect ratios of what they term either ``elongated
-dense core complexes'', or ``bone candidates''.  A more immediate
-example could be the Orion\,A cloud.  \citet{grosschedl_3d_2018}
-showed using YSOs as tracers that Orion\,A is well-described as a
-90\,pc long filament with a dense ``head'', and a lower density 75\,pc
-long ``tail''. Whether the future evolutionary state of this filament,
-or those in the \citet{zucker_physical_2018} sample, could resemble
-something like \cn\ is a topic worthy of future numerical exploration.
-
-%The general shape of the cluster itself seems to include a central
-%overdensity and two tails (the halo).  One tail is leading the
-%cluster's orbit, and is angled toward the center of the galaxy when
-%viewed top-down.  The other tail is trailing the cluster's orbit, and
-%is pointed away from the center of the galaxy.  The elongation of the
-%cluster in both the $(X,Y)$ and $(Z,Y)$ planes is correlated with the
-%direction of the LSR-corrected median cluster velocity.   Possible
-%explanations for this overall morphology are discussed in
-%Section~\ref{subsec:origin}.
-
-TODO: quiver map velocities?
-
-Differential rotation in the galaxy is clearly important. Meingast+21
-established this; in ten different cluster halos, nine showed leading
-and trailing arms oriented in line with the differential rotation of
-the Galaxy.  (See e.g., their Figure 13).
-
-{\it Correlation between mean cluster velocity and elongation axes}
-Figure~\ref{fig:XYZ} shows that after subtracting the local standard
-of rest, the axes along which \cn\ is elongated are correlated with
-its average stellar velocity.
-A similar correlation was noted in Coma Ber by
-\citet{tang_comaber_2019}.
-
-CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
-THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
-LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
-
-Similarly Zari et al 2019 studied the entire Orion complex, and found that
-different regions of the complex (one toward the Belt, one toward 25\,Ori)
-began forming stars roughly simultaneously. No evidence was found of it
-propagating from these initial sites.
-
-% ARE THE MOLECULAR CLOUDS THEMSELVES STRUCTURED INTO CORES AND TAILS?
+rotation in the Galaxy.
+In this context, Psc\,Eri for instance might represent the completion
+of a process that \cn\ is still undergoing.
+
+A second possibility is that the clusters form in larger and more
+dispersed star formation complexes: the stars in the halo need
+not have formed in the same ``clump'' as those in the cluster core.
+In other words, the formation environment might be better described a
+giant molecular filament, rather than a giant molecular cloud.  A
+sample of such filaments has recently been collected and analyzed by
+\citet{zucker_physical_2018}: if the current $\approx$20:1 aspect
+ratio of \cn\ were primordial, its structure would match the aspect
+ratios of what they term either ``elongated dense core complexes'', or
+``bone candidates''.  A more immediate example is the Orion\,A cloud.
+\citet{grosschedl_3d_2018} showed using young stellar objects as tracers that Orion\,A
+is 90\,pc long, and that it has a dense ``head'' and
+a lower density ``tail''.
+
+One way to rule between the two possibilities
+would be 
+to observationally establish an age-size correlation.
+If the $\approx$100\,Myr halos are produced primarily through
+evaporation and expansion, then the
+younger clusters should be significantly smaller.
+A theoretical approach along similar lines
+would be to perform a dynamical analysis
+of whether filaments like Orion\,A, or those
+in the \citet{zucker_physical_2018} sample, are at all expected 
+to evolve into structures
+resembling \cn.
+
+% Possible observational directions could include analyzing whether the
+% Gaia kinematics contain any evidence of cluster expansion,
+% as well as analyzing the mass functions of stars in different
+% regions of the clusters to determine whether they contain
+
+% CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
+% THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
+% LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
+% 
+% Similarly Zari et al 2019 studied the entire Orion complex, and found that
+% different regions of the complex (one toward the Belt, one toward 25\,Ori)
+% began forming stars roughly simultaneously. No evidence was found of it
+% propagating from these initial sites.
 
@@
 From their rotation period analysis of \cn,
-\citet{fritzewski_rotation_2020} have recently
-demonstrated that at ages between 100 and 200\,Myr, young open
-clusters ({\cn, Blanco\,1, the Pleiades, Psc\,Eri, M\,35, and M\,50) all
-show similar rotation period distributions.  The FGK stars
-show a ``slow sequence'', a population of ``rapid rotators'', and
-in some cases stars in the ``void'' between the two
-populations ({\it e.g.,}, Psc\,Eri and \cn\ show more stars in the void
-than the Pleiades).
-The late M dwarfs with detected rotation periods tend to rotate very
-rapidly ($<3$\,days) --- these are beyond the magnitude
-limit of the present analysis.  This section briefly discusses
-the origins of these feature.
+\citet{fritzewski_rotation_2020} have recently demonstrated that at
+ages between 100 and 200\,Myr, young clusters ({\cn, Blanco\,1,
+the Pleiades, Psc\,Eri, M\,35, and M\,50) all show similar rotation
+period distributions.  The FGK stars show a ``slow sequence'', a
+population of ``rapid rotators'', and in some cases stars in the
+``void'' between the two populations ({\it e.g.,}, Psc\,Eri and \cn\
+show more stars in the void than the Pleiades).  The late M dwarfs
+with detected rotation periods tend to rotate very rapidly
+($<3$\,days), although these are beyond the magnitude limit of our
+analysis.  This section briefly discusses the origins of these
+features in the rotation-color diagram.
 
@@
 spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
-\citet{skumanich_time_1972} has not worked in detail, the 
-general idea is correct, and the first-order effect can be understood through magnetic
-breaking \citep{weber_angular_1967}.
-%More systematic efforts at age-dating stars through their rotation periods.
-
-The other features of the rotation-color diagram are in some sense
-affronts to the magnetic breaking model, and the general question of
-what produces the rapid rotators has not yet been definitively
+\citet{skumanich_time_1972} has not worked in detail, the general idea
+is correct, and the first-order effect can be understood through
+magnetic breaking \citep{weber_angular_1967}.
+
+The other features of the rotation-color diagram are indications
+that magnetic breaking is not always the most important effect.
+The question of why the rapid rotators exist, for instance, has not yet been definitively
 answered.  Models that introduce variable core-envelope decoupling
@@
 by {\it e.g.}, \citet{gillen_ngts_2020} and
-\citet{simonian_rapid_2019} have reported that rapid
-rotation in young clusters and in the field is correlated with
-binarity.  Related effects could include tidal synchronization for the
-shortest period binaries, or alternatively disk truncation that when
-combined with magnetic locking {\bf (CITE)} could produce the rotation
-period bimodality.
+\citet{simonian_rapid_2019} have reported that rapid rotation in young
+clusters and in the field is correlated with binarity.  Related
+effects could include tidal synchronization for the shortest period
+binaries, or alternatively disk truncation that when combined with
+magnetic locking {\bf (CITE)} could produce the rotation period
+bimodality.
 
 The lower panels of Figure~\ref{fig:rot} show our attempt at
-identifying unresolved binaries in the \cn\ rotation sample.
-Binaries resolved by Gaia were already excluded in our initial
-definitions of the samples.
-In \cn\, we see that
-the fraction of stars showing signs of binarity in the slow and fast
-rotation subsamples are 30\% (63/210) and 56\% (40/71); the fast
-rotators show a preference for binarity.
+identifying unresolved binaries in the \cn\ rotation sample.  Binaries
+resolved by Gaia were already excluded in our initial definitions of
+the samples.  In \cn\, we see that the fraction of stars showing signs
+of binarity in the slow and fast rotation subsamples are 30\% (63/210)
+and 56\% (40/71); the fast rotators show a preference for binarity.
 This correlation is in line with the earlier findings noted above,
@@
 stars are binaries, and $\approx$9\% have periods below 100\,days
-\citep{raghavan_survey_2010}.  If we assume  that all
-binaries with sub-100\,day periods become tidally synchronized, we
-can only explain a rapid rotator occurrence rate of
-$\approx$5\%.
+\citep{raghavan_survey_2010}.  If we assume  that all binaries with
+sub-100\,day periods become tidally synchronized, we can only explain
+a rapid rotator occurrence rate of $\approx$5\%.
 
 A separate effect is that on the slow rotation sequence,
-Figure~\ref{fig:rot}} shows that the
-binary stars appear to be either preferentially redder, or to have
-faster rotation periods than the single stars.  One likely
-explanation for this could be that the unresolved binaries have a
-component contributing additional red light to the system, skewing
-the color measurement of the primary.  Whether any physical
-effects could be at play remains a question for future
+Figure~\ref{fig:rot}} shows that the binary stars appear to be either
+preferentially redder, or to have faster rotation periods than the
+single stars.  One likely explanation for this could be that the
+unresolved binaries have a component contributing additional red light
+to the system, skewing the color measurement of the primary.  Whether
+any physical effects could be at play remains a question for future
 work.
-
-% \subsubsection{The slowest K dwarfs of NGC\,2516}
-% %FIXME need to rewrite this paragraph? (Or maybe not, because it's
-% %already in the absolute age section...
-% A few conclusions can be drawn by comparing the \cn\ rotation periods
-% with those of the Pleiades and Praesepe.  The first is that in both
-% stages of cleaning, the core and halo of \cn\ each show
-% gyrochronological ``slow sequences''.  These sequences appear to be
-% coeval with each other, and at first order appear to overlap
-% with the Pleiades sequence.
-% 
-%   \paragraph{Stalling}
-%   Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
-%   spindown is not monotonic in time.
-%   We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
-%   much change.
-
-%   \paragraph{Are the ``very slow rotators'' bonafide members?}
-%   Unsure. They are not isotropically distributed around the
-%   cluster... so either triggered star formation (CITE, CITE), or they're
-%   actually field stars.
 
@@
 
-The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) 
-has been shown to carry over beyond
-just equivalent widths, and to the actual abundance of lithium in the
-photosphere, A(Li).
+The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
+been shown to carry over beyond just equivalent widths, and to the
+actual abundance of lithium in the photosphere, A(Li).
 \citet{soderblom_evolution_1993} calculated the relevant LTE curves of
-growth for the Li\,6708\,\AA\ doublet, and used them to
-convert our measured equivalent widths to abundances.  Non-LTE
-corrections affect the abundances by $<0.5\,{\rm dex}$ and do not
-siginificantly affect the global relation between Li abundance and
-stellar temperature or rotation
-\citep{carlsson_1994,lind_departures_2009}.
+growth for the Li\,6708\,\AA\ doublet, and used them to convert our
+measured equivalent widths to abundances.  Non-LTE corrections affect
+the abundances by $<0.5\,{\rm dex}$ and do not siginificantly affect
+the global relation between Li abundance and stellar temperature or
+rotation \citep{carlsson_1994,lind_departures_2009}.
 
@@
 top row of Figure~\ref{fig:rot} shows that the two sequences overlap
-from $0.5<(Bp-Rp)_0<1.2$.  At redder colors
-($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
+from $0.5<\bpmrp<1.2$.  At redder colors
+($1.2<\bpmrp<1.7$) the dispersion in rotation periods increases.
 The maximum rotation periods seen in \cn\ across both sets
@@
 models are empirically calibrated using the Pleiades and Praesepe.  In
-other words, at $(Bp-Rp)_0=1.35$, we observe rotation periods in \cn\
+other words, at $\bpmrp=1.35$, we observe rotation periods in \cn\
 ranging from 7 to 10 days.  The 7 day rotation periods are consistent
@@
 not.  Fitting formulae from the previously cited studies give ages for
-a star with $\{(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}\}$ of 204
+a star with $\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204
 Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
@@
 	%
-  \item {\it Isochronal, rotational, and lithium age-dating show that
-    the halo of \cn\ is coeval with its core.} In short, all
-    the data are consistent with the halo being real.  The Gaia EDR3
-    photometry and astrometry show a main-sequence turnoff that suggest
-    an isochronal age of 150\,Myr for both the core and halo
+  \item {\it Isochronal, rotational, and lithium dating show that
+    the halo of \cn\ is coeval with its core.} In short, all the data
+    are consistent with the halo being real.  The Gaia EDR3 photometry
+    and astrometry show a main-sequence turnoff that suggest an
+    isochronal age of 150\,Myr for both the core and halo
     (Figure~\ref{fig:hr}).  The faintest M dwarfs in the core and halo
     are also brighter than comparison field stars of the same color.
-    Gyrochronally, stars in \cn\ spanning $0.5<(Bp-Rp)_0<1.2$
+    Gyrochronally, stars in \cn\ spanning $0.5<\bpmrp<1.2$
     (spectral types F2V-K3V) overlap with the slow sequence of the
-    Pleaides (Figure~\ref{fig:rot}).  At redder
-    colors, from $1.2<(Bp-Rp)_0<1.7$ (spectral types K3V-K6V,
-    $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
-    dispersion in rotation period.  The upper envelope of the rotation
-    period distribution at these redder colors extends to
-    longer rotation periods than in the Pleiades (11 {\it
-    vs.}\ 8 day rotation periods).
-    This is consistent with \cn\ being slightly older than the Pleiades
-    ({\it i.e.}, $\approx$150\,Myr).
-    The lithium equivalent widths of the kinematic cluster members (Figure~\ref{fig:lithiumcorehalo})
-    are consistent with this assessment.
+    Pleaides (Figure~\ref{fig:rot}).  At redder colors, from
+    $1.2<\bpmrp<1.7$ (spectral types K3V-K6V, $M_\star$$\approx$
+    0.6-0.7\,$M_\odot$), there is a larger dispersion in rotation
+    period.  The upper envelope of the rotation period distribution at
+    these redder colors extends to longer rotation periods than in the
+    Pleiades (11 {\it vs.}\ 8 day rotation periods).  This is
+    consistent with \cn\ being slightly older than the Pleiades ({\it
+    i.e.}, $\approx$150\,Myr).  The lithium equivalent widths of the
+    kinematic cluster members (Figure~\ref{fig:lithiumcorehalo}) are
+    consistent with this assessment.
   %
   \item {\it Bonafide \cn\ members exist out to $\approx$250\,pc in
-    separation and out to $\approx$5\kms\ in
-    tangential velocity separation from the cluster core.} Figure~\ref{fig:physical_x_rotn} shows the
-    overlap between kinematic and rotational cluster members used to
-    make this assessment.
+    separation and out to $\approx$5\kms\ in tangential velocity
+    separation from the cluster core.}
+    Figure~\ref{fig:physical_x_rotn} shows the overlap between
+    kinematic and rotational cluster members used to make this
+    assessment.
   %
@@
     physical and velocity separations from the cluster core.} The
-    right panel of Figure~\ref{fig:physical_x_rotn} quantifies the
-    statement: $\approx$75\% of the ``inner-most'' members show
-    rotation periods consistent with the age of the cluster.  This
-    fraction decreases to $\approx$60\% at physical separations of
-    $\approx$200\,pc.  Given the number density ratios of cluster to
-    field stars at these wide separations ($\approx 10^{-3}$,
-    \citealt{meingast_2021}), the implied contamination fractions are
-    not as bad as one might fear.
-   %
-   \item {\it The rapid rotators show elevated lithium abundances, and elevated
-   binarity fractions}.
-   Each trend has been noted in comparable stellar populations ({\it e.g.,} \citealt{soderblom_evolution_1993}, \citealt{gillen_ngts_2020}).
-   The rough orders of magnitude of the effects are $\approx$1\,dex in lithium abundance 
-   for a factor of 10 in rotation period, and a factor of $\approx2\times$ enhancement
-   in binarity fraction for fast {\it vs.} slowly rotating G and early K dwarfs.
+    right and lower panels of Figure~\ref{fig:physical_x_rotn}
+    quantify the statement: $\approx$75\% of the ``inner-most''
+    members have detected rotation periods.  This fraction decreases
+    to $\approx$60\% at physical separations of $\approx$200\,pc.
+    Based on the rate at which rotation periods are detected at or
+    underneath the gyrochronological slow sequence, field stars appear
+    to contaminate the outermost candidate cluster members at a rate
+    between 40\% and 50\%.  Given the number density ratios of cluster
+    to field stars at these wide separations ($\approx 10^{-3}$,
+    \citealt{meingast_2021}), these contamination fractions are not as
+    bad as one might fear.
+   %
+   \item {\it The rapid rotators show elevated lithium abundances, and
+     elevated binarity fractions}.  Each trend has been noted in
+     comparable stellar populations ({\it e.g.,}
+     \citealt{soderblom_evolution_1993}, \citealt{gillen_ngts_2020}).
+     The rough orders of magnitude of the effects are $\approx$1\,dex
+     in lithium abundance for a factor of 10 in rotation period, and a
+     factor of $\approx2\times$ enhancement in binarity fraction for
+     fast {\it vs.} slowly rotating G and early K dwarfs.
 \end{itemize}
 
-Generally speaking, the existence of the halo itself should not be not very
-surprising.  A star with a 1\kms\ velocity difference from the cluster
-average will move away from the center at a rate of 1\,pc\,Myr$^{-1}$.
-For a $\sim$100\,Myr old cluster, a halo with characteristic size
+The existence of the halo itself should not be not very surprising.  A
+star with a 1\kms\ velocity difference from the cluster average will
+move away from the center at a rate of 1\,pc\,Myr$^{-1}$.  For a
+$\sim$100\,Myr old cluster, a halo with characteristic size
 $\sim$100\,pc is therefore expected, given that the typical velocity
@@
 On the other hand, our ability to reliably identify members of these
-halos is rather surprising.  These stars have
-remained hidden in the background of the Galaxy throughout centuries
-of modern astronomy.  The consequences that these halo stars will have
-in expanding our understanding of planetary and stellar origins are
-only just beginning to be understood.
+halos is rather surprising.  These stars have remained hidden in the
+background of the Galaxy throughout centuries of modern astronomy.
+The consequences that these halo stars will have in expanding our
+understanding of planetary and stellar origins are only just beginning
+to be understood.
 
@@
 The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
-corresponds to an extinction-corrected color cutoff of $(Bp-Rp)_0
+corresponds to an extinction-corrected color cutoff of $\bpmrp
 \approx 0.80$, or spectral types of $\approx$G1V.
@@
 %   velocity space, cross-matched against the rotators.} The plotted
-%   stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
+%   stars are those with $0.5<\bpmrp<1.2$ that meet the crowding
 %   restrictions described in Section~\ref{subsec:tess}: they should

### 2021-04-07 cluster_paper/ms.tex
update ms

@@
   measurements from Gaia-ESO and GALAH, we find that the halo of
-  NGC\,2516 is real and coeval with the core.  Two thirds of
+  NGC\,2516 is real and coeval with the cluster's core.  Two thirds of
   kinematically selected halo members out to 250\,pc from the cluster
@@
 comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
-adopt a PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
+adopt a MSTO age for \cn\ of $175\pm35$\,Myr, where
 our quoted uncertainty is a quadrature sum of the statistical
@@
     (see Section~\ref{subsubsec:cluster}).  The Pleiades
-    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
+    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
     \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
@@
   Equivalent width of the 6708\,\AA\ doublet is shown versus
-  dereddened color for all kinematic \cn\ members with
-  Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
-  spectra comprise slightly over half of the halo stars, due to the
-  non-targeted selection function of that survey.  The gray points are
-  Keck-HIRES measurements of stars with detected transiting planets
-  from the Kepler mission \citep{berger_identifying_2018}.  Points
-  with ${\rm EW}\approx0\,$m\AA\ are non-detections.
+  dereddened color for all kinematic \cn\ members with Gaia-ESO
+  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
+  comprise slightly over half of the halo stars, due to the
+  non-targeted selection function of that survey.  Field stars (gray
+  points) are Keck-HIRES measurements of stars with detected
+  transiting planets from the Kepler mission
+  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
+  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
+  Praesepe detections (650\,Myr) are shown in violet
+  \citep{soderblom_praesepe_1993}.  Points with ${\rm
+  EW}\approx0\,$m\AA\ are non-detections.
   \label{fig:lithiumcorehalo}
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Lithium and rotation in NGC\,2516.}
-    Points show all kinematic \cn\ members with TESS rotation periods
-    in the Gaia-ESO or GALAH surveys, and are colored by equivalent
-    width.  Points with reported equivalent widths below 20\,m\AA\
-    (including non-detections) are shown with gray crosses.  The
-    lithium-rotation correlation for the K-dwarfs has been observed in
-    other young clusters, and is discussed in
-    Section~\ref{subsec:lithium}.
+  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
+  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
+  GALAH surveys, and are colored by equivalent width.  Points with
+  reported equivalent widths below 20\,m\AA\ (including
+  non-detections) are shown with gray crosses.  The lithium-rotation
+  correlation for the K-dwarfs has been observed in other young
+  clusters, and is discussed in Section~\ref{subsec:lithium}.
 	\label{fig:lithiumrot}
@@
 halo population of \cn\ was an analysis of the Li\,\textsc{I}
-6708\,\AA\ doublet.  The method has been reviewed by
+6708\,\AA\ doublet.  This type of age measurement has been reviewed by
 \citet{soderblom_ages_2010}, and its connection to stellar rotation
@@
 Since our main focus is measuring equivalent widths for the
-Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of our
-\nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.  51
-were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
+Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of
+our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.
+51 were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
 lithium, i.e., no lithium flag set.
@@
 cutoff of the survey.  The overall increase of Li EW from
-$0<\bpmrp<1.0$ is thought to be driven by the temperatures in the
-stellar chromospheres: hotter stars fully ionize Li out of its ground
-state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
-depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.},
-later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
-K dwarfs.  For comparison, we have also plotted the Li EWs measured by
+$0<\bpmrp<1.0$ is driven by the temperatures in the stellar
+chromospheres: hotter stars fully ionize Li out of its ground state
+\citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
+depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.}, later
+than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier K
+dwarfs: these stars have burned their lithium.  For comparison, we
+have also plotted the Li EWs measured by
 \citet{berger_identifying_2018} for planet-hosting stars in the Kepler
-field (we interpolated from their effective temperatures using the
+field, and Pleiades members from \citet{bouvier_pleiades_lirot_2018}
+(we interpolated from their effective temperatures using the
 \citet{pecaut_mamajek_2013}
 table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version 2019.3.22.}.  The field star distribution peaks at late F-type
-stars.  The majority of kinematically selected stars in the core and
-the halo show lithium equivalent widths substantially in excess of
-these field stars.
+version 2021.03.02.}.  For clarity, we have omitted the few upper
+limits reported by \citet{bouvier_pleiades_lirot_2018}.  The field
+star distribution peaks at late F-type stars.  The majority of
+kinematically selected stars in the core and the halo show lithium
+equivalent widths substantially in excess of these field stars.
 
@@
 \label{subsec:origin}
-% Should we be calling the halo a tail?
-
-Any theory for the structure of \cn, the Psc\,Eri stream, and
-the strings and halos found by
-\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
-needs to explain a few observations.  First is their aspect
-ratios.  The halo of \cn, for instance,
-extends over at least 500\,pc, despite being only $\approx$25\,pc
-wide.  Second is their axial tilts in the plane of the Milky Way -- why
-their leading and trailing arms tend to conform to the Galaxy's
-differential rotation ({\it e.g.}, Figure~\ref{fig:XYZ} or Figure~13
-of \citetalias{meingast_2021}).  Third is the correlation between the
-mean LSR-subtracted cluster velocity and elongation axes (blue arrows in
-Figure~\ref{fig:XYZ}).  Such a correlation has also been noted in Coma
-Ber by \citet{tang_comaber_2019}.  The other halos may show
+
+Any theory for the structure of \cn, the Psc\,Eri stream, and the
+strings and halos found by \citetalias{kounkel_untangling_2019} and
+\citetalias{meingast_2021} needs to explain a few observations.  First
+is their aspect ratios.  The halo of \cn, for instance, extends over
+at least 500\,pc, despite being only $\approx$25\,pc wide.  Second is
+their axial tilts in the plane of the Milky Way -- why their leading
+and trailing arms tend to conform to the Galaxy's differential
+rotation ({\it e.g.}, Figure~\ref{fig:XYZ} or Figure~13 of
+\citetalias{meingast_2021}).  Third is the correlation between the
+mean LSR-subtracted cluster velocity and elongation axes (blue arrows
+in Figure~\ref{fig:XYZ}).  Such a correlation has also been noted in
+Coma Ber by \citet{tang_comaber_2019}; the other halos may show
 similar trends.
 
-One possible explanation is that these structures are tidal tails
-\citep[{\it e.g.},][]{krumholz_star_2019}. The idea would be that
-stars escape out of the Lagrange points imposed by the Milky Way's
-potential, and then form leading and lagging arms due to differential
-rotation in the Galaxy.
-In this context, Psc\,Eri for instance might represent the completion
-of a process that \cn\ is still undergoing.
+One possible explanation is that the observed halo structures are
+tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea would
+be that as the cluster contracts, stars escape out of the Lagrange
+points imposed by the Milky Way's potential, and form leading and
+lagging arms due to differential rotation in the Galaxy.  Kinematic
+evidence that the core of \cn\ may be undergoing contraction has
+recently been presented by \citet{healy_stellar_2020}, and might
+support this interpretation.  Whether the exact rate of contraction,
+and the corresponding ``evaporation rate'' of the core is sufficient
+to explain the formation of the halo has yet to be assessed.
 
 A second possibility is that the clusters form in larger and more
-dispersed star formation complexes: the stars in the halo need
-not have formed in the same ``clump'' as those in the cluster core.
-In other words, the formation environment might be better described a
+dispersed star formation complexes: the stars in the halo need not
+have formed in the same ``clump'' as those in the cluster core.  In
+other words, the formation environment might be better described a
 giant molecular filament, rather than a giant molecular cloud.  A
@@
 ``bone candidates''.  A more immediate example is the Orion\,A cloud.
-\citet{grosschedl_3d_2018} showed using young stellar objects as tracers that Orion\,A
-is 90\,pc long, and that it has a dense ``head'' and
-a lower density ``tail''.
-
-One way to rule between the two possibilities
-would be 
-to observationally establish an age-size correlation.
-If the $\approx$100\,Myr halos are produced primarily through
-evaporation and expansion, then the
-younger clusters should be significantly smaller.
-A theoretical approach along similar lines
-would be to perform a dynamical analysis
-of whether filaments like Orion\,A, or those
-in the \citet{zucker_physical_2018} sample, are at all expected 
-to evolve into structures
-resembling \cn.
-
-% Possible observational directions could include analyzing whether the
-% Gaia kinematics contain any evidence of cluster expansion,
-% as well as analyzing the mass functions of stars in different
-% regions of the clusters to determine whether they contain
-
-% CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
-% THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
-% LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
-% 
-% Similarly Zari et al 2019 studied the entire Orion complex, and found that
-% different regions of the complex (one toward the Belt, one toward 25\,Ori)
-% began forming stars roughly simultaneously. No evidence was found of it
-% propagating from these initial sites.
-
-% cometary-like cloud,
-% oriented toward the Galactic plane, with two distinct components:
-% a denser and enhanced star-forming (bent) Head, and a
-% lower density and star-formation quieter 75 pc long Tail.
-
-% ``The stars in these halos could have escaped from the cores of the
-% cluster, or they could be remnants of a larger and lower-density star
-% formation complex that has since dispersed.''
-% Another possibility is {\it triggered star formation}.
-
-% One relevant connection may be to the ``Mamajek-2'' stellar group
-% (see Jilinski+2009).
+\citet{grosschedl_3d_2018} showed using young stellar objects as
+tracers that Orion\,A is 90\,pc long, and that it has a dense ``head''
+and a lower density ``tail''.
+
+One way to rule between the two possibilities would be to
+observationally establish an age-size correlation.  If the
+$\approx$100\,Myr halos are produced primarily through evaporation and
+expansion, then the younger clusters should be significantly smaller.
+A theoretical approach along similar lines would be to perform a
+dynamical analysis of whether filaments like Orion\,A, or those in the
+\citet{zucker_physical_2018} sample, are at all expected to evolve
+into structures resembling \cn.
 
@@
 From their rotation period analysis of \cn,
-\citet{fritzewski_rotation_2020} have recently demonstrated that at
-ages between 100 and 200\,Myr, young clusters ({\cn, Blanco\,1,
-the Pleiades, Psc\,Eri, M\,35, and M\,50) all show similar rotation
-period distributions.  The FGK stars show a ``slow sequence'', a
-population of ``rapid rotators'', and in some cases stars in the
-``void'' between the two populations ({\it e.g.,}, Psc\,Eri and \cn\
-show more stars in the void than the Pleiades).  The late M dwarfs
-with detected rotation periods tend to rotate very rapidly
-($<3$\,days), although these are beyond the magnitude limit of our
-analysis.  This section briefly discusses the origins of these
-features in the rotation-color diagram.
+\citet{fritzewski_rotation_2020} have recently demonstrated that
+between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc\,Eri, M\,35, M\,50, and the Pleaides)
+all show similar rotation period
+distributions.  The FGK stars show a ``slow sequence'', a population
+of ``rapid rotators'', and in some cases stars in the ``void'' between
+the two populations ({\it e.g.,}, Psc\,Eri and \cn\ show more stars in
+the void than the Pleiades).  The late M dwarfs with detected rotation
+periods tend to rotate very rapidly ($<3$\,days), although these are
+beyond the magnitude limit of our analysis.  This section briefly
+discusses the origins of these features in the rotation-color diagram.
 
@@
 
-The other features of the rotation-color diagram are indications
-that magnetic breaking is not always the most important effect.
-The question of why the rapid rotators exist, for instance, has not yet been definitively
-answered.  Models that introduce variable core-envelope decoupling
-timescales can reproduce each type of population
-\citep{Irwin_NGC2516_2007}, but the question then shifts to why these
-decoupling timescales should be differ for stars that are otherwise
-identical.  Another possible explanation invokes a spontaneous and
-random change of magnetic field topology that causes the dynamo to
-couple more strongly with the stellar wind
+The other features of the rotation-color diagram are indications that
+magnetic breaking is not always the most important effect.  The
+question of why the rapid rotators exist, for instance, has not yet
+been definitively answered.  Models that introduce variable
+core-envelope decoupling timescales can reproduce each type of
+population \citep{Irwin_NGC2516_2007}, but the question then shifts to
+why these decoupling timescales should be differ for stars that are
+otherwise identical.  Another possible explanation invokes a
+spontaneous and random change of magnetic field topology that causes
+the dynamo to couple more strongly with the stellar wind
 \citep{brown_metastable_2014}.  Other processes could be external the
@@
 A separate effect is that on the slow rotation sequence,
-Figure~\ref{fig:rot}} shows that the binary stars appear to be either
+Figure~\ref{fig:rot} shows that the binary stars appear to be either
 preferentially redder, or to have faster rotation periods than the
@@
 The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
-been shown to carry over beyond just equivalent widths, and to the
-actual abundance of lithium in the photosphere, A(Li).
-\citet{soderblom_evolution_1993} calculated the relevant LTE curves of
-growth for the Li\,6708\,\AA\ doublet, and used them to convert our
-measured equivalent widths to abundances.  Non-LTE corrections affect
-the abundances by $<0.5\,{\rm dex}$ and do not siginificantly affect
-the global relation between Li abundance and stellar temperature or
-rotation \citep{carlsson_1994,lind_departures_2009}.
+carries over beyond just equivalent widths, and to the actual
+abundance of lithium in the photosphere, A(Li).  To our knowledge, the
+first demonstration of this was by \citet{soderblom_evolution_1993},
+who calculated the relevant LTE curves of growth for the
+Li\,6708\,\AA\ doublet, and used them to convert equivalent widths to
+abundances.  Non-LTE corrections affect the abundances by $<0.5\,{\rm
+dex}$ and do not change the overall relationships between Li abundance
+and stellar temperature or rotation
+\citep{carlsson_1994,lind_departures_2009}.
 
@@
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
-In \cn\ specifically, \citet{jeffries_rotation_1998} observed a
+In \cn, \citet{jeffries_rotation_1998} previously observed a
 similar trend that that seen in Figure~\ref{fig:lithiumrot} when
@@
 fixed stellar mass?
-{\bf FIXME: improve this paragraph}
+{\bf FIXME: improve this paragraph... (e.g., see the introductory
+material in Jeffries+21}
 %FIXME IMPROVE THIS PARAGRAPH.
@@
 
-\subsection{Age of NGC\,2516}
+\subsection{The Age of NGC\,2516}
 \label{disc:absage}
 
-Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
-Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
-"""
-Similar to Cummings et al. (2016), Cummings \& Kalirai (2018),
-and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
-\&
-McCullough 2020)
-"""
-
-\paragraph{Gyrochronological Age}
+\paragraph{Age from Gyrochronology}
 Comparing the slow sequence of the Pleiades and \cn\ more closely, the
 top row of Figure~\ref{fig:rot} shows that the two sequences overlap
-from $0.5<\bpmrp<1.2$.  At redder colors
-($1.2<\bpmrp<1.7$) the dispersion in rotation periods increases.
-The maximum rotation periods seen in \cn\ across both sets
-$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
-than the $\approx$8.5day upper limit seen in the Pleiades.  This is
-consistent with \cn\ being slightly older than the Pleiades.
+from $0.5<\bpmrp<1.2$.  At redder colors ($1.2<\bpmrp<1.7$) the
+dispersion in rotation periods increases.  The maximum rotation
+periods seen in \cn\ across both sets $\mathcal{A}$ and $\mathcal{B}$
+extend up to $\approx$11 days, rather than the $\approx$8.5day upper
+limit seen in the Pleiades.  This is consistent with \cn\ being
+slightly older than the Pleiades.
 
@@
 of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
-is, unfortunately, an exercise in tautology.  The aforementioned
-models are empirically calibrated using the Pleiades and Praesepe.  In
+is, unfortunately, an exercise in tautology.  Gyrochronological models
+are empirically calibrated against the Pleiades and Praesepe.  In
 other words, at $\bpmrp=1.35$, we observe rotation periods in \cn\
 ranging from 7 to 10 days.  The 7 day rotation periods are consistent
-with those observed in the Pleiades; the 10 day rotation periods are
-not.  Fitting formulae from the previously cited studies give ages for
-a star with $\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204
-Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
+with the Pleiades; the 10 day rotation periods are not.  Fitting
+formulae from the previously cited studies give ages for a star with
+$\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of
+204 Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
 \citet{spada_competing_2020} model has an absolute scale that is
@@
 older than the Pleaides, and much younger than Praesepe.  Given a
-Pleiades age of $130\pm15$\,Myr (an average of MSTO and
-rotation-corrected lithium-depletion results;
-\citealt{cummings_2018}), a gyrochronology-derived age estimate for \cn\
-between 150 and 200\,Myr would therefore appear reasonable.  Much
+Pleiades age of $125\pm20$\,Myr (an average of MSTO and
+rotation-corrected lithium-depletion results; see {\it e.g.}
+\citealt{stauffer_keck_1998}, \citealt{soderblom_ages_2014}, and
+\citealt{cummings_2018}), a gyrochronological age estimate for
+\cn\ between 150 and 200\,Myr would therefore appear reasonable.  Much
 older, and the issue of why the F and G dwarfs do not spin down
 becomes pressing.
+
+\paragraph{Age from Lithium Depletion}
+
+The lithium depletion boundary for \cn\ has to our knowledge not been
+measured.  Given the distance to the cluster, this is not surprising:
+at $\sim$150\,Myr, the LDB is expected to occur at a stellar mass of
+$\approx$0.088\,$M_\odot$ \citep{soderblom_ages_2014}.  This would
+correspond to $\bpmrp\approx4.7$, well beyond the limits of available
+cluster membership lists and spectroscopy.
+
+An easier {\it relative} measurement to make of lithium depletion is
+from the EW-color diagram (Figure~\ref{fig:lithiumcorehalo}).  The
+\cn\ and Pleiades sequences in this diagram appear indistinguishable.
+A relative lithium-based age estimate for \cn\ would therefore be
+``Pleiades-age''.  The uncertainties on this estimate are large
+because lithium depletion timescales are long for Sun-like stars past
+100\,Myr.  For instance, \citet{jeffries_m35_li_2020} has shown in a
+comparison of the Pleiades and M\,35 ($\approx$150\,Myr) that the two
+are indistinguishable given current measurement, despite the age of
+the latter being expected to be marginally older than the former.
+\cn\ members are on average more depleted than {\it e.g.}, stars in
+IC\,2602, but only marginally so \citep{soderblom_ages_2014}.  On the
+older end though, Figure~\ref{fig:lithiumcorehalo} does show that
+relative to Praesepe ($\approx$650\,Myr), the lithium EWs of \cn\ are
+significantly enhanced.  Given these comparisons, a plausible range of
+ages based on lithium for \cn\ would therefore be between $\approx$100
+and 200\,Myr.
+
+\paragraph{Adopted Age}
+
+As noted in Section~\ref{subsec:hr}, a color-color analysis of the main
+sequence turn-off (MSTO) by \citet{cummings_2018} found that
+\cn\ is 40-60\,Myr older than the Pleiades.  Given a Pleiades age of
+$125\pm20$\,Myr, this yields a MSTO age for \cn\ of
+$175\pm35$\,Myr.  The gyrochronological age we have argued is likely
+super-Pleiades, {\it i.e.}, in the range of 150 to 200\,Myr.  The
+lithium age based on the depletion of the G and K dwarfs is more
+uncertain, but is consistent with the Pleiades, and could be a bit
+older ($\approx$100 to 200\,Myr).  Averaging these three different
+indicators, the age of \cn\ would be $167\pm20$\,Myr, if
+we set the uncertainties to match the absolute Pleiades age
+uncertainty of $\pm 20\,$Myr \citep{soderblom_ages_2014}.  Whether
+this absolute age for \cn\ is particularly meaningful, we leave the
+reader to judge.
 

### 2021-04-07 cluster_paper/collect.sh

@@ -17,7 +17,7 @@ cp ../results/full_kinematics_X_rotation/full_kinematics_X_rotation_fullfaint_ga
 cp ../results/physical_X_rotation/physical_X_rotation_fullfaint.pdf f8a.pdf
 cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f8b.pdf
 
-cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit_showkepfield.pdf f5a.pdf
+cp ../results/lithium/lithiumEW_vs_BpmRp_xmatch_fullfaintkinematic_gaiaeso_galahdr3_corehalosplit_showkepfield_showpleiades_showpraesepe.pdf f5a.pdf
 cp ../results/rotation_X_lithium/rotation_vs_BpmRp_X_lithium_viridis_gaiaeso_galahdr3.pdf f5b.pdf
 
 cp ../results/vtangential_projection/XYZ_with_orbit.pdf f11.pdf

### 2021-04-07 earhart/lithium.py

@@ -4,6 +4,7 @@ Contents:
     get_Randich18_NGC2516
     get_GalahDR3_lithium
     get_GalahDR3_li_EWs
+    get_Bouvier18_pleiades_li_EWs
 """
 import os
 import numpy as np, pandas as pd
@@ -234,3 +235,49 @@ def _make_Randich18_xmatch(datapath, vs_rotators=1, RADIUS=0.5):
         print(f'Got {len(mdf[mdf.subcluster=="halo"])} in halo.')
 
     mdf.to_csv(datapath, index=False)
+
+
+def get_Bouvier18_pleiades_li_EWs():
+
+    hl = fits.open(
+        os.path.join(DATADIR, 'lithium',
+                     'Bouvier_2018_vizier_all148pleiads.fits')
+    )
+
+    t_df = Table(hl[1].data).to_pandas()
+
+    return t_df
+
+
+def get_Barrado01_m35_li_EWs():
+    """
+    https://ui.adsabs.harvard.edu/abs/2001ApJ...549..452B/abstract
+    https://iopscience.iop.org/article/10.1086/319045/fulltext/datafile2.txt?doi=10.1086/319045
+    ascii ApJ MRT
+    """
+    B01_path = os.path.join(DATADIR, 'lithium',
+                            'BarradoNavascues_2001_M35_table2.txt')
+    t = Table.read(B01_path, format='cds')
+
+    df = t.to_pandas()
+
+    return df
+
+
+def get_Soderblom93_praesepe_li_EWs():
+    """
+    http://cdsarc.unistra.fr/viz-bin/cat/J/AJ/106/1080
+    https://ui.adsabs.harvard.edu/?#abs/1993AJ....106.1080S
+    vizier fits binary
+    """
+
+    hl = fits.open(
+        os.path.join(DATADIR, 'lithium',
+                     'Soderblom_1993_praesepe_table2_vizier.fits')
+    )
+
+    t_df = Table(hl[1].data).to_pandas()
+
+    return t_df
+
+

### 2021-04-07 earhart/plotting.py

@@ -1118,7 +1118,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     classes = ['pleiades', 'praesepe', f'{runid}']
     colors = ['gray', 'gray', 'k']
     zorders = [-2, -3, -1]
-    markers = ['s', 'x', 'o']
+    markers = ['s', 'p', 'o']
     lws = [0., 0.1, 0.1]
     mews= [0.5, 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
@@ -1607,7 +1607,9 @@ def plot_randich_lithium(outdir, vs_rotators=1, corehalosplit=0):
 
 
 def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
-                             corehalosplit=0, showkepfield=0, trimx=0):
+                             corehalosplit=0, showkepfield=0,
+                             showpleiades=0, showpraesepe=0,
+                             showm35=0, trimx=0):
     """
     Plot Li EW vs color for a) Randich+18's Gaia ESO lithium stars, b)
     the GALAH DR3 EWs, or c) both, after crossmatching against the
@@ -1634,6 +1636,21 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
         bdf = get_Berger18_lithium()
         bdf['BpmRp0'] = get_interp_BpmRp_from_Teff(bdf['Teff'])
 
+    if showpleiades:
+        from earhart.lithium import get_Bouvier18_pleiades_li_EWs
+        bouvier18_df = get_Bouvier18_pleiades_li_EWs()
+        bouvier18_df['BpmRp0'] = get_interp_BpmRp_from_Teff(bouvier18_df['Teff'])
+
+    if showm35:
+        from earhart.lithium import get_Barrado01_m35_li_EWs
+        b01_df = get_Barrado01_m35_li_EWs()
+        b01_df['BpmRp0'] = get_interp_BpmRp_from_Teff(b01_df['Teff'])
+
+    if showpraesepe:
+        from earhart.lithium import get_Soderblom93_praesepe_li_EWs
+        s93_df = get_Soderblom93_praesepe_li_EWs()
+        s93_df['BpmRp0'] = get_interp_BpmRp_from_Teff(s93_df['Teff'])
+
     #
     # make plot
     #
@@ -1646,9 +1663,40 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
[…]

### 2021-04-08 cluster_paper/ms.tex
update ms

@@
 \submitjournal{AAS journals.}
-\shorttitle{Core \& Halo of NGC\,2516}
+\shorttitle{The Halo of NGC\,2516}
 
@@
   Cluster Difference Imaging Photometric Survey. III.
-  %The Coeval Core \& Halo of NGC\,2516
-  %The Halo of NGC\,2516 Extends To At Least 300\,pc
-  %An Open Cluster With a 300\,pc Halo
   An Open Cluster Spanning 500\,Parsecs
-  %NGC\,2516  300\,pc
 }
@@
 
+%
 % NB. the "classical tidal radius" is from e.g., Piskunov+08 or
@@
 % 7.70pc.
+%
 \begin{abstract}
@@
   rotation rate in a manner consistent with previous observations of
-  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
-  paradigm in which the halos of many open clusters are more populous
-  than their cores.  We highlight the implications for spectroscopic
-  survey targeting, open cluster dispersal, and planet searches around
-  young stars.
+  similar clusters.  Broadly speaking, this work advances a paradigm
+  in which the halos of open clusters are often more populous than
+  their cores.  We highlight the implications for spectroscopic survey
+  targeting, open cluster dispersal, and planet searches around young
+  stars.
 \end{abstract}
@@
 and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
-associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
-an intermediate regime: low-density halos associated with a single
-overdensity, typically an open cluster that was known before Gaia
+associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
+focus on an intermediate regime: low-density halos associated with a
+single overdensity, typically an open cluster that was known before
+Gaia
 \citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
@@
 Our analysis of TESS Sectors 1-13 yielded light curves of
-\numsouthernuniqlcs\ candidate cluster members in the Southern Ecliptic
-hemisphere, which are available on
+\numsouthernuniqlcs\ candidate cluster members in the Southern
+Ecliptic hemisphere, which are available on
 MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
@@
 of stellar masses and evolutionary states.  We also required the
-comparison sample to not overlap with the cluster sample.%, which led to
-%the omission of 1.1\% of the stars drawn from the volume noted above.
+comparison sample to not overlap with the cluster sample.
 
@@
 line-of-sight distances ranging from $\approx$$300$ to
-$\approx$$600\,$pc. 
-Is this structure real? What explains its existence?
-
-% Are these halo stars truly associated with the cluster's core?
-% What is the 
-% The opportunity of verifying and understanding the diffuse
-% population of \cn\ prompted our interest in understanding
-% the reality and origin of the halo.
-
-%Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
-%$\mu_\delta$ projection were reported almost entirely by
-%\citet{meingast_2021}.
+$\approx$$600\,$pc.  Is this structure real? What explains its
+existence?
 
 An initial step in visualizing the reported structure and kinematics
-of the candidate cluster members is to consider them in Cartesian
-coordinates (Figure~\ref{fig:XYZ}).  We computed the physical
-positions by transforming from $(\alpha, \delta, \pi)$ to
-galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
-standard \citep{astropy_2018}.  The direction of galactic rotation is
+of the candidate cluster members is to consider their Cartesian
+coordinates (Figure~\ref{fig:XYZ}).  We computed these positions by
+transforming from $(\alpha, \delta, \pi)$ to galactic ($X,Y,Z$)
+assuming the \texttt{astropy v4.0} coordinate standard
+\citep{astropy_2018}.  The direction of galactic rotation is
 $+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
@@
 
-The general shape of the cluster itself seems to include a central overdensity
-and two tails (the halo).  One tail is leading the cluster's orbit,
-and is angled toward the center of the galaxy when viewed top-down.
-The other tail is trailing the cluster's orbit, and is pointed away
-from the center of the galaxy.  The elongation of the cluster in both
-the $(X,Y)$ and $(Z,Y)$ planes is correlated with the direction of the
-LSR-corrected median cluster velocity.   Possible explanations for
-this overall morphology are discussed in Section~\ref{subsec:origin}.
+The general shape of the cluster itself seems to include a central
+overdensity and two tails (the halo).  One tail is leading the
+cluster's orbit, and is angled toward the center of the galaxy when
+viewed top-down.  The other tail is trailing the cluster's orbit, and
+is pointed away from the center of the galaxy.  The elongation of the
+cluster in both the $(X,Y)$ and $(Z,Y)$ planes is correlated with the
+direction of the LSR-corrected median cluster velocity.   Possible
+explanations for this overall morphology are discussed in
+Section~\ref{subsec:origin}.
 
@@
     atmospheres is less well-calibrated for $M_\star \lesssim
-    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2); these model values
-    are shown with dashed lines.
+    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2);
+    these model values are shown with dashed lines.
     \label{fig:hr}
@@
 
-The first check on whether the candidate cluster members
-share the same age is to analyze the Gaia Hertzsprung-Russell (HR)
-diagrams.  Comparable analyses have previously been performed by
+The first check on whether the candidate cluster members share the
+same age is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.
+Comparable analyses have previously been performed by
 \citetalias{cantatgaudin_gaia_2018},
@@
 Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
-$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$ color.  Stars in
-the core appear consistent with
-having a fixed age and metallicity, and varying mass.  The halo stars
-show a similar sequence, though with greater scatter.  One possible
-explanation for the additional scatter could be that the halo is more
-contaminated by field stars.  For instance, $\approx5$ red
-giants in the halo
-must be field interlopers, because their isochronally implied
-ages would be inconsistent with that of the cluster.
-
-A second possible explanation for scatter in the halo's HR diagram
+$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$
+color.  Stars in the core appear consistent with having a fixed age
+and metallicity, and varying mass.  The halo stars show a similar
+sequence, though with greater scatter.  One possible explanation for
+the additional scatter could be that the halo is more contaminated by
+field stars.  For instance, $\approx5$ red giants in the halo must be
+field interlopers, because their isochronally implied ages would be
+inconsistent with that of the cluster.
+
+Another possible explanation for scatter in the halo's HR diagram
 could be differential reddening across different sightlines.  The
@@
 galactic plane also being further from the Sun by up to 300\,pc
-(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
-reddening is discussed in Appendix~\ref{app:diffred}. Based on the
-available evidence, we expect that both field star contamination and
-differential reddening could play a role in the larger scatter of the
-halo relative to the core.  A third possibility that we have not
-explored is whether the binary fraction could also differ between the
-different regions of the cluster.
+(Figure~\ref{fig:gaia6d}).  An HR diagram binned by galactic latitude
+did show some minimal evidence for differential reddening, and so we
+expect that both field star contamination and differential reddening
+could play a role in the larger scatter of the halo relative to the
+core.  A third possibility that we have not explored is whether the
+binary fraction could also differ between the different regions of the
+cluster.
 
@@
 
-To determine the reddening correction across \cn, we
-queried the 2MASS
-DUST
-service\footnote{
-	\url{http://irsa.ipac.caltech.edu/applications/DUST};
-	query performed using the \texttt{astrobase.services.dust} module
-	\citep{bhatti_astrobase_2018}.
-}
-and retrieved the extinction parameters at the positions of each
-NGC\,2516 member.  The mean and standard
-deviation of the $E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
-and \citet{schlafly_measuring_2011} maps were consistent, so we
-adopted the average from \citet{schlegel_maps_1998}: $E(\mathrm{B}-\mathrm{V})_{\rm
+To determine the reddening correction across \cn, we queried the 2MASS
+DUST service\footnote{
+  \url{http://irsa.ipac.caltech.edu/applications/DUST}; query
+  performed using the \texttt{astrobase.services.dust} module
+  \citep{bhatti_astrobase_2018}.  } and retrieved the extinction
+parameters at the positions of each NGC\,2516 member.  The mean and
+standard deviation of the $E(\mathrm{B}-\mathrm{V})$ values for the
+\citet{schlegel_maps_1998} and \citet{schlafly_measuring_2011} maps
+were consistent, so we adopted the average from
+\citet{schlegel_maps_1998}: $E(\mathrm{B}-\mathrm{V})_{\rm
 S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
@@
 proposed by \citet{bonifacio_search_2000} leads to our adopted
-$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
-the calibration of \citet{stassun_TIC8_2019}, namely $E(\mathrm{Bp}-\mathrm{Rp})=1.31
-E(\mathrm{B}-\mathrm{V})$ and $A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded $E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$,
-which is used in the plots.  To redden the isochrones, we assumed
-$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
-extinction law star-by-star through the PARSEC interface.
+$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$.  Finally, to convert to Gaia
+colors we used the calibration of \citet{stassun_TIC8_2019}, namely
+$E(\mathrm{Bp}-\mathrm{Rp})=1.31 E(\mathrm{B}-\mathrm{V})$ and
+$A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded
+$E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$, which is used in the
+plots.  To redden the isochrones, we assumed $R_V=3.1$, and applied
+the \citet{odonnell_1994} SED-dependent extinction law star-by-star
+through the PARSEC interface.
 
 For the metallicity, we considered a range of super and sub-solar
-metallicities to fit as much of the locus from $0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as
-possible, and settled on the slightly super-solar $[M/H]=0.06$
-\citep{cummings_2011_li_iron}.  Sub-solar metallicities led to model
-predictions too blue along the main sequence by $\approx$0.1\,mag.
-Our adopted metallicity agrees with the spectroscopic metallicity from
+metallicities to fit as much of the locus from
+$0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as possible, and settled on the
+slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
+Sub-solar metallicities led to model predictions too blue along the
+main sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees
+with the spectroscopic metallicity from
 \citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
@@
 Overall, the data and models agree for masses above
-$\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
-colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in the sense that the model
-isochrones have lower luminosities and bluer colors than the
-observations.\footnote{The 100 Myr isochrone does overlap well with the
-lower main sequence, but it fails to fit the upper main sequence and
-red giants.}  The MIST isochrones showed a comparable disagreement
-\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
-between the models and observations have include starspots and
-incomplete molecular line lists\citep[{\it
+$\approx$0.45\,$M_\odot$.  Below this mass, the data and models
+diverge at colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in
+the sense that the model isochrones have lower luminosities and bluer
+colors than the observations.\footnote{The 100 Myr isochrone does
+overlap well with the lower main sequence, but it fails to fit the
+upper main sequence and red giants.}  The MIST isochrones showed a
+comparable disagreement \citep{choi_mesa_2016}.  Proposed explanations
+for the discrepancy between the models and observations have include
+starspots and incomplete molecular line lists\citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
@@
 165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
-PARSEC, MIST, or SYCLIST).  %Their statistical uncertainties ($\approx
-%25$\,Myr) were comparable to the systematic uncertainties.
-
-Our goal is to determine whether the ages of the core and halo
-are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
+PARSEC, MIST, or SYCLIST).
+
+Our goal is to determine whether the ages of the core and halo are
+consistent.  The left panel of Figure~\ref{fig:hr} suggests that
 isochronally, they are: stars past the turnoff in both the halo and
@@
 can be made.  The MSTO stars are consistent with the 178 and 316 Myr
-models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
-consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
-model.  Based on the assumption that the width of the MSTO can be
-attributed to binary stars, we are most interested in the blue edge
-(blue stragglers are, however, a concern).  The blue edge appears most
-compatible with the 178 Myr model.  These results are consistent with
-the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
-and we refer to their work for the more precise model-dependent
-comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
-adopt a MSTO age for \cn\ of $175\pm35$\,Myr, where
-our quoted uncertainty is a quadrature sum of the statistical
-and systematic components from \cite{cummings_2018}. 
-%The relative age of ``slightly older than
-%the Pleiades'' is likely the more robust statement, consistent with
-%reports since \citet{mermilliod_comparative_1981}.
+models, but are bluer than the 100 Myr model.  The red-giant-branch
+(RGB) stars are consistent only with the 178 Myr ($\log_{10} t/{\rm
+yr} = 8.25$) model.  Based on the assumption that the width of the
+MSTO can be attributed to binary stars, we are most interested in the
+blue edge (blue stragglers are, however, a concern).  The blue edge
+appears most compatible with the 178 Myr model.  These results are
+consistent with the MSTO age range of 165 to 195\,Myr found by
+\citet{cummings_2018}, and we refer to their work for the more precise
+model-dependent comparison.  Based on Table~3 of
+\citet{cummings_2018}, we therefore adopt a MSTO age for \cn\ of
+$175\pm35$\,Myr, where our quoted uncertainty is a quadrature sum of
+the statistical and systematic components from \cite{cummings_2018}.
 
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
-    {\it Top row}: \cn\ split into its core and halo.  Sets
-    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
-    (see Section~\ref{subsubsec:cluster}).  The Pleiades
+  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for
+  \cn.} {\it Top row}: \cn\ split into its core and halo.  Sets
+  $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo
+  successive stages of automated cleaning, (see
+  Section~\ref{subsubsec:cluster}).  The Pleiades
   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
   \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
-    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $\bpmrp$$<$1.2.
-    From $1.2<\bpmrp<1.7$, 
-    some NGC\,2516 members have longer rotation periods than Pleaides members.  
-    {\it Bottom row}: \cn\
-    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
-    appear in red (astrometric binaries; 11\% of stars).  Overplotted
-    in orange are sources at least 0.3\,mag above an empirically fit
-    cluster isochrone (photometric binaries; 20\% of stars).
+  Stars in the core and the halo of NGC\,2516 overlap with the
+  Pleiades at $\bpmrp$$<$1.2.  From $1.2<\bpmrp<1.7$, some NGC\,2516
+  members have longer rotation periods than Pleaides members.  {\it
+  Bottom row}: \cn\ colored by binarity indicators.  Sources with
+  $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\% of
+  stars).  Overplotted in orange are sources at least 0.3\,mag above
+  an empirically fit cluster isochrone (photometric binaries; 20\% of
+  stars).
   \label{fig:rot}
@@
 $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
-source list had $Rp<16$.  The difference (65 stars)
-is due to 35 stars from 
-\citet{meingast_2021} which were not available at the time of the
+source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
+from \citet{meingast_2021} which were not available at the time of the
 CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
 $Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
-roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at
-the distance of \cn.
+roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at the
+distance of \cn.
 
@@
 Details regarding our detrending approach are presented in
-Appendix~\ref{app:detrending}.  After applying the initial
-ensemble decorrelation to remove systematic trends, we proceeded with a few
+Appendix~\ref{app:detrending}.  After applying the initial ensemble
+decorrelation to remove systematic trends, we proceeded with a few
 cleaning steps: we masked 0.7 days at the beginning and end of each
@@
 using the periodogram implementations in \texttt{astrobase}, namely
-the traditional Lomb-Scargle, and
-the Stellingwerf phase-dispersion minimization algorithm
+the traditional Lomb-Scargle, and the Stellingwerf phase-dispersion
+minimization algorithm
 \citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
@@
 of automated cleaning criteria, which yielded sets we denote
-$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered
-light curves for which the peak Lomb-Scargle periodogram period was
-below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
-for which no equal-brightness or greater companions were within the
-aperture.  Beyond requiring that the target star be the brightest star
-in the aperture, we also required that at most one companion with
-brightness exceeding one tenth of the target star's brightness be
-present in the aperture according to the Gaia DR2 source catalog.  For
-set $\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out
-of \nautorotdenominator\ stars for which rotation periods might
-plausibly have been detected.  For set $\mathcal{B}$, we additionally
-required that the Lomb-Scargle and Stellingwerf phase-dispersion
-methods yielded identical periods to within 10\%.  This yielded
+$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered light
+curves for which the peak Lomb-Scargle periodogram period was below 15
+days, the normalized Lomb-Scargle power exceeded 0.08, and for which
+no equal-brightness or greater companions were within the aperture.
+Beyond requiring that the target star be the brightest star in the
+aperture, we also required that at most one companion with brightness
+exceeding one tenth of the target star's brightness be present in the
+aperture according to the Gaia DR2 source catalog.  For set
+$\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out of
+\nautorotdenominator\ stars for which rotation periods might plausibly
+have been detected.  For set $\mathcal{B}$, we additionally required
+that the Lomb-Scargle and Stellingwerf phase-dispersion methods
+yielded identical periods to within 10\%.  This yielded
 \nautorotnumeratormatching\ light curves.
@@
 contamination.  For instance, stars above the slow sequence (perhaps
-field contaminants) mostly disappear in set $\mathcal{B}$.
-The ``void'' or ``gap'' underneath the slow sequence is also more
+field contaminants) mostly disappear in set $\mathcal{B}$.  The
+``void'' or ``gap'' underneath the slow sequence is also more
 pronounced in set $\mathcal{B}$.  To facilitate a comparison against
@@
 \cn\ appear gyrochronologically coeval for colors spanning
-$0.5<\bpmrp<1.2$.  The implications of this comparison for the age
-of \cn\ are discussed in Section~\ref{disc:absage}.
-%An important note is that stars redder than $\bpmrp\gtrsim 2.2$
-%(later than $\approx$M2V)  are beyond the magnitude limit of our survey.
+$0.5<\bpmrp<1.2$.  The implications of this comparison for the age of
+\cn\ are discussed in Section~\ref{disc:absage}.
 
@@
 {\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
-photometric binarity, we fitted a spline to Figure~\ref{fig:hr}, fixing
-the nodes by hand.  We then labeled any points brighter than 0.3\,mag above the cluster
-sequence as photometric binaries.  This included 20\% of the overall
-cluster sample.  Many of the resulting candidate binaries, though not
-all of them, appear below the slow rotation sequence.  In set
-$\mathcal{B}$, over the color range where the slow and fast sequence
-are present ($0.5<\bpmrp<1.3$), there are 281 stars, of which 103
-(37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
-and ``slow'' sequences by eye,  the fraction of stars showing signs of
-binarity in the slow and fast subsamples are 30\% (63/210)
-and 56\% (40/71), respectively.  We speculate on possible explanations
-in Section~\ref{disc:rotn}.
+photometric binarity, we fitted a spline to Figure~\ref{fig:hr},
+fixing the nodes by hand.  We then labeled any points brighter than
+0.3\,mag above the cluster sequence as photometric binaries.  This
+included 20\% of the overall cluster sample.  Many of the resulting
+candidate binaries, though not all of them, appear below the slow
+rotation sequence.  In set $\mathcal{B}$, over the color range where
+the slow and fast sequence are present ($0.5<\bpmrp<1.3$), there are
+281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
+281 stars into ``fast'' and ``slow'' sequences by eye,  the fraction
+of stars showing signs of binarity in the slow and fast subsamples are
+30\% (63/210) and 56\% (40/71), respectively.  We speculate on
+possible explanations in Section~\ref{disc:rotn}.
 
@@
   \caption{ 
-  {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
+  {\bf \cn\ members exist hundreds of parsecs from the core, and up
   to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
@@
   shown in orange.  Stars in the core are shown as black points.
-  Computing the 2D-tangential velocity requires a correction for
+  Computing the 2D-tangential velocity requires correcting for
   projection effects (Appendix~\ref{app:vproj}).  In the {\it top
@@
 
-Even with contaminants, our rotation period measurements show
-that the halo extends to separations of $\approx$250\,pc in physical
-space from the cluster core.  The most widely separated F2V-K2V stars
-with rotation periods (blue points in
-Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
-The total length of the structure is therefore 400--500\,pc, depending
-on which members of the halo are chosen as the ``tips'' on either end.
-This agrees with the overall structure of the halo reported by
+Even with contaminants, our rotation period measurements show that the
+halo extends to separations of $\approx$250\,pc in physical space from
+the cluster core.  The most widely separated F2V-K2V stars with
+rotation periods (blue points in Figure~\ref{fig:physical_x_rotn}) are
+separated by $\approx$430\,pc.  The total length of the structure is
+therefore 400--500\,pc, depending on which members of the halo are
+chosen as the ``tips'' on either end.  This agrees with the overall
+structure of the halo reported by
 \citetalias{kounkel_untangling_2019}, with some minor caveats visible
@@
 		\leavevmode
-		%\subfloat{
-			\includegraphics[width=1\textwidth]{f5a.pdf}
-		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
-		%}
+			\includegraphics[width=0.9\textwidth]{f5a.pdf}
 	\end{center}
@@
   EW}\approx0\,$m\AA\ are non-detections.
+  {\bf LGB TODO: the increasing EWs for the Gaia-ESO M dwarfs must be erroneous. What's happening there?}
   \label{fig:lithiumcorehalo}
@@
 		\leavevmode
-		%\subfloat{
-		\includegraphics[width=1\textwidth]{f5b.pdf}
-		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
-		%}
+		\includegraphics[width=0.9\textwidth]{f5b.pdf}
 	\end{center}
@@
 %/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
-%
 % https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
+%
+
 We downloaded\footnote{Via \url{datacentral.org.au/services/download},
@@
 
-%\subsection{Contamination fraction across the halo}
-%\subsection{How did the halo form?}
-%\subsection{Rotation in NGC\,2516}
-%\subsection{The lithium-rotation correlation}
-%\subsection{Age of NGC\,2516}
-
-% \subsection{Contamination fraction across the halo}
-% 
-% The main aim this study is to verify whether stars
-% in the halo of \cn\ are the same age as stars in the core, and
-% secondarily to quantify the contamination rate.
-% Figure~\ref{fig:physical_x_rotn}
-
 \subsection{How did the halo form?}
@@
 \citet{fritzewski_rotation_2020} have recently demonstrated that
-between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc\,Eri, M\,35, M\,50, and the Pleaides)
-all show similar rotation period
+between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc\,Eri,
+M\,35, M\,50, and the Pleaides) all show similar rotation period
 distributions.  The FGK stars show a ``slow sequence'', a population
@@
 binaries, or alternatively disk truncation that when combined with
-magnetic locking {\bf (CITE)} could produce the rotation period
-bimodality.
+magnetic locking \citep[{\it
+e.g.},][]{koenigl_disk_1991,long_locking_2005} could produce the
+rotation period bimodality.
 
@@
 and 56\% (40/71); the fast rotators show a preference for binarity.
-This correlation is in line with the earlier findings noted above,
-though we emphasize many rapid rotators in \cn\ (34\%) have no
-evidence for being binary.
+This correlation is in line with the earlier findings noted above.
+While many rapid rotators in \cn\ (34\%) have no evidence for being
+binary, we emphasize that the cuts we used to select binaries (${\rm
+RUWE}>1.2$; photometric excess exceeding 0.3\,mag) leave open a
+significant fraction of parameter space.  A 0.3\,mag flux excess, for
+instance, corresponds to mass ratios $M_2/M_1 \gtrsim 0.72$, assuming
+the luminosity $L$ scales as $M^{3.5}$.
 
 Comparing the explanations of disk-locking against tidal
-synchronization, the population statistics do seem to strain the
+synchronization, the population statistics seem to strain the
 possibility of tidal synchronization.  A quarter of the \cn\ members
@@
 sub-100\,day periods become tidally synchronized, we can only explain
-a rapid rotator occurrence rate of $\approx$5\%.
+a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity
+fractions in pre-MS stars likely do not change this picture \citep[see
+Sec.~4.4 of][]{duchene_stellar_2013}.
 
@@
 
-The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
-carries over beyond just equivalent widths, and to the actual
-abundance of lithium in the photosphere, A(Li).  To our knowledge, the
-first demonstration of this was by \citet{soderblom_evolution_1993},
-who calculated the relevant LTE curves of growth for the
-Li\,6708\,\AA\ doublet, and used them to convert equivalent widths to
-abundances.  Non-LTE corrections affect the abundances by $<0.5\,{\rm
-dex}$ and do not change the overall relationships between Li abundance
-and stellar temperature or rotation
-\citep{carlsson_1994,lind_departures_2009}.
-
-More recent studies of lithium and rotation since
-\citet{soderblom_evolution_1993} have expanded these results in the
-Pleiades, the Psc-Eri stream, and M\,35, among other clusters
+The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) carries
+over beyond just equivalent widths, and to the actual abundance of
+lithium in the photosphere, A(Li).  To our knowledge, the first
+demonstration of this was by \citet{soderblom_evolution_1993}, who
+calculated the relevant LTE curves of growth for the Li\,6708\,\AA\
+doublet, and used them to convert equivalent widths to abundances.
+Non-LTE corrections affect the abundances by $<0.5\,{\rm dex}$ and do
+not change the overall relationships between Li abundance and stellar
+temperature or rotation \citep{carlsson_1994,lind_departures_2009}.
+
+More recent studies of lithium and rotation have expanded these
+results in the Pleiades, the Psc-Eri stream, and M\,35, among other
+clusters
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
-In \cn, \citet{jeffries_rotation_1998} previously observed a
-similar trend that that seen in Figure~\ref{fig:lithiumrot} when
-analyzing 24 stars in the core of \cn.  Our concatenation of the
+In \cn, \citet{jeffries_rotation_1998} previously observed a similar
+trend that that seen in Figure~\ref{fig:lithiumrot} when analyzing 24
+stars in the core of \cn.  Our concatenation of the
 \citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
 spectra represent a significant expanasion in volume and color range
-since these earlier studies.
-
-% The fact that we see the same trend across the core and halo of \cn\ 
-% {\it a)} supports the interpretation that the halo is coeval with the
-% core, and {\it b)} suggests that the lithium-rotation correlation is
-% not explained by environmental effects such as the density of the cluster
-% during its embedded phase, but is instead tied to physical processes
-% acting on all low-mass pre-main-sequence stars.
+from this earlier analysis.
 
 What causes the correlation between lithium abundance and rotation at
-fixed stellar mass?
-{\bf FIXME: improve this paragraph... (e.g., see the introductory
-material in Jeffries+21}
-%FIXME IMPROVE THIS PARAGRAPH.
-Processes both internal and external to the star have been suggested to explain
-the lithium-rotation trend (CITE see the recent review by Bouvier+20).
-One explanation based in the stellar interior could be that the
-convective mixing efficiency is anticorrelated with the surface
-rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
-possibility could be that stronger magnetic fields in the star's
-interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
-Somers + Pinsonneault 2014).  An external process that might also be
-important is the effect of star-disk magnetic locking during the PMS
-phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
-the star's outer convective zone being locked for longer while the
-radiative core contracts.  The resulting differential rotation and
-rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
-Eggenberger+12).
-We have no preference between these possibilities---we simply note
-that the \cn\ sample could be another helpful data point in
-distinguishing them.
-
-\subsection{The Age of NGC\,2516}
+fixed stellar mass?  This question has yet to be definitively
+answered, but hints likely lie in shared correlations between lithium
+abundance, rotation rate, radius inflation, internal mixing, and
+magnetic field strength
+\citep{chabrier_evolution_2007,somers_measurement_2017,jeffries_m35_li_2020}.
+For instance, one interpretation of the lithium-rotation correlation
+based on pure hydrodynamics is that the rapid rotation leads to less
+efficient convective penetration, which then lowers the mixing
+efficiency at the convective-radiative boundary
+\citep{baraffe_lithium_2017}.  The magnetic field itself may even be
+sufficient to inhibit the convection \citep{ventura_Li_B_1998}.
+Alternatively, the star's structure could be at fault: rapid rotators
+have the strongest magnetic fields and the most spotted surfaces.
+These spots lower the photospheric flux, and drive the stellar radius
+to expand while lowering the core temperatures, which in turn slows
+the rate of lithium-destroying proton capture reactions
+\citep{feiden_magnetic_2013,somers_rotation_2015}.
+
+None of the internal processes noted above explain the initial
+question of why some of the G and K dwarfs rotate rapidly to begin
+with.  One clue noted above is that the rapid rotators tend to be in
+photometric or astrometric binaries.  The majority of such binaries
+have wide ($a\gtrsim200$\,au) or intermediate (0.5 - 100\,au)
+separations \citep{raghavan_survey_2010}.
+
+It is expected that the circumstellar disks in these binaries have
+different properties than those of single stars, due to {\it e.g.} gap
+formation in the circumbinary disk, disk truncation, and faster disk
+clearing \citep{artymowicz_dynamics_1994,moe_impact_2020}.  It would
+therefore be reasonable to assume that disk lifetimes in the binary
+systems are shorter than in single star systems.
+\citet{eggenberger_impact_2012} showed that longer disk lifetimes
+should lead to an extended phase of disk-locking, which leads to a
+greater amount of differential rotation generated in the radiative
+zone, more efficient shear mixing, and ultimately a lower photospheric
+lithium abundance.  This provides a plausible explanation for the
+overall trend that the rapid rotators are often in unresolved
+binaries, and that they are also lithium rich.  A careful
+observational analysis including an assessment of the completeness to
+different separations and mass ratios of binaries would be useful to
+clarify whether or not all of the rapid rotators are in binaries.  If
+not, the scenario presented above would be falsified.
+
+\subsection{The age of NGC\,2516}
 \label{disc:absage}
@@
 top row of Figure~\ref{fig:rot} shows that the two sequences overlap
-from $0.5<\bpmrp<1.2$.  At redder colors ($1.2<\bpmrp<1.7$) the
+from $0.5<\bpmrp<1.2$.  At redder colors, $1.2<\bpmrp<1.7$, the
 dispersion in rotation periods increases.  The maximum rotation
@@
 \citealt{stauffer_keck_1998}, \citealt{soderblom_ages_2014}, and
-\citealt{cummings_2018}), a gyrochronological age estimate for
-\cn\ between 150 and 200\,Myr would therefore appear reasonable.  Much
+\citealt{cummings_2018}), a gyrochronological age estimate for \cn\
+between 150 and 200\,Myr would therefore appear reasonable.  Much
 older, and the issue of why the F and G dwarfs do not spin down
@@
 \paragraph{Age from Lithium Depletion}
-
 The lithium depletion boundary for \cn\ has to our knowledge not been
@@
 \paragraph{Adopted Age}
-
-As noted in Section~\ref{subsec:hr}, a color-color analysis of the main
-sequence turn-off (MSTO) by \citet{cummings_2018} found that
-\cn\ is 40-60\,Myr older than the Pleiades.  Given a Pleiades age of
-$125\pm20$\,Myr, this yields a MSTO age for \cn\ of
-$175\pm35$\,Myr.  The gyrochronological age we have argued is likely
-super-Pleiades, {\it i.e.}, in the range of 150 to 200\,Myr.  The
-lithium age based on the depletion of the G and K dwarfs is more
-uncertain, but is consistent with the Pleiades, and could be a bit
-older ($\approx$100 to 200\,Myr).  Averaging these three different
-indicators, the age of \cn\ would be $167\pm20$\,Myr, if
-we set the uncertainties to match the absolute Pleiades age
-uncertainty of $\pm 20\,$Myr \citep{soderblom_ages_2014}.  Whether
-this absolute age for \cn\ is particularly meaningful, we leave the
-reader to judge.
+As noted in Section~\ref{subsec:hr}, a color-color analysis of the
+main sequence turn-off (MSTO) by \citet{cummings_2018} found that \cn\
+is 40-60\,Myr older than the Pleiades.  Given a Pleiades age of
+$125\pm20$\,Myr, this yields a MSTO age for \cn\ of $175\pm35$\,Myr.
+The gyrochronological age we have argued is likely super-Pleiades,
+{\it i.e.}, in the range of 150 to 200\,Myr.  The lithium age based on
+the depletion of the G and K dwarfs is more uncertain, but is
+consistent with the Pleiades and could be a bit older ($\approx$100 to
+200\,Myr).  Averaging these three different indicators gives an age
+for \cn\ of $167\pm20$\,Myr, if we set the uncertainties to match the
+absolute Pleiades age uncertainty of $\pm 20\,$Myr
+\citep{soderblom_ages_2014}.  Whether this absolute age is at all
+meaningful, we leave the reader to judge.
 
@@
 	%
-  \item {\it Isochronal, rotational, and lithium dating show that
-    the halo of \cn\ is coeval with its core.} In short, all the data
-    are consistent with the halo being real.  The Gaia EDR3 photometry
-    and astrometry show a main-sequence turnoff that suggest an
-    isochronal age of 150\,Myr for both the core and halo
-    (Figure~\ref{fig:hr}).  The faintest M dwarfs in the core and halo
-    are also brighter than comparison field stars of the same color.
-    Gyrochronally, stars in \cn\ spanning $0.5<\bpmrp<1.2$
-    (spectral types F2V-K3V) overlap with the slow sequence of the
-    Pleaides (Figure~\ref{fig:rot}).  At redder colors, from
-    $1.2<\bpmrp<1.7$ (spectral types K3V-K6V, $M_\star$$\approx$
-    0.6-0.7\,$M_\odot$), there is a larger dispersion in rotation
-    period.  The upper envelope of the rotation period distribution at
-    these redder colors extends to longer rotation periods than in the
-    Pleiades (11 {\it vs.}\ 8 day rotation periods).  This is
-    consistent with \cn\ being slightly older than the Pleiades ({\it
-    i.e.}, $\approx$150\,Myr).  The lithium equivalent widths of the
-    kinematic cluster members (Figure~\ref{fig:lithiumcorehalo}) are
-    consistent with this assessment.
+  \item {\it Isochronal, rotational, and lithium dating show that the
+    halo of \cn\ is coeval with its core.} In short, all the data are
+    consistent with the halo being real.  The Gaia EDR3 photometry and
+    astrometry show a main-sequence turnoff that suggest an isochronal
+    age of 150\,Myr for both the core and halo (Figure~\ref{fig:hr}).
+    The faintest M dwarfs in the core and halo are also brighter than
+    comparison field stars of the same color.  Gyrochronally, stars in
+    \cn\ spanning $0.5<\bpmrp<1.2$ (spectral types F2V-K3V) overlap
+    with the slow sequence of the Pleaides (Figure~\ref{fig:rot}).  At
+    redder colors, from $1.2<\bpmrp<1.7$ (spectral types K3V-K6V,
+    $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
+    dispersion in rotation period.  The upper envelope of the rotation
+    period distribution at these redder colors extends to longer
+    rotation periods than in the Pleiades (11 {\it vs.}\ 8 day
+    rotation periods).  This is consistent with \cn\ being slightly
+    older than the Pleiades ({\it i.e.}, $\approx$150\,Myr).  The
+    lithium equivalent widths of the kinematic cluster members
+    (Figure~\ref{fig:lithiumcorehalo}) are consistent with this
+    assessment.
   %
@@
      comparable stellar populations ({\it e.g.,}
-     \citealt{soderblom_evolution_1993}, \citealt{gillen_ngts_2020}).
+     \citealt{soderblom_evolution_1993}; \citealt{jeffries_m35_li_2020}; \citealt{gillen_ngts_2020}).
      The rough orders of magnitude of the effects are $\approx$1\,dex
      in lithium abundance for a factor of 10 in rotation period, and a
-     factor of $\approx2\times$ enhancement in binarity fraction for
-     fast {\it vs.} slowly rotating G and early K dwarfs.
+     factor of $\gtrsim2$ enhancement in binarity fraction for fast
+     {\it vs.} slowly rotating G and early K dwarfs.  We have pointed
+     out in Section~\ref{discussion:lithium} that one plausible
+     connection between the two observations is that the disk-locking
+     phase may be truncated in binary systems, which could lead to
+     less efficient internal mixing and the longer-lasting presence of
+     photospheric lithium \citep{eggenberger_impact_2012}.
+   %
 \end{itemize}
@@
 move away from the center at a rate of 1\,pc\,Myr$^{-1}$.  For a
-$\sim$100\,Myr old cluster, a halo with characteristic size
-$\sim$100\,pc is therefore expected, given that the typical velocity
-dispersions of open clusters are of the order of kilometers per
-second.
+$\sim$100\,Myr cluster, a halo with characteristic size $\sim$100\,pc
+is therefore expected, given that the typical velocity dispersions of
+open clusters are of order kilometers per second.
 
@@
 
-The authors thank X and Y for fruitful discussions.
-
-B.~Tofflemire, % lithium suggestoins
+The authors thank L.~Hillenbrand, B.~Tofflemire, G.~Zhou, E.~Newton, A.~Mann, and A.~Kraus for fruitful discussions.
 
@@
   the kinematically selected NGC\,2516 members.
+  {\bf LGB TODO: Rerun the field measurements with a better grid to fix the aliasing issue.}
   \label{fig:compstar}
@@
 
-\section{Differential Reddening}
-\label{app:diffred}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-			\includegraphics[width=0.75\textwidth]{f2c.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
-    {\it Bottom:} Reported members of the halo, as a function of
-    galactic latitude. Can the additional scatter in the halo be
-    understood through differential reddening?
-    {\bf Maybe}.
-    \label{fig:diffred}
-  }
-\end{figure*}
-
-Figure~\ref{fig:diffred}.
+% \section{Differential Reddening}
+% \label{app:diffred}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 			\includegraphics[width=0.75\textwidth]{f2c.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+%   \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
+%     {\it Bottom:} Reported members of the halo, as a function of
+%     galactic latitude. Can the additional scatter in the halo be
+%     understood through differential reddening?
+%     {\bf Maybe}.
+%     \label{fig:diffred}
+%   }
+% \end{figure*}
+% 
+% Figure~\ref{fig:diffred}.
 

### 2021-04-08 earhart/plotting.py

@@ -1118,7 +1118,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     classes = ['pleiades', 'praesepe', f'{runid}']
     colors = ['gray', 'gray', 'k']
     zorders = [-2, -3, -1]
-    markers = ['s', 'p', 'o']
+    markers = ['s', '+', 'o']
     lws = [0., 0.1, 0.1]
     mews= [0.5, 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
@@ -1693,7 +1693,7 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
         # vizier randomly decided to divide by 10
         ax.scatter(
             s93_df[dets]['BpmRp0'], 10*s93_df[dets]['WLi'], c='violet', alpha=1,
-            zorder=0, s=7, edgecolors='k', marker='P',
+            zorder=0, s=9, edgecolors='k', marker='P',
             linewidths=0.1, label='Praesepe'
         )
 

### 2021-04-08 cluster_paper/ms.tex
update ms

@@
   Cluster Difference Imaging Photometric Survey. III.
-  An Open Cluster Spanning 500\,Parsecs
+  Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
 }
@@
 photoionization, radiation pressure, and supernova shocks disperse the
-gas out of the cloud \citep{krumholz_star_2019}.  Since only $\sim
-1\%$ of the cloud mass is converted into stars, gas dispersal enables
-the majority ($\sim$90\%) of stars in the cluster to escape the
-cluster's gravitational well \citep{lada_embedded_2003}.
+gas out of the cloud \citep{krumholz_star_2019}.  Since only
+$\sim$$1\%$ of the cloud mass is converted into stars, gas dispersal
+enables the majority ($\sim$90\%) of stars in the cluster to escape
+the cluster's gravitational well \citep{lada_embedded_2003}.
 
@@
 and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
-associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
+associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
 focus on an intermediate regime: low-density halos associated with a
@@
 We have recently been making TESS light curves of stars reported to be
-members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
+members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}) and
+searching them for planets \citep{bouma_cluster_2020}.
 Our analysis of TESS Sectors 1-13 yielded light curves of
@@
 (Section~\ref{subsec:hr}), TESS gyrochronology
-(Section~\ref{subsec:tess}), and lithium depletion
+(Section~\ref{subsec:tess}), and relative lithium abundances
 (Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
 discuss the implications of our analysis for NGC\,2516 specifically
-and stellar spin-down and open cluster evolution generally.
+and stellar spin-down more generally.
 Section~\ref{sec:conclusion} gives our conclusions.
@@
 
-\subsection{Rotation at 100-200\,Myr}
+\subsection{Stellar Rotation at 100-200\,Myr}
 \label{disc:rotn}
@@
 are rapidly rotating.  In comparison, in the field half of Sun-like
-stars are binaries, and $\approx$9\% have periods below 100\,days
+stars are binaries, and $\approx$9\% of these binaries have periods below 100\,days
 \citep{raghavan_survey_2010}.  If we assume that all binaries with
-sub-100\,day periods become tidally synchronized, we can only explain
+sub-100\,day periods become tidally synchronized, then we can only explain
 a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity

### 2021-04-12 cluster_paper/ms.tex
update ms
update appendix 1

@@
   in which the halos of open clusters are often more populous than
-  their cores.  We highlight the implications for spectroscopic survey
-  targeting, open cluster dispersal, and planet searches around young
-  stars.
+  their cores.  There are significant implications for spectroscopic
+  survey targeting, open cluster dispersal, and planet searches around
+  young stars.
 \end{abstract}
@@
 and the corresponding ``evaporation rate'' of the core is sufficient
-to explain the formation of the halo has yet to be assessed.
+to explain the formation of the halo has yet to be assessed.  The
+presence or lack of epicyclic overdensities in the tails might also
+help clarify whether this scenario is happening
+\citep{kupper_more_2012}.
 
@@
 \appendix
-\section{Clustering methods and outcomes}
+\section{Clustering Methods and Outcomes}
 \label{app:clustering}
 
-\begin{figure*}[h]
+\begin{figure}[h]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.4\textwidth]{f10.pdf}
+		\includegraphics[width=0.3\textwidth]{f10.pdf}
 	\end{center}
@@
   \caption{ {\bf Different clustering techniques yield different
-  outcomes.} \nkinematic\ unique candidate cluster members found
-  using three different techniques are considered in this study.
-  \citetalias{cantatgaudin_gaia_2018}: \citet{cantatgaudin_gaia_2018},
+  candidate members of \cn.} \nkinematic\ unique candidate cluster
+  members found using three different techniques are considered in
+  this study.  \citetalias{cantatgaudin_gaia_2018}:
+  \citet{cantatgaudin_gaia_2018},
   \citetalias{kounkel_untangling_2019}:
@@
 	}
-\end{figure*}
-
-Figure~\ref{fig:venn} is a Venn diagram showing the overlap between
+\end{figure}
+
+Figure~\ref{fig:venn} is a Venn diagram of
 the three membership catalogs concatenated in this study.  99\% of the
@@
 both.  By comparison, 36\% of the \citetalias{kounkel_untangling_2019}
-sample, and 15\% of the \citetalias{meingast_2021} sample are
-non-overlapping.
-The data---Gaia DR2---used by all the studies was the same.
-What are the differences in methodology that produce these different
-outcomes?
-
-%
-% TODO: the following two paragraphs, WHEN QUOTED, are ENTIRELY FROM
-% the relevant papers. they should probably be nixed in favor of
-% concision.
-%
-\citetalias{cantatgaudin_gaia_2018} applied a procedure that 
-yielded what we call ``the core''.  Their procedure was to first query
-a Gaia DR2 cone around the previously reported RA and dec of the
-cluster, and within $\pm 0.5$\,mas of its previously reported
-distance. 
-% TODO: which was it?
-The outer radius of their cone was either r2 from MWSC (CITE
-Kharchenko 2013), or twice the ``cluster radius'' listed by DAML/Dias (CITE).  No
-proper motion cut was applied.  They then applied an unsupervised
-classification scheme called UPMASK to $G<18$\,mag stars within this
-cone (CITE).  The steps of UPMASK are first to perform a k-means
-clustering in the ``astrometric space'' ($\mu_{\alpha'}, \mu_\delta,
-\pi$).  Then, a ``veto'' step is applied to assess whether the groups
-of stars output from the k-means clustering are or are not more
-concentrated than a random distribution.  This is implemented by
-``comparing the total branch length of the minimum spanning tree
-connecting the stars with the expected branch length for a random
-uniform distribution covering the investigated field of view''.  ``To
-turn this yes/no flag to a membership probability,
-\citet{cantatgaudin_gaia_2018} then redraw new values of
-($\mu_{\alpha'}, \mu_\delta, \pi$) for each source based on the listed
-value, uncertainty, and covariance.  After a certain number of
-redrawings, the final probability is the frequency with which a given
-star passes the veto".  In the case of \cn, this yielded a reported
-``\texttt{r50}'' within which half of the cluster members were found
-to be within 0.496$^\circ$.
-When we selected candidate \cn\ members from the results of
-\citetalias{cantatgaudin_gaia_2018}, we opted to include all candidate members
-with reported membership probability exceeding 10\%.
-While this seems {\it a priori} low, our results ({\bf SECTION XXX})
-show that this ``membership probabilty'' severely underestimates the
-purity of the \citetalias{cantatgaudin_gaia_2018} sample for \cn. 
-%TODO: quantify
-Their false positive rate across the sample is more like 1-5\%.
-
-\citetalias{kounkel_untangling_2019}
-applied a different unsupervised
-clustering method to the 5-dimensional Gaia DR2 data (omitting radial
-velocities, due to their sparsity).  Their selection function (see
-their Section~2) yielded $\approx 2\times 10^7$ stars, mostly within
-$\approx 1$\,kpc and typically with $G<18$\,mag.  Their clustering
-algorithm, which was run on this entire  stellar sample,
- was the ``hierarchical density-based spatial clustering of
-applications with noise'' (HDBSCAN, CITE McInnes17).  The classical
-DBSCAN algorithm ``identifies clusters as overdensities in a
-multi-dimensional space in which the number of sources exceeds the
-required minimum number of points within a neighborhood of a
-particular linking length $\epsilon$.  HDBSCAN does not depend on
-$\epsilon$; instead it condenses the minimum spanning tree by pruning
-off the nodes that do not meet the minimum number of sources in a
-cluster and reanalyzing the nodes that do. Depending on the chosen
-algorithm, it would then either find the most persistent structure
-(through the excess of mass method), or return clusters as the leaves
-of the tree (which results in somewhat more homogeneous clusters). In
-both cases it is more effective at finding structures of varying
-densities in a given data set than DBSCAN.''
-`` The two main parameters that control HDBSCAN are the number of
-sources in a cluster and the number of samples. The former is the
-parameter that rejects groupings that are too small; the latter sets
-the threshold of how conservative the algorithm is in its
-considerations of the background noise (even if the resulting noisy
-groupings do meet the minimum cluster size).  By default, the sample
-size is set to the same value as the cluster size, but it is possible
-to adjust them separately.''
-Regarding membership proabilities,
-\citetalias{kounkel_untangling_2019} did not report continuous membership
-probabilities, instead opting for the binary ``member'' or ``not''.
-
-\citetalias{meingast_2021} did the same as
-\citetalias{kounkel_untangling_2019}, but required A) the tangential
-velocity dispersions to be smaller, and B) deconvolved the distances.
-{\bf FIXME FIXME FIXME}
-{\bf FIXME FIXME FIXME}
-{\bf FIXME FIXME FIXME}
-
-\citet{meingast_2021} reported a 
-Finally, 1577 of the 1860 \citet{meingast_2021} sources (85\%)
-were included in the \citet{kounkel_untangling_2019} sample.
+sample, and 15\% of the \citetalias{meingast_2021} sample do not
+overlap with any of the other catalogs.  The data---Gaia DR2---used by
+all the studies was the same.  What are the differences in methodology
+that produce these different outcomes?
+
+\citetalias{cantatgaudin_gaia_2018} applied a procedure that yielded
+what we colloquially call ``the core''.  Their membership assignment
+algorithm was to first query a Gaia DR2 cone around the previously
+reported $\{\alpha,\delta\}$ of the cluster center, and within $\pm
+0.5$\,mas of its previously reported parallax.  The outer radius of
+their cone was the previously reported angular radius of the cluster
+(0.71$^\circ$; \citealt{Kharchenko_et_al_2013}).  No proper motion cut
+was applied.  \citetalias{cantatgaudin_gaia_2018} then applied an
+unsupervised classification scheme to $G<18$\,mag stars within this
+cone (UPMASK; \citealt{kronemartins_upmask_2014}).  The UPMASK
+algorithm first performs a $k$-means clustering on the astrometric
+parameters, $\{\mu_{\alpha'}, \mu_\delta, \pi\}$.  A ``veto''
+step is then applied to assess whether the groups of stars output from the
+$k$-means clustering are more concentrated than a uniform
+distribution, by comparing the total branch lengths of their
+respective minimum spanning trees.  This binary flag is converted to a
+membership probability by redrawing values for $\{\mu_{\alpha'},
+\mu_\delta, \pi\}$ using the reported uncertainties and
+covariances.  The final membership probability for any given star is
+then the fraction of draws for which the star passes the veto step.
+In the case of \cn, the resulting radius within which half of the
+cluster members were found was 0.50$^\circ$.  We included all
+\citetalias{cantatgaudin_gaia_2018} \cn\ members with reported
+membership probability exceeding 10\% in our list of candidate cluster
+members.
+
+\citetalias{kounkel_untangling_2019} applied a different unsupervised
+clustering method to the 5-D Gaia DR2 positions and proper motions.
+\citetalias{kounkel_untangling_2019} began with a sample of
+$\approx$$2\times 10^7$ stars, mostly within $\approx 1$\,kpc and with
+$G\lesssim18$\,mag.  They then applied the hierarchical density-based
+spatial clustering of applications with noise algorithm (HDBSCAN;
+\citealt{campello_hierarchical_2015}, \citealt{mcinnes_hdbscan_2017})
+on the entire sample,  setting the minimum number of stars per cluster
+to be 40.  \citetalias{kounkel_untangling_2019} then iterated over a
+few different cutoff parallaxes to improve sensitivity to structures
+with distances from the Sun ranging between $\approx$100\,pc and
+$\approx$1000\,pc.  They then fitted ages to the resulting clusters
+using two different methods: isochronal, and a convolutional neural
+network.  Regarding membership proabilities,
+\citetalias{kounkel_untangling_2019} reported the binary ``member'' or
+``not'' from HDBSCAN.
+We included all \citetalias{kounkel_untangling_2019} \cn\ members in
+our list of candidate cluster members.
+
+Finally, \citetalias{meingast_2021} considered ten known open clusters
+within the nearest kiloparsec. One of these clusters was \cn.  Their
+initial  list of $\approx$$3\times10^7$ was similar to that of
+\citetalias{kounkel_untangling_2019}, mainly requiring well-behaved
+astrometry and photometry from Gaia.  Given the mean positions and
+velocities of the clusters determined by
+\citetalias{cantatgaudin_gaia_2018}, \citetalias{meingast_2021} then
+selected all stars with proper-motion correct tangential velocities
+within 1.5\,\kms\ of the cluster mean.  They then applied DBSCAN
+\citep{ester_density-based_1996} on both the observed galactic
+$\{X,Y,Z\}$ coordinates, as well as to a separate deconvolved spatial
+coordinate distribution discussed in their Section~3.3.  The results
+of their procedure, as well as those of
+\citetalias{cantatgaudin_gaia_2018} and
+\citetalias{kounkel_untangling_2019}, are visible at the website of
+S.~Meingast:
+\url{https://homepage.univie.ac.at/stefan.meingast/coronae.html}.
+Their reported membership proabilities were binary, similar to those
+of \citetalias{kounkel_untangling_2019}.  All of the
+\citetalias{meingast_2021} \cn\ members were included in our list of
+candidate cluster members.
 

### 2021-04-12 cluster_paper/ms.tex
update appendix 2

@@
 from \citet{meingast_2021} which were not available at the time of the
-CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
-$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
-roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at the
-distance of \cn.
+CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
+distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
+processing corresponds roughly to $\bpmrp$ of 2.2, or a spectral type
+of $\approx$M2V.
 
@@
   %\texttt{radvel} \citep{fulton_radvel_2018},
-  %\texttt{scikit-learn} \citep{scikit-learn},
+  \texttt{scikit-learn} \citep{scikit-learn},
   \texttt{scipy} \citep{jones_scipy_2001},
@@
 
-\section{Detrending details}
+\section{Considerations for Removing Systemic Variability from TESS Light Curves}
 \label{app:detrending}
 
-In ``detrending'' for our general variability search, our goal was to
-preserve astrophysical variability, while removing systematic
-variability.  One particular concern for the TESS light curves is
-systematic variability at the timescale of the 14-day satellite orbit,
-mostly induced by scattered light from the Earth and Moon.
-
-We therefore turned to the principal components (i.e., the
-eigenvectors) calculated following the procedure described by
-\citet{bouma_cdipsI_2019}. In brief, these vectors are computed using
-a set of ``trend stars'' selected from across each CCD according to
-ad-hoc heuristics that (hopefully) lead them to be dominated by {\it
-systematic} variability (Sec~3.7.2).
-
-The principal component vectors, also referred to as the eigenvectors,
-are rank-ordered by the degree of variance that they predict in the
-training set (of ``trend stars'').
+In ``detrending'' before searching for stellar rotation signals in the
+TESS, our goal was to preserve astrophysical variability, while
+removing systematic variability.  One particular concern for the TESS
+light curves is systematic variability at the timescale of the 14-day
+satellite orbit, mostly induced by scattered light from the Earth and
+Moon.
+
+We opted to use a variant of the principal component analysis approach 
+described by
+\citet{bouma_cdipsI_2019}. This
+principal component analysis focuses on a set of ``trend stars'' selected from
+across each CCD according using ad-hoc heuristics that on average lead
+the light curves to be dominated by systematic variability.  The
+resulting principal component vectors, also referred to as the
+eigenvectors, are rank-ordered by the degree of variance that they
+predict in the training set of trend stars.
 
 We then posit that any given target star's light curve is described as
-a linear combination of the eigenvectors.  Optionally, we also
-considered the inclusion of additional systematic vectors that could
-affect the light curve, such as the CCD temperature, the flux level
-measured in the background annulus, and the centroid positions of the
-stars on the CCDs.  These can be treated as additional ``features'' in
-the linear model.
-
-To determine the coefficients of the linear
-model after the full set of eigenvectors (plus optionally
-``sytematic'' vectors) had been asssembled,  we explored two possible
-methods: ordinary
-least squares, and ridge regression. Ridge regression is the same as
-ordinary least squares, except it includes an L2 norm with a regularization
-coefficient. The regularization coefficient that best applied for any
-given target light curve was solved for using a cross-validation grid
-search, using \texttt{sklearn.linear\_modelRidgeCV} (CITE).
-
+a linear combination of the eigenvectors.  We also considered the
+inclusion of additional systematic vectors that could affect the light
+curve, discussed below.  These can be treated as additional ``features'' in the linear
+model.  To determine the coefficients of the linear model after the
+full set of eigenvectors (plus optional ``sytematic'' vectors) had
+been asssembled,  we explored two possible methods: ordinary least
+squares, and ridge regression. Ridge regression is the same as
+ordinary least squares, except it includes an $\ell^2$ norm with a
+regularization coefficient. The regularization coefficient that best
+applied for any given target light curve was determined using a
+cross-validation grid search, through
+\texttt{sklearn.linear\_modelRidgeCV} \citep{scikit-learn}.
 Each target light curve was mean-subtracted and normalized by its
 standard deviation, as were the eigenvectors. The linear problem was
-then solved numerically, and the light curve was reconstructed by re-adding the
-original mean, and re-multiplying by the standard deviation to ensure
-that the variance of the light curve did not change.
-
-We found that the choice of using ordinary
-least squares versus ridge regression did not seem to significantly
-affect the resulting light curves. In other words, the inclusion (or
-lack thereof) of a regularization term did not strongly alter the
-best-fitting coefficients.
-In the spirit of ``KISS'', we opted to use ordinary least squares.
-
-A few other choices seemed to be more important:
+then solved numerically, and the light curve was reconstructed by
+re-adding the original mean, and re-multiplying by the standard
+deviation to ensure that the variance of the light curve did not
+change.
+
+We found that the choice of using ordinary least squares versus ridge
+regression did not seem to significantly affect the resulting light
+curves. In other words, the inclusion (or lack thereof) of a
+regularization term did not strongly alter the best-fitting
+coefficients.  In the spirit of keeping it simple, we opted to use
+ordinary least squares.  A few other choices seemed to be more
+important:
 
 \begin{itemize}
-
+  %
   \item {\it To smooth, or to not smooth the eigenvectors}.
-  Ideally, the eigenvectors should be smooth in time. They should not contain
-    residuals from {\it e.g.}, eclipsing binaries that snuck their way
-    into the template set, and they should also not be intrinsically noisier
-    than the target star. If either of these is the case (and we found that
-    it sometimes was), it can
-    induce extra variability into the PCA ``detrended'' light curves.
-    To address this problem, we opted to smooth the eigenvectors using
-    a time-windowed filter (with a "biweight" weight scheme, implemented in
-    \texttt{wotan} by \citet{hippke_wotan_2019}; window length 1 day, cval 6).
-  One issue with this is that systematic sharp features (captured e.g., in
-  "spike vectors") no longer are captured, so they end up in the "PCA
-  detrended" light curves. They can then be filtered out
-  ({\it e.g.}, using rolling outlier rejection), and we prefer this approach to having
-  systematic features being {\it injected by} the PCA detrending.
-
+    If the systematic trends are smooth in time, the eigenvectors should be as well. They should not
+    contain residuals from {\it e.g.}, eclipsing binaries 
+    in the template set, and they should also not be
+    intrinsically noisier than the target star. If either of these is
+    the case, extra
+    variability can be introduced into the ``detrended'' light curves.  To address
+    this problem, we opted to smooth the eigenvectors using a
+    sliding time-windowed filter (with a "biweight" weight scheme, implemented
+    in \texttt{wotan} by \citealt{hippke_wotan_2019}).
+    %; window length 1 day, cval 6).
+    One issue with this is that systematic sharp features (captured
+    e.g., in ``spike vectors'') are no longer captured, so they are present
+    in the detrended light curves. Since they can be filtered out in postprocessing
+    ({\it e.g.}, using rolling outlier rejection), we prefer this
+    approach to having systematic features being injected by the
+    PCA detrending.
+  %
   \item {\it How many eigenvectors to use}.
     A larger number always leads to greater whitening.  In
-    \citet{bouma_cdipsI_2019}, we performed a Factor Analysis
+    \citet{bouma_cdipsI_2019}, we performed a factor analysis
     cross-validation to determine the number of eigenvectors to use.
-    The typical number adopted based on this analysis was 10--15.
-    While this approach should in theory prevent over-fitting, in
-    our experience, for stellar rotation it still often lead to
-    distorts the signals, especially for rotation signals with small amplitudes
-    and periodicities of
-    $\gtrsim 3$ days.  (Shorter signals typically are not distorted,
-    since the eigenvectors do not contain the high-frequency
-    content that leads to the distortions).  For the present analysis, we
-    therefore impose the maximum number of eigenvectors to be 5.
-
-  \item {\it Which supplementary systematics vectors to use}.
-    We considered using the \texttt{BGV}, \texttt{CCDTEMP},
-    \texttt{XIC}, \texttt{YIC}, and \texttt{BGV} vectors, packaged
-    with the CDIPS light curves. We found that the background value
-    measured in an annulus centered on the aperture, \texttt{BGV},
-    tended to produce the best independent information from the PCA
-    eigenvectors, and so we adopted it as our only ``supplementary''
-    trend vector.  We opted to not smooth it (in hopes that it would
-    provide direct complement to the smoothed PCA vectors; 1 sharp
-    vector containing literally the background information, plus 5
-    smooth vectors).
+    The typical number adopted based on this analysis was 10 to 15.
+    While this approach should in theory prevent over-fitting, in our
+    experience, for stellar rotation it still often lead to distorts
+    the signals, especially for rotation signals with small amplitudes
+    and periodicities of $\gtrsim 3$ days.  Shorter signals typically
+    are not distorted, since the eigenvectors do not contain the
+    high-frequency content that leads to the distortions.  For the
+    present analysis, we therefore set the maximum number of
+    eigenvectors to be 5.
+%
+  \item {\it Which supplementary systematics vectors to use}.  We
+    considered using the \texttt{BGV}, \texttt{CCDTEMP}, \texttt{XIC},
+    \texttt{YIC}, and \texttt{BGV} vectors, packaged with the CDIPS
+    light curves. We found that the background value measured in an
+    annulus centered on the aperture, \texttt{BGV}, tended to produce
+    the best independent information from the PCA eigenvectors, and so
+    we adopted it as our only ``supplementary'' trend vector.  We
+    opted to not smooth it, assuming that it would provide direct
+    complement to the smoothed PCA vectors.
 \end{itemize}
 
-For every ``target star'', we then decorrelated the raw
-(image-subtracted and background-subtracted) light curve
-using a linear model with ordinary least squares.
+After all these considerations, for every ``target star'', we
+ultimately decorrelated the raw (image-subtracted and
+background-subtracted) light curve using a linear model with ordinary
+least squares.  The vectors considered in the model were the 1
+unsmoothed background flux vector, and 5 smoothed PCA eigenvectors.
 
@@
   Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
-  considered. (Fainter stars in the field sample were not studied as part
-  of the broader CDIPS project).
-  The field stars show a different rotation period distribution than
-  the kinematically selected NGC\,2516 members.
+  considered. (Fainter stars in the field sample were not studied as
+  part of the broader CDIPS project).  The field stars show a
+  different rotation period distribution than the kinematically
+  selected NGC\,2516 members.
   {\bf LGB TODO: Rerun the field measurements with a better grid to fix the aliasing issue.}

### 2021-04-16 cluster_paper/ms.tex
update intro
Co-authored-by: joshwinn <jwinn@astro.princeton.edu>

@@
 \begin{abstract}
-  Recent analyses of the Gaia data have reported the existence of
-  diffuse stellar populations surrounding nearby open clusters.  It is
-  important to verify that these ``halos'', ``tails'', and ``strings''
-  are of similar ages and compositions as stars in the denser part of
-  the cluster.  We present an analysis of NGC\,2516
-  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
-  and an apparent halo of stars spanning 500\,pc.  Combining
-  photometry from Gaia, rotation periods from TESS, and lithium
-  measurements from Gaia-ESO and GALAH, we find that the halo of
-  NGC\,2516 is real and coeval with the cluster's core.  Two thirds of
-  kinematically selected halo members out to 250\,pc from the cluster
-  center have rotation periods consistent with a gyrochronological age
-  of 150\,Myr.  A comparison sample of field stars shows no such
-  trend.  In addition, the lithium abundances of stars in the halo are
-  elevated relative to the field, and are correlated with the stellar
-  rotation rate in a manner consistent with previous observations of
-  similar clusters.  Broadly speaking, this work advances a paradigm
-  in which the halos of open clusters are often more populous than
-  their cores.  There are significant implications for spectroscopic
-  survey targeting, open cluster dispersal, and planet searches around
-  young stars.
+  Recent analyses of the Gaia data have identified diffuse stellar
+  populations surrounding nearby open clusters.  It is important to
+  verify that these ``halos'', ``tails'', and ``strings'' are of
+  similar ages and compositions as stars in the denser part of the
+  cluster.  We present an analysis of NGC\,2516 ($\approx$150\,Myr),
+  which has a classical tidal radius of 10\,pc and an apparent halo of
+  stars spanning 500\,pc.  Combining photometry from Gaia, rotation
+  periods from TESS, and lithium measurements from Gaia-ESO and GALAH,
+  we find that the halo of NGC\,2516 is the same age as the cluster's
+  core.  Two thirds of kinematically selected halo members out to
+  250\,pc from the cluster center have rotation periods consistent
+  with a gyrochronological age of 150\,Myr.  A comparison sample of
+  field stars shows no such trend.  In addition, the lithium
+  abundances of stars in the halo are elevated relative to the field,
+  and are correlated with the stellar rotation rate in a manner
+  consistent with previous observations of similar clusters.  This
+  work supports a new paradigm wherein the halos of open clusters are
+  often more populous than their cores.  We discuss implications for
+  spectroscopic survey targeting, open cluster dispersal, and planet
+  searches around young stars.
 \end{abstract}
@@
 photoionization, radiation pressure, and supernova shocks disperse the
-gas out of the cloud \citep{krumholz_star_2019}.  Since only
-$\sim$$1\%$ of the cloud mass is converted into stars, gas dispersal
-enables the majority ($\sim$90\%) of stars in the cluster to escape
-the cluster's gravitational well \citep{lada_embedded_2003}.
+gas \citep{krumholz_star_2019}.  Since the gas represents about 99\%
+of the mass of the original cloud, gas dispersal enables the majority
+($\sim$90\%) of stars in the cluster to escape the cluster's
+gravitational well \citep{lada_embedded_2003}.
 
 From 10 to 1000\,Myr, the cluster remnants that survive gas dispersal
-suffer an onslaught that almost always leads to dissolution.  Inside
-the cluster supernovae, AGB winds, and close stellar encounters lead
-to additional mass loss \citep{lamers_mass_loss_2010}.  Extrinsic to
-the cluster, collisions with giant molecular clouds
-\citep{spitzer_disruption_1958}, and perturbations from the galactic
-tide in both the radial and vertical dimensions further promote
-stellar escape \citep[{\it
+suffer an onslaught of supernovae, AGB winds, and close stellar
+encounters that almost always leads to dissolution
+\citep{lamers_mass_loss_2010}.  Extrinsic to the cluster, collisions
+with giant molecular clouds \citep{spitzer_disruption_1958}, and
+perturbations from the galactic tide in both the radial and vertical
+dimensions further promote stellar escape \citep[{\it
 e.g.},][]{fukushige_timescale_2000,bergond_gravitational_2001}.
 
-Finding the stars that have dispersed from their natal clusters into
-the galactic field is a pressing topic for a few reasons.  One is to
+Finding the stars that have dispersed from their clusters into the
+galactic field is a pressing topic for a few reasons.  One is to
 understand the spatial extent and duration of star formation events
@@
 
-The detection of dispersed stellar associations, although possible
-before Gaia, has now reached a breakthrough pace \citep[{\it
+Although it has been possible to detect dispersed stellar associations
+for a long time, Gaia is now enabling major advances \citep[{\it
 e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
-The populations found by these studies can be summarized as follows.
-On one end are coeval groups with no discernable cores ({\it e.g.},
-Psc Eri and $\mu$ Tau,
+The populations found by the most recent studies can be summarized as
+follows.  On one end are coeval groups with no discernable cores ({\it
+e.g.}, Psc Eri and $\mu$ Tau,
 \citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
@@
 and under-densities  \citep[{\it e.g.} the Sco-Cen and Vela
-associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
-focus on an intermediate regime: low-density halos associated with a
-single overdensity, typically an open cluster that was known before
+associations,][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here,
+we focus on an intermediate regime: low-density halos associated with
+a single overdensity, typically an open cluster that was known before
 Gaia
 \citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
-In some cases, these halos could correspond to tidal tails, as have
-been reported in the Hyades
-\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
-group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
-\citep{tang_comaber_2019}.
-
-One point of difficulty however is that different clustering methods
-yield differing levels of sensitivity and precision
-\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
-tautologically yields open clusters that are elliptical \citep[{\it
-e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
-and asymmetric structures with number densities down to 100 times less
-than the field around the same regions (see {\it e.g.},
-\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
-Differences between different clustering methods raise the issue of
-how contaminated the resulting samples are, and in some cases whether
-the reported structures are truly coeval stellar populations at all.
-
-We have recently been making TESS light curves of stars reported to be
-members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}) and
-searching them for planets \citep{bouma_cluster_2020}.
-Our analysis of TESS Sectors 1-13 yielded light curves of
-\numsouthernuniqlcs\ candidate cluster members in the Southern
-Ecliptic hemisphere, which are available on
+In some cases, these halos may actually be tidal tails, as have been
+reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
+the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
+Berenices \citep{tang_comaber_2019}.
+
+One point of difficulty however is that different algorithms for
+identifying clusters yield differing levels of sensitivity and
+precision \citep{hunt_clustering_2020}.  For example, using a Gaussian
+Mixture Model tautologically yields open clusters that are elliptical
+\citep[{\it e.g.},][]{wallace_m4_2018}.  Some unsupervised methods
+yield dispersed and asymmetric structures with number densities down
+to 100 times lower than the field around the same regions ({\it
+e.g.}, \citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
+Since different techniques yield quantitatively and
+qualitatively different outcomes, the purity of kinematically selected
+samples is important to verify through independent lines of evidence,
+such as rotation periods and spectroscopic indicators.
+
+As part of a Cluster Difference Imaging Photometric Survey (CDIPS,
+\citealt{bouma_cdipsI_2019}), we have recently been making TESS light
+curves of stars reported to be members of coeval populations and
+searching them for planets \citep{bouma_cluster_2020}.  Our analysis
+of TESS Sectors 1-13 yielded light curves of \numsouthernuniqlcs\
+candidate cluster members in the Southern Ecliptic hemisphere, which
+are available on
 MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
 
-As part of this broader project, we focus here on a modest question:
-in just a single southern open cluster, is the cluster halo truly
-coeval with the core?  We chose NGC\,2516 for this analysis since it
-is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
-to facilitate photometric rotation measurements using TESS.  Its halo
-was recently reported by \citet{kounkel_untangling_2019} to span
-$\approx20\times10\times350$\,pc.  We want to know: is this halo real?
-To what extent can we use Gaia alone to reliably identify age-dated
-needles in the haystack of field stars?  What are the implications,
-observationally and theoretically, if we can identify the dispersed
-halos of open clusters?
+As part of this project, we focus here on a single southern open
+cluster, and ask: is the cluster halo truly coeval with the core?  We
+chose \cn\ for this analysis because it is young (100--200\,Myr) and
+sufficiently nearby ($d\approx400$\,pc) to facilitate photometric
+rotation measurements using TESS.  Its halo was recently reported by
+\citet{kounkel_untangling_2019} to span
+$\approx20\times10\times350$\,pc.  We want to know: is this halo real
+or is it just a collection of interlopers, foreground and background
+stars?  To what extent can we use Gaia alone to reliably identify
+age-dated needles in the haystack of field stars?  What are the
+implications, observationally and theoretically, if we can identify
+the dispersed halos of open clusters?
 
@@
 The latter term implies a particular model for the formation of the
-dispersed group.  ``Halo'' is concise and model agnostic, but suffers
-from the suggestion of spherical symmetry.  The halo of NGC\,2516 is
-most accurately described as consisting of a leading and trailing
-tail. However, the goal of this study is to clarify whether the stars
-in these tails are the same age as stars in the core, and not to make
-a ruling on the origin of the overall structure.  For this reason, we
-adopt the model-agnostic ``halo''.
+dispersed group.  ``Halo'' is concise and model agnostic, but is not
+an ideal term because it connotes spherical symmetry, which is usually
+not the case.  The halo of \cn\ is most accurately described as
+consisting of a leading and trailing tail.  We call it a halo for
+simplicity.
 
@@
 discuss the implications of our analysis for NGC\,2516 specifically
-and stellar spin-down more generally.
-Section~\ref{sec:conclusion} gives our conclusions.
+and stellar spin-down more generally.  Section~\ref{sec:conclusion}
+gives our conclusions.
 
@@
 accuracy of the Gaia-derived results are the current state of the art.
-We therefore adopted what we viewed as the most important clustering
+We therefore adopted what we viewed as the most important Gaia-based
 samples to compare: those of \citet{cantatgaudin_gaia_2018},
@@
   made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}.
-While we considered performing our own clustering analysis on the
-Gaia data, such an effort would in effect only replicate the work of
-these investigators.  We opt to instead use their studies as a starting
-point.
+%While we considered performing our own clustering analysis on the
+%Gaia data, such an effort would in effect only replicate the work of
+%these investigators.  We opt to instead use their studies as a starting
+%point.
 
@@
 
-The authors thank L.~Hillenbrand, B.~Tofflemire, G.~Zhou, E.~Newton, A.~Mann, and A.~Kraus for fruitful discussions.
-
+The authors thank L.~Hillenbrand, B.~Tofflemire, G.~Zhou, E.~Newton,
+K.~Hawkins, A.~Mann, and A.~Kraus for the discussions on young stars,
+rotation, and lithium that encouraged this line of study.
 %

### 2021-04-20 cluster_paper/ms.tex
update ms
Co-authored-by: joshwinn <jwinn@astro.princeton.edu>

@@
 \newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
-\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nhalo}{2{,}192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes
 
@@
 sequence, though with greater scatter.  One possible explanation for
-the additional scatter could be that the halo is more contaminated by
+the additional scatter is that the halo is more contaminated by
 field stars.  For instance, $\approx5$ red giants in the halo must be
@@
 Another possible explanation for scatter in the halo's HR diagram
-could be differential reddening across different sightlines.  The
+is differential reddening across different sightlines.  The
 reported halo spans 20$^\circ$ on-sky, and varies in position from
@@
 purposes of age-dating the cluster we focus on the main-sequence
-turn-off (MSTO), since this is where the models have maximal
-predictive power.
+turn-off (MSTO), since this is where the models have maximum
+sensitivity to age.
 
@@
 consistent.  The left panel of Figure~\ref{fig:hr} suggests that
-isochronally, they are: stars past the turnoff in both the halo and
-core are on a consistent locus.  The right panel of
+isochronally, they are consistent: stars past the turnoff in both the
+halo and core are on the same locus.  The right panel of
 Figure~\ref{fig:hr} demonstrates the precision with which the claim
@@
 Details regarding our detrending approach are presented in
-Appendix~\ref{app:detrending}.  After applying the initial ensemble
-decorrelation to remove systematic trends, we proceeded with a few
-cleaning steps: we masked 0.7 days at the beginning and end of each
-spacecraft orbit, and ran a sliding standard-deviation rejection
+Appendix~\ref{app:detrending}.  After detrending, we proceeded with a
+few cleaning steps: we masked 0.7 days at the beginning and end of
+each spacecraft orbit, and ran a sliding standard-deviation rejection
 window over the light curve, which removed any outlying points within
@@
 from the target star and background-light \citep{Sullivan_2015}.  This
-typically resulted in an aperture radius of either 1 or 1.5 pixels.
+typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
 We recorded the top five periodogram peaks from each method, and their
@@
   {\it bottom panels}, the innermost and outermost 100 stars in
-  position and velocity space are displayed in the space of rotation
-  periods and dereddened colors.
+  position and velocity space (black and green) are shown in the
+  space of rotation periods and dereddened colors.
   \label{fig:physical_x_rotn}
@@
 
-The TESS rotation periods can provide an independent check on the
+The TESS rotation periods provide an independent check on the
 Gaia-derived kinematic cluster memberships.  We explored this by
 cross-matching the stars with detected TESS rotation periods against
-our original target list of \nkinematic\ Gaia DR2 members.
+our original target list of \nkinematic\ Gaia members.
 
@@
 denote $\Delta v^{*}$, per \citetalias{meingast_2021}.  We convert
-these proper motion differences to physical units by dividing by the
-distance to each star, taken as the inverse parallax.
+these proper motion differences to physical units by multiplying by
+the measured parallax.
 
@@
 comparison, we only show stars with $0.5<\bpmrp<1.2$ for which our
-TESS pipeline succeeding in making light curves.  In other words---the
-base sample is the stars for which we could have plausibly detected a
-rotation period.  This color range is preferred since the the slow
-sequence becomes less defined for spectral types later than
+TESS pipeline succeeding in making light curves.  In other words, the
+stars in the base sample are those for which we could have plausibly
+detected a rotation period.  This color range is preferred because the
+the slow sequence becomes less defined for spectral types later than
 $\approx$K4V.  The bins in the histograms (top-right) were chosen to
@@
 
-Even with contaminants, our rotation period measurements show that the
-halo extends to separations of $\approx$250\,pc in physical space from
-the cluster core.  The most widely separated F2V-K2V stars with
-rotation periods (blue points in Figure~\ref{fig:physical_x_rotn}) are
-separated by $\approx$430\,pc.  The total length of the structure is
-therefore 400--500\,pc, depending on which members of the halo are
-chosen as the ``tips'' on either end.  This agrees with the overall
-structure of the halo reported by
+Even with this level of contamination, our rotation period
+measurements show that the halo extends to separations of
+$\approx$250\,pc in physical space from the cluster core.  The most
+widely separated F2V-K2V stars with rotation periods (blue points in
+Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
+The total length of the structure is therefore 400--500\,pc, depending
+on which members of the halo are chosen as the ``tips'' on either end.
+This agrees with the overall structure of the halo reported by
 \citetalias{kounkel_untangling_2019}, with some minor caveats visible
@@
 has recently been surveyed by \citet{bouvier_lithium-rotation_2020}.
-For \cn, two spectroscopic datasets seemed particularly important:
+For \cn, two spectroscopic datasets seemed important:
 Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
@@
 and publicly available photometry.  Since the existence of the \cn\
-halo was not known at the time of target selection, very few ``halo''
+halo was not known at the time of target selection, very few halo
 stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
-ultimately reported stellar parameters (including lithium equivalent
+reported stellar parameters (including lithium equivalent
 widths and metallicities) for 796 stars that they considered possible
@@
 6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
-to the continuum normalized spectra, considering only a $\pm 1\,$\AA
+to the continuum normalized spectra, considering only a $\pm 1\,$\AA\ 
 window centered on the Li doublet.  The equivalent widths were then
@@
 Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
-Gaia-ESO and GALAH results, with the points colored by whether the
+Gaia-ESO and GALAH results, with the points colored according to whether the
 stars are in the core or the halo of the cluster.  The GALAH spectra
@@
 dwarfs.  Similar trends were noted in the Pleiades over three decades
-ago \citep{butler_pleiades_1987}, and are most likely caused by
-intrinsic differences in lithium abundance
+ago \citep{butler_pleiades_1987}, and are thought to be caused by
+differences in lithium abundance
 \citep{soderblom_evolution_1993}.  Alternative explanations, such as
@@
 One possible explanation is that the observed halo structures are
-tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea would
-be that as the cluster contracts, stars escape out of the Lagrange
-points imposed by the Milky Way's potential, and form leading and
-lagging arms due to differential rotation in the Galaxy.  Kinematic
-evidence that the core of \cn\ may be undergoing contraction has
-recently been presented by \citet{healy_stellar_2020}, and might
-support this interpretation.  Whether the exact rate of contraction,
-and the corresponding ``evaporation rate'' of the core is sufficient
-to explain the formation of the halo has yet to be assessed.  The
+tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea is
+that as the cluster contracts, stars escape out of the Lagrange points
+imposed by the Milky Way's potential, and form leading and lagging
+arms due to differential rotation in the Galaxy.  Kinematic evidence
+that the core of \cn\ may be undergoing contraction has recently been
+presented by \citet{healy_stellar_2020}, which might increase the rate
+at which stars escape.  Whether the exact contraction rate, and the
+corresponding evaporation rate of the core are sufficient to
+explain the formation of the halo has yet to be assessed.  The
 presence or lack of epicyclic overdensities in the tails might also
@@
 
-The existence of a slow sequence is the main premise of gyrochronology
+The existence of a slow sequence is the main basis of gyrochronology
 \citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
 spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
-\citet{skumanich_time_1972} has not worked in detail, the general idea
-is correct, and the first-order effect can be understood through
-magnetic breaking \citep{weber_angular_1967}.
+\citet{skumanich_time_1972} is not an accurate enough description, the
+general idea is correct, and the first-order effect can be understood
+through magnetic braking \citep{weber_angular_1967}.
 
 The other features of the rotation-color diagram are indications that
-magnetic breaking is not always the most important effect.  The
+magnetic braking is not always the most important effect.  The
 question of why the rapid rotators exist, for instance, has not yet
@@
 the dynamo to couple more strongly with the stellar wind
-\citep{brown_metastable_2014}.  Other processes could be external the
+\citep{brown_metastable_2014}.  Other processes could be external to the
 star entirely: \citet{qureshi_signature_2018} for instance have
@@
 While many rapid rotators in \cn\ (34\%) have no evidence for being
-binary, we emphasize that the cuts we used to select binaries (${\rm
+binaries, we emphasize that the cuts we used to select binaries (${\rm
 RUWE}>1.2$; photometric excess exceeding 0.3\,mag) leave open a
@@
 Comparing the explanations of disk-locking against tidal
-synchronization, the population statistics seem to strain the
+synchronization, the population statistics seem to rule out the
 possibility of tidal synchronization.  A quarter of the \cn\ members
 are rapidly rotating.  In comparison, in the field half of Sun-like
-stars are binaries, and $\approx$9\% of these binaries have periods below 100\,days
-\citep{raghavan_survey_2010}.  If we assume that all binaries with
-sub-100\,day periods become tidally synchronized, then we can only explain
-a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity
-fractions in pre-MS stars likely do not change this picture \citep[see
-Sec.~4.4 of][]{duchene_stellar_2013}.
+stars are binaries, and only $\approx$9\% of these binaries have
+periods below 100\,days \citep{raghavan_survey_2010}.  If we assume
+that all binaries with sub-100\,day periods become tidally
+synchronized, then we can only explain a rapid rotator occurrence rate
+of $\approx$5\%.  Elevated binarity fractions in pre-MS stars likely
+do not change this picture \citep[see Section~4.4
+of][]{duchene_stellar_2013}.
 
@@
 sufficient to inhibit the convection \citep{ventura_Li_B_1998}.
-Alternatively, the star's structure could be at fault: rapid rotators
+Alternatively, the star's temperature and density profiles could be at fault: rapid rotators
 have the strongest magnetic fields and the most spotted surfaces.
@@
 
-None of the internal processes noted above explain the initial
+None of the internal processes noted above answer the initial
 question of why some of the G and K dwarfs rotate rapidly to begin
@@
 $125\pm20$\,Myr, this yields a MSTO age for \cn\ of $175\pm35$\,Myr.
-The gyrochronological age we have argued is likely super-Pleiades,
-{\it i.e.}, in the range of 150 to 200\,Myr.  The lithium age based on
-the depletion of the G and K dwarfs is more uncertain, but is
-consistent with the Pleiades and could be a bit older ($\approx$100 to
-200\,Myr).  Averaging these three different indicators gives an age
-for \cn\ of $167\pm20$\,Myr, if we set the uncertainties to match the
-absolute Pleiades age uncertainty of $\pm 20\,$Myr
-\citep{soderblom_ages_2014}.  Whether this absolute age is at all
-meaningful, we leave the reader to judge.
+The gyrochronological age we have argued is likely older than the
+Pleiades, {\it i.e.}, in the range of 150 to 200\,Myr.  The lithium
+age based on the depletion of the G and K dwarfs is more uncertain,
+but is consistent with the Pleiades and could be a bit older
+($\approx$100 to 200\,Myr).  Averaging these three different
+indicators gives an age for \cn\ of $167\pm20$\,Myr, if we set the
+uncertainties to match the absolute Pleiades age uncertainty of $\pm
+20\,$Myr \citep{soderblom_ages_2014}.  Whether this absolute age is at
+all meaningful, we leave the reader to judge.
 
@@
 
-The existence of the halo itself should not be not very surprising.  A
+The existence of the halo itself is physically plausible.  A
 star with a 1\kms\ velocity difference from the cluster average will
@@
 We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
-Latham, Sara Seager, Josh Winn, Jon Jenkins) and the many TESS team
+Latham, Sara Seager, Jon Jenkins) and the many TESS team
 members for their efforts to make the mission a continued success.
@@
 
-\section{Considerations for Removing Systemic Variability from TESS Light Curves}
+\section{Considerations for Removing Systematic Variability from TESS Light Curves}
 \label{app:detrending}

### 2021-04-20 earhart/plotting.py

@@ -2673,8 +2673,14 @@ def plot_rotation_X_positions(outdir, basedata='fullfaint', cmapname=None,
 
         axs[0].text(0.03, 0.96, '$\Delta r_{\mathrm{3D}}$: '+f'{mdf0.delta_r_pc.min():.1f}-{mdf0.delta_r_pc.max():.1f} pc', va='top',
                     ha='left', transform=axs[0].transAxes, color='k', bbox=bbox)
-        axs[0].text(0.03, 0.88, '$\Delta r_{\mathrm{3D}}$: '+f'{mdf1.delta_r_pc.min():.1f}-{mdf1.delta_r_pc.max():.1f} pc', va='top',
-                    ha='left', transform=axs[0].transAxes, color='lime', bbox=bbox)
+        _txt = axs[0].text(0.03, 0.88, '$\Delta r_{\mathrm{3D}}$: '+f'{mdf1.delta_r_pc.min():.1f}-{mdf1.delta_r_pc.max():.1f} pc', va='top',
+                    ha='left', transform=axs[0].transAxes, color='forestgreen', bbox=bbox)
+
+        _txt.set_path_effects([
+            pe.Stroke(linewidth=0., foreground='black'),
+            pe.Normal()]
+        )
+
 
         # now the v_tang...
         sort_df = mdf.sort_values(by='delta_mu_prime_km_s')
@@ -2701,8 +2707,14 @@ def plot_rotation_X_positions(outdir, basedata='fullfaint', cmapname=None,
         axs[1].text(0.03, 0.96, txt0, va='top', ha='left',
                     transform=axs[1].transAxes, color='k', bbox=bbox)
         txt1 = '$\Delta v_{\mathrm{2D}}^{*}$: '+f'{mdf1.delta_mu_prime_km_s.min():.1f}-{mdf1.delta_mu_prime_km_s.max():.1f}'+' km$\,$s$^{-1}$'
-        axs[1].text(0.03, 0.86, txt1, va='top', ha='left',
-                    transform=axs[1].transAxes, color='lime', bbox=bbox)
+        _txt = axs[1].text(0.03, 0.86, txt1, va='top', ha='left',
+                    transform=axs[1].transAxes, color='forestgreen', bbox=bbox)
+
+        _txt.set_path_effects([
+            pe.Stroke(linewidth=0., foreground='black'),
+            pe.Normal()]
+        )
+
 
 
     # axes labels
[…]

### 2021-04-20 cluster_paper/ms.tex
update ms
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@
 
+% Cluster Difference Imaging Photometric Survey. III.
+% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
 \title{
-  Cluster Difference Imaging Photometric Survey. III.
-  Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
+  Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open
+  Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference
+  Imaging Photometric Survey.}
 }
@@
   performed using the \texttt{astrobase.services.dust} module
-  \citep{bhatti_astrobase_2018}.  } and retrieved the extinction
-parameters at the positions of each NGC\,2516 member.  The mean and
-standard deviation of the $E(\mathrm{B}-\mathrm{V})$ values for the
-\citet{schlegel_maps_1998} and \citet{schlafly_measuring_2011} maps
-were consistent, so we adopted the average from
-\citet{schlegel_maps_1998}: $E(\mathrm{B}-\mathrm{V})_{\rm
-S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
-the \citet{schlegel_maps_1998} maps overestimate the reddening values
-when the color excess exceeds about 0.10\,mag. Applying the correction
-proposed by \citet{bonifacio_search_2000} leads to our adopted
-$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$.  Finally, to convert to Gaia
+  \citep{bhatti_astrobase_2018}.  } and retrieved the total
+line-of-sight extinction parameters at the positions of each \cn\ 
+member.  The mean and standard deviation of the
+$E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
+and \citet{schlafly_measuring_2011} maps were consistent, so we
+adopted the average from \citet{schlegel_maps_1998}:
+$E(\mathrm{B}-\mathrm{V})_{\rm S}=0.206\pm0.039$.
+\citet{bonifacio_search_2000} found however that the
+\citet{schlegel_maps_1998} maps overestimate the reddening values when
+the color excess exceeds about 0.10\,mag. We therefore applied the
+correction proposed by \citet{bonifacio_search_2000}, and additionally
+corrected for the mean cluster distance by a factor
+$\exp(-\abs{d\sin(b)}/h)$ where $d$ is the average cluster distance,
+$b$ is the average galactic latitude, and $h$ is the scale height of
+the galactic disk, assumed to be 125\,pc.  This yielded our adopted
+$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$, where the uncertainty
+reflects the standard deviation of the individual
+\citet{schlegel_maps_1998} values.  Finally, to convert to Gaia
 colors we used the calibration of \citet{stassun_TIC8_2019}, namely
@@
 \citep{baratella_gaiaeso_2020}, which represented an up-revision from
-an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
-\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
-cluster, we prefer super-solar based on the goodness of fit from the
-photometry.
+an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We
+caution though that other sub-solar metallicities have been reported
+\citep{bailey_rv_2018}, and note that the cluster metallicity and
+reddening are degenerate; if the reddening is lowered, the implied
+metallicity will increase.  While a detailed redetermination of
+these parameters from the Gaia photometry is beyond our scope, our
+adopted values for both the reddening and metallicity are within the
+range of previously reported values in the literature.
 
@@
   stars).
+  Rotation periods for comparison field stars are shown in
+  Appendix~\ref{app:compstar}.
   \label{fig:rot}
@@
 
+The resulting rotation periods and periodogram powers are reported,
+regardless of detection significance, in Table~\ref{tab:maintable}.
 To clean the resulting rotation period measurements, we designed a set
@@
 no equal-brightness or greater companions were within the aperture.
-Beyond requiring that the target star be the brightest star in the
-aperture, we also required that at most one companion with brightness
-exceeding one tenth of the target star's brightness be present in the
-aperture according to the Gaia DR2 source catalog.  For set
-$\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out of
-\nautorotdenominator\ stars for which rotation periods might plausibly
-have been detected.  For set $\mathcal{B}$, we additionally required
-that the Lomb-Scargle and Stellingwerf phase-dispersion methods
-yielded identical periods to within 10\%.  This yielded
-\nautorotnumeratormatching\ light curves.
+These limits were chosen after inspecting the light curves visually,
+to eliminate low-quality rotation periods while retaining high
+completeness.  Beyond requiring that the target star be the brightest
+star in the aperture, we also required that at most one companion with
+brightness exceeding one tenth of the target star's brightness be
+present in the aperture according to the Gaia DR2 source catalog.  For
+set $\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out
+of \nautorotdenominator\ stars for which rotation periods might
+plausibly have been detected.  For set $\mathcal{B}$, we additionally
+required that the Lomb-Scargle and Stellingwerf phase-dispersion
+methods yielded identical periods to within 10\%.  This yielded
+\nautorotnumeratormatching\ light curves.  Boolean flags for each set
+are included in Table~\ref{tab:maintable}.
 
@@
 cluster core.  The average detection rate inside of 25\,pc, 73\%, is
-slightly higher.
+slightly higher.  With $\approx$3.5\,pc, the period detection rate of
+61\% is relatively low, due to the combined effects of crowding and
+saturation.
 
@@
   EW}\approx0\,$m\AA\ are non-detections.
+  The Li measurements for both Core and Halo stars in \cn\ are
+  consistent with a near-Pleiades age.
   {\bf LGB TODO: the increasing EWs for the Gaia-ESO M dwarfs must be erroneous. What's happening there?}
@@
 the void than the Pleiades).  The late M dwarfs with detected rotation
-periods tend to rotate very rapidly ($<3$\,days), although these are
-beyond the magnitude limit of our analysis.  This section briefly
-discusses the origins of these features in the rotation-color diagram.
+periods tend to rotate rapidly ($<3$\,days), though there may also be
+late M dwarfs with slow ($>20$\,days) rotation periods.  Although
+these late M dwarfs are beyond the magnitude limit of our analysis,
+they have been explored in depth for \cn\ by both
+\citet{Irwin_NGC2516_2007} and \citet{fritzewski_rotation_2020}.
+While a detailed comparison of our results with theirs is beyond the
+scope of this work, we emphasize that the main contribution of our
+study beyond these previous analyses is not our sensitivity to new
+ranges of spectral type or rotation period, but rather our sensitivity
+to rotation periods across both the core and halo of \cn, and a
+corresponding expansion in the number of stars for which we can derive
+rotation periods.  In the following section, we therefore focus on
+the origins of the features in the rotation-color diagram, and whether
+our expanded rotation sample can provide any new insights.
 
@@
 population \citep{Irwin_NGC2516_2007}, but the question then shifts to
-why these decoupling timescales should be differ for stars that are
+why these decoupling timescales should differ for stars that are
 otherwise identical.  Another possible explanation invokes a

### 2021-04-20 drivers/cluster_rotation/plot_autorotation.py

@@ -1,10 +1,11 @@
 import os
 import earhart.plotting as ep
 from earhart.paths import RESULTSDIR
+from earhart.priors import AVG_EBpmRp
 
 RUNID_EXTINCTION_DICT = {
 #    'compstar_NGC_2516': 0.1343, # 0.25 for KC19
-    'NGC_2516': 0.1343, # 0.25 for KC19
+    'NGC_2516': AVG_EBpmRp, # 0.25 for KC19
 #     'IC_2602': 0.0799,  # avg E(B-V) from Randich+18, *1.31 per Stassun+2019
 #     'CrA': 0.06389, # KC19 ratio used
 #     'kc19group_113': 0.1386, # 0.258 from KC19 -- take ratio
@@ -23,9 +24,28 @@ for runid, _ in RUNID_EXTINCTION_DICT.items():
         os.mkdir(PLOTDIR)
 
     ep.plot_auto_rotation(
+        PLOTDIR, runid, E_BpmRp, core_halo=0,
+        yscale='linear', cleaning='periodogram_match', emph_binaries=1,
+        talk_aspect=1, xval_absmag=1
+    )
+
+    ep.plot_auto_rotation(
         PLOTDIR, runid, E_BpmRp, core_halo=1, yscale='linear',
         cleaning='periodogram_match', emph_binaries=0, talk_aspect=1
     )
+
+    ep.plot_auto_rotation(
+        PLOTDIR, runid, E_BpmRp, core_halo=0, yscale='linear',
+        cleaning='periodogram_match', emph_binaries=0, talk_aspect=1
+    )
+
+    ep.plot_auto_rotation(
+        PLOTDIR, runid, E_BpmRp, core_halo=0,
+        yscale='linear', cleaning='periodogram_match', emph_binaries=1,
+        talk_aspect=1
+    )
+
+    assert 0
 
     for yscale in ['linear', 'log']:
         for cleaning in ['defaultcleaning', 'periodogram_match',

### 2021-04-20 drivers/plot_lightcurves_rotators.py

@@ -1,0 +1,9 @@
+import os
+import earhart.plotting as ep
+from earhart.paths import RESULTSDIR
+
+PLOTDIR = os.path.join(RESULTSDIR, 'lightcurves_rotators')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+ep.plot_lightcurves_rotators(PLOTDIR)

### 2021-04-20 drivers/plot_lumfunction_vs_position.py

@@ -1,0 +1,9 @@
+import os
+import earhart.plotting as ep
+from earhart.paths import RESULTSDIR
+
+PLOTDIR = os.path.join(RESULTSDIR, 'lumfunction_vs_position')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+ep.plot_lumfunction_vs_position(PLOTDIR)

### 2021-04-20 earhart/extinction.py

@@ -18,6 +18,16 @@ def Bonifacio2000_EBmV_correction(EBmV):
     EBmV_adopted = EBmV
     return EBmV_adopted
 
+
+def get_dist_corr(distance_pc, b, h=125):
+    # h = 125pc, scale height of MW
+    return (
+            1 - np.exp(
+                - np.abs( distance_pc * np.sin(np.deg2rad(b)) ) / h
+            )
+    )
+
+
 def given_S98_EBmV_correct(EBmV, distance_pc=None, b=None):
     """
     EBmV: float or array of E(B-V) color excess values.
@@ -30,10 +40,5 @@ def given_S98_EBmV_correct(EBmV, distance_pc=None, b=None):
 
     elif isinstance(distance_pc, float) and isinstance(b, float):
         h = 125 # pc; scale height of the milky way
-        dist_corr = (
-            1 - np.exp(
-                - np.abs( distance_pc * np.sin(np.deg2rad(b)) ) / h
-            )
-        )
-
+        dist_corr = get_dist_corr(distance_pc, b)
         return dist_corr * Bonifacio2000_EBmV_correction(EBmV)

### 2021-04-20 earhart/plotting.py

@@ -15,12 +15,14 @@ Contents:
         plot_physical_X_rotation (+ histogram_physical_X_rotation)
         plot_slowfast_photbinarity_comparison
         plot_rotation_X_positions
+        plot_lightcurves_rotators
     Lithium:
         plot_lithium_EW_vs_color
         plot_rotation_X_lithium
         plot_galah_dr3_lithium_abundance
         plot_randich_lithium (deprecated)
     Other:
+        plot_lumfunction_vs_position
         plot_gaia_rv_scatter_vs_brightness
         plot_ruwe_vs_apparentmag
         plot_edr3_blending_vs_apparentmag
@@ -1099,7 +1101,8 @@ def plot_skypositions_x_rotn(outdir):
 
 
 def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
-                       cleaning=None, emph_binaries=False, talk_aspect=0):
+                       cleaning=None, emph_binaries=False,
+                       talk_aspect=0, xval_absmag=0):
     """
     Plot rotation periods that satisfy the automated selection criteria
     (specified in helpers.get_autorotation_dataframe)
@@ -1155,7 +1158,8 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
             print(42*'-')
 
         if f'{runid}' not in _cls:
-            xval = df['(BP-RP)0']
+            key = '(BP-RP)0' if not xval_absmag else 'MGmag'
+            xval = df[key]
 
             # NOTE: deprecated; based on the webplotdigitzing approach
             # xval = get_interp_BpmRp_from_Teff(df['teff'])
@@ -1165,9 +1169,15 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
             #     index=False
             # )
         else:
-            xval = (
-                df['phot_bp_mean_mag'] - df['phot_rp_mean_mag'] - E_BpmRp
-            )
+            if not xval_absmag:
+                xval = (
+                    df['phot_bp_mean_mag'] - df['phot_rp_mean_mag'] - E_BpmRp
+                )
+            else:
+                get_xval = lambda _df: np.array(
[…]

### 2021-04-20 cluster_paper/collect.sh
add light curve plotter
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@ -35,6 +35,8 @@ cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear.pdf f
 
 cp ../results/venn/venn.pdf f10.pdf
 
+cp ../results/lightcurves_rotators/lightcurves_rotators_color.pdf f14.pdf
+
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf
 # 

### 2021-04-20 drivers/plot_lightcurves_rotators.py

@@ -6,4 +6,5 @@ PLOTDIR = os.path.join(RESULTSDIR, 'lightcurves_rotators')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+ep.plot_lightcurves_rotators(PLOTDIR, color=1)
 ep.plot_lightcurves_rotators(PLOTDIR)

### 2021-04-20 cluster_paper/ms.tex
add light curve plotter
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>
update ms
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@
 corrected for the mean cluster distance by a factor
-$\exp(-\abs{d\sin(b)}/h)$ where $d$ is the average cluster distance,
+$\exp(-|d\sin(b)|/h)$ where $d$ is the average cluster distance,
 $b$ is the average galactic latitude, and $h$ is the scale height of
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\begin{center}
@@
 
-In ``detrending'' before searching for stellar rotation signals in the
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=\textwidth]{f14.pdf}
+	\end{center}
+	\vspace{-1cm}
+  \caption{ {\bf \cn\ light curves sorted in order
+  of stellar color.} Fifty stars are randomly selected from Set
+  $\mathcal{B}$ (see Section~\ref{subsubsec:cluster}), and the light
+  curves are detrended and cleaned as discussed in
+  Appendix~\ref{app:detrending}.  Each light curve is separated by
+  7.5\% in relative flux.  \label{fig:lightcurves}
+	}
+\end{figure}
+
+In detrending before searching for stellar rotation signals in the
 TESS, our goal was to preserve astrophysical variability, while
@@
 
-We opted to use a variant of the principal component analysis approach 
-described by
-\citet{bouma_cdipsI_2019}. This
-principal component analysis focuses on a set of ``trend stars'' selected from
-across each CCD according using ad-hoc heuristics that on average lead
-the light curves to be dominated by systematic variability.  The
-resulting principal component vectors, also referred to as the
-eigenvectors, are rank-ordered by the degree of variance that they
-predict in the training set of trend stars.
+We opted to use a variant of the principal component analysis (PCA)
+approach discussed by \citet{bouma_cdipsI_2019}. This PCA approach
+uses a set of ``trend stars'' selected from across each CCD according
+using ad-hoc heuristics that on average lead the trend star light
+curves to be dominated by systematic variability.  The resulting
+principal component vectors, also referred to as the eigenvectors, are
+rank-ordered by the degree of variance that they predict in the
+training set of trend stars.
 
@@
 inclusion of additional systematic vectors that could affect the light
-curve, discussed below.  These can be treated as additional ``features'' in the linear
-model.  To determine the coefficients of the linear model after the
-full set of eigenvectors (plus optional ``sytematic'' vectors) had
-been asssembled,  we explored two possible methods: ordinary least
-squares, and ridge regression. Ridge regression is the same as
-ordinary least squares, except it includes an $\ell^2$ norm with a
-regularization coefficient. The regularization coefficient that best
-applied for any given target light curve was determined using a
-cross-validation grid search, through
-\texttt{sklearn.linear\_modelRidgeCV} \citep{scikit-learn}.
-Each target light curve was mean-subtracted and normalized by its
-standard deviation, as were the eigenvectors. The linear problem was
-then solved numerically, and the light curve was reconstructed by
-re-adding the original mean, and re-multiplying by the standard
-deviation to ensure that the variance of the light curve did not
-change.
+curve, discussed below.  These can be treated as additional
+``features'' in the linear model.  To determine the coefficients of
+the linear model after the full set of eigenvectors (plus optional
+``sytematic'' vectors) had been asssembled,  we explored two possible
+methods: ordinary least squares, and ridge regression. Ridge
+regression is the same as ordinary least squares, except it includes
+an $\ell^2$ norm with a regularization coefficient. The regularization
+coefficient that best applied for any given target light curve was
+determined using a cross-validation grid search, through
+\texttt{sklearn.linear\_modelRidgeCV} \citep{scikit-learn}.  Each
+target light curve was mean-subtracted and normalized by its standard
+deviation, as were the eigenvectors. The linear problem was then
+solved numerically, and the light curve was reconstructed by re-adding
+the original mean, and re-multiplying by the standard deviation to
+ensure that the variance of the light curve did not change.
 
@@
 background-subtracted) light curve using a linear model with ordinary
-least squares.  The vectors considered in the model were the 1
+least squares.  The vectors used in the model were the 1
 unsmoothed background flux vector, and 5 smoothed PCA eigenvectors.
+Figure~\ref{fig:lightcurves} shows fifty light curves
+randomly drawn from the resulting ``Set $\mathcal{B}$''.
 

### 2021-04-21 cluster_paper/ms.tex
update ms
Co-authored-by: Jason Curtis <jasonleecurtis@gmail.com>

@@
 \newcommand{\ms}{\,m\,s$^{-1}$}
-\newcommand{\bpmrp}{(\mathrm{Bp}-\mathrm{Rp})_0}
+%\newcommand{\bpmrpo}{(\bpmrp)_0}
+\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
+\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}
 
@@
   open clusters (1160),
-  stellar dynamics (1596),
+  stellar kinematics (1608),
   stellar rotation (1629)
@@
 follows.  On one end are coeval groups with no discernable cores ({\it
-e.g.}, Psc Eri and $\mu$ Tau,
+e.g.}, Psc-Eri and $\mu$ Tau,
 \citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
@@
 the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
-Berenices \citep{tang_comaber_2019}.
+Berenices \citep{tang_comaber_2019,furnkranz_2019}.
 
@@
 are available on
-MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.
+MAST.\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}
 
@@
 cluster, and ask: is the cluster halo truly coeval with the core?  We
-chose \cn\ for this analysis because it is young (100--200\,Myr) and
+chose \cn\ for this analysis because it is young (100-200\,Myr) and
 sufficiently nearby ($d\approx400$\,pc) to facilitate photometric
@@
   Appendix~\ref{app:clustering}.  The field sample is randomly drawn
-  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
+  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
   low volume density of the halo stars makes it difficult to visually
@@
 respectively.  A useful visualization of these samples is
-available online\footnote{
+available online.\footnote{
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}.
+  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}
 %While we considered performing our own clustering analysis on the
@@
   \caption{ {\bf HR diagrams of NGC\,2516.}
-    {\it Left}: Photometric data from Gaia EDR3; no reddening
-    correction has been applied.  The core (black) shows a
+    {\it Left}: Gaia EDR3 photometry for the core (black) shows a
     main-sequence and turnoff consistent with stars of a fixed age and
@@
     8.5\}$.  The arrow represents the average reddening correction
-    applied to the models.  The opacity of the M-dwarf model
-    atmospheres is less well-calibrated for $M_\star \lesssim
-    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2);
+    applied to the models (and to give the approximate spectral
+    types).  The opacity of the M-dwarf model atmospheres is less
+    well-calibrated for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V);
     these model values are shown with dashed lines.
@@
 Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
-$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$
-color.  Stars in the core appear consistent with having a fixed age
-and metallicity, and varying mass.  The halo stars show a similar
-sequence, though with greater scatter.  One possible explanation for
-the additional scatter is that the halo is more contaminated by
-field stars.  For instance, $\approx5$ red giants in the halo must be
-field interlopers, because their isochronally implied ages would be
-inconsistent with that of the cluster.
+$G$ magnitude as a function of observed $\bpmrp$ color.  Spectral
+types are interpolated from the \citet{pecaut_mamajek_2013} table,
+after accounting for reddening as described
+below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
+version 2021.03.02.} Stars in the core appear consistent with having a
+fixed age and metallicity, and varying mass.  The halo stars show a
+similar sequence, though with greater scatter.  One possible
+explanation for the additional scatter is that the halo is more
+contaminated by field stars.  For instance, $\approx5$ red giants in
+the halo must be field interlopers, because their isochronally implied
+ages would be inconsistent with that of the cluster.
 
@@
 We used the web
-interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
-\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
+interface,\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
+\texttt{2021-02-26} \texttt{CMD3.4}, \texttt{YBC} bolometric
 corrections as in \citet{chen_2019_YBC}.} to interpolate these
@@
 colors we used the calibration of \citet{stassun_TIC8_2019}, namely
-$E(\mathrm{Bp}-\mathrm{Rp})=1.31 E(\mathrm{B}-\mathrm{V})$ and
+$E(\bpmrp)=1.31 E(\mathrm{B}-\mathrm{V})$ and
 $A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded
-$E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$, which is used in the
+$E(\bpmrp)=0.134\pm0.025$, which is used in the
 plots.  To redden the isochrones, we assumed $R_V=3.1$, and applied
@@
 metallicities to fit as much of the locus from
-$0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as possible, and settled on the
+$0.5<\bpmrp<1.5$ as possible, and settled on the
 slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
@@
 $\approx$0.45\,$M_\odot$.  Below this mass, the data and models
-diverge at colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in
+diverge at colors redder than $\bpmrp\approx2.2$, in
 the sense that the model isochrones have lower luminosities and bluer
@@
 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\subfloat{
+% 			\includegraphics[width=0.49\textwidth]{f3a.pdf}
+% 			\includegraphics[width=0.49\textwidth]{f3b.pdf}
+% 		}
+% 
+%         \vspace{-0.5cm}
+% 		\subfloat{
+% 			\includegraphics[width=0.49\textwidth]{f3c.pdf}
+% 			\includegraphics[width=0.49\textwidth]{f3d.pdf}
+% 		}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+%   \caption{ {\bf TESS rotations periods versus color for \cn.} {\it
+%   Top row}: The core and halo of \cn.  Sets $\mathcal{A}$ ({\it left})
+%   and $\mathcal{B}$ ({\it right}) undergo successive stages of
+%   cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
+%   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
+%   \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
+%   Stars in the core and the halo of NGC\,2516 overlap with the
+%   Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
+%   members have longer rotation periods than Pleiades members.  {\it
+%   Bottom row}: Binarity indicators in \cn.  Astrometric binaries
+%   ($\mathrm{RUWE}>1.2$; 11\% of stars) appear in red; photometric
+%   binaries ($>$0.3\,mag from an empirical isochrone; 20\% of stars)
+%   are overplotted in orange.  Rotation periods for comparison field
+%   stars are shown in Appendix~\ref{app:compstar}.
+%   \label{fig:rot}
+% 	}
+% \end{figure*}
+
 \begin{figure*}[t]
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f3a.pdf}
-			\includegraphics[width=0.49\textwidth]{f3b.pdf}
+			\includegraphics[width=0.7\textwidth]{f3a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f3c.pdf}
-			\includegraphics[width=0.49\textwidth]{f3d.pdf}
+			\includegraphics[width=0.7\textwidth]{f3b.pdf}
 		}
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for
-  \cn.} {\it Top row}: \cn\ split into its core and halo.  Sets
-  $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo
-  successive stages of automated cleaning, (see
-  Section~\ref{subsubsec:cluster}).  The Pleiades
+  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
+  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
+  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
+  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
@@
   Stars in the core and the halo of NGC\,2516 overlap with the
-  Pleiades at $\bpmrp$$<$1.2.  From $1.2<\bpmrp<1.7$, some NGC\,2516
-  members have longer rotation periods than Pleaides members.  {\it
-  Bottom row}: \cn\ colored by binarity indicators.  Sources with
-  $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\% of
-  stars).  Overplotted in orange are sources at least 0.3\,mag above
-  an empirically fit cluster isochrone (photometric binaries; 20\% of
-  stars).
-  Rotation periods for comparison field stars are shown in
+  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
+  members have longer rotation periods than Pleiades members. Rotation
+  periods for comparison field stars are shown in
   Appendix~\ref{app:compstar}.
   \label{fig:rot}
+	}
+\end{figure*}
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.7\textwidth]{f3c.pdf}
+		}
+
+    \vspace{-0.5cm}
+		\subfloat{
+			\includegraphics[width=0.7\textwidth]{f3d.pdf}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Binarity indicators for the rotators in
+  Figure~\ref{fig:rot}.} Sets $\mathcal{A}$ ({\it top}) and
+  $\mathcal{B}$ ({\it bottom}) undergo the same successive stages of
+  cleaning.  Astrometric binaries ($\mathrm{RUWE}>1.2$; 11\% of stars)
+  appear in red; photometric binaries ($>$0.3\,mag from an empirical
+  isochrone; 20\% of stars) are overplotted in orange.  
+  \label{fig:binarity}
 	}
@@
 distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
-processing corresponds roughly to $\bpmrp$ of 2.2, or a spectral type
+processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
 of $\approx$M2V.
@@
 
-The top row of Figure~\ref{fig:rot} shows the resulting rotation
-periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
-the right.  While set $\mathcal{A}$ is more complete than set
+Figure~\ref{fig:rot} shows the resulting rotation
+periods, with set $\mathcal{A}$ shown on top and $\mathcal{B}$ on
+the bottom.  While set $\mathcal{A}$ is more complete than set
 $\mathcal{B}$, this completeness comes at the expense of greater
@@
 \cn\ appear gyrochronologically coeval for colors spanning
-$0.5<\bpmrp<1.2$.  The implications of this comparison for the age of
+$0.5<\bpmrpo<1.2$.  The implications of this comparison for the age of
 \cn\ are discussed in Section~\ref{disc:absage}.
 
-The bottom row of Figure~\ref{fig:rot} shows the same data, with the
+Figure~\ref{fig:binarity} shows the same data, with the
 points colored with indicators of binarity.  To assess astrometric
@@
 rotation sequence.  In set $\mathcal{B}$, over the color range where
-the slow and fast sequence are present ($0.5<\bpmrp<1.3$), there are
+the slow and fast sequence are present ($0.5<\bpmrpo<1.3$), there are
 281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
@@
   left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
-  2D-tangential velocities for halo members that meet kinematic
-  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
-  rotation periods are shown in blue;  halo members for which rotation
-  periods should have been detectable, but were not detected, are
-  shown in orange.  Stars in the core are shown as black points.
-  Computing the 2D-tangential velocity requires correcting for
-  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
-  right}, the histograms on bin versus 3D separation from the core and
-  2D tangential velocity difference.  Bin widths are chosen to enforce
-  the same number of stars in the denominator ($N=54$) per bin.  In the
-  {\it bottom panels}, the innermost and outermost 100 stars in
-  position and velocity space (black and green) are shown in the
-  space of rotation periods and dereddened colors.
+  2D-tangential velocities for kinematic halo members
+  (Set~$\mathcal{A}$).  Stars with detected rotation periods are shown
+  in blue;  halo members for which rotation periods should have been
+  detectable, but were not detected, are shown in orange.  Stars in
+  the core are shown as black points.  Computing the 2D-tangential
+  velocity requires correcting for projection effects
+  (Appendix~\ref{app:vproj}).  In the {\it top right}, the histograms
+  bin versus 3D separation from the core and 2D tangential velocity
+  difference.  Bin widths are chosen to enforce the same number of
+  stars in the denominator ($N=54$) per bin.  In the {\it bottom
+  panels}, the innermost and outermost 100 stars in position and
+  velocity space (black and green) are shown in the space of rotation
+  periods and dereddened colors.
   \label{fig:physical_x_rotn}
@@
 Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
-comparison, we only show stars with $0.5<\bpmrp<1.2$ for which our
-TESS pipeline succeeding in making light curves.  In other words, the
+comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our
+TESS pipeline succeeded in making light curves.  In other words, the
 stars in the base sample are those for which we could have plausibly
@@
 Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
-The total length of the structure is therefore 400--500\,pc, depending
+The total length of the structure is therefore 400-500\,pc, depending
 on which members of the halo are chosen as the ``tips'' on either end.
@@
   EW}\approx0\,$m\AA\ are non-detections.
-  The Li measurements for both Core and Halo stars in \cn\ are
+  The Li measurements for both core and halo stars in \cn\ are
   consistent with a near-Pleiades age.
@@
 Gaia-ESO spectrum was acquired and stellar parameters were reported.
-15 of these (all with $\bpmrp > 2.0$) are spurious matches based on
+15 of these (all with $\bpmrpo > 2.0$) are spurious matches based on
 the Gaia color and GES effective temperature, and we remove them.
@@
 the galactic plane.  The survey is spatially
-inhomogeneous\footnote{See the footprint at
-\url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
+inhomogeneous.\footnote{See the footprint at
+\url{https://www.galah-survey.org/news/announcing-galah-dr3}} Special
 targeting for stars in the TESS southern continuous viewing zone, and
@@
 the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
-by first focusing in on a 15\AA\ window centered on the 6707.835\AA
+by first focusing in on a 15\AA\ window centered on the 6707.835\AA\ 
 lithium doublet.  We then continuum normalized the spectra using a
@@
 stars are in the core or the halo of the cluster.  The GALAH spectra
-span a color range of $0<\bpmrp<1.5$, due to the $V<14$ brightness
+span a color range of $0<\bpmrpo<1.5$, due to the $V<14$ brightness
 cutoff of the survey.  The overall increase of Li EW from
-$0<\bpmrp<1.0$ is driven by the temperatures in the stellar
+$0<\bpmrpo<1.0$ is driven by the temperatures in the stellar
 chromospheres: hotter stars fully ionize Li out of its ground state
 \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
-depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.}, later
+depletion of stars redder than $\bpmrpo\gtrsim1.4$, {\it i.e.}, later
 than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier K
@@
 \citet{pecaut_mamajek_2013}
-table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version 2021.03.02.}.  For clarity, we have omitted the few upper
+table).  For clarity, we have omitted the few upper
 limits reported by \citet{bouvier_pleiades_lirot_2018}.  The field
@@
 
-Any theory for the structure of \cn, the Psc\,Eri stream, and the
+Any theory for the structure of \cn, the Psc-Eri stream, and the
 strings and halos found by \citetalias{kounkel_untangling_2019} and
@@
 have formed in the same ``clump'' as those in the cluster core.  In
-other words, the formation environment might be better described a
+other words, the formation environment might be better described as a
 giant molecular filament, rather than a giant molecular cloud.  A
@@
 \citet{fritzewski_rotation_2020} have recently demonstrated that
-between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc\,Eri,
-M\,35, M\,50, and the Pleaides) all show similar rotation period
+between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri,
+M\,35, M\,50, and the Pleiades) all show similar rotation period
 distributions.  The FGK stars show a ``slow sequence'', a population
 of ``rapid rotators'', and in some cases stars in the ``void'' between
-the two populations ({\it e.g.,}, Psc\,Eri and \cn\ show more stars in
+the two populations ({\it e.g.,}, Psc-Eri and \cn\ show more stars in
 the void than the Pleiades).  The late M dwarfs with detected rotation
@@
 scope of this work, we emphasize that the main contribution of our
-study beyond these previous analyses is not our sensitivity to new
-ranges of spectral type or rotation period, but rather our sensitivity
-to rotation periods across both the core and halo of \cn, and a
-corresponding expansion in the number of stars for which we can derive
-rotation periods.  In the following section, we therefore focus on
-the origins of the features in the rotation-color diagram, and whether
-our expanded rotation sample can provide any new insights.
+study beyond these previous analyses our sensitivity
+across both the core and halo of \cn, and a
+corresponding expansion in the number of stars that can be analyzed.
+In the following section, we therefore focus on the origins of the
+features in the rotation-color diagram, and whether our expanded
+rotation sample can provide any new insights.
 
 The existence of a slow sequence is the main basis of gyrochronology
-\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
+\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the 
 spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
@@
 been definitively answered.  Models that introduce variable
-core-envelope decoupling timescales can reproduce each type of
-population \citep{Irwin_NGC2516_2007}, but the question then shifts to
-why these decoupling timescales should differ for stars that are
-otherwise identical.  Another possible explanation invokes a
-spontaneous and random change of magnetic field topology that causes
-the dynamo to couple more strongly with the stellar wind
-\citep{brown_metastable_2014}.  Other processes could be external to the
-star entirely: \citet{qureshi_signature_2018} for instance have
-reported that mergers between giant planets and the star could explain
-a non-neglible fraction of the rapid rotators.
+core-envelope decoupling timescales and pre-MS disk lifetimes can
+reproduce each type of population
+\citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015},
+but the question then shifts to why these decoupling timescales should
+differ for stars that are otherwise identical.  Another possible
+explanation invokes a spontaneous and random change of magnetic field
+topology that causes the dynamo to couple more strongly with the
+stellar wind \citep{brown_metastable_2014}.  Other processes could be
+external to the star entirely: \citet{qureshi_signature_2018} for
+instance have reported that mergers between giant planets and the star
+could explain a non-neglible fraction of the rapid rotators.
 
 A separate hypothesis that we can test using the existing data is the
-idea that rapid rotation can be explained through binarity.  Analyses
-by {\it e.g.}, \citet{gillen_ngts_2020} and
-\citet{simonian_rapid_2019} have reported that rapid rotation in young
-clusters and in the field is correlated with binarity.  Related
-effects could include tidal synchronization for the shortest period
-binaries, or alternatively disk truncation that when combined with
-magnetic locking \citep[{\it
-e.g.},][]{koenigl_disk_1991,long_locking_2005} could produce the
-rotation period bimodality.
+idea that rapid rotation can be explained through binarity.  
+Correlations between rapid rotation and binarity in both young
+clusters and the field have preivously been noted
+\citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}
+Related effects could include tidal synchronization for the shortest
+period binaries, or alternatively pre-MS magnetic locking between the
+disk and star \citep[{\it
+e.g.},][]{koenigl_disk_1991,long_locking_2005}.
 
@@
 clusters
-\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
+\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020,hawkins_2020}.
 In \cn, \citet{jeffries_rotation_1998} previously observed a similar
@@
 top row of Figure~\ref{fig:rot} shows that the two sequences overlap
-from $0.5<\bpmrp<1.2$.  At redder colors, $1.2<\bpmrp<1.7$, the
+from $0.5<\bpmrpo<1.2$.  At redder colors, $1.2<\bpmrpo<1.7$, the
 dispersion in rotation periods increases.  The maximum rotation
@@
 is, unfortunately, an exercise in tautology.  Gyrochronological models
-are empirically calibrated against the Pleiades and Praesepe.  In
-other words, at $\bpmrp=1.35$, we observe rotation periods in \cn\
-ranging from 7 to 10 days.  The 7 day rotation periods are consistent
-with the Pleiades; the 10 day rotation periods are not.  Fitting
-formulae from the previously cited studies give ages for a star with
-$\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of
-204 Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
-\citet{spada_competing_2020} model has an absolute scale that is
-well-calibrated at these ages and colors.  For an 8 day rotation
+are empirically calibrated against the Pleiades and Praesepe.  At
+$\bpmrpo=1.35$, we observe rotation periods in \cn\ ranging from 7 to
+10 days.  The 7 day rotation periods are consistent with the Pleiades;
+the 10 day rotation periods are not.  Fitting formulae from the
+previously cited studies give ages for a star with $\{\bpmrpo=1.35,
+(\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204 Myr, 316 Myr,
+and 107 Myr respectively.  Given a Pleiades age of $125\pm20$\,Myr (an
+average of MSTO and rotation-corrected lithium-depletion results; see
+{\it e.g.} \citealt{stauffer_keck_1998},
+\citealt{soderblom_ages_2014}, and \citealt{cummings_2018}), only the
+\citet{spada_competing_2020} model has an absolute scale that seems to
+be well-calibrated at these ages and colors.  For an 8 day rotation
 period of a star with the same color, this model quotes a 193 Myr age;
@@
 the relative statement that \cn\ appears slightly gyrochronologically
-older than the Pleaides, and much younger than Praesepe.  Given a
-Pleiades age of $125\pm20$\,Myr (an average of MSTO and
-rotation-corrected lithium-depletion results; see {\it e.g.}
-\citealt{stauffer_keck_1998}, \citealt{soderblom_ages_2014}, and
-\citealt{cummings_2018}), a gyrochronological age estimate for \cn\
-between 150 and 200\,Myr would therefore appear reasonable.  Much
-older, and the issue of why the F and G dwarfs do not spin down
-becomes pressing.
+older than the Pleiades, and much younger than Praesepe.  A
+gyrochronological age estimate for \cn\ between 150 and 200\,Myr would
+therefore appear reasonable.  Much older, and the issue of why the F
+and G dwarfs do not spin down becomes pressing.  We caution however
+that the spindown rate of FGK stars between 100 and 300 Myr has not
+yet been empirically calibrated, and if ``stalling'' were to occur, it
+could affect our age assessment \citep[see][]{curtis_rup147_2020}.
 
@@
 $\approx$0.088\,$M_\odot$ \citep{soderblom_ages_2014}.  This would
-correspond to $\bpmrp\approx4.7$, well beyond the limits of available
+correspond to $\bpmrpo\approx4.7$, well beyond the limits of available
 cluster membership lists and spectroscopy.
@@
 \paragraph{Adopted Age}
-As noted in Section~\ref{subsec:hr}, a color-color analysis of the
+As noted in Section~\ref{subsec:hr}, a color-color and color-magnitude analysis of the
 main sequence turn-off (MSTO) by \citet{cummings_2018} found that \cn\
@@
   %
-  \item {\it Over-densities observed in the Gaia 3-D positions and 2-D
+  \item {\it Over-densities observed in the Gaia 3D positions and 2D
     velocities revealed a halo of stars spanning $\approx$500 parsecs
@@
     comparison field stars of the same color.  Gyrochronally, stars in
-    \cn\ spanning $0.5<\bpmrp<1.2$ (spectral types F2V-K3V) overlap
-    with the slow sequence of the Pleaides (Figure~\ref{fig:rot}).  At
-    redder colors, from $1.2<\bpmrp<1.7$ (spectral types K3V-K6V,
+    \cn\ spanning $0.5<\bpmrpo<1.2$ (spectral types F2V-K3V) overlap
+    with the slow sequence of the Pleiades (Figure~\ref{fig:rot}).  At
+    redder colors, from $1.2<\bpmrpo<1.7$ (spectral types K3V-K6V,
     $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
@@
     quantify the statement: $\approx$75\% of the ``inner-most''
-    members have detected rotation periods.  This fraction decreases
-    to $\approx$60\% at physical separations of $\approx$200\,pc.
-    Based on the rate at which rotation periods are detected at or
-    underneath the gyrochronological slow sequence, field stars appear
-    to contaminate the outermost candidate cluster members at a rate
-    between 40\% and 50\%.  Given the number density ratios of cluster
+    members have detected rotation periods consistent with a
+    gyrochronological age near that of the Pleiades.  The false
+    positive rate increases to $\approx$50\% at physical separations
+    of $\approx$200\,pc.  Given the number density ratios of cluster
     to field stars at these wide separations ($\approx 10^{-3}$,
@@
      comparable stellar populations ({\it e.g.,}
-     \citealt{soderblom_evolution_1993}; \citealt{jeffries_m35_li_2020}; \citealt{gillen_ngts_2020}).
-     The rough orders of magnitude of the effects are $\approx$1\,dex
-     in lithium abundance for a factor of 10 in rotation period, and a
+     \citealt{soderblom_evolution_1993};
+     \citealt{jeffries_m35_li_2020}; \citealt{gillen_ngts_2020}).  The
+     rough orders of magnitude of the effects are $\approx$1\,dex in
+     lithium abundance for a factor of 10 in rotation period, and a
      factor of $\gtrsim2$ enhancement in binarity fraction for fast
@@
      connection between the two observations is that the disk-locking
-     phase may be truncated in binary systems, which could lead to
-     less efficient internal mixing and the longer-lasting presence of
-     photospheric lithium \citep{eggenberger_impact_2012}.
+     phase may be truncated in binary systems, which could free the
+     stars to spin up on the pre-MS, producing less efficient internal
+     mixing and a longer-lasting presence of photospheric lithium
+     \citep{eggenberger_impact_2012}.
    %
@@
 
-The authors thank L.~Hillenbrand, B.~Tofflemire, G.~Zhou, E.~Newton,
-K.~Hawkins, A.~Mann, and A.~Kraus for the discussions on young stars,
-rotation, and lithium that encouraged this line of study.
-%
-L.G.B. and J.H. acknowledge support by the TESS GI Program, program
-NUMBER, through NASA grant NUMBER.
-L.G.B. was also supported by a XXXX Fellowship from Princeton
-University.
+The authors thank, in reverse-alphabetical order, G.~Zhou,
+B.~Tofflemire, A.~McWilliam, E.~Newton, L.~Hillenbrand, A.~Kraus, and
+K.~Hawkins for the discussions on young stars, rotation, and lithium
+that encouraged this analysis.
+%
+L.G.B. and J.H. acknowledge support by the TESS GI Program, {\bf program
+NUMBER, through NASA grant NUMBER}.
+L.G.B. was also supported by a Charlotte Elizabeth Procter Fellowship
+from Princeton University.
 %
@@
 Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
-2020A-0146; 2020B-NUMBER PI: L{.}~Bouma), which is managed by the
+2020A-0146; 2020B-0029 PI: L{.}~Bouma), which is managed by the
 Association of Universities for Research in Astronomy (AURA) under a
@@
 %
-ACKNOWLEDGE PFS / CAMPANAS.
+% ACKNOWLEDGE PFS / CAMPANAS.
 %
@@
 %
-We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
-Latham, Sara Seager, Jon Jenkins) and the many TESS team
-members for their efforts to make the mission a continued success.
+We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
+S.~Seager, J.~Jenkins) and the many TESS team members for their
+efforts to make the mission a continued success.
 %
@@
 \citetalias{kounkel_untangling_2019} applied a different unsupervised
-clustering method to the 5-D Gaia DR2 positions and proper motions.
+clustering method to the 5D Gaia DR2 positions and proper motions.
 \citetalias{kounkel_untangling_2019} began with a sample of
@@
   Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
-  considered. (Fainter stars in the field sample were not studied as
-  part of the broader CDIPS project).  The field stars show a
+  considered. (Fainter field stars were not reduced during
+  the initial CDIPS reductions).  The field stars (orange) show a
   different rotation period distribution than the kinematically
-  selected NGC\,2516 members.
+  selected \cn\ members.
+  Both field and \cn\ stars are subject to the same ``Set $\mathcal{A}$''
+  cleaning procedures.
   {\bf LGB TODO: Rerun the field measurements with a better grid to fix the aliasing issue.}
@@
 The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
-corresponds to an extinction-corrected color cutoff of $\bpmrp
+corresponds to an extinction-corrected color cutoff of $\bpmrpo
 \approx 0.80$, or spectral types of $\approx$G1V.
@@
 %   velocity space, cross-matched against the rotators.} The plotted
-%   stars are those with $0.5<\bpmrp<1.2$ that meet the crowding
+%   stars are those with $0.5<\bpmrpo<1.2$ that meet the crowding
 %   restrictions described in Section~\ref{subsec:tess}: they should

### 2021-04-21 earhart/plotting.py

@@ -63,7 +63,9 @@ from cdips.utils.gaiaqueries import (
 )
 from cdips.utils.tapqueries import given_source_ids_get_tic8_data
 from cdips.utils.plotutils import rainbow_text
-from cdips.utils.mamajek import get_interp_BpmRp_from_Teff
+from cdips.utils.mamajek import (
+    get_interp_BpmRp_from_Teff, get_SpType_BpmRp_correspondence
+)
 
 from earhart.paths import DATADIR, RESULTSDIR
 from earhart.helpers import (
@@ -867,9 +869,10 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
                 leg.legendHandles[2]._sizes = [1.3*25]
 
 
-    ax.set_ylabel('Absolute $\mathrm{M}_{G}$ [mag]', fontsize='large')
+    ax.set_ylabel('Absolute $\mathrm{M}_{G}$ [mag]', fontsize='medium')
     if color0 == 'phot_bp_mean_mag':
-        ax.set_xlabel('$G_{\mathrm{BP}}-G_{\mathrm{RP}}$ [mag]', fontsize='large')
+        ax.set_xlabel('$G_{\mathrm{BP}}-G_{\mathrm{RP}}$ [mag]',
+                      fontsize='medium')
         c0s = '_Bp_m_Rp'
     elif color0 == 'phot_g_mean_mag':
         ax.set_xlabel('G - Rp [mag]', fontsize='large')
@@ -891,6 +894,36 @@ def plot_hr(outdir, isochrone=None, color0='phot_bp_mean_mag',
         ax.set_ylim([13.7, -4.8])
 
     format_ax(ax)
+    ax.tick_params(axis='x', which='both', top=False)
+
+    #
+    # append SpTypes (ignoring reddening)
+    #
+    if color0 == 'phot_bp_mean_mag':
+        tax = ax.twiny()
+        tax.set_xlabel('Spectral Type')
+
+        xlim = ax.get_xlim()
+        sptypes, BpmRps = get_SpType_BpmRp_correspondence(
+            ['A0V','F0V','G0V','K2V','K5V','M0V','M2V','M4V']
+        )
+        print(sptypes)
+        print(BpmRps)
+
+        xvals = np.linspace(min(xlim), max(xlim), 100)
+        tax.plot(xvals, np.ones_like(xvals), c='k', lw=0) # hidden, but fixes axis.
+        tax.set_xlim(xlim)
+        ax.set_xlim(xlim)
+
[…]

### 2021-04-22 cluster_paper/ms.tex
update plots; update ms

@@
 \newcommand{\nautorotnumeratormatching}{507} % plot_autorotation.py, LSP=SPDM period
-\newcommand{\ncompstardenominator}{820} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
-\newcommand{\ncompstarnumerator}{365} % plot_autorotation.py
-\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
+\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
+\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
+\newcommand{\ncompfrac}{38.0\%} % plot_compstar_rotation.py, based on plot_autorotation
 \newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
@@
 corresponding powers.  Finally, as a check on crowding, we recorded
-the number of stars within the aperture with equal or greater
+the number of stars within the aperture with greater
 brightness than the target star, and with brightness within 1.25 and
@@
 days, the normalized Lomb-Scargle power exceeded 0.08, and for which
-no equal-brightness or greater companions were within the aperture.
+no companions with brightness exceeding one-tenth of the target star
+were in the aperture ({\it i.e.}, excluding visual binaries
+with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
 These limits were chosen after inspecting the light curves visually,
 to eliminate low-quality rotation periods while retaining high
-completeness.  Beyond requiring that the target star be the brightest
-star in the aperture, we also required that at most one companion with
-brightness exceeding one tenth of the target star's brightness be
-present in the aperture according to the Gaia DR2 source catalog.  For
-set $\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out
-of \nautorotdenominator\ stars for which rotation periods might
-plausibly have been detected.  For set $\mathcal{B}$, we additionally
-required that the Lomb-Scargle and Stellingwerf phase-dispersion
-methods yielded identical periods to within 10\%.  This yielded
-\nautorotnumeratormatching\ light curves.  Boolean flags for each set
-are included in Table~\ref{tab:maintable}.
+completeness.  For set $\mathcal{A}$, this yielded \nautorotnumerator\
+light curves, out of \nautorotdenominator\ stars for which rotation
+periods might plausibly have been detected.  For set $\mathcal{B}$, we
+additionally required that the Lomb-Scargle and Stellingwerf
+phase-dispersion methods yielded identical periods to within 10\%.
+This yielded \nautorotnumeratormatching\ light curves.  Boolean flags
+for each set are included in Table~\ref{tab:maintable}.
+A set of field stars was also analyzed for comparison.
+The results are discussed in Appendix~\ref{app:compstar}.
 
@@
   non-targeted selection function of that survey.  Field stars (gray
-  points) are Keck-HIRES measurements of stars with detected
-  transiting planets from the Kepler mission
+  points) are stars with detected
+  transiting planets from Kepler 
   \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 	\begin{center}
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=\textwidth]{f14.pdf}
+		\includegraphics[width=0.8\textwidth]{f14.pdf}
 	\end{center}
@@
 
-\section{Comparison Star Rotation Periods}
+\section{Rotation Periods for Comparison Field Stars}
 \label{app:compstar}
@@
 		\leavevmode
-    \includegraphics[width=0.49\textwidth]{f9.pdf}
+    \includegraphics[width=0.69\textwidth]{f9.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf Rotation in NGC\,2516 compared to the field.}
-  Stars with Gaia $Rp<13$\,mag in the cluster and field samples were
-  considered. (Fainter field stars were not reduced during
-  the initial CDIPS reductions).  The field stars (orange) show a
-  different rotation period distribution than the kinematically
-  selected \cn\ members.
-  Both field and \cn\ stars are subject to the same ``Set $\mathcal{A}$''
-  cleaning procedures.
-  {\bf LGB TODO: Rerun the field measurements with a better grid to fix the aliasing issue.}
+	\caption{ {\bf Rotation in \cn\ compared to the field.}
+  Stars with Gaia $Rp<13$\,mag in the cluster and field samples are
+  displayed; fainter field stars were not analyzed.  Field
+  stars (orange) show a different rotation period distribution than
+  the kinematically selected \cn\ members (black).  Field and \cn\
+  stars are subject to the same ``Set $\mathcal{A}$''
+  period-measurement and cleaning procedures.
   \label{fig:compstar}
@@
 
-To provide a basis for comparison, we also opted to search the
-``calibration'' light curves ($Rp<13$) that were created as a part of
-the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
-corresponded to a sample of \ncalibration\ stars.  Cross-matching these
-against the \nnbhd\ randomly drawn stars in the neighborhood of \cn yielded
-\nnbhdcalibstar\ unique stars, with a cumulative total of
-\nnbhdcaliblc\ TESS
-sectors observed.
-The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
+To verify the significance of the trends seen in Figures~\ref{fig:rot}
+and~\ref{fig:binarity}, we searched the CDIPS calibration light curves
+for rotation periods.  This light curve database comprises all $Rp<13$
+stars that fell on TESS silicon. The need to distinguish field and
+cluster stars is based on our processing capabilities and is discussed
+by \citet{bouma_cdipsI_2019}.  Over the southern sky (Sectors 1-13 of
+TESS), this corresponded to a sample of \ncalibration\ stars.
+Cross-matching these against the \nnbhd\ randomly drawn stars in the
+neighborhood of \cn\ yielded \nnbhdcalibstar\ unique stars.  The
+magnitude cut of $Rp<13$ at the distances of the neighborhood sample
 corresponds to an extinction-corrected color cutoff of $\bpmrpo
-\approx 0.80$, or spectral types of $\approx$G1V.
-This reaches sufficiently far down the main-sequence to enable a
-comparison to the cluster star sample.
+\lesssim 0.80$, or spectral types of $\lesssim$G1V.  This reaches
+sufficiently far down the slow sequence to enable a comparison against
+the cluster star sample.
 
 We performed the same light curve stitching and period-search
-procedure on the field comparison stars.
+procedure discussed in the text on the field stars.
 Imposing the same requirements for crowding resulted in
-\ncompstardenominator\ stars for which rotation periods could have been
-detected.
-Imposing the same Lomb Scargle power cutoff, and period upper limit,
-yielded \ncompstarnumerator\ period detections (\ncompfrac).
-Within the same brightness cutoff, 
-\nautovscompstarnumerator\ of \nautovscompstardenominator\ cluster stars
-yielded period detections (\nautofrac).
-Though the detection fractions are frankly not very different
-(likely because of the brightness cutoff), the period {\it
-vs} color distributions are quite different
+\ncompstardenominator\ stars for which rotation periods could have
+been detected.  Imposing the same Lomb Scargle power cutoff (${\rm
+LSP}>0.08$) and period upper limit ($P<15\,{\rm days}$) yielded
+\ncompstarnumerator\ period detections (\ncompfrac).  Within the same
+brightness cutoff, \nautovscompstarnumerator\ of
+\nautovscompstardenominator\ cluster stars yielded period detections
+(\nautofrac).  While the detection fractions are only mildly
+different (likely due to the brightness cutoff), the period--color
+distributions are extremely different
 (Figure~\ref{fig:compstar}).
+This confirms our expectation that our period measurement procedure is
+not biased in any way that could produce the features we observe in
+\cn.
 
@@
 			\includegraphics[width=0.49\textwidth]{f12a.pdf}
-		}
-
-		\subfloat{
 			\includegraphics[width=0.49\textwidth]{f12b.pdf}
@@
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.5cm}
   \caption{ {\bf Projection effects for co-moving populations across
-  the sky.} The map is colored by the velocity difference a star
+  the sky.} The maps are colored by the proper motion difference a star
   co-moving with \cn\ would exhibit across the equatorial sphere.
@@
   the star denotes the cluster center.
+  The {\it left} and {\it right} panels show the proper motion correction in
+  the declination and right ascension directions.
 	\label{fig:vproj}
@@
 
-Figure~\ref{fig:vproj} shows why the correct approach for comparing
-any star's proper motions against the cluster's mean depends on the
-position of the star being considered.
+To compare any given star's proper motion against the cluster's mean,
+one must consider the position of the star on the sky.  This is
+because two stars sharing identical cartesian velocities will have
+different velocities when projected on the sky, and was emphasized by
+\citet{meingast_2021}.
+Figure~\ref{fig:vproj} shows the magnitude of the correction in
+equatorial tangential velocities.
 

### 2021-04-22 drivers/cluster_rotation/get_auto_rotation_periods.py

@@ -1,3 +1,4 @@
+# Usage: change `runid` below.
 import os, shutil, socket, pickle
 from glob import glob
 import pandas as pd, numpy as np

### 2021-04-22 cluster_paper/maintable_header.tex
update ms

%% \begin{deluxetable}{} command tell LaTeX how many columns
%% there are and how to align them.
%\startlongtable
\begin{deluxetable*}{lll}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)
\tabletypesize{\scriptsize}
%\tabletypesize{\small}

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\tablecaption{Rotation periods and lithium equivalent widths for
candidate NGC\,2516 members.}
\label{tab:maintable}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
%\tablenum{3}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
\tablehead{
  \colhead{Parameter} &
  \colhead{Example Value} &
  \colhead{Description}
}

%% All data must appear between the \startdata and \enddata commands
%
% paste from
% /Users/luke/Dropbox/proj/earhart/results/tables/NGC_2516_Prot_cleaned_header.tex
% via drivers/write_NGC2516_main_table.py
\startdata
             \texttt{source\_id} & 5290824530952816896 &                                         Gaia DR2 source identifier. \\
       \texttt{n\_cdips\_sector} &                   3 &                     Number of TESS sectors with CDIPS light curves. \\
                 \texttt{period} &            1.297662 &                                           Lomb-Scargle best period. \\
                 \texttt{lspval} &            0.171629 &                     Lomb-Scargle periodogram value for best period. \\
                 \texttt{nequal} &                 0.0 &          Number of stars brighter than the target in TESS aperture. \\
                 \texttt{nclose} &                 1.0 &            Number of stars with $\Delta T > 1.25$ in TESS aperture. \\
                 \texttt{nfaint} &                 1.0 &             Number of stars with $\Delta T > 2.5$ in TESS aperture. \\
             \texttt{spdmperiod} &            1.297662 &                                       Stellingwerf PDM best period. \\
                \texttt{spdmval} &            0.826804 &                 Stellingwerf PDM periodogram value for best period. \\
             \texttt{ref\_epoch} &              2015.5 &                                         Reference epoch for RA/dec. \\
                     \texttt{ra} &          120.125821 &                                           Gaia DR2 right ascension. \\
                    \texttt{dec} &          -60.522668 &                                               Gaia DR2 declination. \\
               \texttt{parallax} &            1.980389 &                                                  Gaia DR2 parallax. \\
        \texttt{parallax\_error} &            0.031363 &                                      Gaia DR2 parallax uncertainty. \\
                   \texttt{pmra} &           -4.480769 &                    Gaia DR2 proper motion $\mu_\alpha \cos \delta$. \\
                  \texttt{pmdec} &            8.203306 &                                Gaia DR2 proper motion $\mu_\delta$. \\
     \texttt{phot\_g\_mean\_mag} &           12.958735 &                                             Gaia DR2 $G$ magnitude. \\
    \texttt{phot\_bp\_mean\_mag} &           13.309761 &                                 Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
    \texttt{phot\_rp\_mean\_mag} &           12.406217 &                                 Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
       \texttt{radial\_velocity} &           -0.227623 &                              Gaia DR2 heliocentric radial velocity. \\
\texttt{radial\_velocity\_error} &            1.945767 &                               Gaia DR2 radial velocity uncertainty. \\
             \texttt{subcluster} &                halo &                          Is star in core (CG18) or halo (KC19+M21)? \\
               \texttt{in\_CG18} &               False &                             Star in \citet{cantatgaudin_gaia_2018}. \\
               \texttt{in\_KC19} &                True &                            Star in \citet{kounkel_untangling_2019}. \\
                \texttt{in\_M21} &               False &                                      Star in \citet{meingast_2021}. \\
             \texttt{(Bp-Rp)\_0} &            0.769225 & Gaia Bp-Rp color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
               \texttt{in\_SetA} &                True &       In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
               \texttt{in\_SetB} &                True &         In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
          \texttt{is\_phot\_bin} &                True &                         True if $>0.3$ mag above cluster isochrone. \\
         \texttt{is\_astrm\_bin} &                True &                                                 True if RUWE > 1.2. \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{
Table~\ref{tab:maintable} is published in its entirety in a
machine-readable format.  One entry is shown for guidance regarding
form and content.  This table is a concatenation of all candidate
NGC\,2516 members reported by \citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
based on the Gaia DR2 data.  Different levels of purity and
completeness can be achieved using different cuts on
photometric periods and lithium eqiuvalent widths (see
Figure~\ref{fig:rot}).  Sets $\mathcal{A}$ and $\mathcal{B}$ provide
two levels of cleaning based on rotation periods that may be useful.
Note that when the target star is the only star present in the TESS
aperture, \texttt{nequal}$=0$, \texttt{nclose}$=1$, and
\texttt{nfaint}$=1$.
}
\vspace{-0.5cm}
\end{deluxetable*}

### 2021-04-22 cluster_paper/ms.tex
update ms

@@
   with a gyrochronological age of 150\,Myr.  A comparison sample of
-  field stars shows no such trend.  In addition, the lithium
-  abundances of stars in the halo are elevated relative to the field,
-  and are correlated with the stellar rotation rate in a manner
-  consistent with previous observations of similar clusters.  This
-  work supports a new paradigm wherein the halos of open clusters are
-  often more populous than their cores.  We discuss implications for
-  spectroscopic survey targeting, open cluster dispersal, and planet
-  searches around young stars.
+  field stars shows no such trend.  The lithium abundances of stars in
+  the halo are higher than in the field, and are correlated with the
+  stellar rotation rate and binarity fraction, as has been
+  noted in other young open clusters.  Broadly speaking,
+  this work supports a new paradigm wherein the halos of open clusters
+  are often more populous than their cores.  We discuss implications
+  for spectroscopic survey targeting, open cluster dispersal, and
+  planet searches around young stars.
 \end{abstract}
@@
 As part of a Cluster Difference Imaging Photometric Survey (CDIPS,
-\citealt{bouma_cdipsI_2019}), we have recently been making TESS light
+\citealt{bouma_cdipsI_2019}), we have been making TESS light
 curves of stars reported to be members of coeval populations and
@@
 chose \cn\ for this analysis because it is young (100-200\,Myr) and
-sufficiently nearby ($d\approx400$\,pc) to facilitate photometric
-rotation measurements using TESS.  Its halo was recently reported by
+close enough ($d\approx400$\,pc) to facilitate photometric
+rotation measurements using TESS.  Its halo was reported by
 \citet{kounkel_untangling_2019} to span
@@
 dispersed group.  ``Halo'' is concise and model agnostic, but is not
-an ideal term because it connotes spherical symmetry, which is usually
-not the case.  The halo of \cn\ is most accurately described as
-consisting of a leading and trailing tail.  We call it a halo for
-simplicity.
+an ideal term because it connotes spherical symmetry, which is rarely
+the case.  The halo of \cn\ is most accurately described as consisting
+of a leading and trailing tail.  We call it a halo for simplicity.
 
@@
 
+\input{maintable_header.tex}
 \begin{figure*}[t]
@@
     right and lower panels of Figure~\ref{fig:physical_x_rotn}
-    quantify the statement: $\approx$75\% of the ``inner-most''
+    quantify the statement: $\approx$70\% of the ``inner-most''
     members have detected rotation periods consistent with a
-    gyrochronological age near that of the Pleiades.  The false
-    positive rate increases to $\approx$50\% at physical separations
-    of $\approx$200\,pc.  Given the number density ratios of cluster
-    to field stars at these wide separations ($\approx 10^{-3}$,
+    gyrochronological age near that of the Pleiades. At physical
+    separations of $\approx$200\,pc, this fraction decreases to
+    $\approx$50\%.  Given the number density ratios of cluster to
+    field stars at these wide separations ($\approx 10^{-3}$,
     \citealt{meingast_2021}), these contamination fractions are not as
@@
      \citealt{soderblom_evolution_1993};
+     \citealt{meibom_effect_2007};
      \citealt{jeffries_m35_li_2020}; \citealt{gillen_ngts_2020}).  The
@@
      factor of $\gtrsim2$ enhancement in binarity fraction for fast
-     {\it vs.} slowly rotating G and early K dwarfs.  We have pointed
+     {\it vs{.}} slowly rotating G and early K dwarfs.  We have pointed
      out in Section~\ref{discussion:lithium} that one plausible
@@
      phase may be truncated in binary systems, which could free the
-     stars to spin up on the pre-MS, producing less efficient internal
-     mixing and a longer-lasting presence of photospheric lithium
+     stars to spin up as they contract on the pre-MS.  This spin up
+     would be expected to produce less efficient internal mixing and a
+     longer-lasting presence of photospheric lithium
      \citep{eggenberger_impact_2012}.
@@
 %
-L.G.B. and J.H. acknowledge support by the TESS GI Program, {\bf program
-NUMBER, through NASA grant NUMBER}.
+L.G.B. and J.H. acknowledge support by the TESS GI Program, programs
+G011103 and G022117, through NASA grants 80NSSC19K0386 and {\bf
+XXXXXXXXXXXXX}.
+%
 L.G.B. was also supported by a Charlotte Elizabeth Procter Fellowship
@@
  	{\it Astrometry}:
- 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
+ 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2020_edr3}.
  	{\it Imaging}:
-    Second Generation Digitized Sky Survey,
-    SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
+    Second Generation Digitized Sky Survey.%,
+    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
  	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
@@
 	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
+	AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
  	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
- 	%{\bf VLT (number), UVES and GIRAFFE} (CITE: Pasquini et al 2002)
+ 	VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
 % 	Euler1.2m~(CORALIE),
@@
   of stellar color.} Fifty stars are randomly selected from Set
-  $\mathcal{B}$ (see Section~\ref{subsubsec:cluster}), and the light
-  curves are detrended and cleaned as discussed in
+  $\mathcal{B}$ (see Section~\ref{subsubsec:cluster}). The light
+  curves were detrended and cleaned as discussed in
   Appendix~\ref{app:detrending}.  Each light curve is separated by
@@
   stars (orange) show a different rotation period distribution than
-  the kinematically selected \cn\ members (black).  Field and \cn\
-  stars are subject to the same ``Set $\mathcal{A}$''
+  the kinematically selected \cn\ members (black).  The field and \cn\
+  stars are both subject to the same ``Set $\mathcal{A}$''
   period-measurement and cleaning procedures.
@@
 
-\section{Projection effects for 2D velocities}
+\section{Projection Effects for Tangential Velocities}
 \label{app:vproj}

### 2021-04-22 drivers/write_NGC2516_main_table.py

@@ -16,7 +16,7 @@ from earhart.helpers import (
 )
 from earhart.lithium import _get_lithium_EW_df
 
-include_lithium = False
+include_lithium = False #FIXME: need this to be true!
 N_members = 3298
 
 # made by cluster_rotation.get_auto_rotation_periods
@@ -44,8 +44,34 @@ dropcols = ['level_0', 'source_id_2', 'index', 'datalink_url',
             'a_g_percentile_lower', 'a_g_percentile_upper', 'e_bp_min_rp_val',
             'e_bp_min_rp_percentile_lower', 'e_bp_min_rp_percentile_upper',
             'flame_flags', 'radius_val', 'radius_percentile_lower',
-            'radius_percentile_upper', 'lum_val', 'lum_percentile_lower',
-            'lum_percentile_upper']
+            'radius_percentile_upper', 'lum_val',
+            'lum_percentile_lower', 'lum_percentile_upper',
+            'ra_error', 'dec_error', 'parallax_over_error',
+            'pmra_error', 'pmdec_error', 'ra_dec_corr',
+            'ra_parallax_corr', 'ra_pmra_corr', 'ra_pmdec_corr',
+            'dec_parallax_corr', 'dec_pmra_corr', 'dec_pmdec_corr',
+            'parallax_pmra_corr', 'parallax_pmdec_corr',
+            'pmra_pmdec_corr', 'astrometric_n_obs_al',
+            'astrometric_n_obs_ac', 'astrometric_n_good_obs_al',
+            'astrometric_n_bad_obs_al', 'astrometric_gof_al',
+            'astrometric_chi2_al', 'astrometric_excess_noise',
+            'astrometric_excess_noise_sig',
+            'astrometric_params_solved', 'astrometric_primary_flag',
+            'astrometric_weight_al', 'astrometric_pseudo_colour',
+            'astrometric_pseudo_colour_error', 'mean_varpi_factor_al',
+            'astrometric_matched_observations',
+            'visibility_periods_used', 'astrometric_sigma5d_max',
+            'frame_rotator_object_type', 'matched_observations',
+            'duplicated_source', 'phot_g_n_obs', 'phot_g_mean_flux',
[…]

### 2021-04-22 cluster_paper/maintable_header.tex
update with my own Li EW fitting
the magic M dwarf EW vs Teff relation has disappeared.

@@
 based on the Gaia DR2 data.  Different levels of purity and
-completeness can be achieved using different cuts on
-photometric periods and lithium eqiuvalent widths (see
-Figure~\ref{fig:rot}).  Sets $\mathcal{A}$ and $\mathcal{B}$ provide
-two levels of cleaning based on rotation periods that may be useful.
-Note that when the target star is the only star present in the TESS
-aperture, \texttt{nequal}$=0$, \texttt{nclose}$=1$, and
-\texttt{nfaint}$=1$.
+completeness can be achieved using different cuts on photometric
+periods and lithium eqiuvalent widths (see Figure~\ref{fig:rot}).
+Sets $\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
+cleaning based on rotation periods.  Note that when the target star is
+the only star present in the TESS aperture, \texttt{nequal}$=0$,
+\texttt{nclose}$=1$, and \texttt{nfaint}$=1$.  The light curves are
+available at \url{https://archive.stsci.edu/hlsp/cdips}.
 }

### 2021-04-22 drivers/measure_gaiaeso_Li_EWs.py

@@ -58,8 +58,6 @@ for _, r in df.iterrows():
         print(f'found {csv_outpath}')
         pass
 
-assert 0
-
 #
 # compare with Gaia ESO
 #
@@ -94,8 +92,13 @@ print(f'Got {len(mdf)} calibration stars')
 
 plt.close('all')
 f,ax = plt.subplots(figsize=(4,4))
-ax.scatter(mdf.EWLi, mdf.Fitted_Li_EW_mA)
+ax.scatter(mdf.EWLi, mdf.Fitted_Li_EW_mA, c='k', s=2, zorder=1)
+sel = (mdf.Teff < 4400)
+sdf = mdf[sel]
+ax.scatter(sdf.EWLi, sdf.Fitted_Li_EW_mA, c='r', s=2, label='Teff<4400K',
+           zorder=2)
 ax.plot(np.arange(-50,250,1), np.arange(-50,250,1), 'k--')
+ax.legend()
 ax.set_xlabel('Randich+18 EW Li from Gaia-ESO [mA]')
 ax.set_ylabel('My EW Li from GAIAESODR4 [mA]')
 outpath = '../results/lithium/validate_my_GAIAESODR4_EWs_vs_Randich18.png'
@@ -105,4 +108,7 @@ plt.close('all')
 
 print(42*'-')
 print('Check the following to ensure you understand differences...')
-print(mdf.sort_values(by='Fitted_Li_EW_mA')[['Fitted_Li_EW_mA','EWLi','source_id','Teff']])
+smdf = mdf.sort_values(by='Fitted_Li_EW_mA')[['Fitted_Li_EW_mA','EWLi','source_id','Teff']]
+print(smdf.head(n=20))
+
+import IPython; IPython.embed()

### 2021-04-22 earhart/lithium.py

@@ -16,9 +16,13 @@ from astropy.coordinates import SkyCoord
 
 from earhart.paths import DATADIR, RESULTSDIR
 
-def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99):
+def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99,
+                       use_my_GaiaESO_EWs=1):
     # gaiaeso and galahdr3: booleans for whether to retrieve
 
+    #
+    # Randich+18 measurements from the Gaia-ESO spectra
+    #
     datapath = os.path.join(DATADIR, 'lithium',
                             'randich_fullfaintkinematic_xmatch_20210310.csv')
     if not os.path.exists(datapath):
@@ -35,7 +39,7 @@ def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99):
     from earhart.lithium import get_GalahDR3_li_EWs
     galah_df = get_GalahDR3_li_EWs()
     # the gaussian fit, numerically integrated
-    galah_li_col = 'Fitted_Li_EW_mA'
+    my_li_col = 'Fitted_Li_EW_mA'
 
     s_gaiaeso_df = gaiaeso_df[[gaiaeso_li_col, gaiaeso_li_errcol, 'source_id']]
     s_gaiaeso_df = s_gaiaeso_df.rename({gaiaeso_li_col: 'Li_EW_mA',
@@ -43,13 +47,31 @@ def _get_lithium_EW_df(gaiaeso, galahdr3, EW_CUTOFF_mA=-99):
     s_gaiaeso_df['Li_EW_mA_merr'] = s_gaiaeso_df['Li_EW_mA_perr']
     #s_gaiaeso_df['Li_provenance'] = 'Randich+18'
 
-    s_galah_df = galah_df[[galah_li_col, galah_li_col+"_perr",
-                           galah_li_col+"_merr", 'source_id']]
-    s_galah_df = s_galah_df.rename({galah_li_col: 'Li_EW_mA',
-                                    galah_li_col+"_perr": 'Li_EW_mA_perr',
-                                    galah_li_col+"_merr": 'Li_EW_mA_merr'
+    s_galah_df = galah_df[[my_li_col, my_li_col+"_perr",
+                           my_li_col+"_merr", 'source_id']]
+    s_galah_df = s_galah_df.rename({my_li_col: 'Li_EW_mA',
[…]

### 2021-04-22 earhart/plotting.py

@@ -1816,7 +1816,7 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
         ax.scatter(
             bouvier18_df[dets]['BpmRp0'], bouvier18_df[dets]['WLi'], c='gray', alpha=1,
             zorder=-1, s=7, edgecolors='k', marker='s',
-            linewidths=0.1, label='Pleiades'
+            linewidths=0.2, label='Pleiades'
         )
         # ax.scatter(
         #     bouvier18_df[~dets]['BpmRp0'], bouvier18_df[~dets]['WLi'], c='gray', alpha=0.8,
@@ -1836,9 +1836,9 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
         dets = (s93_df.l_WLi != "<")
         # vizier randomly decided to divide by 10
         ax.scatter(
-            s93_df[dets]['BpmRp0'], 10*s93_df[dets]['WLi'], c='hotpink', alpha=1,
-            zorder=0, s=9, edgecolors='k', marker='s',
-            linewidths=0.1, label='Praesepe'
+            s93_df[dets]['BpmRp0'], 10*s93_df[dets]['WLi'], c='yellow', alpha=1,
+            zorder=0, s=8, edgecolors='k', marker='o',
+            linewidths=0.2, label='Praesepe'
         )
 
 
@@ -1875,8 +1875,6 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
             #     marker=None, markersize=2
             # )
 
-
-
         ax.legend(loc='upper left', handletextpad=0.05)
 
     if not trimx:
@@ -1897,6 +1895,40 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
     ax.set_ylim([-50, 350])
 
     format_ax(ax)
+
+    #
+    # twiny for the SpTypes
+    #
+    tax = ax.twiny()
+    tax.set_xlabel('Spectral Type')
+
+    xlim = ax.get_xlim()
+    sptypes, BpmRps = get_SpType_BpmRp_correspondence(
+        ['A5V','F0V','F5V','G0V','K0V','K3V','K5V','K7V','M0V','M1V','M3V']
+    )
+    print(sptypes)
+    print(BpmRps)
+
+    xvals = np.linspace(min(xlim), max(xlim), 100)
[…]

### 2021-04-23 cluster_paper/maintable_header.tex
update ms

@@
 %% Use Default (12pt)
-\tabletypesize{\scriptsize}
-%\tabletypesize{\small}
+%\tabletypesize{\scriptsize}
+\tabletypesize{\small}
 
@@
                \texttt{source\_id} & 5290824530952816896 &                                                    Gaia DR2 source identifier. \\
+         \texttt{source\_id\_edr3} & 5290824530952816896 &                                                   Gaia EDR3 source identifier. \\
+                 \texttt{in\_SetA} &                True &                  In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
+                 \texttt{in\_SetB} &                True &                    In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
          \texttt{n\_cdips\_sector} &                   3 &                                Number of TESS sectors with CDIPS light curves. \\
@@
                    \texttt{lspval} &            0.171629 &                                Lomb-Scargle periodogram value for best period. \\
+               \texttt{spdmperiod} &            1.297662 &                                                  Stellingwerf PDM best period. \\
+                  \texttt{spdmval} &            0.826804 &                            Stellingwerf PDM periodogram value for best period. \\
                    \texttt{nequal} &                 0.0 &                     Number of stars brighter than the target in TESS aperture. \\
@@
                    \texttt{nfaint} &                 1.0 &                        Number of stars with $\Delta T > 2.5$ in TESS aperture. \\
-             \texttt{spdmperiod} &            1.297662 &                                       Stellingwerf PDM best period. \\
-                \texttt{spdmval} &            0.826804 &                 Stellingwerf PDM periodogram value for best period. \\
-             \texttt{ref\_epoch} &              2015.5 &                                         Reference epoch for RA/dec. \\
                        \texttt{ra} &          120.125821 &                                                      Gaia DR2 right ascension. \\
                       \texttt{dec} &          -60.522668 &                                                          Gaia DR2 declination. \\
+               \texttt{ref\_epoch} &              2015.5 &                           Reference epoch for right ascension and declination. \\
                  \texttt{parallax} &            1.980389 &                                                             Gaia DR2 parallax. \\
@@
                \texttt{(Bp-Rp)\_0} &            0.769225 &            Gaia Bp-Rp color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
-               \texttt{in\_SetA} &                True &       In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
-               \texttt{in\_SetB} &                True &         In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
             \texttt{is\_phot\_bin} &                True &                                    True if $>0.3$ mag above cluster isochrone. \\
-         \texttt{is\_astrm\_bin} &                True &                                                 True if RUWE > 1.2. \\
+           \texttt{is\_astrm\_bin} &                True &                                                  True if Gaia EDR3 RUWE > 1.2. \\
+                     \texttt{ruwe} &            2.221799 &                                                                Gaia EDR3 RUWE. \\
+      \texttt{Li\_EW\_mA\_GaiaESO} &                 NaN &                                    Gaia-ESO Li doublet equivalent width [m\AA] \\
+\texttt{Li\_EW\_mA\_perr\_GaiaESO} &                 NaN & Gaia-ESO Li doublet EW 84$^{\mathrm{th}}$-50$^{\mathrm{th}}$ percentile [m\AA] \\
+\texttt{Li\_EW\_mA\_merr\_GaiaESO} &                 NaN & Gaia-ESO Li doublet EW 50$^{\mathrm{th}}$-16$^{\mathrm{th}}$ percentile [m\AA] \\
+        \texttt{Li\_EW\_mA\_GALAH} &                 NaN &                                       GALAH Li doublet equivalent width [m\AA] \\
+  \texttt{Li\_EW\_mA\_perr\_GALAH} &                 NaN &    GALAH Li doublet EW 84$^{\mathrm{th}}$-50$^{\mathrm{th}}$ percentile [m\AA] \\
+  \texttt{Li\_EW\_mA\_merr\_GALAH} &                 NaN &    GALAH Li doublet EW 50$^{\mathrm{th}}$-16$^{\mathrm{th}}$ percentile [m\AA] \\
 \enddata

### 2021-04-23 cluster_paper/ms.tex
update ms

@@
 Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
-$G$ magnitude as a function of observed $\bpmrp$ color.  Spectral
-types are interpolated from the \citet{pecaut_mamajek_2013} table,
-after accounting for reddening as described
+$G$ magnitude as a function of observed $\bpmrp$ color.  Magnitudes
+are from Gaia EDR3; we performed the Gaia DR2 to EDR3 cross-match
+using the pre-computed \texttt{gaiaedr3.dr2\_neighbourhood} table
+available at the Gaia archive, and required the closest (proper motion
+and epoch-corrected) source to be the single match.  Spectral types
+are interpolated from the \citet{pecaut_mamajek_2013} table, after
+accounting for reddening as described
 below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
@@
 
-\input{maintable_header.tex}
 \begin{figure*}[t]
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f3a.pdf}
-		}
-
-    \vspace{-0.5cm}
+			\includegraphics[width=0.72\textwidth]{f3a.pdf}
+		}
+
+    \vspace{-0.7cm}
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f3b.pdf}
+			\includegraphics[width=0.72\textwidth]{f3b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f3c.pdf}
+			\includegraphics[width=0.72\textwidth]{f3c.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f3d.pdf}
+			\includegraphics[width=0.72\textwidth]{f3d.pdf}
 		}
@@
 these stars is brighter than the CDIPS magnitude limit of
-$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
-source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
+$G_\mathrm{RP}=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
+source list had $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
 from \citet{meingast_2021} which were not available at the time of the
 CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
-distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
+distance of \cn, the $G_\mathrm{RP}=16$\,mag cutoff imposed during the CDIPS
 processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
@@
 
-We removed systematic trends shared between all light curves on each
-CCD in each sector individually, and stitched together the resulting
+We removed systematic trends common to the light curves on each
+CCD in each individual sector, and stitched together the resulting
 light curves before searching for rotation-induced periodicity.
@@
 expectations, was expected to give the optimal balance between light
-from the target star and background-light \citep{Sullivan_2015}.  This
+from the target star and the sky background \citep{Sullivan_2015}.  This
 typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
@@
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.72cm}
   \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
@@
   dereddened color for all kinematic \cn\ members with Gaia-ESO
-  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
+  ($N=459$) or GALAH ($N=107$) spectra available.  The GALAH spectra
   comprise slightly over half of the halo stars, due to the
   non-targeted selection function of that survey.  Field stars (gray
-  points) are stars with detected
-  transiting planets from Kepler 
+  points) are stars with Kepler planets
   \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
   with Li detections are from \citet{bouvier_pleiades_lirot_2018};
-  Praesepe detections (650\,Myr) are shown in violet
+  Praesepe detections (650\,Myr) are shown in yellow
   \citep{soderblom_praesepe_1993}.  Points with ${\rm
-  EW}\approx0\,$m\AA\ are non-detections.
+  EW}\approx0\,$m\AA\ are non-detections; all stars shown
+  later than $\approx$M0 do not have significant lithium.
   The Li measurements for both core and halo stars in \cn\ are
   consistent with a near-Pleiades age.
-  {\bf LGB TODO: the increasing EWs for the Gaia-ESO M dwarfs must be erroneous. What's happening there?}
   \label{fig:lithiumcorehalo}
@@
 
-Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
-and UVES spectrographs based on previously reported literature members
+The Gaia-ESO collaboration chose candidate \cn\ members to observe
+with GIRAFFE and UVES based on previously reported literature members
 and publicly available photometry.  Since the existence of the \cn\
@@
 stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
-reported stellar parameters (including lithium equivalent
-widths and metallicities) for 796 stars that they considered possible
-\cn\ members.  Cross-matching against our kinematic list of
+reported stellar parameters (including lithium equivalent widths and
+metallicities) for 796 stars that they considered possible \cn\
+members.  Cross-matching these against our kinematic list of
 \nkinematic\ candidate members by position and imposing a 0$\farcs$5
-maximum separation limit yields 492 kinematic members for which a
-Gaia-ESO spectrum was acquired and stellar parameters were reported.
-15 of these (all with $\bpmrpo > 2.0$) are spurious matches based on
-the Gaia color and GES effective temperature, and we remove them.
-This yields 477 stars, of which 436 are in the core, and 41 are in the
-halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
-We verified by directly querying the ESO archive that
-\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
-Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
-their results rather than repeating their analysis.
+maximum separation limit yielded 459 kinematic members with available
+spectra, 417 in the core and 42 in the halo of \cn.  The lop-sided
+ratio is due to the Gaia-ESO selection function.  An inspection of the
+effective temperature {\it vs{.}} lithium equivalent width based on
+their Table~2 showed the surprising feature that stars with $T_{\rm
+eff}\lesssim 4000\,{\rm K}$ tend to have non-zero equivalent widths,
+despite the expectation that their lithium should be fully depleted.
+Whether these measurements were significant was not clear, and so we
+opted to download the spectra from the ESO archive to re-measure
+equivalent widths.
 
@@
 lithium, i.e., no lithium flag set.
+We downloaded\footnote{Via \url{datacentral.org.au/services/download},
+using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
+all 107 entries.
 
@@
 
-We downloaded\footnote{Via \url{datacentral.org.au/services/download},
-using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
-all 107 entries.  We then measured the lithium equivalent widths using
-the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
-by first focusing in on a 15\AA\ window centered on the 6707.835\AA\ 
-lithium doublet.  We then continuum normalized the spectra using a
-third-degree Chebyshev series, while excluding any regions that showed
-absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
-6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
-to the continuum normalized spectra, considering only a $\pm 1\,$\AA\ 
-window centered on the Li doublet.  The equivalent widths were then
-evaluated by numerically integrating the fitted model over the same
-window.  Our approach therefore includes the Fe\,6707.44 blend in the
-reported Li equivalent widths---which leads to systematic
-overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
-e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
-uncertainties on the EWs, we repeated the procedure twenty times, but
-added noise to the spectra, drawn from a normal distribution with a
-scale set by the standard deviation of the continuum.  The reported
-uncertainties are then drawn from the $84^{\rm th}$ and $16^{\rm th}$
-percentiles of the resulting EW distribution.  We verified the overall
-scale of our results by comparing the 21 stars that overlapped between
-GALAH and Gaia-ESO.  17 of the 21 measurements agreed within
-1-$\sigma$.  The remaining four were stars with $T_{\rm
-eff}>6200\,{\rm K}$, for which \citet{randich_gaiaeso_2018} reported
-Li EW detections ranging from $50$ to $75\,$m\AA, and for which we
-found either null results or only slightly significant
-($\approx$30m\AA) detections.
+We measured the lithium equivalent widths using the \texttt{specutils}
+package \citep{specutils_v1pt1}.  We did this by first focusing in on
+a 15\AA\ window centered on the 6707.835\AA\ lithium doublet.  We then
+continuum normalized the spectra using a third-degree Chebyshev
+series, while excluding any regions that showed absorption lines ({\it
+e.g.}, Fe\,\textsc{I} is present at 6703.58\,\AA and 6705.10\,\AA).
+We proceeding by fitting a Gaussian to the continuum normalized
+spectra, considering only a $\pm 1\,$\AA\ window centered on the Li
+doublet.  The equivalent widths were then evaluated by numerically
+integrating the fitted model over the same window.  Our approach
+therefore includes the Fe\,6707.44 blend in the reported Li equivalent
+widths---which leads to systematic overestimates of $\approx$10 to
+15\,m\AA\ \citep[{\it e.g.},][]{bouvier_pleiades_lirot_2018}.
+Finally, to derive uncertainties on the EWs, we repeated the procedure
+twenty times, but added noise to the spectra, drawn from a normal
+distribution with a scale set by the standard deviation of the
+continuum.  The reported uncertainties are then drawn from the
+$84^{\rm th}$ and $16^{\rm th}$ percentiles of the resulting EW
+distribution.  We verified the overall scale of our results by
+comparing our measurements with those of \citet{randich_gaiaeso_2018}.
+Most stars with $T_{\rm eff} > 4500\,{\rm K}$ followed a 1-1 relation.
+For cooler stars, our EW measurements are systematically lower, as
+expected given the depletion timescales for stars at these ages
+\citep[{\it e.g.},][]{soderblom_ages_2014}.  We visually verified from
+the spectra that, for the most part, we prefer our measurements.
+
+% the 21 stars that overlapped between GALAH and Gaia-ESO.  17
+% of the 21 measurements agreed within 1-$\sigma$.  The remaining four
+% were stars with $T_{\rm eff}>6200\,{\rm K}$, for which
+% \citet{randich_gaiaeso_2018} reported Li EW detections ranging from
+% $50$ to $75\,$m\AA, and for which we found either null results or only
+% slightly significant ($\approx$30m\AA) detections.
 
@@
 Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
-Gaia-ESO and GALAH results, with the points colored according to whether the
-stars are in the core or the halo of the cluster.  The GALAH spectra
-span a color range of $0<\bpmrpo<1.5$, due to the $V<14$ brightness
-cutoff of the survey.  The overall increase of Li EW from
+Gaia-ESO and GALAH results, with the points colored according to
+whether the stars are in the core or the halo of the cluster.  The
+GALAH spectra span a color range of $0<\bpmrpo<1.5$, due to the $V<14$
+brightness cutoff of the survey.  The overall increase of Li EW from
 $0<\bpmrpo<1.0$ is driven by the temperatures in the stellar
@@
 (we interpolated from their effective temperatures using the
-\citet{pecaut_mamajek_2013}
-table).  For clarity, we have omitted the few upper
-limits reported by \citet{bouvier_pleiades_lirot_2018}.  The field
-star distribution peaks at late F-type stars.  The majority of
+\citet{pecaut_mamajek_2013} table).  For clarity, we have omitted the
+few upper limits reported by \citet{bouvier_pleiades_lirot_2018}.  The
+field star distribution peaks at late F-type stars.  The majority of
 kinematically selected stars in the core and the halo show lithium
@@
 
-One possible explanation is that the observed halo structures are
-tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea is
-that as the cluster contracts, stars escape out of the Lagrange points
-imposed by the Milky Way's potential, and form leading and lagging
-arms due to differential rotation in the Galaxy.  Kinematic evidence
-that the core of \cn\ may be undergoing contraction has recently been
-presented by \citet{healy_stellar_2020}, which might increase the rate
-at which stars escape.  Whether the exact contraction rate, and the
-corresponding evaporation rate of the core are sufficient to
-explain the formation of the halo has yet to be assessed.  The
-presence or lack of epicyclic overdensities in the tails might also
-help clarify whether this scenario is happening
-\citep{kupper_more_2012}.
+The most likely explanation seems to be that the observed halo
+structures are tidal tails \citep[{\it
+e.g.},][]{chumak_tails_2006,krumholz_star_2019}. The idea is that
+stars near the cluster's tidal radius escape slowly due to the
+galactic tide, and subsequently form leading and lagging arms due to
+differential rotation in the Galaxy.  The contraction rate of the
+cluster can affect this process.  Whether the exact contraction rate
+of \cn\ \citep{healy_stellar_2020}, and the corresponding evaporation
+rate of the core are sufficient to explain the formation of the halo
+has yet to be assessed. 
+% The presence or lack of epicyclic
+% overdensities in the tails might also help clarify whether this
+% scenario is happening \citep{kupper_more_2012}.
 
@@
 have formed in the same ``clump'' as those in the cluster core.  In
-other words, the formation environment might be better described as a
+other words, the formation environment might be best described as a
 giant molecular filament, rather than a giant molecular cloud.  A
@@
 
-One way to rule between the two possibilities would be to
-observationally establish an age-size correlation.  If the
+Any differences in the stellar mass function between the core and halo
+could be informative, since the tidal tail explanation predicts that
+the mean tail star is expected to be less massive than the mean core
+star \citep[{\it e.g.},][]{chumak_tails_2006}.
+Figure~\ref{fig:massfunction} shows the histogram of $A_G$-corrected
+absolute magnitudes, with the corresponding
+\citet{pecaut_mamajek_2013} mass estimates.  The stars closest
+to the core do
+indeed seem to have a greater proportion of the $\gtrsim$0.8$M_\odot$
+stars, while the tail stars have a greater proportion of
+$\lesssim$0.4$M_\odot$ stars.  The mean star in the inner separation bin has
+$({\rm M}_G)_0=7.08$ ($\approx$0.69$M_\odot$), while in the outer bin
+the mean star has $({\rm M}_G)_0=7.61$ ($\approx$0.63$M_\odot$).
+While the difference does appear significant given the number of
+stars involved, the theoretical implications are not entirely clear,
+since the ``molecular filament'' explanation might be able to
+accommodate mass differences between the core and tail through density
+gradients across the initial filament.
+
+One possible approach for distinguishing the two explanations could be
+to observationally establish an age-size correlation.  If the
 $\approx$100\,Myr halos are produced primarily through evaporation and
-expansion, then the younger clusters should be significantly smaller.
-A theoretical approach along similar lines would be to perform a
-dynamical analysis of whether filaments like Orion\,A, or those in the
-\citet{zucker_physical_2018} sample, are at all expected to evolve
-into structures resembling \cn.
+expansion, then the younger clusters should be significantly smaller
+\citep{chumak_tails_2006,chumak_realOCs_2006}.  A theoretical approach
+along complementary lines would be to perform a dynamical analysis of
+whether filaments like Orion\,A, or those in the
+\citet{zucker_physical_2018} sample are at all expected to evolve into
+structures resembling \cn.
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.7\textwidth]{f15.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Absolute magnitudes (and masses) for the inner and
+			outermost stars in \cn.}
+		The \nkinematic\ candidate \cn\ members were split into three groups
+		according to their distance from the cluster center.
+		The inner and outermost groups, each containing 1,099 stars, are shown to
+		provide the clearest picture with respect to expectations for mass
+		segregation.
+		The mean halo star does have lower mass, consistent with expectations
+		for tidal tails.
+		\label{fig:massfunction}
+	}
+\end{figure*}
 
@@
 periods tend to rotate rapidly ($<3$\,days), though there may also be
-late M dwarfs with slow ($>20$\,days) rotation periods.  Although
+late M dwarfs with slow ($>20$\,days) rotation periods \citep{stauffer_rotation_2016}.  Although
 these late M dwarfs are beyond the magnitude limit of our analysis,
@@
 magnetic braking is not always the most important effect.  The
-question of why the rapid rotators exist, for instance, has not yet
-been definitively answered.  Models that introduce variable
+question of why the rapid rotators exist, for instance, requires
+additional physics.  Models that introduce variable
 core-envelope decoupling timescales and pre-MS disk lifetimes can
@@
 \citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015},
-but the question then shifts to why these decoupling timescales should
+but the question then shifts to why these timescales should
 differ for stars that are otherwise identical.  Another possible
@@
 clusters and the field have preivously been noted
-\citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}
+\citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}.
 Related effects could include tidal synchronization for the shortest
@@
 doublet, and used them to convert equivalent widths to abundances.
-Non-LTE corrections affect the abundances by $<0.5\,{\rm dex}$ and do
+Non-LTE corrections do
 not change the overall relationships between Li abundance and stellar
@@
 stars in the core of \cn.  Our concatenation of the
-\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
-spectra represent a significant expanasion in volume and color range
+Gaia-ESO and GALAH-DR3
+spectra represents a significant expanasion in volume and color range
 from this earlier analysis.
@@
 What causes the correlation between lithium abundance and rotation at
-fixed stellar mass?  This question has yet to be definitively
-answered, but hints likely lie in shared correlations between lithium
+fixed stellar mass?  Hints likely lie in shared correlations between lithium
 abundance, rotation rate, radius inflation, internal mixing, and
@@
 based on pure hydrodynamics is that the rapid rotation leads to less
-efficient convective penetration, which then lowers the mixing
+efficient convective penetration (``overshoot''), which then lowers the mixing
 efficiency at the convective-radiative boundary
@@
 question of why some of the G and K dwarfs rotate rapidly to begin
-with.  One clue noted above is that the rapid rotators tend to be in
+with.  One clue could be that the rapid rotators tend to be in
 photometric or astrometric binaries.  The majority of such binaries
@@
 It is expected that the circumstellar disks in these binaries have
-different properties than those of single stars, due to {\it e.g.} gap
+different properties than those of single stars, due to {\it e.g.}, gap
 formation in the circumbinary disk, disk truncation, and faster disk
@@
 clarify whether or not all of the rapid rotators are in binaries.  If
-not, the scenario presented above would be falsified.
+not, this scenario would be falsified.
 
@@
 uncertainties to match the absolute Pleiades age uncertainty of $\pm
-20\,$Myr \citep{soderblom_ages_2014}.  Whether this absolute age is at
+20\,$Myr \citep{soderblom_ages_2014}.  Whether this average age is at
 all meaningful, we leave the reader to judge.
@@
     right and lower panels of Figure~\ref{fig:physical_x_rotn}
-    quantify the statement: $\approx$70\% of the ``inner-most''
+    quantify the statement: $\approx$70\% of the inner-most
     members have detected rotation periods consistent with a
@@
     bad as one might fear.
+   %
+   \item {\it The average star in the halo has lower mass (by $\approx$10\%)
+   than the average star in the core.}
+   This is shown in Figure~\ref{fig:massfunction}, and discussed in
+   Section~\ref{subsec:origin}.
+   While the mass function is consistent with expectations for a
+   ``tidal tail'' origin of the halo, the possibility of a primordial halo
+   also merits exploration.
    %
@@
 L.G.B. and J.H. acknowledge support by the TESS GI Program, programs
-G011103 and G022117, through NASA grants 80NSSC19K0386 and {\bf
-XXXXXXXXXXXXX}.
+G011103 and G022117, through NASA grants 80NSSC19K0386 and
+80NSSC19K1728.
 %
@@
 
+\input{maintable_header.tex}
 % \input{TOI837_phot_table.tex}
@@
 	\caption{ {\bf Rotation in \cn\ compared to the field.}
-  Stars with Gaia $Rp<13$\,mag in the cluster and field samples are
+  Stars with Gaia $G_\mathrm{RP}<13$\,mag in the cluster and field samples are
   displayed; fainter field stars were not analyzed.  Field
@@
 and~\ref{fig:binarity}, we searched the CDIPS calibration light curves
-for rotation periods.  This light curve database comprises all $Rp<13$
+for rotation periods.  This light curve database comprises all $G_\mathrm{RP}<13$
 stars that fell on TESS silicon. The need to distinguish field and
@@
 neighborhood of \cn\ yielded \nnbhdcalibstar\ unique stars.  The
-magnitude cut of $Rp<13$ at the distances of the neighborhood sample
+magnitude cut of $G_\mathrm{RP}<13$ at the distances of the neighborhood sample
 corresponds to an extinction-corrected color cutoff of $\bpmrpo

### 2021-04-23 cluster_paper/collect.sh

@@ -32,6 +32,8 @@ cp ../results/vtangential_projection/skymap_tangential_velocity_pmra_cosdec.pdf 
 
 cp ../results/rotation_X_positions/rotation_X_positions_approach2_defaultcleaning.pdf f13.pdf
 
+cp ../results/lumfunction_vs_position/lumfunction_vs_position.pdf f15.pdf
+
 # supplementary
 cp ../results/rotation_X_RUWE/rotation_vs_BpmRp_X_RUWE_viridis_linear.pdf f6.pdf
 

### 2021-04-23 drivers/write_NGC2516_main_table.py

@@ -16,7 +16,7 @@ from earhart.helpers import (
 )
 from earhart.lithium import _get_lithium_EW_df
 
-include_lithium = False #FIXME: need this to be true!
+include_lithium = True
 N_members = 3298
 
 # made by cluster_rotation.get_auto_rotation_periods
@@ -94,6 +94,32 @@ assert len(mdf) == len(df)
 
 is_astrometric_binary = (mdf.ruwe > 1.2)
 df['is_astrometric_binary'] = is_astrometric_binary
+df['ruwe'] = mdf.ruwe
+df['source_id_edr3'] = mdf.source_id_edr3
+
+if include_lithium:
+
+    # get lithium data, first gaia-eso / R+18. rename columns to ensure that
+    # sources with both GaiaESO and GALAH spectra get their EWs reported.
+    ldf0 = _get_lithium_EW_df(1, 0)
+    newcoldict = {}
+    for c in ldf0.columns:
+        if "Li" in c:
+            newcoldict[c] =c+"_GaiaESO"
+    ldf0 = ldf0.rename(newcoldict, axis='columns')
+    mdf = df.merge(ldf0, how='left', on='source_id')
+
+    ldf1 = _get_lithium_EW_df(0, 1)
+    newcoldict = {}
+    for c in ldf1.columns:
+        if "Li" in c:
+            newcoldict[c] =c+"_GALAH"
+    ldf1 = ldf1.rename(newcoldict, axis='columns')
+    outdf = mdf.merge(ldf1, how='left', on='source_id')
+
+    assert len(outdf) == N_members
+
+    df = outdf
 
 
 df = df.drop(['in_match234_alias'], axis=1)
@@ -107,6 +133,19 @@ df = df.rename(columns={
 
 df = df.sort_values(by=['in_SetB'], ascending=False)
 
+orderedcols = [
+'source_id', 'source_id_edr3', 'in_SetA', 'in_SetB', 'n_cdips_sector',
+'period', 'lspval', 'spdmperiod', 'spdmval', 'nequal', 'nclose', 'nfaint',
+'ra', 'dec', 'ref_epoch', 'parallax', 'parallax_error', 'pmra', 'pmdec',
+'phot_g_mean_mag', 'phot_bp_mean_mag', 'phot_rp_mean_mag', 'radial_velocity',
[…]

### 2021-04-26 cluster_paper/maintable_header.tex
update ms

--- 2021-04-22 cluster_paper/maintable_header.tex
@@
 \tablecaption{Rotation periods and lithium equivalent widths for
-candidate NGC\,2516 members.}
+\nkinematic\ candidate NGC\,2516 members.}
 \label{tab:maintable}
@@
 \startdata
-             \texttt{source\_id} & 5290824530952816896 &                                         Gaia DR2 source identifier. \\
-       \texttt{n\_cdips\_sector} &                   3 &                     Number of TESS sectors with CDIPS light curves. \\
-                 \texttt{period} &            1.297662 &                                           Lomb-Scargle best period. \\
-                 \texttt{lspval} &            0.171629 &                     Lomb-Scargle periodogram value for best period. \\
-                 \texttt{nequal} &                 0.0 &          Number of stars brighter than the target in TESS aperture. \\
-                 \texttt{nclose} &                 1.0 &            Number of stars with $\Delta T > 1.25$ in TESS aperture. \\
-                 \texttt{nfaint} &                 1.0 &             Number of stars with $\Delta T > 2.5$ in TESS aperture. \\
-             \texttt{spdmperiod} &            1.297662 &                                       Stellingwerf PDM best period. \\
-                \texttt{spdmval} &            0.826804 &                 Stellingwerf PDM periodogram value for best period. \\
-             \texttt{ref\_epoch} &              2015.5 &                                         Reference epoch for RA/dec. \\
-                     \texttt{ra} &          120.125821 &                                           Gaia DR2 right ascension. \\
-                    \texttt{dec} &          -60.522668 &                                               Gaia DR2 declination. \\
-               \texttt{parallax} &            1.980389 &                                                  Gaia DR2 parallax. \\
-        \texttt{parallax\_error} &            0.031363 &                                      Gaia DR2 parallax uncertainty. \\
-                   \texttt{pmra} &           -4.480769 &                    Gaia DR2 proper motion $\mu_\alpha \cos \delta$. \\
-                  \texttt{pmdec} &            8.203306 &                                Gaia DR2 proper motion $\mu_\delta$. \\
-     \texttt{phot\_g\_mean\_mag} &           12.958735 &                                             Gaia DR2 $G$ magnitude. \\
-    \texttt{phot\_bp\_mean\_mag} &           13.309761 &                                 Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
-    \texttt{phot\_rp\_mean\_mag} &           12.406217 &                                 Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
-       \texttt{radial\_velocity} &           -0.227623 &                              Gaia DR2 heliocentric radial velocity. \\
-\texttt{radial\_velocity\_error} &            1.945767 &                               Gaia DR2 radial velocity uncertainty. \\
-             \texttt{subcluster} &                halo &                          Is star in core (CG18) or halo (KC19+M21)? \\
-               \texttt{in\_CG18} &               False &                             Star in \citet{cantatgaudin_gaia_2018}. \\
+               \texttt{source\_id} & 5290665204845694336 &                                                                   Gaia DR2 source identifier. \\
+         \texttt{source\_id\_edr3} & 5290665204845694336 &                                                                  Gaia EDR3 source identifier. \\
+                 \texttt{in\_SetA} &                True &                                 In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
+                 \texttt{in\_SetB} &                True &                                   In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
+         \texttt{n\_cdips\_sector} &                   7 &                                               Number of TESS sectors with CDIPS light curves. \\
+                   \texttt{period} &            0.228981 &                                                              Lomb-Scargle best period [days]. \\
+                   \texttt{lspval} &            0.657826 &                                               Lomb-Scargle periodogram value for best period. \\
+               \texttt{spdmperiod} &            0.228981 &                                                          Stellingwerf PDM best period [days]. \\
+                  \texttt{spdmval} &            0.372779 &                                           Stellingwerf PDM periodogram value for best period. \\
+                   \texttt{nequal} &                   0 &                                    Number of stars brighter than the target in TESS aperture. \\
+                   \texttt{nclose} &                   1 &                                      Number of stars with $\Delta T > 1.25$ in TESS aperture. \\
+                   \texttt{nfaint} &                   1 &                                       Number of stars with $\Delta T > 2.5$ in TESS aperture. \\
+                       \texttt{ra} &          119.234778 &                                                               Gaia DR2 right ascension [deg]. \\
+                      \texttt{dec} &          -61.015929 &                                                                   Gaia DR2 declination [deg]. \\
+               \texttt{ref\_epoch} &              2015.5 &                                          Reference epoch for right ascension and declination. \\
+                 \texttt{parallax} &            2.458023 &                                                                      Gaia DR2 parallax [mas]. \\
+          \texttt{parallax\_error} &            0.028395 &                                                          Gaia DR2 parallax uncertainty [mas]. \\
+                     \texttt{pmra} &           -4.225826 &                           Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
+                    \texttt{pmdec} &           10.899938 &                                       Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
+       \texttt{phot\_g\_mean\_mag} &           15.778953 &                                                                       Gaia DR2 $G$ magnitude. \\
+      \texttt{phot\_bp\_mean\_mag} &           16.629477 &                                                           Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
+      \texttt{phot\_rp\_mean\_mag} &            14.85896 &                                                           Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
+         \texttt{radial\_velocity} &                 NaN &                                       Gaia DR2 heliocentric radial velocity [km$\,$s${^-1}$]. \\
+  \texttt{radial\_velocity\_error} &                 NaN &                                        Gaia DR2 radial velocity uncertainty [km$\,$s${^-1}$]. \\
+               \texttt{subcluster} &                core &                                                    Is star in core (CG18) or halo (KC19+M21)? \\
+                 \texttt{in\_CG18} &                True &                                                       Star in \citet{cantatgaudin_gaia_2018}. \\
                  \texttt{in\_KC19} &                True &                                                      Star in \citet{kounkel_untangling_2019}. \\
                   \texttt{in\_M21} &               False &                                                                Star in \citet{meingast_2021}. \\
-             \texttt{(Bp-Rp)\_0} &            0.769225 & Gaia Bp-Rp color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
-               \texttt{in\_SetA} &                True &       In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
-               \texttt{in\_SetB} &                True &         In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
-          \texttt{is\_phot\_bin} &                True &                         True if $>0.3$ mag above cluster isochrone. \\
-         \texttt{is\_astrm\_bin} &                True &                                                 True if RUWE > 1.2. \\
+               \texttt{(Bp-Rp)\_0} &            1.636198 & Gaia $G_\mathrm{BP}$-$G_\mathrm{RP}$ color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
+            \texttt{is\_phot\_bin} &               False &                                                   True if $>0.3$ mag above cluster isochrone. \\
+           \texttt{is\_astrm\_bin} &               False &                                                                 True if Gaia EDR3 RUWE > 1.2. \\
+                     \texttt{ruwe} &            0.986349 &                                                                               Gaia EDR3 RUWE. \\
+      \texttt{Li\_EW\_mA\_GaiaESO} &              -1.495 &                          Gaia-ESO Li doublet equivalent width, including the Fe blend [m\AA]. \\
+\texttt{Li\_EW\_mA\_perr\_GaiaESO} &              12.407 &                                     Gaia-ESO Li doublet EW upper uncertainty interval [m\AA]. \\
+\texttt{Li\_EW\_mA\_merr\_GaiaESO} &                 5.0 &                                     Gaia-ESO Li doublet EW lower uncertainty interval [m\AA]. \\
+        \texttt{Li\_EW\_mA\_GALAH} &                 NaN &                             GALAH Li doublet equivalent width, including the Fe blend [m\AA]. \\
+  \texttt{Li\_EW\_mA\_perr\_GALAH} &                 NaN &                                        GALAH Li doublet EW upper uncertainty interval [m\AA]. \\
+  \texttt{Li\_EW\_mA\_merr\_GALAH} &                 NaN &                                        GALAH Li doublet EW lower uncertainty interval [m\AA]. \\
 \enddata

### 2021-04-26 cluster_paper/ms.tex
update ms

@@
 \keywords{
-	stellar ages (1581),
   stellar associations (1582),
   open clusters (1160),
+	stellar ages (1581),
   stellar kinematics (1608),
@@
 \citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
-GALAH DR3 were the most relevant, respectively \citep[{\it
-e.g.},][]{randich_gaiaeso_2018,buder_galah_2020}.  The target
-selection and results from each survey were as follows.
+GALAH DR3 were the most relevant
+\citep{randich_gaiaeso_2018,buder_galah_2020}.  The target selection
+and results from each survey were as follows.
 
@@
 halo was not known at the time of target selection, very few halo
-stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
-reported stellar parameters (including lithium equivalent widths and
-metallicities) for 796 stars that they considered possible \cn\
-members.  Cross-matching these against our kinematic list of
-\nkinematic\ candidate members by position and imposing a 0$\farcs$5
-maximum separation limit yielded 459 kinematic members with available
-spectra, 417 in the core and 42 in the halo of \cn.  The lop-sided
-ratio is due to the Gaia-ESO selection function.  An inspection of the
-effective temperature {\it vs{.}} lithium equivalent width based on
-their Table~2 showed the surprising feature that stars with $T_{\rm
-eff}\lesssim 4000\,{\rm K}$ tend to have non-zero equivalent widths,
-despite the expectation that their lithium should be fully depleted.
-Whether these measurements were significant was not clear, and so we
-opted to download the spectra from the ESO archive to re-measure
-equivalent widths.
+stars are in the sample.  Based on the data,
+\citet{randich_gaiaeso_2018} determined stellar parameters (including
+lithium equivalent widths and metallicities) for 796 stars that they
+considered possible \cn\ members.  Cross-matching these against our
+kinematic list of \nkinematic\ candidate members by position and
+imposing a 0$\farcs$5 maximum separation limit yielded 459 kinematic
+members with available spectra, 417 in the core and 42 in the halo of
+\cn.  The lop-sided ratio is due to the Gaia-ESO selection function.
+An inspection of the effective temperature {\it vs{.}} lithium
+equivalent width based on Table~2 of \citet{randich_gaiaeso_2018}
+showed the surprising feature that stars with $T_{\rm eff}\lesssim
+4000\,{\rm K}$ tend to have non-zero equivalent widths, despite the
+expectation that their lithium should be fully depleted.  Whether
+these measurements were significant was not clear, and so we opted to
+download the spectra from the ESO archive to re-measure equivalent
+widths.
 
@@
 that the ``main'' survey targeted $12<V<14$ stars at
-$\delta<+10^\circ$, provided the stars were at least ten degrees from
-the galactic plane.  The survey is spatially
-inhomogeneous.\footnote{See the footprint at
-\url{https://www.galah-survey.org/news/announcing-galah-dr3}} Special
-targeting for stars in the TESS southern continuous viewing zone, and
-for known open cluster members was also performed.  We identified the
-candidate \cn\ members for which spectra had been obtained by
-searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
-excluding stars with the stellar parameter bit flags 1, 2, 3.  This
-excludes spectra with unreliable broadening, low S/N, and unreliable
-wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
-Since our main focus is measuring equivalent widths for the
-Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of
-our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.
-51 were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
-lithium, i.e., no lithium flag set.
-We downloaded\footnote{Via \url{datacentral.org.au/services/download},
+$\delta<+10^\circ$ in select fields, provided the stars were at least
+ten degrees from the galactic plane.  Special targeting for stars in
+the TESS southern continuous viewing zone, and for known open cluster
+members was also performed.  We identified the candidate \cn\ members
+for which spectra had been obtained by searching the
+\texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after excluding stars
+with the stellar parameter bit flags 1, 2, 3.  This excludes spectra
+with unreliable broadening, low S/N, and unreliable wavelength
+solutions (see Table~6 of \citealt{buder_galah_2020}).
+% Since our main focus is measuring equivalent widths for the
+% Li\,6708\,\AA\ doublet, these were the most relevant bit flags.
+Of our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH
+DR3.  51 were in the core, and 56 were in the ``halo''.  We
+downloaded\footnote{Via \url{datacentral.org.au/services/download},
 using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
@@
 We measured the lithium equivalent widths using the \texttt{specutils}
-package \citep{specutils_v1pt1}.  We did this by first focusing in on
-a 15\AA\ window centered on the 6707.835\AA\ lithium doublet.  We then
+package \citep{specutils_v1pt1}.  We did this by first shifting the
+spectra to the stellar rest frames, and then focused in on a 15\AA\
+window centered on the 6707.835\AA\ lithium doublet.  We then
 continuum normalized the spectra using a third-degree Chebyshev
 series, while excluding any regions that showed absorption lines ({\it
-e.g.}, Fe\,\textsc{I} is present at 6703.58\,\AA and 6705.10\,\AA).
+e.g.}, Fe\,\textsc{I} is present at 6703.58\,\AA\ and 6705.10\,\AA).
 We proceeding by fitting a Gaussian to the continuum normalized
@@
 integrating the fitted model over the same window.  Our approach
-therefore includes the Fe\,6707.44 blend in the reported Li equivalent
+therefore includes the Fe\,6707.44\AA\ blend in the reported Li equivalent
 widths---which leads to systematic overestimates of $\approx$10 to
@@
 $84^{\rm th}$ and $16^{\rm th}$ percentiles of the resulting EW
-distribution.  We verified the overall scale of our results by
-comparing our measurements with those of \citet{randich_gaiaeso_2018}.
-Most stars with $T_{\rm eff} > 4500\,{\rm K}$ followed a 1-1 relation.
-For cooler stars, our EW measurements are systematically lower, as
+distribution, and are included in Table~\ref{tab:maintable}.  We
+verified the overall scale of our results by comparing our
+measurements with those of \citet{randich_gaiaeso_2018}.  Most stars
+with $T_{\rm eff} > 4500\,{\rm K}$ followed a 1-1 relation.  For
+cooler stars, our EW measurements are systematically lower, as
 expected given the depletion timescales for stars at these ages
@@
 absolute magnitudes, with the corresponding
-\citet{pecaut_mamajek_2013} mass estimates.  The stars closest
-to the core do
-indeed seem to have a greater proportion of the $\gtrsim$0.8$M_\odot$
-stars, while the tail stars have a greater proportion of
-$\lesssim$0.4$M_\odot$ stars.  The mean star in the inner separation bin has
-$({\rm M}_G)_0=7.08$ ($\approx$0.69$M_\odot$), while in the outer bin
-the mean star has $({\rm M}_G)_0=7.61$ ($\approx$0.63$M_\odot$).
-While the difference does appear significant given the number of
-stars involved, the theoretical implications are not entirely clear,
-since the ``molecular filament'' explanation might be able to
-accommodate mass differences between the core and tail through density
-gradients across the initial filament.
+\citet{pecaut_mamajek_2013} mass estimates.  The stars closest to the
+cluster center do seem to have a greater proportion of $\gtrsim$0.8$M_\odot$
+stars, while the outermost stars have a greater proportion of
+$\lesssim$0.4$M_\odot$ stars.  The mean star in the inner separation
+bin has $({\rm M}_G)_0=7.08$ ($\approx$0.69$M_\odot$), while in the
+outer bin the mean star has $({\rm M}_G)_0=7.61$
+($\approx$0.63$M_\odot$).  While the difference does appear
+significant given the number of stars involved, the theoretical
+implications are not entirely clear, since the ``molecular filament''
+explanation might be able to accommodate mass differences between the
+core and tail through density gradients across the initial filament.
+Observational incompleteness at the low-mass end seems unlikely to
+alter the result though, since the innermost and outermost are likely
+subject to similar selection effects.
 
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Absolute magnitudes (and masses) for the inner and
-			outermost stars in \cn.}
-		The \nkinematic\ candidate \cn\ members were split into three groups
-		according to their distance from the cluster center.
-		The inner and outermost groups, each containing 1,099 stars, are shown to
-		provide the clearest picture with respect to expectations for mass
-		segregation.
-		The mean halo star does have lower mass, consistent with expectations
-		for tidal tails.
+  \caption{ {\bf Absolute magnitudes and masses for the inner and
+  outermost stars in \cn.} The \nkinematic\ candidate \cn\ members
+  were split into three groups according to their distance from the
+  cluster center.  The inner and outermost groups, each containing
+  1,099 stars, are shown in 1\,mag bins.  The mean star in the
+  innermost sample has a 10\% lower mass than the mean star in the
+  outermost sample, consistent with expectations for tidal tails.
   \label{fig:massfunction}
@@
    %
-   \item {\it The average star in the halo has lower mass (by $\approx$10\%)
-   than the average star in the core.}
+   \item {\it The average star in the outskirts has lower mass (by $\approx$10\%)
+   than the average star in the cluster center.}
    This is shown in Figure~\ref{fig:massfunction}, and discussed in
@@
 
-The existence of the halo itself is physically plausible.  A
+The existence of the halo itself is not particularly surprising.  A
 star with a 1\kms\ velocity difference from the cluster average will
@@
 background of the Galaxy throughout centuries of modern astronomy.
-The consequences that these halo stars will have in expanding our
-understanding of planetary and stellar origins are only just beginning
-to be understood.
+Their discovery likely represents a major step toward understanding
+open cluster dispersal.
+Combined with wide-field photometric surveys, the stars themselves
+will enable advances in understanding stellar
+angular momentum evolution.
+Combined with wide-field spectroscopic
+surveys \citep[{\it e.g.},][]{kollmeier_2017}, the halos could also lead to
+advances in understanding the chemical evolution of low-mass stars.
+Finally, the halo stars also provide an expanded sample of
+age-dated stars to search for planets \citep[{\it
+e.g.},][]{newton_2021}, which could help expand our understanding of
+planetary evolution.
 

### 2021-04-26 drivers/write_NGC2516_main_table.py

@@ -9,6 +9,7 @@ populates results.tables with the table containing:
     * Lithium information.
 """
 import os
+from numpy import array as nparr
 import numpy as np, pandas as pd
 from earhart.paths import DATADIR, RESULTSDIR
 from earhart.helpers import (
@@ -133,6 +134,39 @@ df = df.rename(columns={
 
 df = df.sort_values(by=['in_SetB'], ascending=False)
 
+#
+# clean up the uncertainties on Li EWs
+# minimum 10mA uncertainty everywhere, 20mA for BpmRp0>1.8
+#
+sel = (
+    (~pd.isnull(df.Li_EW_mA_GaiaESO))
+    |
+    (~pd.isnull(df.Li_EW_mA_GALAH))
+)
+df.loc[sel,'Li_EW_mA_perr_GaiaESO'] = np.maximum(nparr(df[sel]['Li_EW_mA_perr_GaiaESO']),5)
+df.loc[sel,'Li_EW_mA_merr_GaiaESO'] = np.maximum(nparr(df[sel]['Li_EW_mA_merr_GaiaESO']),5)
+df.loc[sel,'Li_EW_mA_perr_GALAH'] = np.maximum(nparr(df[sel]['Li_EW_mA_perr_GALAH']),5)
+df.loc[sel,'Li_EW_mA_merr_GALAH'] = np.maximum(nparr(df[sel]['Li_EW_mA_merr_GALAH']),5)
+
+sel = (
+    ((~pd.isnull(df.Li_EW_mA_GaiaESO))
+    |
+    (~pd.isnull(df.Li_EW_mA_GALAH))
+    ) &
+    (df["(Bp-Rp)_0"] > 1.8)
+)
+df.loc[sel,'Li_EW_mA_perr_GaiaESO'] = np.maximum(nparr(df[sel]['Li_EW_mA_perr_GaiaESO']),20)
+df.loc[sel,'Li_EW_mA_merr_GaiaESO'] = np.maximum(nparr(df[sel]['Li_EW_mA_merr_GaiaESO']),20)
+df.loc[sel,'Li_EW_mA_perr_GALAH'] = np.maximum(nparr(df[sel]['Li_EW_mA_perr_GALAH']),20)
+df.loc[sel,'Li_EW_mA_merr_GALAH'] = np.maximum(nparr(df[sel]['Li_EW_mA_merr_GALAH']),20)
+
+#
+# column formatting. nullable integer is Int32Dtype.
[…]

### 2021-04-26 earhart/plotting.py

@@ -3798,7 +3798,7 @@ def plot_lumfunction_vs_position(outdir):
 
             l = "$\Delta r_{\mathrm{3D}}$:" + f'{bin_start:.1f} - {bin_end:.1f} pc'
             bbox = dict(facecolor='white', alpha=1, pad=0, edgecolor='white')
-            ax.text(0.04, 0.96-i*0.04, l, va='top', ha='left',
+            ax.text(0.04, 0.95-i*0.035, l, va='top', ha='left',
                     transform=ax.transAxes, color=colors[i], bbox=bbox,
                     fontsize='small')
 

### 2021-04-26 cluster_paper/maintable_header.tex
update ms

@@
       \texttt{Li\_EW\_mA\_GaiaESO} &              -1.495 &                          Gaia-ESO Li doublet equivalent width, including the Fe blend [m\AA]. \\
-\texttt{Li\_EW\_mA\_perr\_GaiaESO} &              12.407 &                                     Gaia-ESO Li doublet EW upper uncertainty interval [m\AA]. \\
-\texttt{Li\_EW\_mA\_merr\_GaiaESO} &                 5.0 &                                     Gaia-ESO Li doublet EW lower uncertainty interval [m\AA]. \\
+\texttt{Li\_EW\_mA\_perr\_GaiaESO} &              12.407 &                                              Gaia-ESO Li doublet EW upper uncertainty [m\AA]. \\
+\texttt{Li\_EW\_mA\_merr\_GaiaESO} &                 5.0 &                                              Gaia-ESO Li doublet EW lower uncertainty [m\AA]. \\
         \texttt{Li\_EW\_mA\_GALAH} &                 NaN &                             GALAH Li doublet equivalent width, including the Fe blend [m\AA]. \\
-  \texttt{Li\_EW\_mA\_perr\_GALAH} &                 NaN &                                        GALAH Li doublet EW upper uncertainty interval [m\AA]. \\
-  \texttt{Li\_EW\_mA\_merr\_GALAH} &                 NaN &                                        GALAH Li doublet EW lower uncertainty interval [m\AA]. \\
+  \texttt{Li\_EW\_mA\_perr\_GALAH} &                 NaN &                                                 GALAH Li doublet EW upper uncertainty [m\AA]. \\
+  \texttt{Li\_EW\_mA\_merr\_GALAH} &                 NaN &                                                 GALAH Li doublet EW lower uncertainty [m\AA]. \\
 \enddata

### 2021-04-26 cluster_paper/ms.tex
update ms

@@
   this work supports a new paradigm wherein the halos of open clusters
-  are often more populous than their cores.  We discuss implications
+  are often more populous than their cores.  We highlight implications
   for spectroscopic survey targeting, open cluster dispersal, and
@@
   open clusters (1160),
+  stellar rotation (1629),
 	stellar ages (1581),
-  stellar kinematics (1608),
-  stellar rotation (1629)
+  stellar kinematics (1608)
 }
@@
 reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
-the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and in Coma
+the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and Coma
 Berenices \citep{tang_comaber_2019,furnkranz_2019}.
@@
 samples is important to verify through independent lines of evidence,
-such as rotation periods and spectroscopic indicators.
+including rotation periods and spectroscopy.
 
@@
 ``outskirts'', and ``tidal tails'' \citep[{\it
-e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
+e.g.},][]{chumak_tails_2006,davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
 The latter term implies a particular model for the formation of the
-dispersed group.  ``Halo'' is concise and model agnostic, but is not
+dispersed group.  ``Halo'' is model agnostic, but is not
 an ideal term because it connotes spherical symmetry, which is rarely
@@
 (Section~\ref{subsec:hr}), TESS gyrochronology
-(Section~\ref{subsec:tess}), and relative lithium abundances
+(Section~\ref{subsec:tess}), and lithium equivalent widths
 (Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
@@
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f11.pdf}
+		\includegraphics[width=0.99\textwidth]{f11.pdf}
 	\end{center}
@@
   \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.}
-  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
-  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
-  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
-  cluster orbit.  In different projections, the gray line spans
-  different time intervals: the longest time window is visible in the
-  $(Z,X)$ projection.  The large blue arrows denote the median cluster
+  Points are halo members with $\pi/\sigma_\pi>20$; the cluster is
+  most elongated viewed top-down ({\it left}). 
+  The Sun is at $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
+  Gray lines trace the past (solid) and future (dashed)
+  cluster orbit, and blue arrows denote the mean cluster
   velocity after subtracting the local standard of rest: $\{v_{\rm X},
-  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By
-  comparison, $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the
-  inertial frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
-  direction.  The sizes of the blue arrows are internally
-  proportionate. 
+  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms. 
   \label{fig:XYZ}
@@
 
-An initial step in visualizing the reported structure and kinematics
+An initial step in visualizing the structure and kinematics
 of the candidate cluster members is to consider their Cartesian
@@
 below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version 2021.03.02.} Stars in the core appear consistent with having a
+version \texttt{2021/03/02}.} Stars in the core appear consistent with having a
 fixed age and metallicity, and varying mass.  The halo stars show a
@@
 We used the web
-interface,\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
-\texttt{2021-02-26} \texttt{CMD3.4}, \texttt{YBC} bolometric
+interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
+\texttt{2021/02/26} \texttt{CMD3.4}, \texttt{YBC} bolometric
 corrections as in \citet{chen_2019_YBC}.} to interpolate these
@@
 comparable disagreement \citep{choi_mesa_2016}.  Proposed explanations
-for the discrepancy between the models and observations have include
-starspots and incomplete molecular line lists\citep[{\it
+for the discrepancy between the models and observations include
+starspots and incomplete molecular line lists \citep[{\it
 e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
@@
 
-\subsubsection{Methodology}
-\label{subsubsec:cluster}
-
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\subfloat{
-% 			\includegraphics[width=0.49\textwidth]{f3a.pdf}
-% 			\includegraphics[width=0.49\textwidth]{f3b.pdf}
-% 		}
-% 
-%         \vspace{-0.5cm}
-% 		\subfloat{
-% 			\includegraphics[width=0.49\textwidth]{f3c.pdf}
-% 			\includegraphics[width=0.49\textwidth]{f3d.pdf}
-% 		}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-%   \caption{ {\bf TESS rotations periods versus color for \cn.} {\it
-%   Top row}: The core and halo of \cn.  Sets $\mathcal{A}$ ({\it left})
-%   and $\mathcal{B}$ ({\it right}) undergo successive stages of
-%   cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
-%   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
-%   \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
-%   Stars in the core and the halo of NGC\,2516 overlap with the
-%   Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
-%   members have longer rotation periods than Pleiades members.  {\it
-%   Bottom row}: Binarity indicators in \cn.  Astrometric binaries
-%   ($\mathrm{RUWE}>1.2$; 11\% of stars) appear in red; photometric
-%   binaries ($>$0.3\,mag from an empirical isochrone; 20\% of stars)
-%   are overplotted in orange.  Rotation periods for comparison field
-%   stars are shown in Appendix~\ref{app:compstar}.
-%   \label{fig:rot}
-% 	}
-% \end{figure*}
-
-\begin{figure*}[t]
+\begin{figure*}[t!]
 	\begin{center}
@@
 		\subfloat{
-			\includegraphics[width=0.72\textwidth]{f3a.pdf}
+			\includegraphics[width=0.700\textwidth]{f3a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.72\textwidth]{f3b.pdf}
+			\includegraphics[width=0.700\textwidth]{f3b.pdf}
 		}
@@
   members have longer rotation periods than Pleiades members. Rotation
-  periods for comparison field stars are shown in
+  periods for a comparison sample of field stars are shown in
   Appendix~\ref{app:compstar}.
@@
 \end{figure*}
-\begin{figure*}[t]
+\begin{figure*}[t!]
 	\begin{center}
@@
 		\subfloat{
-			\includegraphics[width=0.72\textwidth]{f3c.pdf}
+			\includegraphics[width=0.700\textwidth]{f3c.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.72\textwidth]{f3d.pdf}
+			\includegraphics[width=0.700\textwidth]{f3d.pdf}
 		}
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf Binarity indicators for the rotators in
-  Figure~\ref{fig:rot}.} Sets $\mathcal{A}$ ({\it top}) and
-  $\mathcal{B}$ ({\it bottom}) undergo the same successive stages of
-  cleaning.  Astrometric binaries ($\mathrm{RUWE}>1.2$; 11\% of stars)
-  appear in red; photometric binaries ($>$0.3\,mag from an empirical
-  isochrone; 20\% of stars) are overplotted in orange.  
+  \caption{ {\bf Binarity indicators for rotators in \cn.} Sets
+  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are as in
+  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
+  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
+  above an empirical isochrone; 20\% of stars) are overplotted in
+  orange.  
+  Over $0.5<\bpmrpo<1.3$, set $\mathcal{B}$ has 281 stars.
+  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
+  the fraction of stars showing signs of binarity in the slow and fast
+  subsamples are 30\% (63/210) and 56\% (40/71) respectively.
   \label{fig:binarity}
@@
 \end{figure*}
+
+\subsubsection{Methodology}
+\label{subsubsec:cluster}
 
@@
 $G_\mathrm{RP}=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
-source list had $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
+source list have $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
 from \citet{meingast_2021} which were not available at the time of the
@@
 regardless of detection significance, in Table~\ref{tab:maintable}.
-To clean the resulting rotation period measurements, we designed a set
-of automated cleaning criteria, which yielded sets we denote
+To clean these measurements, we designed 
+automated cleaning criteria, which yielded sets we denote
 $\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered light
@@
 Figure~\ref{fig:binarity} shows the same data, with the
-points colored with indicators of binarity.  To assess astrometric
+points colored by indicators of binarity.  To assess astrometric
 binarity, we used the renormalized unit weight error (RUWE) from Gaia
@@
   left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
-  2D-tangential velocities for kinematic halo members
-  (Set~$\mathcal{A}$).  Stars with detected rotation periods are shown
-  in blue;  halo members for which rotation periods should have been
-  detectable, but were not detected, are shown in orange.  Stars in
-  the core are shown as black points.  Computing the 2D-tangential
-  velocity requires correcting for projection effects
-  (Appendix~\ref{app:vproj}).  In the {\it top right}, the histograms
-  bin versus 3D separation from the core and 2D tangential velocity
-  difference.  Bin widths are chosen to enforce the same number of
-  stars in the denominator ($N=54$) per bin.  In the {\it bottom
-  panels}, the innermost and outermost 100 stars in position and
-  velocity space (black and green) are shown in the space of rotation
-  periods and dereddened colors.
+  2D-tangential velocities for halo members.  Stars with detected
+  rotation periods in Set~$\mathcal{A}$ are shown in blue;  halo
+  members for which rotation periods should have been detectable, but
+  were not detected, are shown in orange.  Stars in the core are shown
+  as black points.  In
+  the {\it top right}, the 
+  ratio of detected to expected stars with rotation period detections is shown
+  versus 3D separation from
+  the core and 2D tangential velocity difference.  Bin widths are
+  chosen to enforce the same number of stars in the denominator
+  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
+  outermost 100 stars with detection rotation periods in position and
+  velocity space.
   \label{fig:physical_x_rotn}
@@
 cluster core.  The average detection rate inside of 25\,pc, 73\%, is
-slightly higher.  With $\approx$3.5\,pc, the period detection rate of
-61\% is relatively low, due to the combined effects of crowding and
-saturation.
+slightly higher.  Within $\approx$3.5\,pc, the period detection rate of
+61\% is relatively low, due to crowding and saturation in the TESS
+images.
 
@@
 periods and did not, this suggests field contaminant rates of
-$\approx$40-50\% for the outermost cluster members in our target list.
+$\approx$50\% for the outermost cluster members in our target list.
 
@@
 This agrees with the overall structure of the halo reported by
-\citetalias{kounkel_untangling_2019}, with some minor caveats visible
-in the top-left panels of Figure~\ref{fig:physical_x_rotn}.
+\citetalias{kounkel_untangling_2019}, with some minor caveats (orange
+circles) visible in the top-left panels of
+Figure~\ref{fig:physical_x_rotn}.
 
@@
 	\end{center}
-	\vspace{-0.72cm}
+	\vspace{-0.700cm}
   \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
   Equivalent width of the 6708\,\AA\ doublet is shown versus
-  dereddened color for all kinematic \cn\ members with Gaia-ESO
+  dereddened color for all candidate \cn\ members with Gaia-ESO
   ($N=459$) or GALAH ($N=107$) spectra available.  The GALAH spectra
   comprise slightly over half of the halo stars, due to the
-  non-targeted selection function of that survey.  Field stars (gray
-  points) are stars with Kepler planets
-  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
+  non-targeted selection function of that survey.  Gray
+  points are field stars with Kepler planets
+  \citep{berger_identifying_2018}.  Pleiades stars (125\,Myr)
   with Li detections are from \citet{bouvier_pleiades_lirot_2018};
@@
   \citep{soderblom_praesepe_1993}.  Points with ${\rm
-  EW}\approx0\,$m\AA\ are non-detections; all stars shown
+  EW}\approx0\,$m\AA\ are non-detections; all stars 
   later than $\approx$M0 do not have significant lithium.
@@
   \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
-  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
-  GALAH surveys, and are colored by equivalent width.  Points with
-  reported equivalent widths below 20\,m\AA\ (including
-  non-detections) are shown with gray crosses.  The lithium-rotation
-  correlation for the K-dwarfs has been observed in other young
-  clusters, and is discussed in Section~\ref{subsec:lithium}.
+  candidate \cn\ members with TESS rotation periods and Gaia-ESO or
+  GALAH spectra, and are colored by equivalent width.  Points with
+  reported equivalent widths below 20\,m\AA\ are shown with gray
+  crosses.  The lithium-rotation correlation for the K-dwarfs has been
+  observed in other young clusters, and is discussed in
+  Section~\ref{discussion:lithium}.
 	\label{fig:lithiumrot}
@@
 
-The third and final approach we took for confirming the youth of the
+The final approach we took for assessing the youth of the
 halo population of \cn\ was an analysis of the Li\,\textsc{I}
@@
 showed the surprising feature that stars with $T_{\rm eff}\lesssim
-4000\,{\rm K}$ tend to have non-zero equivalent widths, despite the
+4000\,{\rm K}$ often had non-zero equivalent widths, despite the
 expectation that their lithium should be fully depleted.  Whether
@@
 
-The GALAH DR3 target selection is discussed in depth by
+The GALAH DR3 target selection is discussed by
 \citet{buder_galah_2020}.  The relevant aspects for our analysis are
-that the ``main'' survey targeted $12<V<14$ stars at
+that the survey targeted $12<V<14$ stars at
 $\delta<+10^\circ$ in select fields, provided the stars were at least
-ten degrees from the galactic plane.  Special targeting for stars in
-the TESS southern continuous viewing zone, and for known open cluster
-members was also performed.  We identified the candidate \cn\ members
+ten degrees from the galactic plane.  We identified the candidate \cn\ members
 for which spectra had been obtained by searching the
@@
 solutions (see Table~6 of \citealt{buder_galah_2020}).
-% Since our main focus is measuring equivalent widths for the
-% Li\,6708\,\AA\ doublet, these were the most relevant bit flags.
 Of our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH
-DR3.  51 were in the core, and 56 were in the ``halo''.  We
+DR3.  51 were in the core, and 56 were in the halo.  We
 downloaded\footnote{Via \url{datacentral.org.au/services/download},
@@
 package \citep{specutils_v1pt1}.  We did this by first shifting the
-spectra to the stellar rest frames, and then focused in on a 15\AA\
-window centered on the 6707.835\AA\ lithium doublet.  We then
-continuum normalized the spectra using a third-degree Chebyshev
-series, while excluding any regions that showed absorption lines ({\it
+spectra to the stellar rest frame, and then focused in on a 15\AA\
+window centered on the 6707.835\AA\ lithium doublet.  We 
+continuum normalized the spectra with a third-degree Chebyshev
+series, excluding any regions that showed absorption lines ({\it
 e.g.}, Fe\,\textsc{I} is present at 6703.58\,\AA\ and 6705.10\,\AA).
-We proceeding by fitting a Gaussian to the continuum normalized
+We proceeded by fitting a Gaussian to the continuum normalized
 spectra, considering only a $\pm 1\,$\AA\ window centered on the Li
@@
 therefore includes the Fe\,6707.44\AA\ blend in the reported Li equivalent
-widths---which leads to systematic overestimates of $\approx$10 to
+widths---which leads to systematic overestimates of 10 to
 15\,m\AA\ \citep[{\it e.g.},][]{bouvier_pleiades_lirot_2018}.
@@
 $84^{\rm th}$ and $16^{\rm th}$ percentiles of the resulting EW
-distribution, and are included in Table~\ref{tab:maintable}.  We
-verified the overall scale of our results by comparing our
-measurements with those of \citet{randich_gaiaeso_2018}.  Most stars
-with $T_{\rm eff} > 4500\,{\rm K}$ followed a 1-1 relation.  For
-cooler stars, our EW measurements are systematically lower, as
-expected given the depletion timescales for stars at these ages
-\citep[{\it e.g.},][]{soderblom_ages_2014}.  We visually verified from
-the spectra that, for the most part, we prefer our measurements.
+distribution, with minima imposed at 5\,m\AA\ for FGK stars, and
+20\,m\AA\ for M dwarfs. They are included in
+Table~\ref{tab:maintable}.  We verified the overall scale of our
+results by comparing our measurements with those of
+\citet{randich_gaiaeso_2018}.  Most stars with $T_{\rm eff} >
+4500\,{\rm K}$ followed a 1-1 relation.  For cooler stars, our EW
+measurements are systematically lower, as expected given the depletion
+timescales for stars at these ages \citep[{\it
+e.g.},][]{soderblom_ages_2014}.  We visually verified from the spectra
+that, for the most part, we prefer our measurements.
 
@@
 (we interpolated from their effective temperatures using the
-\citet{pecaut_mamajek_2013} table).  For clarity, we have omitted the
+\citealt{pecaut_mamajek_2013} table).  For clarity, we have omitted the
 few upper limits reported by \citet{bouvier_pleiades_lirot_2018}.  The
@@
 have formed in the same ``clump'' as those in the cluster core.  In
-other words, the formation environment might be best described as a
+other words, the formation environment might be a
 giant molecular filament, rather than a giant molecular cloud.  A
@@
 could be informative, since the tidal tail explanation predicts that
-the mean tail star is expected to be less massive than the mean core
+the mean tail star should be less massive than the mean core
 star \citep[{\it e.g.},][]{chumak_tails_2006}.
@@
 Observational incompleteness at the low-mass end seems unlikely to
-alter the result though, since the innermost and outermost are likely
+alter the result, since the innermost and outermost regions are 
 subject to similar selection effects.
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf Absolute magnitudes and masses for the inner and
+  \caption{ {\bf Masses and absolute magnitudes for the innermost and
   outermost stars in \cn.} The \nkinematic\ candidate \cn\ members
@@
   cluster center.  The inner and outermost groups, each containing
-  1,099 stars, are shown in 1\,mag bins.  The mean star in the
+  1,099 stars, are counted in 1\,mag bins.  The mean star in the
   innermost sample has a 10\% lower mass than the mean star in the
@@
 scope of this work, we emphasize that the main contribution of our
-study beyond these previous analyses our sensitivity
+study beyond these previous analyses is our sensitivity
 across both the core and halo of \cn, and a
@@
 features in the rotation-color diagram, and whether our expanded
-rotation sample can provide any new insights.
+sample can provide any new insights.
 
@@
 question of why the rapid rotators exist, for instance, requires
-additional physics.  Models that introduce variable
+additional physics.  Models that vary
 core-envelope decoupling timescales and pre-MS disk lifetimes can
@@
 significant fraction of parameter space.  A 0.3\,mag flux excess, for
-instance, corresponds to mass ratios $M_2/M_1 \gtrsim 0.72$, assuming
+instance, corresponds to mass ratios $M_2/M_1 \gtrsim 0.700$, assuming
 the luminosity $L$ scales as $M^{3.5}$.
@@
 Comparing the explanations of disk-locking against tidal
-synchronization, the population statistics seem to rule out the
-possibility of tidal synchronization.  A quarter of the \cn\ members
+synchronization, the population statistics seem to rule out 
+tidal synchronization.  A quarter of the \cn\ members
 are rapidly rotating.  In comparison, in the field half of Sun-like
@@
 over beyond just equivalent widths, and to the actual abundance of
-lithium in the photosphere, A(Li).  To our knowledge, the first
+lithium in the photosphere.  To our knowledge, the first
 demonstration of this was by \citet{soderblom_evolution_1993}, who
-calculated the relevant LTE curves of growth for the Li\,6708\,\AA\
-doublet, and used them to convert equivalent widths to abundances.
+calculated the LTE curve of growth for the Li\,6708\,\AA\
+doublet, and used it to convert equivalent widths to abundances.
 Non-LTE corrections do
@@
 based on pure hydrodynamics is that the rapid rotation leads to less
-efficient convective penetration (``overshoot''), which then lowers the mixing
+efficient convective penetration (``overshoot''), which lowers the mixing
 efficiency at the convective-radiative boundary
@@
 greater amount of differential rotation generated in the radiative
-zone, more efficient shear mixing, and ultimately a lower photospheric
+zone, more efficient shear mixing, and consequently a lower photospheric
 lithium abundance.  This provides a plausible explanation for the
@@
 Fitting a model to substantiate the claim that \cn\ is
-gyrochronologically older than the Pleiades \citep[{\it e.g.}, any
-of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
+gyrochronologically older than the Pleiades \citep[{\it e.g.},][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
 is, unfortunately, an exercise in tautology.  Gyrochronological models
@@
 because lithium depletion timescales are long for Sun-like stars past
-100\,Myr.  For instance, \citet{jeffries_m35_li_2020} has shown in a
-comparison of the Pleiades and M\,35 ($\approx$150\,Myr) that the two
-are indistinguishable given current measurement, despite the age of
-the latter being expected to be marginally older than the former.
-\cn\ members are on average more depleted than {\it e.g.}, stars in
-IC\,2602, but only marginally so \citep{soderblom_ages_2014}.  On the
+100\,Myr.  For example,
+\cn\ FGK stars are on average more depleted than stars in
+IC\,2602 ($\approx40\,{\rm Myr}$), but only marginally so \citep{soderblom_ages_2014}.  On the
 older end though, Figure~\ref{fig:lithiumcorehalo} does show that
@@
   \item {\it Over-densities observed in the Gaia 3D positions and 2D
-    velocities revealed a halo of stars spanning $\approx$500 parsecs
+    velocities revealed a halo of stars spanning $\approx$500\,pc
     in the plane of the galaxy.} The earliest reference to this halo
@@
     halo of \cn\ is coeval with its core.} In short, all the data are
-    consistent with the halo being real.  The Gaia EDR3 photometry and
-    astrometry show a main-sequence turnoff that suggest an isochronal
+    consistent with the halo being real.  The Gaia EDR3 data
+    show a main-sequence turnoff that suggest an isochronal
     age of 150\,Myr for both the core and halo (Figure~\ref{fig:hr}).
     The faintest M dwarfs in the core and halo are also brighter than
-    comparison field stars of the same color.  Gyrochronally, stars in
+    field stars of the same color.  Gyrochronally, stars in
     \cn\ spanning $0.5<\bpmrpo<1.2$ (spectral types F2V-K3V) overlap
     with the slow sequence of the Pleiades (Figure~\ref{fig:rot}).  At
-    redder colors, from $1.2<\bpmrpo<1.7$ (spectral types K3V-K6V,
-    $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
+    redder colors, from $1.2<\bpmrpo<1.7$ (spectral types K3V-K6V), there is a larger
     dispersion in rotation period.  The upper envelope of the rotation
@@
     right and lower panels of Figure~\ref{fig:physical_x_rotn}
-    quantify the statement: $\approx$70\% of the inner-most
-    members have detected rotation periods consistent with a
+    quantify the statement: $\approx$70\% of the innermost
+    candidate members have detected rotation periods consistent with a
     gyrochronological age near that of the Pleiades. At physical
@@
     $\approx$50\%.  Given the number density ratios of cluster to
-    field stars at these wide separations ($\approx 10^{-3}$,
+    field stars at the widest separations ($\approx 10^{-3}$,
     \citealt{meingast_2021}), these contamination fractions are not as
@@
    Section~\ref{subsec:origin}.
-   While the mass function is consistent with expectations for a
+   While this is consistent with expectations for a
    ``tidal tail'' origin of the halo, the possibility of a primordial halo
@@
    %
-   \item {\it The rapid rotators show elevated lithium abundances, and
+   \item {\it The rapidly rotating K dwarfs show elevated lithium abundances, and
      elevated binarity fractions}.  Each trend has been noted in
@@
      factor of $\gtrsim2$ enhancement in binarity fraction for fast
-     {\it vs{.}} slowly rotating G and early K dwarfs.  We have pointed
-     out in Section~\ref{discussion:lithium} that one plausible
-     connection between the two observations is that the disk-locking
+     {\it vs{.}} slow rotators.  
+     Section~\ref{discussion:lithium} discusses one plausible
+     connection between the two observations: the disk-locking
      phase may be truncated in binary systems, which could free the
      stars to spin up as they contract on the pre-MS.  This spin up
-     would be expected to produce less efficient internal mixing and a
+     could result in less efficient internal mixing and a
      longer-lasting presence of photospheric lithium
@@
 Their discovery likely represents a major step toward understanding
-open cluster dispersal.
+the dispersal of open clusters.
 Combined with wide-field photometric surveys, the stars themselves
-will enable advances in understanding stellar
+enable advances in studies of stellar
 angular momentum evolution.
 Combined with wide-field spectroscopic
-surveys \citep[{\it e.g.},][]{kollmeier_2017}, the halos could also lead to
-advances in understanding the chemical evolution of low-mass stars.
-Finally, the halo stars also provide an expanded sample of
+surveys \citep[{\it e.g.},][]{kollmeier_2017}, the halos could also
+improve our
+understanding of the chemical evolution of low-mass stars.
+Finally, the halo stars provide an expanded sample of
 age-dated stars to search for planets \citep[{\it
-e.g.},][]{newton_2021}, which could help expand our understanding of
+e.g.},][]{newton_2021}, which may help in benchmarking our understanding of
 planetary evolution.
@@
 This study was based in part on observations at Cerro Tololo
-Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
+Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
 2020A-0146; 2020B-0029 PI: L{.}~Bouma), which is managed by the
@@
 %
-This paper includes data collected by the TESS mission, which are
+This paper also includes data collected by the TESS mission, which are
 publicly available from the Mikulski Archive for Space Telescopes

### 2021-04-26 drivers/write_NGC2516_main_table.py

@@ -221,11 +221,11 @@ namedesc_dict = {
 'is_astrm_bin': "True if Gaia EDR3 RUWE > 1.2.",
 'ruwe': "Gaia EDR3 RUWE.",
 'Li_EW_mA_GaiaESO': "Gaia-ESO Li doublet equivalent width, including the Fe blend [m\\AA].",
-'Li_EW_mA_perr_GaiaESO': "Gaia-ESO Li doublet EW upper uncertainty interval [m\\AA].",
-'Li_EW_mA_merr_GaiaESO': "Gaia-ESO Li doublet EW lower uncertainty interval [m\\AA].",
+'Li_EW_mA_perr_GaiaESO': "Gaia-ESO Li doublet EW upper uncertainty [m\\AA].",
+'Li_EW_mA_merr_GaiaESO': "Gaia-ESO Li doublet EW lower uncertainty [m\\AA].",
 'Li_EW_mA_GALAH': "GALAH Li doublet equivalent width, including the Fe blend [m\\AA].",
-'Li_EW_mA_perr_GALAH': "GALAH Li doublet EW upper uncertainty interval [m\\AA].",
-'Li_EW_mA_merr_GALAH': "GALAH Li doublet EW lower uncertainty interval [m\\AA]."
+'Li_EW_mA_perr_GALAH': "GALAH Li doublet EW upper uncertainty [m\\AA].",
+'Li_EW_mA_merr_GALAH': "GALAH Li doublet EW lower uncertainty [m\\AA]."
 }
 
 keys = list(namedesc_dict.keys())

### 2021-04-26 earhart/plotting.py

@@ -219,18 +219,19 @@ def plot_full_kinematics(outdir, basedata='bright', show1937=1,
             yv = params[i+1]
             print(i,j,xv,yv)
 
+            alpha = 0.9
             axs[i,j].scatter(
-                nbhd_df[xv], nbhd_df[yv], c='gray', alpha=0.9, zorder=2, s=5,
+                nbhd_df[xv], nbhd_df[yv], c='gray', alpha=0.7*alpha, zorder=2, s=5,
                 rasterized=True, linewidths=0, label='Field', marker='.'
             )
             axs[i,j].scatter(
-                halo_df[xv], halo_df[yv], c='lightskyblue', alpha=1,
+                halo_df[xv], halo_df[yv], c='lightskyblue', alpha=alpha,
                 zorder=3, s=12, rasterized=True, label='Halo',
                 linewidths=0.1, marker='.', edgecolors='k'
             )
             axs[i,j].scatter(
-                core_df[xv], core_df[yv], c='k', alpha=0.9, zorder=4, s=5,
-                rasterized=True, label='Core', marker='.'
+                core_df[xv], core_df[yv], c='k', alpha=alpha, zorder=4, s=5,
+                rasterized=True, label='Core', marker='.', linewidths=0
             )
             if show1937:
                 axs[i,j].plot(
@@ -3342,17 +3343,26 @@ def plot_vtangential_projection(outdir, basedata='fullfaint'):
     set_style()
     plt.close('all')
     fig, axs = plt.subplots(ncols=3, figsize=(8,3))
-    axs[0].scatter(s_halo_df.x_pc, s_halo_df.y_pc, s=1, c='k', rasterized=True)
+    s = 0.8
+    axs[0].scatter(s_halo_df.x_pc, s_halo_df.y_pc, s=s, c='k', rasterized=True)
     delta_x = 0.1
-    axs[0].arrow(0.75, 0.07, delta_x, 0,
+    axs[0].arrow(0.73, 0.07, delta_x, 0,
                  length_includes_head=True, head_width=1e-2,
                  head_length=1e-2,
                  transform=axs[0].transAxes)
-    axs[0].text(0.75+delta_x/2, 0.08, 'Galactic\ncenter', va='bottom',
[…]

### 2021-04-27 cluster_paper/maintable_header.tex
update ms
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@
 completeness can be achieved using different cuts on photometric
-periods and lithium eqiuvalent widths (see Figure~\ref{fig:rot}).
-Sets $\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
-cleaning based on rotation periods.  Note that when the target star is
-the only star present in the TESS aperture, \texttt{nequal}$=0$,
-\texttt{nclose}$=1$, and \texttt{nfaint}$=1$.  The light curves are
-available at \url{https://archive.stsci.edu/hlsp/cdips}.
+periods, periodogram powers, and lithium eqiuvalent widths.  Sets
+$\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
+cleaning (see Section~\ref{subsubsec:cluster}).  Note that when the
+target star is the only star present in the TESS aperture,
+\texttt{nequal}$=0$, \texttt{nclose}$=1$, and \texttt{nfaint}$=1$.
+The light curves are available at
+\url{https://archive.stsci.edu/hlsp/cdips}.
 }

### 2021-04-27 cluster_paper/ms.tex
update ms
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@
 \citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
-abundance recently determined by the Gaia-ESO team
-\citep{baratella_gaiaeso_2020}, which represented an up-revision from
+abundance determined by the Gaia-ESO team
+\citep{baratella_gaiaeso_2020}. The latter result represented an up-revision from
 an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We
@@
   orange.  
-  Over $0.5<\bpmrpo<1.3$, set $\mathcal{B}$ has 281 stars.
+  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has 281 stars.
   Dividing these stars into ``slow'' and ``fast'' sequences by eye,
@@
 to eliminate low-quality rotation periods while retaining high
-completeness.  For set $\mathcal{A}$, this yielded \nautorotnumerator\
+completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
 light curves, out of \nautorotdenominator\ stars for which rotation
-periods might plausibly have been detected.  For set $\mathcal{B}$, we
+periods might plausibly have been detected.  For Set $\mathcal{B}$, we
 additionally required that the Lomb-Scargle and Stellingwerf
@@
 Figure~\ref{fig:rot} shows the resulting rotation
-periods, with set $\mathcal{A}$ shown on top and $\mathcal{B}$ on
-the bottom.  While set $\mathcal{A}$ is more complete than set
+periods, with Set $\mathcal{A}$ shown on top and Set $\mathcal{B}$ on
+the bottom.  While Set $\mathcal{A}$ is more complete than Set
 $\mathcal{B}$, this completeness comes at the expense of greater
 contamination.  For instance, stars above the slow sequence (perhaps
-field contaminants) mostly disappear in set $\mathcal{B}$.  The
+field contaminants) mostly disappear in Set $\mathcal{B}$.  The
 ``void'' or ``gap'' underneath the slow sequence is also more
-pronounced in set $\mathcal{B}$.  To facilitate a comparison against
+pronounced in Set $\mathcal{B}$.  To facilitate a comparison against
 the Pleiades and Praesepe, we used the rotation periods and dereddened
@@
 candidate binaries, though not all of them, appear below the slow
-rotation sequence.  In set $\mathcal{B}$, over the color range where
+rotation sequence.  In Set $\mathcal{B}$, over the color range where
 the slow and fast sequence are present ($0.5<\bpmrpo<1.3$), there are
@@
 \citet{soderblom_ages_2010}, and its connection to stellar rotation
-has recently been surveyed by \citet{bouvier_lithium-rotation_2020}.
+has been discussed by \citet{bouvier_lithium-rotation_2020}.
 For \cn, two spectroscopic datasets seemed important:
@@
 giant molecular filament, rather than a giant molecular cloud.  A
-sample of such filaments has recently been collected and analyzed by
+sample of such filaments has been collected and analyzed by
 \citet{zucker_physical_2018}: if the current $\approx$20:1 aspect
@@
 From their rotation period analysis of \cn,
-\citet{fritzewski_rotation_2020} have recently demonstrated that
+\citet{fritzewski_rotation_2020} demonstrated that
 between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri,
@@
 \citet{Irwin_NGC2516_2007} and \citet{fritzewski_rotation_2020}.
-While a detailed comparison of our results with theirs is beyond the
-scope of this work, we emphasize that the main contribution of our
+A detailed comparison of our results with theirs is beyond the
+scope of this work, but we emphasize that the main contribution of our
 study beyond these previous analyses is our sensitivity
@@
 The other features of the rotation-color diagram are indications that
-magnetic braking is not always the most important effect.  The
+magnetic braking is not always the only important effect.  The
 question of why the rapid rotators exist, for instance, requires
-additional physics.  Models that vary
-core-envelope decoupling timescales and pre-MS disk lifetimes can
-reproduce each type of population
-\citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015},
-but the question then shifts to why these timescales should
-differ for stars that are otherwise identical.  Another possible
-explanation invokes a spontaneous and random change of magnetic field
-topology that causes the dynamo to couple more strongly with the
-stellar wind \citep{brown_metastable_2014}.  Other processes could be
-external to the star entirely: \citet{qureshi_signature_2018} for
-instance have reported that mergers between giant planets and the star
-could explain a non-neglible fraction of the rapid rotators.
+additional physics.  Models that vary core-envelope decoupling
+timescales and pre-MS disk lifetimes can reproduce each type of
+population
+\citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015};
+the question then shifts to why these timescales should differ for
+stars that are otherwise identical.  Another possibility, perhaps
+favored, is that proposed by \citet{matt_mass-dependence_2015}: the
+saturation of the magnetic dynamo yields $\dot{P}\sim P$ when $P >
+P_{\rm sat}$, $\dot{P} \sim P^{-1}$ for $P < P_{\rm sat}$, and a
+maximal spin-down rate in the intermediate regime.  This could
+naturally produce the period bimodality observed at ages of
+$\approx$100\,Myr (though it fails to explain the widening of the
+distribution observed at later times
+\citealt{godoyrivera_stellar_2021}).  Other significant processes
+could be external to the star entirely: \citet{qureshi_signature_2018}
+for instance have reported that mergers between giant planets and the
+star could explain a non-neglible fraction of the rapid rotators.
 
@@
 over beyond just equivalent widths, and to the actual abundance of
-lithium in the photosphere.  To our knowledge, the first
-demonstration of this was by \citet{soderblom_evolution_1993}, who
-calculated the LTE curve of growth for the Li\,6708\,\AA\
-doublet, and used it to convert equivalent widths to abundances.
-Non-LTE corrections do
-not change the overall relationships between Li abundance and stellar
-temperature or rotation \citep{carlsson_1994,lind_departures_2009}.
-
-More recent studies of lithium and rotation have expanded these
-results in the Pleiades, the Psc-Eri stream, and M\,35, among other
+lithium in the photosphere \citep{soderblom_evolution_1993}.  Non-LTE
+corrections do not change the overall relationships between Li
+abundance and stellar temperature or rotation
+\citep{carlsson_1994,lind_departures_2009}.  The correlation has been
+seen in the Pleiades, the Psc-Eri stream, and M\,35, among other
 clusters
 \citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020,hawkins_2020}.
-In \cn, \citet{jeffries_rotation_1998} previously observed a similar
-trend that that seen in Figure~\ref{fig:lithiumrot} when analyzing 24
-stars in the core of \cn.  Our concatenation of the
-Gaia-ESO and GALAH-DR3
-spectra represents a significant expanasion in volume and color range
-from this earlier analysis.
+In \cn, \citet{jeffries_rotation_1998} tentatively observed it as
+well, when analyzing 24 stars in the core of \cn.  Our concatenation
+of the Gaia-ESO and GALAH-DR3 spectra represents a significant
+expanasion in volume and color range from this earlier analysis.
 
@@
 dispersion in rotation periods increases.  The maximum rotation
-periods seen in \cn\ across both sets $\mathcal{A}$ and $\mathcal{B}$
+periods seen in \cn\ across both Sets $\mathcal{A}$ and $\mathcal{B}$
 extend up to $\approx$11 days, rather than the $\approx$8.5day upper
@@
 all the studies was the same.  What are the differences in methodology
-that produce these different outcomes?
+that produce these different membership lists?
 
@@
 \citetalias{kounkel_untangling_2019} applied a different unsupervised
-clustering method to the 5D Gaia DR2 positions and proper motions.
+clustering method using the 5D Gaia DR2 positions and proper motions.
 \citetalias{kounkel_untangling_2019} began with a sample of
@@
 within the nearest kiloparsec. One of these clusters was \cn.  Their
-initial  list of $\approx$$3\times10^7$ was similar to that of
+initial  list of $\approx$$3\times10^7$ stars was similar to that of
 \citetalias{kounkel_untangling_2019}, mainly requiring well-behaved
@@
   Appendix~\ref{app:detrending}.  Each light curve is separated by
-  7.5\% in relative flux.  \label{fig:lightcurves}
+  7.5\% in relative flux.
+  Supplementary plots for the light curves each star are available
+  at \url{http://lgbouma.com/notes}.
+  \label{fig:lightcurves}
 	}
@@
 
-In detrending before searching for stellar rotation signals in the
-TESS, our goal was to preserve astrophysical variability, while
-removing systematic variability.  One particular concern for the TESS
-light curves is systematic variability at the timescale of the 14-day
-satellite orbit, mostly induced by scattered light from the Earth and
-Moon.
+In detrending before searching for stellar rotation signals, our goal
+was to preserve astrophysical variability, while removing systematic
+variability.  One particular concern for the TESS light curves is
+systematic variability at the timescale of the 14-day satellite orbit,
+mostly induced by scattered light from the Earth and Moon.
 
@@
 We found that the choice of using ordinary least squares versus ridge
-regression did not seem to significantly affect the resulting light
+regression did not significantly affect the resulting light
 curves. In other words, the inclusion (or lack thereof) of a
 regularization term did not strongly alter the best-fitting
-coefficients.  In the spirit of keeping it simple, we opted to use
+coefficients.  In the spirit of keeping it simple, we opted for
 ordinary least squares.  A few other choices seemed to be more
@@
     %; window length 1 day, cval 6).
-    One issue with this is that systematic sharp features (captured
-    e.g., in ``spike vectors'') are no longer captured, so they are present
-    in the detrended light curves. Since they can be filtered out in postprocessing
+    One issue with this is that systematic sharp features (``spikes'')
+    are no longer captured, and could be present in the detrended light
+    curves. Since they can be filtered out in postprocessing
     ({\it e.g.}, using rolling outlier rejection), we prefer this
@@
 
-After all these considerations, for every ``target star'', we
+After all these considerations, for every target star, we
 ultimately decorrelated the raw (image-subtracted and
@@
 Figure~\ref{fig:lightcurves} shows fifty light curves
-randomly drawn from the resulting ``Set $\mathcal{B}$''.
+randomly drawn from the resulting Set $\mathcal{B}$.
 
@@
 for rotation periods.  This light curve database comprises all $G_\mathrm{RP}<13$
-stars that fell on TESS silicon. The need to distinguish field and
-cluster stars is based on our processing capabilities and is discussed
+stars that fell on TESS silicon. The need for separate field and
+cluster star magnitude limits is based on our processing capabilities and is discussed
 by \citet{bouma_cdipsI_2019}.  Over the southern sky (Sectors 1-13 of
@@
 corresponds to an extinction-corrected color cutoff of $\bpmrpo
-\lesssim 0.80$, or spectral types of $\lesssim$G1V.  This reaches
+\lesssim 0.80$, or spectral types $\lesssim$G1V.  This reaches
 sufficiently far down the slow sequence to enable a comparison against
@@
 different (likely due to the brightness cutoff), the period--color
-distributions are extremely different
+distributions are very different
 (Figure~\ref{fig:compstar}).
@@
 one must consider the position of the star on the sky.  This is
-because two stars sharing identical cartesian velocities will have
-different velocities when projected on the sky, and was emphasized by
+because two stars sharing identical Cartesian velocities will have
+different velocities when projected on the sky, as was emphasized by
 \citet{meingast_2021}.

### 2021-04-27 cluster_paper/ms.tex
add literature comparison
Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f15.pdf}
+		\includegraphics[width=0.5\textwidth]{f15.pdf}
 	\end{center}
@@
 	}
-\end{figure*}
+\end{figure}
 
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f16.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Comparison against previous analyses of stellar
+  rotation in \cn.} 
+  The bottom panel is a concatenation of Set $\mathcal{B}$ with the
+  results of \citet{Irwin_NGC2516_2007} (I07) and
+  \citet{fritzewski_rotation_2020} (F20).
+  \label{fig:litcomp}
+	}
+\end{figure*}
+
 From their rotation period analysis of \cn,
-\citet{fritzewski_rotation_2020} demonstrated that
-between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri,
-M\,35, M\,50, and the Pleiades) all show similar rotation period
-distributions.  The FGK stars show a ``slow sequence'', a population
-of ``rapid rotators'', and in some cases stars in the ``void'' between
-the two populations ({\it e.g.,}, Psc-Eri and \cn\ show more stars in
-the void than the Pleiades).  The late M dwarfs with detected rotation
-periods tend to rotate rapidly ($<3$\,days), though there may also be
-late M dwarfs with slow ($>20$\,days) rotation periods \citep{stauffer_rotation_2016}.  Although
-these late M dwarfs are beyond the magnitude limit of our analysis,
-they have been explored in depth for \cn\ by both
-\citet{Irwin_NGC2516_2007} and \citet{fritzewski_rotation_2020}.
-A detailed comparison of our results with theirs is beyond the
-scope of this work, but we emphasize that the main contribution of our
-study beyond these previous analyses is our sensitivity
-across both the core and halo of \cn, and a
-corresponding expansion in the number of stars that can be analyzed.
-In the following section, we therefore focus on the origins of the
-features in the rotation-color diagram, and whether our expanded
-sample can provide any new insights.
+\citet{fritzewski_rotation_2020} demonstrated that between 100 and
+200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
+the Pleiades) all show similar rotation period distributions.  The FGK
+stars show a ``slow sequence'', a population of ``rapid rotators'',
+and in some cases stars in the ``void'' between the two populations
+({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
+Pleiades).  The late M dwarfs with detected rotation periods tend to
+rotate rapidly ($<3$\,days), though there may also be late M dwarfs
+with slow ($>20$\,days) rotation periods
+\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
+beyond the magnitude limit of our analysis, they have been explored in
+depth by \citet{Irwin_NGC2516_2007} and
+\citet{fritzewski_rotation_2020}.  A comparison of our results is
+shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
+includes, in order of precedence: rotation periods from our analysis,
+from ``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
+from the rotators in \citet{Irwin_NGC2516_2007}.  The main
+contribution of our study beyond these previous analyses is our
+sensitivity across both the core and halo of \cn, and a corresponding
+expansion in the number of stars that can be analyzed.  In the
+following section, we therefore focus on the origins of the features
+in the rotation-color diagram, and whether our expanded sample can
+provide any new insights.
 
@@
 the question then shifts to why these timescales should differ for
-stars that are otherwise identical.  Another possibility, perhaps
-favored, is that proposed by \citet{matt_mass-dependence_2015}: the
-saturation of the magnetic dynamo yields $\dot{P}\sim P$ when $P >
-P_{\rm sat}$, $\dot{P} \sim P^{-1}$ for $P < P_{\rm sat}$, and a
+stars that are otherwise identical.  Another possibility
+is that proposed by \citet{matt_mass-dependence_2015}: the
+saturation of the magnetic dynamo yields ${\rm d}P/{\rm d}t \sim P$ when $P >
+P_{\rm sat}$, ${\rm d}P/{\rm d}t \sim P^{-1}$ for $P < P_{\rm sat}$, and a
 maximal spin-down rate in the intermediate regime.  This could
@@
 background of the Galaxy throughout centuries of modern astronomy.
-Their discovery likely represents a major step toward understanding
+Their discovery represents an important step toward understanding
 the dispersal of open clusters.
@@
 
-The authors thank, in reverse-alphabetical order, G.~Zhou,
+The authors are indebted, in reverse-alphabetical order, to G.~Zhou,
 B.~Tofflemire, A.~McWilliam, E.~Newton, L.~Hillenbrand, A.~Kraus, and

### 2021-04-27 cluster_paper/collect.sh

@@ -44,6 +44,7 @@ cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear.pdf f
 cp ../results/venn/venn.pdf f10.pdf
 
 cp ../results/lightcurves_rotators/lightcurves_rotators_color.pdf f14.pdf
+cp ../results/litcomp_ngc2516/litcomp_ngc2516.pdf f16.pdf
 
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation.pdf f7a.pdf
 # cp ../results/calc_backintegration_ngc2516/core_halo_to_ngc2516_separation_fix_rvs.pdf f7b.pdf

### 2021-04-27 drivers/plot_litcomp_ngc2516.py

@@ -1,0 +1,10 @@
+import os
+import earhart.plotting as ep
+from earhart.paths import RESULTSDIR
+
+PLOTDIR = os.path.join(RESULTSDIR, 'litcomp_ngc2516')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+ep.plot_litcomp_ngc2516(PLOTDIR)
+

### 2021-04-27 earhart/plotting.py

@@ -23,6 +23,7 @@ Contents:
         plot_randich_lithium (deprecated)
     Other:
         plot_lumfunction_vs_position
+        plot_litcomp_ngc2516
         plot_gaia_rv_scatter_vs_brightness
         plot_ruwe_vs_apparentmag
         plot_edr3_blending_vs_apparentmag
@@ -3171,7 +3172,7 @@ def plot_physical_X_rotation(outdir, basedata=None, show1937=0,
 
         axs[0].set_xlabel('$\Delta r_{\mathrm{3D}}$ [pc]')
         # axs[0].set_ylabel('Fraction in bin with P$_\mathrm{rot}$')
-        fig.text(-0.02,0.5, 'Fraction in bin with measured P$_\mathrm{rot}$', va='center',
+        fig.text(-0.02,0.5, 'Fraction with measured P$_\mathrm{rot}$', va='center',
                  rotation=90)
 
         # axs[0].set_xlim([-10, 410])
@@ -3830,7 +3831,7 @@ def plot_lumfunction_vs_position(outdir):
     #           handletextpad=0.1, fontsize='small')
 
     ax.set_xlabel('Absolute $(\mathrm{M}_{G})_0$ [mag]')
-    ax.set_ylabel('Number in bin')
+    ax.set_ylabel('Count')
 
     format_ax(ax)
 
@@ -3942,4 +3943,278 @@ def plot_lightcurves_rotators(outdir, cleaning='periodogram_match', color=0):
     if color:
         outstr += '_color'
     outpath = os.path.join(outdir, f'lightcurves_rotators{outstr}.png')
+    savefig(fig, outpath)
+
+
+def plot_litcomp_ngc2516(outdir):
+
+    f20path = os.path.join(DATADIR, 'rotation',
+                           'Fritzewski_2020_table4_555_Prot_GKM.fits')
+    hl = fits.open(f20path)
+    d = hl[1].data
+    df_lit = Table(d).to_pandas()
+    b21path = os.path.join(RESULTSDIR, 'tables',
+                           'NGC_2516_Prot_cleaned.csv')
+    df = pd.read_csv(b21path)
+
+    df_b21 = df[df.in_SetA]
+    df_f20 = df_lit[~pd.isnull(df_lit.Prot) &
[…]

### 2021-04-28 cluster_paper/authors.tex
update paper
update ms
Co-authored-by: Gaspar Bakos <gbakos@astro.princeton.edu>

@@
 \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+
+\author[0000-0002-2792-134X]{J. L. Curtis}
+\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
+\affiliation{Department of Astrophysics, American Museum of Natural History, Central Park West, New York, NY 10024, USA}
 
@@
 \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+\affiliation{Institute for Advanced Study, 1 Einstein Drive, Princeton, NJ 08540, USA}

### 2021-04-28 cluster_paper/ms.tex
update ms
Co-authored-by: Gaspar Bakos <gbakos@astro.princeton.edu>

@@
   which has a classical tidal radius of 10\,pc and an apparent halo of
-  stars spanning 500\,pc.  Combining photometry from Gaia, rotation
-  periods from TESS, and lithium measurements from Gaia-ESO and GALAH,
-  we find that the halo of NGC\,2516 is the same age as the cluster's
-  core.  Two thirds of kinematically selected halo members out to
-  250\,pc from the cluster center have rotation periods consistent
-  with a gyrochronological age of 150\,Myr.  A comparison sample of
-  field stars shows no such trend.  The lithium abundances of stars in
-  the halo are higher than in the field, and are correlated with the
-  stellar rotation rate and binarity fraction, as has been
-  noted in other young open clusters.  Broadly speaking,
+  stars spanning 500\,pc ($20^\circ$ on-sky).  Combining photometry
+  from Gaia, rotation periods from TESS, and lithium measurements from
+  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is the same
+  age as the cluster's core.  Two thirds of kinematically selected
+  halo members out to 250\,pc from the cluster center have rotation
+  periods consistent with a gyrochronological age of 150\,Myr.  A
+  comparison sample of field stars shows no such trend.  The lithium
+  abundances of stars in the halo are higher than in the field, and
+  are correlated with the stellar rotation rate and binarity fraction,
+  as has been noted in other young open clusters.  Broadly speaking,
   this work supports a new paradigm wherein the halos of open clusters
@@
 for a long time, Gaia is now enabling major advances \citep[{\it
-e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
+e.g.},][]{de_zeeuw_hipparcos_1999,bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_2019}.
 The populations found by the most recent studies can be summarized as
-follows.  On one end are coeval groups with no discernable cores ({\it
+follows.  On one end are groups with no discernable cores ({\it
 e.g.}, Psc-Eri and $\mu$ Tau,
@@
 dispersed group.  ``Halo'' is model agnostic, but is not
-an ideal term because it connotes spherical symmetry, which is rarely
+an ideal term because it can connote spherical symmetry, which is rarely
 the case.  The halo of \cn\ is most accurately described as consisting
@@
 
-For the metallicity, we considered a range of super and sub-solar
-metallicities to fit as much of the locus from
-$0.5<\bpmrp<1.5$ as possible, and settled on the
-slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
-Sub-solar metallicities led to model predictions too blue along the
-main sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees
-with the spectroscopic metallicity from
-\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
-abundance determined by the Gaia-ESO team
-\citep{baratella_gaiaeso_2020}. The latter result represented an up-revision from
-an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We
-caution though that other sub-solar metallicities have been reported
-\citep{bailey_rv_2018}, and note that the cluster metallicity and
-reddening are degenerate; if the reddening is lowered, the implied
-metallicity will increase.  While a detailed redetermination of
-these parameters from the Gaia photometry is beyond our scope, our
-adopted values for both the reddening and metallicity are within the
-range of previously reported values in the literature.
+For the metallicity, we considered a range of super and sub-solar metallicities
+to fit as much of the locus from $0.5<\bpmrp<1.5$ as possible, and settled by
+eye on the slightly super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.
+Sub-solar metallicities led to model predictions too blue along the main
+sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees with the
+spectroscopic metallicity from \citet[][Sec~4.4.4]{cummings_2011_li_iron}, and
+with the iron abundance determined by the Gaia-ESO team
+\citep{baratella_gaiaeso_2020}. The latter result represented an up-revision
+from an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  We caution
+though that other sub-solar metallicities have been reported
+\citep{bailey_rv_2018}, and note that the cluster metallicity and reddening are
+degenerate; if the reddening is lowered, the implied metallicity will increase.
+While a detailed redetermination of these parameters from the Gaia photometry
+is beyond our scope, our adopted values for both the reddening and metallicity
+are within the range of previously reported values in the literature.
 
@@
 Correlations between rapid rotation and binarity in both young
-clusters and the field have preivously been noted
+clusters and the field have previously been noted
 \citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}.
@@
   7.5\% in relative flux.
-  Supplementary plots for the light curves each star are available
-  at \url{http://lgbouma.com/notes}.
+  Supplementary plots that enable inspection of each individual star
+  are available at \url{https://lgbouma.com/notes}.
   \label{fig:lightcurves}
@@
 
-After all these considerations, for every target star, we
-ultimately decorrelated the raw (image-subtracted and
-background-subtracted) light curve using a linear model with ordinary
-least squares.  The vectors used in the model were the 1
-unsmoothed background flux vector, and 5 smoothed PCA eigenvectors.
-Figure~\ref{fig:lightcurves} shows fifty light curves
-randomly drawn from the resulting Set $\mathcal{B}$.
+After all these considerations, for every target star, we ultimately
+decorrelated the raw (image-subtracted and background-subtracted)
+light curve using a linear model with ordinary least squares.  The
+vectors used in the model were the 1 unsmoothed background flux
+vector, and 5 smoothed PCA eigenvectors.  Figure~\ref{fig:lightcurves}
+shows fifty light curves randomly drawn from the resulting Set
+$\mathcal{B}$.  A supplementary figure set that enables inspection of
+each individual star is available at \url{https://lgbouma.com/notes}.
 

### 2021-04-28 drivers/cluster_rotation/plot_autorotation.py

@@ -14,6 +14,7 @@ RUNID_EXTINCTION_DICT = {
 #     'ScoOB2': 0.161, # KC19 ratio
 }
 
+
 for runid, _ in RUNID_EXTINCTION_DICT.items():
 
     E_BpmRp = RUNID_EXTINCTION_DICT[runid]

### 2021-04-28 earhart/plotting.py

@@ -1161,11 +1161,11 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     classes = ['pleiades', 'praesepe', f'{runid}']
     colors = ['gray', 'gray', 'k']
     zorders = [-2, -3, -1]
-    markers = ['s', '+', 'o']
+    markers = ['X', '+', 'o']
     lws = [0., 0.1, 0.1]
-    mews= [0.5, 0.5, 0.5]
+    mews= [0., 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
-    ss = [5, 6, 10]
+    ss = [15, 12, 10]
     labels = ['Pleaides', 'Praesepe', f'{runid}']
 
     # plot vals
@@ -1817,8 +1817,10 @@ def plot_lithium_EW_vs_color(outdir, gaiaeso=0, galahdr3=0,
         dets = (bouvier18_df.l_WLi != "<")
         ax.scatter(
             bouvier18_df[dets]['BpmRp0'], bouvier18_df[dets]['WLi'], c='gray', alpha=1,
-            zorder=-1, s=7, edgecolors='k', marker='s',
-            linewidths=0.2, label='Pleiades'
+            zorder=-1, s=15, edgecolors='k', marker='X',
+            linewidths=0, label='Pleiades'
+
+
         )
         # ax.scatter(
         #     bouvier18_df[~dets]['BpmRp0'], bouvier18_df[~dets]['WLi'], c='gray', alpha=0.8,

### 2021-04-29 cluster_paper/ms.tex
update ms

@@
 \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another is to
-understand the Sun's birth environment \citep{adams_birth_2010}.  Did
-the Sun form in an open cluster, and if so, what type? Is there any
-hope at identifying the stars that formed near the Sun?  Open clusters
-and their dispersed remnants also provide a test case for ``chemical
-tagging'' \citep[{\it
-e.g.},][]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
-The converse holds too: one can use chemical tagging to test the
-viability of kinematic clustering methods.
+understand the Sun's birth environment \citep{adams_birth_2010}.  The
+plausibility of finding the stars that formed with the Sun, whether
+through chemical tagging
+\citep{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018},
+kinematic analyses, or both, can be tested by first using the
+remnants of open clusters that have only recently
+dissipated into the field.
 
 A separate project that benefits from the new discoveries of dispersed
-stellar populations is that of finding young transiting exoplanets.
+stellar populations is that of finding young transiting exoplanets
+\citep[{\it
+e.g.},][]{Mann_K2_33b_2016,ciardi_k2-136_2018,david_four_2019,livingston_k2-264_2019,bouma_cluster_2020,rizzuto_tess_2020,plavchan_planet_2020,newton_2021,tofflemire_2021,zhou_2021_tois}.
 Young transiting planets are hard to find because young stars are
-rare, and mostly reside in the crowded galactic plane \citep[see {\it
-e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
-dispersed halos of nearby star clusters could be reliably identified,
-this could expand the census of nearby young stars by up to a factor
-of $10$, based on the expected fraction of stars thought to be lost
-during gas dispersal.
+rare, and mostly reside in the crowded galactic plane
+\citep{Kharchenko_et_al_2013,piskunov_global_2018}.  If the dispersed
+halos of nearby star clusters could be reliably identified, this could
+expand the census of nearby young stars by up to a factor of $10$,
+based on the expected fraction of stars thought to be lost during gas
+dispersal.
 
@@
 follows.  On one end are groups with no discernable cores ({\it
-e.g.}, Psc-Eri and $\mu$ Tau,
+e.g.}, Psc-Eri and $\mu$-Tau,
 \citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
@@
 (Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
-discuss the implications of our analysis for NGC\,2516 specifically
-and stellar spin-down more generally.  Section~\ref{sec:conclusion}
+discuss the implications of our analysis for open cluster dispersal,
+stellar spin-down, and the lithium-rotation correlation.  Section~\ref{sec:conclusion}
 gives our conclusions.
@@
 assuming the \texttt{astropy v4.0} coordinate standard
-\citep{astropy_2018}.  The direction of galactic rotation is
-$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
+\citep{astropy_2018}.  The Galaxy rotates in the
+$+\hat{Y}$ direction. The cluster orbit (gray line) was evaluated by taking the
 median parameters for core members for which
@@
   the {\it top right}, the 
-  ratio of detected to expected stars with rotation period detections is shown
+  ratio of detected to expected stars with rotation periods is shown
   versus 3D separation from
@@
   \citet{fritzewski_rotation_2020} (F20).
+  I07 and F20 almost exclusively analyzed stars in the core of \cn.
   \label{fig:litcomp}
@@
  	{\it Imaging}:
-    Second Generation Digitized Sky Survey.%,
+    Second Generation Digitized Sky Survey. %,
     %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).

### 2021-04-29 cluster_paper/ms.tex
update ms

@@
 %% Add "Submitted to " argument.
-\received{---}
+\received{April 29, 2021}
 \revised{---}
@@
 		\leavevmode
-		\includegraphics[width=0.99\textwidth]{f11.pdf}
+		\includegraphics[width=0.99\textwidth]{f2.pdf}
 	\end{center}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f2a.pdf}
-			\includegraphics[width=0.49\textwidth]{f2b.pdf}
+			\includegraphics[width=0.49\textwidth]{f3a.pdf}
+			\includegraphics[width=0.49\textwidth]{f3b.pdf}
 		}
@@
     8.5\}$.  The arrow represents the average reddening correction
-    applied to the models (and to give the approximate spectral
-    types).  The opacity of the M-dwarf model atmospheres is less
-    well-calibrated for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V);
-    these model values are shown with dashed lines.
+    applied to the models.  The opacity of the M-dwarf model
+    atmospheres is less well-calibrated for $M_\star \lesssim
+    0.45M_\odot$ ($\approx$M2V); these model values are shown with
+    dashed lines.
     \label{fig:hr}
@@
 		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f3a.pdf}
+			\includegraphics[width=0.700\textwidth]{f4a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f3b.pdf}
+			\includegraphics[width=0.700\textwidth]{f4b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f3c.pdf}
+			\includegraphics[width=0.700\textwidth]{f5a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f3d.pdf}
+			\includegraphics[width=0.700\textwidth]{f5b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.48\textwidth]{f8a.pdf}
-			\includegraphics[width=0.5\textwidth]{f8b.pdf}
+			\includegraphics[width=0.48\textwidth]{f6a.pdf}
+			\includegraphics[width=0.5\textwidth]{f6b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.9\textwidth]{f13.pdf}
+			\includegraphics[width=0.9\textwidth]{f6c.pdf}
 		}
@@
   ($N=54$) per bin.  The {\it bottom panels} show the innermost and
-  outermost 100 stars with detection rotation periods in position and
+  outermost 100 stars with detected rotation periods in position and
   velocity space.
@@
 		\leavevmode
-			\includegraphics[width=0.9\textwidth]{f5a.pdf}
+			\includegraphics[width=0.9\textwidth]{f7.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f5b.pdf}
+		\includegraphics[width=0.9\textwidth]{f8.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f15.pdf}
+		\includegraphics[width=0.5\textwidth]{f9.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f16.pdf}
+		\includegraphics[width=0.95\textwidth]{f10.pdf}
 	\end{center}
@@
   \citet{fritzewski_rotation_2020} (F20).
-  I07 and F20 almost exclusively analyzed stars in the core of \cn.
+  I07 and F20 analyzed stars in the core of \cn.
   \label{fig:litcomp}
@@
 While many rapid rotators in \cn\ (34\%) have no evidence for being
-binaries, we emphasize that the cuts we used to select binaries (${\rm
-RUWE}>1.2$; photometric excess exceeding 0.3\,mag) leave open a
+binaries, the cuts we used to select binaries (${\rm
+RUWE}>1.2$; photometric excess $>$0.3\,mag) leave open a
 significant fraction of parameter space.  A 0.3\,mag flux excess, for
@@
 periods seen in \cn\ across both Sets $\mathcal{A}$ and $\mathcal{B}$
-extend up to $\approx$11 days, rather than the $\approx$8.5day upper
+extend up to $\approx$11 days, rather than the $\approx$8.5 day upper
 limit seen in the Pleiades.  This is consistent with \cn\ being
@@
 
-The authors are indebted, in reverse-alphabetical order, to G.~Zhou,
-B.~Tofflemire, A.~McWilliam, E.~Newton, L.~Hillenbrand, A.~Kraus, and
-K.~Hawkins for the discussions on young stars, rotation, and lithium
-that encouraged this analysis.
+The authors are grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
+E{.}~Newton, A{.}~Kraus, L{.}~Hillenbrand, and K{.}~Hawkins for the
+discussions on young stars, rotation, and lithium that encouraged this
+analysis.
 %
@@
 		\leavevmode
-		\includegraphics[width=0.3\textwidth]{f10.pdf}
+		\includegraphics[width=0.3\textwidth]{f11.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f14.pdf}
+		\includegraphics[width=0.8\textwidth]{f12.pdf}
 	\end{center}
@@
   7.5\% in relative flux.
-  Supplementary plots that enable inspection of each individual star
-  are available at \url{https://lgbouma.com/notes}.
   \label{fig:lightcurves}
@@
 		\leavevmode
-    \includegraphics[width=0.69\textwidth]{f9.pdf}
+    \includegraphics[width=0.69\textwidth]{f13.pdf}
 	\end{center}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f12a.pdf}
-			\includegraphics[width=0.49\textwidth]{f12b.pdf}
+			\includegraphics[width=0.49\textwidth]{f14a.pdf}
+			\includegraphics[width=0.49\textwidth]{f14b.pdf}
 		}

### 2021-04-29 cluster_paper/collect.sh

@@ -1,53 +1,43 @@
-#cp ../results/cluster_membership/full_kinematics_bright_galactic.pdf f1.pdf
+# main text
 cp ../results/cluster_membership/full_kinematics_fullfaint_galactic.pdf f1.pdf
 
-cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f2a.pdf
-cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f2b.pdf
-cp ../results/hr/hr_colorhalobyglat_Bp_m_Rp_show1937_fullfaint_edr3.pdf f2c.pdf
+cp ../results/vtangential_projection/XYZ_with_orbit.pdf f2.pdf
 
-# cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear_defaultcleaning.pdf f3a.pdf
-# cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_linear_periodogram_match.pdf f3b.pdf
-# cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_linear_defaultcleaning.pdf f3c.pdf
-# cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_linear_periodogram_match.pdf f3d.pdf
+cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f3a.pdf
+cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f3b.pdf
 
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_linear_defaultcleaning.pdf f3a.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_linear_periodogram_match.pdf f3b.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_talkaspect_linear_defaultcleaning.pdf f3c.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_talkaspect_linear_periodogram_match.pdf f3d.pdf
[…]

### 2021-04-29 cluster_paper/ms.tex
AAS journals subm.

@@
   stellar associations (1582),
-  open clusters (1160),
+  open star clusters (1160),
   stellar rotation (1629),
@@
 
-The authors are grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
-E{.}~Newton, A{.}~Kraus, L{.}~Hillenbrand, and K{.}~Hawkins for the
-discussions on young stars, rotation, and lithium that encouraged this
-analysis.
-%
-L.G.B. and J.H. acknowledge support by the TESS GI Program, programs
+L.G.B{.} is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
+E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
+K{.}~Hawkins for the discussions on young stars, rotation, and lithium
+that encouraged this analysis.
+%
+L.G.B{.} and J.H{.} acknowledge support by the TESS GI Program, programs
 G011103 and G022117, through NASA grants 80NSSC19K0386 and
@@
 %
-L.G.B. was also supported by a Charlotte Elizabeth Procter Fellowship
+L.G.B{.} was also supported by a Charlotte Elizabeth Procter Fellowship
 from Princeton University.
@@
 Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
-2020A-0146; 2020B-0029 PI: L{.}~Bouma), which is managed by the
+2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
 Association of Universities for Research in Astronomy (AURA) under a

### 2021-05-03 drivers/plot_skyfield.py
update plotting label bug

@@ -361,8 +361,8 @@ def plot_carina():
 if __name__ == "__main__":
 
     do_references = 0
-    do_carina = 1
-    do_skyfield = 0
+    do_carina = 0
+    do_skyfield = 1
 
     # debugging
     if do_references:

### 2021-05-03 earhart/plotting.py

@@ -3390,6 +3390,18 @@ def plot_vtangential_projection(outdir, basedata='fullfaint'):
         scale_units='xy', scale=1, color='C0',
         width=6e-3, linewidths=4, headwidth=8, zorder=9
     )
+
+    delta_x = 0.1
+    axs[1].arrow(0.73, 0.07, delta_x, 0,
+                 length_includes_head=True, head_width=1e-2,
+                 head_length=1e-2,
+                 transform=axs[1].transAxes)
+
+    bbox = dict(facecolor='white', alpha=1, pad=0, edgecolor='white')
+    axs[1].text(0.73+delta_x/2, 0.085, 'Galactic center', va='bottom',
+                ha='center', transform=axs[1].transAxes, fontsize='xx-small',
+                zorder=12, bbox=bbox)
+
     axs[1].update({'xlabel': 'X [pc]', 'ylabel': 'Z [pc]'})
 
     axs[2].scatter(s_halo_df.y_pc, s_halo_df.z_pc, s=s, c='k', rasterized=True)

### 2021-05-12 cluster_paper/maintable_header.tex
update ms
ms: fix typo

@@
           \texttt{parallax\_error} &            0.028395 &                                                          Gaia DR2 parallax uncertainty [mas]. \\
-                     \texttt{pmra} &           -4.225826 &                           Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
-                    \texttt{pmdec} &           10.899938 &                                       Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
+                     \texttt{pmra} &           -4.225826 &                           Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
+                    \texttt{pmdec} &           10.899938 &                                       Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
        \texttt{phot\_g\_mean\_mag} &           15.778953 &                                                                       Gaia DR2 $G$ magnitude. \\
@@
       \texttt{phot\_rp\_mean\_mag} &            14.85896 &                                                           Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
-         \texttt{radial\_velocity} &                 NaN &                                       Gaia DR2 heliocentric radial velocity [km$\,$s${^-1}$]. \\
-  \texttt{radial\_velocity\_error} &                 NaN &                                        Gaia DR2 radial velocity uncertainty [km$\,$s${^-1}$]. \\
+         \texttt{radial\_velocity} &                 NaN &                                       Gaia DR2 heliocentric radial velocity [km$\,$s$^{-1}$]. \\
+  \texttt{radial\_velocity\_error} &                 NaN &                                        Gaia DR2 radial velocity uncertainty [km$\,$s$^{-1}$]. \\
                \texttt{subcluster} &                core &                                                    Is star in core (CG18) or halo (KC19+M21)? \\
@@
 $\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
-cleaning (see Section~\ref{subsubsec:cluster}).  Note that when the
+cleaning (see Section~\ref{subsubsec:cluster}).  When the
 target star is the only star present in the TESS aperture,
@@
 \url{https://archive.stsci.edu/hlsp/cdips}.
+Supplementary plots enabling the analysis individual stars
+are available at \url{https://lgbouma.com/notes}.
 }

### 2021-05-12 drivers/write_NGC2516_main_table.py

@@ -205,13 +205,13 @@ namedesc_dict = {
 'ref_epoch': "Reference epoch for right ascension and declination.",
 'parallax': "Gaia DR2 parallax [mas].",
 'parallax_error': "Gaia DR2 parallax uncertainty [mas].",
-'pmra': r"Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$].",
-'pmdec': "Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$].",
+'pmra': r"Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$].",
+'pmdec': "Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$].",
 'phot_g_mean_mag': "Gaia DR2 $G$ magnitude.",
 'phot_bp_mean_mag': "Gaia DR2 $G_\mathrm{BP}$ magnitude.",
 'phot_rp_mean_mag': "Gaia DR2 $G_\mathrm{RP}$ magnitude.",
-'radial_velocity': "Gaia DR2 heliocentric radial velocity [km$\,$s${^-1}$].",
-'radial_velocity_error': "Gaia DR2 radial velocity uncertainty [km$\,$s${^-1}$].",
+'radial_velocity': "Gaia DR2 heliocentric radial velocity [km$\,$s$^{-1}$].",
+'radial_velocity_error': "Gaia DR2 radial velocity uncertainty [km$\,$s$^{-1}$].",
 'subcluster': "Is star in core (CG18) or halo (KC19+M21)?",
 'in_CG18': "Star in \\citet{cantatgaudin_gaia_2018}.",
 'in_KC19': "Star in \\citet{kounkel_untangling_2019}.",

### 2021-05-31 planet_paper/ms.tex
update ms
update planet ms

@@
 
+  \paragraph{Steve Shectman on resolution in 3x3 binning mode:}
+	I think what you've done is taken the resolution of the spectrograph with 1x2 binning, which is about 130,000, and divided by three.
+The actual situation is as follows.
+The scale at the detector is 158 microns per arcsecond, so the 0.3 arcsecond slit is 47 microns wide.
+An unbinned pixel is 9 microns, which corresponds to 450 m/s, and the measured FWHM is about 5 unbinned pixels (45 microns or 2.25 km/s).
+So then the resolution is 300,000/2.25 = ~133,000.
+
+	So the unbinned spectra are nicely sampled with about 5 pixels per FWHM, while the binned spectra are somewhat undersampled with about 1.7 pixels per FWHM.
+There's a small loss in resolution caused by the discrete pixel size (it kind of adds 1 pixel in quadrature to the underlying FWHM), so maybe the effective FWHM when the spectra are binned 3x3 is more like 2 pixels instead of 1.7, corresponding to 6 pixels unbinned.
+That would still be a resolution of about 110,000.
+I don't actually have a rigorous measurement of the resolution with 3x3 binning, but I think 110,000 is a reasonable number to quote.
+
   %\input{TOI837_rv_table.tex}
@@
   \subsection{TOI 1937 specifically}
+
+  \subsection{Comparison Stars That We Got Spectra For}
+  They are at
+  /Users/luke/Dropbox/proj/earhart/data/spectra/comparison\_stars.
+
+  There are 6 CHIRON ones, of varying assailability.
+  And 1 PFS unassailable member.
 

### 2021-07-01 cluster_paper_review/referee_response_v0.txt
[WIP] staging for review
(not done)

Dear Dr. Sarajedini,

We thank you for organizing the review of our paper, and also
thank the referee for constructive criticism.  We have considered
the referee's remarks carefully and revised our manuscript
accordingly.

Below, we have reproduced the relevant portions of the referee's
report, alongside our responses. A list of changes is appended to
the manuscript using the trackchanges AASTeX macros. Figures
supporting X AND Y WORDS TODO FIX THESE have also been uploaded for inspection.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
At this stage the referee is more inclined to believe that the authors have
identified a set of young stars in that region of the sky than actual halo
member stars of NGC 2516.

RESPONSE
----------
> We agree that the stars with rotation periods similar to those of the
> Pleiades, and showing Li EWs similar to those of the Pleiades, should indeed
> be considered to have comparable ages to stars in Pleiades.
>
> Figures 4 and 7 show that the "halo" stars from KC19 and M21 overlap both
> with the Pleiades, and with stars in the core of the cluster (those from
> CG18).  They share the same gyrochronological and Li tracks, and therefore
> share the same age to within measurement precision.  The stars in this "halo"
> population are also mostly within 2 km/s of tangential velocity separation
> from the main core population (see panel in top left of Figure 6).
> Spatially, they lead and lag the core of the population, in a similar manner
> to that seen for eight other open clusters by Meingast et al 2021 (Figure 2
> of our manuscript; Figure 8 of Meingast et al 2021, Astronomy & Astrophysics,
> Volume 645, id.A84).
>
> Despite these lines of evidence for the halo stars being associated with NGC
> 2516 (same ages; same tangential velocities; spatial correlation with the
> cluster center), for the sake of caution we analyzed rotation periods of
> randomly draw field stars in this region of sky.  This exercise was described
> in Appendix C, and the results are shown in Figure 13 of the manuscript.  The
> randomly selected field stars (orange points) does not show any evidence for
> an overdensity of Pleiades-age stars.  The candidate "halo" stars selected
> kinematically by CG18, KC19, and M21 do (black points).  As a clarification
> on Figure 13, we have attached "f13_corehalosplit.png" as a supplement in the
> response, which shows the same data, but splitting NGC 2516 into halo and
> core stars.  Stars in the halo (blue) have a very different rotation-color
> distribution than randomly selected field stars (orange).
>
> In short, most of the halo stars i) have similar tangential velocites as core
> stars, ii) are coeval with other halo stars, iii) have the same age as the
> core stars, iv) are spatially correlated with the location of the core stars,
> and iv) have a different rotation-color distribution from a randomly selected
> sample of fields stars (with a large overdensity of rotators falling on the
> same sequence as the core of NGC 2516).
>
> The suggestion that we "have identified a set of young stars in that region
> of the sky [rather] than actual halo member stars of NGC 2516." is equivalent
> to asserting that the clustering techniques used by KC19 and M21 yielded a
> set of stars that is identical to the cluster core according to all age
> measurements, and is spatially and kinematically correlated with the cluster
> core, but are otherwise unrelated.  While the spatial and kinematic
> correlations are tautological -- this is what clustering methods do if they
> are working correctly -- the age measurements have no a priori reason to be
> identical between the core and halo unless the two are in fact related.

> TODO: maybe empahsize the false positive fraction of 50% here.

> Add an uncertainty on the false positive rate? Could it be higher than 50%?

******************************************
REFEREE COMMENT
----------
The authors begin by concatenating a very large collection of stars identified by
CG18 (which they term the core members), and those of KC19 and M21, the last two
having very relaxed criteria for association with NGC 2516.
At this stage, as one can see from the distributions in Fig.1, the sample is
completely bloated with field stars which now clearly have to be excised.
The question is how?

The authors justifiably look first to the HR diagram and understandably find
that that is not very helpful in rejecting non-members because almost all stars in
the extended sample lie close to the ridge line of the canonical cluster members.
(Nevertheless, this is already a basis for rejecting stars below the main sequence
and those significantly above it, including giants. The authors should consider
this.) Incidentally, the Gaia photometry for the brightest stars is likely not good
enough for fitting any isochrones of this age, so in this referee's estimation, the
cluster turnoff is perhaps still the best one can do, keeping in mind the persistence
of photometric problems for bright stars.

RESPONSE
----------
> While the HR diagram is not very helpful in rejecting non-members, it does
> indicate that the distribution of field stars in this region of
> ra/dec/parallax is different from stars in both the core and halo of the
> cluster.  In particular, the lack of a sub-giant branch or red giant branch
> (e.g., at G_BP-G_RP =1.2, M_G =3.0) suggests that the core and halo have a
> different age distribution than the field stars, and supports the suggestion
> that for the most part, the halo population has a single age.
>
> For the ~five halo stars lying significantly above the main sequence, we
> agree that these must be field interlopers.  This is noted at the end of
> Section 3.1, Paragraph 2.  The ~four halo stars, and ~four core stars lying
> well below the main sequence could be there due to poorly determined
> parallaxes, or due to blending in the Gaia BP or RP spectra with unresolved
> white dwarf companions.  We do not consider their presence below the MS to
> support their exclusion from consideration as cluster members.

******************************************
REFEREE COMMENT
----------
The authors then go on to consider the rotation periods and lithium abundances
of the extended sample, to compare them with previously-known distribution of
similarly young stars.

The conceptual problem here is that stars that deviate from the known/expected
rotational distribution for NGC 2516 are again not excised from the sample.  As
prior studies (e.g. Fritzewski+2020) have shown, there are no rotators above
the slow sequence in NGC 2516. So such stars most certainly have to be excised
from further consideration. (Some of these are even older than Praesepe, as the
authors themselves show in Fig.4.) The authors should consider whether other
regions of the color-period diagram merit similar treatment.

RESPONSE
----------

> We agree that stars showing rotation periods above the slow sequence are most
> likely not "genuine", in the sense that either i) they are not cluster
> members, or ii) they are cluster members, but with unresolved binary
> companions contaminating the rotation period measurement.  The latter
> category has precedent in the literature (e.g., Stauffer+2016, AJ 152 115,
> Section 5.1).
>
> Our point in proposing "Set A" and "Set B" is that the exact means by which
> the samples are cleaned is not incredibly important to the central result that
> the halo is the same age as the core, and therefore most likely related to
> it.
>
> Set A comprises 987 stars.  947 of these have (Bp-Rp)_0>0.4, and so make some
> degree of sense to assess gyrochronologically.  873 are below what by eye can
> be identified as the "slow sequence boundary".  While there is subjectivity
> in where this line is drawn, the big-picture result is that ~90% of the "Set
> A" stars are be "genuine" in the sense of being "in or below the slow
> sequence".  
>
> We attach two additional figures, "response_fig2.png" and
> "response_fig3.png" illustrating the point.  Cleaning out these stars to make
> a "perfectly contaminant free" list of "genuine cluster members" is not the
> aim of this work.
>
> (We note also that simply because a star is in or below the slow sequence
> does not guarantee that it is genuine).

******************************************
REFEREE COMMENT
----------
An additional concern is that the authors construct two samples A (where rotation
periods from multiple types of measurements don't agree) and B (where rotation
periods agree), but then go further with sample A (for instance Fig.6).
Had the authors gone forward with the tightest-possible sample, and had then
shown that this tight rotation-lithium sample still has a far greater extent on
sky and in the canonical parallax, proper motion, RV etc. spaces, then readers would
be far more willing to believe the authors' claims about a more extensive cluster.

RESPONSE
----------
> "response_fig2.png" and "response_fig3.png" illustrate that, for the
> rotation-color selected subsample of Set A suggested by the referee in the
> previous comment, the halo is the size claimed.
>
> The suggestion to use "Set B" for this is the suggestion to apply a selection
> function that throws away ~85% of stars.  (We suspect this is mostly because
> the SPDM periods tend to be garbage).  The halo is already sparse, so
> throwing away ~9 in 10 of its members will make it sparser.
>
> Regardless, doing this with Set B gives response_fig4.png.  The "slow
> rotating" (and below) halo stars still span a large range of space, though
> more like ~350 pc tip-to-tip rather than ~500 pc.
>
> Anyway, in Set A (N=937), 216 stars are at least 50 parsecs from the cluster
> center, and 104 are at least 100 parsecs away (23% and 11%).  In Set B
> (N=479), 94 stars are at least 50 parsecs from the cluster center, and 43 are
> at least 100 parsecs away (20% and 9%).

******************************************
REFEREE COMMENT
----------
The referee notes that the prior work of Healy & McCullough 2020, who also used
TESS data to measure a sizeable group of stars in NGC 2516 has been completely
ignored. The authors should include a discussion of the H+M2020 work in relation
to the results in this paper.

RESPONSE
----------
> Healy & McCullough 2020 (HM20) was referenced in the second paragraph of
> Section 4.1, in relation to their discussion of the contraction rate of the
> core NGC 2516, and the corresponding rate at which stars are escaping the
> core.  The form in which Healy & McCullough quantified the contraction rate
> does not readily translate (to our knowledge) to an estimate of the rate at
> which stars escape from the cluster's tidal radius, or an analogous surface.
> So, the reference was rather brief.
>
> The main result of our paper -- that the halo is the same age as the core and
> therefore most easily explained as being associated with NGC2516 -- does not
> clearly translate to the main result of HM20 -- that the stellar spins in the
> core of NGC 2516 are either isotropically oriented, or perhaps mildly
> aligned.  Therefore we did not highlight HM20 prominently in our discussion.
>
> Nonetheless, to address the general idea behind this suggestion, we have i)
> added a reference to HM20 in the introduction, when we discuss TESS and
> NGC2516, and ii) included the HM20 rotation periods in Figure 10.  The
> overall point of the figure has not changed.

******************************************
REFEREE COMMENT
----------
While on the subject, it should also be mentioned that the Introduction of the
paper is only from the galactic structure perspective, and is lacking in
introducing readers to the rotation-lithium perspective. Given that the latter
constitute the bulk of the paper, doing that would clearly help the argument
being laid out.

RESPONSE
----------
> We have added a paragraph in the introduction giving some context to this
> effect -- thank you for the suggestion.

******************************************
REFEREE COMMENT
----------
This referee came away with the general impression that the authors have likely
identified a number of young stars in a vast halo around NGC 2516, but did not
find the association with NGC 2516 itself to be particularly convincing.
Perhaps they can think about how to make their arguments tighter, and the
inclusion criterea stricter in the next iteration, to be able to better convince
the community of the association with NGC 2516.

RESPONSE
----------
> XXX of YYY stars for which we expected to detect rotation showed a detected
> rotation signature...
> In the comparison sample...

******************************************
DATA EDITOR'S COMMENT
----------
The .csv file attached with Table 1 has a large number of significant digits in
most columns. In the revised manuscript, the authors should include a new
version that has only the required precision. This document outlines the
typical uncertainties for Gaia data:

https://gea.esac.esa.int/archive/documentation/GDR2/

Fewer significant digits will result in a smaller file that will be easier to
read, manipulate, and transfer.

RESPONSE
----------
> The CSV file has been updated with an appropriate number of significant
> digits in each column; thank you for the suggestion.

### 2021-07-01 cluster_paper/maintable_header.tex
[WIP] staging for review
(not done)

@@
 \startdata
-               \texttt{source\_id} & 5290665204845694336 &                                                                   Gaia DR2 source identifier. \\
-         \texttt{source\_id\_edr3} & 5290665204845694336 &                                                                  Gaia EDR3 source identifier. \\
-                 \texttt{in\_SetA} &                True &                                 In Set $\mathcal{A}$ (LSP>0.08, P<15d, nequal==0, nclose>=1). \\
-                 \texttt{in\_SetB} &                True &                                   In Set $\mathcal{B}$ (Set $\mathcal{A}$ and periods match). \\
+               \texttt{source\_id} & 5290666991552265088 &                                                                                Gaia DR2 source identifier. \\
+         \texttt{source\_id\_edr3} & 5290666991552265088 &                                                                               Gaia EDR3 source identifier. \\
+                 \texttt{in\_SetA} &                True &                                    In Set $\mathcal{A}$ (LSP$>$0.08, P$<$15d, nequal$==$0, nclose$\geq$1). \\
+                 \texttt{in\_SetB} &                True & In Set $\mathcal{B}$ (Set $\mathcal{A}$ and below $P_\mathrm{rot}-(G_\mathrm{BP}$-$G_\mathrm{RP})_0$ cut). \\
          \texttt{n\_cdips\_sector} &                   7 &                                                            Number of TESS sectors with CDIPS light curves. \\
-                   \texttt{period} &            0.228981 &                                                              Lomb-Scargle best period [days]. \\
-                   \texttt{lspval} &            0.657826 &                                               Lomb-Scargle periodogram value for best period. \\
-               \texttt{spdmperiod} &            0.228981 &                                                          Stellingwerf PDM best period [days]. \\
-                  \texttt{spdmval} &            0.372779 &                                           Stellingwerf PDM periodogram value for best period. \\
+                   \texttt{period} &            7.733189 &                                                                           Lomb-Scargle best period [days]. \\
+                   \texttt{lspval} &            0.505373 &                                                            Lomb-Scargle periodogram value for best period. \\
                    \texttt{nequal} &                   0 &                                                 Number of stars brighter than the target in TESS aperture. \\
@@
                    \texttt{nfaint} &                   1 &                                                    Number of stars with $\Delta T > 2.5$ in TESS aperture. \\
-                       \texttt{ra} &          119.234778 &                                                               Gaia DR2 right ascension [deg]. \\
-                      \texttt{dec} &          -61.015929 &                                                                   Gaia DR2 declination [deg]. \\
+                       \texttt{ra} &          119.494256 &                                                                            Gaia DR2 right ascension [deg]. \\
+                      \texttt{dec} &          -61.060819 &                                                                                Gaia DR2 declination [deg]. \\
                \texttt{ref\_epoch} &              2015.5 &                                                       Reference epoch for right ascension and declination. \\
-                 \texttt{parallax} &            2.458023 &                                                                      Gaia DR2 parallax [mas]. \\
-          \texttt{parallax\_error} &            0.028395 &                                                          Gaia DR2 parallax uncertainty [mas]. \\
-                     \texttt{pmra} &           -4.225826 &                           Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
-                    \texttt{pmdec} &           10.899938 &                                       Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
-       \texttt{phot\_g\_mean\_mag} &           15.778953 &                                                                       Gaia DR2 $G$ magnitude. \\
-      \texttt{phot\_bp\_mean\_mag} &           16.629477 &                                                           Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
-      \texttt{phot\_rp\_mean\_mag} &            14.85896 &                                                           Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
+                 \texttt{parallax} &            2.462602 &                                                                                   Gaia DR2 parallax [mas]. \\
+          \texttt{parallax\_error} &            0.017613 &                                                                       Gaia DR2 parallax uncertainty [mas]. \\
+                     \texttt{pmra} &           -5.366865 &                                        Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
+                    \texttt{pmdec} &           10.959816 &                                                    Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
+       \texttt{phot\_g\_mean\_mag} &           14.592845 &                                                                                    Gaia DR2 $G$ magnitude. \\
+      \texttt{phot\_bp\_mean\_mag} &           15.189163 &                                                                        Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
+      \texttt{phot\_rp\_mean\_mag} &           13.867151 &                                                                        Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
          \texttt{radial\_velocity} &                 NaN &                                                    Gaia DR2 heliocentric radial velocity [km$\,$s$^{-1}$]. \\
@@
                  \texttt{in\_KC19} &                True &                                                                   Star in \citet{kounkel_untangling_2019}. \\
-                  \texttt{in\_M21} &               False &                                                                Star in \citet{meingast_2021}. \\
-               \texttt{(Bp-Rp)\_0} &            1.636198 & Gaia $G_\mathrm{BP}$-$G_\mathrm{RP}$ color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
+                  \texttt{in\_M21} &                True &                                                                             Star in \citet{meingast_2021}. \\
+               \texttt{(Bp-Rp)\_0} &            1.187693 &              Gaia $G_\mathrm{BP}$-$G_\mathrm{RP}$ color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$)=0.1343 \\
             \texttt{is\_phot\_bin} &               False &                                                                True if $>0.3$ mag above cluster isochrone. \\
            \texttt{is\_astrm\_bin} &               False &                                                                              True if Gaia EDR3 RUWE > 1.2. \\
-                     \texttt{ruwe} &            0.986349 &                                                                               Gaia EDR3 RUWE. \\
-      \texttt{Li\_EW\_mA\_GaiaESO} &              -1.495 &                          Gaia-ESO Li doublet equivalent width, including the Fe blend [m\AA]. \\
-\texttt{Li\_EW\_mA\_perr\_GaiaESO} &              12.407 &                                              Gaia-ESO Li doublet EW upper uncertainty [m\AA]. \\
+                     \texttt{ruwe} &            0.954835 &                                                                                            Gaia EDR3 RUWE. \\
+      \texttt{Li\_EW\_mA\_GaiaESO} &             144.259 &                                       Gaia-ESO Li doublet equivalent width, including the Fe blend [m\AA]. \\
+\texttt{Li\_EW\_mA\_perr\_GaiaESO} &               6.301 &                                                           Gaia-ESO Li doublet EW upper uncertainty [m\AA]. \\
 \texttt{Li\_EW\_mA\_merr\_GaiaESO} &                 5.0 &                                                           Gaia-ESO Li doublet EW lower uncertainty [m\AA]. \\

### 2021-07-01 cluster_paper/ms.tex
update plotting label bug
[WIP] staging for review
(not done)

@@
 
-\documentclass[12pt,twocolumn,tighten]{aastex63}
+%\documentclass[12pt,twocolumn,tighten]{aastex63}
+\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
@@
 \newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
-\newcommand{\nautorotnumeratormatching}{507} % plot_autorotation.py, LSP=SPDM period
+\newcommand{\nautorotnumeratormatching}{892} % plot_autorotation.py, Set A, with manual Prot-color cut
 \newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
@@
 chose \cn\ for this analysis because it is young (100-200\,Myr) and
-close enough ($d\approx400$\,pc) to facilitate photometric
-rotation measurements using TESS.  Its halo was reported by
+close enough ($d\approx400$\,pc) to facilitate \replaced{photometric
+rotation measurements}{measurements of stellar rotation periods} using TESS.  \added{\citet{healy_stellar_2020}
+in fact already used TESS rotation periods to study the stellar inclination
+distribution of stars in the core of NGC\,2516.}
+\replaced{Its halo}{The cluster's halo however} was reported by
 \citet{kounkel_untangling_2019} to span
-$\approx20\times10\times350$\,pc.  We want to know: is this halo real
-or is it just a collection of interlopers, foreground and background
-stars?  To what extent can we use Gaia alone to reliably identify
-age-dated needles in the haystack of field stars?  What are the
-implications, observationally and theoretically, if we can identify
-the dispersed halos of open clusters?
+$\approx20\times10\times350$\,pc\added{, which is a larger volume than
+has been considered in previous rotation analyses
+\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}}.
+We want to know: is this halo real or is it just a collection of
+interlopers, foreground and background stars?  To what extent can we
+use Gaia alone to reliably identify age-dated needles in the haystack
+of field stars?  What are the implications, observationally and
+theoretically, if we can identify the dispersed halos of open
+clusters?
+
+\added{To assess whether the stars in the cluster's halo are the same
+ages as stars in the cluster's core, we use three different age-dating
+techniques: isochrones, gyrochrones, and lithium depletion
+\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
+Gyrochronology has now been empirically calibrated for FGK stars with
+ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
+e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
+Lithium equivalent widths can be used for relative age-dating of dwarf
+stars over similar age ranges, though for a narrower range of spectral
+types \citep{sestito_2005}.  One complication when deriving relative
+lithium ages of the K dwarfs comes from the correlation between
+lithium abundance and stellar rotation rate, which has recently been
+reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
+that we explore in the latter portions of our analysis.  }
 
@@
 
-\begin{figure*}[t]
+\begin{figure*}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f1.pdf}
+		\includegraphics[width=\textwidth]{f1.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.99\textwidth]{f2.pdf}
+		\includegraphics[width=\textwidth]{f2.pdf}
 	\end{center}
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[tp!]
 	\begin{center}
@@
   and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
-  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
+  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
+  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
+  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
@@
   \label{fig:rot}
-	}
-\end{figure*}
-\begin{figure*}[t!]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f5a.pdf}
-		}
-
-    \vspace{-0.5cm}
-		\subfloat{
-			\includegraphics[width=0.700\textwidth]{f5b.pdf}
-		}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Binarity indicators for rotators in \cn.} Sets
-  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are as in
-  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
-  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
-  above an empirical isochrone; 20\% of stars) are overplotted in
-  orange.  
-  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has 281 stars.
-  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
-  the fraction of stars showing signs of binarity in the slow and fast
-  subsamples are 30\% (63/210) and 56\% (40/71) respectively.
-  \label{fig:binarity}
 	}
@@
 
-We then measured the rotation period from the resulting light curve
-using the periodogram implementations in \texttt{astrobase}, namely
-the traditional Lomb-Scargle, and the Stellingwerf phase-dispersion
-minimization algorithm
-\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
+We then measured the rotation period from the resulting light curve using
+the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
+\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
+the Stellingwerf phase-dispersion minimization algorithm
+\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{\citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
 We used the CDIPS aperture radius that, based on theoretical
-expectations, was expected to give the optimal balance between light
-from the target star and the sky background \citep{Sullivan_2015}.  This
-typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
-We recorded the top five periodogram peaks from each method, and their
-corresponding powers.  Finally, as a check on crowding, we recorded
-the number of stars within the aperture with greater
-brightness than the target star, and with brightness within 1.25 and
-2.5 TESS magnitudes of the target star.
+expectations, was expected to give the optimal balance between light from
+the target star and the sky background \citep{Sullivan_2015}.  This
+typically resulted in a circular aperture radius of either 1 or 1.5
+pixels.  We recorded the top five periodogram peaks\deleted{ from each
+method}, and their corresponding powers.  Finally, as a check on
+crowding, we recorded the number of stars within the aperture with
+greater brightness than the target star, and with brightness within 1.25
+and 2.5 TESS magnitudes of the target star.
 
@@
 automated cleaning criteria, which yielded sets we denote
-$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered light
+$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
+$\mathcal{A}$}, we considered light
 curves for which the peak Lomb-Scargle periodogram period was below 15
@@
 periods might plausibly have been detected.  For Set $\mathcal{B}$, we
-additionally required that the Lomb-Scargle and Stellingwerf
-phase-dispersion methods yielded identical periods to within 10\%.
-This yielded \nautorotnumeratormatching\ light curves.  Boolean flags
-for each set are included in Table~\ref{tab:maintable}.
-A set of field stars was also analyzed for comparison.
-The results are discussed in Appendix~\ref{app:compstar}.
+additionally required that the Lomb-Scargle \replaced{and Stellingwerf
+phase-dispersion methods yielded identical periods to within 10\%}{period
+fell below the boundary drawn in Figure~\ref{fig:rot}, which was
+constructed the filter using the Pleiades polynomial from
+\citet{curtis_rup147_2020}, with hard cuts at the red and blue ends.}.
+This yielded \nautorotnumeratormatching\ light curves\added{in Set
+$\mathcal{B}$}. 
+Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
+set of field stars was also analyzed for comparison.  The results are
+discussed in Appendix~\ref{app:compstar}.
+\added{The justification for the additional cut in Set $\mathcal{B}$ is
+that \cn\ is already known to have few if any stars above the slow
+sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
+are most likely therefore either not cluster members, or they are cluster
+members, but with unresolved binary companions contaminating the rotation
+period measurement \citep[{\it
+e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
+assessing which are rotationally consistent with being in the cluster,
+this is a viable filter; for the purpose of fully exploring whether any
+rotational outliers could be bonafide cluster members, Set $\mathcal{A}$
+would be the better cut.}
 
@@
 
-Figure~\ref{fig:rot} shows the resulting rotation
-periods, with Set $\mathcal{A}$ shown on top and Set $\mathcal{B}$ on
-the bottom.  While Set $\mathcal{A}$ is more complete than Set
-$\mathcal{B}$, this completeness comes at the expense of greater
-contamination.  For instance, stars above the slow sequence (perhaps
-field contaminants) mostly disappear in Set $\mathcal{B}$.  The
-``void'' or ``gap'' underneath the slow sequence is also more
-pronounced in Set $\mathcal{B}$.  To facilitate a comparison against
-the Pleiades and Praesepe, we used the rotation periods and dereddened
-colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
-from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
-respectively.  The first order conclusion is that the Pleiades and
-\cn\ appear gyrochronologically coeval for colors spanning
-$0.5<\bpmrpo<1.2$.  The implications of this comparison for the age of
-\cn\ are discussed in Section~\ref{disc:absage}.
-
-Figure~\ref{fig:binarity} shows the same data, with the
-points colored by indicators of binarity.  To assess astrometric
-binarity, we used the renormalized unit weight error (RUWE) from Gaia
-EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
-1.2 as an astrometric binary (11\% of the overall cluster sample; see
-{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
-photometric binarity, we fitted a spline to Figure~\ref{fig:hr},
-fixing the nodes by hand.  We then labeled any points brighter than
-0.3\,mag above the cluster sequence as photometric binaries.  This
-included 20\% of the overall cluster sample.  Many of the resulting
-candidate binaries, though not all of them, appear below the slow
-rotation sequence.  In Set $\mathcal{B}$, over the color range where
-the slow and fast sequence are present ($0.5<\bpmrpo<1.3$), there are
-281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
-281 stars into ``fast'' and ``slow'' sequences by eye,  the fraction
-of stars showing signs of binarity in the slow and fast subsamples are
-30\% (63/210) and 56\% (40/71), respectively.  We speculate on
-possible explanations in Section~\ref{disc:rotn}.
+\begin{figure*}[t!]
+	\begin{center}
+		\leavevmode
+    \includegraphics[width=0.700\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
+  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
+  $\mathcal{B}$ is shown} as in
+  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
+  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
+  above an empirical isochrone; 20\% of stars) are overplotted in
+  orange.  
+  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
+  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
+  the fraction of stars showing signs of binarity in the slow and fast
+  subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
+  \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
+  \label{fig:binarity}
+	}
+\end{figure*}
+
+Figure~\ref{fig:rot} shows the resulting rotation periods, with Set
+$\mathcal{A}$ shown on top and Set $\mathcal{B}$ on the bottom.  While
+Set $\mathcal{A}$ is more complete than Set $\mathcal{B}$, this
+completeness comes at the expense of greater contamination.  \deleted{For
+instance, stars above the slow sequence (perhaps field contaminants)
+mostly disappear in Set $\mathcal{B}$.  The ``void'' or ``gap''
+underneath the slow sequence is also more pronounced in Set
+$\mathcal{B}$.}  To facilitate a comparison against the Pleiades and
+Praesepe, we used the rotation periods and dereddened colors from Table~5
+of \citet{curtis_rup147_2020}, which drew on data from
+\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019} respectively.
+The first order conclusion is that the Pleiades and \cn\ appear
+gyrochronologically coeval for colors spanning $0.5<\bpmrpo<1.2$.  The
+implications of this comparison for the age of \cn\ are discussed in
+Section~\ref{disc:absage}.
+
+\explain{Figure~\ref{fig:binarity} has been modified to account for the
+revised definition of Set B.}
+
+Figure~\ref{fig:binarity} shows the same data, with the points colored by
+indicators of binarity.  To assess astrometric binarity, we used the
+renormalized unit weight error (RUWE) from Gaia EDR3.  For plotting
+purposes, we labeled anything with RUWE exceeding 1.2 as an astrometric
+binary (11\% of the overall cluster sample; see {\it e.g.},
+\citealt{belokurov_unresolved_2020}).  To assess photometric binarity, we
+fitted a spline to Figure~\ref{fig:hr}, fixing the nodes by hand.  We
+then labeled any points brighter than 0.3\,mag above the cluster sequence
+as photometric binaries.  This included 20\% of the overall cluster
+sample.  Many of the resulting candidate binaries, though not all of
+them, appear below the slow rotation sequence.  In Set $\mathcal{B}$,
+over the color range where the slow and fast sequence are present
+($0.5<\bpmrpo<1.3$), there are \replaced{281}{523} stars, of which
+\replaced{103 (37\%)}{174 (33\%)} show signs of binarity.  Dividing these
+\replaced{281}{523} stars into ``fast'' and ``slow'' sequences by eye,
+the fraction of stars showing signs of binarity in the slow and fast
+subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and \replaced{56\%
+(40/71)}{51\% (68/134)}, respectively.  We speculate on possible
+explanations in Section~\ref{disc:rotn}.
 
@@
 
-\begin{figure*}[t]
+\begin{figure*}[tp]
 	\begin{center}
@@
   2D-tangential velocities for halo members.  Stars with detected
-  rotation periods in Set~$\mathcal{A}$ are shown in blue;  halo
+  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
   members for which rotation periods should have been detectable, but
@@
   outermost 100 stars with detected rotation periods in position and
-  velocity space.
+  velocity space\added{ (from Set~$\mathcal{A})$}.
   \label{fig:physical_x_rotn}
@@
 
+\explain{This paragraph contained a typo: the sentence containg ``at
+least 25\,pc from the cluster core'' was meant to refer to the entire
+cluster. }
+
 The results are shown in the top two rows of
 Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
-comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our
-TESS pipeline succeeded in making light curves.  In other words, the
-stars in the base sample are those for which we could have plausibly
-detected a rotation period.  This color range is preferred because the
-the slow sequence becomes less defined for spectral types later than
-$\approx$K4V.  The bins in the histograms (top-right) were chosen to
-ensure that the same number of stars were in each bin.  Rotation
-periods are detected for 66\% of stars at least 25\,pc from the
-cluster core.  The average detection rate inside of 25\,pc, 73\%, is
-slightly higher.  Within $\approx$3.5\,pc, the period detection rate of
-61\% is relatively low, due to crowding and saturation in the TESS
-images.
+comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our TESS
+pipeline succeeded in making light curves.  In other words, the stars in
+the base sample are those for which we could have plausibly detected a
+rotation period.  This color range is preferred because the the slow
+sequence becomes less defined for spectral types later than $\approx$K4V.
+The bins in the histograms (top-right) were chosen to ensure that the
+same number of stars were in each bin.  Rotation periods
+\added{consistent with a gyrochronological age of 150\,Myr} are detected
+for \replaced{66\%}{68\%} of stars \replaced{at least 25\,pc from the
+cluster core}{reported to be in the cluster}.  The average detection rate
+inside of 25\,pc, \replaced{73\%}{72\%}, is slightly higher\added{ than
+outside of 25\,pc, where the detection rate is 63\%}. Within
+$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\%} is
+relatively low, due to crowding and saturation in the TESS images.
 
@@
 cluster.  Of the outermost 100 stars in position space with detected
-rotation periods, $\approx$8 appear as outliers above the slow
-sequence, compared to just a few for the innermost 100 stars.  For the
-outermost 100 stars in velocity space, $\approx$13 appear as outliers.
-After accounting for the stars that were expected to show rotation
-periods and did not, this suggests field contaminant rates of
-$\approx$50\% for the outermost cluster members in our target list.
+rotation periods\added{ in Set $\mathcal{A}$}, $\approx$8 appear as
+outliers above the slow sequence, compared to just a few for the
+innermost 100 stars.  For the outermost 100 stars in velocity space,
+$\approx$13 appear as outliers.  After accounting for the stars that were
+expected to show rotation periods and did not, this suggests field
+contaminant rates of $\approx$50\% for the outermost cluster members in
+our \added{initial} target list.
 
@@
 halo population of \cn\ was an analysis of the Li\,\textsc{I}
-6708\,\AA\ doublet.  This type of age measurement has been reviewed by
+6708\,\AA\ doublet.  \deleted{This type of age measurement has been reviewed by
 \citet{soderblom_ages_2010}, and its connection to stellar rotation
-has been discussed by \citet{bouvier_lithium-rotation_2020}.
+has been discussed by \citet{bouvier_lithium-rotation_2020}.}
 For \cn, two spectroscopic datasets seemed important:
@@
   The bottom panel is a concatenation of Set $\mathcal{B}$ with the
-  results of \citet{Irwin_NGC2516_2007} (I07) and
-  \citet{fritzewski_rotation_2020} (F20).
-  I07 and F20 analyzed stars in the core of \cn.
+  results of \replaced{\citet{Irwin_NGC2516_2007} (I07) and
+  \citet{fritzewski_rotation_2020} (F20)}{\citet{Irwin_NGC2516_2007}
+  (I07), \citet{fritzewski_rotation_2020} (F20), and
+  \citet{healy_stellar_2020} (HM20)}.
+  \replaced{I07 and F20}{I07, F20, and HM20} analyzed stars in the
+  core of \cn.
   \label{fig:litcomp}
@@
 \end{figure*}
+
+\explain{Figure~\ref{fig:litcomp} has been updated to include rotation
+periods from \citet{healy_stellar_2020}.}
 
@@
 includes, in order of precedence: rotation periods from our analysis,
-from ``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
-from the rotators in \citet{Irwin_NGC2516_2007}.  The main
+from \replaced{``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
+from the rotators in \citet{Irwin_NGC2516_2007}}{``Class 1 and
+2'' \citet{fritzewski_rotation_2020} sources, from
+Table 1 of \citet{healy_stellar_2020}, and
+from \citet{Irwin_NGC2516_2007}.}.  The main
 contribution of our study beyond these previous analyses is our
@@
 
-The lower panels of Figure~\ref{fig:rot} show our attempt at
-identifying unresolved binaries in the \cn\ rotation sample.  Binaries
-resolved by Gaia were already excluded in our initial definitions of
-the samples.  In \cn\, we see that the fraction of stars showing signs
-of binarity in the slow and fast rotation subsamples are 30\% (63/210)
-and 56\% (40/71); the fast rotators show a preference for binarity.
-This correlation is in line with the earlier findings noted above.
-While many rapid rotators in \cn\ (34\%) have no evidence for being
-binaries, the cuts we used to select binaries (${\rm
-RUWE}>1.2$; photometric excess $>$0.3\,mag) leave open a
-significant fraction of parameter space.  A 0.3\,mag flux excess, for
-instance, corresponds to mass ratios $M_2/M_1 \gtrsim 0.700$, assuming
-the luminosity $L$ scales as $M^{3.5}$.
+The lower panels of Figure~\ref{fig:rot} show our attempt at identifying
+unresolved binaries in the \cn\ rotation sample.  Binaries resolved by
+Gaia were already excluded in our initial definitions of the samples.  In
+\cn\, we see that the fraction of stars showing signs of binarity in the
+slow and fast rotation subsamples are \replaced{30\% (63/210) and 56\%
+(40/71)}{26\% (106/289) and 51\% (68/134)}; the fast rotators show a
+preference for binarity.  This correlation is in line with the earlier
+findings noted above.  While many rapid rotators in \cn\
+(\replaced{34\%}{66\%}) have no evidence for being binaries, the cuts we
+used to select binaries (${\rm RUWE}>1.2$; photometric excess
+$>$0.3\,mag) leave open a significant fraction of parameter space.  A
+0.3\,mag flux excess, for instance, corresponds to mass ratios $M_2/M_1
+\gtrsim 0.70$, assuming the luminosity $L$ scales as $M^{3.5}$.
 

### 2021-07-01 cluster_paper/collect.sh

@@ -6,10 +6,9 @@ cp ../results/vtangential_projection/XYZ_with_orbit.pdf f2.pdf
 cp ../results/hr/hr_Bp_m_Rp_fullfaint_edr3.pdf f3a.pdf
 cp ../results/hr/hr_parsec_Bp_m_Rp_parsec_fullfaint_edr3.pdf f3b.pdf
 
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_linear_defaultcleaning.pdf f4a.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_linear_periodogram_match.pdf f4b.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_talkaspect_linear_defaultcleaning.pdf f5a.pdf
-cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_talkaspect_linear_periodogram_match.pdf f5b.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_showPleiadesQuad_linear_defaultcleaning.pdf f4a.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_corehalosplit_talkaspect_linear_defaultcleaning_cutProtColor.pdf f4b.pdf
+cp ../results/rotation/NGC_2516/NGC_2516_rotation_vs_BpmRp_emphbinaries_talkaspect_linear_defaultcleaning_cutProtColor.pdf f5.pdf
 
 cp ../results/physical_X_rotation/physical_X_rotation_fullfaint.pdf f6a.pdf
 cp ../results/physical_X_rotation/histogram_physical_X_rotation_fullfaint.pdf f6b.pdf

### 2021-07-01 cluster_paper_review/collect_figures.sh

@@ -1,0 +1,1 @@
+cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear_corehalosplit.png f13_corehalosplit.png

### 2021-07-01 drivers/cluster_rotation/plot_autorotation.py

@@ -24,6 +24,25 @@ for runid, _ in RUNID_EXTINCTION_DICT.items():
     if not os.path.exists(PLOTDIR):
         os.mkdir(PLOTDIR)
 
+    c = 'defaultcleaning_cutProtColor'
+    ep.plot_auto_rotation(
+        PLOTDIR, runid, E_BpmRp, core_halo=1, yscale='linear', cleaning=c,
+        emph_binaries=0, talk_aspect=1, showPleiadesQuad=0
+    )
+    ep.plot_auto_rotation(
+        PLOTDIR, runid, E_BpmRp, core_halo=0, yscale='linear', cleaning=c,
+        emph_binaries=1, talk_aspect=1, showPleiadesQuad=0
+    )
+
+    for c in ['defaultcleaning', 'harderlsp', 'periodogram_match']:
+        for e in [0,1]:
+            ep.plot_auto_rotation(
+                PLOTDIR, runid, E_BpmRp, core_halo=1, yscale='linear', cleaning=c,
+                emph_binaries=e, talk_aspect=1, showPleiadesQuad=1
+            )
+
+    assert 0
+
     # pleiades and praesepe only
     ep.plot_auto_rotation(
         PLOTDIR, runid, E_BpmRp, core_halo=0,

### 2021-07-01 drivers/plot_compstar_rotation.py

@@ -7,4 +7,5 @@ if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
 for yscale in ['linear', 'log']:
+    ep.plot_compstar_rotation(PLOTDIR, yscale=yscale, corehalosplit=1)
     ep.plot_compstar_rotation(PLOTDIR, yscale=yscale)

### 2021-07-01 drivers/write_NGC2516_main_table.py

@@ -30,7 +30,7 @@ assert len(df) == N_members
 from earhart.priors import AVG_EBpmRp
 df['(Bp-Rp)_0'] = df.phot_bp_mean_mag - df.phot_rp_mean_mag - AVG_EBpmRp
 
-cleanings = ['defaultcleaning', 'periodogram_match', 'match234_alias']
+cleanings = ['defaultcleaning', 'defaultcleaning_cutProtColor', 'periodogram_match', 'match234_alias']
 
 # add the "in_{CLEANING}" columns
 for c in cleanings:
@@ -124,10 +124,11 @@ if include_lithium:
 
 
 df = df.drop(['in_match234_alias'], axis=1)
+df = df.drop(['in_periodogram_match'], axis=1)
 
 df = df.rename(columns={
     'in_defaultcleaning': 'in_SetA',
-    'in_periodogram_match': 'in_SetB',
+    'in_defaultcleaning_cutProtColor': 'in_SetB',
     'is_phot_binary': 'is_phot_bin',
     'is_astrometric_binary': 'is_astrm_bin'
 })
@@ -169,7 +170,7 @@ df['nfaint'] = df['nfaint'].astype(pd.Int32Dtype())
 
 orderedcols = [
 'source_id', 'source_id_edr3', 'in_SetA', 'in_SetB', 'n_cdips_sector',
-'period', 'lspval', 'spdmperiod', 'spdmval', 'nequal', 'nclose', 'nfaint',
+'period', 'lspval', 'nequal', 'nclose', 'nfaint',
 'ra', 'dec', 'ref_epoch', 'parallax', 'parallax_error', 'pmra', 'pmdec',
 'phot_g_mean_mag', 'phot_bp_mean_mag', 'phot_rp_mean_mag', 'radial_velocity',
 'radial_velocity_error', 'subcluster', 'in_CG18', 'in_KC19', 'in_M21',
@@ -180,58 +181,55 @@ orderedcols = [
 
 df = df[orderedcols]
 
-outpath = os.path.join(RESULTSDIR, 'tables', 'NGC_2516_Prot_cleaned.csv')
-df.to_csv(outpath, index=False)
-print(f'Wrote {outpath}')
-
 #
 # make the header table for the paper
 #
 namedesc_dict = {
-'source_id': "Gaia DR2 source identifier.",
-'source_id_edr3': "Gaia EDR3 source identifier.",
[…]

### 2021-07-01 earhart/helpers.py

@@ -13,6 +13,7 @@ get_gaia_basedata
     _get_median_ngc2516_core_params
 get_denis_xmatch
 append_phot_binary_column
+PleaidesQuadProtModel
 """
 import os, collections, pickle
 import numpy as np, pandas as pd
@@ -795,11 +796,14 @@ def get_autorotation_dataframe(runid='NGC_2516', verbose=1, returnbase=0,
     if runid == 'NGC_2516':
         df = append_phot_binary_column(df)
 
-    if cleaning in ['defaultcleaning', 'periodogram_match', 'match234_alias']:
+    if cleaning in ['defaultcleaning', 'periodogram_match',
+                    'match234_alias','harderlsp', 'defaultcleaning_cutProtColor']:
         # automatic selection criteria for viable rotation periods
         NEQUAL_CUTOFF = 0 # could also do 1
         NCLOSE_CUTOFF = 1
         LSP_CUTOFF = 0.08 # 0.08 standard
+        if cleaning == 'harderlsp':
+            LSP_CUTOFF = 0.15
         sel = (
             (df.period < 15)
             &
@@ -809,6 +813,16 @@ def get_autorotation_dataframe(runid='NGC_2516', verbose=1, returnbase=0,
             &
             (df.nclose <= NCLOSE_CUTOFF)
         )
+        if cleaning == 'defaultcleaning_cutProtColor':
+            from earhart.priors import AVG_EBpmRp
+            BpmRp0 = (
+                df['phot_bp_mean_mag'] - df['phot_rp_mean_mag'] - AVG_EBpmRp
+            )
+            Prot_boundary = PleaidesQuadProtModel(BpmRp0)
+            sel &= (
+                df.period < Prot_boundary
+            )
+
     elif cleaning in ['nocleaning']:
         NEQUAL_CUTOFF = 99999
         NCLOSE_CUTOFF = 99999
@@ -819,10 +833,8 @@ def get_autorotation_dataframe(runid='NGC_2516', verbose=1, returnbase=0,
     else:
         raise ValueError(f'Got cleaning == {cleaning}, not recognized.')
 
-    if cleaning == 'defaultcleaning':
-        pass
-
-    elif cleaning == 'nocleaning':
+    if cleaning in ['defaultcleaning', 'nocleaning', 'harderlsp',
[…]

### 2021-07-01 earhart/plotting.py

@@ -1146,7 +1146,8 @@ def plot_skypositions_x_rotn(outdir):
 
 def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
                        cleaning=None, emph_binaries=False,
-                       refcluster_only=False, talk_aspect=0, xval_absmag=0):
+                       refcluster_only=False, talk_aspect=0, xval_absmag=0,
+                       showPleiadesQuad=0):
     """
     Plot rotation periods that satisfy the automated selection criteria
     (specified in helpers.get_autorotation_dataframe)
@@ -1172,7 +1173,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
     lws = [0., 0.1, 0.1]
     mews= [0., 0.5, 0.5]
     _s = 3 if runid != 'VelaOB2' else 1.2
-    ss = [15, 12, 10]
+    ss = [15, 12, 8]
     labels = ['Pleaides', 'Praesepe', f'{runid}']
 
     # plot vals
@@ -1259,7 +1260,7 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
                     xval,
                     df[ykey],
                     c=_col, alpha=1, zorder=z, s=s, edgecolors='k',
-                    marker=m, linewidths=_lw, label=f"{l.replace('_','')}"
+                    marker=m, linewidths=_lw, label=f"{l.replace('_',' ')}"
                 )
 
         if emph_binaries and f'{runid}' in _cls:
@@ -1300,6 +1301,17 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
                 c='red', alpha=1, zorder=9, s=s, edgecolors='k',
                 marker='o', linewidths=_lw, label="Astrometric binary"
             )
+
+    if showPleiadesQuad:
+        # 0.6 to 1.3 is actual ok range. from Curtis+2020
+        xmod = np.arange(0.3, 2.4+0.01, 0.01)
+        from earhart.helpers import PleaidesQuadProtModel
+        Protmod = PleaidesQuadProtModel(xmod)
+
+        ax.plot(
[…]

### 2021-07-02 cluster_paper/ms.tex
update ms

@@
 
-%\documentclass[12pt,twocolumn,tighten]{aastex63}
-\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
+\documentclass[12pt,twocolumn,tighten]{aastex63}
+%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
@@
 \newcommand{\nautorotnumeratormatching}{892} % plot_autorotation.py, Set A, with manual Prot-color cut
-\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
-\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
-\newcommand{\ncompfrac}{38.0\%} % plot_compstar_rotation.py, based on plot_autorotation
-\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
-\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
-\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py
+
+%\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
+%\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
+
+\newcommand{\ncompstardenominator}{656} % plot_compstar_rotation.py BpmRp0 from 0.4 to 0.8
+%\newcommand{\ncompstarnumerator}{223} % plot_compstar_rotation.py, ditto, with Prot detected in field star sample
+\newcommand{\ncompstarnumerator}{107} % plot_compstar_rotation.py, ditto, with Prot meeting 'set B' criteria in field star sample
+\newcommand{\ncompfrac}{16.3\%} % plot_compstar_rotation.py, based on plot_autorotation for defaultcleaning_cutProtColor
+
+\newcommand{\nautovscompstardenominator}{327} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
+\newcommand{\nautovscompstarnumerator}{164} % plot_autorotation.py
+\newcommand{\nautofrac}{50.1\%} % plot_autorotation.py
 
@@
 \added{To assess whether the stars in the cluster's halo are the same
-ages as stars in the cluster's core, we use three different age-dating
+ages as stars in the cluster's core, we apply three different age-dating
 techniques: isochrones, gyrochrones, and lithium depletion
@@
 
-\begin{figure*}[tp]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=\textwidth]{f1.pdf}
+		\includegraphics[width=0.9\textwidth]{f1.pdf}
 	\end{center}
@@
   \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
-  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}
+  made by \citet{meingast_2021}, last accessed \texttt{2021/07/01}.}
 %While we considered performing our own clustering analysis on the
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
+  \caption{ {\bf TESS rotation periods and dereddened Gaia colors
   across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
@@
 \end{figure*}
-
-\subsubsection{Methodology}
-\label{subsubsec:cluster}
-
-We began our rotation analysis by considering all \nkinematic\
-candidate members of \cn.  For each source, we retrieved all available
-CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
-stars with at least one sector from a CDIPS light curve.  Each of
-these stars is brighter than the CDIPS magnitude limit of
-$G_\mathrm{RP}=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
-source list have $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
-from \citet{meingast_2021} which were not available at the time of the
-CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
-distance of \cn, the $G_\mathrm{RP}=16$\,mag cutoff imposed during the CDIPS
-processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
-of $\approx$M2V.
-
-We removed systematic trends common to the light curves on each
-CCD in each individual sector, and stitched together the resulting
-light curves before searching for rotation-induced periodicity.
-Details regarding our detrending approach are presented in
-Appendix~\ref{app:detrending}.  After detrending, we proceeded with a
-few cleaning steps: we masked 0.7 days at the beginning and end of
-each spacecraft orbit, and ran a sliding standard-deviation rejection
-window over the light curve, which removed any outlying points within
-$\pm3\times$MAD of the median in each window.
-
-We then measured the rotation period from the resulting light curve using
-the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
-\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
-the Stellingwerf phase-dispersion minimization algorithm
-\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{\citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
-We used the CDIPS aperture radius that, based on theoretical
-expectations, was expected to give the optimal balance between light from
-the target star and the sky background \citep{Sullivan_2015}.  This
-typically resulted in a circular aperture radius of either 1 or 1.5
-pixels.  We recorded the top five periodogram peaks\deleted{ from each
-method}, and their corresponding powers.  Finally, as a check on
-crowding, we recorded the number of stars within the aperture with
-greater brightness than the target star, and with brightness within 1.25
-and 2.5 TESS magnitudes of the target star.
-
-The resulting rotation periods and periodogram powers are reported,
-regardless of detection significance, in Table~\ref{tab:maintable}.
-To clean these measurements, we designed 
-automated cleaning criteria, which yielded sets we denote
-$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
-$\mathcal{A}$}, we considered light
-curves for which the peak Lomb-Scargle periodogram period was below 15
-days, the normalized Lomb-Scargle power exceeded 0.08, and for which
-no companions with brightness exceeding one-tenth of the target star
-were in the aperture ({\it i.e.}, excluding visual binaries
-with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
-These limits were chosen after inspecting the light curves visually,
-to eliminate low-quality rotation periods while retaining high
-completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
-light curves, out of \nautorotdenominator\ stars for which rotation
-periods might plausibly have been detected.  For Set $\mathcal{B}$, we
-additionally required that the Lomb-Scargle \replaced{and Stellingwerf
-phase-dispersion methods yielded identical periods to within 10\%}{period
-fell below the boundary drawn in Figure~\ref{fig:rot}, which was
-constructed the filter using the Pleiades polynomial from
-\citet{curtis_rup147_2020}, with hard cuts at the red and blue ends.}.
-This yielded \nautorotnumeratormatching\ light curves\added{in Set
-$\mathcal{B}$}. 
-Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
-set of field stars was also analyzed for comparison.  The results are
-discussed in Appendix~\ref{app:compstar}.
-\added{The justification for the additional cut in Set $\mathcal{B}$ is
-that \cn\ is already known to have few if any stars above the slow
-sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
-are most likely therefore either not cluster members, or they are cluster
-members, but with unresolved binary companions contaminating the rotation
-period measurement \citep[{\it
-e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
-assessing which are rotationally consistent with being in the cluster,
-this is a viable filter; for the purpose of fully exploring whether any
-rotational outliers could be bonafide cluster members, Set $\mathcal{A}$
-would be the better cut.}
-
-\subsubsection{Rotation-Color Diagrams}
-
-\begin{figure*}[t!]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-    \includegraphics[width=0.700\textwidth]{f5.pdf}
+		\includegraphics[width=0.480\textwidth]{f5.pdf}
 	\end{center}
@@
 	}
-\end{figure*}
+\end{figure}
+
+\subsubsection{Methodology}
+\label{subsubsec:cluster}
+
+We began our rotation analysis by considering all \nkinematic\
+candidate members of \cn.  For each source, we retrieved all available
+CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
+stars with at least one sector from a CDIPS light curve.  Each of
+these stars is brighter than the CDIPS magnitude limit of
+$G_\mathrm{RP}=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
+source list have $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
+from \citet{meingast_2021} which were not available at the time of the
+CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
+distance of \cn, the $G_\mathrm{RP}=16$\,mag cutoff imposed during the CDIPS
+processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
+of $\approx$M2V.
+
+We removed systematic trends common to the light curves on each
+CCD in each individual sector, and stitched together the resulting
+light curves before searching for rotation-induced periodicity.
+Details regarding our detrending approach are presented in
+Appendix~\ref{app:detrending}.  After detrending, we proceeded with a
+few cleaning steps: we masked 0.7 days at the beginning and end of
+each spacecraft orbit, and ran a sliding standard-deviation rejection
+window over the light curve, which removed any outlying points within
+$\pm3\times$MAD of the median in each window.
+
+We then measured the rotation period from the resulting light curve using
+the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
+\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
+the Stellingwerf phase-dispersion minimization algorithm
+\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{ \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
+We used the CDIPS aperture radius that, based on theoretical
+expectations, was expected to give the optimal balance between light from
+the target star and the sky background \citep{Sullivan_2015}.  This
+typically resulted in a circular aperture radius of either 1 or 1.5
+pixels.  We recorded the top five periodogram peaks\deleted{ from each
+method}, and their corresponding powers.  Finally, as a check on
+crowding, we recorded the number of stars within the aperture with
+greater brightness than the target star, and with brightness within 1.25
+and 2.5 TESS magnitudes of the target star.
+
+The resulting rotation periods and periodogram powers are reported,
+regardless of detection significance, in Table~\ref{tab:maintable}.
+To clean these measurements, we designed 
+automated cleaning criteria, which yielded sets we denote
+$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
+$\mathcal{A}$}, we considered light
+curves for which the peak Lomb-Scargle periodogram period was below 15
+days, the normalized Lomb-Scargle power exceeded 0.08, and for which
+no companions with brightness exceeding one-tenth of the target star
+were in the aperture ({\it i.e.}, excluding visual binaries
+with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
+These limits were chosen after inspecting the light curves visually,
+to eliminate low-quality rotation periods while retaining high
+completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
+\replaced{light curves}{stars}, out of \nautorotdenominator\ stars \replaced{for which rotation
+periods might plausibly have been detected}{that met the initial crowding requirements}.  For Set $\mathcal{B}$, we
+additionally required that the Lomb-Scargle \replaced{and Stellingwerf
+phase-dispersion methods yielded identical periods to within 10\%}{period
+fell below the boundary drawn in Figure~\ref{fig:rot}, which was
+constructed using the Pleiades polynomial from
+\citet{curtis_rup147_2020}}.
+This yielded \nautorotnumeratormatching\ light curves. 
+Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
+set of field stars was also analyzed for comparison.  The results are
+discussed in Appendix~\ref{app:compstar}.
+\added{The rationale behind the additional cut in Set $\mathcal{B}$ is
+that \cn\ is already known to have few if any stars above the slow
+sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
+are therefore most likely either not cluster members, or they are cluster
+members, but with unresolved binary companions contaminating the rotation
+period measurement \citep[{\it
+e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
+assessing which stars are rotationally consistent with being in the cluster,
+this is a viable filter; for the purpose of exploring whether any
+rotational outliers might be bonafide cluster members, Set $\mathcal{A}$
+would be the better cut.}
+
+\subsubsection{Rotation-Color Diagrams}
 
@@
   the {\it top right}, the 
-  ratio of detected to expected stars with rotation periods is shown
+  ratio of detected to expected stars with rotation periods \added{in Set $\mathcal{B}$} is shown
   versus 3D separation from
@@
 pipeline succeeded in making light curves.  In other words, the stars in
-the base sample are those for which we could have plausibly detected a
+the base sample are those for which we would have expected to detect a
 rotation period.  This color range is preferred because the the slow
@@
 \added{consistent with a gyrochronological age of 150\,Myr} are detected
-for \replaced{66\%}{68\%} of stars \replaced{at least 25\,pc from the
+for \replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc from the
 cluster core}{reported to be in the cluster}.  The average detection rate
-inside of 25\,pc, \replaced{73\%}{72\%}, is slightly higher\added{ than
-outside of 25\,pc, where the detection rate is 63\%}. Within
-$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\%} is
-relatively low, due to crowding and saturation in the TESS images.
+inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is slightly higher\added{ than
+outside of 25\,pc, where the detection rate is 63\% (181/288)}. Within
+$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\% (26/44)} is
+\replaced{relatively}{anomalously} low, due to crowding and saturation in the TESS images.
 
@@
 
-Even with this level of contamination, our rotation period
+\replaced{Even with}{Despite} this level of contamination, the rotation period
 measurements show that the halo extends to separations of
@@
 		\leavevmode
-			\includegraphics[width=0.9\textwidth]{f7.pdf}
+			\includegraphics[width=0.95\textwidth]{f7.pdf}
 	\end{center}
@@
   \label{fig:lithiumcorehalo}
-  }
-\end{figure*}
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f8.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
-  candidate \cn\ members with TESS rotation periods and Gaia-ESO or
-  GALAH spectra, and are colored by equivalent width.  Points with
-  reported equivalent widths below 20\,m\AA\ are shown with gray
-  crosses.  The lithium-rotation correlation for the K-dwarfs has been
-  observed in other young clusters, and is discussed in
-  Section~\ref{discussion:lithium}.
-	\label{fig:lithiumrot}
   }
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f8.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
+		candidate \cn\ members with TESS rotation periods\added{ (Set $\mathcal{A}$)} and Gaia-ESO or
+		GALAH spectra, and are colored by equivalent width.  Points with
+		reported equivalent widths below 20\,m\AA\ are shown with gray
+		crosses.  The lithium-rotation correlation for the K-dwarfs has been
+		observed in other young clusters, and is discussed in
+		Section~\ref{discussion:lithium}.
+		\label{fig:lithiumrot}
+	}
+\end{figure*}
+
 We can go one step further, and match our sample of kinematically
@@
 Table 1 of \citet{healy_stellar_2020}, and
-from \citet{Irwin_NGC2516_2007}.}.  The main
+from \citet{Irwin_NGC2516_2007}}.  The main
 contribution of our study beyond these previous analyses is our
@@
 dispersion in rotation periods increases.  The maximum rotation
-periods seen in \cn\ across both Sets $\mathcal{A}$ and $\mathcal{B}$
+periods seen in \cn\ \deleted{across both Sets $\mathcal{A}$ and
+$\mathcal{B}$}
 extend up to $\approx$11 days, rather than the $\approx$8.5 day upper
@@
     separations of $\approx$200\,pc, this fraction decreases to
-    $\approx$50\%.  Given the number density ratios of cluster to
-    field stars at the widest separations ($\approx 10^{-3}$,
-    \citealt{meingast_2021}), these contamination fractions are not as
-    bad as one might fear.
+    $\approx$50\%.
+\replaced{Given the number density ratios of cluster to field stars at the widest separations ($\approx 10^{-3}$, \citealt{meingast_2021}), these contamination fractions are not as bad as one might fear.}{While the latter contamination fraction may seem high, a comparison sample of field stars (Appendix~\ref{app:compstar}) has a detection rate three times lower, and a  distribution in the period-color plane inconsistent with that of a population with a single age.}
    %
@@
   the kinematically selected \cn\ members (black).  The field and \cn\
-  stars are both subject to the same ``Set $\mathcal{A}$''
+  stars are both subject to the same \deleted{``Set $\mathcal{A}$''}
   period-measurement and cleaning procedures.
+  \added{The dashed line is as in Figure~\ref{fig:rot}}.
   \label{fig:compstar}
@@
 
+\explain{Numbers in this paragraph have been updated per the revised
+definition of Set $\mathcal{B}$. }
+
 We performed the same light curve stitching and period-search
 procedure discussed in the text on the field stars.
-Imposing the same requirements for crowding resulted in
-\ncompstardenominator\ stars for which rotation periods could have
-been detected.  Imposing the same Lomb Scargle power cutoff (${\rm
-LSP}>0.08$) and period upper limit ($P<15\,{\rm days}$) yielded
-\ncompstarnumerator\ period detections (\ncompfrac).  Within the same
-brightness cutoff, \nautovscompstarnumerator\ of
-\nautovscompstardenominator\ cluster stars yielded period detections
-(\nautofrac).  While the detection fractions are only mildly
-different (likely due to the brightness cutoff), the period--color
-distributions are very different
-(Figure~\ref{fig:compstar}).
-This confirms our expectation that our period measurement procedure is
-not biased in any way that could produce the features we observe in
-\cn.
+\replaced{Imposing}{Within $0.4<\bpmrpo<0.8$}, the same requirements
+for crowding resulted in \ncompstardenominator\ \added{field} stars
+for which rotation periods could have been detected.  Imposing the
+same Lomb Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
+upper limit ($P<15\,{\rm days}$)}{color-period cutoff as that used to
+define Set $\mathcal{B}$} yielded \ncompstarnumerator\ period
+detections (\ncompfrac).  Within the same \replaced{brightness
+cutoff}{color range}, \nautovscompstarnumerator\ of
+\nautovscompstardenominator\ \replaced{cluster stars}{kinematically
+identified candidate cluster members} yielded period detections
+(\nautofrac).  \replaced{While the detection fractions are only mildly different
+(likely due to the brightness cutoff), the }{The }period--color distributions
+are \replaced{very}{also quite} different\added{: the candidate
+cluster members show an overdensity coincident with the expected
+gyrochronal age, while the field stars show a much broader rotation
+period distribution} (Figure~\ref{fig:compstar}).  This confirms our
+expectation that our period measurement procedure is not biased in any
+way that could produce the features we observe in \cn.
 

### 2021-07-02 cluster_paper/collect.sh

@@ -25,7 +25,7 @@ cp ../results/litcomp_ngc2516/litcomp_ngc2516.pdf f10.pdf
 # appendix
 cp ../results/venn/venn.pdf f11.pdf
 cp ../results/lightcurves_rotators/lightcurves_rotators_color.pdf f12.pdf
-cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear.pdf f13.pdf
+cp ../results/compstar_rotation/compstar_rotation_NGC_2516_vs_BpmRp_linear_showPleiadesQuad.pdf f13.pdf
 
 cp ../results/vtangential_projection/skymap_tangential_velocity_pmdec.pdf f14a.pdf
 cp ../results/vtangential_projection/skymap_tangential_velocity_pmra_cosdec.pdf f14b.pdf

### 2021-07-02 drivers/cluster_rotation/plot_autorotation.py

@@ -4,8 +4,8 @@ from earhart.paths import RESULTSDIR
 from earhart.priors import AVG_EBpmRp
 
 RUNID_EXTINCTION_DICT = {
-#    'compstar_NGC_2516': 0.1343, # 0.25 for KC19
-    'NGC_2516': AVG_EBpmRp, # 0.25 for KC19
+    #'compstar_NGC_2516': 0.1343, # 0.25 for KC19
+     'NGC_2516': AVG_EBpmRp, # 0.25 for KC19
 #    'IC_2602': 0.0799,  # avg E(B-V) from Randich+18, *1.31 per Stassun+2019
 #    'CrA': 0.05025, # AV_to_EBpmRp, using A_V=0.11892 from KC19
 #    'kc19group_113': 0.1386, # 0.258 from KC19 -- take ratio
@@ -40,8 +40,6 @@ for runid, _ in RUNID_EXTINCTION_DICT.items():
                 PLOTDIR, runid, E_BpmRp, core_halo=1, yscale='linear', cleaning=c,
                 emph_binaries=e, talk_aspect=1, showPleiadesQuad=1
             )
-
-    assert 0
 
     # pleiades and praesepe only
     ep.plot_auto_rotation(

### 2021-07-02 drivers/plot_compstar_rotation.py

@@ -6,6 +6,16 @@ PLOTDIR = os.path.join(RESULTSDIR, 'compstar_rotation')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-for yscale in ['linear', 'log']:
-    ep.plot_compstar_rotation(PLOTDIR, yscale=yscale, corehalosplit=1)
-    ep.plot_compstar_rotation(PLOTDIR, yscale=yscale)
+#for yscale in ['linear', 'log']:
+for yscale in ['linear']:
+    for cleaning in [
+        'defaultcleaning','defaultcleaning_cutProtColor'
+    ]:
+        ep.plot_compstar_rotation(
+            PLOTDIR, yscale=yscale, corehalosplit=1,
+            cleaning=cleaning
+        )
+        ep.plot_compstar_rotation(
+            PLOTDIR, yscale=yscale, showPleiadesQuad=1,
+            cleaning=cleaning
+        )

### 2021-07-02 earhart/plotting.py

@@ -1387,7 +1387,8 @@ def plot_auto_rotation(outdir, runid, E_BpmRp, core_halo=0, yscale='linear',
 
 
 def plot_compstar_rotation(outdir, E_BpmRp=AVG_EBpmRp, yscale=None,
-                           corehalosplit=0):
+                           corehalosplit=0, showPleiadesQuad=0,
+                           cleaning='defaultcleaning'):
     """
     Plot rotation periods that satisfy the automated selection criteria
     (specified in helpers.get_autorotation_dataframe)
@@ -1432,7 +1433,7 @@ def plot_compstar_rotation(outdir, E_BpmRp=AVG_EBpmRp, yscale=None,
                 raise NotImplementedError
 
         else:
-            df = get_autorotation_dataframe(_cls, cleaning='defaultcleaning')
+            df = get_autorotation_dataframe(_cls, cleaning=cleaning)
             df = df[df.phot_rp_mean_mag < 13]
             print(42*'-')
             print(f'Applying E(Bp-Rp) = {E_BpmRp:.4f}')
@@ -1469,8 +1470,14 @@ def plot_compstar_rotation(outdir, E_BpmRp=AVG_EBpmRp, yscale=None,
                 |
                 ( (df.period > window1[0]) & (df.period < window1[1]) )
             )
-            print(f'Before adhoc window puring, N={len(df)} field rot')
-            print(f'After adhoc window puring, N={len(df[sel0])} field rot')
+
+            BpmRp0 = (
+                df['phot_bp_mean_mag'] - df['phot_rp_mean_mag'] - AVG_EBpmRp
+            )
+            bpmrp_sel = (BpmRp0 < 0.8) & (BpmRp0 > 0.4)
+
+            print(f'Before adhoc window puring, N={len(df[bpmrp_sel])} field rot')
+            print(f'After adhoc window puring, N={len(df[sel0 & bpmrp_sel])} field rot')
             xval = xval[sel0]
             yval = yval[sel0]
 
@@ -1490,6 +1497,16 @@ def plot_compstar_rotation(outdir, E_BpmRp=AVG_EBpmRp, yscale=None,
                 xval[sel], yval[sel], c='lightskyblue', alpha=0.9, zorder=z, s=8,
[…]

### 2021-07-02 cluster_paper_review/referee_response_v0.txt
refere response ready for revision

@@
 
-We thank you for organizing the review of our paper, and also
-thank the referee for constructive criticism.  We have considered
-the referee's remarks carefully and revised our manuscript
-accordingly.
+We thank you for organizing the review of our paper, and also thank
+the referee for constructive criticism.  We have considered the
+referee's remarks carefully and revised our manuscript accordingly.
 
@@
 report, alongside our responses. A list of changes is appended to
-the manuscript using the trackchanges AASTeX macros. Figures
-supporting X AND Y WORDS TODO FIX THESE have also been uploaded for inspection.
+the manuscript using the trackchanges AASTeX macros. A figure
+expanding on the comparison between the field and cluster star
+rotation periods has also been uploaded for inspection
+(f13_corehalosplit.png).
 
@@
 ----------
-At this stage the referee is more inclined to believe that the authors have
-identified a set of young stars in that region of the sky than actual halo
-member stars of NGC 2516.
+At this stage the referee is more inclined to believe that the
+authors have identified a set of young stars in that region of the
+sky than actual halo member stars of NGC 2516.
 
@@
 ----------
-> We agree that the stars with rotation periods similar to those of the
-> Pleiades, and showing Li EWs similar to those of the Pleiades, should indeed
-> be considered to have comparable ages to stars in Pleiades.
+> We agree that the stars with rotation periods similar to those of
+> the Pleiades, and showing Li EWs similar to those of the Pleiades,
+> should indeed be considered to have comparable ages as stars in
+> Pleiades.
 >
-> Figures 4 and 7 show that the "halo" stars from KC19 and M21 overlap both
-> with the Pleiades, and with stars in the core of the cluster (those from
-> CG18).  They share the same gyrochronological and Li tracks, and therefore
-> share the same age to within measurement precision.  The stars in this "halo"
-> population are also mostly within 2 km/s of tangential velocity separation
-> from the main core population (see panel in top left of Figure 6).
-> Spatially, they lead and lag the core of the population, in a similar manner
-> to that seen for eight other open clusters by Meingast et al 2021 (Figure 2
-> of our manuscript; Figure 8 of Meingast et al 2021, Astronomy & Astrophysics,
-> Volume 645, id.A84).
+> Figures 4 and 7 show that the stars from KC19 and M21 (the "halo")
+> overlap both with the Pleiades, and also with stars in the core of
+> the cluster (those from CG18).  They share the same
+> gyrochronological and Li tracks, and therefore share the same age
+> to within measurement precision.  The stars in this "halo"
+> population are also mostly within 2 km/s of tangential velocity
+> separation from the main core population (see panel in top left of
+> Figure 6).  Spatially, they lead and lag the core of the cluster
+> population, in a similar manner to that seen for eight other open
+> clusters by Meingast et al 2021 (Figure 2 of our manuscript;
+> Figure 8 of Meingast et al 2021, Astronomy & Astrophysics, Volume
+> 645, id.A84).
 >
-> Despite these lines of evidence for the halo stars being associated with NGC
-> 2516 (same ages; same tangential velocities; spatial correlation with the
-> cluster center), for the sake of caution we analyzed rotation periods of
-> randomly draw field stars in this region of sky.  This exercise was described
-> in Appendix C, and the results are shown in Figure 13 of the manuscript.  The
-> randomly selected field stars (orange points) does not show any evidence for
-> an overdensity of Pleiades-age stars.  The candidate "halo" stars selected
-> kinematically by CG18, KC19, and M21 do (black points).  As a clarification
-> on Figure 13, we have attached "f13_corehalosplit.png" as a supplement in the
-> response, which shows the same data, but splitting NGC 2516 into halo and
-> core stars.  Stars in the halo (blue) have a very different rotation-color
-> distribution than randomly selected field stars (orange).
+> Despite these lines of evidence for the halo stars being
+> associated with NGC 2516 -- same ages; same tangential velocities;
+> spatial correlation with the cluster center --  for the sake of
+> caution we analyzed rotation periods of randomly draw field stars
+> in this region of sky.  This exercise was described in Appendix C,
+> and the results are shown in Figure 13 of the manuscript.  The
+> randomly selected field stars (orange points) does not show any
+> evidence for an overdensity of Pleiades-age stars.  Their period
+> detection rates within the relevant ~150 Myr gyrochrone regime are
+> also three times lower than the kinematically selected sample (see
+> added text in Appendix C).   As a clarification on Figure 13, we
+> have attached "f13_corehalosplit.png" as a supplement in the
+> response, which shows the same data, but splitting NGC 2516 into
+> halo and core stars.  Stars in the halo (blue) have a very
+> different rotation-color distribution than randomly selected field
+> stars (orange).
 >
-> In short, most of the halo stars i) have similar tangential velocites as core
-> stars, ii) are coeval with other halo stars, iii) have the same age as the
-> core stars, iv) are spatially correlated with the location of the core stars,
-> and iv) have a different rotation-color distribution from a randomly selected
-> sample of fields stars (with a large overdensity of rotators falling on the
-> same sequence as the core of NGC 2516).
+> In short, most of the halo stars i) have similar tangential
+> velocites as core stars, ii) are coeval with other halo stars,
+> iii) have the same age as the core stars, iv) are spatially
+> correlated with the location of the core stars, and v) have a
+> different rotation-color distribution from a randomly selected
+> sample of fields stars (with a large overdensity of rotators
+> falling on the same sequence as the core of NGC 2516).
 >
-> The suggestion that we "have identified a set of young stars in that region
-> of the sky [rather] than actual halo member stars of NGC 2516." is equivalent
-> to asserting that the clustering techniques used by KC19 and M21 yielded a
-> set of stars that is identical to the cluster core according to all age
-> measurements, and is spatially and kinematically correlated with the cluster
-> core, but are otherwise unrelated.  While the spatial and kinematic
-> correlations are tautological -- this is what clustering methods do if they
-> are working correctly -- the age measurements have no a priori reason to be
-> identical between the core and halo unless the two are in fact related.
-
-> TODO: maybe empahsize the false positive fraction of 50% here.
-
-> Add an uncertainty on the false positive rate? Could it be higher than 50%?
+> The suggestion that we "have identified a set of young stars in
+> that region of the sky [rather] than actual halo member stars of
+> NGC 2516." is equivalent to asserting that the clustering
+> techniques used by KC19 and M21 yielded a set of stars that is
+> identical to the cluster core according to all age measurements,
+> and is spatially and kinematically correlated with the cluster
+> core, but is otherwise unrelated.  While the spatial and kinematic
+> correlations are tautological -- this is what clustering methods
+> should do if they are working correctly -- the age measurements
+> have no a priori reason to be identical between the core and halo
+> unless the two populations are in fact related.
 
@@
 ----------
-The authors begin by concatenating a very large collection of stars identified by
-CG18 (which they term the core members), and those of KC19 and M21, the last two
-having very relaxed criteria for association with NGC 2516.
-At this stage, as one can see from the distributions in Fig.1, the sample is
-completely bloated with field stars which now clearly have to be excised.
-The question is how?
+The authors begin by concatenating a very large collection of stars
+identified by CG18 (which they term the core members), and those of
+KC19 and M21, the last two having very relaxed criteria for
+association with NGC 2516.  At this stage, as one can see from the
+distributions in Fig.1, the sample is completely bloated with field
+stars which now clearly have to be excised.  The question is how?
 
-The authors justifiably look first to the HR diagram and understandably find
-that that is not very helpful in rejecting non-members because almost all stars in
-the extended sample lie close to the ridge line of the canonical cluster members.
-(Nevertheless, this is already a basis for rejecting stars below the main sequence
-and those significantly above it, including giants. The authors should consider
-this.) Incidentally, the Gaia photometry for the brightest stars is likely not good
-enough for fitting any isochrones of this age, so in this referee's estimation, the
-cluster turnoff is perhaps still the best one can do, keeping in mind the persistence
-of photometric problems for bright stars.
+The authors justifiably look first to the HR diagram and
+understandably find that that is not very helpful in rejecting
+non-members because almost all stars in the extended sample lie
+close to the ridge line of the canonical cluster members.
+(Nevertheless, this is already a basis for rejecting stars below the
+main sequence and those significantly above it, including giants.
+The authors should consider this.) Incidentally, the Gaia photometry
+for the brightest stars is likely not good enough for fitting any
+isochrones of this age, so in this referee's estimation, the cluster
+turnoff is perhaps still the best one can do, keeping in mind the
+persistence of photometric problems for bright stars.
 
@@
 ----------
-> While the HR diagram is not very helpful in rejecting non-members, it does
-> indicate that the distribution of field stars in this region of
-> ra/dec/parallax is different from stars in both the core and halo of the
-> cluster.  In particular, the lack of a sub-giant branch or red giant branch
-> (e.g., at G_BP-G_RP =1.2, M_G =3.0) suggests that the core and halo have a
-> different age distribution than the field stars, and supports the suggestion
-> that for the most part, the halo population has a single age.
+> While we agree the HR diagram is not very helpful in rejecting
+> non-members, it does indicate that the distribution of field stars
+> in this region of ra/dec/parallax is different from stars in both
+> the core and halo of the cluster.  In particular, the lack of a
+> sub-giant branch or red giant branch (e.g., at G_BP-G_RP =1.2, M_G
+> =3.0)  in the core and halo suggests that both populations have a
+> different age distribution than the field stars, and is consistent
+> with the suggestion that the halo population has a single age.
 >
-> For the ~five halo stars lying significantly above the main sequence, we
-> agree that these must be field interlopers.  This is noted at the end of
-> Section 3.1, Paragraph 2.  The ~four halo stars, and ~four core stars lying
-> well below the main sequence could be there due to poorly determined
-> parallaxes, or due to blending in the Gaia BP or RP spectra with unresolved
-> white dwarf companions.  We do not consider their presence below the MS to
-> support their exclusion from consideration as cluster members.
+> For the ~five halo stars lying significantly above the main
+> sequence, we agree that these must be field interlopers.  This is
+> noted at the end of Section 3.1, Paragraph 2.  The ~four halo
+> stars, and ~four core stars lying well below the main sequence
+> could have poorly determined parallaxes, or could be blended in
+> the Gaia BP or RP spectra with unresolved white dwarf companions.
+> We do not consider their presence below the MS to support their
+> exclusion from consideration as cluster members.
 
@@
 ----------
-The authors then go on to consider the rotation periods and lithium abundances
-of the extended sample, to compare them with previously-known distribution of
-similarly young stars.
+The authors then go on to consider the rotation periods and lithium
+abundances of the extended sample, to compare them with
+previously-known distribution of similarly young stars.
 
-The conceptual problem here is that stars that deviate from the known/expected
-rotational distribution for NGC 2516 are again not excised from the sample.  As
-prior studies (e.g. Fritzewski+2020) have shown, there are no rotators above
-the slow sequence in NGC 2516. So such stars most certainly have to be excised
-from further consideration. (Some of these are even older than Praesepe, as the
-authors themselves show in Fig.4.) The authors should consider whether other
-regions of the color-period diagram merit similar treatment.
+The conceptual problem here is that stars that deviate from the
+known/expected rotational distribution for NGC 2516 are again not
+excised from the sample.  As prior studies (e.g. Fritzewski+2020)
+have shown, there are no rotators above the slow sequence in NGC
+2516. So such stars most certainly have to be excised from further
+consideration. (Some of these are even older than Praesepe, as the
+authors themselves show in Fig.4.) The authors should consider
+whether other regions of the color-period diagram merit similar
+treatment.
 
@@
 
-> We agree that stars showing rotation periods above the slow sequence are most
-> likely not "genuine", in the sense that either i) they are not cluster
-> members, or ii) they are cluster members, but with unresolved binary
-> companions contaminating the rotation period measurement.  The latter
-> category has precedent in the literature (e.g., Stauffer+2016, AJ 152 115,
+> We agree that stars showing rotation periods above the slow
+> sequence are most likely not "genuine", in the sense that they are
+> most likely either not cluster members, or they are cluster
+> members, but with unresolved binary companions contaminating the
+> rotation period measurement (e.g., Stauffer+2016, AJ 152 115,
 > Section 5.1).
 >
-> Our point in proposing "Set A" and "Set B" is that the exact means by which
-> the samples are cleaned is not incredibly important to the central result that
-> the halo is the same age as the core, and therefore most likely related to
-> it.
+> At the referee's suggestion, we have therefore opted to redefine
+> Set B, in order to excise the stars above the slow sequence from
+> consideration.
 >
-> Set A comprises 987 stars.  947 of these have (Bp-Rp)_0>0.4, and so make some
-> degree of sense to assess gyrochronologically.  873 are below what by eye can
-> be identified as the "slow sequence boundary".  While there is subjectivity
-> in where this line is drawn, the big-picture result is that ~90% of the "Set
-> A" stars are be "genuine" in the sense of being "in or below the slow
-> sequence".  
+> The definition of Set A remains the same in the resubmitted
+> manuscript; Set B is now Set A, with the added requirement that
+> the rotation periods fall roughly below the expected slow sequence
+> (see revised Figure 4, top).
 >
-> We attach two additional figures, "response_fig2.png" and
-> "response_fig3.png" illustrating the point.  Cleaning out these stars to make
-> a "perfectly contaminant free" list of "genuine cluster members" is not the
-> aim of this work.
+> This change ultimately clarifies our results for the following
+> reason.  Set A comprises 987 stars.  947 of these have
+> (Bp-Rp)_0>0.4, and so make some degree of sense to assess
+> gyrochronologically.  892 of the latter stars are below what by
+> eye can be identified as the "slow sequence boundary".  While
+> there is subjectivity in where this line is drawn, the big-picture
+> result is that roughly 90% of the "Set A" stars are in or below
+> the slow sequence.  The arguments we initially drew using Set A
+> are therefore still applicable.
 >
-> (We note also that simply because a star is in or below the slow sequence
-> does not guarantee that it is genuine).
+> This change we feel improves the quality of the manuscript for a
+> separate reason as well: manual inspection revealed that many of
+> our Stellingwerf Phase Dispersion Minimization (SPDM) rotation
+> periods were less congruous with the data than the Lomb Scargle
+> periods (We wanted to know: what were half of them different? Our
+> suspicion is that the SPDM method is more suspectible to the gaps
+> int he TESS data).  We therefore have also opted to remove the
+> SPDM periods from Table 1, since their primary purpose was in the
+> definition of Set B.
 
@@
 ----------
-An additional concern is that the authors construct two samples A (where rotation
-periods from multiple types of measurements don't agree) and B (where rotation
-periods agree), but then go further with sample A (for instance Fig.6).
-Had the authors gone forward with the tightest-possible sample, and had then
-shown that this tight rotation-lithium sample still has a far greater extent on
-sky and in the canonical parallax, proper motion, RV etc. spaces, then readers would
-be far more willing to believe the authors' claims about a more extensive cluster.
+An additional concern is that the authors construct two samples A
+(where rotation periods from multiple types of measurements don't
+agree) and B (where rotation periods agree), but then go further
+with sample A (for instance Fig.6).  Had the authors gone forward
+with the tightest-possible sample, and had then shown that this
+tight rotation-lithium sample still has a far greater extent on sky
+and in the canonical parallax, proper motion, RV etc. spaces, then
+readers would be far more willing to believe the authors' claims
+about a more extensive cluster.
 
@@
 ----------
-> "response_fig2.png" and "response_fig3.png" illustrate that, for the
-> rotation-color selected subsample of Set A suggested by the referee in the
-> previous comment, the halo is the size claimed.
+> The top two rows of Figure 6 now use the redefined Set B --
+> rotational outliers have been excised.
 > 
-> The suggestion to use "Set B" for this is the suggestion to apply a selection
-> function that throws away ~85% of stars.  (We suspect this is mostly because
-> the SPDM periods tend to be garbage).  The halo is already sparse, so
-> throwing away ~9 in 10 of its members will make it sparser.
->
-> Regardless, doing this with Set B gives response_fig4.png.  The "slow
-> rotating" (and below) halo stars still span a large range of space, though
-> more like ~350 pc tip-to-tip rather than ~500 pc.
->
-> Anyway, in Set A (N=937), 216 stars are at least 50 parsecs from the cluster
-> center, and 104 are at least 100 parsecs away (23% and 11%).  In Set B
-> (N=479), 94 stars are at least 50 parsecs from the cluster center, and 43 are
-> at least 100 parsecs away (20% and 9%).
+> An alternative visualization of Fig 6 in the canonical "parallax,
+> proper motion, RV etc." space has been added in the new Appendix
+> E.  We preferred the visualizations we adopted since this
+> visualization approach gives the false impression of near-perfect
+> rotation period detection near the center of the cluster.
 
@@
 ----------
-The referee notes that the prior work of Healy & McCullough 2020, who also used
-TESS data to measure a sizeable group of stars in NGC 2516 has been completely
-ignored. The authors should include a discussion of the H+M2020 work in relation
-to the results in this paper.
+The referee notes that the prior work of Healy & McCullough 2020,
+who also used TESS data to measure a sizeable group of stars in NGC
+2516 has been completely ignored. The authors should include a
+discussion of the H+M2020 work in relation to the results in this
+paper.
 
@@
 ----------
-> Healy & McCullough 2020 (HM20) was referenced in the second paragraph of
-> Section 4.1, in relation to their discussion of the contraction rate of the
-> core NGC 2516, and the corresponding rate at which stars are escaping the
-> core.  The form in which Healy & McCullough quantified the contraction rate
-> does not readily translate (to our knowledge) to an estimate of the rate at
-> which stars escape from the cluster's tidal radius, or an analogous surface.
-> So, the reference was rather brief.
+> Healy & McCullough 2020 (HM20) was referenced in the second
+> paragraph of Section 4.1, in relation to their discussion of the
+> contraction rate of the core NGC 2516, and the corresponding rate
+> at which stars are escaping the core.  The form in which Healy &
+> McCullough quantified the contraction rate does not readily
+> translate (to our knowledge) to an estimate of the rate at which
+> stars escape from the cluster's tidal radius, or an analogous
+> surface.  So, the reference was rather brief.
 >
-> The main result of our paper -- that the halo is the same age as the core and
-> therefore most easily explained as being associated with NGC2516 -- does not
-> clearly translate to the main result of HM20 -- that the stellar spins in the
-> core of NGC 2516 are either isotropically oriented, or perhaps mildly
-> aligned.  Therefore we did not highlight HM20 prominently in our discussion.
+> The main result of our paper -- that the halo is the same age as
+> the core and therefore most easily explained as being associated
+> with NGC2516 -- does not clearly translate to the main result of
+> HM20 -- that the stellar spins in the core of NGC 2516 are either
+> isotropically oriented, or perhaps mildly aligned.  Therefore we
+> did not highlight HM20 prominently in our discussion.
 >
-> Nonetheless, to address the general idea behind this suggestion, we have i)
-> added a reference to HM20 in the introduction, when we discuss TESS and
-> NGC2516, and ii) included the HM20 rotation periods in Figure 10.  The
-> overall point of the figure has not changed.
+> Nonetheless, to address the general idea behind this suggestion,
+> we have i) added a reference to HM20 in the introduction, when we
+> discuss TESS and NGC2516, and ii) included the HM20 rotation
+> periods in Figure 10.  The overall point of the figure has not
+> changed.
 
@@
 ----------
-While on the subject, it should also be mentioned that the Introduction of the
-paper is only from the galactic structure perspective, and is lacking in
-introducing readers to the rotation-lithium perspective. Given that the latter
-constitute the bulk of the paper, doing that would clearly help the argument
-being laid out.
+While on the subject, it should also be mentioned that the
+Introduction of the paper is only from the galactic structure
+perspective, and is lacking in introducing readers to the
+rotation-lithium perspective. Given that the latter constitute the
+bulk of the paper, doing that would clearly help the argument being
+laid out.
 
@@
 ----------
-> We have added a paragraph in the introduction giving some context to this
-> effect -- thank you for the suggestion.
+> We have added a paragraph in the introduction giving some context
+> to this effect -- thank you for the suggestion.
 
@@
 ----------
-This referee came away with the general impression that the authors have likely
-identified a number of young stars in a vast halo around NGC 2516, but did not
-find the association with NGC 2516 itself to be particularly convincing.
-Perhaps they can think about how to make their arguments tighter, and the
-inclusion criterea stricter in the next iteration, to be able to better convince
-the community of the association with NGC 2516.
+This referee came away with the general impression that the authors
+have likely identified a number of young stars in a vast halo around
+NGC 2516, but did not find the association with NGC 2516 itself to
+be particularly convincing.  Perhaps they can think about how to
+make their arguments tighter, and the inclusion criterea stricter in
+the next iteration, to be able to better convince the community of
+the association with NGC 2516.
 
@@
 ----------
-> XXX of YYY stars for which we expected to detect rotation showed a detected
-> rotation signature...
-> In the comparison sample...
+> As discussed above, the "halo" stars i) have similar tangential
+> velocites as core stars, ii) have the same age as other halo
+> stars, iii) have the same age as the core stars, iv) are spatially
+> correlated with the location of the core stars, and v) have a
+> different rotation-color distribution from a randomly selected
+> sample of fields stars (with a large overdensity of rotators
+> falling on the same sequence as the core of NGC 2516).
+>
+> Points (i) and (iv) are the definitions of being associated
+> kinematically and spatially with the cluster core.  Points (ii),
+> (iii), and (v) have no reason to be the case if the "halo" stars
+> were a random population of young stars in the field.  Why would
+> this population itself be coeval?  Why would it happen to share an
+> age with the cluster core?  Is it not a simpler presumption that
+> the stars that are inematically and spatially linked with the
+> cluster core, and that have the same age as the cluster core, are
+> in fact related to the cluster?
 
@@
 ----------
-The .csv file attached with Table 1 has a large number of significant digits in
-most columns. In the revised manuscript, the authors should include a new
-version that has only the required precision. This document outlines the
-typical uncertainties for Gaia data:
+The .csv file attached with Table 1 has a large number of
+significant digits in most columns. In the revised manuscript, the
+authors should include a new version that has only the required
+precision. This document outlines the typical uncertainties for Gaia
+data:
 
@@
 
-Fewer significant digits will result in a smaller file that will be easier to
-read, manipulate, and transfer.
+Fewer significant digits will result in a smaller file that will be
+easier to read, manipulate, and transfer.
 
@@
 ----------
-> The CSV file has been updated with an appropriate number of significant
-> digits in each column; thank you for the suggestion.
+> The CSV file has been updated with an appropriate number of
+> significant digits in each column; thank you for the suggestion.

### 2021-07-02 cluster_paper/ms.tex
refere response ready for revision

@@
 
-\documentclass[12pt,twocolumn,tighten]{aastex63}
+%\documentclass[12pt,twocolumn,tighten]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
+\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
 \usepackage{amsmath,amstext,amssymb}
@@
 The results are shown in the top two rows of
-Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
-comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our TESS
-pipeline succeeded in making light curves.  In other words, the stars in
-the base sample are those for which we would have expected to detect a
-rotation period.  This color range is preferred because the the slow
-sequence becomes less defined for spectral types later than $\approx$K4V.
-The bins in the histograms (top-right) were chosen to ensure that the
-same number of stars were in each bin.  Rotation periods
-\added{consistent with a gyrochronological age of 150\,Myr} are detected
-for \replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc from the
-cluster core}{reported to be in the cluster}.  The average detection rate
-inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is slightly higher\added{ than
-outside of 25\,pc, where the detection rate is 63\% (181/288)}. Within
-$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\% (26/44)} is
-\replaced{relatively}{anomalously} low, due to crowding and saturation in the TESS images.
+Figure~\ref{fig:physical_x_rotn}.  \added{Alternative visualizations
+in the canoncial space of observed positions, parallaxes, and proper
+motions are presented in Appendix~\ref{app:gaia6d_x_rotn}.}  In
+\added{all} these figures, to ensure a fair comparison, we only show
+stars with $0.5<\bpmrpo<1.2$ for which our TESS pipeline succeeded in
+making light curves.  In other words, the stars in the base sample are
+those for which we would have expected to detect a rotation period.
+This color range is preferred because the the slow sequence becomes
+less defined for spectral types later than $\approx$K4V.  The bins in
+the histograms (top-right) were chosen to ensure that the same number
+of stars were in each bin.  Rotation periods \added{consistent with a
+gyrochronological age of 150\,Myr} are detected for
+\replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc
+from the cluster core}{reported to be in the cluster}.  The average
+detection rate inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is
+slightly higher\added{ than outside of 25\,pc, where the detection
+rate is 63\% (181/288)}. Within $\approx$3.5\,pc, the period detection
+rate of \replaced{61\%}{59\% (26/44)} is
+\replaced{relatively}{anomalously} low, due to crowding and saturation
+in the TESS images.
 
@@
     assessment.
+   	\added{The majority of the halo members do however exist within
+   	tangential velocity separations of $\approx$2\kms.}
   %
@@
 
-\begin{figure}[b]
+\begin{figure}[btp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f12.pdf}
+		\includegraphics[width=\textwidth]{f12.pdf}
 	\end{center}
@@
 
-% \section{Kinematics $\otimes$ Rotation}
-% \label{app:gaia6d_x_rotn}
-% 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=0.95\textwidth]{f4.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-%   \caption{ {\bf Gaia-based components of NGC\,2516 in position and
-%   velocity space, cross-matched against the rotators.} The plotted
-%   stars are those with $0.5<\bpmrpo<1.2$ that meet the crowding
-%   restrictions described in Section~\ref{subsec:tess}: they should
-%   have been sufficiently bright and non-crowded to enable rotation
-%   period detection.  We caution that the appearance of fewer
-%   non-rotators being present toward the core is due the layering of
-%   the plot: quantitatively, stars toward the cluster center do not all
-%   show rotation periods in our analysis (see
-%   Figure~\ref{fig:hist_physical_x_rotn}).
-%   \label{fig:gaia6d_x_rotn}
-% 	}
-% \end{figure*}
-% 
-% Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
-% data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
-% in this diagram correspond to ``Set $\mathcal{A}$'' as described in
-% Section~\ref{subsec:tess}.  Many stars in the outer region of the
-% cluster show rotation periods consistent with a young age.  One issue
-% with the visualization as displayed however is that it hides the
-% amount of non-rotators toward the cluster center: we did not detect
-% rotation periods for roughly one in four stars at the cluster
-% center (see Figures~\ref{fig:physical_x_rotn}
-% and~\ref{fig:hist_physical_x_rotn}).
+\section{Kinematics $\otimes$ Rotation}
+\label{app:gaia6d_x_rotn}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f15.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Gaia-based components of NGC\,2516 in position and
+  velocity space, cross-matched against the rotators.} The plotted
+  cluster members are those with $0.5<\bpmrpo<1.2$ that meet the
+  crowding restrictions described in Section~\ref{subsec:tess}: they
+  should have been sufficiently bright and non-crowded to enable
+  rotation period detection. Stars in Set $\mathcal{B}$ are shown in blue; those which should have
+  had rotation periods detected but did not are in orange.
+  We caution that the appearance of fewer
+  non-rotators being present toward the core is due the layering of
+  the plot: quantitatively, stars toward the cluster center do not all
+  show rotation periods in our analysis (see
+  Figure~\ref{fig:physical_x_rotn}).
+  \label{fig:gaia6d_x_rotn}
+	}
+\end{figure*}
+
+\added{Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization
+of the data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation
+periods in this diagram correspond to ``Set $\mathcal{B}$'' as
+described in Section~\ref{subsec:tess}.  One issue with the
+visualization as displayed however is that it hides the amount of
+non-rotators toward the cluster center: we did not detect rotation
+periods for roughly one in four stars at the cluster center (see
+upper-right panel in Figure~\ref{fig:physical_x_rotn}).  }
 % 

### 2021-07-02 drivers/plot_full_kinematics_X_rotation.py

@@ -6,8 +6,12 @@ PLOTDIR = os.path.join(RESULTSDIR, 'full_kinematics_X_rotation')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+ep.plot_full_kinematics_X_rotation(PLOTDIR, basedata='fullfaint',
+                                   show1937=0, galacticframe=1) # deeper
+assert 0
+
 for show1937 in [0,1]:
-    for galacticframe in [0,1]:
+    for galacticframe in [1,0]:
         ep.plot_full_kinematics_X_rotation(PLOTDIR, basedata='fullfaint', show1937=show1937, galacticframe=galacticframe) # deeper
         ep.plot_full_kinematics_X_rotation(PLOTDIR, basedata='bright', show1937=show1937, galacticframe=galacticframe) # standard
         ep.plot_full_kinematics_X_rotation(PLOTDIR, basedata='fullfaint_edr3', show1937=show1937, galacticframe=galacticframe) # deeper + EDR3

### 2021-07-02 earhart/helpers.py

@@ -783,6 +783,8 @@ def get_autorotation_dataframe(runid='NGC_2516', verbose=1, returnbase=0,
         'match234_alias': requires LS and SPDM periods to agree to within 10%
             (up to 1x,2x,3x,4x harmonic).
         'nocleaning': P<99d.
+        'defaultcleaning_cutProtColor': add Prot-color plane cut to
+            defaultcleaning.
     """
 
     assert isinstance(cleaning, str)

### 2021-07-02 earhart/plotting.py

@@ -2758,7 +2758,10 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
 
     nbhd_df, core_df, halo_df, full_df, trgt_df = get_gaia_basedata(basedata)
 
-    rot_df, lc_df = get_autorotation_dataframe(runid='NGC_2516', returnbase=True)
+    rot_df, lc_df = get_autorotation_dataframe(
+        runid='NGC_2516', returnbase=True,
+        cleaning='defaultcleaning_cutProtColor'
+    )
 
     dfs = [nbhd_df, core_df, halo_df, trgt_df, rot_df]
     nbhd_df.source_id = nbhd_df.source_id.astype(np.int64)
@@ -2784,14 +2787,24 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
         xl, yl = r'$\alpha$ [deg]', r'$\delta$ [deg]'
 
     params = [xkey, ykey, 'parallax', 'pmra', 'pmdec', rvkey]
+    # # whether to limit axis by 5/95th percetile
+    # qlimd = {
+    #     xkey: 0, ykey: 0, 'parallax': 0, 'pmra': 0, 'pmdec': 0, rvkey: 0
+    # }
+    # # whether to limit axis by 99th percentile
+    # nnlimd = {
+    #     xkey:0, ykey:0, 'parallax':0, 'pmra':0, 'pmdec':0, rvkey:0
+    # }
+
     # whether to limit axis by 5/95th percetile
     qlimd = {
-        xkey: 0, ykey: 0, 'parallax': 0, 'pmra': 0, 'pmdec': 0, rvkey: 0
+        xkey: 0, ykey: 0, 'parallax': 0, 'pmra': 1, 'pmdec': 1, rvkey: 1
     }
     # whether to limit axis by 99th percentile
     nnlimd = {
-        xkey:0, ykey:0, 'parallax':0, 'pmra':0, 'pmdec':0, rvkey:0
+        xkey: 1, ykey: 1, 'parallax': 1, 'pmra': 0, 'pmdec': 0, rvkey: 0
     }
+
     ldict = {
         xkey: xl, ykey: yl,
         'parallax': r'$\pi$ [mas]', 'pmra': r"$\mu_{{\alpha'}}$ [mas/yr]",
@@ -2825,21 +2838,22 @@ def plot_full_kinematics_X_rotation(outdir, basedata='bright', show1937=0,
[…]

### 2021-07-02 cluster_paper_review/referee_response_v0.txt
ditto

@@
 > periods were less congruous with the data than the Lomb Scargle
-> periods (We wanted to know: what were half of them different? Our
-> suspicion is that the SPDM method is more suspectible to the gaps
-> int he TESS data).  We therefore have also opted to remove the
-> SPDM periods from Table 1, since their primary purpose was in the
-> definition of Set B.
+> periods (Our suspicion is that the SPDM method is more suspectible
+> to the gaps in the TESS data).  We therefore have also opted to
+> remove the SPDM periods from Table 1, since their primary purpose
+> was in the definition of Set B.
 
@@
 > The top two rows of Figure 6 now use the redefined Set B --
-> rotational outliers have been excised.
+> rotational outliers have been excised.  For the reasons discussed
+> above, the conclusions do not change.
 > 
@@
 > proper motion, RV etc." space has been added in the new Appendix
-> E.  We preferred the visualizations we adopted since this
+> E.  We preferred the visualizations we adopted since the latter
 > visualization approach gives the false impression of near-perfect
 > rotation period detection near the center of the cluster.
+> Nonetheless, we agree that it has pedagogical value.
 

### 2021-07-07 cluster_paper_review/referee_response_v1.txt
add v1 ref response

--- 2021-07-02 cluster_paper_review/referee_response_v0.txt
@@
 > the Pleiades, and showing Li EWs similar to those of the Pleiades,
-> should indeed be considered to have comparable ages as stars in
-> Pleiades.
+> are likely to have comparable ages as stars in the Pleiades.
 >
@@
 > spatial correlation with the cluster center --  for the sake of
-> caution we analyzed rotation periods of randomly draw field stars
+> caution we analyzed rotation periods of randomly drawn field stars
 > in this region of sky.  This exercise was described in Appendix C,
@@
 > the core and halo of the cluster.  In particular, the lack of a
-> sub-giant branch or red giant branch (e.g., at G_BP-G_RP =1.2, M_G
-> =3.0)  in the core and halo suggests that both populations have a
+> sub-giant branch or red giant branch (e.g., at G_BP-G_RP =1.2,
+> M_G=3.0) in the core and halo suggests that both populations have a
 > different age distribution than the field stars, and is consistent
@@
 >
-> For the ~five halo stars lying significantly above the main
-> sequence, we agree that these must be field interlopers.  This is
-> noted at the end of Section 3.1, Paragraph 2.  The ~four halo
-> stars, and ~four core stars lying well below the main sequence
-> could have poorly determined parallaxes, or could be blended in
-> the Gaia BP or RP spectra with unresolved white dwarf companions.
-> We do not consider their presence below the MS to support their
-> exclusion from consideration as cluster members.
+> We agree that the ~five halo stars lying significantly above the
+> main sequence must be field interlopers.  This is noted at the end
+> of Section 3.1, Paragraph 2.  The ~four halo stars, and ~four core
+> stars lying well below the main sequence could have poorly
+> determined parallaxes, or could be blended in the Gaia BP or RP
+> spectra with unresolved white dwarf companions.  We do not consider
+> their presence below the MS to support their exclusion from
+> consideration as cluster members.
 
@@
 > periods were less congruous with the data than the Lomb Scargle
-> periods (Our suspicion is that the SPDM method is more suspectible
+> periods. Our suspicion is that the SPDM method is more susceptible
 > to the gaps in the TESS data).  We therefore have also opted to
@@
 > the core and therefore most easily explained as being associated
-> with NGC2516 -- does not clearly translate to the main result of
+> with NGC 2516 -- does not clearly translate to the main result of
 > HM20 -- that the stellar spins in the core of NGC 2516 are either
@@
 > we have i) added a reference to HM20 in the introduction, when we
-> discuss TESS and NGC2516, and ii) included the HM20 rotation
+> discuss TESS and NGC 2516, and ii) included the HM20 rotation
 > periods in Figure 10.  The overall point of the figure has not
@@
 be particularly convincing.  Perhaps they can think about how to
-make their arguments tighter, and the inclusion criterea stricter in
+make their arguments tighter, and the inclusion criteria stricter in
 the next iteration, to be able to better convince the community of
@@
 > this population itself be coeval?  Why would it happen to share an
-> age with the cluster core?  Is it not a simpler presumption that
-> the stars that are inematically and spatially linked with the
-> cluster core, and that have the same age as the cluster core, are
-> in fact related to the cluster?
+> age with the cluster core?  There are no known nearby populations
+> that it can plausibly be linked to (we searched the Kounkel & Covey
+> 2019 catalog).  So would it therefore constitute a new stand-alone
+> ~150 Myr population?  The simpler explanation is that stars that are
+> kinematically and spatially linked with the cluster core, and that
+> have the same age as the cluster core, are in fact related to the
+> cluster.
 

### 2021-07-08 cluster_paper_review/referee_response_20210707.txt
update ms
AAS journals resubmit

--- 2021-07-07 cluster_paper_review/referee_response_v1.txt
@@
 report, alongside our responses. A list of changes is appended to
-the manuscript using the trackchanges AASTeX macros. A figure
-expanding on the comparison between the field and cluster star
-rotation periods has also been uploaded for inspection
-(f13_corehalosplit.png).
+the manuscript using the trackchanges AASTeX macros
+(ms_trackchanges.pdf). A figure expanding on the comparison between
+the field and cluster star rotation periods has also been uploaded
+for inspection (f13_corehalosplit.png).
 
@@
 > periods. Our suspicion is that the SPDM method is more susceptible
-> to the gaps in the TESS data).  We therefore have also opted to
+> to the gaps in the TESS data.  We therefore have also opted to
 > remove the SPDM periods from Table 1, since their primary purpose
-> was in the definition of Set B.
+> was in the definition of Set B.  One validation of our adopted Lomb
+> Scargle periods is through the top-right panel of Figure 10, which
+> shows that almost all stars that overlap with previously reported
+> periods in the literature are consistent.
 

### 2021-07-08 cluster_paper/ms.tex
update ms
AAS journals resubmit

@@
 
-%\documentclass[12pt,twocolumn,tighten]{aastex63}
+\documentclass[12pt,twocolumn,tighten]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
+%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
 \usepackage{amsmath,amstext,amssymb}
@@
 \received{April 29, 2021}
-\revised{---}
+\revised{July 7, 2021}
 \accepted{---}
@@
   1,099 stars, are counted in 1\,mag bins.  The mean star in the
-  innermost sample has a 10\% lower mass than the mean star in the
+  innermost sample has a 10\% higher mass than the mean star in the
   outermost sample, consistent with expectations for tidal tails.
@@
 
-\subsection{Stellar Rotation at 100-200\,Myr}
+\subsection{Stellar rotation at 100-200\,Myr}
 \label{disc:rotn}
@@
 
-\section{Kinematics $\otimes$ Rotation}
+\section{Positions and Kinematics of the Rotators}
 \label{app:gaia6d_x_rotn}
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf Gaia-based components of NGC\,2516 in position and
+  \caption{ {\bf Core and halo of NGC\,2516 in position and
   velocity space, cross-matched against the rotators.} The plotted
@@
 described in Section~\ref{subsec:tess}.  One issue with the
-visualization as displayed however is that it hides the amount of
+visualization as displayed however is that the layering hides the 
 non-rotators toward the cluster center: we did not detect rotation

### 2021-07-16 cluster_paper/ARXIV_VERSION/ms.tex
ngc2516: AJ accepted

--- 2021-07-08 cluster_paper/ms.tex
@@
 \documentclass[12pt,twocolumn,tighten]{aastex63}
-%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
 \usepackage{amsmath,amstext,amssymb}
@@
 
-%\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
-%\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
-
 \newcommand{\ncompstardenominator}{656} % plot_compstar_rotation.py BpmRp0 from 0.4 to 0.8
@@
 \newcommand{\ms}{\,m\,s$^{-1}$}
-%\newcommand{\bpmrpo}{(\bpmrp)_0}
 \newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
@@
 \revised{July 7, 2021}
-\accepted{---}
-\submitjournal{AAS journals.}
+\accepted{July 16, 2021}
+\submitjournal{AAS Journals}
 \shorttitle{The Halo of NGC\,2516}
@@
 
-% Cluster Difference Imaging Photometric Survey. III.
-% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
 \title{
@@
 chose \cn\ for this analysis because it is young (100-200\,Myr) and
-close enough ($d\approx400$\,pc) to facilitate \replaced{photometric
-rotation measurements}{measurements of stellar rotation periods} using TESS.  \added{\citet{healy_stellar_2020}
+close enough ($d\approx400$\,pc) to facilitate measurements of stellar rotation periods using TESS.  \citet{healy_stellar_2020}
 in fact already used TESS rotation periods to study the stellar inclination
-distribution of stars in the core of NGC\,2516.}
-\replaced{Its halo}{The cluster's halo however} was reported by
+distribution of stars in the core of NGC\,2516.
+The cluster's halo however was reported by
 \citet{kounkel_untangling_2019} to span
-$\approx20\times10\times350$\,pc\added{, which is a larger volume than
+$\approx20\times10\times350$\,pc, which is a larger volume than
 has been considered in previous rotation analyses
-\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}}.
+\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}.
 We want to know: is this halo real or is it just a collection of
@@
 
-\added{To assess whether the stars in the cluster's halo are the same
+To assess whether the stars in the cluster's halo are the same
 ages as stars in the cluster's core, we apply three different age-dating
@@
 Gyrochronology has now been empirically calibrated for FGK stars with
-ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
-e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
+ages spanning $\approx$10\,Myr to $\approx$4\,Gyr \citep[{\it
+e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020,barnes_rotation_2016}.
 Lithium equivalent widths can be used for relative age-dating of dwarf
@@
 reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
-that we explore in the latter portions of our analysis.  }
+that we explore in the latter portions of our analysis.
 
@@
   and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
-  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
+  automated cleaning ; Set
   $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
-  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
+  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster}).  The Pleiades
   \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
-			$\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
-			$\mathcal{B}$ is shown} as in
+	\caption{ {\bf Binarity indicators for rotators in \cn.} Set
+			$\mathcal{B}$ is shown as in
 		Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
@@
 		orange.  
-		Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
+		Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has 523 stars.
 		Dividing these stars into ``slow'' and ``fast'' sequences by eye,
 		the fraction of stars showing signs of binarity in the slow and fast
-		subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
-		\replaced{56\% (40/71)}{51\% (68/134)}, respectively.
+		subsamples are 27\% (106/289) and
+		51\% (68/134), respectively.
 		\label{fig:binarity}
@@
 We then measured the rotation period from the resulting light curve using
-the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
-\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
-the Stellingwerf phase-dispersion minimization algorithm
-\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{ \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
+the Lomb-Scargle periodogram implementation in
+\texttt{astrobase} \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}.
 We used the CDIPS aperture radius that, based on theoretical
@@
 typically resulted in a circular aperture radius of either 1 or 1.5
-pixels.  We recorded the top five periodogram peaks\deleted{ from each
-method}, and their corresponding powers.  Finally, as a check on
+pixels.  We recorded the top five periodogram peaks, and their corresponding powers.  Finally, as a check on
 crowding, we recorded the number of stars within the aperture with
@@
 automated cleaning criteria, which yielded sets we denote
-$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
-$\mathcal{A}$}, we considered light
+$\mathcal{A}$ and $\mathcal{B}$.  For Set
+$\mathcal{A}$, we considered light
 curves for which the peak Lomb-Scargle periodogram period was below 15
@@
 completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
-\replaced{light curves}{stars}, out of \nautorotdenominator\ stars \replaced{for which rotation
-periods might plausibly have been detected}{that met the initial crowding requirements}.  For Set $\mathcal{B}$, we
-additionally required that the Lomb-Scargle \replaced{and Stellingwerf
-phase-dispersion methods yielded identical periods to within 10\%}{period
+stars, out of \nautorotdenominator\ stars that met the initial crowding requirements.  For Set $\mathcal{B}$, we
+additionally required that the Lomb-Scargle period
 fell below the boundary drawn in Figure~\ref{fig:rot}, which was
 constructed using the Pleiades polynomial from
-\citet{curtis_rup147_2020}}.
+\citet{curtis_rup147_2020}.
 This yielded \nautorotnumeratormatching\ light curves. 
@@
 discussed in Appendix~\ref{app:compstar}.
-\added{The rationale behind the additional cut in Set $\mathcal{B}$ is
+The rationale behind the additional cut in Set $\mathcal{B}$ is
 that \cn\ is already known to have few if any stars above the slow
@@
 rotational outliers might be bonafide cluster members, Set $\mathcal{A}$
-would be the better cut.}
+would be the better cut.
 
@@
 Set $\mathcal{A}$ is more complete than Set $\mathcal{B}$, this
-completeness comes at the expense of greater contamination.  \deleted{For
-instance, stars above the slow sequence (perhaps field contaminants)
-mostly disappear in Set $\mathcal{B}$.  The ``void'' or ``gap''
-underneath the slow sequence is also more pronounced in Set
-$\mathcal{B}$.}  To facilitate a comparison against the Pleiades and
+completeness comes at the expense of greater contamination.    To facilitate a comparison against the Pleiades and
 Praesepe, we used the rotation periods and dereddened colors from Table~5
@@
 Section~\ref{disc:absage}.
-
-\explain{Figure~\ref{fig:binarity} has been modified to account for the
-revised definition of Set B.}
 
@@
 over the color range where the slow and fast sequence are present
-($0.5<\bpmrpo<1.3$), there are \replaced{281}{523} stars, of which
-\replaced{103 (37\%)}{174 (33\%)} show signs of binarity.  Dividing these
-\replaced{281}{523} stars into ``fast'' and ``slow'' sequences by eye,
+($0.5<\bpmrpo<1.3$), there are 523 stars, of which
+174 (33\%) show signs of binarity.  Dividing these
+523 stars into ``fast'' and ``slow'' sequences by eye,
 the fraction of stars showing signs of binarity in the slow and fast
-subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and \replaced{56\%
-(40/71)}{51\% (68/134)}, respectively.  We speculate on possible
+subsamples are 27\% (106/289) and 51\% (68/134), respectively.  We speculate on possible
 explanations in Section~\ref{disc:rotn}.
@@
   2D-tangential velocities for halo members.  Stars with detected
-  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
+  rotation periods in Set~$\mathcal{B}$ are shown in blue;  halo
   members for which rotation periods should have been detectable, but
@@
   the {\it top right}, the 
-  ratio of detected to expected stars with rotation periods \added{in Set $\mathcal{B}$} is shown
+  ratio of detected to expected stars with rotation periods in Set $\mathcal{B}$ is shown
   versus 3D separation from
@@
   outermost 100 stars with detected rotation periods in position and
-  velocity space\added{ (from Set~$\mathcal{A})$}.
+  velocity space (from Set~$\mathcal{A})$.
   \label{fig:physical_x_rotn}
@@
 
-\explain{This paragraph contained a typo: the sentence containg ``at
-least 25\,pc from the cluster core'' was meant to refer to the entire
-cluster. }
-
 The results are shown in the top two rows of
-Figure~\ref{fig:physical_x_rotn}.  \added{Alternative visualizations
+Figure~\ref{fig:physical_x_rotn}.  Alternative visualizations
 in the canoncial space of observed positions, parallaxes, and proper
-motions are presented in Appendix~\ref{app:gaia6d_x_rotn}.}  In
-\added{all} these figures, to ensure a fair comparison, we only show
+motions are presented in Appendix~\ref{app:gaia6d_x_rotn}.  In
+all these figures, to ensure a fair comparison, we only show
 stars with $0.5<\bpmrpo<1.2$ for which our TESS pipeline succeeded in
@@
 the histograms (top-right) were chosen to ensure that the same number
-of stars were in each bin.  Rotation periods \added{consistent with a
-gyrochronological age of 150\,Myr} are detected for
-\replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc
-from the cluster core}{reported to be in the cluster}.  The average
-detection rate inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is
-slightly higher\added{ than outside of 25\,pc, where the detection
-rate is 63\% (181/288)}. Within $\approx$3.5\,pc, the period detection
-rate of \replaced{61\%}{59\% (26/44)} is
-\replaced{relatively}{anomalously} low, due to crowding and saturation
+of stars were in each bin.  Rotation periods consistent with a
+gyrochronological age of 150\,Myr are detected for
+68\% (471/692) of stars reported to be in the cluster.  The average
+detection rate inside of 25\,pc, 72\% (290/404), is
+slightly higher than outside of 25\,pc, where the detection
+rate is 63\% (181/288). Within $\approx$3.5\,pc, the period detection
+rate of 59\% (26/44) is
+anomalously low, due to crowding and saturation
 in the TESS images.
@@
 cluster.  Of the outermost 100 stars in position space with detected
-rotation periods\added{ in Set $\mathcal{A}$}, $\approx$8 appear as
+rotation periods in Set $\mathcal{A}$, $\approx$8 appear as
 outliers above the slow sequence, compared to just a few for the
@@
 contaminant rates of $\approx$50\% for the outermost cluster members in
-our \added{initial} target list.
-
-\replaced{Even with}{Despite} this level of contamination, the rotation period
+our initial target list.
+
+Despite this level of contamination, the rotation period
 measurements show that the halo extends to separations of
@@
 halo population of \cn\ was an analysis of the Li\,\textsc{I}
-6708\,\AA\ doublet.  \deleted{This type of age measurement has been reviewed by
-\citet{soderblom_ages_2010}, and its connection to stellar rotation
-has been discussed by \citet{bouvier_lithium-rotation_2020}.}
+6708\,\AA\ doublet.  
 For \cn, two spectroscopic datasets seemed important:
@@
 
-% the 21 stars that overlapped between GALAH and Gaia-ESO.  17
-% of the 21 measurements agreed within 1-$\sigma$.  The remaining four
-% were stars with $T_{\rm eff}>6200\,{\rm K}$, for which
-% \citet{randich_gaiaeso_2018} reported Li EW detections ranging from
-% $50$ to $75\,$m\AA, and for which we found either null results or only
-% slightly significant ($\approx$30m\AA) detections.
-
-% The outliers:
-%    Fitted_Li_EW_mA     EWLi            source_id  Teff
-%    24           -1.490   76.100  5290730488348194304  6786
-%    34           19.485   71.300  5290935341107132800  6702
-%    15           20.251   50.867  5290722310730727040  6197
-%    25           29.378   52.500  5290724681552584192  6511
-
 Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
@@
 	\caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
-		candidate \cn\ members with TESS rotation periods\added{ (Set $\mathcal{A}$)} and Gaia-ESO or
+		candidate \cn\ members with TESS rotation periods (Set $\mathcal{A}$) and Gaia-ESO or
 		GALAH spectra, and are colored by equivalent width.  Points with
@@
   The bottom panel is a concatenation of Set $\mathcal{B}$ with the
-  results of \replaced{\citet{Irwin_NGC2516_2007} (I07) and
-  \citet{fritzewski_rotation_2020} (F20)}{\citet{Irwin_NGC2516_2007}
+  results of \citet{Irwin_NGC2516_2007}
   (I07), \citet{fritzewski_rotation_2020} (F20), and
-  \citet{healy_stellar_2020} (HM20)}.
-  \replaced{I07 and F20}{I07, F20, and HM20} analyzed stars in the
+  \citet{healy_stellar_2020} (HM20).
+  I07, F20, and HM20 analyzed stars in the
   core of \cn.
@@
 \end{figure*}
-
-\explain{Figure~\ref{fig:litcomp} has been updated to include rotation
-periods from \citet{healy_stellar_2020}.}
 
@@
 includes, in order of precedence: rotation periods from our analysis,
-from \replaced{``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
-from the rotators in \citet{Irwin_NGC2516_2007}}{``Class 1 and
+from ``Class 1 and
 2'' \citet{fritzewski_rotation_2020} sources, from
 Table 1 of \citet{healy_stellar_2020}, and
-from \citet{Irwin_NGC2516_2007}}.  The main
+from \citet{Irwin_NGC2516_2007}.  The main
 contribution of our study beyond these previous analyses is our
@@
 \cn\, we see that the fraction of stars showing signs of binarity in the
-slow and fast rotation subsamples are \replaced{30\% (63/210) and 56\%
-(40/71)}{26\% (106/289) and 51\% (68/134)}; the fast rotators show a
+slow and fast rotation subsamples are 26\% (106/289) and 51\% (68/134); the fast rotators show a
 preference for binarity.  This correlation is in line with the earlier
 findings noted above.  While many rapid rotators in \cn\
-(\replaced{34\%}{66\%}) have no evidence for being binaries, the cuts we
+(66\%) have no evidence for being binaries, the cuts we
 used to select binaries (${\rm RUWE}>1.2$; photometric excess
@@
 dispersion in rotation periods increases.  The maximum rotation
-periods seen in \cn\ \deleted{across both Sets $\mathcal{A}$ and
-$\mathcal{B}$}
+periods seen in \cn\ 
 extend up to $\approx$11 days, rather than the $\approx$8.5 day upper
@@
 \citet{spada_competing_2020} model has an absolute scale that seems to
-be well-calibrated at these ages and colors.  For an 8 day rotation
+be well-calibrated at these ages and colors.  For an 8-day rotation
 period of a star with the same color, this model quotes a 193 Myr age;
@@
     assessment.
-   	\added{The majority of the halo members do however exist within
-   	tangential velocity separations of $\approx$2\kms.}
+   	The majority of the halo members do however exist within
+   	tangential velocity separations of $\approx$2\kms.
   %
@@
     $\approx$50\%.
-\replaced{Given the number density ratios of cluster to field stars at the widest separations ($\approx 10^{-3}$, \citealt{meingast_2021}), these contamination fractions are not as bad as one might fear.}{While the latter contamination fraction may seem high, a comparison sample of field stars (Appendix~\ref{app:compstar}) has a detection rate three times lower, and a  distribution in the period-color plane inconsistent with that of a population with a single age.}
+While the latter contamination fraction may seem high, a comparison sample of field stars (Appendix~\ref{app:compstar}) has a detection rate three times lower, and a  distribution in the period-color plane inconsistent with that of a population with a single age.
    %
@@
   the kinematically selected \cn\ members (black).  The field and \cn\
-  stars are both subject to the same \deleted{``Set $\mathcal{A}$''}
+  stars are both subject to the same 
   period-measurement and cleaning procedures.
-  \added{The dashed line is as in Figure~\ref{fig:rot}}.
+  The dashed line is as in Figure~\ref{fig:rot}.
   \label{fig:compstar}
@@
 
-\explain{Numbers in this paragraph have been updated per the revised
-definition of Set $\mathcal{B}$. }
-
 We performed the same light curve stitching and period-search
 procedure discussed in the text on the field stars.
-\replaced{Imposing}{Within $0.4<\bpmrpo<0.8$}, the same requirements
-for crowding resulted in \ncompstardenominator\ \added{field} stars
+Within $0.4<\bpmrpo<0.8$, the same requirements
+for crowding resulted in \ncompstardenominator\ field stars
 for which rotation periods could have been detected.  Imposing the
-same Lomb Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
-upper limit ($P<15\,{\rm days}$)}{color-period cutoff as that used to
-define Set $\mathcal{B}$} yielded \ncompstarnumerator\ period
-detections (\ncompfrac).  Within the same \replaced{brightness
-cutoff}{color range}, \nautovscompstarnumerator\ of
-\nautovscompstardenominator\ \replaced{cluster stars}{kinematically
-identified candidate cluster members} yielded period detections
-(\nautofrac).  \replaced{While the detection fractions are only mildly different
-(likely due to the brightness cutoff), the }{The }period--color distributions
-are \replaced{very}{also quite} different\added{: the candidate
+same Lomb-Scargle power cutoff (${\rm LSP}>0.08$) and color-period cutoff as that used to
+define Set $\mathcal{B}$ yielded \ncompstarnumerator\ period
+detections (\ncompfrac).  Within the same color range, \nautovscompstarnumerator\ of
+\nautovscompstardenominator\ kinematically
+identified candidate cluster members yielded period detections
+(\nautofrac).  The period--color distributions
+are also quite different: the candidate
 cluster members show an overdensity coincident with the expected
 gyrochronal age, while the field stars show a much broader rotation
-period distribution} (Figure~\ref{fig:compstar}).  This confirms our
+period distribution (Figure~\ref{fig:compstar}).  This confirms our
 expectation that our period measurement procedure is not biased in any
@@
 
-\added{Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization
+Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization
 of the data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation
@@
 periods for roughly one in four stars at the cluster center (see
-upper-right panel in Figure~\ref{fig:physical_x_rotn}).  }
+upper-right panel in Figure~\ref{fig:physical_x_rotn}).  
 % 

### 2021-07-16 cluster_paper/ms.tex
ngc2516: AJ accepted

@@
 \revised{July 7, 2021}
-\accepted{---}
-\submitjournal{AAS journals.}
+\accepted{July 16, 2021}
+\submitjournal{AAS Journals.}
 \shorttitle{The Halo of NGC\,2516}
@@
 for which rotation periods could have been detected.  Imposing the
-same Lomb Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
+same Lomb-Scargle power cutoff (${\rm LSP}>0.08$) and \replaced{period
 upper limit ($P<15\,{\rm days}$)}{color-period cutoff as that used to

### 2021-07-16 planet_paper/ms.tex
ngc2516: AJ accepted

@@
 
+- Get a spectrum of the secondary?
+  Andrew Mann suggests this is possible.
+     The primary is G=13.0, secondary G=17.6.  The age test would be
+     based on Halpha.  
+     ``If there's no halpha you can be sure it's old, but if there is
+     some it's kinda inconclusive''
+     ``10-40\% of M3s are inactive, assuming that the SpT is right ''
+     ``get the M dwarf spectrum in good seeing and confirm age, or
+     don't bother''
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-09-16 cluster_paper/ms.tex
update ms for proofing

@@
 In some cases, these halos may actually be tidal tails, as have been
-reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019},
+reported in the Hyades \citep{meingast_hyades_2019,roser_hyades_2019,jerabkova_800_2021},
 the Ursa Major moving group \citep{gagne_lowmassUMA_2020}, and Coma
@@
 
-The most likely explanation seems to be that the observed halo
-structures are tidal tails \citep[{\it
-e.g.},][]{chumak_tails_2006,krumholz_star_2019}. The idea is that
-stars near the cluster's tidal radius escape slowly due to the
-galactic tide, and subsequently form leading and lagging arms due to
-differential rotation in the Galaxy.  The contraction rate of the
-cluster can affect this process.  Whether the exact contraction rate
-of \cn\ \citep{healy_stellar_2020}, and the corresponding evaporation
-rate of the core are sufficient to explain the formation of the halo
-has yet to be assessed. 
+% UPDATED IN PROOFING
+The most likely explanation seems to be that the observed halo structures are
+tidal tails \citep[{\it
+e.g.},][]{chumak_tails_2006,krumholz_star_2019,dinnbier_tidaltwo_2020}.  As
+discussed by \citet{dinnbier_tidalone_2020}, the tail could either be
+mostly made of stars released during the gas expulsion phase, or it
+could be made of
+stars released during the long-term evaporation of the cluster.  In their
+N-body simulations, \citet{dinnbier_tidaltwo_2020} found that for the tails to
+be as rich as those of \cn, they would need to be dominated by stars
+released during an impulsive gas expulsion phase.  The observed
+morphology and kinematics of the \cn\ halo also appear to qualitatively agree
+with these N-body predictions.
+
+% % ACCEPTED
+% The most likely explanation seems to be that the observed halo
+% structures are tidal tails \citep[{\it
+% e.g.},][]{chumak_tails_2006,krumholz_star_2019}. The idea is that
+% stars near the cluster's tidal radius escape slowly due to the
+% galactic tide, and subsequently form leading and lagging arms due to
+% differential rotation in the Galaxy.  The contraction rate of the
+% cluster can affect this process.  Whether the exact contraction rate
+% of \cn\ \citep{healy_stellar_2020}, and the corresponding evaporation
+% rate of the core are sufficient to explain the formation of the halo
+% has yet to be assessed. 
 % The presence or lack of epicyclic
@@
 systems are shorter than in single star systems.
-\citet{eggenberger_impact_2012} showed that longer disk lifetimes
+\citet{bouvier_lithium_2008} and
+\citet{eggenberger_impact_2012} have shown that longer disk lifetimes
 should lead to an extended phase of disk-locking, which leads to a

### 2021-09-16 earhart/extinction.py

@@ -4,6 +4,10 @@ Tools for working out the extinction.
 
 import numpy as np
 
+def EBmV_to_AV(EBmV):
+    R_V = 3.1
+    return 3.1*EBmV
+
 def AV_to_EBpmRp(A_V):
     """
     Convert A_V to E(Bp-Rp). NOTE: assumes A_V has been "corrected" for
